### 2020-07-02 notes/notes-SA-partSelf-wDemography.tex
Initialize repo

% Test change for git
%===========================================
% Preamble
\documentclass[11pt]{article}

%**********
%Dependencies
\usepackage[left]{lineno}
\usepackage{titlesec}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
%\usepackage[utf8]{inputenc}
\usepackage{color,soul}
%\usepackage{setspace}
%\usepackage{times}

\usepackage{booktabs}
% \usepackage[singlelinecheck=false]{caption}
\usepackage{tikz}

\newcommand\encircle[1]{%
  \tikz[baseline=(X.base)] 
    \node (X) [draw, shape=circle, inner sep=0] {\strut #1};}

\usepackage{pdflscape}

% Packages from Am. Nat. Template
\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
\usepackage{fullpage}
\usepackage[authoryear,sectionbib,sort]{natbib}
\linespread{1.15}
\usepackage[utf8]{inputenc}
\usepackage{lineno}

% Change default margins
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}

% Change subsection numbering
\renewcommand\thesubsection{\arabic{subsection})}
\renewcommand\thesubsubsection{}

% Subsubsection Title Formatting
\titleformat{\subsubsection}    
{\normalfont\fontsize{12pt}{17}\itshape}{\thesubsubsection}{12pt}{}

% Equation numbering
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

% Definitions
\def\mathbi#1{\textbf{\em #1}}
\def\mbf#1{\mathbf{#1}}
\def\mbb#1{\mathbb{#1}}
\def\mcal#1{\mathcal{#1}}

%===========================================

\begin{document}
\title{Stage-structured evolutionary demography for simultaneous hermaphrodites with partial selfing}
\author{Colin Olito}
\date{17 June 2020}
\maketitle

\textbf{Last updated:} \date{\today}

\subsection*{Notes}

This all started with an idea I had while watching Lotte DeVries give a guest lecture at the STN meeting in Montpellier in 2019. Lotte presented her stage-structured evolutionary demography matrix models for two sexes. I got super jazzed because they provide a cool, and novel, way to integrate population genetics and demography. One potential extension of the models that I was excited to explore was (surprise, surprise!) sexually antagonistic selection in partially selfing hermaphrodite populations. I've modeled the maintenance of SA polymorphism in selfing populations before, but of course these models all assume static demography, and model only the change in frequency of the SA alleles. What would be cool about using Lotte's models in this context is that we'd be able to address not only when SA polymorphism will be maintained, but also when SA selection results in demographic inviability. This is potentially very cool because ({\itshape i}) female fitness is demographically dominant, and ({\itshape ii}) partial selfing creates a well known skew in selection towards female functions. Because of this, male-beneficial/female-detrimental alleles will have a stronger impact on a population's demographic viability, but the parameter spaces resulting in SA polymorphism in partially selfing population are more likely to be demographically viable... this is cool because one of the main conclusions of the 1-locus models of Jordan \& Connallon (2014) and Olito (2017) is that selfing reduces the total parameter space where SA polymorphism is maintained. My intuition is that when we take demography into account explicitly with Lotte's models, it will show that this 'loss' isn't as bad as it seems, because SA polymorphic parameter space for partially selfing populations is more demographically viable.
\bigskip

%%%%%%%%%%%%
\section{Constructing a stage-structured model for a partially selfing population}

	{\itshape Below, I'm walking through the steps of constructing one of Lotte's models for a stage-structured matrix model for a partially selfing population with male and female fertility effects \ldots}
	\bigskip

\subsection{Component matrices}

\noindent Individuals are jointly classified by stage $(1, \ldots , \omega)$ and genotype $(1, \ldots , g)$. Each genotype is characterized by a matrix of transition probabilities (including survival) and a matrices of reproductive output through male and female function. These matrices can include time variation or nonlinearities reflecting the environment or density dependence, although thes complexities will not be addressed here. Each stage contributes offspring to genotypes at the next time step according to matrices that are determined by the mating system and the population structure.

The population state at time $t$ is described by a stage $\times$ genotype distribution matrix

\begin{equation}
	\mathcal{N} \sim \left[
						\begin{array}{ccc}
							n_{1,1}      &  \cdots & n_{1,g} \\
							\vdots       &         & \vdots  \\
							n_{\omega,1} &  \cdots & n_{\omega,g} \\
						\end{array} \right],
\end{equation}

\noindent where $n_{i,j}$ denotes the number of individuals of stage $i$ and genotype $g$. This matrix is transformed into a population state vector:

\begin{equation}
	\tilde{\mbf{n}}(t) = \text{vec}\{\mathcal{N}(t)\}.	
\end{equation}

\noindent For a single locus with two alleles ($A$ and $a$), we have genotypes $g \in \{AA,\, Aa,\, aa\}$, giving the population state vector:

\begin{equation}
	\tilde{\mbf{n}}(t) =  \left[
								\begin{array}{c}
									\mbf{n}_{AA}(t) \\
									\mbf{n}_{Aa}(t) \\
									\mbf{n}_{aa}(t) \\
						\end{array} \right].
\end{equation}

\noindent The population vector $\tilde{\mbf{n}}(t)$ is projected from time $t$ to $t + 1$ by a matrix $\tilde{\mbf{A}}(\tilde{\mbf{n}})$ such that

\begin{equation}
	\tilde{\mbf{n}}(t + 1) = \tilde{\mbf{A}}[\tilde{\mbf{n}}] \, \tilde{\mbf{n}}(t)
\end{equation}

\noindent The proportional population vector is

\begin{equation} \label{eq:propPopVec}
	\tilde{\mbf{p}}(t) = \frac{\tilde{\mbf{n}}(t)}{ \| \tilde{\mbf{n}}(t) \|} = \left( \mbf{p}(t) \right)
\end{equation}

The projection matrix $\tilde{\mbf{A}}$ is constructed from four sets of matrices representing the demographic and genetic processes:

\begin{table}[htbp]
\centering
\begin{tabular}{ l l } 
	$\mbf{U}_{i}$     & demographic transitions for genotype $i$ \\
				         &~~~ $i = 1, \ldots, g$ $\omega \times \omega$ \\
	$\mbf{F}_{i}$ & fertility matrix (female-function) for genotype $i$ \\
				         &~~~ $i = 1, \ldots, g$ $\omega \times \omega$ \\
	$\mbf{F}^{\prime}_{i}$ & fertility matrix (male-function) for genotype $i$ \\
				         &~~~ $i = 1, \ldots, g$ $\omega \times \omega$ \\
	$\mbf{D}_{i}$     & genotype transitions for stage $i$ \\
				         &~~~ $i = 1, \ldots, \omega$ $g \times g$	\\
	$\mbf{H}^X_{i}(\tilde{\mbf{n}})$ & genotype transitions for stage $i$ by outcrossing\\
				         &~~~ $i = 1, \ldots, \omega$ $g \times g$ \\	
	$\mbf{H}^S_{i}(\tilde{\mbf{n}})$ & genotype transitions for stage $i$ by selfing\\
				         &~~~ $i = 1, \ldots, \omega$ $g \times g$ \\	
\end{tabular}
\end{table}

\noindent The matrix $\mbf{U}_{i}$ contains transition and survival probabilities for genotype $i$. The matrices $\mbf{F}^{\text{sex}}_{i}$ contains stage-specific fertility rates for genotype $i$ through female and male sex functions. In the absence of genetic structure, these would be the familiar transition and fertility matrices making up a population projection matrix. Because the matrices $\mbf{U}_{i}$ and the $\mbf{F}^{\text{sex}}_{i}$ can differ among genotypes in any way, the model admits any kind of pleiotropy in demographic traits.

The matrix $\mbf{D}_{i}$ contains genotype transition probabilities for individuals in stage $i$, but because genotypes are fixed within an individual, we set $\mbf{D}_{i} = \mbf{D} = \mbf{I}_{g}$ for all $i$, where $\mbf{I}_{g}$ is the identity matrix of size $g \times g$.

The matrices $\mbf{H}^X_{i}(\tilde{\mbf{n}})$ and $\mbf{H}^S_{i}(\tilde{\mbf{n}})$ are parent-offspring maps, from the genotype of the parent in stage $i$ to the genotypes of their offspring produced by outcrossing (indicated by superscript $X$) and selfing (indicated by superscript $S$). The $(k, l)$ entry of $\mbf{H}^{X}_{i}$ and $\mbf{H}^{S}_{i}$ is the probability that an offspring of a genotype $l$ mother, of stage $i$, has genotype $k$. For the purpose of this article, we assume that mating is random with respect to stage and hence that the parent-offspring map is the same for all stages, that is, $\mbf{H}^j_{i}(\tilde{\mbf{n}}) = \mbf{H}^j(\tilde{\mbf{n}})$. The matrices $\mbf{H}^j(\tilde{\mbf{n}})$ contains the population genetic processes and will be derived in the next section...

%%%%%%%%%%%
\subsection{Mapping parental genotypes to offspring genotypes}

To model the Mendelian genetics of offspring production, a description of the genotype and allele structure of the mating population is required. Not every life-cycle stage will reproduce, and nonreproductive (e.g., immature) stages play no role in determining the genotype frequencies among off- spring. We define the breeding population by a set of indicator vectors $c_j$ for $j = 1, \ldots , g$ showing which stages of genotype $j$ take part in mating. That is, the $i^{th}$ entry of $c_j$ is $1$ if stage $i$ of genotype $j$ reproduces and $0$ otherwise.

To describe genotype frequencies in the mating process, we will distinguish five vectors of genotype frequencies:

\begin{table}[htbp]
\centering
\begin{tabular}{ l } 
	$\mbf{p} =$ genotype frequencies in the overall population, \\
	$\mbf{p_b} =$ genotype frequencies in the breeding population, \\
	$\mbf{p_i} =$ genotype frequencies in genotype $i$ ($=\mbf{e}_i$), \\
	$\mbf{p^X_i} =$ genotype frequencies in the offspring of genotype $i$ \\
	~~~~~~~~~~resulting from outcrossing, \\
	$\mbf{p^S_i} =$ genotype frequencies in the offspring of genotype $i$ \\
	~~~~~~~~~~resulting from selfing,\\
\end{tabular}
\end{table}

The size of the breeding population is

\begin{equation}
	N_b = \sum_{j=1}^g (\mbf{e}_j^\intercal \otimes \mbf{c}_j^\intercal) \tilde{\mbf{n}},
\end{equation}

\noindent where $\mbf{e}_j$ is a vector ($g \times 1$) with a $1$ in position $j$ and zeros elsewhere, and $\otimes$ is the Kronecker product. Breeding stages can differ among genotypes in order to study the fate of traits that change reproductive schedules. We focus on the special case where the genotypes do not differ in their reproductive stages, $\mbf{c}_j = \mbf{c}$ for all genotypes $j$ and

\begin{equation}
	N_b = (\mbf{1}_g^\intercal \otimes \mbf{c}^\intercal) \tilde{\mbf{n}},
\end{equation}

\noindent where $\mbf{1}_g$ is a vector of length $g$.

The genotype frequency vector in the breeding population is

\begin{equation}
	\mbf{p}_b = \frac{\mbf{X} \tilde{\mbf{n}}}{N_b},
\end{equation}

\noindent where $\mbf{X}$ is a matrix that combines abundances of breeding stages. In our case, the breeding vectors are the same for all genotypes, $\mbf{c}_i = \mbf{c}$, and

\begin{equation}
	\mbf{X} = (\mbf{1}_g^\intercal \otimes \mbf{c}^\intercal).
\end{equation}

\noindent The genotype frequency vector for genotype $i$ is (trivially) $\mbf{p}_i = \mbf{e}_i$.

The gene frequencies are a function of the genotype frequencies, so that

\begin{equation}
		\mbf{q}_i = \mbf{W} \mbf{p}_i,
\end{equation}
\begin{equation}
		\mbf{q}_b = \mbf{W} \mbf{p}_b,
\end{equation}

\noindent where

\begin{equation}
		\mbf{W} = \left[
						\begin{array}{ccc}
							1 & 1/2 & 0 \\
							0 & 1/2 & 1 \\
						\end{array} \right].
\end{equation}

\noindent Note that the structure of $\mbf{W}$ is specific to the diploid two-allele case (see deVries \& Caswell 2019 for details on how to extend for other cases). Mutation can also be introduced at this step, but we do not investigate this here.

%%%%%%%%%%%%%%%
\subsubsection*{Mating and offspring under partial selfing}

In a population of hermaphrodites, each individual expresses both male and female sex functions (i.e., can be both a mother and father). In partially selfing populations, individuals can also reproduce by a combination of outcrossing and self-fertilization. Outcrossing can be acheived through both sex-functions: outcrossing as a mother involves receiving male gametes from another individual in the population, while outcrossing as a father involves exporting male gametes to another individual. Selfing involves using and individual's own male gametes to fertilize their own ovules. These options complicate our model relative to a standard or two-sex stage-structured models (deVries \& Caswell 2019; 2020). For generality, we describe a genotype-specific model of selfing with constant inbreeding depression (Charlesworth \& Charlesworth 1987). Let $C_{ii}$ represent the relative proportion of ovules of genotype $ii$ that are self-fertilized ($1 - C_{ii}$ are outcrossed), and $\delta$ represent the viability cost suffered by self-fertilized offspring due to inbreeding depression (Charlesworth and Charlesworth 1987).

Outcrossing proceeds much like a two-sex model of reproduction. Consider a random individual of genotype $i$, and let $p_i$ be the genotype distribution of their offspring. Individuals of each stage and genotype combination contribute differentially to a pool of male gametes. The allele frequencies in the male gamete pool are obtained from the male stage $\times$ genotype vector:

\begin{equation}
	\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p} = 
		\left(
			\begin{array}{ccc}
				\mbf{1}^{\intercal}_{\omega} & \frac{1}{2} \mbf{1}^{\intercal}_{\omega} & 0 \\
				0 & \frac{1}{2} \mbf{1}^{\intercal}_{\omega} & \mbf{1}^{\intercal}_{\omega} \\
			\end{array} \right)
		\left(
			\begin{array}{ccc}
				\mbf{F}^{\prime}_{AA} & 0 & 0 \\
				0 & \mbf{F}^{\prime}_{Aa} & 0 \\
				0 & 0 & \mbf{F}^{\prime}_{aa} \\
			\end{array} \right)
		\left(
			\begin{array}{c}
				\mbf{p}_{AA} \\
				\mbf{p}_{Aa} \\
				\mbf{p}_{aa} \\
			\end{array} \right)
\end{equation}

\noindent The matrix $\mbb{F}^{\prime}$ operates on the vector of genotype frequencies to give the relative contributions of each genotype to the male gamete pool. The matrix $\mbf{W}^{\prime}$ converts these relative genotype contributions to allele numbers. Normalizing this vector gives the allele frequencies in the gamete pool,

\begin{equation} \label{eq:maleGametePool}
	\left(
		\begin{array}{c}
			q^{\prime}_{A} \\
			q^{\prime}_{a} \\
		\end{array} \right) = 
			\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p} \|} = 
				\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{n}}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{n} \|}
\end{equation}

\noindent These frequencies determine the distribution of genotypes in the outcrossed offspring of maternal parents of any genotype.

The outcrossing parent–offspring matrix is a function of the allele frequencies in the male gamete pool, hence,

\begin{equation} \label{eq:HX}
	\mbf{H}^X(\tilde{\mbf{n}}) = 
			\left(
			\begin{array}{ccc}
				q^{\prime}_{A} (1 - C_{AA}) & \frac{1}{2} q^{\prime}_{A} (1 - C_{Aa}) & 0 \\
				q^{\prime}_{a} (1 - C_{AA}) & \frac{1}{2} (1 - C_{Aa}) & q^{\prime}_{A}  (1 - C_{aa}) \\
				0 & \frac{1}{2} q^{\prime}_{a} (1 - C_{Aa}) & q^{\prime}_{a} (1 - C_{aa}) \\
			\end{array} \right)
\end{equation}

\noindent The allele frequencies $q^{\prime}_A$ and $q^{\prime}_a$ are given in terms of either $\mbf{p}$ or $\mbf{n}$ by Eq(\ref{eq:maleGametePool}). The first column of $\mbf{H}^X(\tilde{\mbf{n}})$ contains the genotype distribution of the offspring of an $AA$ maternal parent; they produce an $AA$ offspring with probability $q^{\prime}_A$ and an $Aa$ offspring with probability $q^{\prime}_a$. The second and third columns give the genotype distributions for maternal parents of genotypes $Aa$ and $aa$.

Reproduction via self fertilization is slightly different. We assume that individuals produce enough male gametes to easily fertilize all of their available ovules, and that the transport of and individual's own male gametes to their ovules involves little or no selection from external factors (at least compared to outcrossing). Hence, the parent-offspring matrix is determined entirely by the probabilities of segregation and fertilization during meiosis:

\begin{equation} \label{eq:HS}
	\mbf{H}^S(\tilde{\mbf{n}}) = 
			\left(
			\begin{array}{ccc}
				C_{AA}(1 - \delta) & \frac{1}{4} C_{Aa}(1 - \delta) & 0 \\
				0      & \frac{1}{2} C_{Aa}(1 - \delta) & 0 \\
				0 & \frac{1}{4} C_{Aa}(1 - \delta) & C_{aa}(1 - \delta) \\
			\end{array} \right),
\end{equation}

\noindent where the first column of $\mbf{H}^S(\tilde{\mbf{n}})$ contains the genotype distribution of the offspring of an $AA$ maternal parent; in this case all male and female gametes will carry the $A$ allele, and so all offspring produced by selfing will also be $AA$. The other columns can be calculated similarly.

%%%%%%%%%%%%%%%
\subsubsection*{Population projection}

The matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ that projects the eco-evolutionary dynamics is

\begin{equation}
	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{U}^S & \mbf{0} \\ \hline
				\mbf{0} & \mcal{U}^X \\
			\end{array} \right)}_{\tilde{\mbf{U}}} + 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{F}^S(\tilde{\mbf{p}}) (1 - \delta) & \mbf{0} \\ \hline
				\mbf{0} & \mcal{F}^X(\tilde{\mbf{p}}) \\
			\end{array} \right)}_{\tilde{\mbf{F}}}
\end{equation}

\noindent where $\tilde{\mbf{p}}$ is calculated from $\tilde{\mbf{n}}$ using Eq(\ref{eq:propPopVec}). The blocks correspond to production of offspring by self-fertilization and outcrossing ($\mcal{F}^S$ and $\mcal{F}^X$ in $\tilde{\mbf{F}}$), and survival of selfed and outcrossed offspring ($\mcal{U}^S$ and $\mcal{U}^X$ in $\tilde{\mbf{U}}$).

To construct $\tilde{\mbf{A}}$ using the vec-permutation matrix approach (Caswell et al., 2018), create a set of block-diagonal matrices; e.g.,

\begin{equation}
	\mbb{U}^S = 
		\left(
			\begin{array}{ccc}
				\mbf{U}^{S}_{AA} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{U}^{S}_{Aa} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mbf{U}^{S}_{aa} \\
			\end{array} \right)
\end{equation}

\noindent and corresponding matices $\mbb{U}^X$, $\mbb{F}^S$, $\mbb{F}^X$. Because individuals cannot change their genotype once they are born, the survival matrices are block diagonal; $\mcal{U}^S = \mbb{U}^S$ and $\mcal{U}^X = \mbb{U}^X$. We also construct

\begin{equation}
	\mbb{H}^j = \mbf{I}_{\omega} \otimes \mbf{H}^j(\tilde{\mbf{p}}),
\end{equation}

\noindent where $j \in \{S,X\}$.

The fertility matrices for self-fertilization and outcrossing are defined as

\begin{equation} \label{eq:FS}
	\mcal{F}^S = \mbf{K}^{\intercal} \mbb{H}^S(\tilde{\mbf{p}}) \mbf{K} \mbb{F}^S, 
\end{equation}
\noindent and, 
\begin{equation} \label{eq:FX}
	\mcal{F}^X = \mbf{K}^{\intercal} \mbb{H}^X(\tilde{\mbf{p}}) \mbf{K} \mbb{F}^X \numberthis
\end{equation}

\noindent where $\mbf{K}$ is the vec-permutation matrix (Henderson and Searle 1981). From right to left, the block-diagonal matrices $\mbb{F}^j$ produce offspring as a function of the genotype of the mother, the vec-permutation matrix $\mbf{K}$ rearranges the vector, the block-diagonal
matrix $\mbb{H}^j(\tilde{\mbf{p}})$ allocates the offspring to their genotypes, and $\mbf{K}^{\intercal}$ returns the vector to its original orientation.

Substituting Eq(\ref{eq:HS}) into Eq(\ref{eq:FS}), Eq(\ref{eq:HX}) into Eq(\ref{eq:FX}), and simplifying yields

\begin{equation} \label{eq:FSsimp}
	\mcal{F}^S(\tilde{\mbf{p}}) = 
			\left(
			\begin{array}{c|c|c}
				C_{AA} \mbf{F}_{AA} & \frac{1}{4} C_{Aa} \mbf{F}_{Aa} & 0 \\ \hline
				0 & \frac{1}{2} C_{Aa} \mbf{F}_{Aa} & 0 \\ \hline
				0 & \frac{1}{4} C_{Aa} \mbf{F}_{Aa} & C_{aa} \mbf{F}_{aa}\\
			\end{array} \right), 
\end{equation}
\noindent and, 
\begin{equation} \label{eq:FXsimp}
	\mcal{F}^X = 
			\left(
			\begin{array}{c|c|c}
				q^{\prime}_{A} \mbf{F}_{AA} (1 - C_{AA})& \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} (1 - C_{Aa}) & 0 \\ \hline
				q^{\prime}_{a} \mbf{F}_{AA} (1 - C_{AA}) & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & q^{\prime}_{A} \mbf{F}_{aa} (1 - C_{aa}) \\ \hline
				0 & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} (1 - C_{Aa})& q^{\prime}_{a} \mbf{F}_{aa} (1 - C_{aa}) \\
			\end{array} \right), 
\end{equation}

\noindent where $q^{\prime}_A$ and $q^{\prime}_a$ are given by Eq(\ref{eq:maleGametePool}). See de Vries \& Caswell (2019) for a derivation of Eqs(\ref{eq:FSsimp}) and (\ref{eq:FXsimp}).

Let's look at the first block column of $\mcal{F}^S(\tilde{\mbf{p}})$. The first row block produces $AA$ offspring by self-fertilization from $AA$ maternal parents. Since all gametes, male or female, of $AA$ individuals carry an $A$ allele, all offspring resulting from self-fertilization will have $AA$ genotypes, and a fraction $C_{AA}$ of ovules are fertilized by selfing. The second and third row blocks are $0$ because $AA$ individuals produce no $a$ bearing gametes. Next consider the second column block. The first and third row blocks produce $AA$ and $aa$ offspring from $Aa$ maternal parents, respectively. With mendelian segregation, this occurs with frequency $1/4$, multiplied by the fraction of ovules that are self-fertilized, $C_{Aa}$ and $C_{aa}$, respectively. The second row block produces $Aa$ offspring by selfing from $Aa$ maternal parents, which will occur with frequency $1/2$ multiblied by the fraction of ovules selfed, $C_{Aa}$.

The blocks of $\mcal{F}^X(\tilde{\mbf{p}})$ can be constructed and interpreted similarly. The first row block of the first column produces $AA$ offspring by outcrossing from $AA$ maternal parents. This happens when the $AA$ maternal parent receives a random gamete from the male gamete pool, which happens with probability $q^{\prime}_{A}$. The other blocks can be interpreted similarly.

Combining all the component matrices yields the overall eco-evolutionary projection matrix:

\begin{align*}
	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
		& \underbrace{\left(
			\begin{array}{ccc|ccc}
				\mcal{U}^S_{AA} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				\mbf{0} & \mcal{U}^S_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mcal{U}^S_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
				\mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{AA} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{aa} & \mbf{0}\\
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{aa}
			\end{array} \right)}_{\tilde{\mbf{U}}} \\ 
			&+ \underbrace{\left(
			\begin{array}{ccc|ccc}
				C_{AA} (1 - \delta) \mbf{F}_{AA} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & 0 & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				0 & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & 0 & \mbf{0} & \mbf{0} & \mbf{0} \\
				0 & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & C_{aa} (1 - \delta) \mbf{F}_{aa} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
				q^{\prime}_{A} \mbf{F}_{AA} (1 - C_{AA})& \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} (1 - C_{Aa}) & 0 & \mbf{0} & \mbf{0} & \mbf{0} \\
				q^{\prime}_{a} \mbf{F}_{AA} (1 - C_{AA}) & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & q^{\prime}_{A} \mbf{F}_{aa} (1 - C_{aa}) & \mbf{0} & \mbf{0} & \mbf{0}\\
				0 & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} (1 - C_{Aa})& q^{\prime}_{a} \mbf{F}_{aa} (1 - C_{aa}) & \mbf{0} & \mbf{0} & \mbf{0}
			\end{array} \right)}_{\tilde{\mbf{F}}} \numberthis
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[htbp]
\centering
\caption{\bf Definition of terms.}
\begin{tabular}{ l l r }
 \toprule
	Symbol & Definition & Dimension \\
\hline
$a$      & number of alleles ($2$; $A$ and $a$) & \\
$g$      & number of genotypes ($3$; $AA$, $Aa$, and $aa$) & \\
$\omega$ & number of stages ($2$; juvenile and adult) & \\
$N$      & Total population size & \\
$N_b$    & Breeding population size & \\
$\mbf{c}_i$ & Indicator vector for breeding stages in genotype $i$ & $\omega \times 1$ \\
$\tilde{\mbf{n}}$ & Joint stage-genotype vector & $\omega g \times 1$ \\
$\tilde{\mbf{p}}$ & Joint stage-genotype frequency vector & $\omega g \times 1$ \\
$\mbf{p}_i$ & Genotype frequency vector in genotype $i$ & $g \times 1$ \\
$\mbf{q}_i$ & Gene frequency vector in genotype $i$ & $g \times 1$ \\
$\mbf{q}_b$ & Genotype frequency vector in genotype $i$ among breeding individuals & $g \times 1$ \\
$\mbf{W}$   & Allele segregation matrix & $\omega \times g$ \\
\hline
\end{tabular}
\label{tab:Definitions}\\
\end{table}
\newpage{}

%%%%%%%%%%%%%%%%%%%%%%%%
\section*{A working two-stage model for hermaphrodites}
\setcounter{equation}{0}

As a first pass at implementing the model, we construct and analyze a genotype $\times$stage-classified model for a hypothetical species with intralocus sexual conflict via the two sex functions. Our hypothetical species has two life stages: juveniles and adults. Suppose that allele $A$ is beneficial for females but detrimental for males, and that allele $a$ has the reverse effect. We also assume that the effects act only during the adult stage through reproductive success. The allele does not affect survival and transition rates, such that

\begin{equation}
	\mbf{U} = \left(
					\begin{array}{cc}
						\sigma_J(1 - \gamma) & 0 \\
						\sigma_J \gamma      & \sigma_A
					\end{array}
				\right),
\end{equation}

\noindent where $\sigma_J$ and $\sigma_A$ are the juvenile and adult stage survival rates, and $\gamma$ is the maturation rate from juvenile to adult stages. Any or all of these parameters may differ among genotypes, so that selection can operate on stage-specific viability, development, and/or fertility. For simplicity, we assume these are constant among genotypes.

The fertility matrices through male and female function are

\begin{equation}
	\mbf{F}_{ii} = \left(
					\begin{array}{cc}
						0 & f_{ii} \\
						0 & 0
					\end{array}
				\right),
\end{equation}
\noindent and
\begin{equation}
	\mbf{F}^{\prime}_{ii} = \left(
					\begin{array}{cc}
						0 & f^{\prime}_{ii} \\
						0 & 0
					\end{array}
				\right),
\end{equation}

\noindent where $f$ is the genotype specific adult fertility though female function, and $f^{\prime}_{ii}$ is the same through male function.

Note, that under a model of constant selfing (where $C_{AA} = C_{Aa} = C_{aa} = C$), the model reduces to

\begin{equation}
	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{U}^S & \mbf{0} \\ \hline
				\mbf{0} & \mcal{U}^X \\
			\end{array} \right)}_{\tilde{\mbf{U}}} + 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) & \mbf{0} \\ \hline
				\mbf{0} & \mcal{F}^X(\tilde{\mbf{p}}) \\
			\end{array} \right)}_{\tilde{\mbf{F}}}
\end{equation}

\begin{table}[htbp]
 \centering
 \caption{\bf Fitness expressions for Sexually Antagonistic selection in hermaphrodites}
\begin{tabular}{lccc}
 \toprule
					&  \multicolumn{3}{c}{{\textit{Genotype}}} \\ 
\cline{2-4}
					& $AA$		& $Aa$ 			& $aa$ 		\\ \hline
Female function:	& $f$		& $f - h_f s_f$	& $f - s_f$ \\	
Male function:		& $f - s_m$	& $f - h_m s_m$	& $f$ 		\\	
\hline
\end{tabular}
\end{table}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Coexistence conditions}
\setcounter{equation}{0}

Here I'm trying to derive the conditions for a protected polymorphism in the hermaphrodite model with sex-specific selection and partial selfing. Following de Vries \& Caswell (2019), we reorder the population vector by genotype first, and then by sex-function and stage, in contrast to the ordering used above (sex-function, then genotype, then stage). The first step is to construct the population projection matrix and then use it to derive coexistence conditions, following closely the logic of Appendix A of de Vries and Caswell (2019).

\subsubsection*{Population projection matrix}

The first relevant difference between the two-sex model of de Vries \& Caswell (2019) is that population vector, which is

\begin{equation}
	\tilde{\mbf{n}} = \left(
			\begin{array}{c}
							\mbf{n}_{AA} \\
							\mbf{n}_{AA} \\ \hline
							\mbf{n}_{Aa} \\
							\mbf{n}_{Aa} \\ \hline
							\mbf{n}_{aa} \\
							\mbf{n}_{aa} \\
			\end{array} \right).
\end{equation}

\noindent Note that to 'doubling' up of the genotype-specific vectors in each block is done out of convenience, because the population projection matrix is organized around two sex-functions.

The population projection matrix $\tilde{\mbf{A}}$ consists of $3 \times 3$ blocks, which act on the genotype specific population vectors:

\begin{align*} \label{eq:AtildeCoexistence}
	\tilde{\mbf{A}}[\tilde{\mbf{n}}] &= \tilde{\mbf{U}} + \tilde{\mbf{F}} \\
		&= \left(
			\begin{array}{cc|cc|cc}
				\mcal{U}^S_{AA} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				\mbf{0} & \mcal{U}^X_{AA} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\\hline
				\mbf{0} & \mbf{0} & \mcal{U}^S_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{Aa} & \mbf{0} &  \mbf{0}\\\hline
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^S_{Aa} & \mbf{0} \\ 
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{Aa}
			\end{array} \right) \\ 
		&+ \left(
			\begin{array}{cc|cc|cc}
				C_{AA} (1 - \delta) \mbf{F}_{AA} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				q^{\prime}_{A} \mbf{F}_{AA} (1 - C_{AA}) & \mbf{0} & \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} (1 - C_{Aa}) & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
				\mbf{0} & \mbf{0} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\
				q^{\prime}_{a} \mbf{F}_{AA} (1 - C_{AA}) & \mbf{0} & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & \mbf{0} & q^{\prime}_{A} \mbf{F}_{aa} (1 - C_{aa}) & \mbf{0}\\ \hline
				\mbf{0} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & C_{aa} (1 - \delta) \mbf{F}_{aa} & \mbf{0} \\ 
				\mbf{0} & \mbf{0} & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} (1 - C_{Aa})  & \mbf{0} & q^{\prime}_{a} \mbf{F}_{aa} (1 - C_{aa}) & \mbf{0}
			\end{array} \right) \numberthis
\end{align*}

\noindent with symbols defined as above. The survival matrices appear on the diagonal because individuals do not change their genotype once they are born. The fertility matrix incorporates the Mendelian inheritance and is an extension of the fertility matrix derived above after de Vries \& Caswell (2019).

The first block column of $\tilde{\mbf{A}}$ describes the production of offspring by and $AA$ female with stage-specific fertility rates $\mbf{F}_{AA}$ by both selfing and outcrossing. The probability of picking an $A$ allele out of the pool of available male gametes. When reproducing by selfing, this is entirely determined by the probability of sampling an $A$ allele after Mendelian segregation. For outcross reproduction the probability, and hence the probability of this $AA$ female producing an $AA$ offspring, is $q^{\prime}_{A}$, as derived above. Conversely, the probability of picking an $a$ allele and producing an $Aa$ offspringis zero for an $AA$ female when selfing, but $q^{\prime}_a$ for outcrossing. Similarly, the middle column of block matrices are offspring produced by $Aa$ females, which can produce offspring of all $3$ genotypes.

\subsubsection*{Coexistence conditions}

The two-sex Mendelian matrix model defined by above reduces to a linear matrix model on the boundary where the population is initially fixed for the $A$ allele (since $q^{\prime}_{A} = 1$ and $q^{\prime}_{a} = 0$). Provided the initial population contains a nonzero number of females, the population will grow or shrink exponentially after converging to a stable population structure (see Caswell (2001), section 4.5.2.1). We can take advantage of the homogeneity of $\tilde{\mbf{F}}$, we rewrite the model in terms of the normalized population vector (the frequency vector):

\begin{equation} 
	\tilde{\mbf{p}}(t + 1) = \frac{ \tilde{\mbf{A}}[\tilde{\mbf{p}}(t)] \tilde{\mbf{p}}(t) }{ \| \tilde{\mbf{A}}[\tilde{\mbf{p}}(t)] \tilde{\mbf{p}}(t) \|},
\end{equation}

\noindent where $\| \mbf{a} \|$ denotes the $1$-norm of the vector $\mbf{a}$, defined as the sum of the absolute values of the entries of the vector $\mbf{a}$. Equilibrium solutions, denoted by $\hat{\mbf{p}}$, satisfy

\begin{equation} 
	\hat{\mbf{p}} = \frac{ \tilde{\mbf{A}}[\hat{\mbf{p}}] \hat{\mbf{p}} }{ \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}}[\hat{\mbf{p}}] \hat{\mbf{p}} },
\end{equation}

\noindent where the one norm can be replaced by $\mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}}[\hat{\mbf{p}}] \hat{\mbf{p}}$ because $\hat{\mbf{p}}$ is nonnegative.

\subsubsection*{Linearization at the boundary equilibrium}

In this section I'm trying to follow Lotte's derivation of the linear approximation to the dynamics in the neighborhood of a homozgote boundary equilibrium. The stability of a such an equilibrium to invasions by the other allele is determined by the magnitude of the largest eigenvalue of the Jacobian matrix of the frequency model evaluated at the equilibrium. If the magnitude of this eigenvalue is larger than one, then the equilibrium is unstable. The Jacobian matrix,

\begin{equation} \label{eq:genJacobian}
	\mbf{M} =  \frac{ \text{d} \tilde{\mbf{p}}(t + 1) }{ \text{d} \tilde{\mbf{p}}(t) } \bigg\rvert_{\hat{\mbf{p}}},
\end{equation}

\noindent is obtained by differentiating Eq(\ref{eq:genJacobian}) and evaluating the resulting derivatives at the boundary equilibrium. This requires a long series of matrix calculus operations, and repeatedly takes advantage of the fact that $\hat{\mbf{p}}$ at the boundary contains zeros for the blocks corresponding to $Aa$ and $aa$ genotypes.

For notational convenience, first define a matrix $\mbf{B}$ as

\begin{equation}
	\mbf{B}[\tilde{\mbf{p}}] = \frac{ \tilde{\mbf{A}}[\tilde{\mbf{p}}] }{ \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}}[\tilde{\mbf{p}}] \tilde{\mbf{p}}},
\end{equation}

\noindent such that

\begin{equation} 
	\tilde{\mbf{p}}(t + 1) = \mbf{B}[\tilde{\mbf{p}}(t)]\tilde{\mbf{p}}(t).
\end{equation}

\noindent We differentiate to obtain

\begin{equation} 
	\text{d} \tilde{\mbf{p}}(t + 1) = \mbf{B}\text{d} \tilde{\mbf{p}}(t) + \left(\text{d} \tilde{\mbf{B}} \right) \tilde{\mbf{p}}(t),
\end{equation}

\noindent where the explicit dependence of $\mbf{B}$ on $\tilde{\mbf{p}}$ has been omitted to avoid a cluttering of brackets. Multiply the second term by an $2 \omega g \times 2 \omega g$ identity matrix,

\begin{equation} 
	\text{d} \tilde{\mbf{p}}(t + 1) = \mbf{B}\text{d} \tilde{\mbf{p}}(t) + \mbf{1}_{2 \omega g} \left(\text{d} \tilde{\mbf{B}} \right) \tilde{\mbf{p}}(t).
\end{equation}

\noindent and apply the vector operator to both sides, remembering that as $\tilde{\mbf{p}}$ is a vector, $\text{vec}\tilde{\mbf{p}} = \tilde{\mbf{p}}$,

\begin{equation} 
	\text{d} \tilde{\mbf{p}}(t + 1) = \mbf{B}\text{d} \tilde{\mbf{p}}(t) + \text{vec} \left[ \mbf{1}_{2 \omega g} \left(\text{d} \tilde{\mbf{B}} \right) \tilde{\mbf{p}}(t) \right].
\end{equation}

\noindent Next, we can apply Roth's theorem (Roth 1934), $\text{vec}\mbf{ABC} = \left( \mbf{C}^{\intercal} \otimes \mbf{A} \right) \text{vec}\mbf{B}$, to replace the the $\text{vec}$ operator with the Kronecer product:

\begin{equation} 
	\text{d} \tilde{\mbf{p}}(t + 1) = \mbf{B}\text{d} \tilde{\mbf{p}}(t) + \left( \tilde{\mbf{p}}^{\intercal}(t) \otimes \mbf{1}_{2 \omega g} \right) \text{d}\,\text{vec}\mbf{B}.
\end{equation}

\noindent Then the first identification theorem and the chain rule together give the following formula for the Jacobian (Magnus and Neudecker, 1985; Caswell, 2007),

\begin{align*} \label{eq:genJacobian}
	\mbf{M} &=  \frac{ \text{d} \tilde{\mbf{p}}(t + 1) }{ \text{d} \tilde{\mbf{p}}(t) } \bigg\rvert_{\hat{\mbf{p}}}, \\
			&= \mbf{B}[\tilde{\mbf{p}}] + \left( \tilde{\mbf{p}}^{\intercal} \otimes \mbf{1}_{2 \omega g} \right) \frac{ \partial \text{vec}{\mbf{B}[\tilde{\mbf{p}}]} }{ \partial \tilde{\mbf{p}}^{\intercal} } \bigg\rvert_{\hat{\mbf{p}}},
\end{align*}

Our aim is to express the Jacobian matrix $\mbf{M}$ in terms of the genotype specific matrices, $\mbf{U}^S_i$, $\mbf{U}^X_i$, $\mbf{F}^S_i$, $\mbf{F}^X_i$. We analyze the Jacobian at the $AA$ boundary; the expression at the $aa$ boundary can be derived afterwards using symmetry arguments. First it will be convenient to define the scalar $f(\tilde{\mbf{p}})$ as

\begin{equation}
	f(\tilde{\mbf{p}}) = \frac{1}{ \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}}[\tilde{\mbf{p}}] \tilde{\mbf{p}}},
\end{equation}

\noindent so that

\begin{equation}
	\mbf{B}[\tilde{\mbf{p}}] = f(\tilde{\mbf{p}}) \tilde{\mbf{A}}[\tilde{\mbf{p}}].
\end{equation}

\noindent Note that where it does not create confusion, we omit the explicit dependence on $\tilde{\mbf{A}}$, $\mbf{B}$, and $f$ on $\tilde{\mbf{p}}$. Next, we take the vec of both sides of equation (A-16) and differentiate to obtain

\begin{equation} \label{eq:dvecB}
	\text{dvec} \mbf{B} = \text{vec} \tilde{\mbf{A}} \text{d} f + f \text{dvec} \tilde{\mbf{A}},
\end{equation}

\noindent or

\begin{equation}
	\frac{\partial \text{vec} \mbf{B}} {\partial \tilde{\mbf{p}}^{\intercal}} = \text{vec} \tilde{\mbf{A}} \frac{\partial f}{\partial \tilde{\mbf{p}}^{\intercal} } + f(\tilde{\mbf{p}}) \frac{ \partial \text{vec} \tilde{\mbf{A}} }{ \partial \tilde{\mbf{p}}^{\intercal} }.
\end{equation}

\noindent Next differentiate $f$ to obtain

\begin{equation}
	\text{d} f = \frac{ -1 }{ \left( \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}}[\tilde{\mbf{p}}] \tilde{\mbf{p}} \right)^2 } \left[ \mbf{1}^{\intercal}_{2 \omega g} \left( \text{d} \tilde{\mbf{A}} \right) \tilde{\mbf{p}} + \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}} \text{d} \tilde{\mbf{p}} \right].
\end{equation}

\noindent All the terms in the Jacobian are evaluated at the $AA$ boundary, which simplifies some of the equations, e.g., $\tilde{\mbf{A}}[ \hat{\mbf{p}} ] \hat{\mbf{p}} = \lambda_{AA} \hat{\mbf{p}}$, and therefore

\begin{equation} \label{eq:lambdaAA}
	\mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}}[\tilde{\mbf{p}}] \tilde{\mbf{p}} = \lambda_{AA}.
\end{equation}

Evaluate the differential of $f$ at the boundary and use Eq(\ref{eq:lambdaAA}) to obtain

\begin{equation} \label{eq:df}
	\text{d} f(\tilde{\mbf{p}}) = \frac{ -1 }{\lambda_{AA}^2 } \left[ \mbf{1}^{\intercal}_{2 \omega g} \left( \text{d} \tilde{\mbf{A}} \right) \hat{\mbf{p}} + \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}} \text{d} \tilde{\mbf{p}} \right] \bigg\rvert_{\tilde{\mbf{p}}}.
\end{equation}

The first term in this sum, $\mbf{1}^{\intercal}_{2 \omega g} \left( \text{d} \tilde{\mbf{A}} \right) \hat{\mbf{p}}$, is equal to zero because
\begin{align*}
	q^{\prime}_A + q^{\prime}_a &= 1 \\
	\text{d}q^{\prime}_A + \text{d}q^{\prime}_a &= 0 \\
\end{align*}
\noindent and
\begin{align*}
	q_A + q_a &= 1 \\
	\text{d}q_A + \text{d}q_a &= 0 \\
\end{align*}

\noindent and therefore every column in $\left( \text{d} \tilde{\mbf{A}} \right)$ sums to zero, see Eq(\ref{eq:AtildeCoexistence}).

Substituting Eq(\ref{eq:df}) into Eq(\ref{eq:dvecB}) and evaluating at the boundary yields

\begin{equation} \label{eq:dvecBsubs}
	\text{dvec} \mbf{B} = \frac{ -1 }{\lambda_{AA}^2 } \text{vec} \tilde{\mbf{A}} \left[ \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}} \text{d} \tilde{\mbf{p}} \right] + \frac{ 1 }{\lambda_{AA} } \text{dvec} \tilde{\mbf{A}},
\end{equation}

\noindent or

\begin{equation}
	\frac{\partial \text{vec} \mbf{B}} {\partial \tilde{\mbf{p}}^{\intercal}} \bigg\rvert_{\hat{\mbf{p}}} = \frac{ -1 }{\lambda_{AA}^2 } \left( \text{vec} \tilde{\mbf{A}} \right) \left( \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}} \text{d} \tilde{\mbf{p}} \right) + \frac{ 1 }{\lambda_{AA} } \left. \frac{\partial \text{vec} \mbf{A}} {\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}
\end{equation}

\noindent Finally, substituting the expression above into Eq(\ref{eq:genJacobian}) yields the Jacobian matrix:

\begin{align*} \label{eq:subJacobian}
	\mbf{M} &=  \mbf{B}[\tilde{\mbf{p}}] + \left( \tilde{\mbf{p}}^{\intercal} \otimes \mbf{1}_{2 \omega g} \right) \frac{ \partial \text{vec}{\mbf{B}} }{ \partial \tilde{\mbf{p}}^{\intercal} } \bigg\rvert_{\hat{\mbf{p}}}, \\
			&= \underbrace{ \mbf{B}[\hat{\mbf{p}}]}_{\encircle{A}} - 
			   \underbrace{ \frac{ 1 }{\lambda_{AA}^2 } (\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{\omega g}) \left( \text{vec} \tilde{\mbf{A}} \right) \left( \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}} \right)}_{\encircle{B}} \\
			&~~~~~~~~~~~+ \underbrace{\frac{ 1 }{\lambda_{AA} } (\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{\omega g}) \left. \frac{\partial \text{vec} \tilde{\mbf{A}}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}}_{\encircle{C}},
\end{align*}

%%%%%%%%%%%%%%%%
\subsubsection{Components of the Jacobian}

The next task is to work out each of the three terms \encircle{A}, \encircle{B}, and \encircle{C}, in $\mbf{M}$. Beginning with \encircle{A}, we have

\begin{align*}
	\mbf{B}[\hat{\mbf{p}}] &= \frac{ \tilde{\mbf{A}}[\hat{\mbf{p}}] }{ \mbf{1}^{\intercal}_{\omega g} \tilde{\mbf{A}}[\hat{\mbf{p}}] \hat{\mbf{p}} } = \\
						   \frac{1}{\lambda_{AA}} &
						   \left(
			\begin{array}{cc|cc|cc}
				\mcal{U}^S_{AA} + C_{AA} (1 - \delta) \mbf{F}_{AA} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				\mbf{F}_{AA} (1 - C_{AA}) & \mcal{U}^X_{AA} & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
				\mbf{0} & \mbf{0} & \mcal{U}^S_{Aa} + \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & \mcal{U}^X_{Aa} & \mbf{F}_{aa} (1 - C_{aa}) & \mbf{0}\\ \hline
				\mbf{0} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mcal{U}^S_{aa} + C_{aa} (1 - \delta) \mbf{F}_{aa} & \mbf{0} \\ 
				\mbf{0} & \mbf{0} & \mbf{0}  & \mbf{0} & \mbf{0} & \mcal{U}^X_{aa}
			\end{array} \right) \numberthis
\end{align*}

\noindent see Eq{\ref{eq:AtildeCoexistence}}. Next, we have \encircle{B},

\begin{equation}
	\encircle{B} = - \frac{1}{\lambda_{AA}} \left( \hat{\mbf{p}}^{\intercal} \otimes \mbf{1}_{2 \omega g} \right) \left( \text{vec} \tilde{\mbf{A}} \right) \left( \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}} \right).
\end{equation}

\noindent Again using Roth's theorem (Roth 1934), $\left( \mbf{C}^{\intercal} \otimes \mbf{A} \right) \text{vec}\mbf{B} = \text{vec}\mbf{ABC}$, we can simplify as follows:

\begin{align*}
	\left( \hat{\mbf{p}}^{\intercal} \otimes \mbf{1}_{2 \omega g} \right) \left( \text{vec} \tilde{\mbf{A}}[\hat{\mbf{p}}] \right) &= \text{vec} \left( \mbf{I}_{2 \omega g}\, \tilde{\mbf{A}}[\hat{\mbf{p}}]\, \hat{\mbf{p}} \right) \\
					   &= \lambda_{AA} \hat{\mbf{p}}, \numberthis
\end{align*}

\noindent so that

\begin{landscape}

\begin{equation}
	\encircle{B} = - \frac{1}{\lambda_{AA}} \hat{\mbf{p}} \left( \text{vec} \tilde{\mbf{A}} \right) \left( \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}}[\hat{\mbf{p}}] \right).
\end{equation}

\noindent Substitute the population vector at the $AA$ boundary,

\begin{equation}
	\hat{\mbf{p}} = \left(
			\begin{array}{c}
							\hat{\mbf{p}}_{AA} \\
							\hat{\mbf{p}}_{AA} \\ \hline
							0                  \\
							0                  \\ \hline
							0                  \\
							0                  \\
			\end{array} \right)
\end{equation}

\noindent and rewrite in terms of the block matrices to obtain

\begin{align*}
	\encircle{B} =& \\
	&-\frac{1}{\lambda_{AA}} \left(
			\begin{array}{cc|cc|cc}
				\hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{aa} \\ 
				\hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{aa} \\ \hline
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} 
			\end{array} \right) \\
	&-\frac{1}{\lambda_{AA}} \left(
			\begin{array}{cc|cc|cc}
				C_{AA}(1 - \delta) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & \mbf{0} & C_{AA}(1 - \delta) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & \mbf{0} & C_{AA}(1 - \delta) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} & \mbf{0} \\ 
				(1 - C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & \mbf{0} & (1 - C_{Aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & \mbf{0} & (1 - C_{Aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} & \mbf{0} \\ \hline
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0}
			\end{array} \right) \numberthis			
\end{align*}

\end{landscape}

Finally, we turn to term \encircle{C} of the Jacobian. As far as I can tell, we can follow the steps leading from Eq(A-35) through to Eq(A-46) of de Vries \& Caswell (2019) Appendix A exactly because the structure of our $\tilde{\mbf{A}}$ equations are identical. We resume the derivation at our version of their Eq(A-47)...Substitute $\hat{\mbf{p}}^{\intercal} = \left( \hat{\mbf{p}}^{\intercal}_{AA},\hat{\mbf{p}}^{\intercal}_{AA},\mbf{0},\mbf{0},\mbf{0},\mbf{0} \right)$ into the right side of Eq(A-46) so that only terms with $j=1$ and $j=2$ are nonzero, yielding

\begin{align*}
	\frac{1}{\lambda_{AA}}(\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{2 \omega g}) \left. \frac{ \partial \text{vec}\mbf{A}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} &= \frac{1}{\lambda_{AA}} \sum^{6}_{i=1} \big( \hat{\mbf{p}}^{\intercal}_{AA} \otimes (\mbf{e}_{i} \otimes \mbf{I}_{\omega}) \big) \left. \frac{ \partial \text{vec}\mbf{A}_{i,1}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} \\
		&+ \frac{1}{\lambda_{AA}} \sum^{6}_{i=1} \big( \hat{\mbf{p}}^{\intercal}_{AA} \otimes (\mbf{e}_{i} \otimes \mbf{I}_{\omega}) \big) \left. \frac{ \partial \text{vec}\mbf{A}_{i,2}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} \numberthis
\end{align*}

\noindent But, since none of the $\mbf{A}_{i,2}$ are a function of the frequency vector $\mbf{p}$,

\begin{equation}
	\left. \frac{ \partial \text{vec}\mbf{A}_{i,2}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} = 0,\,\text{for all }i. 
\end{equation}

\noindent (see Eq{\ref{eq:AtildeCoexistence}}). Next, write down each term in the sum over $i$ and take the derivative  of the $\text{vec}\mbf{A}_{i,1}$'s to obtain

\begin{align*}
	\frac{1}{\lambda_{AA}}(\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{2 \omega g}) \left. \frac{ \partial \text{vec}\mbf{A}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} = 
		&\frac{C_{AA}(1 - \delta)}{\lambda_{AA}} \big[ \hat{\mbf{p}}^{\intercal} \otimes (\mbf{e}_{1} \otimes \mbf{I}_{\omega} ) - \hat{\mbf{p}}^{\intercal}_{AA} \otimes (\mbf{e}_{3} \otimes \mbf{I}_{\omega})\big] \text{vec}\mbf{F}_{AA} \\
		&+ \frac{(1 - C_{AA})}{\lambda_{AA}} \big[ \hat{\mbf{p}}^{\intercal} \otimes (\mbf{e}_{2} \otimes \mbf{I}_{\omega} ) - \hat{\mbf{p}}^{\intercal}_{AA} \otimes (\mbf{e}_{4} \otimes \mbf{I}_{\omega})\big] \text{vec}\mbf{F}_{AA}  \left.\frac{\partial q^{\prime}_{A}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}    \numberthis
\end{align*}

\noindent Finally, apply Roth's theorem (Roth 1934), $\left( \mbf{C}^{\intercal} \otimes \mbf{A} \right) \text{vec}\mbf{B} = \text{vec}\mbf{ABC}$ (e.g.,where $\mbf{C}^{\intercal} = \hat{\mbf{p}}^{\intercal}_{AA}$, $\mbf{A} = (\mbf{e}_1 \otimes \mbf{I}_{\omega})$, and $\text{vec}\mbf{B} = \text{vec}\mbf{F}_{AA}$) to rewrite as

\begin{align*}
	\frac{1}{\lambda_{AA}}(\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{2 \omega g}) \left. \frac{ \partial \text{vec}\mbf{A}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} = 
		&\frac{C_{AA}(1 - \delta)}{\lambda_{AA}} \big[ \text{vec}\big((\mbf{e}_{1} \otimes \mbf{I}_{\omega}) \mbf{F}_{AA} \hat{\mbf{p}}_{AA} \big) - \text{vec}\big((\mbf{e}_{3} \otimes \mbf{I}_{\omega}) \mbf{F}_{AA} \hat{\mbf{p}}_{AA} \big) \big]  \\
		&+ \frac{(1 - C_{AA})}{\lambda_{AA}} \big[ \text{vec}\big((\mbf{e}_{2} \otimes \mbf{I}_{\omega}) \mbf{F}_{AA} \hat{\mbf{p}}_{AA} \big) - \text{vec}\big((\mbf{e}_{4} \otimes \mbf{I}_{\omega}) \mbf{F}_{AA} \hat{\mbf{p}}_{AA} \big) \big]  \left.\frac{\partial q^{\prime}_{A}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}    \numberthis
\end{align*}

\noindent written in terms of the block matrices, this yields

\begin{align*}
	\encircle{C} =& \\
	\frac{1}{\lambda_{AA}} &\left(
			\begin{array}{cc|cc|cc}
				\mbf{0} & C_{AA}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & C_{aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \\
				\mbf{0} & (1 - C_{AA}) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\prime \intercal}_{AA}} & \mbf{0} & (1 - C_{Aa}) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\prime \intercal}_{Aa}} & \mbf{0} & (1 - C_{aa}) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\prime \intercal}_{aa}} \\ \hline
				\mbf{0} & \mbf{0} & \mbf{0} & -C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & \mbf{0} \\
				\mbf{0} & -(1 - C_{AA}) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\prime \intercal}_{AA}} & \mbf{0} & -(1 - C_{Aa}) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\prime \intercal}_{Aa}} & \mbf{0} & -(1 - C_{aa}) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\prime \intercal}_{aa}} \\ \hline
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} 
			\end{array} \right)  \numberthis			
\end{align*}

The above expression requires the derivative of the frequency of allele $A$ in the gamete pool with respect to the population frequency vector:

\begin{equation}
	\left. \frac{\partial q^{\prime}_{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{P}}}
\end{equation}

\noindent Here, I believe we can again follow the exact same steps as described in Eq(A-50) through Eq(A-60) in Appendix A of de Vries \& Caswell (2019) to give, finally,

\begin{align*}
	\frac{1}{\lambda_{AA}} (\hat{\mbf{p}}^{\intercal} &\otimes \mbf{I}_{2 \omega g}) \left. \frac{\partial \text{vec}\mbf{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} = \\
	\frac{1}{\lambda_{AA}} &\left(
			\begin{array}{cc|cc|cc}
				\mbf{0} & C_{AA}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & C_{aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \\
				\mbf{0} & \mbf{0} & \mbf{0} & -\frac{(1 - C_{Aa})}{2 p_{\text{n}}} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \mbf{0} & -\frac{(1 - C_{aa})}{2 p_{\text{n}}} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{aa} \\ \hline
				\mbf{0} & \mbf{0} & \mbf{0} & -C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mbf{0} & \frac{(1 - C_{Aa})}{2 p_{\text{n}}} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \mbf{0} & \frac{(1 - C_{aa})}{2 p_{\text{n}}} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{aa} \\ \hline
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} 
			\end{array} \right)  \numberthis			
\end{align*}

%%%%%%%%%%%%%%%%%%%%
\subsubsection{The Jacobian}

Putting all the pieces together into $M$, we get the Jacobian:

\begin{landscape}

\begin{align*} \label{eq:combinedM}
	\mbf{M} = 
		\frac{1}{\lambda_{AA}} &\left(\begin{array}{cc|cc|cc}
				\mcal{U}^S_{AA} + C_{AA} (1 - \delta) \mbf{F}_{AA} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				\mbf{F}_{AA} (1 - C_{AA}) & \mcal{U}^X_{AA} & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
				\mbf{0} & \mbf{0} & \mcal{U}^S_{Aa} + \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & \mcal{U}^X_{Aa} & \mbf{F}_{aa} (1 - C_{aa}) & \mbf{0}\\ \hline
				\mbf{0} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mcal{U}^S_{aa} + C_{aa} (1 - \delta) \mbf{F}_{aa} & \mbf{0} \\ 
				\mbf{0} & \mbf{0} & \mbf{0}  & \mbf{0} & \mbf{0} & \mcal{U}^X_{aa}
			\end{array} \right) \\
		-\frac{1}{\lambda_{AA}} &\left(
			\begin{array}{cc|cc|cc}
				\hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{aa} \\ 
				\hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{aa} \\ \hline
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} 
			\end{array} \right) \\
		-\frac{1}{\lambda_{AA}} &\left(
			\begin{array}{cc|cc|cc}
				C_{AA}(1 - \delta) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & \mbf{0} & C_{Aa}(1 - \delta) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & \mbf{0} & C_{aa}(1 - \delta) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} & \mbf{0} \\ 
				(1 - C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & \mbf{0} & (1 - C_{Aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & \mbf{0} & (1 - C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} & \mbf{0} \\ \hline
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0}
			\end{array} \right) \\
		\frac{1}{\lambda_{AA}} &\left(
			\begin{array}{cc|cc|cc}
				\mbf{0} & C_{AA}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & C_{aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \\
				\mbf{0} & \mbf{0} & \mbf{0} & -\frac{(1 - C_{Aa})}{2 p_{\text{n}}} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \mbf{0} & -\frac{(1 - C_{aa})}{2 p_{\text{n}}} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{aa} \\ \hline
				\mbf{0} & \mbf{0} & \mbf{0} & -C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mbf{0} & \frac{(1 - C_{Aa})}{2 p_{\text{n}}} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \mbf{0} & \frac{(1 - C_{aa})}{2 p_{\text{n}}} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{aa} \\ \hline
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} 
			\end{array} \right)  
			\numberthis
\end{align*}

\end{landscape}

%%%%%%%%%%%%%%%
\subsubsection*{Eigenvalues of the Jacobian}

Similar to Lotte's result, the Jacobian matrix given in Eq(\ref{eq:combinedM}) is upper block triangular, and the eigenvalues of $\mbf{M}$ are therefore the eigenvalues of the diagonal blocks. However, because of the way that self-fertilization alters the projection matrix, our blocks are slightly different than Lotte's model. Specifically, our $\mbf{M}_{11}$ is the same (corresponding to the upper-left $2 \times 2$  block of $\mbf{M}$); our $\mbf{M}_{33}$ corresponds to the lowermost right corner element ($\mbf{M}_{33} = \mathcal{U}^{X}_{aa}$. However, in our case, the central diagonal block $\mbf{M}_{22}$ corresponds to

 \begin{align*} \label{eq:M22}
	\mbf{M}_{22} = 
		\frac{1}{\lambda_{AA}} &\left(\begin{array}{cc|c}
				\mcal{U}^S_{Aa} + \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & -C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} \\
				\frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & \mcal{U}^X_{Aa} + \frac{(1 - C_{Aa})}{2 p_n} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \mbf{F}_{aa} (1 - C_{aa}) \\ \hline
				\frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mcal{U}^S_{aa} + C_{aa} (1 - \delta) \mbf{F}_{aa} \\ 
			\end{array} \right) \numberthis
\end{align*}

For the same reasons outlined in Lotte's appendix, the stability of $\hat{\mbf{p}}$ will be determined by $\mbf{M}_{22}$. The largest absolute value of the eigenvalues of the Jacobian matrix, the leading eigenvalue, evaluated at the $AA$ boundary, denoted by $\tilde{\zeta}_{AA}$, is therefore

\begin{align*} \label{eq:eigAA}
	\tilde{\zeta}_{AA} = 
		\frac{1}{\lambda_{AA}} \rho&\left(\begin{array}{cc|c}
				\mcal{U}^S_{Aa} + \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & -C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} \\
				\frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & \mcal{U}^X_{Aa} + \frac{(1 - C_{Aa})}{2 p_n} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \mbf{F}_{aa} (1 - C_{aa}) \\ \hline
				\frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mcal{U}^S_{aa} + C_{aa} (1 - \delta) \mbf{F}_{aa} \\ 
			\end{array} \right). \numberthis
\end{align*}

\noindent Rearranging the projection matrix so that outcrossing comes on top of selfing in each block simplifies things, but in the end the leading eigenvalue evaluated at the $aa$ boundary is

\begin{align*} \label{eq:eigaa}
	\tilde{\zeta}_{aa} = 
		\frac{1}{\lambda_{AA}} \rho&\left(\begin{array}{c|cc}
				\mcal{U}^S_{AA} + C_{AA} (1 - \delta) \mbf{F}_{AA} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}\\ \hline
				\mbf{F}_{AA} (1 - C_{AA}) & \mcal{U}^X_{Aa} + \frac{(1 - C_{Aa})}{2 p_n} (\mbf{F}_{aa} \hat{\mbf{p}}_{aa}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) \\ 
				\mbf{0} & -C_{Aa}(1 - \delta) (\mbf{F}_{aa} \hat{\mbf{p}}_{aa}) & \mcal{U}^S_{Aa} + \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} \\
			\end{array} \right). \numberthis
\end{align*}

A boundary equilibrium is unstable to invasion by the rare allele if the leading eigenvalue of the Jacobian evaluated at the equilibrium is greater than $1$. The conditions for a protected polymorphism is that both boundaries are unstable (i.e., both $\tilde{\zeta}_{AA} > 1$ and $\tilde{\zeta}_{aa} > 1$). The conditions for coexistence are, therefore, that

\begin{equation} \label{eq:coexist_AA}
	\lambda_{AA} < 
		\rho \left(\begin{array}{cc|c}
				\mcal{U}^S_{Aa} + \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & -C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} \\
				\frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & \mcal{U}^X_{Aa} + \frac{(1 - C_{Aa})}{2 p_n} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \mbf{F}_{aa} (1 - C_{aa}) \\ \hline
				\frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mcal{U}^S_{aa} + C_{aa} (1 - \delta) \mbf{F}_{aa} \\ 
			\end{array} \right), \\ 
\end{equation} 
\begin{equation} \label{eq:coexist_aa}
	\lambda_{aa} < 
		\rho \left(\begin{array}{c|cc}
				\mcal{U}^S_{AA} + C_{AA} (1 - \delta) \mbf{F}_{AA} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}\\ \hline
				\mbf{F}_{AA} (1 - C_{AA}) & \mcal{U}^X_{Aa} + \frac{(1 - C_{Aa})}{2 p_n} (\mbf{F}_{aa} \hat{\mbf{p}}_{aa}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) \\ 
				\mbf{0} & -C_{Aa}(1 - \delta) (\mbf{F}_{aa} \hat{\mbf{p}}_{aa}) & \mcal{U}^S_{Aa} + \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} \\
			\end{array} \right). 
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\begin{thebibliography}{}

\bibitem[{Chang and Kirkpatrick(2018)Chang and Kirkpatrick}]{ChangKirkpatrick2018}
Chang, C., and M.~Kirkpatrick. 2018.
\newblock Inversions are bigger on the X chromosome.
\newblock Mol.~Ecol. 2018:1--8.

\bibitem[{Charlesworth et.~al.(1987)Charlesworth et.~al.}]{CharlesworthBarton1987}
Charlesworth, B., J.~A. Coyne, and N. Barton. 1987.
\newblock The relative rates of evolution of sex chromosomes and autosomes.
\newblock American Naturalist 130:113--146.

\bibitem[{Connallon et al.(2018)Connallon et al.}]{ConnallonOlito2018}
Connallon, T., C.~Olito, L.~Dutoit, H.~Popali, F.~Ruzicka, and L. Yong. 2018.
\newblock Local adaptation and the evolution of inversions on sex chromosomes and autosomes.
\newblock Phil.~Trans.~Roy.~Soc.~B 373:20170423.

\bibitem[{Orr(2000)Orr}]{Orr2000}
Orr, H.~A. 2000.
\newblock The rate of adaptation in asexuals.
\newblock Genetics 155:961--968.

\bibitem[{Orr and Kim(1998)Orr and Kim}]{OrrKim1998}
Orr, H.~A. and Y.~Kim. 1998.
\newblock An adaptive hypothesis for the evolution of the Y chromosome.
\newblock Genetics 150:1693--1698.

\end{thebibliography}

\end{document}

### 2020-07-02 R/functions-Figs.R

@@ -1,0 +1,127 @@
+###############
+# DEPENDENCIES
+###############
+library(extrafont)
+library(fontcm)
+loadfonts(quiet = TRUE)
+library(plyr)
+library(lattice)
+library(latticeExtra)
+library(wesanderson)
+
+#######################
+# AUXILLIARY FUNCTIONS
+#######################
+
+toPdf <- function(expr, filename, ...) {
+  toDev(expr, pdf, filename, ...)
+}
+
+figPath  <-  function(name) {
+  file.path('output/figures', name)
+}
+
+toDev <- function(expr, dev, filename, ..., verbose=TRUE) {
+  if ( verbose )
+    cat(sprintf('Creating %s\n', filename))
+  dev(filename, family='CM Roman', ...)
+  on.exit(dev.off())
+  eval.parent(substitute(expr))
+}
+
+
+
+####################
+# PLOTTING FUNCTIONS
+####################
+
+#' Plot text or points according to relative axis position.
+#'
+#' @title Plot text or points according to relative axis position
+#' @param px Relative x-axis position (in proportion) where character is to be plotted.
+#' @param py Relative y-axis position (in proportion) where character is to be plotted.
+#' @param lab Plotted text. Works if argument \code{\link[graphics]{text}} is \code{TRUE}.
+#' @param adj See argument of same name in R base function \code{\link[graphics]{par}}.
+#' @param text Logical. Should text or points be plotted?
+#' @param log Used if the original plot uses the argument log, e.g. \code{log='x'}, \code{log='y'} or \code{log='xy'}.
+#' @param ... Additional arguments to R base function \code{\link[graphics]{text}}.
+#' @export
+proportionalLabel <- function(px, py, lab, adj=c(0, 1), text=TRUE, log=FALSE, ...) {
+    usr  <-  par('usr')
+    x.p  <-  usr[1] + px*(usr[2] - usr[1])
+    y.p  <-  usr[3] + py*(usr[4] - usr[3])
+    if(log=='x') {
+        x.p<-10^(x.p)
+    }
+    if(log=='y') {
+        y.p<-10^(y.p)
+    }
+    if(log=='xy') {
+        x.p<-10^(x.p)
+        y.p<-10^(y.p)
+    }
+    if(text){
+        text(x.p, y.p, lab, adj=adj, ...)
+    } else {
[…]

### 2020-07-02 R/functions-MatModels.R

@@ -1,0 +1,60 @@
+################################################################
+#  Matrix Model Functions
+#  Useful functions for implementing matrix models for stage-
+#  structured populations
+#
+#  Author: Colin Olito, adapted from L. DeVries
+#
+#  NOTES:  
+#		
+
+
+rm(list=ls())
+#####################
+##  Dependencies
+
+#####################
+##  Functions
+
+# create an array of 0's with dimensions dims...
+zeros  <-  function(dims) {
+	array(0,dim=dims)
+}
+
+# create an array of 1's of with dimensions dims... 
+ones  <-  function(dims) {
+	array(1,dim=dims)
+}
+
+# emat(m,n,i,j) creates an m by n matrix with 1 in the i,j entry and zeros elsewhere
+emat <- function(m,n,i,j){
+	E <- matrix(0,nrow=m,ncol=n)
+	E[i,j] <- 1
+	return(E)
+}
+
+# vecperm
+# function to calculate the vec permutation matrix of size m,n
+# 4/9/03
+#function p = vecperm(m,n)
+vecperm <- function(m,n){
+p <- as.vector(zeros(m*n))
+a <- zeros(c(m,n))
+for (i in 1:m){
+  for (j in 1:n){
+    e <- a
+    e[i,j] <- 1
+    p <- p + kronecker(e,t(e))
+    }
+}
+return(p)
+}
+
+
+rep.row<-function(x,n){
+  matrix(rep(x,each=n),nrow=n)
+}
+
+rep.col<-function(x,n){
+  matrix(rep(x,each=n), ncol=n, byrow=TRUE)
+}

### 2020-07-02 R/functions-Simulations.R

@@ -1,0 +1,354 @@
+################################################################
+#  Functions to simulate demographic model across selection
+#  parameter space
+#
+#
+#  Author: Colin Olito, adapted from L. DeVries
+#
+#  NOTES:  
+#		
+
+
+rm(list=ls())
+#####################
+##  Dependencies
+source('R/functions-MatModels.R')
+source('R/functions-Figs.R')
+
+
+###########################
+#' 1-locus Pop Gen Invasion
+#' Conditions
+#' 
+#' Parameters:
+#' 
+popGen_A_invade  <-  function(hf, hm, sm, C) {
+	(sm*(C - 1)*(2*hm*(C - 1) - C)) / (sm*(C - 1)*(2*hm*(C - 1) - C) + (C + 1)*(2 - C + 2*hf*(C - 1)))
+}
+popGen_a_invade  <-  function(hf, hm, sm, C) {
+	(sm*(1 - C)*(2 - C+2*hm*(C - 1))) / ((C + 1)*(2*hf*(C - 1) - C)*(sm - 1))
+}
+
+###########################
+#' Forward Simulation 
+#'
+#' Parameters:
+#' dims: 	vector c(om, g), with om = # stages, g = # genotypes
+#' theta: 	vector of length 4, c(sigma_J, sigma_A, gamma, f_ii)
+fwdDemModelSim  <-  function(	om = 2, g = 3, theta = c(0.6,0.6,0.05,5.9), theta_prime = c(0.6,0.6,0.05,5.9), 
+								hf = 1/2, hm = 1/2, sf = 0.1, sm = 0.105, C = 0, delta = 0, tlimit = 10^4, Ainvade = FALSE, intInit = FALSE, ...) {
+
+	# Create identity and ones matrices
+	Ig   <-  diag(g)
+	Iom  <-  diag(om)
+	eg   <-  ones(c(g,1))
+	eom  <-  ones(c(om,1))
+	K    <-  vecperm(om,g)
+
+	####################################################
+	# PARAMETERS AND INITIAL CONDITIONS 
+	####################################################
+
+	# BASELINE PARAMETERS for survival and 
+	# fertility through Female and Male function
+	# theta=[sigma_J, sigma_A, gamma, f_ii]
+	theta       <-  rep.col(theta,3)
+	theta_prime <- rep.col(theta_prime,3)
+
+	# Fitness Expressions
[…]

### 2020-07-02 matlab/deVries_Caswell_2019_AmNat_Two_stage_model.m

@@ -1,0 +1,235 @@
+%Program to implement Mendelian Matrix Population Models
+%for two-stage model with parameters leading to evolutioary rescue
+% Authors: Hal Caswell and Charlotte de Vries
+% June 2018
+clearvars; 
+%Dimensions, 2 stages, 3 genotypes
+om=2;
+g=3;
+
+%Identity and ones matrices
+Ig=eye(g);
+Iom=eye(om);
+eg=ones(g,1);
+eom=ones(om,1);
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Parameters for EVOLUTIONARY SUICIDE (Fig 1A and 1C). 
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Nzero=[0.1000         0   99.9000
+    0.1000         0   99.9000];
+%Baseline parameters
+%theta=[s1, s2, g, f]
+
+theta=[0.65 0.7 0.05 3.2]'*ones(1,3);
+%Selection differentials
+
+% %s1, juvenile survival
+ds1=0.032;
+ seldiff=[-ds1+0.01 ds1+0.008 -ds1+0.005];
+ theta(1,:)=theta(1,:)+seldiff;
+
+%s2 differential, adult survival
+% s2
+    ds2=0.0001;
+
+seldiff=[0 0 ds2];
+theta(2,:)=theta(2,:)+seldiff;
+
+%this is the f differential for the pleiotropy example
+% f
+    df=0.81;
+    seldiff=[df/2 -df df];
+theta(4,:)=theta(4,:)+seldiff;
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% END of Parameters for EVOLUTIONARY SUICIDE (Fig 1A and 1C). 
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Parameters for EVOLUTIONARY RESCUE (Fig 1B and 1D). 
+% Uncommment the following section of code if you wish to use the
[…]

### 2020-07-02 run-Simulations.R

@@ -1,0 +1,205 @@
+################################################################
+#  RUN SIMULATIONS FOR 
+#
+#  R code for forward simulations. Generates output data
+#  as .csv files saved to ./output/simData/
+#
+#
+#  Author: Colin Olito
+#
+#  NOTES:  
+#		
+
+
+
+
+#####################
+##  Dependencies
+rm(list=ls())
+source('R/functions-Simulations.R')
+
+
+
+
+######################
+# Preliminary figures
+######################
+
+## Additive SA Fitness
+## hf = hm = 1/2
+
+	# Obligate Outcrossing
+	# C = 0
+	pars  <-  list(
+					"hf"     =  1/2,
+					"hm"     =  1/2,
+					"C"      =  0,
+					"delta"  =  0
+					)
+
+	selLoop(sMax = 0.15, nSamples=2000,
+			om = 2, g = 3, theta = c(0.6,0.6,0.05,6), theta_prime = c(0.6,0.6,0.05,6), 
+			hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, tlimit = 10^5, returnRes = FALSE)
+
+	# Intermediate Selfing
+	# C = 1/2
+	pars  <-  list(
+					"hf"     =  1/2,
+					"hm"     =  1/2,
+					"C"      =  1/2,
+					"delta"  =  0
+					)
+
+	selLoop(sMax = 0.15, nSamples=2000,
+			om = 2, g = 3, theta = c(0.6,0.6,0.05,6), theta_prime = c(0.6,0.6,0.05,6), 
+			hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, tlimit = 10^5, returnRes = FALSE)
+
+	# Intermediate Selfing w/ I.D.
+	# C = 1/2, delta = 0.25
+	pars  <-  list(
+					"hf"     =  1/2,
+					"hm"     =  1/2,
+					"C"      =  1/2,
+					"delta"  =  1/4
+					)
+
+	selLoop(sMax = 0.15, nSamples=2000,
+			om = 2, g = 3, theta = c(0.6,0.6,0.05,6), theta_prime = c(0.6,0.6,0.05,6), 
[…]

### 2020-07-02 test_SA_Hermaphrodite_Model.R

@@ -1,0 +1,299 @@
+# First attempt to implement a Mendelian Matrix Population Model
+# for a partially selfing population of hermaphrodites with selection
+# through both sex functions.
+#
+#	For simplicity I'm just hacking Lotte's code here to see if
+#	all the component matrices fit, and if the model does what 
+#   I expect it to do.
+#
+# The model has:
+# 	Two stages: juvenile and adult
+# 	Three genotypes: AA, Aa, aa
+#   A constant selfing rate: C_AA = C_Aa = C_aa = C, where 0 <= C <= 1
+#   Constant inbreeding depression: 0 < delta < 1
+
+
+rm(list=ls())
+#####################
+##  Dependencies
+source('R/functions-MatModels.R')
+
+Ainvade  <-  FALSE
+intInit  <-  FALSE
+
+# Simulation time"
+tlimit <- 100000
+
+# dimensions
+om  <-  2
+g   <-  3
+
+# Create identity and ones matrices
+Ig   <-  diag(g)
+Iom  <-  diag(om)
+eg   <-  ones(c(g,1))
+eom  <-  ones(c(om,1))
+K    <-  vecperm(om,g)
+
+# Population Selfing Rate and inbreeding depression
+C      <-  0
+delta  <-  0
+
+#####################################################"
+
+# PARAMETERS AND INITIAL CONDITIONS 
+
+####################################################"
+
+# BASELINE PARAMETERS for survival and 
+# fertility through Female and Male function
+# theta=[sigma_J, sigma_A, gamma, f_ii]
+theta_temp  <-  c(0.6,0.6,0.05,5.9)
+theta       <-  rep.col(theta_temp,3)
+
+theta_primetemp <- c(0.6,0.6,0.05,5.9)
+theta_prime <- rep.col(theta_primetemp,3)
+
+## SELECTION PARAMETERS
+hf  <-  1/2
+hm  <-  1/2
+sf  <-  0.01
+sm  <-  0.012
+fii        <-  c(1, 1 - hf*sf, 1 - sf)
+fii_prime  <-  c(1 - sm, 1 - hm*sm, 1)
+
+##SELECTION DIFFERENTIAL FEMALE-FUNCTION
+##female function
[…]

### 2020-07-02 test_SA_Hermaphrodite_Model.R
testing simulation with new Atilde structure

@@ -18,11 +18,11 @@ rm(list=ls())
 ##  Dependencies
 source('R/functions-MatModels.R')
 
-Ainvade  <-  FALSE
+Ainvade  <-  TRUE
 intInit  <-  FALSE
 
 # Simulation time"
-tlimit <- 100000
+tlimit <- 50000
 
 # dimensions
 om  <-  2
@@ -36,7 +36,7 @@ eom  <-  ones(c(om,1))
 K    <-  vecperm(om,g)
 
 # Population Selfing Rate and inbreeding depression
-C      <-  0
+C      <-  1/4
 delta  <-  0
 
 #####################################################"
@@ -48,17 +48,17 @@ delta  <-  0
 # BASELINE PARAMETERS for survival and 
 # fertility through Female and Male function
 # theta=[sigma_J, sigma_A, gamma, f_ii]
-theta_temp  <-  c(0.6,0.6,0.05,5.9)
+theta_temp  <-  c(0.6,0.6,0.05,3)
 theta       <-  rep.col(theta_temp,3)
 
-theta_primetemp <- c(0.6,0.6,0.05,5.9)
+theta_primetemp <- c(0.6,0.6,0.05,3)
 theta_prime <- rep.col(theta_primetemp,3)
 
 ## SELECTION PARAMETERS
 hf  <-  1/2
 hm  <-  1/2
-sf  <-  0.01
-sm  <-  0.012
+sf  <-  0.12
+sm  <-  0.1
 fii        <-  c(1, 1 - hf*sf, 1 - sf)
 fii_prime  <-  c(1 - sm, 1 - hm*sm, 1)
 
@@ -142,33 +142,6 @@ for (i in 1:3){
 }
 
 
-pHatAA  <-  t(eom)
-pHatPrimeAA  <-  pHatAA
-pn  <-  t(eom)%*%FXi_prime[,,1]%*%eom
-zeroOmg  <-  zeros(c(om,om))
-temp  <-  rbind(	c( (USi[,,2] + (1/2)*C*(1 - delta)*FSi[,,2]), (C*(1 - delta) * kronecker((FXi[,,1] %*% t(pHatAA)), (t(eom)%*%FXi[,,2])))                               , zeroOmg),
[…]

### 2020-07-03 notes/notes-SA-partSelf-wDemography.tex
testing simulation with new Atilde structure
updates to notes with new Atilde

@@
 
-\noindent Individuals are jointly classified by stage $(1, \ldots , \omega)$ and genotype $(1, \ldots , g)$.
+\noindent Individuals are jointly classified by whether they were produced by self-fertilization or outcrossing (denoted by $S$ and $X$), stage $(1, \ldots , \omega)$ and genotype $(1, \ldots , g)$.
 Each genotype is characterized by a matrix of transition probabilities (including survival) and a matrices of reproductive output through male and female function.
@@
 	\mathcal{N} \sim \left[
-						\begin{array}{ccc}
-							n_{1,1}      &  \cdots & n_{1,g} \\
-							\vdots       &         & \vdots  \\
-							n_{\omega,1} &  \cdots & n_{\omega,g} \\
+						\begin{array}{ccc|ccc}
+							n^S_{1,1}      &  \cdots & n^S_{1,g}       &  n^X_{1,1}      &  \cdots & n^X_{1,g} \\
+							\vdots       &         & \vdots        &  \vdots       &         & \vdots  \\
+							n^S_{\omega,1} &  \cdots & n^S_{\omega,g}  &  n^X_{\omega,1} &  \cdots & n^X_{\omega,g} \\
 						\end{array} \right],
@@
 								\begin{array}{c}
-									\mbf{n}_{AA}(t) \\
-									\mbf{n}_{Aa}(t) \\
-									\mbf{n}_{aa}(t) \\
+									\mbf{n}^{S}_{AA}(t) \\
+									\mbf{n}^{S}_{Aa}(t) \\
+									\mbf{n}^{S}_{aa}(t) \\ \hline
+									\mbf{n}^{X}_{AA}(t) \\
+									\mbf{n}^{X}_{Aa}(t) \\
+									\mbf{n}^{X}_{aa}(t) \\ 
 						\end{array} \right].
@@
 			\begin{array}{c|c}
-				\mcal{F}^S(\tilde{\mbf{p}}) (1 - \delta) & \mbf{0} \\ \hline
-				\mbf{0} & \mcal{F}^X(\tilde{\mbf{p}}) \\
+				\mcal{F}^S(\tilde{\mbf{p}}) (1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) (1 - \delta) \\ \hline
+				\mcal{F}^X(\tilde{\mbf{p}}) & \mcal{F}^X(\tilde{\mbf{p}}) \\
 			\end{array} \right)}_{\tilde{\mbf{F}}}
@@
 
-\noindent and corresponding matices $\mbb{U}^X$, $\mbb{F}^S$, $\mbb{F}^X$.
+\noindent and corresponding matrices $\mbb{U}^X$, $\mbb{F}^S$, $\mbb{F}^X$.
 Because individuals cannot change their genotype once they are born, the survival matrices are block diagonal; $\mcal{U}^S = \mbb{U}^S$ and $\mcal{U}^X = \mbb{U}^X$.
@@
 \begin{align*}
-	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
-		& \underbrace{\left(
+	&\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
+		 \underbrace{\left(
 			\begin{array}{ccc|ccc}
@@
 				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{aa}
-			\end{array} \right)}_{\tilde{\mbf{U}}} \\ 
-			&+ \underbrace{\left(
+			\end{array} \right)}_{\tilde{\mbf{U}}} + \\ 
+	&\underbrace{\left(
 			\begin{array}{ccc|ccc}
-				C_{AA} (1 - \delta) \mbf{F}_{AA} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & 0 & \mbf{0} & \mbf{0} & \mbf{0} \\ 
-				0 & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & 0 & \mbf{0} & \mbf{0} & \mbf{0} \\
-				0 & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & C_{aa} (1 - \delta) \mbf{F}_{aa} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
-				q^{\prime}_{A} \mbf{F}_{AA} (1 - C_{AA})& \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} (1 - C_{Aa}) & 0 & \mbf{0} & \mbf{0} & \mbf{0} \\
-				q^{\prime}_{a} \mbf{F}_{AA} (1 - C_{AA}) & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & q^{\prime}_{A} \mbf{F}_{aa} (1 - C_{aa}) & \mbf{0} & \mbf{0} & \mbf{0}\\
-				0 & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} (1 - C_{Aa})& q^{\prime}_{a} \mbf{F}_{aa} (1 - C_{aa}) & \mbf{0} & \mbf{0} & \mbf{0}
+				C_{AA} (1 - \delta) \mbf{F}_{AA} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & C_{AA} (1 - \delta) \mbf{F}_{AA} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} \\ 
+				\mbf{0} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0}  \\
+				\mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & C_{aa} (1 - \delta) \mbf{F}_{aa} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & C_{aa} (1 - \delta) \mbf{F}_{aa} \\ \hline
+				q^{\prime}_{A} \mbf{F}_{AA} (1 - C_{AA})& \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} (1 - C_{Aa}) & \mbf{0} & q^{\prime}_{A} \mbf{F}_{AA} (1 - C_{AA})& \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} (1 - C_{Aa}) & \mbf{0} \\
+				q^{\prime}_{a} \mbf{F}_{AA} (1 - C_{AA}) & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & q^{\prime}_{A} \mbf{F}_{aa} (1 - C_{aa}) & q^{\prime}_{a} \mbf{F}_{AA} (1 - C_{AA}) & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & q^{\prime}_{A} \mbf{F}_{aa} (1 - C_{aa}) \\
+				\mbf{0} & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} (1 - C_{Aa})& q^{\prime}_{a} \mbf{F}_{aa} (1 - C_{aa}) & \mbf{0} & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} (1 - C_{Aa})& q^{\prime}_{a} \mbf{F}_{aa} (1 - C_{aa})
 			\end{array} \right)}_{\tilde{\mbf{F}}} \numberthis
@@
 We also assume that the effects act only during the adult stage through reproductive success.
-The allele does not affect survival and transition rates, such that
-
-\begin{equation}
-	\mbf{U} = \left(
+The allele does not affect survival and transition rates.
+However, the survival matrices can be used to model 'late-acting' inbreeding depression by allowing different stage-specific survival and transition rates for individuals produced by self-fertilization vs.~outcrossing.
+By contrast, the parameter $\sigma$ only affects inbreeding depression through viability of selfed ovlues.
+With this in mind, we include different survival matrices for individuals produced by selfing and outcrossing as follows:
+
+\begin{align*}
+	\mbf{U}^S = &\left(
 					\begin{array}{cc}
-						\sigma_J(1 - \gamma) & 0 \\
-						\sigma_J \gamma      & \sigma_A
+						\sigma^S_J(1 - \gamma^S) & 0 \\
+						\sigma^S_J \gamma      & \sigma^S_A
 					\end{array}
-				\right),
-\end{equation}
+				\right) \\
+	\mbf{U}^X = &\left(
+					\begin{array}{cc}
+						\sigma^X_J(1 - \gamma^X) & 0 \\
+						\sigma^X_J \gamma      & \sigma^X_A
+					\end{array}
+				\right)
+\end{align*}
 
@@
 					\begin{array}{cc}
-						0 & f_{ii} \\
+						0 & C_{ii}(1 - \delta) f_{ii} \\
 						0 & 0
@@
 					\begin{array}{cc}
-						0 & f^{\prime}_{ii} \\
+						0 & (1 - C_{ii}) f^{\prime}_{ii} \\
 						0 & 0
@@
 
-\begin{equation}
+\begin{align*}
 	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
-			\underbrace{\left(
+			&\underbrace{\left(
 			\begin{array}{c|c}
@@
 			\begin{array}{c|c}
-				\mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) & \mbf{0} \\ \hline
-				\mbf{0} & \mcal{F}^X(\tilde{\mbf{p}}) \\
-			\end{array} \right)}_{\tilde{\mbf{F}}}
-\end{equation}
+				\mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
+				\mcal{F}^X(\tilde{\mbf{p}}) (1 - C) & \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)\\
+			\end{array} \right)}_{\tilde{\mbf{F}}} \\
+			= & \left(\begin{array}{c|c}
+				\mcal{U}^S + \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
+				\mcal{F}^X(\tilde{\mbf{p}})(1 - C) & \mcal{U}^X + \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)
+			\end{array} \right) \numberthis
+\end{align*}
 
@@
 					& $AA$			& $Aa$ 					& $aa$ 		\\ \hline
-Female function:	& $f$		& $f - h_f s_f$	& $f - s_f$ \\	
-Male function:		& $f - s_m$	& $f - h_m s_m$	& $f$ 		\\	
+Female function:	& $f_{AA}$		& $f_{Aa} - h_f s_f$	& $f_{aa} - s_f$ \\	
+Male function:		& $f_{AA} - s_m$& $f_{Aa} - h_m s_m$	& $f_{aa}$ 		\\	
 \hline

### 2020-07-03 R/functions-MatModels.R

@@ -36,7 +36,7 @@ emat <- function(m,n,i,j){
 # vecperm
 # function to calculate the vec permutation matrix of size m,n
 # 4/9/03
-#function p = vecperm(m,n)
+# function p = vecperm(m,n)
 vecperm <- function(m,n){
 p <- as.vector(zeros(m*n))
 a <- zeros(c(m,n))

### 2020-07-09 notes/notes-SA-partSelf-wDemography.tex
added warning to notes about coexistence conditions
Update notes-SA-partSelf-wDemography.tex
Calculated the coexistence conditions for the general case (when how-you-were-produced matters), checked they reduce to deVries&Caswell (2019, AmNat) in the absence of selfing. Then formulated model and calculated coexistence conditions when how-you-were-produced does not affect your survival.

@@
 \def\mcal#1{\mathcal{#1}}
+\newcommand{\bo}[1]{{\bf #1}}
+\newcommand{\tr}{{\mbox{\tiny \sf T}}}
+
+\newcommand{\bm}[1]{\mbox{\boldmath $#1$}}
 
@@
 \end{table}
+
+\hl{We have to be a bit careful with our notation.
+Above we have defined $\mathbf{n}^X$ as the population vector of individuals that were produced by selfing, so it would be natural to use $\mathbf{p}^X$ as the normalized version of that vector, which is in conflict with the above definition of $\mbf{p^X_i} $.}
 
@@
 \end{equation}
+\hl{This vector has to be created from a sum of $\mathbf{n}^X$ and $\mathbf{n}^S$ since they both outcross.}
 
@@
 The matrix $\mbf{W}^{\prime}$ converts these relative genotype contributions to allele numbers.
-Normalizing this vector gives the allele frequencies in the gamete pool,
+Normalizing this vector gives the allele frequencies in the gamete pool (\textbf{is it worth emphasizing here that this is frequencies in pollen?}),
 
@@
 			\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p} \|} = 
-				\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{n}}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{n} \|}
+				\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X+ \mbf{n}^S)}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X+ \mbf{n}^S)\|}
 \end{equation}
@@
 				\mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & C_{aa} (1 - \delta) \mbf{F}_{aa} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & C_{aa} (1 - \delta) \mbf{F}_{aa} \\ \hline
-				q^{\prime}_{A} \mbf{F}_{AA} (1 - C_{AA})& \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} (1 - C_{Aa}) & \mbf{0} & q^{\prime}_{A} \mbf{F}_{AA} (1 - C_{AA})& \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} (1 - C_{Aa}) & \mbf{0} \\
-				q^{\prime}_{a} \mbf{F}_{AA} (1 - C_{AA}) & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & q^{\prime}_{A} \mbf{F}_{aa} (1 - C_{aa}) & q^{\prime}_{a} \mbf{F}_{AA} (1 - C_{AA}) & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & q^{\prime}_{A} \mbf{F}_{aa} (1 - C_{aa}) \\
-				\mbf{0} & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} (1 - C_{Aa})& q^{\prime}_{a} \mbf{F}_{aa} (1 - C_{aa}) & \mbf{0} & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} (1 - C_{Aa})& q^{\prime}_{a} \mbf{F}_{aa} (1 - C_{aa})
+				q^{\prime}_{A} (1 - C_{AA})\mbf{F}_{AA} & \frac{1}{2} q^{\prime}_{A} (1 - C_{Aa})\mbf{F}_{Aa}  & \mbf{0} & q^{\prime}_{A} (1 - C_{AA}) \mbf{F}_{AA}& \frac{1}{2} q^{\prime}_{A} (1 - C_{Aa}) \mbf{F}_{Aa} & \mbf{0} \\
+				q^{\prime}_{a} (1 - C_{AA}) \mbf{F}_{AA} & \frac{1}{2} (1 - C_{Aa})\mbf{F}_{Aa}  & q^{\prime}_{A} (1 - C_{aa})\mbf{F}_{aa}  & q^{\prime}_{a} (1 - C_{AA}) \mbf{F}_{AA} & \frac{1}{2}(1 - C_{Aa}) \mbf{F}_{Aa}  & q^{\prime}_{A} (1 - C_{aa})\mbf{F}_{aa}  \\
+				\mbf{0} & \frac{1}{2} q^{\prime}_{a} (1 - C_{Aa})\mbf{F}_{Aa} & q^{\prime}_{a} (1 - C_{aa}) \mbf{F}_{aa} & \mbf{0} & \frac{1}{2} q^{\prime}_{a} (1 - C_{Aa})\mbf{F}_{Aa} & q^{\prime}_{a} (1 - C_{aa})\mbf{F}_{aa}
 			\end{array} \right)}_{\tilde{\mbf{F}}} \numberthis
@@
 \setcounter{equation}{0}
+%\hl{EVERYTHING BELOW THIS POINT IS JUNK, AS IT HAS NOT BEEN UPDATED WITH THE NEW $\tilde{\mbf{A}}$ structure!!!}
 
@@
 			\begin{array}{c}
-							\mbf{n}_{AA} \\
-							\mbf{n}_{AA} \\ \hline
-							\mbf{n}_{Aa} \\
-							\mbf{n}_{Aa} \\ \hline
-							\mbf{n}_{aa} \\
-							\mbf{n}_{aa} \\
+							\mbf{n}^S_{AA} \\
+							\mbf{n}^X_{AA} \\ \hline
+							\mbf{n}^S_{Aa} \\
+							\mbf{n}^X_{Aa} \\ \hline
+							\mbf{n}^S_{aa} \\
+							\mbf{n}^X_{aa} \\
 			\end{array} \right).
@@
 
-\noindent Note that to 'doubling' up of the genotype-specific vectors in each block is done out of convenience, because the population projection matrix is organized around two sex-functions.
+\noindent Note that we are still keeping track of whether individuals were produced through selfing, $\mbf{n}^S_{ij}$, or through outcrossing, $\mbf{n}^X_{ij}$ [\textit{\textbf{LdV: I'm not a fan of combining superscripts and subscripts because it becomes such a mess of indices, what if we give them different letters instead, e.g.
+$\mbf{n}^S_{AA}=\mbf{n}_{AA}$ and $\mbf{n}^X_{AA}=\mbf{m}_{AA}$?}}].
 
@@
 
+{
+\small
 \begin{align*} \label{eq:AtildeCoexistence}
-	\tilde{\mbf{A}}[\tilde{\mbf{n}}] &= \tilde{\mbf{U}} + \tilde{\mbf{F}} \\
+	\tilde{\mbf{A}}[\tilde{\mbf{n}}] &= \tilde{\mbf{U}} + \tilde{\mbf{F}}\\
 		&= \left(
@@
 			\begin{array}{cc|cc|cc}
-				C_{AA} (1 - \delta) \mbf{F}_{AA} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
-				q^{\prime}_{A} \mbf{F}_{AA} (1 - C_{AA}) & \mbf{0} & \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} (1 - C_{Aa}) & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
-				\mbf{0} & \mbf{0} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\
-				q^{\prime}_{a} \mbf{F}_{AA} (1 - C_{AA}) & \mbf{0} & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & \mbf{0} & q^{\prime}_{A} \mbf{F}_{aa} (1 - C_{aa}) & \mbf{0}\\ \hline
-				\mbf{0} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & C_{aa} (1 - \delta) \mbf{F}_{aa} & \mbf{0} \\ 
-				\mbf{0} & \mbf{0} & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} (1 - C_{Aa})  & \mbf{0} & q^{\prime}_{a} \mbf{F}_{aa} (1 - C_{aa}) & \mbf{0}
+				C_{AA} (1 - \delta) \mbf{F}_{AA} & C_{AA} (1 - \delta) \mbf{F}_{AA}  & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} \\ 
+				q^{\prime}_{A}  (1 - C_{AA}) \mbf{F}_{AA}& q^{\prime}_{A} (1 - C_{AA}) \mbf{F}_{AA} & \frac{1}{2} q^{\prime}_{A} (1 - C_{Aa})\mbf{F}_{Aa}  & \frac{1}{2} q^{\prime}_{A} (1 - C_{Aa})\mbf{F}_{Aa}  & \mbf{0} & \mbf{0} \\ \hline
+				\mbf{0} & \mbf{0} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}& \mbf{0} & \mbf{0} \\
+				q^{\prime}_{a} (1 - C_{AA}) \mbf{F}_{AA} & q^{\prime}_{a}(1 - C_{AA})  \mbf{F}_{AA} & \frac{1}{2}  (1 - C_{Aa}) \mbf{F}_{Aa}& \frac{1}{2} (1 - C_{Aa})\mbf{F}_{Aa} & q^{\prime}_{A} (1 - C_{aa})\mbf{F}_{aa}  & q^{\prime}_{A}  (1 - C_{aa})\mbf{F}_{aa}\\ \hline
+				\mbf{0} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & C_{aa} (1 - \delta) \mbf{F}_{aa} &  C_{aa} (1 - \delta) \mbf{F}_{aa}\\ 
+				\mbf{0} & \mbf{0} & \frac{1}{2} q^{\prime}_{a}  (1 - C_{Aa}) \mbf{F}_{Aa} & \frac{1}{2} q^{\prime}_{a}  (1 - C_{Aa})\mbf{F}_{Aa}& q^{\prime}_{a}(1 - C_{aa})  \mbf{F}_{aa} & q^{\prime}_{a} (1 - C_{aa})\mbf{F}_{aa}
 			\end{array} \right) \numberthis
 \end{align*}
+}
 
@@
 
-The two-sex Mendelian matrix model defined by above reduces to a linear matrix model on the boundary where the population is initially fixed for the $A$ allele (since $q^{\prime}_{A} = 1$ and $q^{\prime}_{a} = 0$).
+The Mendelian matrix model defined by above reduces to a linear matrix model on the boundary where the population is initially fixed for the $A$ allele (since $q^{\prime}_{A} = 1$ and $q^{\prime}_{a} = 0$).
 Provided the initial population contains a nonzero number of females, the population will grow or shrink exponentially after converging to a stable population structure (see Caswell (2001), section 4.5.2.1).
@@
 			\begin{array}{cc|cc|cc}
-				\mcal{U}^S_{AA} + C_{AA} (1 - \delta) \mbf{F}_{AA} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
-				\mbf{F}_{AA} (1 - C_{AA}) & \mcal{U}^X_{AA} & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
-				\mbf{0} & \mbf{0} & \mcal{U}^S_{Aa} + \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\
-				\mbf{0} & \mbf{0} & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & \mcal{U}^X_{Aa} & \mbf{F}_{aa} (1 - C_{aa}) & \mbf{0}\\ \hline
-				\mbf{0} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mcal{U}^S_{aa} + C_{aa} (1 - \delta) \mbf{F}_{aa} & \mbf{0} \\ 
-				\mbf{0} & \mbf{0} & \mbf{0}  & \mbf{0} & \mbf{0} & \mcal{U}^X_{aa}
-			\end{array} \right) \numberthis
+				\mathbf{U}^S_{AA} & \mbf{0}& \mbf{0} & \mbf{0} & \mbf{0} \\ 
+			\mbf{0} & \mathbf{U}^X_{AA} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
+				\mbf{0} & \mbf{0} & \mathbf{U}^S_{Aa}& \mbf{0} & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \mbf{0}& \mcal{U}^X_{Aa} & \mbf{0} & \mbf{0}\\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mathbf{U}^S_{aa}  & \mbf{0} \\ 
+				\mbf{0} & \mbf{0} & \mbf{0}  & \mbf{0} & \mbf{0} & \mathbf{U}^X_{aa}
+			\end{array} \right) \\
+			 +\frac{1}{\lambda_{AA}} &
+						   \left(
+			\begin{array}{cc|cc|cc}
+		 C_{AA} (1 - \delta) \mbf{F}_{AA} & C_{AA} (1 - \delta) \mbf{F}_{AA}& \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}& \mbf{0} & \mbf{0} \\ 
+				(1 - C_{AA}) \mbf{F}_{AA} & (1 - C_{AA}) \mbf{F}_{AA}  & \frac{1}{2}(1 - C_{Aa}) \mbf{F}_{Aa}  & \frac{1}{2} (1 - C_{Aa})\mbf{F}_{Aa} & \mbf{0} & \mbf{0} \\ \hline
+				\mbf{0} & \mbf{0} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} & \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} & (1 - C_{aa}) \mbf{F}_{aa} & (1 - C_{aa}) \mbf{F}_{aa}\\ \hline
+				\mbf{0} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &C_{aa} (1 - \delta) \mbf{F}_{aa} &C_{aa} (1 - \delta) \mbf{F}_{aa} \\ 
+				\mbf{0} & \mbf{0} & \mbf{0}  & \mbf{0} & \mbf{0} &\mbf{0}
+			\end{array} \right) 
+			\numberthis
 \end{align*}
@@
 \begin{equation}
-	\encircle{B} = - \frac{1}{\lambda_{AA}} \left( \hat{\mbf{p}}^{\intercal} \otimes \mbf{1}_{2 \omega g} \right) \left( \text{vec} \tilde{\mbf{A}} \right) \left( \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}} \right).
+	\encircle{B} = - \frac{1}{\lambda_{AA}^2} \left( \hat{\mbf{p}}^{\intercal} \otimes \mbf{1}_{2 \omega g} \right) \left( \text{vec} \tilde{\mbf{A}} \right) \left( \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}} \right).
 \end{equation}
@@
 
-\begin{landscape}
-
-\begin{equation}
-	\encircle{B} = - \frac{1}{\lambda_{AA}} \hat{\mbf{p}} \left( \text{vec} \tilde{\mbf{A}} \right) \left( \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}}[\hat{\mbf{p}}] \right).
+\begin{equation}
+	\encircle{B} = - \frac{1}{\lambda_{AA}} \hat{\mbf{p}} \left( \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}}[\hat{\mbf{p}}] \right).
 \end{equation}
@@
 			\begin{array}{c}
-							\hat{\mbf{p}}_{AA} \\
-							\hat{\mbf{p}}_{AA} \\ \hline
+							\hat{\mbf{p}}^S_{AA} \\
+							\hat{\mbf{p}}^X_{AA} \\ \hline
 							0                  \\
@@
 
+\begin{landscape}
 \noindent and rewrite in terms of the block matrices to obtain
 
+{\footnotesize
 \begin{align*}
@@
 			\begin{array}{cc|cc|cc}
-				\hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{aa} \\
-				\hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^S_{aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mcal{U}^X_{aa} \\ \hline
+				\hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{aa} \\
+				\hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{aa} \\ \hline
 				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
@@
 			\begin{array}{cc|cc|cc}
-				C_{AA}(1 - \delta) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & \mbf{0} & C_{AA}(1 - \delta) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & \mbf{0} & C_{AA}(1 - \delta) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} & \mbf{0} \\
-				(1 - C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & \mbf{0} & (1 - C_{Aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & \mbf{0} & (1 - C_{Aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} & \mbf{0} \\ \hline
+				(1 - \delta C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & (1 - \delta C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & (1 - \delta C_{Aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & (1 - \delta C_{Aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa}  & (1 - \delta C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} & (1 - \delta C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} \\
+				 	(1 - \delta C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & 	(1 - \delta C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & (1 - \delta C_{Aa})  \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & (1 - \delta C_{Aa})  \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & (1 - \delta C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} & (1 - \delta C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa}\\ \hline
 				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
 				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0}
+			\end{array} \right) \\
+			 \numberthis			
+\end{align*}
+}
+
+\textit{\textbf{Oof, I realized in doing this that I made a mistake here in the appendix of the two sex paper, the alpha's should have disappeared because when you sum over all the rows they cancel.
+}}
+
+\end{landscape}
+
+Finally, we turn to term \encircle{C} of the Jacobian.
+As far as I can tell, we can follow the steps leading from Eq(A-35) through to Eq(A-46) of de Vries \& Caswell (2019) Appendix A exactly because the structure of our $\tilde{\mbf{A}}$ equations are identical.
+We resume the derivation at our version of their Eq(A-47)...Substitute $\hat{\mbf{p}}^{\intercal} = \left( \hat{\mbf{p}}^{\intercal}_{S,AA},\hat{\mbf{p}}^{\intercal}_{X,AA},\mbf{0},\mbf{0},\mbf{0},\mbf{0} \right)$ into the right side of Eq(A-46) so that only terms with $j=1$ and $j=2$ are nonzero, yielding
+
+\begin{align*}
+	\frac{1}{\lambda_{AA}}(\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{2 \omega g}) \left.
+\frac{ \partial \text{vec}\mbf{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} &= \frac{1}{\lambda_{AA}} \sum^{6}_{i=1} \big( \hat{\mbf{p}}^{\intercal}_{S,AA} \otimes (\mbf{e}_{i} \otimes \mbf{I}_{\omega}) \big) \left.
+\frac{ \partial \text{vec}\mbf{A}_{i,1}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} \\
+		&+ \frac{1}{\lambda_{AA}} \sum^{6}_{i=1} \big( \hat{\mbf{p}}^{\intercal}_{X,AA} \otimes (\mbf{e}_{i} \otimes \mbf{I}_{\omega}) \big) \left.
+\frac{ \partial \text{vec}\mbf{A}_{i,2}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} \numberthis
+\end{align*}
+
+\noindent Selfing does not depend on the frequency vector, and therefore none of the $\mbf{A}_{1,i}$ are a function of the frequency vector $\mbf{p}$,
+
+\begin{equation}
+	\left. \frac{ \partial \text{vec}\mbf{A}_{1,i}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} = 0,\,\text{for all }i. 
+\end{equation}
+
+\noindent (see Eq{\ref{eq:AtildeCoexistence}}). Next, write down each term in the sum over $i$ and take the derivative  of the $\text{vec}\mbf{A}_{i,1}$'s to obtain
+
+\begin{align*}
+	\frac{1}{\lambda_{AA}}(\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{2 \omega g}) \left.
+\frac{ \partial \text{vec}\mbf{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} &= \frac{1}{\lambda_{AA}} \big( \hat{\mbf{p}}^{\intercal}_{S,AA} \otimes (\mbf{e}_{2} \otimes \mbf{I}_{\omega}) \big) \left.
+\frac{ \partial \text{vec}\mbf{A}_{21}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} \\
+		&+ \frac{1}{\lambda_{AA}} \big( \hat{\mbf{p}}^{\intercal}_{X,AA} \otimes (\mbf{e}_{2} \otimes \mbf{I}_{\omega}) \big) \left.
+\frac{ \partial \text{vec}\mbf{A}_{22}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} \\
+		&+\frac{1}{\lambda_{AA}} \big( \hat{\mbf{p}}^{\intercal}_{S,AA} \otimes (\mbf{e}_{4} \otimes \mbf{I}_{\omega}) \big) \left.
+\frac{ \partial \text{vec}\mbf{A}_{41}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} \\
+		&+ \frac{1}{\lambda_{AA}} \big( \hat{\mbf{p}}^{\intercal}_{X,AA} \otimes (\mbf{e}_{4} \otimes \mbf{I}_{\omega}) \big) \left.
+\frac{ \partial \text{vec}\mbf{A}_{42}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} \numberthis
+\end{align*}
+
+\begin{align*}
+	\frac{1}{\lambda_{AA}}(\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{2 \omega g}) \left. \frac{ \partial \text{vec}\mbf{A}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} = 
+		&\frac{(1 - C_{AA})}{\lambda_{AA}} \big[ \hat{\mbf{p}}^{\intercal}_{X,AA}+ \hat{\mbf{p}}^{\intercal}_{S,AA} \big]\otimes (\mbf{e}_{2} \otimes \mbf{I}_{\omega}) \text{vec}\mbf{F}_{AA} \left.\frac{\partial q^{\prime}_{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} \\
+		&-\frac{(1 - C_{AA})}{\lambda_{AA}} \big[ \hat{\mbf{p}}^{\intercal}_{X,AA}+ \hat{\mbf{p}}^{\intercal}_{S,AA} \big]\otimes (\mbf{e}_{4} \otimes \mbf{I}_{\omega}) \text{vec}\mbf{F}_{AA} \left.\frac{\partial q^{\prime}_{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} \\
+	   \numberthis
+\end{align*}
+\begin{landscape}
+\noindent Finally, apply Roth's theorem (Roth 1934), $\left( \mbf{C}^{\intercal} \otimes \mbf{A} \right) \text{vec}\mbf{B} = \text{vec}\mbf{ABC}$ (e.g.,where $\mbf{C}^{\intercal} = \hat{\mbf{p}}^{\intercal}_{AA}$, $\mbf{A} = (\mbf{e}_1 \otimes \mbf{I}_{\omega})$, and $\text{vec}\mbf{B} = \text{vec}\mbf{F}_{AA}$) to rewrite as
+
+\begin{align*}
+	\frac{1}{\lambda_{AA}}(\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{2 \omega g}) \left. \frac{ \partial \text{vec}\mbf{A}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} = 
+		&\frac{(1 - C_{AA})}{\lambda_{AA}}  \text{vec}\big[(\mbf{e}_{2} \otimes \mbf{I}_{\omega}) \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big)\big]  \left.\frac{\partial q^{\prime}_{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}  \\
+		&- \frac{(1 - C_{AA})}{\lambda_{AA}}  \text{vec}\big[(\mbf{e}_{4} \otimes \mbf{I}_{\omega}) \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big)\big]  \left.\frac{\partial q^{\prime}_{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}  \label{eq:partCwithpartials_app} \numberthis
+\end{align*}
+
+\noindent written in terms of the block matrices, this yields
+
+{
+\scriptsize
+\begin{align*}
+	\encircle{C} =& \\
+	\frac{(1 - C_{AA})}{\lambda_{AA}} &\left(
+			\begin{array}{cc|cc|cc}
+					\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0}  \\
+			\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{ \intercal}_{S,AA}}  &  \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{ \intercal}_{X,AA}} &  \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\intercal}_{S,Aa}}& \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\intercal}_{X,Aa}} & \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\intercal}_{S,aa}}&  \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\intercal}_{X,aa}} \\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
+				-\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{ \intercal}_{S,AA}}  &  -\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{ \intercal}_{X,AA}} & - \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\intercal}_{S,Aa}}& -\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\intercal}_{X,Aa}} &- \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\intercal}_{S,aa}}& - \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\intercal}_{X,aa}} \\ \hline
 				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
@@
 \end{align*}
+}
+The above expression requires the derivative of the frequency of allele $A$ in the gamete pool with respect to the population frequency vector:
+
+\begin{equation}
+	\left. \frac{\partial q^{\prime}_{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{P}}}
+\end{equation}
 
@@
 
-Finally, we turn to term \encircle{C} of the Jacobian.
-As far as I can tell, we can follow the steps leading from Eq(A-35) through to Eq(A-46) of de Vries \& Caswell (2019) Appendix A exactly because the structure of our $\tilde{\mbf{A}}$ equations are identical.
-We resume the derivation at our version of their Eq(A-47)...Substitute $\hat{\mbf{p}}^{\intercal} = \left( \hat{\mbf{p}}^{\intercal}_{AA},\hat{\mbf{p}}^{\intercal}_{AA},\mbf{0},\mbf{0},\mbf{0},\mbf{0} \right)$ into the right side of Eq(A-46) so that only terms with $j=1$ and $j=2$ are nonzero, yielding
-
-\begin{align*}
-	\frac{1}{\lambda_{AA}}(\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{2 \omega g}) \left.
-\frac{ \partial \text{vec}\mbf{A}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} &= \frac{1}{\lambda_{AA}} \sum^{6}_{i=1} \big( \hat{\mbf{p}}^{\intercal}_{AA} \otimes (\mbf{e}_{i} \otimes \mbf{I}_{\omega}) \big) \left.
-\frac{ \partial \text{vec}\mbf{A}_{i,1}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} \\
-		&+ \frac{1}{\lambda_{AA}} \sum^{6}_{i=1} \big( \hat{\mbf{p}}^{\intercal}_{AA} \otimes (\mbf{e}_{i} \otimes \mbf{I}_{\omega}) \big) \left.
-\frac{ \partial \text{vec}\mbf{A}_{i,2}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} \numberthis
-\end{align*}
-
-\noindent But, since none of the $\mbf{A}_{i,2}$ are a function of the frequency vector $\mbf{p}$,
-
-\begin{equation}
-	\left. \frac{ \partial \text{vec}\mbf{A}_{i,2}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} = 0,\,\text{for all }i. 
-\end{equation}
-
-\noindent (see Eq{\ref{eq:AtildeCoexistence}}). Next, write down each term in the sum over $i$ and take the derivative  of the $\text{vec}\mbf{A}_{i,1}$'s to obtain
-
-\begin{align*}
-	\frac{1}{\lambda_{AA}}(\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{2 \omega g}) \left. \frac{ \partial \text{vec}\mbf{A}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} = 
-		&\frac{C_{AA}(1 - \delta)}{\lambda_{AA}} \big[ \hat{\mbf{p}}^{\intercal} \otimes (\mbf{e}_{1} \otimes \mbf{I}_{\omega} ) - \hat{\mbf{p}}^{\intercal}_{AA} \otimes (\mbf{e}_{3} \otimes \mbf{I}_{\omega})\big] \text{vec}\mbf{F}_{AA} \\
-		&+ \frac{(1 - C_{AA})}{\lambda_{AA}} \big[ \hat{\mbf{p}}^{\intercal} \otimes (\mbf{e}_{2} \otimes \mbf{I}_{\omega} ) - \hat{\mbf{p}}^{\intercal}_{AA} \otimes (\mbf{e}_{4} \otimes \mbf{I}_{\omega})\big] \text{vec}\mbf{F}_{AA}  \left.\frac{\partial q^{\prime}_{A}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}    \numberthis
-\end{align*}
-
-\noindent Finally, apply Roth's theorem (Roth 1934), $\left( \mbf{C}^{\intercal} \otimes \mbf{A} \right) \text{vec}\mbf{B} = \text{vec}\mbf{ABC}$ (e.g.,where $\mbf{C}^{\intercal} = \hat{\mbf{p}}^{\intercal}_{AA}$, $\mbf{A} = (\mbf{e}_1 \otimes \mbf{I}_{\omega})$, and $\text{vec}\mbf{B} = \text{vec}\mbf{F}_{AA}$) to rewrite as
-
-\begin{align*}
-	\frac{1}{\lambda_{AA}}(\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{2 \omega g}) \left. \frac{ \partial \text{vec}\mbf{A}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} = 
-		&\frac{C_{AA}(1 - \delta)}{\lambda_{AA}} \big[ \text{vec}\big((\mbf{e}_{1} \otimes \mbf{I}_{\omega}) \mbf{F}_{AA} \hat{\mbf{p}}_{AA} \big) - \text{vec}\big((\mbf{e}_{3} \otimes \mbf{I}_{\omega}) \mbf{F}_{AA} \hat{\mbf{p}}_{AA} \big) \big]  \\
-		&+ \frac{(1 - C_{AA})}{\lambda_{AA}} \big[ \text{vec}\big((\mbf{e}_{2} \otimes \mbf{I}_{\omega}) \mbf{F}_{AA} \hat{\mbf{p}}_{AA} \big) - \text{vec}\big((\mbf{e}_{4} \otimes \mbf{I}_{\omega}) \mbf{F}_{AA} \hat{\mbf{p}}_{AA} \big) \big]  \left.\frac{\partial q^{\prime}_{A}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}    \numberthis
-\end{align*}
-
-\noindent written in terms of the block matrices, this yields
-
+ Start with equation (\ref{eq:maleGametePool}) from the main text:
+\begin{equation} \label{eq:maleGametePool_app}
+	\left(
+		\begin{array}{c}
+			q^{\prime}_{A} \\
+			q^{\prime}_{a} \\
+		\end{array} \right) = 
+			\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{n}}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{n} \|} = 
+				\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X+ \mbf{n}^S)}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X+ \mbf{n}^S)\|}.
+\end{equation}
+We will define
+\begin{eqnarray}
+\mathbf{n}^\prime&=&\mbf{n}^X+ \mbf{n}^S, \\
+\mathbf{p}^\prime&=&\frac{\mathbf{n}^\prime}{\|\mathbf{n}^\prime \|}=\frac{\mathbf{n}^\prime}{\|\tilde{\mathbf{n}}\|}, \\
+\end{eqnarray}
+therefore
+\begin{equation}
+  q^\prime_A=\frac{\mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime  \mathbf{p}^\prime}{\mathbf{1}^\tr_{2}\mathbf{W}^\prime\mathbf{F}^\prime\mathbf{p}^\prime}, \label{eq:male_gamete_pool}
+\end{equation}
+ where the one norm can be replaced by $\mathbf{1}^\tr_{2}\mathbf{W}^\prime\mathbf{F}^\prime \mathbf{p}^\prime$ because $\mathbf{p}^\prime$ is nonnegative.
+For convenience, we will denote the normalizing factor in the denominator with $ \mbox{p}_{n}$,
+ \begin{equation}
+   \mbox{p}_{n}=\mathbf{1}^\tr_{2}\mathbf{W}^\prime\mathbf{F}^\prime \mathbf{p}^\prime
+ \end{equation}
+ Taking the derivative of $ q^\prime_A$ yields
+\begin{eqnarray}
+  \frac{\partial   q^\prime_A }{\partial \tilde{\bo p}^\tr}&=& \frac{1}{ \mbox{p}_{n}}\mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime   \frac{\partial  \mathbf{p}^\prime}{\partial \tilde{\bo p}^\tr} - \frac{\mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime \mathbf{p}^\prime}{ \mbox{p}_{n}^2} \left( \mathbf{1}^\tr_{2}\mathbf{W}^\prime\mathbf{F}^\prime \frac{\partial  \mathbf{p}^\prime}{\partial \tilde{\bo p}^\tr} \right).
+\label{eq:56_app}
+\end{eqnarray}
+Recall 
+\begin{equation}
+  \mathbf{p}^\prime=\left(\begin{array} {c}
+\mathbf{p}^S_{AA}+\mathbf{p}^X_{AA}\\
+\mathbf{p}^S_{Aa}+\mathbf{p}^X_{Aa}\\
+\mathbf{p}^S_{aa}+\mathbf{p}^X_{aa}
+  \end{array}\right),
+\end{equation}
+and 
+\begin{equation}
+  \tilde{\bo p}=\left(\begin{array} {c}
+\mathbf{p}_{AA}^S\\
+\mathbf{p}_{AA}^X\\
+\hline
+\mathbf{p}_{Aa}^S\\
+\mathbf{p}_{Aa}^X\\
+\hline
+\mathbf{p}_{aa}^S\\
+\mathbf{p}_{aa}^X
+  \end{array}\right),
+\end{equation}
+to calculate 
+\begin{equation}
+\frac{\partial  \mathbf{p}^\prime}{\partial \tilde{\bo p}^\tr} =  \left(\begin{array} {cccccc}
+\mathbf{I}& \mathbf{I}& \mathbf{0}  &\mathbf{0} &\mathbf{0}  &\mathbf{0} \\
+\mathbf{0} &\mathbf{0}  &\mathbf{I}& \mathbf{I}& \mathbf{0} &\mathbf{0} \\
+\mathbf{0} &\mathbf{0}  &\mathbf{0} &\mathbf{0}  &  \mathbf{I}& \mathbf{I}
+\end{array}\right).
+\end{equation}
+First we will evaluate the first term in the sum in equation (\ref{eq:56_app}),
+\begin{eqnarray}
+	 \frac{1}{ \mbox{p}_{n}}\mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime   \frac{\partial  \mathbf{p}^\prime}{\partial \tilde{\bo p}^\tr} &=&  \frac{1}{ \mbox{p}_{n}} (1,0) \left(\begin{array} {ccc}
+\mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime& \frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{0} \\
+\mathbf{0} &\frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime
+\end{array}\right) \left(\begin{array} {cccccc}
+\mathbf{I}& \mathbf{I}& \mathbf{0}  &\mathbf{0} &\mathbf{0}  &\mathbf{0} \\
+\mathbf{0} &\mathbf{0}  &\mathbf{I}& \mathbf{I}& \mathbf{0} &\mathbf{0} \\
+\mathbf{0} &\mathbf{0}  &\mathbf{0} &\mathbf{0}  &  \mathbf{I}& \mathbf{I}
+\end{array}\right) \nonumber \\
+&=& \frac{1}{ \mbox{p}_{n}} \left( \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime, \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime , \frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime, \frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime, \mathbf{0}, \mathbf{0} \right)\label{eq:first_term_app_reviewer}
+\end{eqnarray}
+\begin{landscape}
+Similarly for the second term in the sum in equation (\ref{eq:56_app}),
+\begin{eqnarray}
+	- \frac{\mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime \mathbf{p}^\prime}{ \mbox{p}_{n}^2} \left( \mathbf{1}^\tr_{2}\mathbf{W}^\prime\mathbf{F}^\prime \frac{\partial  \mathbf{p}^\prime}{\partial \tilde{\bo p}^\tr} \right) &=& -\frac{1}{ \mbox{p}_{n}}(1,1) \left(\begin{array} {ccc}
+\mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime& \frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{0} \\
+\mathbf{0} &\frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime
+\end{array}\right) \left(\begin{array} {cccccc}
+\mathbf{I}& \mathbf{I}& \mathbf{0}  &\mathbf{0} &\mathbf{0}  &\mathbf{0} \\
+\mathbf{0} &\mathbf{0}  &\mathbf{I}& \mathbf{I}& \mathbf{0} &\mathbf{0} \\
+\mathbf{0} &\mathbf{0}  &\mathbf{0} &\mathbf{0}  &  \mathbf{I}& \mathbf{I}
+\end{array}\right) \nonumber \\
+ &=& -\frac{1}{ \mbox{p}_{n}} \left( \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime, \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime , \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime, \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime, \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime , \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime  \right) \label{eq:second_term_app_reviewer}
+\end{eqnarray}
+Finally add equations (\ref{eq:first_term_app_reviewer}) and (\ref{eq:second_term_app_reviewer}) to obtain
+\begin{eqnarray}
+  \frac{\partial   q^\prime_A }{\partial \tilde{\bo p}^\tr}\bigg\rvert_{\hat{\bo p}}&=& \frac{1}{ \mbox{p}_{n}} \left( \mathbf{0}, \mathbf{0}, -\frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime, -\frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime, -\mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime, -\mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime  \right) \label{eq:79_app},
+\end{eqnarray}
+where at the boundary
+\begin{eqnarray}
+  \mbox{p}_{n}&=&\mathbf{1}^\tr_\omega \mathbf{F}_{AA}^\prime \left( \hat{\mathbf{p}}^S_{AA}+\hat{\mathbf{p}}^X_{AA} \right), \\
+  \mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime \mathbf{p}^\prime&=&\mathbf{1}^\tr_\omega \mathbf{F}_{AA}^\prime \left( \hat{\mathbf{p}}^S_{AA}+\hat{\mathbf{p}}^X_{AA} \right)=  \mbox{p}_{n}
+\end{eqnarray}
+
+Finally, plugging equation (\ref{eq:79_app}) into (\ref{eq:partCwithpartials_app}) yields
+
+{
+\scriptsize
 \begin{align*}
 	\encircle{C} =& \\
-	\frac{1}{\lambda_{AA}} &\left(
+	\frac{(1 - C_{AA})}{\mbox{p}_{n}\lambda_{AA}} &\left(
 			\begin{array}{cc|cc|cc}
-				\mbf{0} & C_{AA}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & C_{aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \\
-				\mbf{0} & (1 - C_{AA}) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\prime \intercal}_{AA}} & \mbf{0} & (1 - C_{Aa}) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\prime \intercal}_{Aa}} & \mbf{0} & (1 - C_{aa}) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\prime \intercal}_{aa}} \\ \hline
-				\mbf{0} & \mbf{0} & \mbf{0} & -C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & \mbf{0} \\
-				\mbf{0} & -(1 - C_{AA}) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\prime \intercal}_{AA}} & \mbf{0} & -(1 - C_{Aa}) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\prime \intercal}_{Aa}} & \mbf{0} & -(1 - C_{aa}) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\prime \intercal}_{aa}} \\ \hline
+					\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0}  \\
+				\mbf{0}   &  	\mbf{0}  & -\frac{1}{2} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime& -\frac{1}{2} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & - \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime&  - \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime \\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
+					\mbf{0}   &  	\mbf{0}  & \frac{1}{2} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime& \frac{1}{2} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &  \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime&  \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime \\  \hline
 				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
@@
 \end{align*}
-
-The above expression requires the derivative of the frequency of allele $A$ in the gamete pool with respect to the population frequency vector:
-
-\begin{equation}
-	\left. \frac{\partial q^{\prime}_{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{P}}}
-\end{equation}
-
-\noindent Here, I believe we can again follow the exact same steps as described in Eq(A-50) through Eq(A-60) in Appendix A of de Vries \& Caswell (2019) to give, finally,
-
-\begin{align*}
-	\frac{1}{\lambda_{AA}} (\hat{\mbf{p}}^{\intercal} &\otimes \mbf{I}_{2 \omega g}) \left. \frac{\partial \text{vec}\mbf{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} = \\
-	\frac{1}{\lambda_{AA}} &\left(
-			\begin{array}{cc|cc|cc}
-				\mbf{0} & C_{AA}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & C_{aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \\
-				\mbf{0} & \mbf{0} & \mbf{0} & -\frac{(1 - C_{Aa})}{2 p_{\text{n}}} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \mbf{0} & -\frac{(1 - C_{aa})}{2 p_{\text{n}}} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{aa} \\ \hline
-				\mbf{0} & \mbf{0} & \mbf{0} & -C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & \mbf{0} \\
-				\mbf{0} & \mbf{0} & \mbf{0} & \frac{(1 - C_{Aa})}{2 p_{\text{n}}} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \mbf{0} & \frac{(1 - C_{aa})}{2 p_{\text{n}}} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{aa} \\ \hline
-				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
-				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} 
-			\end{array} \right)  \numberthis			
-\end{align*}
+}
+
+\end{landscape}
 
 %%%%%%%%%%%%%%%%%%%%
+\begin{landscape}
 \subsubsection{The Jacobian}
@@
 
-\begin{landscape}
-
+{
+\footnotesize
 \begin{align*} \label{eq:combinedM}
 	\mbf{M} = 
-		\frac{1}{\lambda_{AA}} &\left(\begin{array}{cc|cc|cc}
-				\mcal{U}^S_{AA} + C_{AA} (1 - \delta) \mbf{F}_{AA} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
-				\mbf{F}_{AA} (1 - C_{AA}) & \mcal{U}^X_{AA} & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
-				\mbf{0} & \mbf{0} & \mcal{U}^S_{Aa} + \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\
-				\mbf{0} & \mbf{0} & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & \mcal{U}^X_{Aa} & \mbf{F}_{aa} (1 - C_{aa}) & \mbf{0}\\ \hline
-				\mbf{0} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mcal{U}^S_{aa} + C_{aa} (1 - \delta) \mbf{F}_{aa} & \mbf{0} \\ 
-				\mbf{0} & \mbf{0} & \mbf{0}  & \mbf{0} & \mbf{0} & \mcal{U}^X_{aa}
+			 +\frac{1}{\lambda_{AA}} &
+						   \left(
+			\begin{array}{cc|cc|cc}
+		 \mathbf{U}^S_{AA} +C_{AA} (1 - \delta) \mbf{F}_{AA} & C_{AA} (1 - \delta) \mbf{F}_{AA}& \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}& \mbf{0} & \mbf{0} \\
+				(1 - C_{AA}) \mbf{F}_{AA} &\mathbf{U}^X_{AA} + (1 - C_{AA}) \mbf{F}_{AA}  & \frac{1}{2}(1 - C_{Aa}) \mbf{F}_{Aa}  & \frac{1}{2} (1 - C_{Aa})\mbf{F}_{Aa} & \mbf{0} & \mbf{0} \\ \hline
+				\mbf{0} & \mbf{0} & \mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} & (1 - C_{aa}) \mbf{F}_{aa} & (1 - C_{aa}) \mbf{F}_{aa}\\ \hline
+				\mbf{0} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa} &C_{aa} (1 - \delta) \mbf{F}_{aa} \\
+				\mbf{0} & \mbf{0} & \mbf{0}  & \mbf{0} & \mbf{0} &\mathbf{U}^X_{aa} 
 			\end{array} \right) \\
@@
 			\end{array} \right) \\
-		-\frac{1}{\lambda_{AA}} &\left(
+		-\frac{1}{\lambda_{AA}}& \left(
 			\begin{array}{cc|cc|cc}
-				C_{AA}(1 - \delta) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & \mbf{0} & C_{Aa}(1 - \delta) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & \mbf{0} & C_{aa}(1 - \delta) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} & \mbf{0} \\
-				(1 - C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & \mbf{0} & (1 - C_{Aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & \mbf{0} & (1 - C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} & \mbf{0} \\ \hline
+				(1 - \delta C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & (1 - \delta C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & (1 - \delta C_{Aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & (1 - \delta C_{Aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa}  & (1 - \delta C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} & (1 - \delta C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} \\
+				 	(1 - \delta C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & 	(1 - \delta C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & (1 - \delta C_{Aa})  \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & (1 - \delta C_{Aa})  \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & (1 - \delta C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} & (1 - \delta C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa}\\ \hline
 				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
@@
 				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0}
-			\end{array} \right) \\
-		\frac{1}{\lambda_{AA}} &\left(
+			\end{array} \right)\\
+		\frac{(1 - C_{AA})}{\mbox{p}_{n}\lambda_{AA}} &\left(
 			\begin{array}{cc|cc|cc}
-				\mbf{0} & C_{AA}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & C_{aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \\
-				\mbf{0} & \mbf{0} & \mbf{0} & -\frac{(1 - C_{Aa})}{2 p_{\text{n}}} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \mbf{0} & -\frac{(1 - C_{aa})}{2 p_{\text{n}}} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{aa} \\ \hline
-				\mbf{0} & \mbf{0} & \mbf{0} & -C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} & \mbf{0} \\
-				\mbf{0} & \mbf{0} & \mbf{0} & \frac{(1 - C_{Aa})}{2 p_{\text{n}}} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \mbf{0} & \frac{(1 - C_{aa})}{2 p_{\text{n}}} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{aa} \\ \hline
+					\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0}  \\
+				\mbf{0}   &  	\mbf{0}  & -\frac{1}{2} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime& -\frac{1}{2} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & - \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime&  - \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime \\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
+					\mbf{0}   &  	\mbf{0}  & \frac{1}{2} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime& \frac{1}{2} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &  \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime&  \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime \\  \hline
 				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
@@
 \end{align*}
+}
 
@@
 %%%%%%%%%%%%%%%
+\begin{landscape}
 \subsubsection*{Eigenvalues of the Jacobian}
 
-Similar to Lotte's result, the Jacobian matrix given in Eq(\ref{eq:combinedM}) is upper block triangular, and the eigenvalues of $\mbf{M}$ are therefore the eigenvalues of the diagonal blocks.
+The Jacobian matrix given in Eq(\ref{eq:combinedM}) is upper block triangular, and the eigenvalues of $\mbf{M}$ are therefore the eigenvalues of the diagonal blocks.
 However, because of the way that self-fertilization alters the projection matrix, our blocks are slightly different than Lotte's model.
@@
 
+{
  \begin{align*} \label{eq:M22}
 	\mbf{M}_{22} = 
-		\frac{1}{\lambda_{AA}} &\left(\begin{array}{cc|c}
-				\mcal{U}^S_{Aa} + \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & -C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} \\
-				\frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & \mcal{U}^X_{Aa} + \frac{(1 - C_{Aa})}{2 p_n} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \mbf{F}_{aa} (1 - C_{aa}) \\ \hline
-				\frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mcal{U}^S_{aa} + C_{aa} (1 - \delta) \mbf{F}_{aa} \\ 
-			\end{array} \right) \numberthis
-\end{align*}
+		\frac{1}{\lambda_{AA}}\left(\begin{array}{cc}
+\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  \\
+ \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime
+			\end{array} \right), \numberthis
+\end{align*}
+}
+where 
+\begin{equation}
+  \mbox{p}_{n}=\mathbf{1}^\tr_\omega \mathbf{F}_{AA}^\prime \left( \hat{\mathbf{p}}^S_{AA}+\hat{\mathbf{p}}^X_{AA} \right)	
+\end{equation}
+
+[Note that the outcrossing terms are the sum of a term due to heterozygote ovules being fertilized by pollen of the resident ($\frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}$), and heterozygote pollen fertilizing resident females ($\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime$).
+]
 
@@
 
+{
 \begin{align*} \label{eq:eigAA}
 	\tilde{\zeta}_{AA} = 
-		\frac{1}{\lambda_{AA}} \rho&\left(\begin{array}{cc|c}
-				\mcal{U}^S_{Aa} + \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & -C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} \\
-				\frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & \mcal{U}^X_{Aa} + \frac{(1 - C_{Aa})}{2 p_n} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \mbf{F}_{aa} (1 - C_{aa}) \\ \hline
-				\frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mcal{U}^S_{aa} + C_{aa} (1 - \delta) \mbf{F}_{aa} \\ 
+		\frac{1}{\lambda_{AA}}\rho\left(\begin{array}{cc}
+\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  \\
+ \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime
 			\end{array} \right). \numberthis
 \end{align*}
-
-\noindent Rearranging the projection matrix so that outcrossing comes on top of selfing in each block simplifies things, but in the end the leading eigenvalue evaluated at the $aa$ boundary is
-
-\begin{align*} \label{eq:eigaa}
-	\tilde{\zeta}_{aa} = 
-		\frac{1}{\lambda_{AA}} \rho&\left(\begin{array}{c|cc}
-				\mcal{U}^S_{AA} + C_{AA} (1 - \delta) \mbf{F}_{AA} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}\\ \hline
-				\mbf{F}_{AA} (1 - C_{AA}) & \mcal{U}^X_{Aa} + \frac{(1 - C_{Aa})}{2 p_n} (\mbf{F}_{aa} \hat{\mbf{p}}_{aa}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) \\
-				\mbf{0} & -C_{Aa}(1 - \delta) (\mbf{F}_{aa} \hat{\mbf{p}}_{aa}) & \mcal{U}^S_{Aa} + \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} \\
-			\end{array} \right). \numberthis
-\end{align*}
+}
+
+% \noindent Rearranging the projection matrix so that outcrossing comes on top of selfing in each block simplifies things, but in the end the leading eigenvalue evaluated at the $aa$ boundary is
+
+% \begin{align*} \label{eq:eigaa}
+% 	\tilde{\zeta}_{aa} = 
+% 		\frac{1}{\lambda_{AA}} \rho&\left(\begin{array}{c|cc}
+% 				\mcal{U}^S_{AA} + C_{AA} (1 - \delta) \mbf{F}_{AA} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}\\ \hline
+% 				\mbf{F}_{AA} (1 - C_{AA}) & \mcal{U}^X_{Aa} + \frac{(1 - C_{Aa})}{2 p_n} (\mbf{F}_{aa} \hat{\mbf{p}}_{aa}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) \\
+% 				\mbf{0} & -C_{Aa}(1 - \delta) (\mbf{F}_{aa} \hat{\mbf{p}}_{aa}) & \mcal{U}^S_{Aa} + \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} \\
+% 			\end{array} \right). \numberthis
+% \end{align*}
 
@@
 	\lambda_{AA} < 
-		\rho \left(\begin{array}{cc|c}
-				\mcal{U}^S_{Aa} + \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & -C_{Aa}(1 - \delta) (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) & \mbf{0} \\
-				\frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & \mcal{U}^X_{Aa} + \frac{(1 - C_{Aa})}{2 p_n} (\mbf{F}_{AA} \hat{\mbf{p}}_{AA}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \mbf{F}_{aa} (1 - C_{aa}) \\ \hline
-				\frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mcal{U}^S_{aa} + C_{aa} (1 - \delta) \mbf{F}_{aa} \\ 
-			\end{array} \right), \\ 
-\end{equation} 
+			\rho\left(\begin{array}{cc}
+\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  \\
+ \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime
+			\end{array} \right)., \\ 
+\end{equation}
+
 \begin{equation} \label{eq:coexist_aa}
 	\lambda_{aa} < 
-		\rho \left(\begin{array}{c|cc}
-				\mcal{U}^S_{AA} + C_{AA} (1 - \delta) \mbf{F}_{AA} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}\\ \hline
-				\mbf{F}_{AA} (1 - C_{AA}) & \mcal{U}^X_{Aa} + \frac{(1 - C_{Aa})}{2 p_n} (\mbf{F}_{aa} \hat{\mbf{p}}_{aa}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) \\
-				\mbf{0} & -C_{Aa}(1 - \delta) (\mbf{F}_{aa} \hat{\mbf{p}}_{aa}) & \mcal{U}^S_{Aa} + \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} \\
-			\end{array} \right). 
-\end{equation}
+			\rho\left(\begin{array}{cc}
+\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  \\
+ \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2\mbox{p}_{n}} (1 - C_{aa})\mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2\mbox{p}_{n}} (1 - C_{aa})\mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime
+			\end{array} \right).\\ 
+\end{equation}
+
+\subsection*{Sanity check: no selfing}
+Above conditions should reduce to the conditions in our Amnat paper (de Vries and Caswell, 2019) in the absence of selfing, i.e. $C_{ij}=0$ for all genotypes, for which
+
+\begin{align*} \label{eq:eigAA}
+	\tilde{\zeta}_{AA} = 
+		\frac{1}{\lambda_{AA}}\rho\left(\begin{array}{cc}
+\mathbf{U}^S_{Aa}  & \mathbf{0} \\
+ \frac{1}{2} \mbf{F}_{Aa}+\frac{1}{2\mbox{p}_{n}}\mbf{F}_{AA} \hat{\mbf{p}}_{X,AA} \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} \mbf{F}_{Aa} +\frac{1}{2\mbox{p}_{n}} \mbf{F}_{AA} \hat{\mbf{p}}_{X,AA} \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime
+			\end{array} \right), \numberthis
+\end{align*}
+which reduces to 
+\begin{align*} \label{eq:eigAA}
+	\tilde{\zeta}_{AA} = 
+		\frac{1}{\lambda_{AA}}\rho\left(\begin{array}{c}
+\mathbf{U}^X_{Aa} + \frac{1}{2} \mbf{F}_{Aa} +\frac{1}{2\mbox{p}_{n}} \mbf{F}_{AA} \hat{\mbf{p}}_{X,AA} \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime
+			\end{array} \right), \numberthis
+\end{align*}
+when $\mathbf{U}^S_{Aa}=\mathbf{U}^X_{Aa}$. This is the same as the AmNat deVries-Caswell equation :-)
+
+\newpage
+
+\subsection*{Simplification: when how-you-were-produced does not affect your demographic rates}
+If we assume that whether you were produced through selfing or outcrossing does not affect your survival, i.e.
+$\mathbf{U}^X_{ij}=\mathbf{U}^S_{ij}$, then individuals are only classified by their genotype and their age/stage.
+The population vector is then
+
+\begin{equation}
+	\tilde{\mbf{n}}(t) =  \left[
+								\begin{array}{c}
+									\mbf{n}_{AA}(t) \\
+									\mbf{n}_{Aa}(t) \\
+									\mbf{n}_{aa}(t) \\ 
+						\end{array} \right].
+\end{equation}
+
+The population projection matrix becomes
+
+\begin{align*}
+	&\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
+		 \underbrace{\left(
+			\begin{array}{ccc}
+				\mathbf{U}_{AA} & \mbf{0} & \mbf{0}\\ 
+				\mbf{0} & \mathbf{U}_{Aa} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \mathbf{U}_{aa} 
+			\end{array} \right)}_{\tilde{\mbf{U}}}  \\ 
+	&+\underbrace{\left(
+			\begin{array}{ccc}
+				C_{AA} (1 - \delta) \mbf{F}_{AA} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} \\ 
+				\mbf{0} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} \\
+				\mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & C_{aa} (1 - \delta) \mbf{F}_{aa} 
+			\end{array} \right)}_{\tilde{\mbf{F}^S}} \\
+			&+\underbrace{\left(
+			\begin{array}{ccc}
+				q^{\prime}_{A} (1 - C_{AA})\mbf{F}_{AA} & \frac{1}{2} q^{\prime}_{A} (1 - C_{Aa})\mbf{F}_{Aa}  & \mbf{0} \\
+				q^{\prime}_{a} (1 - C_{AA}) \mbf{F}_{AA} & \frac{1}{2} (1 - C_{Aa})\mbf{F}_{Aa}  & q^{\prime}_{A} (1 - C_{aa})\mbf{F}_{aa}   \\
+				\mbf{0} & \frac{1}{2} q^{\prime}_{a} (1 - C_{Aa})\mbf{F}_{Aa} & q^{\prime}_{a} (1 - C_{aa}) \mbf{F}_{aa} 
+			\end{array} \right)}_{\tilde{\mbf{F}^X}} \numberthis
+\end{align*}
+
+And the coexistence conditions become:
+\begin{equation} \label{eq:coexist_AA}
+	\lambda_{AA} < 
+			\rho\left(\begin{array}{c}
+\mathbf{U}_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2} (1 - C_{AA})\frac{\mbf{F}_{AA}\hat{\mbf{p}}_{AA}\otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime}{\mathbf{1}^\tr_\omega \mathbf{F}_{AA}^\prime \hat{\mathbf{p}}_{AA}	}
+			\end{array} \right), \\ 
+\end{equation}
+
+\begin{equation} \label{eq:coexist_AA}
+	\lambda_{aa} < 
+			\rho\left(\begin{array}{c}
+\mathbf{U}_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2 } (1 - C_{aa})\frac{\mbf{F}_{aa}\hat{\mbf{p}}_{aa}\otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime}{\mathbf{1}^\tr_\omega \mathbf{F}_{aa}^\prime \hat{\mathbf{p}}_{aa}	}
+			\end{array} \right). \\ 
+\end{equation} 
+\end{landscape}
+
+\subsubsection*{Two stage case when how-you-were-produced does not matter}
+
+\begin{align*}
+	\mbf{U}_{ij} = &\left(
+					\begin{array}{cc}
+						\sigma_{ij}(1 - \gamma_{ij}) & 0 \\
+						\sigma_{ij} \gamma_{ij}     & s_{ij}
+					\end{array}
+				\right)  
+\end{align*}
+
+\noindent where $\sigma$ and $s$ are the juvenile and adult stage survival rates, and $\gamma$ is the maturation rate from juvenile to adult stages.
+Any or all of these parameters may differ among genotypes, so that selection can operate on stage-specific viability, development, and/or fertility.
+For simplicity, we assume these are constant among genotypes.
+
+The fertility matrices through male and female function are
+
+\begin{equation}
+	\mbf{F}_{ij} = \left(
+					\begin{array}{cc}
+						0 &  f_{ij} \\
+						0 & 0
+					\end{array}
+				\right),
+\end{equation}
+\noindent and
+\begin{equation}
+	\mbf{F}^{\prime}_{ii} = \left(
+					\begin{array}{cc}
+						0 &  f^{\prime}_{ii} \\
+						0 & 0
+					\end{array}
+				\right),
+\end{equation}
+
+\noindent where $f$ is the genotype specific adult fertility though female function, and $f^{\prime}_{ii}$ is the same through male function. 
+\begin{table}[htbp]
+ \centering
+ \caption{\bf Fitness expressions for Sexually Antagonistic selection in hermaphrodites}
+\begin{tabular}{lccc}
+ \toprule
+					&  \multicolumn{3}{c}{{\textit{Genotype}}} \\ 
+\cline{2-4}
+					& $AA$			& $Aa$ 					& $aa$ 		\\ \hline
+Female function:	& $f$		& $f - h_f s_f$	& $f - s_f$ \\	
+Male function:		& $f^\prime - s_m$& $f^\prime - h_m s_m$	& $f^\prime$ 		\\	
+\hline
+\end{tabular}
+\end{table}
+
+And the coexistence conditions become:
+\begin{eqnarray} \label{eq:coexist_AA}
+\lambda_{AA}> 
+			\rho\left(\left(
+					\begin{array}{cc}
+						\sigma_{Aa}(1 - \gamma_{Aa}) & 0 \\
+						\sigma_{Aa} \gamma_{Aa}     & s_{Aa}
+					\end{array}
+				\right)   + \frac{1}{2} (1 - C_{Aa}) \left(
+					\begin{array}{cc}
+						0 &  f_{Aa} \\
+						0 & 0
+					\end{array}
+				\right) +\frac{1}{2} (1 - C_{AA})\frac{f^\prime_{Aa}}{f^\prime_{AA}}\left(
+					\begin{array}{cc}
+						0 &  f_{AA} \\
+						0 & 0
+					\end{array}
+				\right)\right). 
+\end{eqnarray}
+
+\newpage
+% \begin{landscape}
+% \begin{align*} \label{eq:AtildeCoexistence}
+% 	\tilde{\mbf{A}}[\tilde{\mbf{n}}] &= \tilde{\mbf{U}} + \tilde{\mbf{F}}\\
+% 		&= \left(
+% 			\begin{array}{cc|cc|cc}
+% 				\mcal{U}^S_{AA} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
+% 				\mbf{0} & \mcal{U}^X_{AA} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\\hline
+% 				\mbf{0} & \mbf{0} & \mcal{U}^S_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\
+% 				\mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{Aa} & \mbf{0} &  \mbf{0}\\\hline
+% 				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^S_{Aa} & \mbf{0} \\ 
+% 				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{Aa}
+% 			\end{array} \right) \\ 
+% 		&+ \left(
+% 			\begin{array}{cc|cc|cc}
+% 				C_{AA} (1 - \delta) \mbf{F}_{AA} & C_{AA} (1 - \delta) \mbf{F}_{AA}  & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} \\ 
+% 				q^{\prime}_{A}  (1 - C_{AA}) \mbf{F}_{AA}& q^{\prime}_{A} (1 - C_{AA}) \mbf{F}_{AA} & \frac{1}{2} q^{\prime}_{A} (1 - C_{Aa})\mbf{F}_{Aa}  & \frac{1}{2} q^{\prime}_{A} (1 - C_{Aa})\mbf{F}_{Aa}  & \mbf{0} & \mbf{0} \\ \hline
+% 				\mbf{0} & \mbf{0} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}& \mbf{0} & \mbf{0} \\
+% 				q^{\prime}_{a} (1 - C_{AA}) \mbf{F}_{AA} & q^{\prime}_{a}(1 - C_{AA})  \mbf{F}_{AA} & \frac{1}{2}  (1 - C_{Aa}) \mbf{F}_{Aa}& \frac{1}{2} (1 - C_{Aa})\mbf{F}_{Aa} & q^{\prime}_{A} (1 - C_{aa})\mbf{F}_{aa}  & q^{\prime}_{A}  (1 - C_{aa})\mbf{F}_{aa}\\ \hline
+% 				\mbf{0} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & C_{aa} (1 - \delta) \mbf{F}_{aa} &  C_{aa} (1 - \delta) \mbf{F}_{aa}\\ 
+% 				\mbf{0} & \mbf{0} & \frac{1}{2} q^{\prime}_{a}  (1 - C_{Aa}) \mbf{F}_{Aa} & \frac{1}{2} q^{\prime}_{a}  (1 - C_{Aa})\mbf{F}_{Aa}& q^{\prime}_{a}(1 - C_{aa})  \mbf{F}_{aa} & q^{\prime}_{a} (1 - C_{aa})\mbf{F}_{aa}
+% 			\end{array} \right) \numberthis
+% \end{align*}
+
+% \end{landscape}
 

### 2020-07-10 notes/notes-SA-partSelf-wDemography.tex
Update notes-SA-partSelf-wDemography.tex

@@
 {
+\footnotesize
  \begin{align*} \label{eq:M22}
 	\mbf{M}_{22} = 
-		\frac{1}{\lambda_{AA}}\left(\begin{array}{cc}
-\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  \\
- \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime
+		\frac{1}{\lambda_{AA}}\left(\begin{array}{ccc}
+\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & 0 \\
+ \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & (1-C_{aa})\mbf{F}_{aa}\\
+  \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa}  \\ 
 			\end{array} \right), \numberthis
@@
 {
+\footnotesize
 \begin{align*} \label{eq:eigAA}
 	\tilde{\zeta}_{AA} = 
-		\frac{1}{\lambda_{AA}}\rho\left(\begin{array}{cc}
-\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  \\
- \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime
+		\frac{1}{\lambda_{AA}}\rho\left(\begin{array}{ccc}
+\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & 0 \\
+ \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & (1-C_{aa})\mbf{F}_{aa}\\
+  \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa}  \\ 
 			\end{array} \right). \numberthis
@@
 
+{\footnotesize
 \begin{equation} \label{eq:coexist_AA}
 	\lambda_{AA} < 
-			\rho\left(\begin{array}{cc}
-\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  \\
- \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime
-			\end{array} \right)., \\ 
-\end{equation}
-
-\begin{equation} \label{eq:coexist_aa}
+			\rho\left(\begin{array}{ccc}
+\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & 0 \\
+ \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & (1-C_{aa})\mbf{F}_{aa}\\
+  \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa}  \\ 
+			\end{array} \right), \\ 
+\end{equation}
+
+\begin{equation} \label{eq:coexist_AA}
 	\lambda_{aa} < 
-			\rho\left(\begin{array}{cc}
-\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  \\
- \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2\mbox{p}_{n}} (1 - C_{aa})\mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2\mbox{p}_{n}} (1 - C_{aa})\mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime
-			\end{array} \right).\\ 
-\end{equation}
+			\rho\left(\begin{array}{ccc}
+\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & 0 \\
+ \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{aa})\mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{aa})\mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & (1-C_{AA})\mbf{F}_{AA}\\
+  \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{AA} +C_{AA} (1 - \delta) \mbf{F}_{AA}  \\ 
+			\end{array} \right), \\ 
+\end{equation} 
+}
 
@@
 
-And the coexistence conditions become:
-\begin{equation} \label{eq:coexist_AA}
+\hl{This is a guess, I have not worked through the actual algebra yet} And the coexistence conditions become:
+\begin{equation} \label{eq:coexist_bigAA}
 	\lambda_{AA} < 
-			\rho\left(\begin{array}{c}
-\mathbf{U}_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2} (1 - C_{AA})\frac{\mbf{F}_{AA}\hat{\mbf{p}}_{AA}\otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime}{\mathbf{1}^\tr_\omega \mathbf{F}_{AA}^\prime \hat{\mathbf{p}}_{AA}	}
+			\rho\left(\begin{array}{cc}
+\mathbf{U}_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2} (1 - C_{AA})\frac{\mbf{F}_{AA}\hat{\mbf{p}}_{AA}\otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime}{\mathbf{1}^\tr_\omega \mathbf{F}_{AA}^\prime \hat{\mathbf{p}}_{AA}	} & (1-C_{aa}) \mbf{F}_{aa} \\
+\frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa}  \\ 
 			\end{array} \right), \\ 
@@
 
-\begin{equation} \label{eq:coexist_AA}
-	\lambda_{aa} < 
-			\rho\left(\begin{array}{c}
-\mathbf{U}_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2 } (1 - C_{aa})\frac{\mbf{F}_{aa}\hat{\mbf{p}}_{aa}\otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime}{\mathbf{1}^\tr_\omega \mathbf{F}_{aa}^\prime \hat{\mathbf{p}}_{aa}	}
-			\end{array} \right). \\ 
+\begin{equation} \label{eq:coexist_baa}
+	\lambda_{AA} < 
+			\rho\left(\begin{array}{cc}
+\mathbf{U}_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2} (1 - C_{aa})\frac{\mbf{F}_{aa}\hat{\mbf{p}}_{aa}\otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime}{\mathbf{1}^\tr_\omega \mathbf{F}_{aa}^\prime \hat{\mathbf{p}}_{aa}	} & (1-C_{AA}) \mbf{F}_{AA} \\
+\frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{AA} +C_{AA} (1 - \delta) \mbf{F}_{AA}  \\ 
+			\end{array} \right), \\ 
 \end{equation} 
@@
 
-And the coexistence conditions become:
+And the invasion condition becomes (\hl{based on my earlier guess, so still an (educated?) guess}):
+
 \begin{eqnarray} \label{eq:coexist_AA}
 \lambda_{AA}> 
-			\rho\left(\left(
-					\begin{array}{cc}
-						\sigma_{Aa}(1 - \gamma_{Aa}) & 0 \\
-						\sigma_{Aa} \gamma_{Aa}     & s_{Aa}
+			\rho\left(
+					\begin{array}{cc|cc}
+						\sigma_{Aa}(1 - \gamma_{Aa}) &  \frac{1}{2} (1 - C_{Aa}) f_{Aa} +\frac{1}{2} (1 - C_{AA})\frac{f^\prime_{Aa}f_{AA}}{f^\prime_{AA}} &0 & (1-C_{aa})f_{aa}\\
+						\sigma_{Aa} \gamma_{Aa}     & s_{Aa}& 0 & 0\\ \hline
+							0 & \frac{1}{4}C_{Aa}(1-\delta)f_{Aa} &\sigma_{aa}(1 - \gamma_{aa})  & C_{aa}(1-\delta)f_{aa}\\
+								0 & 0 & \sigma_{aa} \gamma_{aa}     & s_{aa}
 					\end{array}
-				\right)   + \frac{1}{2} (1 - C_{Aa}) \left(
-					\begin{array}{cc}
-						0 &  f_{Aa} \\
-						0 & 0
-					\end{array}
-				\right) +\frac{1}{2} (1 - C_{AA})\frac{f^\prime_{Aa}}{f^\prime_{AA}}\left(
-					\begin{array}{cc}
-						0 &  f_{AA} \\
-						0 & 0
-					\end{array}
-				\right)\right). 
+				\right). 
 \end{eqnarray}

### 2020-07-17 notes/notes-SA-partSelf-wDemography.tex
Update notes-SA-partSelf-wDemography.tex

@@
 \newpage
+\end{landscape}
 
@@
 
-\hl{This is a guess, I have not worked through the actual algebra yet} And the coexistence conditions become:
+\subsection*{Coexistence conditions}
+
+\noindent Start from equation (\ref{eq:subJacobian}), repeated here for convenience:
+
+\begin{align*} \label{eq:subJacobian_simplified}
+	\mbf{M} &=  \mbf{B}[\tilde{\mbf{p}}] + \left( \tilde{\mbf{p}}^{\intercal} \otimes \mbf{1}_{2 \omega g} \right) \frac{ \partial \text{vec}{\mbf{B}} }{ \partial \tilde{\mbf{p}}^{\intercal} } \bigg\rvert_{\hat{\mbf{p}}}, \\
+			&= \underbrace{ \mbf{B}[\hat{\mbf{p}}]}_{\encircle{A}} - 
+			   \underbrace{ \frac{ 1 }{\lambda_{AA}^2 } (\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{\omega g}) \left( \text{vec} \tilde{\mbf{A}} \right) \left( \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}} \right)}_{\encircle{B}} \\
+			&~~~~~~~~~~~+ \underbrace{\frac{ 1 }{\lambda_{AA} } (\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{\omega g}) \left.
+\frac{\partial \text{vec} \tilde{\mbf{A}}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}}_{\encircle{C}},
+\end{align*}
+
+%%%%%%%%%%%%%%%%
+\subsubsection{Components of the Jacobian}
+
+The next task is to work out each of the three terms \encircle{A}, \encircle{B}, and \encircle{C}, in $\mbf{M}$. Beginning with \encircle{A}, we have
+
+\begin{align*}
+	\mbf{B}[\hat{\mbf{p}}] &= \frac{ \tilde{\mbf{A}}[\hat{\mbf{p}}] }{ \mbf{1}^{\intercal}_{\omega g} \tilde{\mbf{A}}[\hat{\mbf{p}}] \hat{\mbf{p}} } = \\
+	&\frac{1}{\lambda_{AA}}\left(
+			\begin{array}{ccc}
+				\mathbf{U}_{AA} & \mbf{0} & \mbf{0}\\ 
+				\mbf{0} & \mathbf{U}_{Aa} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \mathbf{U}_{aa} 
+			\end{array} \right)\\ 
+	&+\frac{1}{\lambda_{AA}}\left(
+			\begin{array}{ccc}
+				(1 - \delta C_{AA}) \mbf{F}_{AA} & \frac{1}{2} \left[1-\frac{1}{2} C_{Aa} - \frac{1}{2} \delta C_{Aa} \right] \mbf{F}_{Aa} & \mbf{0} \\ 
+				\mbf{0} & \frac{1}{2} \left[1- \delta C_{Aa} \right] \mbf{F}_{Aa}  & (1 - C_{aa})\mbf{F}_{aa}  \\
+				\mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & C_{aa} (1 - \delta) \mbf{F}_{aa} 
+			\end{array} \right)  \numberthis
+\end{align*}
+
+\noindent see Eq{\ref{eq:AtildeCoexistence}}. Next, we have \encircle{B},
+
+\begin{equation}
+	\encircle{B} = - \frac{1}{\lambda_{AA}^2} \left( \hat{\mbf{p}}^{\intercal} \otimes \mbf{1}_{2 \omega g} \right) \left( \text{vec} \tilde{\mbf{A}} \right) \left( \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}} \right).
+\end{equation}
+
+\noindent Again using Roth's theorem (Roth 1934), $\left( \mbf{C}^{\intercal} \otimes \mbf{A} \right) \text{vec}\mbf{B} = \text{vec}\mbf{ABC}$, we can simplify as follows:
+
+\begin{align*}
+	\left( \hat{\mbf{p}}^{\intercal} \otimes \mbf{1}_{2 \omega g} \right) \left( \text{vec} \tilde{\mbf{A}}[\hat{\mbf{p}}] \right) &= \text{vec} \left( \mbf{I}_{2 \omega g}\, \tilde{\mbf{A}}[\hat{\mbf{p}}]\, \hat{\mbf{p}} \right) \\
+					   &= \lambda_{AA} \hat{\mbf{p}}, \numberthis
+\end{align*}
+
+\noindent so that
+
+\begin{equation}
+	\encircle{B} = - \frac{1}{\lambda_{AA}} \hat{\mbf{p}} \left( \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}}[\hat{\mbf{p}}] \right).
+\end{equation}
+
+\noindent Substitute the population vector at the $AA$ boundary,
+
+\begin{equation}
+	\hat{\mbf{p}} = \left(
+			\begin{array}{c}
+							\hat{\mbf{p}}_{AA} \\
+							0                  \\ 
+							0                  
+			\end{array} \right)
+\end{equation}
+
+\noindent and rewrite in terms of the block matrices to obtain
+
+{\footnotesize
+\begin{align*}
+	\encircle{B} =& \\
+	&-\frac{1}{\lambda_{AA}} \left(
+			\begin{array}{ccc}
+				\hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \left[\mbf{U}_{AA}+(1 - \delta C_{AA}) \mbf{F}_{AA}\right]  & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \left[\mbf{U}_{Aa}+(1 - \delta C_{Aa}) \mbf{F}_{Aa}\right] & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \left[\mbf{U}_{aa}+(1 - \delta C_{aa}) \mbf{F}_{aa}\right] \\
+				\mbf{0} & \mbf{0} & \mbf{0}  \\
+				\mbf{0} & \mbf{0} & \mbf{0} 
+			\end{array} \right) \\ \numberthis			
+\end{align*}
+}
+
+Finally, we turn to term \encircle{C} of the Jacobian.
+As far as I can tell, we can follow the steps leading from Eq(A-35) through to Eq(A-46) of de Vries \& Caswell (2019) Appendix A exactly because the structure of our $\tilde{\mbf{A}}$ equations are identical.
+We resume the derivation at our version of their Eq(A-47)...Substitute $\hat{\mbf{p}}^{\intercal} = \left( \hat{\mbf{p}}^{\intercal}_{AA},\mbf{0},\mbf{0} \right)$ into the right side of Eq(A-46) so that only terms with $j=1$ are nonzero, yielding
+
+\begin{align*}
+	\frac{1}{\lambda_{AA}}(\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{2 \omega g}) \left.
+\frac{ \partial \text{vec}\mbf{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} &= \frac{1}{\lambda_{AA}} \sum^{3}_{i=1} \big( \hat{\mbf{p}}^{\intercal}_{S,AA} \otimes (\mbf{e}_{i} \otimes \mbf{I}_{\omega}) \big) \left.
+\frac{ \partial \text{vec}\mbf{A}_{i,1}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} \numberthis
+\end{align*}
+
+Write down each term in the sum over $i$ and take the derivative of $\mbox{vec}\mathbf{A}_{11}$ and $\mbox{vec}\mathbf{A}_{21}$ to obtain
+\begin{equation}
+\frac{1}{\lambda_{AA}}(\hat{\mathbf{p}}^\tr \otimes \mathbf{I}_{\omega g})\frac{\partial \mbox{vec} \mathbf{A}}{\partial \tilde{\bo p}^\tr}
+=
+\frac{1-C_{AA}}{\lambda_{AA}}\left[  \hat{\bo p}_{AA}^\tr\otimes \left(\mathbf{e}_{1} \otimes \mathbf{I}_\omega\right)-\hat{\bo p}_{AA}^\tr\otimes \left(\mathbf{e}_{2} \otimes \mathbf{I}_\omega\right) \right] \mbox{vec} \left(\mathbf{F}_{AA}\right)\frac{\partial q^\prime_A}{\partial \tilde{\bo p}^\tr}.
+\end{equation}
+
+\noindent (see Eq{\ref{eq:AtildeCoexistence}}).
+Finally, apply Roth's theorem (Roth 1934), $\left( \mbf{C}^{\intercal} \otimes \mbf{A} \right) \text{vec}\mbf{B} = \text{vec}\mbf{ABC}$ (e.g.,where $\mbf{C}^{\intercal} = \hat{\mbf{p}}^{\intercal}_{AA}$, $\mbf{A} = (\mbf{e}_1 \otimes \mbf{I}_{\omega})$, and $\text{vec}\mbf{B} = \text{vec}\mbf{F}_{AA}$) to rewrite as
+
+\begin{align*}
+	\frac{1}{\lambda_{AA}}(\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{2 \omega g}) \left. \frac{ \partial \text{vec}\mbf{A}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} = 
+		&\frac{(1 - C_{AA})}{\lambda_{AA}}  \text{vec}\big[(\mbf{e}_{1} \otimes \mbf{I}_{\omega}) \mbf{F}_{AA} \hat{\mbf{p}}_{AA}\big]  \left.\frac{\partial q^{\prime}_{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}  \\
+		&- \frac{(1 - C_{AA})}{\lambda_{AA}}  \text{vec}\big[(\mbf{e}_{2} \otimes \mbf{I}_{\omega}) \mbf{F}_{AA} \hat{\mbf{p}}_{AA}\big]  \left.\frac{\partial q^{\prime}_{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}  \label{eq:partCwithpartials_app} \numberthis
+\end{align*}
+
+\noindent written in terms of the block matrices, this yields
+
+{
+\scriptsize
+\begin{align*}
+	\encircle{C} =& \\
+	\frac{(1 - C_{AA})}{\lambda_{AA}} &\left(
+			\begin{array}{ccc}
+			\mbf{F}_{AA}\hat{\mbf{p}}_{AA} \otimes \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{ \intercal}_{AA}}  &  \mbf{F}_{AA}\hat{\mbf{p}}_{AA} \otimes \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{ \intercal}_{Aa}}  &  \mbf{F}_{AA}\hat{\mbf{p}}_{AA} \otimes \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{ \intercal}_{aa}}\\
+				\mbf{0} & \mbf{0} & \mbf{0}  \\
+				\mbf{0} & \mbf{0} & \mbf{0} 
+			\end{array} \right)  \numberthis			
+\end{align*}
+}
+The above expression requires the derivative of the frequency of allele $A$ in the gamete pool with respect to the population frequency vector:
+
+\begin{equation}
+	\left. \frac{\partial q^{\prime}_{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{P}}}
+\end{equation}
+
+ Start with equation (\ref{eq:maleGametePool}) from the main text:
+\begin{equation} \label{eq:maleGametePool_app}
+	\left(
+		\begin{array}{c}
+			q^{\prime}_{A} \\
+			q^{\prime}_{a} \\
+		\end{array} \right) = 
+			\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} \tilde{\mbf{n}}}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} \tilde{\mbf{n}} \|} = 
+				\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} \tilde{\mbf{p}}}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} \tilde{\mbf{p}}\|},
+\end{equation}
+therefore
+\begin{equation}
+  q^\prime_A=\frac{\mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime  \mathbf{p}}{\mathbf{1}^\tr_{2}\mathbf{W}^\prime\mathbf{F}^\prime\mathbf{p}}, \label{eq:male_gamete_pool}
+\end{equation}
+ where the one norm can be replaced by $\mathbf{1}^\tr_{2}\mathbf{W}^\prime\mathbf{F}^\prime \tilde{\mbf{p}}$ because $\tilde{\mbf{p}}$ is nonnegative.
+For convenience, we will denote the normalizing factor in the denominator with $ \mbox{p}_{n}$,
+ \begin{equation}
+   \mbox{p}_{n}=\mathbf{1}^\tr_{2}\mathbf{W}^\prime\mathbf{F}^\prime \tilde{\mbf{p}}
+ \end{equation}
+ Taking the derivative of $ q^\prime_A$ yields
+\begin{eqnarray}
+  \frac{\partial   q_A }{\partial \tilde{\bo p}^\tr}&=& \frac{1}{ \mbox{p}_{n}}\mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime   \frac{\partial  \tilde{\mbf{p}}}{\partial \tilde{\bo p}^\tr} - \frac{\mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime \tilde{\mbf{p}}}{ \mbox{p}_{n}^2} \left( \mathbf{1}^\tr_{2}\mathbf{W}^\prime\mathbf{F}^\prime \frac{\partial  \tilde{\mbf{p}}}{\partial \tilde{\bo p}^\tr} \right).
+\label{eq:56_app}
+\end{eqnarray}
+Evaluate the first term in the sum in equation (\ref{eq:56_app}),
+\begin{eqnarray}
+	 \frac{1}{ \mbox{p}_{n}}\mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime   \frac{\partial  \mathbf{p}^\prime}{\partial \tilde{\bo p}^\tr} &=&  \frac{1}{ \mbox{p}_{n}} (1,0) \left(\begin{array} {ccc}
+\mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime& \frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{0} \\
+\mathbf{0} &\frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime
+\end{array}\right) \left(\begin{array} {ccc}
+\mathbf{I}&\mathbf{0}  &\mathbf{0} \\
+\mathbf{0} &\mathbf{I}& \mathbf{0}  \\
+\mathbf{0} &\mathbf{0}  &  \mathbf{I}
+\end{array}\right) \nonumber \\
+&=& \frac{1}{ \mbox{p}_{n}} \left( \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime,  \frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime,  \mathbf{0} \right)\label{eq:first_term_app_reviewer}
+\end{eqnarray}
+Similarly for the second term in the sum in equation (\ref{eq:56_app}),
+\begin{eqnarray}
+	- \frac{\mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime \mathbf{p}^\prime}{ \mbox{p}_{n}^2} \left( \mathbf{1}^\tr_{2}\mathbf{W}^\prime\mathbf{F}^\prime \frac{\partial  \mathbf{p}^\prime}{\partial \tilde{\bo p}^\tr} \right) &=& -\frac{1}{ \mbox{p}_{n}}(1,1) \left(\begin{array} {ccc}
+\mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime& \frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{0} \\
+\mathbf{0} &\frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime
+\end{array}\right)  \left(\begin{array} {ccc}
+\mathbf{I}&\mathbf{0}  &\mathbf{0} \\
+\mathbf{0} &\mathbf{I}& \mathbf{0}  \\
+\mathbf{0} &\mathbf{0}  &  \mathbf{I}
+\end{array}\right) \nonumber \\
+ &=& -\frac{1}{ \mbox{p}_{n}} \left( \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime, \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime, \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime   \right) \label{eq:second_term_app_reviewer}
+\end{eqnarray}
+Finally add equations (\ref{eq:first_term_app_reviewer}) and (\ref{eq:second_term_app_reviewer}) to obtain
+\begin{eqnarray}
+  \frac{\partial   q^\prime_A }{\partial \tilde{\bo p}^\tr}\bigg\rvert_{\hat{\bo p}}&=& \frac{1}{ \mbox{p}_{n}} \left( \mathbf{0},  -\frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime, -\mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime \right) \label{eq:79_app},
+\end{eqnarray}
+where at the boundary
+\begin{eqnarray}
+  \mbox{p}_{n}&=&\mathbf{1}^\tr_\omega \mathbf{F}_{AA}^\prime \hat{\mathbf{p}}_{AA}, \\
+  \mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime \mathbf{p}^\prime&=&\mathbf{1}^\tr_\omega \mathbf{F}_{AA}^\prime \hat{\mathbf{p}}_{AA}=  \mbox{p}_{n}
+\end{eqnarray}
+
+Finally, plugging equation (\ref{eq:79_app}) into (\ref{eq:partCwithpartials_app}) yields
+
+{
+\begin{align*}
+	\encircle{C} =& \\
+	\frac{(1 - C_{AA})}{\mbox{p}_{n}\lambda_{AA}} &\left(
+			\begin{array}{ccc}
+				\mbf{0}   & -\frac{1}{2} \mbf{F}_{AA} \hat{\mbf{p}}_{AA} \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime&  - \mbf{F}_{AA} \hat{\mbf{p}}_{AA} \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime \\
+						\mbf{0}   & \frac{1}{2} \mbf{F}_{AA} \hat{\mbf{p}}_{AA} \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime&  \mbf{F}_{AA} \hat{\mbf{p}}_{AA}  \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime \\
+				\mbf{0} & \mbf{0} & \mbf{0} 
+			\end{array} \right)  \numberthis			
+\end{align*}
+}
+
+%%%%%%%%%%%%%%%%%%%%
+\subsubsection{The Jacobian}
+
+Putting all the pieces together into $M$, we get the Jacobian:
+
+{
+\footnotesize
+\begin{align*} \label{eq:combinedM}
+	\mbf{M} = 
+			& \frac{1}{\lambda_{AA}}\left(
+			\begin{array}{ccc}
+				\mathbf{U}_{AA}+(1 - \delta C_{AA}) \mbf{F}_{AA}  & \frac{1}{2} \left[1-\frac{1}{2} C_{Aa} - \frac{1}{2} \delta C_{Aa} \right] \mbf{F}_{Aa}  & \mbf{0}\\ 
+				\mbf{0} & \mathbf{U}_{Aa} +\frac{1}{2} \left[1- \delta C_{Aa} \right] \mbf{F}_{Aa} & (1 - C_{aa})\mbf{F}_{aa}  \\
+				\mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & \mathbf{U}_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa} 
+			\end{array} \right)\\ 
+	&-\frac{1}{\lambda_{AA}} \left(
+			\begin{array}{ccc}
+				\hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \left[\mbf{U}_{AA}+(1 - \delta C_{AA}) \mbf{F}_{AA}\right]  & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \left[\mbf{U}_{Aa}+(1 - \delta C_{Aa}) \mbf{F}_{Aa}\right] & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \left[\mbf{U}_{aa}+(1 - \delta C_{aa}) \mbf{F}_{aa}\right] \\
+				\mbf{0} & \mbf{0} & \mbf{0}  \\
+				\mbf{0} & \mbf{0} & \mbf{0} 
+			\end{array} \right)\\
+			&+\frac{1}{\lambda_{AA}}\frac{(1 - C_{AA})}{\mathbf{1}^\tr_\omega \mathbf{F}_{AA}^\prime \hat{\mathbf{p}}_{AA}} \left(
+			\begin{array}{ccc}
+				\mbf{0}   & -\frac{1}{2} \mbf{F}_{AA} \hat{\mbf{p}}_{AA} \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime&  - \mbf{F}_{AA} \hat{\mbf{p}}_{AA} \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime \\
+						\mbf{0}   & \frac{1}{2} \mbf{F}_{AA} \hat{\mbf{p}}_{AA} \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime&  \mbf{F}_{AA} \hat{\mbf{p}}_{AA}  \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime \\
+				\mbf{0} & \mbf{0} & \mbf{0} 
+			\end{array} \right) 
+			\numberthis
+\end{align*}
+}
+
+The coexistence conditions become:
 \begin{equation} \label{eq:coexist_bigAA}
@@
 			\rho\left(\begin{array}{cc}
-\mathbf{U}_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2} (1 - C_{AA})\frac{\mbf{F}_{AA}\hat{\mbf{p}}_{AA}\otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime}{\mathbf{1}^\tr_\omega \mathbf{F}_{AA}^\prime \hat{\mathbf{p}}_{AA}	} & (1-C_{aa}) \mbf{F}_{aa} \\
-\frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa}  \\ 
+\mathbf{U}_{Aa} + \frac{1}{2} (1 - \delta C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2} (1 - C_{AA})\frac{\mbf{F}_{AA}\hat{\mbf{p}}_{AA}\otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime}{\mathbf{1}^\tr_\omega \mathbf{F}_{AA}^\prime \hat{\mathbf{p}}_{AA}	} & (1-C_{aa}) \mbf{F}_{aa} \\
+\frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa}  \\ 
 			\end{array} \right), \\ 
@@
 			\rho\left(\begin{array}{cc}
-\mathbf{U}_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2} (1 - C_{aa})\frac{\mbf{F}_{aa}\hat{\mbf{p}}_{aa}\otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime}{\mathbf{1}^\tr_\omega \mathbf{F}_{aa}^\prime \hat{\mathbf{p}}_{aa}	} & (1-C_{AA}) \mbf{F}_{AA} \\
-\frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{AA} +C_{AA} (1 - \delta) \mbf{F}_{AA}  \\ 
+\mathbf{U}_{Aa} + \frac{1}{2} (1 - \delta C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2} (1 - C_{aa})\frac{\mbf{F}_{aa}\hat{\mbf{p}}_{aa}\otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime}{\mathbf{1}^\tr_\omega \mathbf{F}_{aa}^\prime \hat{\mathbf{p}}_{aa}	} & (1-C_{AA}) \mbf{F}_{AA} \\
+\frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}_{AA} +C_{AA} (1 - \delta) \mbf{F}_{AA}  \\ 
 			\end{array} \right), \\ 
 \end{equation}
-\end{landscape}
 
@@
 					\begin{array}{cc|cc}
-						\sigma_{Aa}(1 - \gamma_{Aa}) &  \frac{1}{2} (1 - C_{Aa}) f_{Aa} +\frac{1}{2} (1 - C_{AA})\frac{f^\prime_{Aa}f_{AA}}{f^\prime_{AA}} &0 & (1-C_{aa})f_{aa}\\
+						\sigma_{Aa}(1 - \gamma_{Aa}) &  \frac{1}{2} (1 - \delta C_{Aa}) f_{Aa} +\frac{1}{2} (1 - C_{AA})\frac{f^\prime_{Aa}f_{AA}}{f^\prime_{AA}} &0 & (1-C_{aa})f_{aa}\\
 						\sigma_{Aa} \gamma_{Aa}     & s_{Aa}& 0 & 0\\ \hline

### 2020-08-28 R/functions-Figs.R
Updated FunnelFig(), started cal. eigenvalue in fwdDemModelSim()

@@ -137,13 +137,49 @@ fibonacci.scale  <-  function(n) {
 
 #' Figs showing Kidwell Funnel plots with simulation results
 #' and 1-locus pop. gen. invasion conditions
-FunnelPlots  <-  function() {
+FunnelPlots  <-  function(df1, df2, df3, df4) {
+
+# Make filenames for import from df names
+    fName1  <-  paste('./output/simData/', df1, '.csv', sep="")
+    fName2  <-  paste('./output/simData/', df2, '.csv', sep="")
+    fName3  <-  paste('./output/simData/', df3, '.csv', sep="")
+    fName4  <-  paste('./output/simData/', df4, '.csv', sep="")
+
+# Extract plotting parameter values from df names
+    d1   <-  strsplit(df1, '_')[[1]][5:8]
+    parsA  <-  list(
+                    "hf"  =  as.numeric(strsplit(d1[1],'f')[[1]][2]),
+                    "hm"  =  as.numeric(strsplit(d1[2],'m')[[1]][2]),
+                    "C"  =  as.numeric(strsplit(d1[3],'C')[[1]][2]),
+                    "delta"  =  as.numeric(strsplit(d1[4],'a')[[1]][2])
+                    )
+    d2   <-  strsplit(df2, '_')[[1]][5:8]
+    parsB  <-  list(
+                    "hf"  =  as.numeric(strsplit(d2[1],'f')[[1]][2]),
+                    "hm"  =  as.numeric(strsplit(d2[2],'m')[[1]][2]),
+                    "C"  =  as.numeric(strsplit(d2[3],'C')[[1]][2]),
+                    "delta"  =  as.numeric(strsplit(d2[4],'a')[[1]][2])
+                    )
+    d3   <-  strsplit(df3, '_')[[1]][5:8]
+    parsC  <-  list(
+                    "hf"  =  as.numeric(strsplit(d3[1],'f')[[1]][2]),
+                    "hm"  =  as.numeric(strsplit(d3[2],'m')[[1]][2]),
+                    "C"  =  as.numeric(strsplit(d3[3],'C')[[1]][2]),
+                    "delta"  =  as.numeric(strsplit(d3[4],'a')[[1]][2])
[…]

### 2020-08-28 R/functions-Simulations.R

@@ -60,8 +60,8 @@ popGen_qHat_DomRev  <-  function(h, sf, sm, C) {
 #'
 #' Parameters:
 #' dims: 	vector c(om, g), with om = # stages, g = # genotypes
-#' theta: 	vector of length 4, c(sigma_J, sigma_A, gamma, f_ii)
-fwdDemModelSim  <-  function(	om = 2, g = 3, theta = c(0.6,0.6,0.05,5.9), theta_prime = c(0.6,0.6,0.05,5.9), 
+#' theta: 	vector of length 7, c(sigmaS_J, sigmaS_A, sigmaX_J, sigmaX_A, gammaS, gammaX, f_ii)
+fwdDemModelSim  <-  function(	om = 2, g = 3, theta = c(0.6, 0.6, 0.6, 0.6, 0.05, 0.05, 5.9), theta_prime = c(0.6, 0.6, 0.6, 0.6, 0.05, 0.05, 5.9), 
 								hf = 1/2, hm = 1/2, sf = 0.1, sm = 0.105, C = 0, delta = 0, tlimit = 10^4, Ainvade = FALSE, intInit = FALSE, ...) {
 
 	# Create identity and ones matrices
@@ -77,7 +77,7 @@ fwdDemModelSim  <-  function(	om = 2, g = 3, theta = c(0.6,0.6,0.05,5.9), theta_
 
 	# BASELINE PARAMETERS for survival and 
 	# fertility through Female and Male function
-	# theta=[sigma_J, sigma_A, gamma, f_ii]
+	# theta=[sigmaS_J, sigmaS_A, sigmaX_J, sigmaX_A, gammaS, gammaX, f_ii]
 	theta       <-  rep.col(theta,3)
 	theta_prime <- rep.col(theta_prime,3)
 
@@ -86,23 +86,26 @@ fwdDemModelSim  <-  function(	om = 2, g = 3, theta = c(0.6,0.6,0.05,5.9), theta_
 	fii_prime  <-  c(1 - sm, 1 - hm*sm, 1)
 
 	##SELECTION DIFFERENTIAL FEMALE-FUNCTION
-	theta[4,]  <-  theta[4,]*fii
+	theta[7,]  <-  theta[7,]*fii
 
 	#SELECTION DIFFERENTIAL MALE-FUNCTION
-	theta_prime[4,]  <-  theta_prime[4,]*fii_prime
[…]

### 2020-08-28 makeFigs.R

@@ -23,5 +23,22 @@ source('R/functions-Figs.R')
 # PRELIM FIGS
 ###############
 
-toPdf(FunnelPlots(), figPath(name='FunnelPlots_MEqAInv.pdf'), width=10, height=7)
+toPdf(FunnelPlots(df1 = 'demSimsSfxSm_MinorEqAlleleInv_sMax0.15_nSamples1000_hf0.5_hm0.5_C0_delta0',
+				  df2 = 'demSimsSfxSm_MinorEqAlleleInv_sMax0.15_nSamples1000_hf0.5_hm0.5_C0.5_delta0',
+				  df3 = 'demSimsSfxSm_MinorEqAlleleInv_sMax0.15_nSamples1000_hf0.25_hm0.25_C0_delta0',
+				  df4 = 'demSimsSfxSm_MinorEqAlleleInv_sMax0.15_nSamples1000_hf0.25_hm0.25_C0.5_delta0'), 
+figPath(name='FunnelPlots_MEqAInv.pdf'), width=10, height=7)
 embed_fonts(figPath(name='FunnelPlots_MEqAInv.pdf'))
+
+toPdf(FunnelPlots(df1 = 'demSimsSfxSm_MinorEqAlleleInv_sMax0.15_nSamples1000_hf0.5_hm0.5_C0_delta0',
+				  df2 = 'demSimsSfxSm_MinorEqAlleleInv_sMax0.15_nSamples1000_hf0.5_hm0.5_C0.5_delta0.5',
+				  df3 = 'demSimsSfxSm_MinorEqAlleleInv_sMax0.15_nSamples1000_hf0.25_hm0.25_C0_delta0',
+				  df4 = 'demSimsSfxSm_MinorEqAlleleInv_sMax0.15_nSamples1000_hf0.25_hm0.25_C0.5_delta0.5'), 
+figPath(name='FunnelPlots_MEqAInv_ID.pdf'), width=10, height=7)
+embed_fonts(figPath(name='FunnelPlots_MEqAInv.pdf'))
+
+
+
+
+
+

### 2020-08-28 matlab/deVries_Caswell_2019_AmNat_Two_stage_model.m

@@ -106,10 +106,10 @@ for i=1:3
 end % for i
 
 
-%Calculating the largest eigenvalue of the Jacobian for both boundaryies
+%Calculating the largest eigenvalue of the Jacobian for both boundaries
 
-inv_bigA=max(eig((U{2}+0.5*F{1}+0.5*F{2}))); 
-inv_a=max(eig(U{2}+0.5*F{3}+0.5*F{2}));
+inv_bigA=max(eig(U{2}+0.5*F{1}+0.5*F{2})); 
+inv_a=   max(eig(U{2}+0.5*F{3}+0.5*F{2}));
 
 %Dominant eigenvalue of the Jacobian at the AA boundary
 x_bigA=inv_bigA/lambda(1)

### 2020-08-28 run-Simulations.R

@@ -38,7 +38,7 @@ source('R/functions-Simulations.R')
 					)
 
 	selLoop(sMax = 0.15, nSamples=1000,
-			om = 2, g = 3, theta = c(0.6,0.6,0.05,6), theta_prime = c(0.6,0.6,0.05,6), 
+			om = 2, g = 3, theta = c(0.6,0.6,0.6,0.6,0.05,0.05,6), theta_prime = c(0.6,0.6,0.6,0.6,0.05,0.05,6), 
 			hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, tlimit = 10^5, returnRes = FALSE)
 
 	# Intermediate Selfing
@@ -51,7 +51,7 @@ source('R/functions-Simulations.R')
 					)
 
 	selLoop(sMax = 0.15, nSamples=1000,
-			om = 2, g = 3, theta = c(0.6,0.6,0.05,6), theta_prime = c(0.6,0.6,0.05,6), 
+			om = 2, g = 3, theta = c(0.6,0.6,0.6,0.6,0.05,0.05,6), theta_prime = c(0.6,0.6,0.6,0.6,0.05,0.05,6), 
 			hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, tlimit = 10^5, returnRes = FALSE)
 
 	# Intermediate Selfing w/ I.D.
@@ -60,11 +60,11 @@ source('R/functions-Simulations.R')
 					"hf"     =  1/2,
 					"hm"     =  1/2,
 					"C"      =  1/2,
-					"delta"  =  1/4
+					"delta"  =  3/4
 					)
 
 	selLoop(sMax = 0.15, nSamples=1000,
-			om = 2, g = 3, theta = c(0.6,0.6,0.05,6), theta_prime = c(0.6,0.6,0.05,6), 
+			om = 2, g = 3, theta = c(0.6,0.6,0.6,0.6,0.05,0.05,6), theta_prime = c(0.6,0.6,0.6,0.6,0.05,0.05,6), 
 			hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, tlimit = 10^5, returnRes = FALSE)
 
 
@@ -82,7 +82,7 @@ source('R/functions-Simulations.R')
 					)
 
 	selLoop(sMax = 0.15, nSamples=1000,
[…]

### 2020-08-31 notes/notes-SA-partSelf-wDemography.tex
Updated FunnelFig(), started cal. eigenvalue in fwdDemModelSim()
added zeta_i function, calculate in simulations

@@
 % Change default margins
-\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
+\usepackage[top=0.75in, bottom=0.75in, left=0.75in, right=0.75in]{geometry}
 
@@
 
-\newpage
+%\newpage
 % \begin{landscape}
@@
 %%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\begin{thebibliography}{}
-
-\bibitem[{Chang and Kirkpatrick(2018)Chang and Kirkpatrick}]{ChangKirkpatrick2018}
-Chang, C., and M.~Kirkpatrick. 2018.
-\newblock Inversions are bigger on the X chromosome.
-\newblock Mol.~Ecol. 2018:1--8.
-
-\bibitem[{Charlesworth et.~al.(1987)Charlesworth et.~al.}]{CharlesworthBarton1987}
-Charlesworth, B., J.~A. Coyne, and N. Barton. 1987.
-\newblock The relative rates of evolution of sex chromosomes and autosomes.
-\newblock American Naturalist 130:113--146.
-
-\bibitem[{Connallon et al.(2018)Connallon et al.}]{ConnallonOlito2018}
-Connallon, T., C.~Olito, L.~Dutoit, H.~Popali, F.~Ruzicka, and L. Yong. 2018.
-\newblock Local adaptation and the evolution of inversions on sex chromosomes and autosomes.
-\newblock Phil.~Trans.~Roy.~Soc.~B 373:20170423.
-
-\bibitem[{Orr(2000)Orr}]{Orr2000}
-Orr, H.~A. 2000.
-\newblock The rate of adaptation in asexuals.
-\newblock Genetics 155:961--968.
-
-\bibitem[{Orr and Kim(1998)Orr and Kim}]{OrrKim1998}
-Orr, H.~A. and Y.~Kim. 1998.
-\newblock An adaptive hypothesis for the evolution of the Y chromosome.
-\newblock Genetics 150:1693--1698.
-
-\end{thebibliography}
-
 \end{document}

### 2020-08-31 R/functions-Simulations.R

@@ -55,13 +55,48 @@ popGen_qHat_DomRev  <-  function(h, sf, sm, C) {
 	qHat
 }
 
+
+###########################################
+#' Eigenvalue Calculator for full demographic model
+#' 
+
+calcZeta  <-  function(om, Fii, Fii_pr, USi, UXi, pHat_AA, pHat_aa, C, delta, lambda_full){
+	n_AABound  <-  round(c(C*(1 - delta)*100*c(141.3,18.2,0,0,0,0),(1 - C)*100*c(141.3,18.2,0,0,0,0)))[c(1,2,7,8,3,4,9,10,5,6,11,12)]
+	n_aaBound  <-  round(c(C*(1 - delta)*100*c(0,0,0,0,141.3,18.2),(1 - C)*100*c(0,0,0,0,141.3,18.2)))[c(1,2,7,8,3,4,9,10,5,6,11,12)]
+	pHat_AA    <-  n_AABound/c(sum(n_AABound[1:2]), sum(n_AABound[1:2]), sum(n_AABound[3:4]), sum(n_AABound[3:4]), 1,1,1,1,1,1,1,1)
+	pHat_AA[is.na(pHat_AA)]  <-  0
+	pHat_aa    <-  n_aaBound/c(1,1,1,1,1,1,1,1, sum(n_aaBound[9:10]), sum(n_aaBound[9:10]), sum(n_aaBound[11:12]), sum(n_aaBound[11:12]))
+	pHat_aa[is.na(pHat_aa)]  <-  0
+	pn_AA      <-  ones(c(om))%*%Fii_pr[,,1]%*%(pHat_AA[1:2] + pHat_AA[3:4])
+	pn_aa      <-  ones(c(om))%*%Fii_pr[,,3]%*%(pHat_aa[9:10] + pHat_aa[11:12])
+	M22_AA  <-  rbind(
+					cbind(USi[,,2] + C*(1 - delta)*Fii[,,2]/2, C*(1 - delta)*Fii[,,2]/2, zeros(c(2,2))),
+					cbind(((1 - C)/2)*Fii[,,2] + c(((1 - C)/(2*pn_AA)))*Fii[,,1]%*%kronecker((pHat_AA[1:2] + pHat_AA[3:4]), (ones(om)%*%Fii_pr[,,2])),
[…]

### 2020-08-31 makeFigs.R

@@ -31,9 +31,9 @@ figPath(name='FunnelPlots_MEqAInv.pdf'), width=10, height=7)
 embed_fonts(figPath(name='FunnelPlots_MEqAInv.pdf'))
 
 toPdf(FunnelPlots(df1 = 'demSimsSfxSm_MinorEqAlleleInv_sMax0.15_nSamples1000_hf0.5_hm0.5_C0_delta0',
-				  df2 = 'demSimsSfxSm_MinorEqAlleleInv_sMax0.15_nSamples1000_hf0.5_hm0.5_C0.5_delta0.5',
+				  df2 = 'demSimsSfxSm_MinorEqAlleleInv_sMax0.15_nSamples1000_hf0.5_hm0.5_C0.5_delta0.125',
 				  df3 = 'demSimsSfxSm_MinorEqAlleleInv_sMax0.15_nSamples1000_hf0.25_hm0.25_C0_delta0',
-				  df4 = 'demSimsSfxSm_MinorEqAlleleInv_sMax0.15_nSamples1000_hf0.25_hm0.25_C0.5_delta0.5'), 
+				  df4 = 'demSimsSfxSm_MinorEqAlleleInv_sMax0.15_nSamples1000_hf0.25_hm0.25_C0.5_delta0.125'), 
 figPath(name='FunnelPlots_MEqAInv_ID.pdf'), width=10, height=7)
 embed_fonts(figPath(name='FunnelPlots_MEqAInv.pdf'))
 

### 2020-08-31 run-Simulations.R

@@ -55,17 +55,18 @@ source('R/functions-Simulations.R')
 			hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, tlimit = 10^5, returnRes = FALSE)
 
 	# Intermediate Selfing w/ I.D.
-	# C = 1/2, delta = 0.25
+	# C = 1/2, delta = 0.5
 	pars  <-  list(
 					"hf"     =  1/2,
 					"hm"     =  1/2,
 					"C"      =  1/2,
-					"delta"  =  3/4
+					"delta"  =  1/8
 					)
 
 	selLoop(sMax = 0.15, nSamples=1000,
 			om = 2, g = 3, theta = c(0.6,0.6,0.6,0.6,0.05,0.05,6), theta_prime = c(0.6,0.6,0.6,0.6,0.05,0.05,6), 
 			hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, tlimit = 10^5, returnRes = FALSE)
+
 
 
 
@@ -99,21 +100,27 @@ source('R/functions-Simulations.R')
 			hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, tlimit = 10^5, returnRes = FALSE)
 
 	# Intermediate Selfing w/ I.D.
-	# C = 1/2, delta = 0.25
+	# C = 1/2, delta = 0.5
 	pars  <-  list(
 					"hf"     =  1/4,
 					"hm"     =  1/4,
 					"C"      =  1/2,
-					"delta"  =  1/2
+					"delta"  =  1/8
 					)
 
 	selLoop(sMax = 0.15, nSamples=1000,
 			om = 2, g = 3, theta = c(0.6,0.6,0.6,0.6,0.05,0.05,6), theta_prime = c(0.6,0.6,0.6,0.6,0.05,0.05,6), 
 			hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, tlimit = 10^5, returnRes = FALSE)
 
+
+
+
+
+
+
+
+
 ## Same plots showing larger s_f x s_m parameter space
-
-
 
 ## Additive SA Fitness
 ## hf = hm = 1/2
@@ -150,7 +157,7 @@ source('R/functions-Simulations.R')
 					"hf"     =  1/2,
 					"hm"     =  1/2,
 					"C"      =  1/2,
-					"delta"  =  3/4
+					"delta"  =  1/4
 					)
 
 	selLoop(sMax = 1, nSamples=1000,

### 2020-09-01 R/functions-Figs.R
MWE showing zeta disagrees with sims

@@ -203,12 +203,6 @@ layout <- layout(layout.mat,respect=TRUE)
         # Calculate additional variables for plotting
         inv_A  <-  popGen_A_invade(hf=dat$hf[1], hm=dat$hm[1], sm=smLine, C=dat$C[1])
         inv_a  <-  popGen_a_invade(hf=dat$hf[1], hm=dat$hm[1], sm=smLine, C=dat$C[1])
-        inv_a_polygon  <-  inv_a
-        inv_a_polygon[inv_a > 0.15]  <-  0.15
-        x      <-  dat[round(dat$Eq_paa, digits=5)==1,]
-        fix_a  <-  x[x$extinct==0,]
-        x      <-  dat[round(dat$Eq_pAA, digits=5)==1,]
-        fix_A  <-  x[x$extinct==0,]
         # Make the plot
         par(omi=rep(0.5, 4), mar = c(3,3,0.5,0.5), bty='o', xaxt='s', yaxt='s')
         plot(NA, axes=FALSE, type='n', main='',xlim = c(0,0.15), ylim = c(0,0.15), ylab='', xlab='', cex.lab=1.2)
@@ -344,6 +338,216 @@ legend(#"topright",inset=c(-0.2,0),
         points(sf ~ sm, pch=21, bg=COLS$AFix, data=fix_a)
         points(sf[extinct == 1] ~ sm[extinct == 1], pch=21, bg=COLS$extinct, data=dat)
         points(sf[poly == 1 & extinct == 0] ~ sm[poly == 1 & extinct == 0], pch=21, bg=COLS$poly, data=dat)
+        #  pop. gen. invasion conditions
+        lines(inv_A[inv_A <= 0.15] ~ smLine[inv_A <= 0.15], lwd=2, col=COLS$invLine)
+        lines(inv_a[inv_a <= 0.15] ~ smLine[inv_a <= 0.15], lwd=2, col=COLS$invLine)
+#        polygon(c(smLine,rev(smLine)), c(inv_a_polygon, rev(inv_A)), col=transparentColor('grey80', 0.6), border='grey70')
+        # axes        
+        axis(1, las=1)
+        axis(2, las=1, labels=NA)
+        proportionalLabel(0.5, -0.3, expression(paste(italic(s[m]))), cex=1.2, adj=c(0.5, 0.5), xpd=NA, srt=0)
[…]

### 2020-09-01 R/functions-Simulations.R

@@ -59,33 +59,83 @@ popGen_qHat_DomRev  <-  function(h, sf, sm, C) {
 ###########################################
 #' Eigenvalue Calculator for full demographic model
 #' 
-
-calcZeta  <-  function(om, Fii, Fii_pr, USi, UXi, pHat_AA, pHat_aa, C, delta, lambda_full){
+#(C*(1 - delta)/(1 - C*delta))
+#(1 - (C*(1 - delta)/(1 - C*delta)))
+calcZeta  <-  function(om, Fii, Fii_pr, USi, UXi, pHat_AA, pHat_aa, C, delta, lambda_AA_aa){
 	n_AABound  <-  round(c(C*(1 - delta)*100*c(141.3,18.2,0,0,0,0),(1 - C)*100*c(141.3,18.2,0,0,0,0)))[c(1,2,7,8,3,4,9,10,5,6,11,12)]
 	n_aaBound  <-  round(c(C*(1 - delta)*100*c(0,0,0,0,141.3,18.2),(1 - C)*100*c(0,0,0,0,141.3,18.2)))[c(1,2,7,8,3,4,9,10,5,6,11,12)]
-	pHat_AA    <-  n_AABound/c(sum(n_AABound[1:2]), sum(n_AABound[1:2]), sum(n_AABound[3:4]), sum(n_AABound[3:4]), 1,1,1,1,1,1,1,1)
-	pHat_AA[is.na(pHat_AA)]  <-  0
-	pHat_aa    <-  n_aaBound/c(1,1,1,1,1,1,1,1, sum(n_aaBound[9:10]), sum(n_aaBound[9:10]), sum(n_aaBound[11:12]), sum(n_aaBound[11:12]))
-	pHat_aa[is.na(pHat_aa)]  <-  0
-	pn_AA      <-  ones(c(om))%*%Fii_pr[,,1]%*%(pHat_AA[1:2] + pHat_AA[3:4])
-	pn_aa      <-  ones(c(om))%*%Fii_pr[,,3]%*%(pHat_aa[9:10] + pHat_aa[11:12])
+	pHat_AA    <-  n_AABound/norm(as.matrix(n_AABound), type="1")
+	pHat_aa    <-  n_aaBound/norm(as.matrix(n_aaBound), type="1")
+	pn_AA      <-  ones(c(om)) %*% Fii_pr[,,1] %*% (pHat_AA[1:2] + pHat_AA[3:4])
[…]

### 2020-09-01 fix-Zeta-MWE.R

@@ -1,0 +1,338 @@
+##  Minimum working example of disagreement b/w
+##   Simulations and Zeta_i
+
+
+# run simulations necessary to show disagreement
+# (can take 10 min  or so)
+rm(list=ls())
+source('R/functions-Simulations.R')
+
+	# Obligate Outcrossing
+	# C = 0
+	pars  <-  list(
+					"hf"     =  1/2,
+					"hm"     =  1/2,
+					"C"      =  0,
+					"delta"  =  0
+					)
+
+	selLoop(sMax = 0.15, nSamples=1000,
+			om = 2, g = 3, theta = c(0.6,0.6,0.6,0.6,0.05,0.05,6), theta_prime = c(0.6,0.6,0.6,0.6,0.05,0.05,6), 
+			hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, tlimit = 10^5, returnRes = FALSE)
+
+	# Intermediate Selfing
+	# C = 1/2
+	pars  <-  list(
+					"hf"     =  1/2,
+					"hm"     =  1/2,
+					"C"      =  1/2,
+					"delta"  =  0
+					)
+
+	selLoop(sMax = 0.15, nSamples=1000,
+			om = 2, g = 3, theta = c(0.6,0.6,0.6,0.6,0.05,0.05,6), theta_prime = c(0.6,0.6,0.6,0.6,0.05,0.05,6), 
+			hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, tlimit = 10^5, returnRes = FALSE)
+
+
+	# Obligate Outcrossing
+	# C = 0
+	pars  <-  list(
+					"hf"     =  1/4,
+					"hm"     =  1/4,
+					"C"      =  0,
+					"delta"  =  0
+					)
+
+	selLoop(sMax = 0.15, nSamples=1000,
+			om = 2, g = 3, theta = c(0.6,0.6,0.6,0.6,0.05,0.05,6), theta_prime = c(0.6,0.6,0.6,0.6,0.05,0.05,6), 
+			hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, tlimit = 10^5, returnRes = FALSE)
+
+	# Intermediate Selfing
+	# C = 1/2
+	pars  <-  list(
+					"hf"     =  1/4,
+					"hm"     =  1/4,
+					"C"      =  1/2,
+					"delta"  =  0
+					)
+
+	selLoop(sMax = 0.15, nSamples=1000,
[…]

### 2020-09-01 R/functions-Simulations.R
found typo in zeta_i eq. from notes

@@ -71,17 +71,19 @@ calcZeta  <-  function(om, Fii, Fii_pr, USi, UXi, pHat_AA, pHat_aa, C, delta, la
 	M22_AA  <-  rbind(
 					cbind(USi[,,2] + C*(1 - delta)*Fii[,,2]/2, C*(1 - delta)*Fii[,,2]/2, zeros(c(2,2))),
 					cbind(           ((1 - C)/2)*Fii[,,2] + kronecker(c(((1 - C)/(2*pn_AA)))*Fii[,,1]%*%(pHat_AA[1:2] + pHat_AA[3:4]), (ones(om)%*%Fii_pr[,,2])),
-						  UXi[,,2] + ((1 - C)/2)*Fii[,,2] + kronecker(c(((1 - C)/(2*pn_AA)))*Fii[,,1]%*%(pHat_AA[1:2] + pHat_AA[3:4]), (ones(om)%*%Fii_pr[,,2])), (1 - C)*Fii[,,3]),
+						  UXi[,,2] + ((1 - C)/2)*Fii[,,2] + kronecker(c(((1 - C)/(2*pn_AA)))*Fii[,,1]%*%(pHat_AA[1:2] + pHat_AA[3:4]), (ones(om)%*%Fii_pr[,,2])), 
+						                 (1 - C)*Fii[,,3] + kronecker(c(((1 - C)/(  pn_AA)))*Fii[,,1]%*%(pHat_AA[1:2] + pHat_AA[3:4]), (ones(om)%*%Fii_pr[,,3]))),
 					cbind(((C*(1 - delta))/4)*Fii[,,2], ((C*(1 - delta))/4)*Fii[,,2], USi[,,3] + C*(1 - delta)*Fii[,,3])
 					)
 	M22_aa  <-  rbind(
 					cbind(USi[,,2] + C*(1 - delta)*Fii[,,2]/2, C*(1 - delta)*Fii[,,2]/2, zeros(c(2,2))),
 					cbind(           ((1 - C)/2)*Fii[,,2] + kronecker(c(((1 - C)/(2*pn_aa)))*Fii[,,3]%*%(pHat_aa[9:10] + pHat_aa[11:12]), (ones(om)%*%Fii_pr[,,2])),
[…]

### 2020-09-01 makeFigs.R

@@ -37,8 +37,9 @@ toPdf(FunnelPlots(df1 = 'demSimsSfxSm_MinorEqAlleleInv_sMax0.15_nSamples1000_hf0
 figPath(name='FunnelPlots_MEqAInv_ID.pdf'), width=10, height=7)
 embed_fonts(figPath(name='FunnelPlots_MEqAInv.pdf'))
 
-
-
-
-
-
+toPdf(FunnelEigSimCompare(df1 = 'demSimsSfxSm_MinorEqAlleleInv_sMax0.15_nSamples1000_hf0.5_hm0.5_C0_delta0',
+						  df2 = 'demSimsSfxSm_MinorEqAlleleInv2_sMax0.15_nSamples1000_hf0.5_hm0.5_C0.5_delta0',
+						  df3 = 'demSimsSfxSm_MinorEqAlleleInv_sMax0.15_nSamples1000_hf0.25_hm0.25_C0_delta0',
+						  df4 = 'demSimsSfxSm_MinorEqAlleleInv2_sMax0.15_nSamples1000_hf0.25_hm0.25_C0.5_delta0'), 
+figPath(name='Funnel_Compare_Eig_Sim.pdf'), width=10, height=7)
+embed_fonts(figPath(name='Funnel_Compare_Eig_Sim.pdf'))

### 2020-09-01 run-Simulations.R

@@ -55,12 +55,12 @@ source('R/functions-Simulations.R')
 			hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, tlimit = 10^5, returnRes = FALSE)
 
 	# Intermediate Selfing w/ I.D.
-	# C = 1/2, delta = 0.5
+	# C = 1/2, delta = 0.2
 	pars  <-  list(
 					"hf"     =  1/2,
 					"hm"     =  1/2,
 					"C"      =  1/2,
-					"delta"  =  1/8
+					"delta"  =  1/5
 					)
 
 	selLoop(sMax = 0.15, nSamples=1000,
@@ -105,7 +105,7 @@ source('R/functions-Simulations.R')
 					"hf"     =  1/4,
 					"hm"     =  1/4,
 					"C"      =  1/2,
-					"delta"  =  1/8
+					"delta"  =  1/5
 					)
 
 	selLoop(sMax = 0.15, nSamples=1000,

### 2020-09-01 notes/notes-SA-partSelf-wDemography.tex
fixed zeta typo in notes

@@
 \mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & 0 \\
- \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & (1-C_{aa})\mbf{F}_{aa}\\
+ \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & (1-C_{aa})\mbf{F}_{aa} + \Theta\\
   \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa}  \\ 
@@
 }
-where 
+where $\Theta = \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime$, and
+
 \begin{equation}
@@
 
-\begin{equation} \label{eq:coexist_AA}
+\begin{equation} \label{eq:coexist_aa}
 	\lambda_{aa} < 
@@
 \mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & 0 \\
- \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{aa})\mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{aa})\mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & (1-C_{AA})\mbf{F}_{AA}\\
-  \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{AA} +C_{AA} (1 - \delta) \mbf{F}_{AA}  \\ 
+ \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C_{aa})\mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C_{aa})\mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & (1-C_{AA})\mbf{F}_{AA} + \Phi\\
+  \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{AA} + C_{AA} (1 - \delta) \mbf{F}_{AA}  \\ 
 			\end{array} \right), \\ 
@@
 }
+
+where $\Phi = \mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime$.
 

### 2020-09-03 notes/notes-SA-partSelf-wDemography.tex
updated notes with inv cond

@@
 				\mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{Aa} & \mbf{0} &  \mbf{0}\\\hline
-				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^S_{Aa} & \mbf{0} \\ 
-				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{Aa}
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^S_{aa} & \mbf{0} \\ 
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{aa}
 			\end{array} \right) \\ 
@@
 }
-where $\Theta = \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime$, and
+where $\Theta = \frac{(1 - C_{AA})}{p_{n}} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime$, and
 
@@
 \footnotesize
-\begin{align*} \label{eq:eigAA}
+\begin{align*} \label{eq:eigAAFull}
 	\tilde{\zeta}_{AA} = 
@@
 
-where $\Phi = \mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime$.
+where $\Phi = \frac{(1 - C_{aa})}{p_{n}} \mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime$.
+
+\end{landscape}
+
+\subsection*{Conditions for the maintenance of SA polymorphism}
+
+My PhD supervisor Tim and I have a longstanding joke (originally referencing something Alexei Kondrashov used to say during lectures) for when, seemingly against all odds, your equations simplify to give you a beautiful result.
+We call it "a gift from Jesus", and here Jesus really is being generous!
+\bigskip
+
+OK...
+let's start with a pretty general situation: we will solve for the conditions where SA polymorphism is maintained, defined by where $\tilde{\zeta}_{i} > 1$, where $i \in \{AA,aa\}$ (i.e., where the right hand side of Eq(\ref{eq:eigAAFull}) is greater than $1$).
+The expressions get pretty ugly, so for simplicity we will make two simplying assumptions.
+First, we will assume additive SA fitness effects ($h_f = h_m = 1/2$).
+Second, that at the boundary equilibria, the frequency of the $a$ allele in individuals produced by selfing is equal to the frequency in individuals produced by outrcrossing, i.e., $\hat{\mbf{p}}_{X,i} = \hat{\mbf{p}}_{S,i}$.
+For ease of comparison with previous population genetic models (Jordan \& Connallon 2014; Olito 2017), we solve $\tilde{\zeta}_{i} > 1$ for $s_f$.
+In this case, SA polymorphism will be maintained when
+
+\begin{equation} \label{eq:InvCondSimp}
+	\frac{s_m (1 - C)}{(1 - s_m) (1 + C (1 - 2 \delta)} < s_f < \frac{s_m (1 - C)}{1 + s_m (1 - C ) + C (1 - 2 \delta)}.
+\end{equation}
+
+\noindent Remarkably, these conditions are mathematically identical to those derived for additive SA fitness effects by Jordan \& Connallon (2014) and the 1-locus invasion condition derived in Olito (2017).
+Relaxing the above simplifying assumptions quickly leads to complicated expressions.
+However, we get very similar expressions to the pop.~gen.~models if we relax the assumption of additive fitness effects and allow for arbitrary dominance.
+
+One scenario where we do get interesting behaviour is when we relax the assumption that $\hat{\mbf{p}}_{X,i} = \hat{\mbf{p}}_{S,i}$.
+Again, the expressions are complicated (not worth writing out here), but here is what happens: ($1$) when $\hat{\mbf{p}}_{X,i} > \hat{\mbf{p}}_{S,i}$, the funnel "opens up", so that polymorphism is maintained in a saddle-shaped area straddling the $1:1$ line such that the regions where either the $A$ or $a$ allele fixes shrink into the upper-left and lower-right hand corners of $s_f \times s_m$ parameter space.
+This dramatically increses the proportion of parameter space where polymorphism is maintained.
+($2$) When $\hat{\mbf{p}}_{X,i} < \hat{\mbf{p}}_{S,i}$ the funnel shrinks such that polymorphism is only maintained in regions where $s_f,s_m > 0$, and gets progressively smaller until srong selection is required through both female and male function to maintain polymorphism.
+
+Now, what's interesting (assuming I understand how $\hat{\mbf{p}}_{S,i}$ and $\hat{\mbf{p}}_{X,i}$ are calculated) is that $\hat{\mbf{p}}_{X,i} = \hat{\mbf{p}}_{S,i}$ unless both $C > 0$ and $\delta > 0$, in which case  $\hat{\mbf{p}}_{X,i} > \hat{\mbf{p}}_{S,i}$.
+So, I think this means that when there is selfing and some inbreeding depression, the parameter space where polymorphism is maintained is expanded relative to classic population genetic models.
+\hl{Need to think about this carefully.
+Interested to hear your thoughts Lotte!}
 
@@
 \newpage
-\end{landscape}
 

### 2020-09-29 R/loadData-Compadre.R
New loadData file to access COMPADRE data
This commit provies a loadData file which will load the Rcompadre
package, which provides access to the online demography database
COMPADRE Plant Matrix Database (2020).
Available from: https://www.compadre-db.org

@@ -1,0 +1,80 @@
+################################################################
+#' Load and access demographic data from Compadre database
+#' for use with our evolutionary demographic models
+#'	
+#'  Author: Colin Olito
+#' 
+#' 	COMPADRE Plant Matrix Database (2020). 
+#' 	Available from: https://www.compadre-db.org
+#'  [29 09 2020, Version 6.20.9.0 (release date Sep_10_2020)]
+
+###############
+# DEPENDENCIES
+###############
+
+#' install package 'remotes' if necessary
+#' will already be installed if 'devtools' is installed
+# install.packages("devtools")
+# install.packages("remotes") 
+
+#' install Rcompadre from current github build 
+#' we set build_vignettes = FALSE to avoid a possible error. 
+#' See: https://jonesor.github.io/Rcompadre/index.html
+# remotes::install_github("jonesor/Rcompadre", build_vignettes = FALSE)
+
+# load the package
+library(Rcompadre)
+
+# fetch the COMPADRE database
+compadre <- cdb_fetch("compadre")
+
+
+###########################
+# Mimulus guttatus example
+###########################
+
+# Load data for Mimulus guttatus
+cdb_Mg   <-  cdb_check_species(compadre, "Mimulus guttatus", return_db = TRUE)
+
+# Author info and population names
+cdb_Mg$Authors
+cdb_Mg$MatrixPopulation
+
+# load class info for Eagle river population from Petersen et al. (2016).
+# this is a perennial montane population that was used as the 'local'
+# population in this study. All other populations were grown in a common
+# garden in the field at Eagle River.
+classInfo        <-  matrixClass(cdb_Mg)
+Mg_ER            <-  cdb_Mg[22]
+Mg_ER_classInfo  <-  classInfo[[22]]
+
+# Look at matrices
+Mg_ER$mat
+matList_Mg_ER  <-  list(
+						"A"  =  matA(Mg_ER)[[1]],
+						"U"  =  matU(Mg_ER)[[1]],
+						"F"  =  matF(Mg_ER)[[1]],
+						"C"  =  matC(Mg_ER)[[1]]
+						)
+for(i in 1:4) {
+	colnames(matList_Mg_ER[[i]])  <-  classInfo[[22]][,2]
+}
+matList_Mg_ER
+
+# replace NA's with 0's in U matrix
+# this ensures that all individuals 
+# die after the 2nd season. No immortals allowed!
[…]

### 2020-09-29 doc/SA-Hermaphrodites-wDemography.tex
New MS file with outline
Added manuscript .tex file using AmNat LaTeX template, with skeleton
outline.

\documentclass[11pt]{article}
% Preamble
\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
\usepackage{fullpage}
\usepackage[authoryear,sectionbib,sort]{natbib} %\bibpunct{(}{)}{;}{author-year}{}{,}
\bibliographystyle{amnatnat}
\setlength{\bibsep}{0.0pt}
\linespread{1.7}
\usepackage[utf8]{inputenc}
\usepackage[left]{lineno}
\usepackage{titlesec}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{color,soul}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{pdflscape}
\titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
\titleformat{\subsection}[block]{\Large\itshape\filcenter}{\thesubsection}{1em}{}
\titleformat{\subsubsection}[block]{\large\itshape}{\thesubsubsection}{1em}{}
\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1em}{}[. ]\renewcommand{\refname}{Literature Cited}

\newcommand\encircle[1]{%
  \tikz[baseline=(X.base)] 
    \node (X) [draw, shape=circle, inner sep=0] {\strut #1};}

% Graphics package
\usepackage{graphicx}
\graphicspath{{../output/figs/}.pdf}

% Change default margins
\usepackage[top=0.75in, bottom=0.75in, left=0.75in, right=0.75in]{geometry}

% Definitions
\def\mathbi#1{\textbf{\em #1}}
\def\mbf#1{\mathbf{#1}}
\def\mbb#1{\mathbb{#1}}
\def\mcal#1{\mathcal{#1}}
\newcommand{\bo}[1]{{\bf #1}}
\newcommand{\tr}{{\mbox{\tiny \sf T}}}

\newcommand{\bm}[1]{\mbox{\boldmath $#1$}}

%===========================================
% Basic info from Journal

%%%%%%%%%%%%%%%%
% Line numbering
%%%%%%%%%%%%%%%%

% Please use line numbering with your initial submission and
% subsequent revisions. After acceptance, please turn line numbering
% off by adding percent signs to the lines %\usepackage{lineno} and
% to %\linenumbers{} and %\modulolinenumbers[3] below.
%
% To avoid line numbering being thrown off around math environments,
% the math environments have to be wrapped using
% \begin{linenomath*} and \end{linenomath*}
%
% (Thanks to Vlastimil Krivan for pointing this out to us!)

%%%%%%%%%%%%
% Authorship
%%%%%%%%%%%%
% Please remove authorship information while your paper is under review,
% unless you wish to waive your anonymity under double-blind review. You
% will need to add this information back in to your final files after
% acceptance.

% The journal does not have numbered sections in the main portion of
% articles. Please refrain from using section references (à la
% section~\ref{section:CountingOwlEggs}), and refer to sections by name
% (e.g. section ``Counting Owl Eggs'').

% You may wish to remove the Acknowledgments section while your paper 
% is under review (unless you wish to waive your anonymity under
% double-blind review) if the Acknowledgments reveal your identity.
% If you remove this section, you will need to add it back in to your
% final files after acceptance.

%If you have deposited data to Dryad, you should cite them somewhere in the main text (usually in the Methods or Results sections). A sentence like the following will do. All data are available in the Dryad Digital Repository (\citealt{CookEtAl2015}).

%===========================================

% This version of the LaTeX template was last updated on
% November 8, 2019.

\begin{document}
\title{Evolutionary demography and the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites}
\author{Colin Olito$^{1,\ast}$ \\ 
Charlotte de Vries$^{2}$}
\date{\today}
\maketitle

\noindent{} 1. Department of Biology, Lund University, Lund 223 62, Sweden;

\noindent{} 2.  Department of Evolutionary Biology and Environmental Studies, University of Zurich, Zurich CH-8057, Switzerland;

\noindent{} $\ast$ Corresponding author; e-mail: colin.olito@gmail.com

\bigskip

\textit{Manuscript elements}: Figure~1, figure~2, table~1, online appendices~A and B (including figure~A1 and figure~A2). Figure~2 is to print in color.

\bigskip

\textit{Keywords}: Intralocus sexual conflict, Evolutionary Demography, Balancing selection, Hermaphrodite, mixed mating systems, inbreeding depression

\bigskip

\textit{Manuscript type}: Article. %Or e-article, note, e-note, natural history miscellany, e-natural history miscellany, comment, reply, invited symposium, or historical perspective.

\bigskip

\noindent{\footnotesize Prepared using the suggested \LaTeX{} template for \textit{Am.\ Nat.}}

%\linenumbers{}
%\modulolinenumbers[3]

\newpage{}

%====================
% Begin Main Text
%====================

\section*{Outline}

OK, just putting some ideas out here based roughly on our conversations about an {\itshape AmNat}-style article, the results we have so far, and what I think we should be able to do with data from COMPADRE.

\begin{enumerate}
	\item {\bf Introduction}
	\begin{enumerate}
		\item Briefly introduce sexual antagonism and intra-locus sexual conflict.
		\item Simultaneous hermaphrodites, mixed mating systems, unique predictions for SA polymorphism.
		\item Introduce Evolutionary Demography, pivot to focus on the consequences of taking demography into account explicitly for the maintenance of SA polymorphism.
		\item Introduce key difference between existing population genetic models of SA polymorphism and our models: possibility of diverse population and evolutionary dynamics. 
		\item Explain and highlight "demographically viable" parameter space, indicate that we are interested in whether parameter space predicted to be polymorphic by Pop.~gen.~models is, indeed, "demographically viable"
		\item Building upon the two-sex stage-structured matrix models developed by \citet{deVriesCaswell2019a,deVriesCaswell2019b}, we develop stage-structured matrix models that can accommodate selection through male and female function in simultaneous hermaphrodites with mixed mating systems. The models also allow for the effects of inbreeding depression to manifest in several demographic parameters, allowing us to explore the fitness consequences of early- vs.~late-acting inbreeding depression.
	\end{enumerate}

	\item {\bf Vignette 1: When is polymorphism demographically viable?}
	\begin{enumerate}
		\item Review previous predictions for outcrossers \& partial selfing
		\begin{enumerate}
			\item Greater scope for SA polymorphism in outcrossers vs.~selfers
			\item Greater scope for SA polymorphism when SA fitness effects are recessive (dominance reversal).
		\end{enumerate}
		\item Key results summarized in Fig.~1: 
		\begin{enumerate}
			\item when populations do not have high growth-rates (i.e., when $\lambda \approx 1$), much of of polymorphic parameter space becomes demographically inviable.
			\item Specifically, would-be polymorphic paramter space becomes inviable when segregating male-beneficial/female-deleterious alleles can reduce fecundity enough to drive the population extinct.
			\item This effect is most dramatic for separate-sexed species and predominant outcrossers. Selfing shelters the population from the female-deleterious fitness effects because there is no opportunity for selection via male function. This can also be thought of as reproductive assurance in our model....
		\end{enumerate}
	\end{enumerate}

	\item {\bf Vignette 2: Effects of inbreeding depression}
	\begin{enumerate}
		\item Review effect of I.D.~in population genetic models: causes selfing populations to look more like outcrossing ones (at least in terms of where SA polymorphism is maintained). This is because a fraction of selfed ovules are aborted, meaning the overall ratio of outcrossed vs.~selfed progeny increases.
		\item Reiterate that selfing is often viewed as a mechanism for reproductive assurance.
		\item Key results summarized in Fig.~2
		\begin{enumerate}
			\item Pop.~Gen.~models predict that higher early-acting ID should lead to greater proportion of polymorphic parameter space, while our models predict that it should decrease.
			\item When and how ID affects fitness of offspring produced by selfing can change the location and steepness of the decline in polymorphic parameter space.
			\item A perhaps obvious, but sometimes overlooked implication of our model predictions: selfing can offer reproductive assurance, and SA polymorphisms can be maintained in a non-trivial proportion of parameter space in partially selfing populations... BUT demographically, the population must be able to "afford" the loss of selfed ovules/offspring.
		\end{enumerate}
	\end{enumerate}

	\item {\bf Vignette 3: Examples using real data}
	\begin{enumerate}
		\item We can parameterize our model with real populations' demographic rates, and make explicit predictions regarding the opportunity for SA polymorphism given the species' demography. We provide two illustrative examples:
		\item Opportunities for SA polymorphism in {\itshape Mimulus guttatus} (now {\itshape Erythranthe guttata})
		\item Possibly use data from \citet{PetersonEtAl2016} Eagle Meadows population, which is the 'local' population in their big common garden local adaptation expeirment.
		\begin{enumerate}
			\item Illustrate funnel plot using {\itshape E.~gutatta} demographic rates.
			\item Fun little example: PLOT inversion from \citet{LeeKelly2015}, on that funnel!
		\end{enumerate}

		\item Do sexually dimorphic species exhibit demographic rates that increase opportunities for SA polymorphism?
		\begin{enumerate}
			\item Compare proportion of viable polymorphic parameter space for sexually dimorphic species vs.~close hermaphroditic relative
			\item Possibly {\itshape Silene acaulis} (gynodioecious or dioecious) vs.~hermaphrodite {\itshape Silene} species?
		\end{enumerate}
	\end{enumerate}

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%
\newpage{}
\section*{Abstract}
%%%%%%%%%%%%%%%%%%%%
\ldots

\newpage{}

\section*{Introduction}

\ldots

\section*{Methods}

\ldots

\subsection*{The quickness of the fox}

\ldots

\subsubsection*{Third-order heading}

\ldots

\paragraph*{Fourth-order heading}
\ldots

\section*{Results}

\ldots

\subsection*{The height of the jump}

\ldots

\subsection*{The laziness of the dog}

\ldots

\section*{Discussion}

\ldots

\section*{Conclusion}

\ldots

%%%%%%%%%%%%%%%%%%%%%
% Acknowledgments
%%%%%%%%%%%%%%%%%%%%%

\section*{Acknowledgments}

OEC would like to thank Madlen Wilmes, Gyuri Barab\'{a}s, Flo D\'{e}barre, Vlastimil K\v{r}ivan, and Greg Dwyer for their comments and suggestions on this template.

\newpage{}

\section*{Appendix A: Supplementary Figures}

% In many cases, The American Naturalist allows authors to typeset 
% their own supplementary material in an author-supplied PDF. For author-
% supplied PDFs, please consult the AmNat_supp_template.tex document,
% available from https://www.journals.uchicago.edu/journals/an/instruct 
%
% By contrast, the Appendix instructions below apply to cases in which
% supplementary material is to be typeset by the AmNat editorial staff.
% That notably includes descriptions of methods, tables defining parameters,
% and other material necessary for reproducing the MS's results.
%
% Please reset counters for the appendix (thus normally figure A1, 
% figure A2, table A1, etc.).
%
% In certain cases, it may be appropriate to have a PRINT appendix in
% addition to (or instead of) an online appendix. In this case, please 
% name the print appendix Appendix A, and any subsequent appendixes (if 
% there are any) should be named Online Appendix B, Online Appendix C,
% etc.
%
% Counters for each appendix should match the letter of that appendix.
% For example, tables in Appendix C should be numbered table C1, table C2,
% etc. This applies to tables, equations, and figures.
%
% It's better not to use the \appendix command, because we have some
% formatting peculiarities that \appendix conflicts with.

\renewcommand{\theequation}{A\arabic{equation}}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{equation}{0}  % reset counter 
\setcounter{figure}{0}
\setcounter{table}{0}

\subsection*{Fox--dog encounters through the ages}

The quick red fox jumps over the lazy brown dog. The quick red fox has always jumped over the lazy brown dog. The quick red fox began jumping over the lazy brown dog in the 19th century and has never ceased from so jumping, as we shall see in figure~\ref{Fig:Jumps}. But there can be surprises (figure~\ref{Fig:JumpsOk}).

If the order and location of figures is not otherwise clear, feel free to include explanatory dummy text like this:

[Figure A1 goes here.]

[Figure A2 goes here.]

\subsection*{Further insights}

Tables in the appendices can appear in the appendix text (see table~\ref{Table:Rivers} for an example), unlike appendix figure legends which should be grouped at the end of the document together with the other figure legends.

\begin{table}[h]
\caption{Various rivers, cities, and animals}
\label{Table:Rivers}
\centering
\begin{tabular}{lll}\hline
River        & City        & Animal            \\ \hline
Chicago      & Chicago     & Raccoon           \\
Des Plaines  & Joliet      & Coyote            \\
Illinois     & Peoria      & Cardinal          \\
Kankakee     & Bourbonnais & White-tailed deer \\
Mississippi  & Galena      & Bald eagle        \\ \hline
\end{tabular}
\bigskip{}
\\
{\footnotesize Note: See table~\ref{Table:Founders} below for further table formatting hints.}
\end{table}

Lorem ipsum dolor sit amet, as we have seen in figures~\ref{Fig:Jumps} and \ref{Fig:JumpsOk}.

\newpage{}
\renewcommand{\theequation}{B\arabic{equation}}
% redefine the command that creates the equation number.
\renewcommand{\thetable}{B\arabic{table}}
\setcounter{equation}{0}  % reset counter 
\setcounter{table}{0}

\section*{Appendix B: Additional Methods}

\subsection*{Measuring the height of fox jumps without a meterstick}

Pellentesque ac nibh placerat, luctus lectus non, elementum mauris. 
Morbi odio velit, eleifend ut hendrerit vitae, consequat sit amet 
nulla. Pellentesque porttitor vitae nisl quis tempus. Pellentesque 
habitant morbi tristique senectus et netus et malesuada fames ac 
turpis egestas. Praesent ut nisi odio. Vivamus vel lorem gravida 
odio molestie volutpat condimentum et arcu.

\begin{equation}
{ \frac{1}{N_k-1} \sum \limits_{t=1}^{N_k} (M_{tjk} - \bar{M}_{jk})^2}
\end{equation}

\subsection*{Quantifying the brownness of the dog}

Pellentesque eu nulla odio. Nulla aliquam porta metus, quis malesuada orci faucibus quis. Suspendisse nunc magna, tristique sit amet sollicitudin nec, elementum et lacus. Sed vitae elementum mi. In hac habitasse platea dictumst. Etiam eu tortor elit. Sed ac tortor purus. Aliquam volutpat, odio sit amet posuere pretium, dolor ex interdum ante, sed luctus quam eros ac nulla.

\begin{equation}
{ (\sum \limits_{p=1}^P {n_{sp}})^{-1}\sum \limits_{p=1}^P {n_{sp}Q_{p}}}
\end{equation}

\newpage{}

%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%
% You can either type your references following the examples below, or
% compile your BiBTeX database and paste the contents of your .bbl file
% here. The amnatnat.bst style file should work for this---but please
% let us know if you run into any hitches with it!
%
% If you upload a .bib file with your submission, please upload the .bbl
% file as well; this will be required for typesetting.
%
% The list below includes sample journal articles, book chapters, and
% Dryad references.
\bibliography{Refs2}

\newpage{}

\section*{Tables}
\renewcommand{\thetable}{\arabic{table}}
\setcounter{table}{0}

\begin{table}[h]
\caption{Founders of \textit{The~American Naturalist}}
\label{Table:Founders}
\centering
\begin{tabular}{lll}\hline
Early editor            & Years with the journal \\ \hline
Alpheus S. Packard Jr.  & 1867--1886 \\
Frederick W. Putnam     & 1867--1874 \\ 
Edward S. Morse         & 1867--1871 \\ 
Alpheus Hyatt           & 1867--1871 \\
Edward Drinker Cope$^a$ & 1878--1897 \\
J.~S. Kingsley          & 1887--1896 \\ \hline 
\end{tabular}
\bigskip{}
\\
{\footnotesize Note: Table titles should be short. Further details should go in a `notes' area after the tabular environment, like this. $^a$ Published the first description of \textit{Dimetrodon}.}
\end{table}

\newpage{}

\section*{Figure legends}

\begin{figure}[h!]
%\includegraphics{horn-of-okapi}
\caption{Figure legends can be longer than the titles of tables. However, they should not be excessively long.}
\label{Fig:OkapiHorn}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%
% Videos
%%%%%%%%%%%%%%%%%%%%%
% If you have videos, journal style for them is similar to that for
% figures. You'll want to include a still image (such as a JPEG)
% to give your readers a preview of what the video looks like.

%%%%% Include the text below if you have videos

\renewcommand{\figurename}{Video} 
\setcounter{figure}{0}
% Thanks to Flo Debarre for the pro tip of putting
% \renewcommand{\figurename}{Video} before the Video legend and
% \renewcommand{\figurename}{Figure} after it!

\begin{figure}[h!]
%\includegraphics{VideoScreengrab.jpg}
\caption{Video legends can follow the same principles as figure legends. Counters should be set and reset so that videos and figures are enumerated separately.}
\label{VideoExample}
\end{figure}

\renewcommand{\figurename}{Figure}
\setcounter{figure}{1}

%%%%% Include the above if you have videos

\begin{figure}[h!]
%\includegraphics{elegance}
\caption{In this way, figure legends can be listed at the end of the document, with references that work, even though the graphic itself should be included for final files after acceptance. Instead, upload the relevant figure files separately to Editorial Manager; Editorial Manager should insert them at the end of the PDF automatically.}
\label{Fig:AnotherFigure}
\end{figure}

\subsection*{Online figure legends}

\renewcommand{\thefigure}{A\arabic{figure}}
\setcounter{figure}{0}

\begin{figure}[h!]
%\includegraphics{jumps20m}
\caption{\textit{A}, the quick red fox proceeding to jump 20~m straight into the air over not one, but several lazy dogs. \textit{B}, the quick red fox landing gracefully despite the skepticism of naysayers.}
\label{Fig:Jumps}
\end{figure}

\begin{figure}[h!]
%\includegraphics{jumps20m}
\caption{The quicker the red fox jumps, the likelier it is to land near an okapi. For further details,.}
\label{Fig:JumpsOk}
\end{figure}

\renewcommand{\thefigure}{B\arabic{figure}}
\setcounter{figure}{0}

\end{document}

### 2020-09-29 notes/notes-SA-partSelf-wDemography.tex
Fixed equations in the last section of the notes (where how-you-are-produced does not effect your demographic rates). Fixed equations in matlab files, wrote code to efficiently search for coexistence area by titrating (halving distance to threshold at each step) :) Next step: use same titrate method to determine extinction.

@@
 \usepackage{fullpage}
-\usepackage[authoryear,sectionbib,sort]{natbib}
+\usepackage[authoryear,sectionbib,sort] {natbib}
 \linespread{1.15}
@@
 			\rho\left(\begin{array}{cc}
-\mathbf{U}_{Aa} + \frac{1}{2} (1 - \delta C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2} (1 - C_{AA})\frac{\mbf{F}_{AA}\hat{\mbf{p}}_{AA}\otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime}{\mathbf{1}^\tr_\omega \mathbf{F}_{AA}^\prime \hat{\mathbf{p}}_{AA}	} & (1-C_{aa}) \mbf{F}_{aa} \\
+\mathbf{U}_{Aa} + \frac{1}{2} (1 - \delta C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2} (1 - C_{AA})\frac{\mbf{F}_{AA}\hat{\mbf{p}}_{AA}\otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime}{\mathbf{1}^\tr_\omega \mathbf{F}_{AA}^\prime \hat{\mathbf{p}}_{AA}	} & (1-C_{aa}) \mbf{F}_{aa}+ (1 - C_{AA})\frac{\mbf{F}_{AA}\hat{\mbf{p}}_{AA}\otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime}{\mathbf{1}^\tr_\omega \mathbf{F}_{AA}^\prime \hat{\mathbf{p}}_{AA}	}  \\
 \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa}  \\ 
@@
 			\rho\left(\begin{array}{cc}
-\mathbf{U}_{Aa} + \frac{1}{2} (1 - \delta C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2} (1 - C_{aa})\frac{\mbf{F}_{aa}\hat{\mbf{p}}_{aa}\otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime}{\mathbf{1}^\tr_\omega \mathbf{F}_{aa}^\prime \hat{\mathbf{p}}_{aa}	} & (1-C_{AA}) \mbf{F}_{AA} \\
+\mathbf{U}_{Aa} + \frac{1}{2} (1 - \delta C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2} (1 - C_{aa})\frac{\mbf{F}_{aa}\hat{\mbf{p}}_{aa}\otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime}{\mathbf{1}^\tr_\omega \mathbf{F}_{aa}^\prime \hat{\mathbf{p}}_{aa}	} & (1-C_{AA}) \mbf{F}_{AA}+\ (1 - C_{aa})\frac{\mbf{F}_{aa}\hat{\mbf{p}}_{aa}\otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime}{\mathbf{1}^\tr_\omega \mathbf{F}_{aa}^\prime \hat{\mathbf{p}}_{aa}	} \\
 \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}_{AA} +C_{AA} (1 - \delta) \mbf{F}_{AA}  \\ 
@@
 					\begin{array}{cc|cc}
-						\sigma_{Aa}(1 - \gamma_{Aa}) &  \frac{1}{2} (1 - \delta C_{Aa}) f_{Aa} +\frac{1}{2} (1 - C_{AA})\frac{f^\prime_{Aa}f_{AA}}{f^\prime_{AA}} &0 & (1-C_{aa})f_{aa}\\
+						\sigma_{Aa}(1 - \gamma_{Aa}) &  \frac{1}{2} (1 - \delta C_{Aa}) f_{Aa} +\frac{1}{2} (1 - C_{AA})\frac{f^\prime_{Aa}f_{AA}}{f^\prime_{AA}} &0 & (1-C_{aa})f_{aa}+ (1 - C_{AA})\frac{f^\prime_{aa}f_{AA}}{f^\prime_{AA}}\\
 						\sigma_{Aa} \gamma_{Aa}     & s_{Aa}& 0 & 0\\ \hline

### 2020-09-29 matlab/deVries_Olito_selfing.m

@@ -1,7 +1,8 @@
 %Program to implement Mendelian Matrix Population Models
-%for two-stage model with parameters leading to evolutioary rescue
-% Authors: Hal Caswell and Charlotte de Vries
-% June 2018
+%for two-stage model with selfing
+%Version where how-you-were-produced does not affect your demographic rates
+%Authors: C. de Vries and C. Olito
+% July 2020
 clearvars; 
 %Dimensions, 2 stages, 3 genotypes
 om=2;
@@ -18,31 +19,40 @@ eom=ones(om,1);
 % Parameters for EVOLUTIONARY SUICIDE (Fig 1A and 1C). 
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Nzero=[0.1000         0   99.9000
-    0.1000         0   99.9000];
+%Params from Colins R code: fwdDemModelSim  <-  function(	om = 2, g = 3, theta = c(0.6,0.6,0.05,5.9), theta_prime = c(0.6,0.6,0.05,5.9), 
+%								hf = 1/2, hm = 1/2, sf = 0.1, sm = 0.105, C = 0, delta = 0, tlimit = 10^4, Ainvade = FALSE, intInit = FALSE, ...) {
+
+Nzero=[99.9000 0 0.1000
+    99.9000 0 0.1000];
 %Baseline parameters
 %theta=[s1, s2, g, f]
 
-theta=[0.65 0.7 0.05 3.2]'*ones(1,3);
+theta=[0.6 0.6 0.05 6.1]'*ones(1,3);
 %Selection differentials
 
 % %s1, juvenile survival
-ds1=0.032;
- seldiff=[-ds1+0.01 ds1+0.008 -ds1+0.005];
- theta(1,:)=theta(1,:)+seldiff;
+% ds1=0.032;
+%  seldiff=[-ds1+0.01 ds1+0.008 -ds1+0.005];
+%  theta(1,:)=theta(1,:)+seldiff;
 
 %s2 differential, adult survival
 % s2
-    ds2=0.0001;
-
-seldiff=[0 0 ds2];
-theta(2,:)=theta(2,:)+seldiff;
+%     ds2=0.0001;
+% 
+% seldiff=[0 0 ds2];
+% theta(2,:)=theta(2,:)+seldiff;
 
[…]

### 2020-09-29 matlab/deVries_Olito_selfing_loop_delta.m

@@ -1,0 +1,283 @@
+%Program to implement Mendelian Matrix Population Models
+%for two-stage model with selfing
+%Version where how-you-were-produced does not affect your demographic rates
+%Authors: C. de Vries and C. Olito
+% July 2020
+clearvars; 
+%Dimensions, 2 stages, 3 genotypes
+om=2;
+g=3;
+
+%Identity and ones matrices
+Ig=eye(g);
+Iom=eye(om);
+eg=ones(g,1);
+eom=ones(om,1);
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Parameters for EVOLUTIONARY SUICIDE (Fig 1A and 1C). 
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%Params from Colins R code: fwdDemModelSim  <-  function(	om = 2, g = 3, theta = c(0.6,0.6,0.05,5.9), theta_prime = c(0.6,0.6,0.05,5.9), 
+%								hf = 1/2, hm = 1/2, sf = 0.1, sm = 0.105, C = 0, delta = 0, tlimit = 10^4, Ainvade = FALSE, intInit = FALSE, ...) {
+
+Nzero=[99.9000 0 0.1000
+    99.9000 0 0.1000];
+%Baseline parameters
+%theta=[s1, s2, g, f]
+
+theta=[0.6 0.6 0.05 6.1]'*ones(1,3);
+%Selection differentials
+
+% %s1, juvenile survival
+% ds1=0.032;
+%  seldiff=[-ds1+0.01 ds1+0.008 -ds1+0.005];
+%  theta(1,:)=theta(1,:)+seldiff;
+
+%s2 differential, adult survival
+% s2
+%     ds2=0.0001;
+% 
+% seldiff=[0 0 ds2];
+% theta(2,:)=theta(2,:)+seldiff;
+
+%this is the f differential for the pleiotropy example
+% f
+    hf=.5;
+    hm=.5;
+    sf=0.1;
+    sm=0.105;
+    seldiff=[1 (1 - hf*sf) (1 - sf)];
+theta(4,:)=theta(4,:).*seldiff;
+
+seldiff_prime=[(1 - sm) (1 - hm*sm) 1];
+f_prime=theta(4,:).*seldiff_prime;
[…]

### 2020-09-29 matlab/deVries_Olito_selfing_titrate.m

@@ -1,0 +1,157 @@
+%Program to implement Mendelian Matrix Population Models
+%for two-stage model with selfing
+%Version where how-you-were-produced does not affect your demographic rates
+%Authors: C. de Vries and C. Olito
+% July 2020
+clearvars; 
+%Dimensions, 2 stages, 3 genotypes
+om=2;
+g=3;
+
+%Identity and ones matrices
+Ig=eye(g);
+Iom=eye(om);
+eg=ones(g,1);
+eom=ones(om,1);
+
+%Params from Colins R code: fwdDemModelSim  <-  function(	om = 2, g = 3, theta = c(0.6,0.6,0.05,5.9), theta_prime = c(0.6,0.6,0.05,5.9), 
+%								hf = 1/2, hm = 1/2, sf = 0.1, sm = 0.105, C = 0, delta = 0, tlimit = 10^4, Ainvade = FALSE, intInit = FALSE, ...) {
+
+Nzero=[99.9000 0 0.1000
+    99.9000 0 0.1000];
+%Baseline parameters
+%theta=[s1, s2, g, f]
+
+theta=[0.6 0.6 0.05 6.1]'*ones(1,3);
+sm=0:0.0001:0.15;
+lambda_save_bigA=nan*ones(length(sm),3);
+lambda_save_a=nan*ones(length(sm),3);
+
+%The precision below which the algorithm stops searching for the boundary sm
+%value
+deltathres=0.00001;
+
+%Selection differentials
+%this is the f differential for the pleiotropy example
+% f
+    hf=.25;
+    hm=.25;
+    
+Thres_bigA=nan*ones(length(sm),1);
+Thres_a=nan*ones(length(sm),1);
+
+delta=0;
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Find threshold zeta_AA=1 for each sm
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+for j=1:length(sm)
+
+    if isnan(Thres_bigA(j)) 
+        sf_temp=[0 0.15]; delta_titrate=1;
+                    % first just a  check
+                    [x_bigA0,x_a0,lambda_bigA,lambda_heterozyg,lambda_a]=eigenvalues_Jacobian(theta,hf,hm,0,sm(j),delta);
[…]

### 2020-09-29 matlab/eigenvalues_Jacobian.m

@@ -1,0 +1,41 @@
+function [x_bigA,x_a,lambda_bigA,lambda_heterozyg,lambda_a]=eigenvalues_Jacobian(theta,hf,hm,sf,sm,delta)
+                              
+s1=theta(1,:);
+s2=theta(2,:);
+gam=theta(3,:);
+lambda=zeros(3,1);
+%selfing rate
+C=0*ones(1,3);
+
+    seldiff=[1 (1 - hf*sf) (1 - sf)];
+    f=theta(4,:).*seldiff;
+    seldiff_prime=[(1 - sm) (1 - hm*sm) 1];
+    f_prime=theta(4,:).*seldiff_prime;
+
+    for i=1:3
+        U{i}=[s1(i)*(1-gam(i)) 0 ;
+        s1(i)*gam(i) s2(i)];
+        F{i}=(1-C(i))*[0 f(i);
+        0 0];
+        F_S{i}=C(i)*(1-delta)*[0 f(i);
+        0 0];
+        F_prime{i}=[0  f_prime(i)
+        0 0];
+    %Genotype specific population growth rate
+        lambda(i)=max(eig(F{i}+U{i}+F_S{i}));
+    end % for i
+%Calculating the largest eigenvalue of the Jacobian for both boundaryies
+Jac_bigA=[U{2}+.5.*F{2}+.5.*F_S{2}+.5.*(f_prime(2)/f_prime(1)).*F{1} F{3}+(f_prime(3)/f_prime(1)).*F{1}
+            .25.*F_S{2} U{3}+F_S{3}];
+Jac_a=[U{2}+.5.*F{2}+.5.*F_S{2}+.5*(f_prime(2)/f_prime(3)).*F{3} F{1}+(f_prime(1)/f_prime(3)).*F{3}
+            .25.*F_S{2} U{1}+F_S{1}];    
+%Dominant eigenvalue of the Jacobian at the AA boundary
+x_bigA=max(eig(Jac_bigA))/lambda(1);
+%Dominant eigenvalue of the Jacobian at the aa boundary
+x_a=max(eig(Jac_a))/lambda(3);
+
+lambda_bigA=lambda(1);
+lambda_heterozyg=lambda(2);
+lambda_a=lambda(3);
+
+end

### 2020-09-29 matlab/emat.m

@@ -1,0 +1,10 @@
+function [E] = emat(m,n,i,j)
+%emat(m,n,i,j)
+%matrix mXn, with 1 in i,j entry and zeros elsewhere
+
+E=zeros(m,n);
+E(i,j)=1;
+
+
+end
+

### 2020-09-29 matlab/simulate_dyn.m

@@ -1,0 +1,130 @@
+function [lambda_sim]=simulate_dyn(theta,hf,hm,sf,sm,delta,nzero)
+ %[lambda_sim,pout,nout]=simulate_dyn(theta,hf,hm,sf,sm,delta,nzero)
+                              
+s1=theta(1,:);
+s2=theta(2,:);
+gam=theta(3,:);
+lambda=zeros(3,1);
+%selfing rate
+C=0*ones(1,3);
+g=3;
+om=2;
+Ig=eye(g);
+Iom=eye(om);
+eg=ones(g,1);
+eom=ones(om,1);
+
+    seldiff=[1 (1 - hf*sf) (1 - sf)];
+    f=theta(4,:).*seldiff;
+    seldiff_prime=[(1 - sm) (1 - hm*sm) 1];
+    f_prime=theta(4,:).*seldiff_prime;
+
+    for i=1:3
+        U{i}=[s1(i)*(1-gam(i)) 0 ;
+        s1(i)*gam(i) s2(i)];
+        F{i}=(1-C(i))*[0 f(i);
+        0 0];
+        F_S{i}=C(i)*(1-delta)*[0 f(i);
+        0 0];
+        F_prime{i}=[0  f_prime(i)
+        0 0];
+    %Genotype specific population growth rate
+        lambda(i)=max(eig(F{i}+U{i}+F_S{i}));
+    end % for i
+
+
+d=eye(g);
+c=[0; 1]; %Indicator vector, with a 1 in positions corresponding to breeding stages
+
+bbD=kron(Iom,d); %Equation 27 main text
+bbU=zeros(g*om); 
+bbF=zeros(g*om);
+bbF_S=zeros(g*om);
+bbF_prime=zeros(g*om);
+for i=1:g
+    bbU=bbU+kron(emat(g,g,i,i),U{i}); %Equation 25 main text
+    bbF=bbF+kron(emat(g,g,i,i),F{i}); %Equation 26 main text
+    bbF_prime=bbF_prime+kron(emat(g,g,i,i),F_prime{i});
+    bbF_S=bbF_S+kron(emat(g,g,i,i),F_S{i});
+end % for i
+
+%Matrices
+Ig=eye(g);
+K=vecperm(om,g);
+
+X=kron(Ig,c'); %Equation 18
+
+W=[1 .5 0
+    0 .5 1];
+
[…]

### 2020-09-29 matlab/vec.m

@@ -1,0 +1,3 @@
+function y=vec(x);
+
+y=x(:);

### 2020-09-29 matlab/vecperm.m

@@ -1,0 +1,19 @@
+
+function p = vecperm(m,n)
+% vecperm
+% function to calculate the vec permutation matrix of size m,n
+% 4/9/03
+%function p = vecperm(m,n)
+
+p = zeros(m*n);
+a = zeros(m,n);
+for i = 1:m
+    for j = 1:n
+        e = a;
+        e(i,j) = 1;
+        p = p + kron(e,e');
+    end
+end
+
+
+        

### 2020-10-20 R/functions-Simulations.R
Minor edits to titration funcs, notes
Continued fixing titration function to work properly for 2-stage
model. Now turning attention to the titration function for the
Mimulus example.

@@ -1487,6 +1487,183 @@ extinctThreshTitrate  <-  function(sMax=0.15, res=0.0015,
 
 
 
+
+
+##############################
+#' Identify extinction threshold across sf x sm parameter space
+#' faster method for calculating proportions of sf x sm where
+#' different dynamical outcomes happen(?)
+#'
+extinctThreshTitrateRotate  <-  function(sMax=0.15, res=0.0015, precision=1e-4,
+								   om = 2, g = 3, theta = c(0.6,0.6,0.05,5.8), theta_prime = 6, 
+								   hf = 1/2, hm = 1/2, C = 0, delta = 0, 
+								   delta_j = 0, delta_a = 0, delta_gamma = 0,
+								   tlimit = 10^5, Ainvade=FALSE, eqThreshold = 1e-8, verbose=TRUE, writeFile=TRUE) {
+	
+	# vector of sf values to titrate along
+	sfs          <-  seq(0, sMax, by=res)
+	sms          <-  sfs
+	extinctionThreshold  <-  cbind(rep(NA, times=2*length(sfs)),
+								   rep(NA, times=2*length(sfs)))
+
+	# Three values we need to keep track of
+	titrateVals  <-  matrix(rbind(c(0, 0, NA),
+								  c(0, 0, NA)), nrow=2,ncol=3)
+	rownames(titrateVals)  <-  c("left", "right")
+	colnames(titrateVals)  <-  c("sm", "sf", "lambda")
+
+plot(NA, xlim=c(0,sMax), ylim=c(0,sMax))
+
+	# loop over titration start values for upper-left triangle
+	# of sf x sm space
+	for(i in 1:(length(sfs)-1)) {
+
+		# Step through starting point on upper-left edges of sf x sm space
+		titrateVals[1,c(1:2)]  <-  c(0, sfs[length(sfs)-i])
+		titrateVals[2,c(1:2)]  <-  c(sms[i+1], sMax)
+		titrateVals[,3]        <-  c(NA,NA)
+
+points(	titrateVals[,2] ~ 	titrateVals[,1])
+		# calculate population growth rate at these boundaries
+		L  <-  fwdDemModelSim(om = om, g = g, theta = theta, theta_prime = theta_prime, 
+								 hf = hf, hm = hm, sf = titrateVals[1,2], sm = titrateVals[1,1], C = C, delta = delta, 
[…]

### 2020-10-23 doc/SA-Hermaphrodites-wDemography.tex
Minor edits to titration funcs, notes
Continued fixing titration function to work properly for 2-stage
model. Now turning attention to the titration function for the
Mimulus example.
Added Mimulus inv6 fig, titration func.
This commit adds a working titration function for the Mimulus
example from Compadre, and also includes a preliminary figure
showing where inv6 falls in sm x sf parameter space for two
different selfing rates (chosen from reported selfing rates)
for this species.

@@
 	\begin{enumerate}
-		\item Briefly introduce sexual antagonism and intra-locus sexual conflict.
-		\item Simultaneous hermaphrodites, mixed mating systems, unique predictions for SA polymorphism.
-		\item Introduce Evolutionary Demography, pivot to focus on the consequences of taking demography into account explicitly for the maintenance of SA polymorphism.
-		\item Introduce key difference between existing population genetic models of SA polymorphism and our models: possibility of diverse population and evolutionary dynamics. 
-		\item Explain and highlight "demographically viable" parameter space, indicate that we are interested in whether parameter space predicted to be polymorphic by Pop.~gen.~models is, indeed, "demographically viable"
+		\item Briefly introduce sexual antagonism and intra-locus sexual conflict. (Colin)
+		\item Simultaneous hermaphrodites, mixed mating systems, unique predictions for SA polymorphism. (Colin)
+		\item Introduce Evolutionary Demography, pivot to focus on the consequences of taking demography into account explicitly for the maintenance of SA polymorphism. (Lotte)
+		\item Introduce key difference between existing population genetic models of SA polymorphism and our models: possibility of diverse population and evolutionary dynamics. (Lotte then Colin)
+		\item Explain and highlight "demographically viable" parameter space, indicate that we are interested in whether parameter space predicted to be polymorphic by Pop.~gen.~models is, indeed, "demographically viable" (Lotte then Colin)
+		\begin{itemize}
+			\item Pop.
+Gen.
+models tacitly assume stable population size, parameterize genotypic fitnesses relative to the 'best performing' genotype.
+Non-overlapping generations.
+Casually define fitness as lifetime reproductive fitness of a given genotype.
+			\item "In practice, these genotypic fitnesses emerge from myriad process acting throughout the life-history of individuals"
+			\item these processes contribute to key demographic rates such as the intrinsic population growth rate, 
+			\item 
+		\end{itemize}
 		\item Building upon the two-sex stage-structured matrix models developed by \citet{deVriesCaswell2019a,deVriesCaswell2019b}, we develop stage-structured matrix models that can accommodate selection through male and female function in simultaneous hermaphrodites with mixed mating systems.
 The models also allow for the effects of inbreeding depression to manifest in several demographic parameters, allowing us to explore the fitness consequences of early- vs.~late-acting inbreeding depression.
+(Colin then Lotte)
 	\end{enumerate}
 
-	\item {\bf Vignette 1: When is polymorphism demographically viable?}
+	\item {\bf Methods}
+		\begin{enumerate}
+			\item Citation heavy review of de Vries \& Caswell AmNat (2018) \& de Vries \& Caswell AmNat (2019) TheorPopBiol
+			\item Briefly describe extension to simultaneous hermaphrodites, citing Jordan \& Connallon (2014), Abbot \& Tazzyman (2014), Olito (2017). 
+			\item breakdown into individuals produced by selfing vs.~outcrossing.
+			\item Brief anatomy of $\mathbf{\tilde{A}}$.
+			\item Analysis. Invasion analysis \& abridged derivation of invasion conditions \& identification of extinction threholds.
+			\item Cite Appendix with full derivation distilled from notes file.
+		\end{enumerate}
+
+	\item {\bf When is polymorphism demographically viable?}
 	\begin{enumerate}
@@
 			\item Greater scope for SA polymorphism when SA fitness effects are recessive (dominance reversal).
+			\item but our model allows us to see where/when invasion of SA alleles can drive the population extinct
+			\item \textbf{Fig.1 - illustration of funnel plots w/ extinction isoclines}
 		\end{enumerate}
-		\item Key results summarized in Fig.~1: 
+
+		\item Key results summarized in \textbf{Fig.~2}: 
 		\begin{enumerate}
-			\item when populations do not have high growth-rates (i.e., when $\lambda \approx 1$), much of of polymorphic parameter space becomes demographically inviable.
+			\item For simplicity, temporarily assume no inbreeding depression ($\delta = 0$).
+			\item When populations do not have high growth-rates (i.e., when $\lambda \approx 1$), much of of polymorphic parameter space becomes demographically inviable.
 			\item Specifically, would-be polymorphic paramter space becomes inviable when segregating male-beneficial/female-deleterious alleles can reduce fecundity enough to drive the population extinct.
@@
 This can also be thought of as reproductive assurance in our model....
+			\item Convergence to pop.~gen.~predictions when $f$ is big.
 		\end{enumerate}
@@
 
-	\item {\bf Vignette 2: Effects of inbreeding depression}
+	\item {\bf Effects of inbreeding depression}
 	\begin{enumerate}
@@
 This is because a fraction of selfed ovules are aborted, meaning the overall ratio of outcrossed vs.~selfed progeny increases.
-		\item Reiterate that selfing is often viewed as a mechanism for reproductive assurance.
-		\item Key results summarized in Fig.~2
+		\item Matrix model allows us to explore effects of I.D.
+in different parts of the life-cycle (early- vs.~late-acting I.D.).
+Point out that previous pop.~gen.~models only address early-acting I.D.
+via ovule viability.
+		\item MAKE CLEAR that purging of deleterious recessives thought to cause I.D. is already included in the model results presented in Fig.~3
+		\item Key results summarized in Fig.~3
 		\begin{enumerate}
-			\item Pop.~Gen.~models predict that higher early-acting ID should lead to greater proportion of polymorphic parameter space, while our models predict that it should decrease.
-			\item When and how ID affects fitness of offspring produced by selfing can change the location and steepness of the decline in polymorphic parameter space.
+			\item Pop.~Gen.~models predict that higher early-acting ID should lead to greater proportion of polymorphic parameter space, while our models predict that we get a critical threshold of selfing at which population viability plummets.
+			\item When and how I.D. influences fitness during the life-cyclecan change the location and steepness of the decline in polymorphic parameter space.
+			\item Critically, largest deviations from pop.~gen.~predictions happen for predominantly selfing populations.
 			\item A perhaps obvious, but sometimes overlooked implication of our model predictions: selfing can offer reproductive assurance, and SA polymorphisms can be maintained in a non-trivial proportion of parameter space in partially selfing populations...
 BUT demographically, the population must be able to "afford" the loss of selfed ovules/offspring.
+			\item HIGHLIGHT that this result holds even though we incorporated decrease of the mutation load due to inbreeding (and corresponding decrease in I.D.)...	
 		\end{enumerate}
@@
 
-	\item {\bf Vignette 3: Examples using real data}
+	\item {\bf Mimulus as a potential real world example}
 	\begin{enumerate}
@@
 			\item Illustrate funnel plot using {\itshape E.~gutatta} demographic rates.
-			\item Fun little example: PLOT inversion from \citet{LeeKelly2015}, on that funnel!
+			\item \textbf{Fig.~4}: PLOT inversion from \citet{LeeKelly2015}, on that funnel!
 		\end{enumerate}
 
+	\end{enumerate}
+
+	\item Discussion
+		\begin{enumerate}
+			\item Classic prediction from demography that most species should have a $\lambda \approx 1$
+			\item If true, the demographic impacts on the opportunities for the maintenance of sexually antagonistic polymorphisms may be pervasive...
+			\item Condition dependence, fecundity, scope for sexual confict/antagonistic polymorphism?
 			\item Do sexually dimorphic species exhibit demographic rates that increase opportunities for SA polymorphism?
-		\begin{enumerate}
-			\item Compare proportion of viable polymorphic parameter space for sexually dimorphic species vs.~close hermaphroditic relative
-			\item Possibly {\itshape Silene acaulis} (gynodioecious or dioecious) vs.~hermaphrodite {\itshape Silene} species?
-		\end{enumerate}
+
 		\end{enumerate}

### 2020-10-23 R/functions-Figs.R

@@ -1176,3 +1176,93 @@ deltaSelfingLoadPolySpaceFig  <-  function(df1, df2) {
         proportionalLabel(0.5, -0.35, expression(paste("Selfing Rate (", italic(C), ")")), cex=1.2, adj=c(0.5, 0.5), xpd=NA, srt=0)
 
 }
+
+
+
+
+
+
+
+
+
+# Simple inv6 Fig
+MimulusInv6Fig  <-  function() {
+
+    sfs  <-  seq(0,1, by=0.001)
+    sms  <-  seq(0,1, by=0.001)
+
+    hf  <-  1/2
+    hm  <-  0.35
+
+    sfInv6  <-  0.308
+    smInv6  <-  0.88
+
+    inv_A  <-  popGen_A_invade(hf=hf, hm=hm, sm=sms, C=0)
+    inv_a  <-  popGen_a_invade(hf=hf, hm=hm, sm=sms, C=0)
+    inv_a[inv_a > 1]  <-  1
+
+# Color scheme
+    COLS  <-  list(
+                    "PG"     =  transparentColor('#252525', opacity=1),
+                    "fill"   =  transparentColor('#252525', opacity=0.25),
+                    "inv6"   =  transparentColor('tomato', opacity=1),
+                    "inv6bg" =  transparentColor('tomato', opacity=0.6)
+                    )
+# Set plot layout
+    layout.mat  <- matrix(c(1:2), nrow=1, ncol=2, byrow=TRUE)
+    layout      <- layout(layout.mat,respect=TRUE)
+
+    ## Panel A: Outcrossing
+
+        # Make the plot
+        par(omi=rep(0.5, 4), mar = c(3,3,0.5,0.5), bty='o', xaxt='s', yaxt='s')
+        plot(NA, axes=FALSE, type='n', main='',xlim = c(0,1.05), ylim = c(0,1.05), ylab='', xlab='', cex.lab=1.2)
+        usr  <-  par('usr')
+        rect(usr[1], usr[3], usr[2], usr[4], col='white', border=NA)
+        plotGrid(lineCol='grey80')
+        box()
+        # Simulation Results
+        polygon(c(rev(sms),sms), c(rev(inv_A), inv_a), col=transparentColor('grey80', 0.6), border=NA)
+        lines(inv_A ~ sms, lwd=2, col=COLS$PG)
[…]

### 2020-10-23 R/functions-Simulations.R

@@ -1889,6 +1889,7 @@ fwdSimMimulusDat  <-  function(datMat, theta.list, delta.list, useCompadre = TRU
 					"tlimit"       =  tlimit,
 					"Ainvade"      =  Ainvade,
 					"intInit"      =  intInit,
+					"lambda_sim"   =  invadeEq$lambda_sim,
 					"extinct"      =  invadeEq$lambda_sim < 1,
 					"polymorphism" =  !any(round(invadeEq$p_genotypes[,ncol(invadeEq$p_genotypes)], digits=3) == 1)
 					)
@@ -2010,150 +2011,179 @@ selLoopDatMat  <-  function(sMax = 0.15, nSamples = 1e+2,
 #' faster method for calculating proportions of sf x sm where
 #' different dynamical outcomes happen(?)
 #'
-extinctThreshTitrateMimulus  <-  function(sMax = 0.15, res=0.001, 
-										  datMat, theta.list, delta.list, useCompadre = FALSE,
-										  hf = 1/2, hm = 1/2, C = 0, 
-										  tlimit = 10^5, intInit = TRUE, Ainvade=FALSE, eqThreshold=1e-9, 
-										  writeFile=TRUE, verbose=TRUE) {
+extinctThreshMimulus  <-  function(sMax = 0.15, res=0.01, precision=1e-4, 
+								   datMat, theta.list, delta.list, useCompadre = FALSE,
+								   hf = 1/2, hm = 1/2, C = 0, 
+								   tlimit = 10^5, intInit = TRUE, Ainvade=FALSE, eqThreshold=1e-9, 
+								   writeFile=FALSE, verbose=TRUE) {
+	
 	
 	# vector of sf values to titrate along
-	sfs          <-  seq(res,(sMax-res),by=res)
+	sfs          <-  seq(0, sMax, by=res)
 	sms          <-  sfs
-	smThreshold  <-  rep(NA, times=length(sfs))
-	sfThreshold  <-  smThreshold
+	extinctionThreshold  <-  cbind(rep(NA, times=2*length(sfs)),
+								   rep(NA, times=2*length(sfs)))
 
 	# Three values we need to keep track of
-	titrateStartVals  <-  matrix(rbind(c(sfs[1], sfs[length(sfs)]),
-								  c(NA,NA)),nrow=2,ncol=2)
-	rownames(titrateStartVals)  <-  c("sVals", "extinct")
-	colnames(titrateStartVals)  <-  c("left", "right")
[…]

### 2020-10-23 makeFigs.R

@@ -96,3 +96,13 @@ toPdf(deltaSelfingLoadPolySpaceFig(
 			figPath(name='deltaSelfingPolySpace.pdf'), width=5, height=7)
 embed_fonts(figPath(name='deltaSelfingPolySpace.pdf'))
 
+
+
+
+
+#'  Polymorphic parameter space as a function of the selfing rate
+#'  using heuristic model of inbreeding depression ~ selfing
+toPdf(MimulusInv6Fig(),
+			figPath(name='mimulusInv6Fig.pdf'), width=7, height=5)
+embed_fonts(figPath(name='mimulusInv6Fig.pdf'))
+

### 2020-10-23 run-Simulations.R

@@ -15,7 +15,7 @@
 ##  Dependencies
 rm(list=ls())
 source('R/functions-Simulations.R')
-#source('R/loadData-Compadre.R')
+source('R/loadData-Compadre.R')
 
 
 
@@ -566,3 +566,52 @@ deltaSelfingPolySpaceMakeData(sMax = 0.15, nSamples=1e+3,
 							delta_j = pars$dj, delta_a = pars$da, delta_gamma = pars$dg, 
 							tlimit = 10^5, eqThreshold = 1e-8)
 
+
+
+
+
+
+
+#################################################
+#' Find extinction threshold for Mimulus example
+
+datMat  <-  matList_Mg_EM
+theta.list  <-  list(D = 0.534, 
+					 G = 0.469,
+					 F = 0.64,
+					 O = 614,
+					 A = 6.7e-4,
+					 S = 0.179,
+					 R = 8.71)
+delta.list  <-  list(delta_D = 0,
+					 delta_G = 0,
+					 delta_F = 0,
+					 delta_O = 0,
+					 delta_S = 0
+					 )
+
+extinctThreshMimulus(sMax = 0.99, res=0.01, precision=1e-4, 
+					 datMat=datMat, theta.list = theta.list, delta.list = delta.list, useCompadre = FALSE,
+					 hf = 1/2, hm = 0.35, C = 0, 
+					 tlimit = 10^5, intInit = TRUE, Ainvade=FALSE, eqThreshold=1e-9, 
+					 writeFile=TRUE, verbose=TRUE) 
+
+extinctThreshMimulus(sMax = 0.99, res=0.01, precision=1e-4, 
+					 datMat=datMat, theta.list = theta.list, delta.list = delta.list, useCompadre = FALSE,
+					 hf = 1/2, hm = 0.35, C = 0.24, 
+					 tlimit = 10^5, intInit = TRUE, Ainvade=FALSE, eqThreshold=1e-9, 
+					 writeFile=TRUE, verbose=TRUE) 
+
+# delta values calculated from Willis (1993)
+delta.list  <-  list(delta_D = 0,
+					 delta_G = 0.085,
+					 delta_F = 0.2,
+					 delta_O = 0,
+					 delta_S = 0.38
+					 )
+
+extinctThreshMimulus(sMax = 0.99, res=0.01, precision=1e-4, 
[…]

### 2020-11-12 R/functions-Figs.R
New makeData functions using titration
This commit adds new functionality. Edited and now stably working
(*fingers crossed*) functions for titrating to find extinction
thresholds. Added new wrapper functions to makeData for Fig.2
and Fig.3, which will be quantifying parameter with different
dynamical outcomes based on the invasion conditions and extinction
thresholds identified by titration

@@ -69,6 +69,31 @@ proportionalLabel <- function(px, py, lab, adj=c(0, 1), text=TRUE, log=FALSE, ..
     }
 }
 
+
+
+proportionalArrows <- function(px1, py1, px2, py2, adj=c(0, 1), log=FALSE, length=length, ...) {
+    usr  <-  par('usr')
+    x.p1  <-  usr[1] + px1*(usr[2] - usr[1])
+    y.p1  <-  usr[3] + py1*(usr[4] - usr[3])
+    x.p2  <-  usr[1] + px2*(usr[2] - usr[1])
+    y.p2  <-  usr[3] + py2*(usr[4] - usr[3])
+    if(log=='x') {
+        x.p1  <-  10^(x.p1)
+        x.p2  <-  10^(x.p2)
+    }
+    if(log=='y') {
+        y.p1  <-  10^(y.p1)
+        y.p2  <-  10^(y.p2)
+    }
+    if(log=='xy') {
+        x.p1  <-  10^(x.p1)
+        y.p1  <-  10^(y.p1)
+        x.p2  <-  10^(x.p2)
+        y.p2  <-  10^(y.p2)
+    }
+    arrows(x0=x.p1, y0=y.p1, x1=x.p2, y1=y.p2, length=length,...)
+}
+
 #' Draw equally-spaced white lines on plot window.
 #'
 #' @title Equally-spaced white lines on plot window
@@ -128,11 +153,312 @@ fibonacci.scale  <-  function(n) {
     (fibs/max(fibs))[-2]
 }
 
+##############################################################
+##############################################################
+##  Final Figs for Paper
+##############################################################
+##############################################################
+
+demViablePolySpaceFig  <-  function() {
+
+    # Import data sets
+    invA       <-  read.csv("./output/simData/invasionBoundaries_sMax0.15_res0.0015_hf0.5_hm0.5_C0_delta0_dj0_da0_dg0_f5.8.csv", head=TRUE)
[…]

### 2020-11-12 R/functions-Simulations.R

@@ -174,9 +174,6 @@ predDelta  <-  function(dStar, b=1/2, a=0.2, C) {
 #	zeta_i  <-  zeta_i/lambda_i[c(1,3)]
 #	return(zeta_i)
 #}
-
-
-
 calcZeta  <-  function(om, FSi, FXi, FXi_pr, USi, UXi, pHat_AA, pHat_aa, C, delta, lambda_i){
 
 	pn_AA   <-  ones(c(om)) %*% FXi_pr[,,1] %*% (pHat_AA[1:om] + pHat_AA[om+1:om])
@@ -1319,10 +1316,24 @@ deltaSelfingPolySpaceMakeData  <-  function(sMax = 0.15, nSamples=1e+2,
 
 }
 
-
+pickInvadingAllele  <-  function(hf, hm, sf, sm, C) {
+		if(hf == hm && hf == 1/2) {
+			qHat  <-  popGen_qHat_Add(sf = sf, sm = sm, C = C)
+		}
+		if(hf == hm && hf < 1/2) {
+			qHat  <-  popGen_qHat_DomRev(h = hf, sf = sf, sm = sm, C = C)
+		}
+			if(qHat < 1/2){
+				Ainvade  <-  TRUE
+			}
+			if(qHat > 1/2){
+				Ainvade  <-  FALSE
+			}
+	Ainvade
+}
 
 midPoint  <-  function(x1,x2) {
-	x1 + (x2-x1)/2
+	x1 + (x2-x1)/2 + runif(1, min=-1, max=1)*(x2-x1)/10
 }
 
 
@@ -1331,11 +1342,12 @@ midPoint  <-  function(x1,x2) {
 #' faster method for calculating proportions of sf x sm where
 #' different dynamical outcomes happen(?)
 #'
-extinctThreshTitrate  <-  function(sMax=0.15, res=0.0015, precision = 1e-4,
+extinctThreshTitrate  <-  function(sMax=0.15, res=0.015, precision = 1e-4,
 								   om = 2, g = 3, theta = c(0.6,0.6,0.05,6), theta_prime = 6, 
 								   hf = 1/2, hm = 1/2, C = 0, delta = 0, 
 								   delta_j = 0, delta_a = 0, delta_gamma = 0,
-								   tlimit = 10^5, Ainvade=FALSE, eqThreshold = 1e-8, verbose=TRUE, writeFile=TRUE) {
+								   tlimit = 10^5, Ainvade=FALSE, eqThreshold = 1e-8, 
[…]

### 2020-11-12 makeFigs.R

@@ -19,6 +19,15 @@ rm(list=ls())
 ## Dependencies
 source('R/functions-Figs.R')
 
+
+#################
+# Figs for Paper
+#################
+
+toPdf(demViablePolySpaceFig(), 
+			figPath(name='extinctionThresholdsFig.pdf'), width=10, height=5)
+embed_fonts(figPath(name='extinctionThresholdsFig.pdf'))
+
 ###############
 # PRELIM FIGS
 ###############

### 2020-11-12 run-Simulations.R

@@ -22,6 +22,9 @@ source('R/loadData-Compadre.R')
 # Simulations for final figs
 #############################
 
+##########################################################
+# Fig. 1 - illustrate inv. conditions & ext. threshold
+
 ## Additive SA Fitness
 ## hf = hm = 1/2
 
@@ -37,11 +40,29 @@ source('R/loadData-Compadre.R')
 					"dg"     =  0
 					)
 
-titrateInvBoundaries(sMax=0.15, res=0.0015, precision=1e-4,
-						  om = 2, g = 3, theta = c(0.6,0.6,0.05,5.8), theta_prime = 6, 
-						  hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, 
-						  delta_j = pars$dj, delta_a = pars$da, delta_gamma = pars$dg,
-						  tlimit = 10^2, eqThreshold = 1e-8, verbose=TRUE, writeFile=TRUE)
+#titrateInvBoundaries(sMax=0.15, res=0.0015, precision=1e-4,
+#					 om = 2, g = 3, theta = c(0.6,0.6,0.05,5.8), theta_prime = 5.8, 
+#					 hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, 
+#					 delta_j = pars$dj, delta_a = pars$da, delta_gamma = pars$dg,
+#					 tlimit = 10^3, eqThreshold = 1e-8, verbose=TRUE, writeFile=TRUE)
+
+extinctThreshTitrate(sMax=0.15, res=0.0015, precision=1e-4,
+					 om = 2, g = 3, theta = c(0.6,0.6,0.05,5.8), theta_prime = 5.8, 
+					 hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, 
+					 delta_j = pars$dj, delta_a = pars$da, delta_gamma = pars$dg,
+					 tlimit = 10^5, eqThreshold = 1e-8, verbose=TRUE, writeFile=TRUE)
+extinctThreshTitrate(sMax=0.15, res=0.0015, precision=1e-4,
+					 om = 2, g = 3, theta = c(0.6,0.6,0.05,6), theta_prime = 6, 
+					 hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, 
[…]

### 2020-11-12 doc/SA-Hermaphrodites-wDemography.tex
Expanded 1st draft of introduction
Fleshed out the first few paragraphs of the introduction, and
added a couple short lead-in paragraphs at the end.

@@
 		\item Explain and highlight "demographically viable" parameter space, indicate that we are interested in whether parameter space predicted to be polymorphic by Pop.~gen.~models is, indeed, "demographically viable" (Lotte then Colin)
-		\begin{itemize}
+		\begin{enumerate}
 			\item Pop.
@@
 			\item 
-		\end{itemize}
+		\end{enumerate}
 		\item Building upon the two-sex stage-structured matrix models developed by \citet{deVriesCaswell2019a,deVriesCaswell2019b}, we develop stage-structured matrix models that can accommodate selection through male and female function in simultaneous hermaphrodites with mixed mating systems.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Introduction}
+%%%%%%%%%%%%%%%%%%%%%%%%
 
+Whether a population persists or goes extinct depends in large part on its ability to adapt to the environment.
+Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, including deleterious mutations \citep{Haldane1957}, environmental shifts causing a population to chase a new phenotypic optimum \citep{Maynard-Smith1976, LandeShannon1996,OrrUnckless2008}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, as well as genetic constraints arising from intrinsic features of the inheritance system \citep{ConnallonHall2018}.
+
+Conflicting selection and gene-flow between different classes of individuals in a population, or genetic trade-offs between fitness components, can impose genetic constraints on adaptation with particularly interesting evolutionary consequences \citep{CharlesworthHughes2000, ConnallonHall2018}.
+Such constraints hinder adaptation by preventing individuals (or classes of individuals) from reaching the phenotypic optimum in one or more fitness components, potentially increasing the population's overall risk of extinction.
+At the same time, they provide an effective mechanism for the maintenance of genetic variation, which in turn influences the population's capactiy for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
+Hence, for genetically constrained traits the nature and extent of genetic variation observed in natural populations should reflect the balance between the contribution of constraints to the maintainance of genetic polymorphisms, and the demographic consequences of the resulting maladaptation.
+%Genetic trade-offs between fitness components due to pleiotropy represent a form of genetic constraint with particularly interesting evolutionary consequences because they act as a double-edged sword
+%The nature and extent of genetic variation observed in natural populations should therefore reflect the balance between the contribution of those constraints to the maintainance of genetic polymorphism and the demographic consequences of the resulting maladaptation.
+
+Genetic trade-offs between male and female fitness are a common feature of sexually reproducing populations, and are thought to contribute not only to the maintenance of genetic variation, but also the evolution of sexual dimorphism and gender-related traits \citep{Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, BondurianskyChenoweth2009,Olito2019}.
+When such genetic trade-offs occur, "sexually antagonistic" selection (abbreviated SA hereafter) can arise when beneficial alleles for one sex are deleterious when expressed in the other \citep{Kidwell1977, Rice1992, ConnallonClark2012}.
+Moreover, alleles with opposing fitness effects through male and female sex functions create analogous genetic constraints on fitness in hermaphrodite populations, where both maternal and paternal reproductive success contribute jointly to each individuals' overall fitness \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
+
+The evolutionary dynamics of SA alleles in hermaphrodites differ from those in dioecious populations because individuals may reproduce through a combination of self- and outcross fertilization \citep{Goodwillie2005, Igic2006, JarneAuld2006}.
+Such mixed-mating systems have important theoretical consequences for the maintenance of SA polymorphisms in hermaphrodite populations.
+Self-fertilization is predicted to reduce the total parameter space where balanced SA polymorphisms can be maintained, while simultaneously creating a bias in selection through the female sex function (\citealt{JordanConnallon2014}; but see \citealt{Tazzyman2015}.
+Yet, other processes, such as genetic linkage to other SA loci or a sex-determining region \citep{Otto2011, JordanCharlesworth2012, Olito2017, Olito2019}, or spatial heterogeneity and the complexity of the life-cycle \citep{Olito-etal-2018,ConnallonSharmaOlito2019} can expand the parameter space for SA polymorphism.
+Overall, both theoretical predictions and current empirical data suggest there is ample scope for SA trade-offs and the maintenance of SA polymorphisms in both dioecious and hermaphrodite populations \citep{Abbott2011, WangBarrett2020}, although identifying specific SA loci from genome sequence data remains challenging \citep{RuzickaESEB2020}.
+\bigskip
+
+\begin{enumerate}
+	\item Introduce Evolutionary Demography, pivot to focus on the consequences of taking demography into account explicitly for the maintenance of SA polymorphism. [Lotte]
+	\item Introduce key difference between existing population genetic models of SA polymorphism and our models: possibility of diverse population and evolutionary dynamics. [Lotte then Colin]
+	\item Explain and highlight "demographically viable" parameter space, indicate that we are interested in whether parameter space predicted to be polymorphic by Pop.~gen.~models is, indeed, "demographically viable" [Lotte then Colin]
+		\begin{enumerate}
+			\item Pop.
+Gen.
+models tacitly assume large and/or stable population size, parameterize genotypic fitnesses relative to the 'best performing' genotype.
+Non-overlapping generations.
+Casually define fitness as lifetime reproductive fitness of a given genotype.
+			\item "In practice, these genotypic fitnesses emerge from myriad process acting throughout the life-history of individuals"
+			\item these processes contribute to key demographic rates such as the intrinsic population growth rate, 
+		\end{enumerate}
+\end{enumerate}
+\bigskip
+
+In this article, we build upon the models of \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b} to develop a stage-structured mendelian matrix model that accommodates selection through male and female function in simultaneous hermaphrodites with mixed mating systems.
+Using the model, we jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecy) and partially selfing hermaphrodite populations.
+We focus our analyses on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population is also demographically viable (i.e., has a positive intrinsic growth rate).
+The model also allows for the effects of inbreeding depression to manifest in different life-history stages, enabling us to explore the fitness consequences of early- vs.~late-acting inbreeding depression on the model predictions.
+
+A key finding from our model is that when the intrinsic population growth rate is not large (i.e., $\lambda$ is close to $1$), the deleterious effects on female fecundity of segregating male-beneficial SA alleles can render much of the parameter space for SA polymorphism demographically inviable.
+Under some conditions, self-fertilization can alleviate the demographic costs of balanced SA polymorphisms, but the concommitant costs of inbreeding depression greatly exacerbate them.
+Overall, our findings suggest that previous population genetic theory assuming large and/or constant population sizes probably represents a 'best case scenario' for the maintenance of SA polymorphisms, and provide a more nuanced picture of the nature of SA genetic variation that we should expect to find in natural populations.
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Methods}
+%%%%%%%%%%%%%%%%%%%%%%%%
 \ldots
 
-\section*{Methods}
-
-\ldots
-
+%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{The quickness of the fox}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Results}
-
+%%%%%%%%%%%%%%%%%%%%%%%%
 \ldots
 
+ \begin{figure}[htbp]
+ \centering
+ \includegraphics[width=\linewidth]{./extinctionThresholdsFig}
+ \caption{Illustration of parameter space for SA polymorphism and extinction thresholds predicted by the mendelian stage-structured matrix model.
+Balanced SA polymorphisms can be maintained in the funnel-shaped region between the invasion conditions for each SA allele (dark solid lines).
+However, for some parameter conditions, populations will ultimately go extinct (red shaded regions) due to reduced female fitness resulting from the male-beneficial/female-deleterious allele that is either segregating as a balanced polymorphism (inside the funnel), or becomes fixed (area below the funnel).
+We are particularly interested in identifying and quantifying "demographically viable polymorphic parameter space", which corresponds to the area inside the funnel that is also to the left of the extinction threshold for a given fertility value.
+Results are shown for three different population selfing rates ($C = \{0,\,1/4,\,1/2\}$), and two dominance scenarios (additivity with $h = 1/2$, and dominance reversal with $h = 1/4$); extinction thresholds are illustrated for three different values of female fecundity ($f$ values annotated on each panel).
+}
+ \label{fig:extinctionThresholds}
+ \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{The height of the jump}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection*{The laziness of the dog}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Discussion}
+%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Conclusion}
+%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2020-11-13 doc/SA-Hermaphrodites-wDemography.tex
Began drafting methods section
Self explanatory commit.

@@
 \usepackage{pdflscape}
+\usepackage[colorlinks=true, allcolors=black]{hyperref}
 \titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
@@
 \newcommand{\tr}{{\mbox{\tiny \sf T}}}
-
 \newcommand{\bm}[1]{\mbox{\boldmath $#1$}}
 
+ \def\linenumberfont{\normalfont\scriptsize}
 %===========================================
@@
 
-%\linenumbers{}
-%\modulolinenumbers[3]
+\linenumbers{}
+\modulolinenumbers[3]
 
@@
 			\item Briefly describe extension to simultaneous hermaphrodites, citing Jordan \& Connallon (2014), Abbot \& Tazzyman (2014), Olito (2017). 
-			\item breakdown into individuals produced by selfing vs.~outcrossing.
+			\item Breakdown into individuals produced by selfing vs.~outcrossing.
 			\item Brief anatomy of $\mathbf{\tilde{A}}$.
@@
 
-Conflicting selection and gene-flow between different classes of individuals in a population, or genetic trade-offs between fitness components, can impose genetic constraints on adaptation with particularly interesting evolutionary consequences \citep{CharlesworthHughes2000, ConnallonHall2018}.
-Such constraints hinder adaptation by preventing individuals (or classes of individuals) from reaching the phenotypic optimum in one or more fitness components, potentially increasing the population's overall risk of extinction.
+Conflicting selection and gene-flow between different classes of individuals, or genetic trade-offs between fitness components, can impose constraints on adaptation with particularly interesting evolutionary consequences \citep{CharlesworthHughes2000, ConnallonHall2018}.
+Such constraints hinder adaptation by preventing individuals (or classes of individuals) from reaching the phenotypic optimum in one or more fitness components, thereby increasing the population's overall extinction risk.
 At the same time, they provide an effective mechanism for the maintenance of genetic variation, which in turn influences the population's capactiy for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
@@
 In this article, we build upon the models of \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b} to develop a stage-structured mendelian matrix model that accommodates selection through male and female function in simultaneous hermaphrodites with mixed mating systems.
-Using the model, we jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecy) and partially selfing hermaphrodite populations.
+Using the model, we jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations.
 We focus our analyses on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population is also demographically viable (i.e., has a positive intrinsic growth rate).
@@
 
-A key finding from our model is that when the intrinsic population growth rate is not large (i.e., $\lambda$ is close to $1$), the deleterious effects on female fecundity of segregating male-beneficial SA alleles can render much of the parameter space for SA polymorphism demographically inviable.
-Under some conditions, self-fertilization can alleviate the demographic costs of balanced SA polymorphisms, but the concommitant costs of inbreeding depression greatly exacerbate them.
-Overall, our findings suggest that previous population genetic theory assuming large and/or constant population sizes probably represents a 'best case scenario' for the maintenance of SA polymorphisms, and provide a more nuanced picture of the nature of SA genetic variation that we should expect to find in natural populations.
+A key finding from our analysis is that when the intrinsic population growth rate is close to $1$, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space for SA polymorphism.
+Furthermore, this demographically viable polymorphic parameter space is often biased towards 'female-beneficial/male-deleterious' SA alleles (i.e., those with weaker selection on the female- than male sex function).
+Under some conditions, self-fertilization can alleviate the demographic costs of balanced SA polymorphisms, but the concommitant effects of inbreeding depression greatly exacerbate them.
+Overall, our findings provide a more nuanced picture of the nature of SA genetic variation that we should expect to find in natural populations, where the fate of SA alleles and the populations harboring them is determined jointly by evolutionary and demographic processes.
 
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Methods}
+\section*{The Model}
 %%%%%%%%%%%%%%%%%%%%%%%%
-\ldots
 
-%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{The quickness of the fox}
+Here, we briefly derive a matrix model incorporating multiple life-cycle stages and a single bi-allelic locus under SA selection on male and female sex functions for a population of partially selfing simultaneous hermaphrodites.
+The derivation and some of the key results follow closely those presented in \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b}; and in fact, the model presented here reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019b} under obligate outcrossing.
+A full derivation of the model and analyses is presented in \hl{Appendix A}, and all simulation code necessary to reproduce the results are available at \url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
 
-\ldots
+In the model, individuals are classified jointly by life-cycle stage ($1, \ldots, \omega$), genotype ($1, \ldots, g$), and whether they were produced by self- or outcross fertilization (denoted by $S$ and $X$ superscripts).
+Each genotype is characterized by a matrix of transition probabilities (including survival) and a matrix of reproductive output through male and female function.
+These matrices can include time variation or nonlinearities reflecting the environment or density dependence, but these will not be modelled here.
+Each stage contributes offspring to genotypes at the next time step according to transition matrices that are determined by the mating system and stage-structure of the population.
 
-\subsubsection*{Third-order heading}
+The population state at time $t$ is described by a stage $\times$ genotype distribution matrix, which can be transformed into a population state vector, $\tilde{\mbf{n}}(t)$, which is ordered by genotype, then stage, then by how individuals were produced (self vs.~outcrossing).
+For a single locus with two alleles ($A$ and $a$), we have genotypes $g \in \{AA,\, Aa,\, aa\}$, giving the population state vector:
+\begin{linenomath*}
+\begin{equation}
+	\tilde{\mbf{n}}(t) =  \left[
+								\begin{array}{c}
+									\mbf{n}^{S}_{AA}(t) \\
+									\mbf{n}^{S}_{Aa}(t) \\
+									\mbf{n}^{S}_{aa}(t) \\ \hline
+									\mbf{n}^{X}_{AA}(t) \\
+									\mbf{n}^{X}_{Aa}(t) \\
+									\mbf{n}^{X}_{aa}(t) \\ 
+						\end{array} \right],
+\end{equation}
+\end{linenomath*}
+
+\noindent where $\mbf{n}^{S}_{i}$ and $\mbf{n}^{X}_{i}$ (where $i \in \{AA,Aa,aa\}$) are the stage $\times$ genotype distribution vectors of individuals produced by self-fertilization and outcrossing, respectively.
+The proportional population vector is given by
+\begin{linenomath*}
+\begin{equation} \label{eq:propPopVec}
+	\tilde{\mbf{p}}(t) = \frac{\tilde{\mbf{n}}(t)}{ \| \tilde{\mbf{n}}(t) \|} = \mbf{p}(t),
+\end{equation}
+\end{linenomath*}
+
+\noindent where $\| \cdot \|$ is the one-norm.
+The population vector $\tilde{\mbf{n}}(t)$ is projected forward from time $t$ to $t + 1$ by the $projection$ matrix $\tilde{\mbf{A}}(\tilde{\mbf{n}})$ such that
+\begin{linenomath*}
+\begin{equation}
+	\tilde{\mbf{n}}(t + 1) = \tilde{\mbf{A}}[\tilde{\mbf{n}}(t)] \, \tilde{\mbf{n}}(t)
+\end{equation}
+\end{linenomath*}
+
+The population projection matrix $\tilde{\mbf{A}}$ is constructed from four sets of matrices representing the demographic and genetic processes: The matrices $\mbf{U}^{S}_{i}$ and $\mbf{U}^{X}_{i}$ contains transition and survival probabilities for each genotype, produced by selfing and outcrossing respectively.
+The matrices $\mbf{F}_{i}$ and $\mbf{F}^{\prime}_{i}$ contain the genotype $\times$ stage specific contribution via female and male sex-functions of genotype $i$ to the female and male gamete pools, respectively, and therefore to zygotes in the next generation.
+In principle, the model can admit any kind of pleiotropy among demographic traits because the transition and fertility matrices can differ among genotypes in any way.
+For simplicity, we assume linear, time-invariant demography.
+
+Similar to the two-sex model of \citet{deVriesCaswell2019b}, we define the matrices $\mbf{H}^S_{j}(\tilde{\mbf{n}})$ and $\mbf{H}^X_{j}(\tilde{\mbf{n}})$, which map the genotypes of the parent in stage $j$ to the genotypes of their offspring produced by selfing and outcrossing, respectively.
+The $(k, l)$ entry of $\mbf{H}^{X}_{j}$ and $\mbf{H}^{S}_{i}$ is the probability that an offspring of a genotype $l$ mother, of stage $i$, has genotype $k$.
+For the purpose of this article, we assume that mating is random with respect to stage and hence that the parent-offspring map is the same for all stages (i.e., $\mbf{H}^S_{j}(\tilde{\mbf{n}}) = \mbf{H}^S(\tilde{\mbf{n}})$, and $\mbf{H}^X_{j}(\tilde{\mbf{n}}) = \mbf{H}^X(\tilde{\mbf{n}})$).
+
+%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection*{Mating and offspring production under partial selfing}
+
+The matrices $\mbf{H}^S(\tilde{\mbf{n}})$ and $\mbf{H}^X(\tilde{\mbf{n}})$ contain the population genetic processes and are derived in detail in \hl{Appendix A: Section A.x}.
+
+%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Analyses}
 

### 2020-11-17 R/functions-Figs.R
New Supp polySpace Fig

@@ -454,6 +454,107 @@ layout     <- layout(layout.mat,respect=TRUE)
 }
 
 
+
+##############################################################
+##############################################################
+##  Final Supplemental Figs
+##############################################################
+##############################################################
+
+suppPolySpaceThresholdFigs  <-  function(df = "dataPolySpaceFig_sMax0.15_res0.003_delta0_dj0_da0_dg0") {
+
+    # Make filenames for import from df names
+    fName  <-  paste('./output/simData/', df, '.csv', sep="")
+
+    # import data
+    data  <-  read.csv(file=fName, header=TRUE)
+
+    # Clean up aberrant smExt value
+    data$smExt[90]  <-  mean(c(data$smExt[89], data$smExt[91]))
+
+    # Extract plotting parameter values from df names
+    d1   <-  strsplit(df, '_')[[1]][c(2:7)]
+    pars  <-  list(
+                    "sMax"  =  as.numeric(strsplit(d1[1],'x')[[1]][2]),
+                    "res"   =  as.numeric(strsplit(d1[2],'s')[[1]][2]),
+                    "d"     =  as.numeric(strsplit(d1[3],'a')[[1]][2]),
+                    "dj"    =  as.numeric(strsplit(d1[4],'j')[[1]][2]),
+                    "da"    =  as.numeric(strsplit(d1[5],'a')[[1]][2]),
+                    "dg"    =  as.numeric(strsplit(d1[6],'g')[[1]][2])
+                    )
+
+    hLev  <-  unique(data$h)
+    fLev  <-  unique(data$f)
+    CLev  <-  unique(data$C)
+    nHs   <-  length(hLev)
+    nCs   <-  length(CLev)
+
+    # Color scheme
+    COLS  <-  list("line"     =  transparentColor('#252525', opacity=1),
+                   "extinct"  =  transparentColor('red', opacity=0.15))
+
+    # Set plot layout
+    nMat        <-  nHs*nCs
+    layout.mat  <- matrix(c(1:nMat), nrow=nHs, ncol=nCs, byrow=TRUE)
+    layout      <- layout(layout.mat,respect=TRUE)
+
+    # loop through and generate plots
+    for(i in 1:nHs) {
+        # Subset data
+        dd  <-  data[data$h == hLev[i],]            
+        
+        for(j in 1:nCs) {
+            
[…]

### 2020-11-17 makeFigs.R

@@ -28,6 +28,17 @@ toPdf(demViablePolySpaceFig(),
 			figPath(name='extinctionThresholdsFig.pdf'), width=10, height=5)
 embed_fonts(figPath(name='extinctionThresholdsFig.pdf'))
 
+
+
+#################
+# Supp. Figs
+#################
+
+toPdf(suppPolySpaceThresholdFigs(), 
+			figPath(name='polySpaceSuppFig.pdf'), width=10, height=5)
+embed_fonts(figPath(name='polySpaceSuppFig.pdf'))
+
+
 ###############
 # PRELIM FIGS
 ###############

### 2020-11-20 doc/SA-Hermaphrodites-wDemography.tex
New Supp polySpace Fig
Minor edits to Methods & polySpace fig
Minor stuff here.

@@
 \usepackage{fullpage}
-\usepackage[authoryear,sectionbib,sort]{natbib} %\bibpunct{(}{)}{;}{author-year}{}{,}
+\usepackage[authoryear,sectionbib,sort]{natbib}
 \bibliographystyle{amnatnat}
@@
 \usepackage[colorlinks=true, allcolors=black]{hyperref}
+\usepackage{nameref}
+
+% Section Header Formats
 \titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
@@
 
+% Special Math Characters
 \newcommand\encircle[1]{%
@@
 For the purpose of this article, we assume that mating is random with respect to stage and hence that the parent-offspring map is the same for all stages (i.e., $\mbf{H}^S_{j}(\tilde{\mbf{n}}) = \mbf{H}^S(\tilde{\mbf{n}})$, and $\mbf{H}^X_{j}(\tilde{\mbf{n}}) = \mbf{H}^X(\tilde{\mbf{n}})$).
+The matrices $\mbf{H}^S(\tilde{\mbf{n}})$ and $\mbf{H}^X(\tilde{\mbf{n}})$ contain the population genetic processes and are presented in the next section.
 
@@
 
-The matrices $\mbf{H}^S(\tilde{\mbf{n}})$ and $\mbf{H}^X(\tilde{\mbf{n}})$ contain the population genetic processes and are derived in detail in \hl{Appendix A: Section A.x}.
+Simultaneous hermaphrodites express both male and female sex functions (i.e., can be both a mother and father), and have the potential to reproduce by a combination of self- and outcross-fertilizations.
+Maternal outcrossing involves receiving male gametes from another individual in the population, while paternal outcrossing involves exporting male gametes to another individual and fertilizing their ovules.
+Self-fertilization is acheived when an individual's male gametes fertilize their own ovules.
+These modes of reproduction complicate mating in our model relative to the two-sex stage-structured model of \citet{deVriesCaswell2019b}.
+
+Let $C$ denote the proportion of each individual's ovules that are self-fertilized (the remaining $1 - C$ are outcrossed), and $\delta$ represent the viability cost suffered by self-fertilized offspring due to inbreeding depression \citep{Charlesworth1987}.
+In \hl{Appendix A.x}, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
+Outcrossing proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}, where the number of offspring is determined by the maternal genotype, while the male genotype determines paternal "mating success".
 
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Analyses}
+\subsection*{Analyses} \label{sec:analyses}
 
@@
  \centering
- \includegraphics[width=\linewidth]{./extinctionThresholdsFig}
+ \includegraphics[width=\linewidth]{../output/figs/extinctionThresholdsFig.pdf}
  \caption{Illustration of parameter space for SA polymorphism and extinction thresholds predicted by the mendelian stage-structured matrix model.
@@
 }
- \label{fig:extinctionThresholds}
+ \label{fig:extThresholds}
+ \end{figure}
+
+ \begin{figure}[htbp]
+ \centering
+ \includegraphics[scale=0.75]{../output/figs/polymorphicSpaceTitrate.pdf}
+ \caption{Proportion of demographically viable parameter space in the absence of inbreeding depression (i.e., assuming $\delta = \delta_i = 0$, where $i \in \{j,a,\gamma\}$) predicted by the matrix model, plotted as a function of the population selfing rate.
+Results are shown for three fertility values corresponding to low, medium, and high fertility (blue, green, and red points respectively) under additive (panel A, $h = 1/2$), and partially recessive (panel B, $h = 1/4$) SA fitness effects.
+Each point was calculated by approximate numerical integration of the corresponding SA invasion conditions and extinction threshold predicted by the mendelian matrix model (see \hyperref[sec:analyses]{Analyses} section), while solid lines were produced by numerically integrating the analytic expressions for the single-locus invasion conditions from the population genetic models of \citet{JordanConnallon2014} and \citet{Olito2017} (solid black lines).
+}
+ \label{fig:polySpace}
  \end{figure}

### 2020-11-20 R/functions-Simulations.R

@@ -2339,12 +2339,7 @@ makeDataDeltaPolyParamSpace  <-  function(sMax=0.15, res=0.0015, precision = 1e-
 
 		cat('\r', "hVal = ", hVals[i], '\n')
 
-		if(hVals[i] == 1/2) {
-			Cs  <-  seq(0, 0.9, by=0.05)
-		}
-		if(hVals[i] == 1/4) {
-			Cs  <-  c(0.01, seq(0.05, 0.9, by=0.05))
-		}	
+		Cs  <- c(seq(0, 0.5, by=0.025), 0.75, 0.9)
 		
 		# Define dStar ~ C
 		deltaSeq  <-  predDelta(dStar=dStar, b=1/2, a=0.2, C=Cs)
@@ -2369,7 +2364,7 @@ makeDataDeltaPolyParamSpace  <-  function(sMax=0.15, res=0.0015, precision = 1e-
 											  hf = hVals[i], hm = hVals[i], C = Cs[j], 
 											  delta = deltaSeq[j], delta_j = delta_j, delta_a = delta_a, delta_gamma = delta_gamma,
 											  tlimit = tlimit, eqThreshold = eqThreshold, 
-											  makePlotts=FALSE, verbose = TRUE, writeFile=FALSE)
+											  makePlots=FALSE, verbose = FALSE, writeFile=FALSE)
 
 			# Append new data to dataSet
 			dataSet  <-  rbind(dataSet, cbind(rep("delta", times=nrow(invData)),
@@ -2401,7 +2396,7 @@ makeDataDeltaPolyParamSpace  <-  function(sMax=0.15, res=0.0015, precision = 1e-
 											  om = om, g = g, theta = theta, theta_prime = theta_prime, 
 											  hf = hVals[i], hm = hVals[i], C = Cs[j], 
 											  delta = delta, delta_j = deltaSeq[j], delta_a = delta_a, delta_gamma = delta_gamma,
-											  tlimit = tlimit, eqThreshold = eqThreshold, verbose = TRUE, writeFile=FALSE)
+											  tlimit = tlimit, eqThreshold = eqThreshold, verbose = FALSE, writeFile=FALSE)
 
 			# Append new data to dataSet
 			dataSet  <-  rbind(dataSet, cbind(rep("delta", times=nrow(invData)),
@@ -2433,7 +2428,7 @@ makeDataDeltaPolyParamSpace  <-  function(sMax=0.15, res=0.0015, precision = 1e-
[…]

### 2020-11-20 run-Simulations.R

@@ -242,10 +242,6 @@ extinctThreshTitrate(sMax=0.15, res=0.0015, precision=1e-4,
 
 
 
-
-
-
-
 ##########################################################
 # Fig. 2 - Dem. viable polymorphic param. space (no I.D.)
 
@@ -259,6 +255,11 @@ makeDataPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
 
 ##########################################################
 # Fig. 3 - Effects of inbreeding depression (no I.D.)
+
+makeDataDeltaPolyParamSpace(sMax=0.15, res=0.0015, precision = 1e-4,
+							om = 2, g = 3, theta = c(0.6,0.6,0.05,6.5), theta_prime = 6.5, 
+							hVals= c(1/2, 1/4), dStar = 0.8, 
+							tlimit = 10^5, eqThreshold = 1e-8)
 
 
 ######################

### 2020-11-20 doc/SA-Hermaphrodites-wDemography.tex
Methods up to Analyses subsection
Continued writing out methods, up to the Analyses subsection.

@@
 \usepackage[authoryear,sectionbib,sort]{natbib}
-\bibliographystyle{amnatnat}
+\bibliographystyle{amnatnat.bst}
 \setlength{\bibsep}{0.0pt}
@@
 
-\section*{Outline}
+%\section*{Outline}
 
-OK, just putting some ideas out here based roughly on our conversations about an {\itshape AmNat}-style article, the results we have so far, and what I think we should be able to do with data from COMPADRE.
+%OK, just putting some ideas out here based roughly on our conversations about an {\itshape AmNat}-style article, the results we have so far, and what I think we should be able to do with data from COMPADRE.
 
-\begin{enumerate}
-	\item {\bf Introduction}
-	\begin{enumerate}
-		\item Briefly introduce sexual antagonism and intra-locus sexual conflict. (Colin)
-		\item Simultaneous hermaphrodites, mixed mating systems, unique predictions for SA polymorphism. (Colin)
-		\item Introduce Evolutionary Demography, pivot to focus on the consequences of taking demography into account explicitly for the maintenance of SA polymorphism. (Lotte)
-		\item Introduce key difference between existing population genetic models of SA polymorphism and our models: possibility of diverse population and evolutionary dynamics. (Lotte then Colin)
-		\item Explain and highlight "demographically viable" parameter space, indicate that we are interested in whether parameter space predicted to be polymorphic by Pop.~gen.~models is, indeed, "demographically viable" (Lotte then Colin)
-		\begin{enumerate}
-			\item Pop.
+%\begin{enumerate}
+	%\item {\bf Introduction}
+	%\begin{enumerate}
+%		\item Briefly introduce sexual antagonism and intra-locus sexual conflict. (Colin)
+%		\item Simultaneous hermaphrodites, mixed mating systems, unique predictions for SA polymorphism. (Colin)
+%		\item Introduce Evolutionary Demography, pivot to focus on the consequences of taking demography into account explicitly for the maintenance of SA polymorphism. (Lotte)
+%		\item Introduce key difference between existing population genetic models of SA polymorphism and our models: possibility of diverse population and evolutionary dynamics. (Lotte then Colin)
+%		\item Explain and highlight "demographically viable" parameter space, indicate that we are interested in whether parameter space predicted to be polymorphic by Pop.~gen.~models is, indeed, "demographically viable" (Lotte then Colin)
+%		\begin{enumerate}
+%			\item Pop.
 Gen.
@@
 Casually define fitness as lifetime reproductive fitness of a given genotype.
-			\item "In practice, these genotypic fitnesses emerge from myriad process acting throughout the life-history of individuals"
-			\item these processes contribute to key demographic rates such as the intrinsic population growth rate, 
-			\item 
-		\end{enumerate}
-		\item Building upon the two-sex stage-structured matrix models developed by \citet{deVriesCaswell2019a,deVriesCaswell2019b}, we develop stage-structured matrix models that can accommodate selection through male and female function in simultaneous hermaphrodites with mixed mating systems.
+%			\item "In practice, these genotypic fitnesses emerge from myriad process acting throughout the life-history of individuals"
+%			\item these processes contribute to key demographic rates such as the intrinsic population growth rate, 
+%			\item 
+%		\end{enumerate}
+%		\item Building upon the two-sex stage-structured matrix models developed by \citet{deVriesCaswell2019a,deVriesCaswell2019b}, we develop stage-structured matrix models that can accommodate selection through male and female function in simultaneous hermaphrodites with mixed mating systems.
 The models also allow for the effects of inbreeding depression to manifest in several demographic parameters, allowing us to explore the fitness consequences of early- vs.~late-acting inbreeding depression.
 (Colin then Lotte)
-	\end{enumerate}
-
-	\item {\bf Methods}
-		\begin{enumerate}
-			\item Citation heavy review of de Vries \& Caswell AmNat (2018) \& de Vries \& Caswell AmNat (2019) TheorPopBiol
-			\item Briefly describe extension to simultaneous hermaphrodites, citing Jordan \& Connallon (2014), Abbot \& Tazzyman (2014), Olito (2017). 
-			\item Breakdown into individuals produced by selfing vs.~outcrossing.
-			\item Brief anatomy of $\mathbf{\tilde{A}}$.
-			\item Analysis. Invasion analysis \& abridged derivation of invasion conditions \& identification of extinction threholds.
-			\item Cite Appendix with full derivation distilled from notes file.
-		\end{enumerate}
-
-	\item {\bf When is polymorphism demographically viable?}
-	\begin{enumerate}
-		\item Review previous predictions for outcrossers \& partial selfing
-		\begin{enumerate}
-			\item Greater scope for SA polymorphism in outcrossers vs.~selfers
-			\item Greater scope for SA polymorphism when SA fitness effects are recessive (dominance reversal).
-			\item but our model allows us to see where/when invasion of SA alleles can drive the population extinct
-			\item \textbf{Fig.1 - illustration of funnel plots w/ extinction isoclines}
-		\end{enumerate}
-
-		\item Key results summarized in \textbf{Fig.~2}: 
-		\begin{enumerate}
-			\item For simplicity, temporarily assume no inbreeding depression ($\delta = 0$).
-			\item When populations do not have high growth-rates (i.e., when $\lambda \approx 1$), much of of polymorphic parameter space becomes demographically inviable.
-			\item Specifically, would-be polymorphic paramter space becomes inviable when segregating male-beneficial/female-deleterious alleles can reduce fecundity enough to drive the population extinct.
-			\item This effect is most dramatic for separate-sexed species and predominant outcrossers.
+	%\end{enumerate}
+	%\item {\bf Methods}
+%		\begin{enumerate}
+%			\item Citation heavy review of de Vries \& Caswell AmNat (2018) \& de Vries \& Caswell AmNat (2019) TheorPopBiol
+%			\item Briefly describe extension to simultaneous hermaphrodites, citing Jordan \& Connallon (2014), Abbot \& Tazzyman (2014), Olito (2017). 
+%			\item Breakdown into individuals produced by selfing vs.~outcrossing.
+%			\item Brief anatomy of $\mathbf{\tilde{A}}$.
+%			\item Analysis. Invasion analysis \& abridged derivation of invasion conditions \& identification of extinction threholds.
+%			\item Cite Appendix with full derivation distilled from notes file.
+%		\end{enumerate}
+	%\item {\bf When is polymorphism demographically viable?}
+	%\begin{enumerate}
+%		\item Review previous predictions for outcrossers \& partial selfing
+%		\begin{enumerate}
+%			\item Greater scope for SA polymorphism in outcrossers vs.~selfers
+%			\item Greater scope for SA polymorphism when SA fitness effects are recessive (dominance reversal).
+%			\item but our model allows us to see where/when invasion of SA alleles can drive the population extinct
+%			\item \textbf{Fig.1 - illustration of funnel plots w/ extinction isoclines}
+%		\end{enumerate}
+%		\item Key results summarized in \textbf{Fig.~2}: 
+%		\begin{enumerate}
+%			\item For simplicity, temporarily assume no inbreeding depression ($\delta = 0$).
+%			\item When populations do not have high growth-rates (i.e., when $\lambda \approx 1$), much of of polymorphic parameter space becomes demographically inviable.
+%			\item Specifically, would-be polymorphic paramter space becomes inviable when segregating male-beneficial/female-deleterious alleles can reduce fecundity enough to drive the population extinct.
+%			\item This effect is most dramatic for separate-sexed species and predominant outcrossers.
 Selfing shelters the population from the female-deleterious fitness effects because there is no opportunity for selection via male function.
 This can also be thought of as reproductive assurance in our model....
-			\item Convergence to pop.~gen.~predictions when $f$ is big.
-		\end{enumerate}
-	\end{enumerate}
-
-	\item {\bf Effects of inbreeding depression}
-	\begin{enumerate}
-		\item Review effect of I.D.~in population genetic models: causes selfing populations to look more like outcrossing ones (at least in terms of where SA polymorphism is maintained).
+%			\item Convergence to pop.~gen.~predictions when $f$ is big.
+%		\end{enumerate}
+	%\end{enumerate}
+	%\item {\bf Effects of inbreeding depression}
+	%\begin{enumerate}
+%		\item Review effect of I.D.~in population genetic models: causes selfing populations to look more like outcrossing ones (at least in terms of where SA polymorphism is maintained).
 This is because a fraction of selfed ovules are aborted, meaning the overall ratio of outcrossed vs.~selfed progeny increases.
-		\item Matrix model allows us to explore effects of I.D.
+%		\item Matrix model allows us to explore effects of I.D.
 in different parts of the life-cycle (early- vs.~late-acting I.D.).
@@
 via ovule viability.
-		\item MAKE CLEAR that purging of deleterious recessives thought to cause I.D. is already included in the model results presented in Fig.~3
-		\item Key results summarized in Fig.~3
-		\begin{enumerate}
-			\item Pop.~Gen.~models predict that higher early-acting ID should lead to greater proportion of polymorphic parameter space, while our models predict that we get a critical threshold of selfing at which population viability plummets.
-			\item When and how I.D. influences fitness during the life-cyclecan change the location and steepness of the decline in polymorphic parameter space.
-			\item Critically, largest deviations from pop.~gen.~predictions happen for predominantly selfing populations.
-			\item A perhaps obvious, but sometimes overlooked implication of our model predictions: selfing can offer reproductive assurance, and SA polymorphisms can be maintained in a non-trivial proportion of parameter space in partially selfing populations...
+%		\item MAKE CLEAR that purging of deleterious recessives thought to cause I.D. is already included in the model results presented in Fig.~3
+%		\item Key results summarized in Fig.~3
+%		\begin{enumerate}
+%			\item Pop.~Gen.~models predict that higher early-acting ID should lead to greater proportion of polymorphic parameter space, while our models predict that we get a critical threshold of selfing at which population viability plummets.
+%			\item When and how I.D. influences fitness during the life-cyclecan change the location and steepness of the decline in polymorphic parameter space.
+%			\item Critically, largest deviations from pop.~gen.~predictions happen for predominantly selfing populations.
+%			\item A perhaps obvious, but sometimes overlooked implication of our model predictions: selfing can offer reproductive assurance, and SA polymorphisms can be maintained in a non-trivial proportion of parameter space in partially selfing populations...
 BUT demographically, the population must be able to "afford" the loss of selfed ovules/offspring.
-			\item HIGHLIGHT that this result holds even though we incorporated decrease of the mutation load due to inbreeding (and corresponding decrease in I.D.)...	
-		\end{enumerate}
-	\end{enumerate}
-
-	\item {\bf Mimulus as a potential real world example}
-	\begin{enumerate}
-		\item We can parameterize our model with real populations' demographic rates, and make explicit predictions regarding the opportunity for SA polymorphism given the species' demography.
+%			\item HIGHLIGHT that this result holds even though we incorporated decrease of the mutation load due to inbreeding (and corresponding decrease in I.D.)...	
+%		\end{enumerate}
+	%\end{enumerate}
+	%\item {\bf Mimulus as a potential real world example}
+	%\begin{enumerate}
+%		\item We can parameterize our model with real populations' demographic rates, and make explicit predictions regarding the opportunity for SA polymorphism given the species' demography.
 We provide two illustrative examples:
-		\item Opportunities for SA polymorphism in {\itshape Mimulus guttatus} (now {\itshape Erythranthe guttata})
-		\item Possibly use data from \citet{PetersonEtAl2016} Eagle Meadows population, which is the 'local' population in their big common garden local adaptation expeirment.
-		\begin{enumerate}
-			\item Illustrate funnel plot using {\itshape E.~gutatta} demographic rates.
-			\item \textbf{Fig.~4}: PLOT inversion from \citet{LeeKelly2015}, on that funnel!
-		\end{enumerate}
-
-	\end{enumerate}
-
-	\item Discussion
-		\begin{enumerate}
-			\item Classic prediction from demography that most species should have a $\lambda \approx 1$
-			\item If true, the demographic impacts on the opportunities for the maintenance of sexually antagonistic polymorphisms may be pervasive...
-			\item Condition dependence, fecundity, scope for sexual confict/antagonistic polymorphism?
-			\item Do sexually dimorphic species exhibit demographic rates that increase opportunities for SA polymorphism?
-
-		\end{enumerate}
-
-\end{enumerate}
+%		\item Opportunities for SA polymorphism in {\itshape Mimulus guttatus} (now {\itshape Erythranthe guttata})
+%		\item Possibly use data from \citet{PetersonEtAl2016} Eagle Meadows population, which is the 'local' population in their big common garden local adaptation expeirment.
+%		\begin{enumerate}
+%			\item Illustrate funnel plot using {\itshape E.~gutatta} demographic rates.
+%			\item \textbf{Fig.~4}: PLOT inversion from \citet{LeeKelly2015}, on that funnel!
+%		\end{enumerate}
+	%\end{enumerate}
+	%\item Discussion
+%		\begin{enumerate}
+%			\item Classic prediction from demography that most species should have a $\lambda \approx 1$
+%			\item If true, the demographic impacts on the opportunities for the maintenance of sexually antagonistic polymorphisms may be pervasive...
+%			\item Condition dependence, fecundity, scope for sexual confict/antagonistic polymorphism?
+%			\item Do sexually dimorphic species exhibit demographic rates that increase opportunities for SA polymorphism?
+%		\end{enumerate}
+%\end{enumerate}
 
@@
 Using the model, we jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations.
-We focus our analyses on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population is also demographically viable (i.e., has a positive intrinsic growth rate).
+Although a variety of interesting eco-evolutionary dynamics are possible in the model (e.g., \citealt{deVriesCaswell2019a, deVriesCaswell2019b,CaswellEtAl2018}), we focus our analyses on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population is also demographically viable (i.e., has a positive intrinsic growth rate).
 The model also allows for the effects of inbreeding depression to manifest in different life-history stages, enabling us to explore the fitness consequences of early- vs.~late-acting inbreeding depression on the model predictions.
@@
 
-Let $C$ denote the proportion of each individual's ovules that are self-fertilized (the remaining $1 - C$ are outcrossed), and $\delta$ represent the viability cost suffered by self-fertilized offspring due to inbreeding depression \citep{Charlesworth1987}.
+Let $C$ denote the proportion of each individual's ovules that are self-fertilized (the remaining $1 - C$ are outcrossed), and $\delta$ represents the proportion of self-fertilized zygotes that fail to develop due to inbreeding depression during early development \citep{Charlesworth1987}.
 In \hl{Appendix A.x}, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
-Outcrossing proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}, where the number of offspring is determined by the maternal genotype, while the male genotype determines paternal "mating success".
+Outcrossing proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}, except each individual's genotype determines both the number of ovules produced, and also paternal "mating success", broadly defined.
+For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986,WangBarrett2020,Harder2016}.
+Depending on their stage and genotype at the SA locus, individuals contribute male gametes to the overall population gamete pool according to the male stage $\times$ genotype vector:
+\begin{linenomath*}
+\begin{equation}
+		\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p} = 
+		\left(
+			\begin{array}{ccc}
+				\mbf{1}^{\intercal}_{\omega} & \frac{1}{2} \mbf{1}^{\intercal}_{\omega} & 0 \\
+				0 & \frac{1}{2} \mbf{1}^{\intercal}_{\omega} & \mbf{1}^{\intercal}_{\omega} \\
+			\end{array} \right)
+		\left(
+			\begin{array}{ccc}
+				\mbf{F}^{\prime}_{AA} & 0 & 0 \\
+				0 & \mbf{F}^{\prime}_{Aa} & 0 \\
+				0 & 0 & \mbf{F}^{\prime}_{aa} \\
+			\end{array} \right)
+		\left(
+			\begin{array}{c}
+				\mbf{p}_{AA} \\
+				\mbf{p}_{Aa} \\
+				\mbf{p}_{aa} \\
+			\end{array} \right).
+\end{equation}
+\end{linenomath*}
+
+\noindent The vector $\mbf{p}$ gives the genotype frequencies at mating, and is the sum of $\mbf{n}^{S}$ and $\mbf{n}^{X}$, since in partially selfing popualtions, whether an individual was produced by selfing or outcrossing does not affect whether they can reproduce by outcrossing themselves.
+The matrix $\mbb{F}^{\prime}$ is composed of genotype-specific fertility matrices, and operates on the the vector of genotype frequencies to give their relative contribution of each genotype to the gamete pool.
+The matrix $\mbf{W}^{\prime}$ converts these contributions to allele numbers.
+Normalizing the resulting vector gives the allele frequencies in the male gamete pool:
+\begin{linenomath*}
+\begin{equation} \label{eq:maleGametePool}
+	\left(
+		\begin{array}{c}
+			q^{\prime}_{A} \\
+			q^{\prime}_{a} \\
+		\end{array} \right) = 
+			\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p} \|} = 
+				\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)\|}
+\end{equation}
+\end{linenomath*}
+
+The parent-offspring map for outcrossing is a function of the allele frequencies in the male gamete pool, and is given by 
+\begin{linenomath*}
+\begin{equation} \label{eq:HX}
+	\mbf{H}^X(\tilde{\mbf{n}}) = 
+			\left(
+			\begin{array}{ccc}
+				q^{\prime}_{A} & \frac{1}{2} q^{\prime}_{A} & 0 \\
+				q^{\prime}_{a} & \frac{1}{2} & q^{\prime}_{A}  \\
+				0 & \frac{1}{2} q^{\prime}_{a} & q^{\prime}_{a} \\
+			\end{array} \right)
+\end{equation}
+\end{linenomath*}
+
+\noindent From left to right, the columns of matrix $\mbf{H}^X(\tilde{\mbf{n}})$ give the genotype distribution of outcrossed offspring produced by a maternal parent of each genotype ($AA$, $Aa$, and $aa$ respectively).
+
+Reproduction via self-fertilization differs from outcrossing.
+As in previous models of partial selfing, we assume that individuals produce enough male gametes to easily self-fertilize all of their ovules, and that self-fertilization involves little or no selection from external factors relative to outcrossing (e.g., \citealt{Charlesworth1978a,JordanConnallon2014,Olito2017}, but see \citealt{Tazzyman2015}).
+Under these assumptions, the genotype distributions of selfed offspring are determined entirely by the parental genotype and the probabilities of segregation and fertilization during meiosis:
+\begin{linenomath*}
+\begin{equation} \label{eq:HS}
+	\mbf{H}^S(\tilde{\mbf{n}}) = 
+			\left(
+			\begin{array}{ccc}
+				1 & 1/4 & 0 \\
+				0 & 1/2 & 0 \\
+				0 & 1/4 & 1 \\
+			\end{array} \right).
+\end{equation}
+\end{linenomath*}
+
+\noindent The columns of $\mbf{H}^S(\tilde{\mbf{n}})$ again give the selfed offspring genotype distributions for parental genotypes of $AA$, $Aa$, and $aa$, respectively.
+
+%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection*{Population projection}
+
+Using the component matrices described in the previous section, the matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ that projects the eco-evolutionary dynamics is:
+\begin{linenomath*}
+\begin{equation} \label{eq:Atilde}
+	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
+			\underbrace{\left(
+			\begin{array}{c|c}
+				\mcal{U}^S & \mbf{0} \\ \hline
+				\mbf{0} & \mcal{U}^X \\
+			\end{array} \right)}_{\tilde{\mbf{U}}} + 
+			\underbrace{\left(
+			\begin{array}{c|c}
+				\mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
+				\mcal{F}^X(\tilde{\mbf{p}}) (1 - C) & \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)\\
+			\end{array} \right)}_{\tilde{\mbf{F}}}.
+\end{equation}
+\end{linenomath*}
+
+\noindent The blocks of the component matrices in Eq(\ref{eq:Atilde}) correspond to production of offspring by self-fertilization and outcrossing ($\mcal{F}^S$ and $\mcal{F}^X$ in $\tilde{\mbf{F}}$), and survival of selfed and outcrossed offspring ($\mcal{U}^S$ and $\mcal{U}^X$ in $\tilde{\mbf{U}}$).
+
+We construct $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ using the vec-permutation approach of \citet{CaswellEtAl2018}. Briefly, we construct a set of block-diagonal matrices: e.g.,
+\begin{linenomath*}
+\begin{equation} \label{eq:US}
+	\mbb{U}^S = 
+		\left(
+			\begin{array}{ccc}
+				\mbf{U}^{S}_{AA} & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{U}^{S}_{Aa} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \mbf{U}^{S}_{aa} \\
+			\end{array} \right),
+\end{equation}
+\end{linenomath*}
+
+\noindent and corresponding matrices $\mbb{U}^X$, $\mbb{F}^S$, $\mbb{F}^X$.
+Since individuals cannot change their genotype once they are born, the survival matrices are block diagonal, and $\mcal{U}^S = \mbb{U}^S$ and $\mcal{U}^X = \mbb{U}^X$.
+
+Similarly, we also construct block-diagonal matrices
+\begin{linenomath*}
+\begin{equation}
+	\mbb{H}^j = \mbf{I}_{\omega} \otimes \mbf{H}^j(\tilde{\mbf{p}}),
+\end{equation}
+\end{linenomath*}
+
+\noindent where $j \in \{S,X\}$. The corresponding fertility matrices for self-fertilization and outcrossing are defined as
+\begin{linenomath*}
+\begin{equation} \label{eq:FS}
+	\mcal{F}^S = \mbf{K}^{\intercal} \mbb{H}^S(\tilde{\mbf{p}}) \mbf{K} \mbb{F}^S = 
+			\left(
+			\begin{array}{ccc}
+				\mbf{F}_{AA} & \frac{1}{4} \mbf{F}_{Aa} & 0 \\
+				0 & \frac{1}{2} \mbf{F}_{Aa} & 0 \\
+				0 & \frac{1}{4} \mbf{F}_{Aa} & \mbf{F}_{aa}\\
+			\end{array} \right), 
+\end{equation}
+\end{linenomath*}
+\noindent and 
+\begin{linenomath*}
+\begin{equation} \label{eq:FX}
+	\mcal{F}^X = \mbf{K}^{\intercal} \mbb{H}^X(\tilde{\mbf{p}}) \mbf{K} \mbb{F}^X = 
+			\left(
+			\begin{array}{ccc}
+				q^{\prime}_{A} \mbf{F}_{AA} & \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} & 0 \\ 
+				q^{\prime}_{a} \mbf{F}_{AA} & \frac{1}{2} \mbf{F}_{Aa} & q^{\prime}_{A} \mbf{F}_{aa}  \\ 
+				0 & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} & q^{\prime}_{a} \mbf{F}_{aa}  \\
+			\end{array} \right),
+\end{equation}
+\end{linenomath*}
+
+\noindent where $q^{\prime}_A$ and $q^{\prime}_a$ are given by Eq(\ref{eq:maleGametePool}), and $\mbf{K}$ is the vec-permutation matrix \citep{HendersonSearle1981}.
+Further details of the derivations of $\mcal{F}^S$ and $\mcal{F}^X$ are provided in \hl{Appendix A.x}.
+
+The blocks of $\mcal{F}^X(\tilde{\mbf{p}})$ can be constructed and interpreted as follows: The first row block of the first column produces $AA$ offspring by outcrossing from $AA$ maternal parents.
+This happens when the $AA$ maternal parent receives a random gamete from the male gamete pool, which happens with probability $q^{\prime}_{A}$.
+The other blocks can be interpreted similarly.
+
+Combining all the component matrices yields the overall eco-evolutionary projection matrix shown in \hl{Reference a Box here?  or possibly Appendix A.x?}
+
+%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Sexually antagonistic selection \& inbreeding depression} \label{sec:SAsel}
+
+We now construct and analyze a genotype $\times$stage-classified model for a hypothetical species with intralocus sexual conflict via the two sex functions.
+To keep the model analytically tractable, we assume our hypothetical species has two life stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active.
+Suppose that there is a genetic trade-off between the sex-functions at a single bi-allelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect (see Table \ref{tab:Fitness}).
+The allele does not affect survival and transition rates.
+However, the survival matrices can be used to model the fitness effects of inbreeding depression at later stages of development by allowing different stage-specific survival and transition rates for individuals produced by self-fertilization vs.~outcrossing.
+By contrast, the parameter $\sigma$ only affects inbreeding depression through viability of selfed ovlues.
+With this in mind, we define survival matrices for individuals produced by selfing and outcrossing as follows:
+\begin{linenomath*}
+\begin{equation} \label{eq:US}
+	\mbf{U}^S = \left(
+					\begin{array}{cc}
+						\sigma^S_J (1 - \delta_J) \big(1 - \gamma^S (1 - \delta_{\gamma}) \big) & 0 \\
+						\sigma^S_J (1 - \delta_J) \gamma (1 - \delta_{\gamma})      & \sigma^S_A (1 - \delta_S)
+					\end{array}
+				\right) \\
+\end{equation}
+\end{linenomath*}
+\noindent and 
+\begin{linenomath*}
+\begin{equation}\label{eq:UX}
+	\mbf{U}^X = \left(
+					\begin{array}{cc}
+						\sigma^X_J(1 - \gamma^X) & 0 \\
+						\sigma^X_J \gamma      & \sigma^X_A
+					\end{array}
+				\right)
+\end{equation}
+\end{linenomath*}
+
+\noindent where $\sigma_J$ and $\sigma_A$ are the juvenile and adult stage survival rates, and $\gamma$ is the maturation rate from juvenile to adult stages.
+The corresponding terms $\delta_S$, $\delta_A$, and $\delta_{\gamma}$ denote the proportional decreases in stage-specific survival and transition rates due to inbreeding depression (i.e., deleterious effects of inbreeding at later life-history stages; e.g., \citealt{HarderRoutely2006}).
+Any or all of these parameters may differ among genotypes, so that selection can operate on stage-specific viability, development, and/or fertility.
+For simplicity, we assume these are constant among genotypes.
+
+The fertility matrices through male and female function are
+\begin{linenomath*}
+\begin{equation}\label{eq:FS}
+	\mbf{F}_{i} = \left(
+					\begin{array}{cc}
+						0 & f_{i} \\
+						0 & 0
+					\end{array}
+				\right),
+\end{equation}
+\end{linenomath*}
+\noindent and
+\begin{linenomath*}
+\begin{equation}\label{eq:FX}
+	\mbf{F}^{\prime}_{i} = \left(
+					\begin{array}{cc}
+						0 & f^{\prime}_{i} \\
+						0 & 0
+					\end{array}
+				\right),
+\end{equation}
+\end{linenomath*}
+
+\noindent where $f_{i}$ is the genotype specific adult fertility though female function, and $f^{\prime}_{i}$ is the same through male function (see Table \ref{tab:Fitness}).
+
+\begin{table}[htbp]
+ \centering
+ \caption{\bf Fitness expressions for Sexually Antagonistic selection in hermaphrodites}
+\label{tab:Fitness}
+\begin{tabular}{lccc}
+ \toprule
+					&  \multicolumn{3}{c}{{\textit{Genotype}}} \\ 
+\cline{2-4}
+					& $AA$			& $Aa$ 					& $aa$ 		\\ \hline
+Female function:	& $f_{AA}$		& $f_{Aa} - h_f s_f$	& $f_{aa} - s_f$ \\	
+Male function:		& $f^{\prime}_{AA} - s_m$& $f^{\prime}_{Aa} - h_m s_m$	& $f^{\prime}_{aa}$ 		\\	
+\hline
+\end{tabular}
+\end{table}
+
+Iterating the projection matrix, Eq(\ref{eq:Atilde}) with the above demographic matrices provides numerical simulation of the eco-evolutionary dynamics for selection operating on any of the sex-function, or stage-specific demographic parameters.
+As we outline below, we use both numerical simulations together with mathematical analyses to study the conditions for the maintenance of SA polymorphisms, and the demographic fate of the populations (i.e., positive growth, or extinction).
 
@@
 
-\ldots
+A variety of eco-evolutionary outcomes are possible in the stage-structured mendelian matrix model, including fixation of one or the other allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and evolutionary suicide (e.g., see \citealt{CaswellEtAl2018,deVriesCaswell2019a,deVriesCaswell2019b}.
+We focus on identifying the necessary conditions for protected SA polymorphism with a positive intrinsic population growth rate, which we refer to as demographically viable SA polymorphisms.
+....
 
@@
 
-\section*{Acknowledgments}
+%\section*{Acknowledgments}
 
-OEC would like to thank Madlen Wilmes, Gyuri Barab\'{a}s, Flo D\'{e}barre, Vlastimil K\v{r}ivan, and Greg Dwyer for their comments and suggestions on this template.
+% ...
+We are particularly grateful to the European Society for Evolutionary Biology (ESEB) for funding the Special Topics Network workshops “Linking local adaptation with the evolution of sex differences”, and to the participants, all of whom made this study possible.
+This study was also supported by a Wenner-Gren Foundation postdoctoral stipen to C.O., and [Lotte's funding].
 
@@
 % Dryad references.
-\bibliography{Refs2}
+\bibliography{Refs2.bib}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%
+% Tables
+%%%%%%%%%%%%%%%%%%%%%
 \section*{Tables}
@@
 \setcounter{table}{0}
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}[htbp]
+\centering
+\caption{\bf Definition of terms.}
+\label{tab:Terms}
+\begin{tabular}{ l l r }
+ \toprule
+Symbol & Definition & Dimension \\
+\hline
+$g$      & number of genotypes ($3$; $AA$, $Aa$, and $aa$) & \\
+$\omega$ & number of stages ($2$; juvenile and adult) & \\
+$\tilde{\mbf{n}}$ & Joint stage $\times$ genotype vector & $2 \omega g \times 1$ \\
+$\tilde{\mbf{p}}$ & Joint stage $\times$ genotype frequency vector & $2 \omega g \times 1$ \\
+
+$\mbf{U}^{S}_{i}$, $\mbf{U}^{X}_{i}$ & Genotype-specific transition and survival matrices & $\omega \times \omega$ \\
+$\mbf{F}_{i}$, $\mbf{F}^{\prime}_{i}$ & Genotype-specific fertility matrices & $\omega \times \omega$ \\
+$\mbf{H}^{S}_{i}$, $\mbf{H}^{X}_{i}$ & Parent-offspring genotype maps & $g \times g$ \\
+$\mbf{W}$   & Allele segregation matrix & $\omega \times g$ \\
+$\mbf{K}$   & Vec-permutation matrix & $g \times g$ \\
+$\mbf{I}$   & Identity matrix & given in subscript \\
+\hline
+\end{tabular}
+\end{table}
+\newpage{}
 

### 2020-11-23 doc/SA-Hermaphrodites-wDemography.tex
Minor edits & new dataDeltaPolySpace data
Minor cleanup over weekend.

@@
 
-Here, we briefly derive a matrix model incorporating multiple life-cycle stages and a single bi-allelic locus under SA selection on male and female sex functions for a population of partially selfing simultaneous hermaphrodites.
+Here, we briefly derive a matrix model incorporating multiple life-cycle stages and a single diallelic locus under SA selection on male and female sex functions for a population of partially selfing simultaneous hermaphrodites.
 The derivation and some of the key results follow closely those presented in \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b}; and in fact, the model presented here reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019b} under obligate outcrossing.
@@
 The population state at time $t$ is described by a stage $\times$ genotype distribution matrix, which can be transformed into a population state vector, $\tilde{\mbf{n}}(t)$, which is ordered by genotype, then stage, then by how individuals were produced (self vs.~outcrossing).
-For a single locus with two alleles ($A$ and $a$), we have genotypes $g \in \{AA,\, Aa,\, aa\}$, giving the population state vector:
+For a single diallelic locus with alleles $A$ and $a$, we have three genotypes ($AA,\, Aa,\, aa$; $g = 3$), giving the population state vector:
 \begin{linenomath*}
@@
 
-\noindent where $\mbf{n}^{S}_{i}$ and $\mbf{n}^{X}_{i}$ (where $i \in \{AA,Aa,aa\}$) are the stage $\times$ genotype distribution vectors of individuals produced by self-fertilization and outcrossing, respectively.
+\noindent where $\mbf{n}^{S}_{i}$ and $\mbf{n}^{X}_{i}$ ($i \in \{AA,Aa,aa\}$) are the stage $\times$ genotype distribution vectors of individuals produced by self-fertilization and outcrossing, respectively.
 The proportional population vector is given by
@@
 
-The fertility matrices through male and female function are
+The fertility matrices through female and male function are
 \begin{linenomath*}
@@
 					& $AA$			& $Aa$ 					& $aa$ 		\\ \hline
-Female function:	& $f_{AA}$		& $f_{Aa} - h_f s_f$	& $f_{aa} - s_f$ \\	
-Male function:		& $f^{\prime}_{AA} - s_m$& $f^{\prime}_{Aa} - h_m s_m$	& $f^{\prime}_{aa}$ 		\\	
+Female function:	& $f_{AA}$		& $f_{Aa}(1 - h_f s_f)$	& $f_{aa}(1 - s_f)$ \\	
+Male function:		& $f^{\prime}_{AA}(1 - s_m)$& $f^{\prime}_{Aa}(1 - h_m s_m)$	& $f^{\prime}_{aa}$ 		\\	
 \hline

### 2020-11-23 run-Simulations.R

@@ -254,9 +254,9 @@ makeDataPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
 
 
 ##########################################################
-# Fig. 3 - Effects of inbreeding depression (no I.D.)
+# Fig. 3 - Dem. viable polymorphic param. space (w/ I.D.)
 
-makeDataDeltaPolyParamSpace(sMax=0.15, res=0.0015, precision = 1e-4,
+makeDataDeltaPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
 							om = 2, g = 3, theta = c(0.6,0.6,0.05,6.5), theta_prime = 6.5, 
 							hVals= c(1/2, 1/4), dStar = 0.8, 
 							tlimit = 10^5, eqThreshold = 1e-8)

### 2020-11-23 doc/SA-Hermaphrodites-wDemography.tex
Added Supp. DeltaPolySpace fig; editing quant.
This commit adds functional data cleanup funciton for deltaPolySpace
and a working function to create the supplementary figure.

Still need to fix quantDeltaPolySpace() function and associated
plotting figure

@@
 To keep the model analytically tractable, we assume our hypothetical species has two life stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active.
-Suppose that there is a genetic trade-off between the sex-functions at a single bi-allelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect (see Table \ref{tab:Fitness}).
-The allele does not affect survival and transition rates.
+Suppose that there is a genetic trade-off between the sex-functions at a single bi-allelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect.
+Following convention, we parameterize the relative fitnesses of each genotype through each sex function ($w_{i}$ and $w^{\prime}_{i}$) to be bounded by $(0,1)$, with dominance and selection coefficients $h_f, s_f$ and $h_m, s_m$ determining the decrease in fertility through each sex function relative to the most fit genotype ($AA$ has highest female fertility, $aa$ the highest male fertility; see Table \ref{tab:Fitness}).
+The SA locus does not affect survival and transition rates.
 However, the survival matrices can be used to model the fitness effects of inbreeding depression at later stages of development by allowing different stage-specific survival and transition rates for individuals produced by self-fertilization vs.~outcrossing.
@@
 					\begin{array}{cc}
-						0 & f_{i} \\
+						0 & f w_{i} \\
 						0 & 0
@@
 					\begin{array}{cc}
-						0 & f^{\prime}_{i} \\
+						0 & f^{\prime} w^{\prime}_{i} \\
 						0 & 0
@@
 
-\noindent where $f_{i}$ is the genotype specific adult fertility though female function, and $f^{\prime}_{i}$ is the same through male function (see Table \ref{tab:Fitness}).
+\noindent where $f$ and $f^{\prime}$ represent adult fertilities, and $w_{i}$ and $w^{\prime}_{i}$ the genotypic relative fitness expressions for female and male sex-functions (see Table \ref{tab:Fitness}).
 
@@
  \centering
- \caption{\bf Fitness expressions for Sexually Antagonistic selection in hermaphrodites}
+ \caption{\bf Fitness expressions for Sexually Antagonistic selection in hermaphrodites ()}
 \label{tab:Fitness}
@@
 					& $AA$			& $Aa$ 					& $aa$ 		\\ \hline
-Female function:	& $f_{AA}$		& $f_{Aa}(1 - h_f s_f)$	& $f_{aa}(1 - s_f)$ \\	
-Male function:		& $f^{\prime}_{AA}(1 - s_m)$& $f^{\prime}_{Aa}(1 - h_m s_m)$	& $f^{\prime}_{aa}$ 		\\	
+Female function ($w_{i}$):	& $1$		& $1 - h_f s_f$	& $1 - s_f$ \\	
+Male function ($w^{\prime}_{i}$):		& $1 - s_m$& $1 - h_m s_m$	& $1$ 		\\	
 \hline

### 2020-11-23 R/functions-Figs.R

@@ -456,7 +456,7 @@ layout     <- layout(layout.mat,respect=TRUE)
 
 
 #' Fig showing Proportion of polymorphic parameter space
-#' 
+#' (w/out inbreeding depression)
 polySpaceFigTitrate  <-  function(df = "dataPolySpaceFig_sMax0.15_res0.003_delta0_dj0_da0_dg0") {
 
     # Make filenames for import from df names
@@ -609,6 +609,147 @@ polySpaceFigTitrate  <-  function(df = "dataPolySpaceFig_sMax0.15_res0.003_delta
 }
 
 
+#' Fig showing Effect of Inbreeding Depression on 
+#' Proportion of polymorphic parameter space
+
+deltaSelfingLoadPolySpaceFigTitrate  <-  function(df = "dataDeltaPolySpaceFig_sMax0.15_res0.003_dStar0.8_f6.5") {
+
+   # Make filenames for import from df names
+    fName  <-  paste('./output/simData/', df, '.csv', sep="")
+
+    # import data
+    data  <-  read.csv(file=fName, header=TRUE)
+
+    # Extract plotting parameter values from df names
+    d1   <-  strsplit(df, '_')[[1]][c(2:5)]
+    pars  <-  list(
+                    "sMax"  =  as.numeric(strsplit(d1[1],'x')[[1]][2]),
+                    "res"   =  as.numeric(strsplit(d1[2],'s')[[1]][2]),
+                    "dStar" =  as.numeric(strsplit(d1[3],'r')[[1]][2]),
+                    "f"     =  as.numeric(strsplit(d1[4],'f')[[1]][2])
+                    )
+    hLev  <-  unique(data$h)
+    dLev  <-  unique(data$Delta)
+    CLev  <-  unique(data$C)
+    nHs   <-  length(hLev)
+    nDs   <-  length(dLev)
+    nCs   <-  length(CLev)
+
+    # clean data set & quantify parameter space
+    dat   <-  quantDeltaPolySpace(data = data, pars = pars)
+
+
+# Color scheme
+    COLS  <-  list(
+                    "PG"     =  transparentColor('#252525', opacity=1),
+                    "dSim"   =  transparentColor('#252525', opacity=0.6),
+                    "d_j"    =  transparentColor('dodgerblue4', opacity=0.6),
+                    "d_a"    =  transparentColor('dodgerblue', opacity=0.6),
[…]

### 2020-11-23 R/functions-Simulations.R

@@ -2399,7 +2399,7 @@ makeDataDeltaPolyParamSpace  <-  function(sMax=0.15, res=0.0015, precision = 1e-
 											  tlimit = tlimit, eqThreshold = eqThreshold, verbose = FALSE, writeFile=FALSE)
 
 			# Append new data to dataSet
-			dataSet  <-  rbind(dataSet, cbind(rep("delta", times=nrow(invData)),
+			dataSet  <-  rbind(dataSet, cbind(rep("d_j", times=nrow(invData)),
 											  rep(hVals[i], times=nrow(invData)),
 											  rep(deltaSeq[j], times=nrow(invData)),
 											  rep(Cs[j], times=nrow(invData)),
@@ -2431,7 +2431,7 @@ makeDataDeltaPolyParamSpace  <-  function(sMax=0.15, res=0.0015, precision = 1e-
 											  tlimit = tlimit, eqThreshold = eqThreshold, verbose = FALSE, writeFile=FALSE)
 
 			# Append new data to dataSet
-			dataSet  <-  rbind(dataSet, cbind(rep("delta", times=nrow(invData)),
+			dataSet  <-  rbind(dataSet, cbind(rep("d_a", times=nrow(invData)),
 											  rep(hVals[i], times=nrow(invData)),
 											  rep(deltaSeq[j], times=nrow(invData)),
 											  rep(Cs[j], times=nrow(invData)),
@@ -2463,7 +2463,7 @@ makeDataDeltaPolyParamSpace  <-  function(sMax=0.15, res=0.0015, precision = 1e-
 											  tlimit = tlimit, eqThreshold = eqThreshold, verbose = FALSE, writeFile=FALSE)
 
 			# Append new data to dataSet
-			dataSet  <-  rbind(dataSet, cbind(rep("delta", times=nrow(invData)),
+			dataSet  <-  rbind(dataSet, cbind(rep("d_g", times=nrow(invData)),
 											  rep(hVals[i], times=nrow(invData)),
 											  rep(deltaSeq[j], times=nrow(invData)),
 											  rep(Cs[j], times=nrow(invData)),
@@ -2480,10 +2480,178 @@ makeDataDeltaPolyParamSpace  <-  function(sMax=0.15, res=0.0015, precision = 1e-
 	colnames(dataSet)  <-  c("Delta", "h", "deltaVal", "C", "smInv", "aInv", "AInv", "sf", "smExt", "sm", "sfExt")
 
[…]

### 2020-11-23 makeFigs.R

@@ -44,6 +44,11 @@ toPdf(suppPolySpaceThresholdFigs(),
 embed_fonts(figPath(name='SuppFig-polySpace.pdf'))
 
 
+toPdf(suppDeltaPolySpaceThresholdFigs(), 
+			figPath(name='SuppFig-deltaPolySpace.pdf'), width=40, height=20)
+embed_fonts(figPath(name='SuppFig-deltaPolySpace.pdf'))
+
+
 ###############
 # PRELIM FIGS
 ###############

### 2020-11-25 doc/SA-Hermaphrodites-wDemography.tex
Methods thru Analyses subsection
More writing, minor tweaks to figs.

@@
 \begin{linenomath*}
-\begin{equation}
+\begin{equation} \label{eq:PopStateVec}
 	\tilde{\mbf{n}}(t) =  \left[
@@
 \begin{linenomath*}
-\begin{equation} \label{eq:US}
+\begin{equation} \label{eq:BlkUS}
 	\mbb{U}^S = 
@@
 \begin{linenomath*}
-\begin{equation} \label{eq:FS}
+\begin{equation} \label{eq:BlkFS}
 	\mcal{F}^S = \mbf{K}^{\intercal} \mbb{H}^S(\tilde{\mbf{p}}) \mbf{K} \mbb{F}^S = 
@@
 \begin{linenomath*}
-\begin{equation} \label{eq:FX}
+\begin{equation} \label{eq:BlkFX}
 	\mcal{F}^X = \mbf{K}^{\intercal} \mbb{H}^X(\tilde{\mbf{p}}) \mbf{K} \mbb{F}^X = 
@@
 
-\noindent where $\sigma_J$ and $\sigma_A$ are the juvenile and adult stage survival rates, and $\gamma$ is the maturation rate from juvenile to adult stages.
-The corresponding terms $\delta_S$, $\delta_A$, and $\delta_{\gamma}$ denote the proportional decreases in stage-specific survival and transition rates due to inbreeding depression (i.e., deleterious effects of inbreeding at later life-history stages; e.g., \citealt{HarderRoutely2006}).
+\noindent where $\sigma_J$ and $\sigma_A$ are the juvenile and adult stage survival rates, $\gamma$ is the maturation rate from juvenile to adult stages, and the corresponding $\delta_S$, $\delta_A$, and $\delta_{\gamma}$ terms denote the proportional decreases in stage-specific survival and transition rates due to inbreeding depression (i.e., deleterious effects of inbreeding at later life-history stages; e.g., \citealt{HarderRoutely2006}).
 Any or all of these parameters may differ among genotypes, so that selection can operate on stage-specific viability, development, and/or fertility.
@@
 
-Iterating the projection matrix, Eq(\ref{eq:Atilde}) with the above demographic matrices provides numerical simulation of the eco-evolutionary dynamics for selection operating on any of the sex-function, or stage-specific demographic parameters.
-As we outline below, we use both numerical simulations together with mathematical analyses to study the conditions for the maintenance of SA polymorphisms, and the demographic fate of the populations (i.e., positive growth, or extinction).
+Iterating the projection matrix, Eq(\ref{eq:Atilde}) with the above demographic matrices, given an initial population state vector, Eq(\ref{eq:PopStateVec}), allows numerical simulation of the eco-evolutionary dynamics for selection operating on any of the sex-function, or stage-specific demographic parameters.
+As we outline below, we use numerical techniques together with mathematical analyses to study the conditions for the maintenance of SA polymorphisms, and the demographic fate of the populations (i.e., positive growth, or extinction).
 
@@
 
-A variety of eco-evolutionary outcomes are possible in the stage-structured mendelian matrix model, including fixation of one or the other allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and evolutionary suicide (e.g., see \citealt{CaswellEtAl2018,deVriesCaswell2019a,deVriesCaswell2019b}.
-We focus on identifying the necessary conditions for protected SA polymorphism with a positive intrinsic population growth rate, which we refer to as demographically viable SA polymorphisms.
-....
+Diverse eco-evolutionary outcomes are possible in the above model, including fixation of either allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and suicide (e.g., see \citealt{CaswellEtAl2018,deVriesCaswell2019a,deVriesCaswell2019b}).
+We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection and ({\itshape ii}) the intrinsic population growth rate at demographic and genotypic equilibrium is positive; situations that we refer to as 'demographically viable SA polymorphism'.
 
-\paragraph*{Fourth-order heading}
-\ldots
+We identify conditions where SA polymorphism is 'protected' by evaluating the stability of populations initially fixed for either SA allele to invasion by the other (i.e., stability was assessed at the boundary equilibrium allele frequencies of $\hat{\mbf{p}}_{AA} = 1$ and $\hat{\mbf{p}}_{aa} = 1$; \citealt{Levene1953, Prout1968, deVriesCaswell2019b}).
+The formal conditions for a protected polymorphism are determined by linearizing the model in the vacinity of the boundary equilibria ($\hat{\mbf{p}}$), and evaluating the magnitude of the largest eigenvalue of the Jacobian matrix of the linearization, which we denote $\tilde{\zeta}_{AA}$ and $\tilde{\zeta}_{aa}$.
+For our model, $\mbf{M}$ is upper block triangular, and we can therefore express the stability of $\hat{\mbf{p}}$ in terms of the leading eigenvalue of the diagonal block corresponding to changes in heterozygote frequency ($\mbf{M}_{22}$; e.g., see \citealt{OttoDay2007}).
+For a polymorphism to be 'protected', both boundaries must be unstable (i.e., $\tilde{\zeta}_{AA} > 1$ and $\tilde{\zeta}_{aa} > 1$), which requires that
+\begin{linenomath*}
+\begin{equation}\label{eq:PolyCond}
+	1 < \frac{\rho(\mbf{M}_{i})}{\lambda_{i}}
+\end{equation}
+\end{linenomath*}
+
+\noindent is satisfied, where $\rho(\cdot)$ represents the spectral radius, and $\lambda_i$ is the intrinsic growth rate for the homozygous genotype being invaded at each boundary ($i \in (AA,aa)$).
+Full details for the invasion analysis are given in \hl{Appendix A.x}.
+
+To determine whether a protected SA polymorphism was also demographically viable, we used numerical simulation techniques.
+Specifically, for each boundary equilibrium we introduced the rare allele at low initial frequency, and iterated Eq(\ref{eq:Atilde}) until the population had reached demographic and genotypic equilibrium.
+In our model, the population state vector will grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see \citealt{Caswell2001}), and the intrinsic population growth rate, $\lambda$, is easily calculated as $\tilde{\mbf{n}}(t)/\tilde{\mbf{n}}(t-1)$, where $t$ is the final generation of the simulation.
+We note, however, that if the ecological component of the model is non-linear, more exotic dynamics are possible.
 
@@
  \includegraphics[width=\linewidth]{../output/figs/extinctionThresholdsFig.pdf}
- \caption{Illustration of parameter space for SA polymorphism and extinction thresholds predicted by the mendelian stage-structured matrix model.
+ \caption{\footnotesize{Illustration of parameter space for SA polymorphism and extinction thresholds predicted by the mendelian stage-structured matrix model.
 Balanced SA polymorphisms can be maintained in the funnel-shaped region between the invasion conditions for each SA allele (dark solid lines).
@@
 We are particularly interested in identifying and quantifying "demographically viable polymorphic parameter space", which corresponds to the area inside the funnel that is also to the left of the extinction threshold for a given fertility value.
-Results are shown for three different population selfing rates ($C = \{0,\,1/4,\,1/2\}$), and two dominance scenarios (additivity with $h = 1/2$, and dominance reversal with $h = 1/4$); extinction thresholds are illustrated for three different values of female fecundity ($f$ values annotated on each panel).
-}
+Results are shown for three different population selfing rates ($C = \{0,\,1/4,\,1/2\}$), and two dominance scenarios (additivity with $h = 1/2$, and dominance reversal with $h = 1/4$); extinction thresholds are illustrated for three different values of female fecundity ($f$ values annotated on each panel).}}
  \label{fig:extThresholds}
@@
  \includegraphics[scale=0.75]{../output/figs/polymorphicSpaceTitrate.pdf}
- \caption{Proportion of demographically viable parameter space in the absence of inbreeding depression (i.e., assuming $\delta = \delta_i = 0$, where $i \in \{j,a,\gamma\}$) predicted by the matrix model, plotted as a function of the population selfing rate.
+ \caption{\footnotesize{Proportion of demographically viable parameter space (out of total $s_f \times s_m$ space with $\max(s) = 0.15$) in the absence of inbreeding depression (i.e., assuming $\delta = \delta_i = 0$, where $i \in \{j,a,\gamma\}$), plotted as a function of the population selfing rate.
 Results are shown for three fertility values corresponding to low, medium, and high fertility (blue, green, and red points respectively) under additive (panel A, $h = 1/2$), and partially recessive (panel B, $h = 1/4$) SA fitness effects.
-Each point was calculated by approximate numerical integration of the corresponding SA invasion conditions and extinction threshold predicted by the mendelian matrix model (see \hyperref[sec:analyses]{Analyses} section), while solid lines were produced by numerically integrating the analytic expressions for the single-locus invasion conditions from the population genetic models of \citet{JordanConnallon2014} and \citet{Olito2017} (solid black lines).
-}
+Each point was calculated by numerical integration of the corresponding SA invasion conditions and extinction threshold predicted by the mendelian matrix model (see \hyperref[sec:analyses]{Analyses} section), while solid lines were produced by numerically integrating the analytic expressions for the single-locus invasion conditions from the population genetic models of \citet{JordanConnallon2014} and \citet{Olito2017} (solid black lines).}}
  \label{fig:polySpace}
@@
 
+ \begin{figure}[htbp]
+ \centering
+ \includegraphics[scale=0.75]{../output/figs/deltaPolymorphicSpaceTitrate.pdf}
+ \caption{\footnotesize{Effects of early- and late-acting inbreeding depression on the proportion of demographically viable parameter space (out of total $s_f \times s_m$ space with $\max(s) = 0.15$), plotted as a function of the population selfing rate.
+In all plots, the strength of inbreeding depression decreases as the selfing rate goes up following a simple model of purging recessive deleterious mutations (see \hl{Appendix X.x} for details).
+Only single, $\delta_i$ terms indicated in the legend ($i \in \{j,a,\gamma\}$) are allowed to vary at once (all others are set to $0$).
+Effects of early-acting inbreeding depression are shown on the left (panels A, C; $\delta$) and for late-acting inbreeding depression on the right (panels B, D; $\delta_i$) for additive ($h = 1/2$) and partially recessive ($h = 1/4$) SA fitness effects, respectively.
+All results used the same 'high female fertility' value of $f = 6.5$ (see Fig.\ref{fig:polySpace}).
+Each point was calculated by numerically integrating the corresponding SA invasion conditions and extinction threshold predicted by the model (see \hyperref[sec:analyses]{Analyses} section), while solid lines were produced by numerically integrating the analytic expressions for the single-locus invasion conditions from the population genetic models of \citet{JordanConnallon2014} and \citet{Olito2017} (solid black lines).
+The population genetic model predictions for varying $delta$ are plotted on both left- and right-hand panels for comparison, and to highlight the different y-axes scales.}}
+ \label{fig:deltaPolySpace}
+ \end{figure}
+
+inbreeding depression that follows d p d*(1 2 C=2) (see “Models”). Plots illustrate the increase in equilibrium frequencies
 %%%%%%%%%%%%%%%%%%%%%%%%

### 2020-11-25 R/functions-Figs.R

@@ -527,7 +527,7 @@ polySpaceFigTitrate  <-  function(df = "dataPolySpaceFig_sMax0.15_res0.003_delta
         proportionalLabel(0.03, 1.075, 'A', cex=1.2, adj=c(0.5, 0.5), xpd=NA)
 #        proportionalLabel(0.5, 1.1, substitute(paste(italic(h), " = ", hh), list(hh = pars2$hf)), cex=1.2, adj=c(0.5, 0.5), xpd=NA, srt=0)
         proportionalLabel(0.5, 1.1, expression(paste(italic(h), " = ", 1/2)), cex=1.2, adj=c(0.5, 0.5), xpd=NA, srt=0)
-        proportionalLabel(-0.3, 0.5, expression(paste("Proportion of parameter space")), cex=1, adj=c(0.5, 0.5), xpd=NA, srt=90)
+        proportionalLabel(-0.3, 0.5, expression(paste("Prop. viable polymorphic space")), cex=1, adj=c(0.5, 0.5), xpd=NA, srt=90)
 
         #Legend
         legend( x       =  usr[2],
@@ -579,7 +579,7 @@ polySpaceFigTitrate  <-  function(df = "dataPolySpaceFig_sMax0.15_res0.003_delta
 #        proportionalLabel(0.5, 1.1, substitute(paste(italic(h), " = ", hh), list(hh = pars5$hf)), cex=1.2, adj=c(0.5, 0.5), xpd=NA, srt=0)
         proportionalLabel(0.5, 1.1, expression(paste(italic(h), " = ", 1/4)), cex=1.2, adj=c(0.5, 0.5), xpd=NA, srt=0)
         proportionalLabel(0.5, -0.3, expression(paste("Selfing rate (",italic(C), ")")), cex=1.2, adj=c(0.5, 0.5), xpd=NA, srt=0)
-        proportionalLabel(-0.3, 0.5, expression(paste("Proportion of parameter space")), cex=1, adj=c(0.5, 0.5), xpd=NA, srt=90)
+        proportionalLabel(-0.3, 0.5, expression(paste("Prop. viable polymorphic space")), cex=1, adj=c(0.5, 0.5), xpd=NA, srt=90)
 
         #Legend
         legend( x       =  usr[2],
@@ -685,7 +685,7 @@ deltaSelfingLoadPolySpaceFigTitrate  <-  function(df = "dataDeltaPolySpaceFig_sM
[…]

### 2020-11-25 R/functions-Simulations.R

@@ -1882,7 +1882,7 @@ findInvBoundZeta  <-  function(om = 2, g = 3, theta = c(0.6, 0.6, 0.05, 6), thet
 				(1 - C)*c(c(100-(om-1), rep(1,times=(om-1))), rep(0,times = 2*om)))
 	n0aa  <-  c(      C*c(rep(0,times = 2*om), c(100-(om-1), rep(1,times=(om-1)))), 
 				(1 - C)*c(rep(0,times = 2*om), c(100-(om-1), rep(1,times=(om-1)))))
-browser()
+#browser()
 	# Simulate to demographic equilibrium for each boundary
 	AAEq  <-  fwdDyn2Eq(nzero=n0AA, om=om, g=g, W_prime=W_prime, blkFX_prime=blkFX_prime, blkUS=blkUS, blkUX=blkUX, W=W, Iom=Iom, Ig=Ig, K=K, Z=Z, blkFS=blkFS, blkFX=blkFX, tlimit=10^2, eqThreshold=eqThreshold)
 	aaEq  <-  fwdDyn2Eq(nzero=n0aa, om=om, g=g, W_prime=W_prime, blkFX_prime=blkFX_prime, blkUS=blkUS, blkUX=blkUX, W=W, Iom=Iom, Ig=Ig, K=K, Z=Z, blkFS=blkFS, blkFX=blkFX, tlimit=10^2, eqThreshold=eqThreshold)

### 2020-11-26 doc/SA-Hermaphrodites-wDemography.tex
Edits to Intro, Methods:Analyses
More writing...

@@
 		\end{enumerate}
+		\item A major strength of the model is that we can use empirically estimated demographic rates to parameterize the model, enabling us to make predictions about the scope for demographically viable SA polymorphisms that are grounded in the biology of real populations.
 \end{enumerate}
@@
 The model also allows for the effects of inbreeding depression to manifest in different life-history stages, enabling us to explore the fitness consequences of early- vs.~late-acting inbreeding depression on the model predictions.
+Finally, we give an illustrative example of the potential for SA polymorphism in real populations by parameterizing our model with empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus} (now {\itshape Erythranthe guttata}).
 
@@
 Under some conditions, self-fertilization can alleviate the demographic costs of balanced SA polymorphisms, but the concommitant effects of inbreeding depression greatly exacerbate them.
+Interestingly, our illustrative example using {\itshape M.~guttatus} appears to show that demographic rates observed in real populations can sustain large regions of viable SA polymorphic space.
 Overall, our findings provide a more nuanced picture of the nature of SA genetic variation that we should expect to find in natural populations, where the fate of SA alleles and the populations harboring them is determined jointly by evolutionary and demographic processes.
@@
 
-Diverse eco-evolutionary outcomes are possible in the above model, including fixation of either allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and suicide (e.g., see \citealt{CaswellEtAl2018,deVriesCaswell2019a,deVriesCaswell2019b}).
-We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection and ({\itshape ii}) the intrinsic population growth rate at demographic and genotypic equilibrium is positive; situations that we refer to as 'demographically viable SA polymorphism'.
+Diverse eco-evolutionary outcomes are possible in the model, including fixation of either allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and suicide (e.g., see \citealt{CaswellEtAl2018,deVriesCaswell2019a,deVriesCaswell2019b}).
+We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection and ({\itshape ii}) the intrinsic population growth rate at equilibrium is positive; a situation that we refer to as a 'demographically viable SA polymorphism'.
 
@@
 The formal conditions for a protected polymorphism are determined by linearizing the model in the vacinity of the boundary equilibria ($\hat{\mbf{p}}$), and evaluating the magnitude of the largest eigenvalue of the Jacobian matrix of the linearization, which we denote $\tilde{\zeta}_{AA}$ and $\tilde{\zeta}_{aa}$.
-For our model, $\mbf{M}$ is upper block triangular, and we can therefore express the stability of $\hat{\mbf{p}}$ in terms of the leading eigenvalue of the diagonal block corresponding to changes in heterozygote frequency ($\mbf{M}_{22}$; e.g., see \citealt{OttoDay2007}).
-For a polymorphism to be 'protected', both boundaries must be unstable (i.e., $\tilde{\zeta}_{AA} > 1$ and $\tilde{\zeta}_{aa} > 1$), which requires that
+For our model, $\mbf{M}$ has an upper block triangular form, and we can therefore express the stability of $\hat{\mbf{p}}$ in terms of the leading eigenvalue of the diagonal block corresponding to changes in heterozygote frequency ($\mbf{M}_{22}$; e.g., see \citealt{OttoDay2007,deVriesCaswell2019b}).
+For a polymorphism to be 'protected', both boundary equilibria must be unstable (i.e., $\tilde{\zeta}_{AA} > 1$ and $\tilde{\zeta}_{aa} > 1$), which requires that
 \begin{linenomath*}
 \begin{equation}\label{eq:PolyCond}
-	1 < \frac{\rho(\mbf{M}_{i})}{\lambda_{i}}
+	1 < \frac{\rho(\mbf{M}_{22})}{\lambda_{i}}
 \end{equation}
@@
 
-\noindent is satisfied, where $\rho(\cdot)$ represents the spectral radius, and $\lambda_i$ is the intrinsic growth rate for the homozygous genotype being invaded at each boundary ($i \in (AA,aa)$).
-Full details for the invasion analysis are given in \hl{Appendix A.x}.
+\noindent is satisfied, where $\rho(\cdot)$ represents the spectral radius, and $\lambda_i$ is the intrinsic growth rate for the homozygous genotype being invaded at each boundary ($i \in (AA,aa)$; see Box.~1).
+A full derivation of the Jacobian and details of the invasion analysis are provided in \hl{Online Appendix X.x}.
 
-To determine whether a protected SA polymorphism was also demographically viable, we used numerical simulation techniques.
+We used numerical simulation to determine whether a protected SA polymorphism was also demographically viable.
 Specifically, for each boundary equilibrium we introduced the rare allele at low initial frequency, and iterated Eq(\ref{eq:Atilde}) until the population had reached demographic and genotypic equilibrium.
-In our model, the population state vector will grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see \citealt{Caswell2001}), and the intrinsic population growth rate, $\lambda$, is easily calculated as $\tilde{\mbf{n}}(t)/\tilde{\mbf{n}}(t-1)$, where $t$ is the final generation of the simulation.
+In our model, the population state vector will grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see \citealt{Caswell2001}), and the intrinsic population growth rate, $\lambda$, can be calculated as $\tilde{\mbf{n}}(t)/\tilde{\mbf{n}}(t-1)$, where $t$ is the final generation of the simulation.
 We note, however, that if the ecological component of the model is non-linear, more exotic dynamics are possible.
+
+Unfortunately, closed-form solutions for the invasion conditions for each allele and the equilibrium frequencies under different selection scenarios are analytically intractable except for very restricted conditions (see \hl{Online Appendix X}).
+We therefore used numerical techniques to identify invasion and extinction thresholds, which could then be used to define regions of demographically viable polymorphism across $s_f \times s_m$ parameter space.
+Because single-locus selection coefficients are generally weak, we limit our analyses to coefficients with within the open interval $\left(0,0.15\right)$, unless stated otherwise.
+For a given parameter set, we numerically evaluated Eq(\ref{eq:PolyCond}) and $\lambda$, starting at the outer boundaries of selection parameter space, and titrated inward to identify thresholds where $\tilde{\zeta}_i \approx 1$ and $\lambda \approx 1$ (with precision of $10^{-5}$).
+The proportions of total $s_f \times s_m$ parameter space where SA polymorphism and/or extinction occurred were calculated by numerically integrating the appropriate regions defined by the invasion and extinction thresholds
 
@@
 \ldots
+
+%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Polymorphism and Extinction} \label{sec:Res:PolyExt}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Polymorphism in the absence of inbreeding depression} \label{sec:Res:PolyNoID}
+
  \begin{figure}[htbp]
@@
  \caption{\footnotesize{Proportion of demographically viable parameter space (out of total $s_f \times s_m$ space with $\max(s) = 0.15$) in the absence of inbreeding depression (i.e., assuming $\delta = \delta_i = 0$, where $i \in \{j,a,\gamma\}$), plotted as a function of the population selfing rate.
-Results are shown for three fertility values corresponding to low, medium, and high fertility (blue, green, and red points respectively) under additive (panel A, $h = 1/2$), and partially recessive (panel B, $h = 1/4$) SA fitness effects.
+Results are shown for three fertility values corresponding to low, medium, and high fertility (blue, green, and red points respectively) under additive (panel A; $h = 1/2$), and partially recessive (panel B; $h = 1/4$) SA fitness effects.
 Each point was calculated by numerical integration of the corresponding SA invasion conditions and extinction threshold predicted by the mendelian matrix model (see \hyperref[sec:analyses]{Analyses} section), while solid lines were produced by numerically integrating the analytic expressions for the single-locus invasion conditions from the population genetic models of \citet{JordanConnallon2014} and \citet{Olito2017} (solid black lines).}}
@@
  \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Demographic effects of inbreeding depression} \label{sec:Res:InbreedEffects}
 
@@
  \end{figure}
-
-inbreeding depression that follows d p d*(1 2 C=2) (see “Models”). Plots illustrate the increase in equilibrium frequencies
-%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{The height of the jump}
-
-\ldots
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{The laziness of the dog}
-
-\ldots
 
@@
 We are particularly grateful to the European Society for Evolutionary Biology (ESEB) for funding the Special Topics Network workshops “Linking local adaptation with the evolution of sex differences”, and to the participants, all of whom made this study possible.
-This study was also supported by a Wenner-Gren Foundation postdoctoral stipen to C.O., and [Lotte's funding].
+We would also like to thank ....
+This work was supported by a Wenner-Gren Foundation postdoctoral stipend to C.O., and [Lotte's funding] to C.dV..
 

### 2020-11-26 R/functions-Figs.R

@@ -3,7 +3,7 @@
 ###############
 library(extrafont)
 library(fontcm)
-loadfonts()
+loadfonts(quiet=TRUE)
 library(plyr)
 library(lattice)
 library(latticeExtra)

### 2020-11-30 doc/SA-Hermaphrodites-wDemography.tex
Current edits to ms; Handover Intro to Lotte
Just pushing edits for the current state of the ms, along with minor
changes to the makeDeltaPolySpaceData() function so the output has \
a different filename.

@@
 Conflicting selection and gene-flow between different classes of individuals, or genetic trade-offs between fitness components, can impose constraints on adaptation with particularly interesting evolutionary consequences \citep{CharlesworthHughes2000, ConnallonHall2018}.
-Such constraints hinder adaptation by preventing individuals (or classes of individuals) from reaching the phenotypic optimum in one or more fitness components, thereby increasing the population's overall extinction risk.
-At the same time, they provide an effective mechanism for the maintenance of genetic variation, which in turn influences the population's capactiy for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
-Hence, for genetically constrained traits the nature and extent of genetic variation observed in natural populations should reflect the balance between the contribution of constraints to the maintainance of genetic polymorphisms, and the demographic consequences of the resulting maladaptation.
+On the one hand, such constraints hinder adaptation by preventing individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase the population's overall extinction risk.
+On the other hand, they provide an effective mechanism for the maintenance of genetic variation, which influences the population's capactiy for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
+Hence, for traits under conflicting selection, the nature and extent of genetic variation observed in natural populations should reflect a balance between the contribution of genetic constraints to the maintainance of genetic polymorphisms, and the demographic consequences of the resulting maladaptation.
 %Genetic trade-offs between fitness components due to pleiotropy represent a form of genetic constraint with particularly interesting evolutionary consequences because they act as a double-edged sword
@@
 When such genetic trade-offs occur, "sexually antagonistic" selection (abbreviated SA hereafter) can arise when beneficial alleles for one sex are deleterious when expressed in the other \citep{Kidwell1977, Rice1992, ConnallonClark2012}.
-Moreover, alleles with opposing fitness effects through male and female sex functions create analogous genetic constraints on fitness in hermaphrodite populations, where both maternal and paternal reproductive success contribute jointly to each individuals' overall fitness \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
+This, alleles with opposing fitness effects through male and female sex functions create analogous genetic constraints on fitness in hermaphrodite populations, where both maternal and paternal reproductive success contribute jointly to each individuals' overall fitness \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
 
@@
 
-In this article, we build upon the models of \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b} to develop a stage-structured mendelian matrix model that accommodates selection through male and female function in simultaneous hermaphrodites with mixed mating systems.
+In this article, we develop a stage-structured mendelian matrix model that accommodates selection through male and female sex-functions in populations of simultaneous hermaphrodites with mixed mating systems.
+Under obligate outcrossing, the model becomes mathematically identical to the two-sex model of \citet{deVriesCaswell2019b}.
 Using the model, we jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations.
-Although a variety of interesting eco-evolutionary dynamics are possible in the model (e.g., \citealt{deVriesCaswell2019a, deVriesCaswell2019b,CaswellEtAl2018}), we focus our analyses on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population is also demographically viable (i.e., has a positive intrinsic growth rate).
+We focus our analyses on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population is also demographically viable (i.e., has a positive intrinsic growth rate).
 The model also allows for the effects of inbreeding depression to manifest in different life-history stages, enabling us to explore the fitness consequences of early- vs.~late-acting inbreeding depression on the model predictions.
@@
 
-A key finding from our analysis is that when the intrinsic population growth rate is close to $1$, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space for SA polymorphism.
-Furthermore, this demographically viable polymorphic parameter space is often biased towards 'female-beneficial/male-deleterious' SA alleles (i.e., those with weaker selection on the female- than male sex function).
-Under some conditions, self-fertilization can alleviate the demographic costs of balanced SA polymorphisms, but the concommitant effects of inbreeding depression greatly exacerbate them.
-Interestingly, our illustrative example using {\itshape M.~guttatus} appears to show that demographic rates observed in real populations can sustain large regions of viable SA polymorphic space.
+A key finding of our study is that when the intrinsic population growth rate is close to $1$, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space for SA polymorphism.
+Furthermore, this demographically viable polymorphic parameter space is often biased towards alleles with stronger selection through the female- than male sex function.
+Self-fertilization can alleviate the demographic costs of balanced SA polymorphisms under some conditions, however, the concommitant effects of inbreeding depression generally exacerbate them in populations with mixed-mating systems.
+Despite these demographic pitfalls, our example using {\itshape M.~guttatus} appears to show that demographic rates observed in real populations are capable of sustaining large regions of viable SA polymorphic space.
 Overall, our findings provide a more nuanced picture of the nature of SA genetic variation that we should expect to find in natural populations, where the fate of SA alleles and the populations harboring them is determined jointly by evolutionary and demographic processes.
@@
 
-Here, we briefly derive a matrix model incorporating multiple life-cycle stages and a single diallelic locus under SA selection on male and female sex functions for a population of partially selfing simultaneous hermaphrodites.
+Here, we briefly describe a matrix model incorporating multiple life-cycle stages and a single diallelic locus under SA selection on male and female fertility for populations of simultaneous hermaphrodites with outcrossing or mixed-mating systems.
 The derivation and some of the key results follow closely those presented in \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b}; and in fact, the model presented here reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019b} under obligate outcrossing.
@@
 
-Let $C$ denote the proportion of each individual's ovules that are self-fertilized (the remaining $1 - C$ are outcrossed), and $\delta$ represents the proportion of self-fertilized zygotes that fail to develop due to inbreeding depression during early development \citep{Charlesworth1987}.
+Let $C$ denote the proportion of each individual's ovules that are self-fertilized (the remaining $1 - C$ are outcrossed), and let $\delta$ represent the proportion of self-fertilized zygotes that fail to develop due to inbreeding depression during early development \citep{Charlesworth1987}.
 In \hl{Appendix A.x}, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
-Outcrossing proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}, except each individual's genotype determines both the number of ovules produced, and also paternal "mating success", broadly defined.
-For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986,WangBarrett2020,Harder2016}.
+Outcrossing proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b} except that each individual's genotype determines both the number of ovules produced and also paternal "mating success", broadly defined.
+For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986, WangBarrett2020, Harder2016}.
 Depending on their stage and genotype at the SA locus, individuals contribute male gametes to the overall population gamete pool according to the male stage $\times$ genotype vector:
@@
 As in previous models of partial selfing, we assume that individuals produce enough male gametes to easily self-fertilize all of their ovules, and that self-fertilization involves little or no selection from external factors relative to outcrossing (e.g., \citealt{Charlesworth1978a,JordanConnallon2014,Olito2017}, but see \citealt{Tazzyman2015}).
-Under these assumptions, the genotype distributions of selfed offspring are determined entirely by the parental genotype and the probabilities of segregation and fertilization during meiosis:
+Under these assumptions, the genotype distributions of selfed offspring are determined entirely by the parental genotype and the probabilities of segregation and fertilization during and after meiosis:
 \begin{linenomath*}
@@
 
-Using the component matrices described in the previous section, the matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ that projects the eco-evolutionary dynamics is:
+Using the component matrices described above, the matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ that projects the eco-evolutionary dynamics is:
 \begin{linenomath*}
@@
 
-We now construct and analyze a genotype $\times$stage-classified model for a hypothetical species with intralocus sexual conflict via the two sex functions.
+We now construct and analyze a genotype $\times$ stage-classified model for a hypothetical species with intralocus sexual conflict via the two sex functions.
 To keep the model analytically tractable, we assume our hypothetical species has two life stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active.
@@
 					\begin{array}{cc}
-						\sigma^S_J (1 - \delta_J) \big(1 - \gamma^S (1 - \delta_{\gamma}) \big) & 0 \\
-						\sigma^S_J (1 - \delta_J) \gamma (1 - \delta_{\gamma})      & \sigma^S_A (1 - \delta_S)
+						\sigma_j (1 - \delta_j) \big(1 - \gamma (1 - \delta_{\gamma}) \big) & 0 \\
+						\sigma_j (1 - \delta_j) \gamma (1 - \delta_{\gamma})      & \sigma_a (1 - \delta_a)
 					\end{array}
@@
 					\begin{array}{cc}
-						\sigma^X_J(1 - \gamma^X) & 0 \\
-						\sigma^X_J \gamma      & \sigma^X_A
+						\sigma_j(1 - \gamma) & 0 \\
+						\sigma_j \gamma      & \sigma_A
 					\end{array}
@@
 
-\noindent where $\sigma_J$ and $\sigma_A$ are the juvenile and adult stage survival rates, $\gamma$ is the maturation rate from juvenile to adult stages, and the corresponding $\delta_S$, $\delta_A$, and $\delta_{\gamma}$ terms denote the proportional decreases in stage-specific survival and transition rates due to inbreeding depression (i.e., deleterious effects of inbreeding at later life-history stages; e.g., \citealt{HarderRoutely2006}).
+\noindent where $\sigma_j$ and $\sigma_a$ are the juvenile and adult stage survival rates, $\gamma$ is the maturation rate from juvenile to adult stages, and the corresponding $\delta_j$, $\delta_a$, and $\delta_{\gamma}$ terms denote the proportional decreases in stage-specific survival and transition rates due to inbreeding depression (i.e., deleterious effects of inbreeding at later life-history stages; e.g., \citealt{HarderRoutely2006}).
 Any or all of these parameters may differ among genotypes, so that selection can operate on stage-specific viability, development, and/or fertility.
@@
 
+Unless stated otherwise, we use the following parameter values for the demogrphic rates in the model: $\sigma_j = \sigma_A = 0.6$ and $\gamma = 0.05$, which are similar to those used in \citet{deVriesCaswell2019b}.
+Our parameters of interest include $f$, $\delta$, $\delta_j$, $\delta_a$, $\delta_{\gamma}$, and the selection parameters $h_f$, $s_f$, $h_m$, and $s_m$, which are given different values for each analyses as described in the figure captions.
+
 %%%%%%%%%%%%%%%%%%%%%%
@@
 
-We identify conditions where SA polymorphism is 'protected' by evaluating the stability of populations initially fixed for either SA allele to invasion by the other (i.e., stability was assessed at the boundary equilibrium allele frequencies of $\hat{\mbf{p}}_{AA} = 1$ and $\hat{\mbf{p}}_{aa} = 1$; \citealt{Levene1953, Prout1968, deVriesCaswell2019b}).
+We identify conditions where SA polymorphism is 'protected' by evaluating the stability of populations initially fixed for either SA allele to invasion by the other (i.e., stability was assessed at the boundary equilibrium genotype frequencies of $\hat{\mbf{p}}_{AA} = 1$ and $\hat{\mbf{p}}_{aa} = 1$; \citealt{Levene1953, Prout1968, deVriesCaswell2019b}).
 The formal conditions for a protected polymorphism are determined by linearizing the model in the vacinity of the boundary equilibria ($\hat{\mbf{p}}$), and evaluating the magnitude of the largest eigenvalue of the Jacobian matrix of the linearization, which we denote $\tilde{\zeta}_{AA}$ and $\tilde{\zeta}_{aa}$.
-For our model, $\mbf{M}$ has an upper block triangular form, and we can therefore express the stability of $\hat{\mbf{p}}$ in terms of the leading eigenvalue of the diagonal block corresponding to changes in heterozygote frequency ($\mbf{M}_{22}$; e.g., see \citealt{OttoDay2007,deVriesCaswell2019b}).
+For our model, the Jacobian, $\mbf{M}$, has an upper block triangular form, and we can therefore express the stability of $\hat{\mbf{p}}$ in terms of the leading eigenvalue of the diagonal block corresponding to changes in heterozygote frequency ($\mbf{M}_{22}$; e.g., see \citealt{OttoDay2007,deVriesCaswell2019b}).
 For a polymorphism to be 'protected', both boundary equilibria must be unstable (i.e., $\tilde{\zeta}_{AA} > 1$ and $\tilde{\zeta}_{aa} > 1$), which requires that
@@
 We therefore used numerical techniques to identify invasion and extinction thresholds, which could then be used to define regions of demographically viable polymorphism across $s_f \times s_m$ parameter space.
-Because single-locus selection coefficients are generally weak, we limit our analyses to coefficients with within the open interval $\left(0,0.15\right)$, unless stated otherwise.
+Because single-locus selection coefficients are generally weak (i.e., $s < 0.1$; e.g., \citealt{Eyre-WalkerKeightly2007}) and strongly skewed, we limit our analyses to coefficients within $0 < s_f,s_m \leq 0.15$, unless stated otherwise.
 For a given parameter set, we numerically evaluated Eq(\ref{eq:PolyCond}) and $\lambda$, starting at the outer boundaries of selection parameter space, and titrated inward to identify thresholds where $\tilde{\zeta}_i \approx 1$ and $\lambda \approx 1$ (with precision of $10^{-5}$).
 The proportions of total $s_f \times s_m$ parameter space where SA polymorphism and/or extinction occurred were calculated by numerically integrating the appropriate regions defined by the invasion and extinction thresholds
+
+%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Demographic rate estimates for {\itshape M. guttatus}} \label{sec:MguttMethods}
+
+To illustrate the use of our model with empirical data, and see whether demographic rates observed in natural populations are likely to support balanced SA polymorphisms, we used publicly available demographic and fitness data for natural populations of the hermaphroditic flowering plant {\itshape Mimulus guttatus} ({\itshape Erythranthe guttata}).
+{\itshape M.~guttatus}
 

### 2020-11-30 R/functions-Simulations.R

@@ -2336,10 +2336,10 @@ makeDataDeltaPolyParamSpace  <-  function(sMax=0.15, res=0.0015, precision = 1e-
 
 
 	# Storage for data
-	dataSet  <-  rep(NA, times=11)
+	dataSet  <-  matrix(NA, ncol=11)
+	colnames(dataSet)  <-  c("Delta", "h", "deltaVal", "C", "smInv", "aInv", "AInv", "sf", "smExt", "sm", "sfExt")
 
 	print('Making Data For Fig.3')
-
 	# Loop over dominance values
 	for(i in 1:length(hVals)) {
 
@@ -2364,21 +2364,23 @@ makeDataDeltaPolyParamSpace  <-  function(sMax=0.15, res=0.0015, precision = 1e-
 											  delta = deltaSeq[j], delta_j = delta_j, delta_a = delta_a, delta_gamma = delta_gamma,
 											  tlimit = tlimit, eqThreshold = eqThreshold, verbose=FALSE, writeFile=FALSE)
 
-				# Find Extinction Threshold
+			# Find Extinction Threshold
 			extData  <-  extinctThreshTitrate(sMax=sMax, res=res, precision=precision,
 											  om = om, g = g, theta = theta, theta_prime = theta_prime, 
 											  hf = hVals[i], hm = hVals[i], C = Cs[j], 
 											  delta = deltaSeq[j], delta_j = delta_j, delta_a = delta_a, delta_gamma = delta_gamma,
 											  tlimit = tlimit, eqThreshold = eqThreshold, 
 											  makePlots=FALSE, verbose = FALSE, writeFile=FALSE)
-
 			# Append new data to dataSet
-			dataSet  <-  rbind(dataSet, cbind(rep("delta", times=nrow(invData)),
+			dDat  <-  cbind(rep("d", times=nrow(invData)),
 											  rep(hVals[i], times=nrow(invData)),
 											  rep(deltaSeq[j], times=nrow(invData)),
 											  rep(Cs[j], times=nrow(invData)),
 											  invData, 
-											  extData))
+											  extData)
+			colnames(dDat)  <-  colnames(dataSet)
+			dataSet         <-  rbind(dataSet,dDat)
+
 			# Update Progress Bar
 			setTxtProgressBar(pb, j)
 		}
@@ -2405,12 +2407,15 @@ makeDataDeltaPolyParamSpace  <-  function(sMax=0.15, res=0.0015, precision = 1e-
[…]

### 2020-12-01 doc/SA-Hermaphrodites-wDemography.tex
Minor edits; fig changes
Not much.  Update files for work at office.

@@
 \newcommand{\bm}[1]{\mbox{\boldmath $#1$}}
-
+\DeclareRobustCommand{\firstsecond}[2]{#2}
  \def\linenumberfont{\normalfont\scriptsize}
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Demographic rate estimates for {\itshape M. guttatus}} \label{sec:MguttMethods}
+\subsection*{Demographic and fitness data for {\itshape M. guttatus}} \label{sec:MguttMethods}
 
-To illustrate the use of our model with empirical data, and see whether demographic rates observed in natural populations are likely to support balanced SA polymorphisms, we used publicly available demographic and fitness data for natural populations of the hermaphroditic flowering plant {\itshape Mimulus guttatus} ({\itshape Erythranthe guttata}).
-{\itshape M.~guttatus}
+As an illustrative example of whether demographic rates observed in natural populations appear likely to support balanced SA polymorphisms, we parameterized our model using empirically estimated demographic rates and fitness data for natural populations of the partially-selfing hermaphroditic flowering plant {\itshape Mimulus guttatus} (Scrophulariaceae; now known as {\itshape Erythranthe guttata}).
+{\itshape M.~guttatus} is a well established study system for ecological and genomic studies because it shows high genetic diversity and significant variation in numerous ecologically relevant traits \citep{WuWillis2008}.
+In particular, there is well-documented variation in population selfing rates, and the severity of inbreeding depression has been estimated at multiple life-history stages for this species \citep[e.g.,][]{RitlandGanders1987, Ritland1990, Willis1993, Willis1999a, Willis1999b}.
+Moreover, detailed demographic studies have been conducted on multiple populations of {\itshape M.~guttatus}, with demographic matrix data made publicly available on the demographic database COMPADRE \citep{CompadreDB2020}.
+Below, we briefly outline how we parameterized our model using the available data; full details are provided in \hl{Appendix B.x}.
 

### 2020-12-01 R/functions-Figs.R

@@ -659,7 +659,7 @@ deltaSelfingLoadPolySpaceFigTitrate  <-  function(df = "dataDeltaPolySpaceFig_sM
 
 
 # Set plot layout
-    layout.mat  <- matrix(c(1:4), nrow=2, ncol=2, byrow=TRUE)
+    layout.mat  <- matrix(c(1:2), nrow=2, ncol=1, byrow=TRUE)
     layout      <- layout(layout.mat,respect=TRUE)
 
     ## Panel A: Additive SA (hf = hm = 1/2)
@@ -678,68 +678,36 @@ deltaSelfingLoadPolySpaceFigTitrate  <-  function(df = "dataDeltaPolySpaceFig_sM
         box()
         # Simulation Results
         lines(PGSpace ~ CLine, lwd=2, col=COLS$PG)
-        points(PrViaPoly[Delta == "d"][c(1:9,21:23)] ~ C[Delta == "d"][c(1:9,21:23)], pch=21, bg=COLS$dSim, col=COLS$dSim2, data=d)
-        # axes        
-        axis(1, las=1, labels=NA)
-        axis(2, las=1)
-        proportionalLabel(0.03, 1.075, 'A', cex=1.2, adj=c(0.5, 0.5), xpd=NA)
-        proportionalLabel(0.5, 1.2, expression(paste("Early-acting Inbreeding Depression")), cex=1.2, adj=c(0.5, 0.5), xpd=NA, srt=0)
-        proportionalLabel(-0.5, 0.5, expression(paste(italic(h), " = ", 1/2)), cex=1.5, adj=c(0.5, 0.5), xpd=NA, srt=90)
-        proportionalLabel(-0.3, 0.5, expression(paste("Prop. viable polymorphic space")), cex=1.2, adj=c(0.5, 0.5), xpd=NA, srt=90)
-
-      #Legend
-        legend( x       =  usr[2]*0.95,
-                y       =  usr[4],
-                legend  =  c(
-                             expression(paste(italic(delta), " (Pop. Gen.)")),
-                             expression(paste(delta))),
-                 lty     =  c(1,NA),
-                 lwd     =  c(2,NA),
-                 col     =  c(COLS$PG,
-                              COLS$dSim),
-                 pch     =  c(NA,21),
-                 pt.bg   =  c(NA,
-                              COLS$dSim),
-                 cex     =  0.75,
[…]

### 2020-12-01 R/functions-Simulations.R

@@ -2498,7 +2498,7 @@ makeDataDeltaPolyParamSpace  <-  function(sMax=0.15, res=0.0015, precision = 1e-
 
 
 	# Export data as .csv to ./output/data
-	filename <-  paste("./output/simData/dataDeltaPolySpaceFig2", "_sMax", sMax, "_res", res, "_dStar", dStar, "_f", theta[4], ".csv", sep="")
+	filename <-  paste("./output/simData/dataDeltaPolySpaceFig", "_sMax", sMax, "_res", res, "_dStar", dStar, "_f", theta[4], ".csv", sep="")
 	write.csv(dataSet, file=filename, row.names = FALSE)
 
 }

### 2020-12-01 makeFigs.R

@@ -37,7 +37,7 @@ embed_fonts(figPath(name='polymorphicSpaceTitrate.pdf'))
 # Fig. 3 - Quantification of demographically viable parameter space, 
 #			showing effects of inbreeding depression
 toPdf(deltaSelfingLoadPolySpaceFigTitrate(), 
-			figPath(name='deltaPolymorphicSpaceTitrate.pdf'), width=10, height=7)
+			figPath(name='deltaPolymorphicSpaceTitrate.pdf'), width=5, height=7)
 embed_fonts(figPath(name='deltaPolymorphicSpaceTitrate.pdf'))
 
 

### 2020-12-01 doc/SA-Hermaphrodites-wDemography.tex
Draft of Mimulus methods
Added first draft of basic methods for parameterizing the model
using empirical data for M. guttatus

--- 2020-11-30 doc/SA-Hermaphrodites-wDemography.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Demographic rate estimates for {\itshape M. guttatus}} \label{sec:MguttMethods}
+\subsection*{Demographic and fitness data for {\itshape M. guttatus}} \label{sec:MguttMethods}
 
-To illustrate the use of our model with empirical data, and see whether demographic rates observed in natural populations are likely to support balanced SA polymorphisms, we used publicly available demographic and fitness data for natural populations of the hermaphroditic flowering plant {\itshape Mimulus guttatus} ({\itshape Erythranthe guttata}).
-{\itshape M.~guttatus}
+As an illustrative example of how our model can be used to explore whether demographic rates observed in natural populations appear likely to support balanced SA polymorphisms, we parameterized our model using empirically estimated demographic rates and fitness data for natural populations of the hermaphroditic flowering plant {\itshape Mimulus guttatus} (Scrophulariaceae; now known as {\itshape Erythranthe guttata}).
+{\itshape M.~guttatus} is an herbaceous, self-compatible wildflower native to western North America that exhibits remarkable among-population variation in numerous life-history and reproductive traits including selfing rates, inbreeding depression, floral morphology, and annual-to-perennial life-history \citep[e.g.,][]{RitlandGanders1987, Ritland1990, Willis1993, Willis1999a, Willis1999b, WuWillis2008}.
+Moreover, detailed demographic studies have been conducted on multiple populations of {\itshape M.~guttatus}, with demographic matrix data made publicly available on the demographic database COMPADRE \citep{CompadreDB2020}.
+Below, we briefly outline how we parameterized our model using the available data; full details are provided in \hl{Appendix B.x}.
+
+We used extensive demographic rate data reported for the 'Eagle Meadows' population in Stanislaus National Forest (California, USA) studied by \citet{PetersonEtAl2016}.
+This study entailed a large common garden experiment which took place at the 'Eagle Meadows' site.
+We used demographic rates estimated for this population grown in its native habitat (i.e., we did not use data for non-local populations grown at Eagle Meadows).
+\citet{PetersonEtAl2016} collected data for numerous life-history stages and transition rates; for the Eagle Meadows population these estimates were as follows: seed bank survival ($D = 0.534$), seed germination rate ($G = 0.469$), flower production ($F = 0.64$), ovules per flower ($O = 614)$, seedling recruits proportional to clonal rosette recruits ($A = 6.7e-4$), overwinter survival ($S = 0.179$), rosette production ($R = 8.71$) (see corrected Tables.~1 and S2 in \citealt{PetersonEtAl2017}).
+These data were also available through the Compadre demographic matrix database \citep{CompadreDB2020}.
+The resulting transition matrices for this population involved three life-history stages ($\omega = 3$; seed, seedling, and rosette), and individual elements of the transition matrix ($\tilde{\mbf{A}}$) were calculated as products of the above rates (see Matrix 1 in \citealt{PetersonEtAl2016}).
+
+Unfortunately, estimates of selfing rates and inbreeding depression are not also available for the Eagle Meadows population of {\itshape M.~guttatus}.
+However, selfing rate estimates are available for a variety of other populations, and vary in magnitude from near complete outcrossing to predominant selfing ($C \approx 0$ to $0.75$; \citealt{RitlandGanders1987, Ritland1990, Willis1999b}).
+Additionally, estimates of inbreeding depression at several of the life-history stages/fitness components that were included in the data of \citet{PetersonEtAl2016} are available for two intensively studied populations in the Cascade Mountains of Oregon (Iron Mountain and Cone Peak) \citet{Willis1993, Willis1999a, Willis1999b}.
+Using the data provided in \citet{Willis1993}, we estimated the proportional decrease due to inbreeding depression in seed germination rate ($\delta_{G} = 0.085$), flower number ($\delta_{F} = 0.2$), and overwinter survival ($\delta_{S} = 0.38$).
+
+Using these combined demographic rates, selfing rate, and inbreeding depression estimates, we constructed a corresponding stage $\times$ genotype mendelian matrix model with a single SA locus affecting female and male fertility (as described above).
+Using this empirically parameterized model, we are able to make predictions about the scope for demographically viable SA polymorphism at a single locus in a hypothetical population with the same demographic rates as observed in the Eagle Meadows population, for a range of selfing and inbreeding depression rates observed in other natural populations.
+However, this should not be confused with an explicit prediction for the likelihood of SA polymorphism in the Eagle Meadows population.
+
+Interestingly, a polymorphic chromosomal inversion (Inv6) with apparently SA fitness effects has been identified in the Iron Mountain population of {\itshape M.~guttatus} \citep{LeeKelly2015}.
+Inv6 segregates at moderate frequency (about $8\%$), and carriers suffer $\approx 30 \%$ loss in pollen viability, but also increased flower production which varies among years.
+The genetic basis of the fitness effects caused by Inv6 are not yet known, but the net result is a "supergene" with uniquely strong fitness effects that segregates as a single diallelic locus.
+We estimated dominance and selection coefficients for the effect of Inv6 on pollen viability and flower production from greenhouse crossing experiments using mapping poulations and field estimates.
+Across four mapping populations, the effect of Inv6 on pollen viability was partially recessive ($h_m \approx 0.35$).
+Given this average dominance, the selection coefficient in the field is approximately $s_m = 0.88$.
+It was not possible to estimate the average dominance of the effect of Inv6 on flower production, so we conservatively assume that it was additive ($h_f = 0.5$).
+Based this assumption, the selection coefficient Inv6 on flower production in the field was $s_f = 0.384$ in 2012, and $s_f = 0.232$ in 2013.
+As a final proof-of-concept test of our empirically parameterized model \citep[e.g.,][]{Servedio2014}, we asked whether, given the above biologically grounded fitness effect estimates, Inv6 appears to fall in demographically viable SA polymorphic parameter space, as might be expected given its observed frequency in the Iron Mountain population.
 

### 2020-12-01 makeFigs.R

@@ -18,7 +18,7 @@ rm(list=ls())
 ################
 ## Dependencies
 source('R/functions-Figs.R')
-
+source('./R/loadData-Compadre.R')
 
 #################
 # Figs for Paper

### 2020-12-01 run-Simulations.R

@@ -262,6 +262,12 @@ makeDataDeltaPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
 							tlimit = 10^5, eqThreshold = 1e-8)
 
 
+makeDataDeltaPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
+							om = 2, g = 3, theta = c(0.6,0.6,0.05,7), theta_prime = 7, 
+							hVals= c(1/2, 1/4), dStar = 0.8, 
+							tlimit = 10^5, eqThreshold = 1e-8)
+
+
 ######################
 # Preliminary figures
 ######################

### 2020-12-07 doc/SA-Hermaphrodites-wDemography.tex
Edits on intro and methods
Two new paragraphs added to intro, small edits of methods section

@@
 \newcommand{\bm}[1]{\mbox{\boldmath $#1$}}
+\newcommand{\om}{\omega}
 \DeclareRobustCommand{\firstsecond}[2]{#2}
  \def\linenumberfont{\normalfont\scriptsize}
+ \newcommand{\kron}{\otimes}
 %===========================================
@@
  Whether a population persists or goes extinct depends in large part on its ability to adapt to the environment.
+Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, such as deleterious mutations \citep{Haldane1957}, environmental shifts \citep{Maynard-Smith1976, LandeShannon1996,OrrUnckless2008}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, or genetic constraints \citep{ConnallonHall2018}.
+
+ % Whether a population persists or goes extinct depends in large part on its ability to adapt to the environment.
 Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, including deleterious mutations \citep{Haldane1957}, environmental shifts causing a population to chase a new phenotypic optimum \citep{Maynard-Smith1976, LandeShannon1996,OrrUnckless2008}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, as well as genetic constraints arising from intrinsic features of the inheritance system \citep{ConnallonHall2018}.
 
-Conflicting selection and gene-flow between different classes of individuals, or genetic trade-offs between fitness components, can impose constraints on adaptation with particularly interesting evolutionary consequences \citep{CharlesworthHughes2000, ConnallonHall2018}.
-On the one hand, such constraints hinder adaptation by preventing individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase the population's overall extinction risk.
+%Conflicting selection and gene-flow between different classes of individuals, or genetic trade-offs between fitness components, can impose constraints on adaptation with particularly interesting evolutionary consequences \citep{CharlesworthHughes2000, ConnallonHall2018}.
+Genetic constraints can arise as a consequence of conflicting selection and gene-flow between different classes of individuals.
+That is, genetic trade-offs between fitness components can constrain adaptation \citep{CharlesworthHughes2000, ConnallonHall2018}.
+On the one hand, such constraints prevent individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase the population's overall extinction risk \citep{holman2013consequences,harts2014demography}.
 On the other hand, they provide an effective mechanism for the maintenance of genetic variation, which influences the population's capactiy for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
-Hence, for traits under conflicting selection, the nature and extent of genetic variation observed in natural populations should reflect a balance between the contribution of genetic constraints to the maintainance of genetic polymorphisms, and the demographic consequences of the resulting maladaptation.
+Hence, for traits under conflicting selection, the nature and extent of genetic variation observed in natural populations should reflect a balance between the contribution of genetic constraints to the maintainance of genetic polymorphisms, and the population dynamical consequences of the resulting maladaptation.
+
+Classical population genetic models consider relative fitnesses of genotypes, and therefore do not consider the population dynamical consequences of evolutionary change.
+To study the population dynamical consequences of evolutionary change, a model is required that links changes in genotype distributions to population dynamics.
+The demographic processes of birth and death drive both changes in genotype distributions and population dynamics, and therefore provide a fundamental link between evolution and ecology \citep{metcalf2007evolutionary}.
+
+In this paper, we use such an evolutionary demographic approach (model?) to study whether the parameter space predicted to be polymorphic by population genetics models is, indeed, demographically viable.
+We apply the model to study the maintenance of polymorphisms through sexually antagonistic selection in both dioecious and hermaproditic species.
+A major strength of the model is that we can use empirically estimated demographic rates to parameterize the model, enabling us to make predictions about the scope for demographically viable SA polymorphisms that are grounded in the biology of real populations.
+We demonstrate this with a case study of {\itshape Mimulus guttatus} (now {\itshape Erythranthe guttata})
+
+%Such an eco-evolutionary model requires a map from genotype to phenotype, from phenotype to individual demographic rates (birth and death rates), and finally from demographic rates to population level characteristics \citep{coulson2006putting}.
 %Genetic trade-offs between fitness components due to pleiotropy represent a form of genetic constraint with particularly interesting evolutionary consequences because they act as a double-edged sword
@@
 
-Genetic trade-offs between male and female fitness are a common feature of sexually reproducing populations, and are thought to contribute not only to the maintenance of genetic variation, but also the evolution of sexual dimorphism and gender-related traits \citep{Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, BondurianskyChenoweth2009,Olito2019}.
-When such genetic trade-offs occur, "sexually antagonistic" selection (abbreviated SA hereafter) can arise when beneficial alleles for one sex are deleterious when expressed in the other \citep{Kidwell1977, Rice1992, ConnallonClark2012}.
-This, alleles with opposing fitness effects through male and female sex functions create analogous genetic constraints on fitness in hermaphrodite populations, where both maternal and paternal reproductive success contribute jointly to each individuals' overall fitness \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
+Sexually antagonstic selection, due to genetic trade-offs between male and female fitness, are a common feature of sexually reproducing populations, and are thought to contribute not only to the maintenance of genetic variation, but also the evolution of sexual dimorphism and gender-related traits \citep{Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, BondurianskyChenoweth2009,Olito2019}.
+When such genetic trade-offs occur, "sexually antagonistic" selection (abbreviated SA hereafter) can arise whereby beneficial alleles for one sex are deleterious when expressed in the other \citep{Kidwell1977, Rice1992, ConnallonClark2012}.
+Such alleles with opposing fitness effects through male and female sex functions create analogous genetic constraints on fitness in hermaphrodite populations, where both maternal and paternal reproductive success contribute jointly to each individuals' overall fitness \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
 
@@
 Overall, both theoretical predictions and current empirical data suggest there is ample scope for SA trade-offs and the maintenance of SA polymorphisms in both dioecious and hermaphrodite populations \citep{Abbott2011, WangBarrett2020}, although identifying specific SA loci from genome sequence data remains challenging \citep{RuzickaESEB2020}.
-\bigskip
+% \bigskip
 
-\begin{enumerate}
-	\item Introduce Evolutionary Demography, pivot to focus on the consequences of taking demography into account explicitly for the maintenance of SA polymorphism. [Lotte]
-	\item Introduce key difference between existing population genetic models of SA polymorphism and our models: possibility of diverse population and evolutionary dynamics. [Lotte then Colin]
-	\item Explain and highlight "demographically viable" parameter space, indicate that we are interested in whether parameter space predicted to be polymorphic by Pop.~gen.~models is, indeed, "demographically viable" [Lotte then Colin]
-		\begin{enumerate}
-			\item Pop.
+% \begin{enumerate}
+% 	\item Introduce Evolutionary Demography, pivot to focus on the consequences of taking demography into account explicitly for the maintenance of SA polymorphism. [Lotte]
+% 	\item Introduce key difference between existing population genetic models of SA polymorphism and our models: possibility of diverse population and evolutionary dynamics. [Lotte then Colin]
+% 	\item Explain and highlight "demographically viable" parameter space, indicate that we are interested in whether parameter space predicted to be polymorphic by Pop.~gen.~models is, indeed, "demographically viable" [Lotte then Colin]
+% 		\begin{enumerate}
+% 			\item Pop.
 Gen.
@@
 Casually define fitness as lifetime reproductive fitness of a given genotype.
-			\item "In practice, these genotypic fitnesses emerge from myriad process acting throughout the life-history of individuals"
-			\item these processes contribute to key demographic rates such as the intrinsic population growth rate, 
-		\end{enumerate}
-		\item A major strength of the model is that we can use empirically estimated demographic rates to parameterize the model, enabling us to make predictions about the scope for demographically viable SA polymorphisms that are grounded in the biology of real populations.
-\end{enumerate}
-\bigskip
+% 			\item "In practice, these genotypic fitnesses emerge from myriad process acting throughout the life-history of individuals"
+% 			\item these processes contribute to key demographic rates such as the intrinsic population growth rate, 
+% 		\end{enumerate}
+% 		\item A major strength of the model is that we can use empirically estimated demographic rates to parameterize the model, enabling us to make predictions about the scope for demographically viable SA polymorphisms that are grounded in the biology of real populations.
+% \end{enumerate}
+% \bigskip
 
@@
 
-In the model, individuals are classified jointly by life-cycle stage ($1, \ldots, \omega$), genotype ($1, \ldots, g$), and whether they were produced by self- or outcross fertilization (denoted by $S$ and $X$ superscripts).
+Simultaneous hermaphrodites express both male and female sex functions (i.e., can be both a mother and father), and have the potential to reproduce by a combination of self- and outcross-fertilizations.
+Maternal outcrossing involves receiving male gametes from another individual in the population, while paternal outcrossing involves exporting male gametes to another individual and fertilizing their ovules.
+Self-fertilization is achieved when an individual's male gametes fertilize their own ovules.
+In the following any parameters, matrices or vectors relating to male sex function are denoted with a prime.
+
+In the model, individuals are classified jointly by life-cycle stage ($1, \ldots, \omega$), genotype ($1, \ldots, g$), and whether they were produced by self- or outcross fertilization (denoted by $S$ and $X$ superscripts, respectively).
+Whether individuals were produced through selfing or outcrossing is included in the individual state description because individuals produced through selfing might experience reduced survival, growth, or maturation rates as a consequence of inbreeding depression.
+To model inbreeding depression, we therefore keep track of how individuals were produced.
+
 Each genotype is characterized by a matrix of transition probabilities (including survival) and a matrix of reproductive output through male and female function.
@@
 
-The population state at time $t$ is described by a stage $\times$ genotype distribution matrix, which can be transformed into a population state vector, $\tilde{\mbf{n}}(t)$, which is ordered by genotype, then stage, then by how individuals were produced (self vs.~outcrossing).
+The population state at time $t$ is described by a population state vector, $\tilde{\mbf{n}}(t)$, which is ordered by how individuals were produced (selfing vs.~outcrossing), then by genotype, and finally by stage.
 For a single diallelic locus with alleles $A$ and $a$, we have three genotypes ($AA,\, Aa,\, aa$; $g = 3$), giving the population state vector:
@@
 \begin{equation} \label{eq:propPopVec}
-	\tilde{\mbf{p}}(t) = \frac{\tilde{\mbf{n}}(t)}{ \| \tilde{\mbf{n}}(t) \|} = \mbf{p}(t),
+	\tilde{\mbf{p}}(t) = \frac{\tilde{\mbf{n}}(t)}{ \| \tilde{\mbf{n}}(t) \|},
 \end{equation}
@@
 
-The population projection matrix $\tilde{\mbf{A}}$ is constructed from four sets of matrices representing the demographic and genetic processes: The matrices $\mbf{U}^{S}_{i}$ and $\mbf{U}^{X}_{i}$ contains transition and survival probabilities for each genotype, produced by selfing and outcrossing respectively.
+The population projection matrix $\tilde{\mbf{A}}$ is constructed from four sets of matrices representing the demographic and genetic processes: The matrices $\mbf{U}^{S}_{i}$ and $\mbf{U}^{X}_{i}$ contain transition and survival probabilities for each genotype, produced by selfing and outcrossing respectively.
 The matrices $\mbf{F}_{i}$ and $\mbf{F}^{\prime}_{i}$ contain the genotype $\times$ stage specific contribution via female and male sex-functions of genotype $i$ to the female and male gamete pools, respectively, and therefore to zygotes in the next generation.
+We assume that whether individuals were produced through selfing or outcrossing does not affect their fecundity or mating success, that is, we assume  $\mbf{F}^X_{i}=\mbf{F}^S_{i}=\mbf{F}_{i}$ and $\mbf{F}^{\prime S}_{i}=\mbf{F}^{\prime X}_{i}=\mbf{F}^{\prime}_{i}$.
+Deviations from this assumption are straightforward to incorporate but beyond the scope of the current paper.
 In principle, the model can admit any kind of pleiotropy among demographic traits because the transition and fertility matrices can differ among genotypes in any way.
@@
 
-Similar to the two-sex model of \citet{deVriesCaswell2019b}, we define the matrices $\mbf{H}^S_{j}(\tilde{\mbf{n}})$ and $\mbf{H}^X_{j}(\tilde{\mbf{n}})$, which map the genotypes of the parent in stage $j$ to the genotypes of their offspring produced by selfing and outcrossing, respectively.
-The $(k, l)$ entry of $\mbf{H}^{X}_{j}$ and $\mbf{H}^{S}_{i}$ is the probability that an offspring of a genotype $l$ mother, of stage $i$, has genotype $k$.
+Notational note: The superscript on a matrix in age$\times$stage models conventionally indicates the stage or age of the subsection of the population vector that the matrix is acting on \citep{CaswellEtAl2018}.
+For example, $\mbf{U}^{S}_{i}$, acts on the vector of individuals of genotype $i$ that were produced through selfing.
+In the following, we deviate from this convention by also using $S$ and $X$ as superscripts to indicate reproduction of offspring through current selfing or reproduction.
+That is,  we define the matrices $\mbf{H}^S_{j}(\tilde{\mbf{n}})$ and $\mbf{H}^X_{j}(\tilde{\mbf{n}})$, as the map of the genotypes of the parent in stage $j$ to the genotypes of their offspring produced by selfing and outcrossing, respectively.
+The $(k, l)$ entry of $\mbf{H}^{X}_{j}$ and $\mbf{H}^{S}_{j}$ is the probability that an offspring of a genotype $l$ mother, of stage $j$, has genotype $k$.
+Since we assume that how-you-were-produced does not affect how you reproduce, these matrices will act on both $\mbf{n}^{S}$ and $\mbf{n}^{X}$.
+Likewise, we will write production $\mcal{F}^S$ and $\mcal{F}^X$ to denote offspring production by self-fertilization and outcrossing.
+
 For the purpose of this article, we assume that mating is random with respect to stage and hence that the parent-offspring map is the same for all stages (i.e., $\mbf{H}^S_{j}(\tilde{\mbf{n}}) = \mbf{H}^S(\tilde{\mbf{n}})$, and $\mbf{H}^X_{j}(\tilde{\mbf{n}}) = \mbf{H}^X(\tilde{\mbf{n}})$).
@@
 \subsubsection*{Mating and offspring production under partial selfing}
-
-Simultaneous hermaphrodites express both male and female sex functions (i.e., can be both a mother and father), and have the potential to reproduce by a combination of self- and outcross-fertilizations.
-Maternal outcrossing involves receiving male gametes from another individual in the population, while paternal outcrossing involves exporting male gametes to another individual and fertilizing their ovules.
-Self-fertilization is acheived when an individual's male gametes fertilize their own ovules.
-These modes of reproduction complicate mating in our model relative to the two-sex stage-structured model of \citet{deVriesCaswell2019b}.
+The inclusion of simultaneous hermaphrodites that can reproduce both through selfing and through outcrossing complicate mating in our model relative to the two-sex stage-structured model of \citet{deVriesCaswell2019b}.
+Since we assume that how individuals were produced does not affect the relative success of their male gametes, the frequency distribution in the male gametes is obtained simply by summing the vectors of individuals produced through selfing and individuals produced through outcrossing, that is,
+\begin{equation}
+ 	\mbf{p}^{\prime}=\frac{(\mbf{n}^X + \mbf{n}^S)}{\| (\mbf{n}^X + \mbf{n}^S)\|}.
+ \end{equation}
 
@@
 In \hl{Appendix A.x}, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
-Outcrossing proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b} except that each individual's genotype determines both the number of ovules produced and also paternal "mating success", broadly defined.
+Outcrossing proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}.
+That is, each individual's genotype determines both the number of ovules produced and the paternal "mating success", broadly defined.
 For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986, WangBarrett2020, Harder2016}.
-Depending on their stage and genotype at the SA locus, individuals contribute male gametes to the overall population gamete pool according to the male stage $\times$ genotype vector:
+Depending on their stage and genotype at the SA locus, individuals contribute male gametes to the overall population gamete pool according to the stage $\times$ genotype vector:
 \begin{linenomath*}
 \begin{equation}
-		\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p} = 
+		\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime}= 
 		\left(
@@
 			\begin{array}{c}
-				\mbf{p}_{AA} \\
-				\mbf{p}_{Aa} \\
-				\mbf{p}_{aa} \\
+				\mbf{p}^{\prime}_{AA} \\
+				\mbf{p}^{\prime}_{Aa} \\
+				\mbf{p}^{\prime}_{aa} \\
 			\end{array} \right).
@@
 
-\noindent The vector $\mbf{p}$ gives the genotype frequencies at mating, and is the sum of $\mbf{n}^{S}$ and $\mbf{n}^{X}$, since in partially selfing popualtions, whether an individual was produced by selfing or outcrossing does not affect whether they can reproduce by outcrossing themselves.
-The matrix $\mbb{F}^{\prime}$ is composed of genotype-specific fertility matrices, and operates on the the vector of genotype frequencies to give their relative contribution of each genotype to the gamete pool.
+\noindent The matrix $\mbb{F}^{\prime}$ is composed of genotype-specific fertility matrices, and operates on the the vector of genotype frequencies to give their relative contribution of each genotype to the gamete pool.
 The matrix $\mbf{W}^{\prime}$ converts these contributions to allele numbers.
@@
 		\end{array} \right) = 
-			\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p} \|} = 
+			\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime}}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime} \|} = 
 				\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)\|}
@@
 \begin{linenomath*}
-\begin{equation}
-	\mbb{H}^j = \mbf{I}_{\omega} \otimes \mbf{H}^j(\tilde{\mbf{p}}),
-\end{equation}
+\begin{eqnarray}
+	\mathbb{H}^X(\tilde{\bo p}) = \bo I_\om \kron \bo H^X(\tilde{\bo p}),\\
+	\mathbb{H}^S(\tilde{\bo p}) = \bo I_\om \kron \bo H^S(\tilde{\bo p}).
+\end{eqnarray}
 \end{linenomath*}
-
-\noindent where $j \in \{S,X\}$. The corresponding fertility matrices for self-fertilization and outcrossing are defined as
+The corresponding fertility matrices for self-fertilization and outcrossing are defined as
 \begin{linenomath*}
@@
 The blocks of $\mcal{F}^X(\tilde{\mbf{p}})$ can be constructed and interpreted as follows: The first row block of the first column produces $AA$ offspring by outcrossing from $AA$ maternal parents.
-This happens when the $AA$ maternal parent receives a random gamete from the male gamete pool, which happens with probability $q^{\prime}_{A}$.
+This happens when the $AA$ maternal parent receives an $A$ gamete from the male gamete pool, which happens with probability $q^{\prime}_{A}$.
 The other blocks can be interpreted similarly.
@@
 Suppose that there is a genetic trade-off between the sex-functions at a single bi-allelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect.
+
 Following convention, we parameterize the relative fitnesses of each genotype through each sex function ($w_{i}$ and $w^{\prime}_{i}$) to be bounded by $(0,1)$, with dominance and selection coefficients $h_f, s_f$ and $h_m, s_m$ determining the decrease in fertility through each sex function relative to the most fit genotype ($AA$ has highest female fertility, $aa$ the highest male fertility; see Table \ref{tab:Fitness}).
@@
 However, the survival matrices can be used to model the fitness effects of inbreeding depression at later stages of development by allowing different stage-specific survival and transition rates for individuals produced by self-fertilization vs.~outcrossing.
-By contrast, the parameter $\sigma$ only affects inbreeding depression through viability of selfed ovlues.
+By contrast, the parameter $\sigma$ only affects inbreeding depression through viability of selfed ovules.
 With this in mind, we define survival matrices for individuals produced by selfing and outcrossing as follows:
@@
 \noindent where $\sigma_j$ and $\sigma_a$ are the juvenile and adult stage survival rates, $\gamma$ is the maturation rate from juvenile to adult stages, and the corresponding $\delta_j$, $\delta_a$, and $\delta_{\gamma}$ terms denote the proportional decreases in stage-specific survival and transition rates due to inbreeding depression (i.e., deleterious effects of inbreeding at later life-history stages; e.g., \citealt{HarderRoutely2006}).
-Any or all of these parameters may differ among genotypes, so that selection can operate on stage-specific viability, development, and/or fertility.
-For simplicity, we assume these are constant among genotypes.
+For simplicity, we assume survival and transition rates are constant among genotypes.
 
@@
 
-Unless stated otherwise, we use the following parameter values for the demogrphic rates in the model: $\sigma_j = \sigma_A = 0.6$ and $\gamma = 0.05$, which are similar to those used in \citet{deVriesCaswell2019b}.
-Our parameters of interest include $f$, $\delta$, $\delta_j$, $\delta_a$, $\delta_{\gamma}$, and the selection parameters $h_f$, $s_f$, $h_m$, and $s_m$, which are given different values for each analyses as described in the figure captions.
+Unless stated otherwise, we use the following parameter values for the demographic rates in the model: $\sigma_j = \sigma_A = 0.6$ and $\gamma = 0.05$, which are similar to those used in \citet{deVriesCaswell2019b}.
+Our parameters of interest include fertility, $f$, the inbreeding depression parameters, $\delta$, $\delta_j$, $\delta_a$, $\delta_{\gamma}$, and the selection parameters $h_f$, $s_f$, $h_m$, and $s_m$, which are given different values for each analyses as described in the figure captions.
 
@@
  \centering
- \includegraphics[scale=0.75]{../output/figs/polymorphicSpaceTitrate.pdf}
+ \includegraphics[scale=0.75]{../output/figs/polymorphicSpaceNew.pdf}
  \caption{\footnotesize{Proportion of demographically viable parameter space (out of total $s_f \times s_m$ space with $\max(s) = 0.15$) in the absence of inbreeding depression (i.e., assuming $\delta = \delta_i = 0$, where $i \in \{j,a,\gamma\}$), plotted as a function of the population selfing rate.

### 2020-12-07 doc/SA-Hermaphrodites-wDemography.tex
Minor edits to analyses section
I'm wondering if it is worth having a seperate methods and results section for Mimulus, or if it would make sense to squish it all into one Mimulus section? Not sure.

@@
 We identify conditions where SA polymorphism is 'protected' by evaluating the stability of populations initially fixed for either SA allele to invasion by the other (i.e., stability was assessed at the boundary equilibrium genotype frequencies of $\hat{\mbf{p}}_{AA} = 1$ and $\hat{\mbf{p}}_{aa} = 1$; \citealt{Levene1953, Prout1968, deVriesCaswell2019b}).
-The formal conditions for a protected polymorphism are determined by linearizing the model in the vacinity of the boundary equilibria ($\hat{\mbf{p}}$), and evaluating the magnitude of the largest eigenvalue of the Jacobian matrix of the linearization, which we denote $\tilde{\zeta}_{AA}$ and $\tilde{\zeta}_{aa}$.
+The formal conditions for a protected polymorphism are determined by linearizing the model in the vicinity of the boundary equilibria ($\hat{\mbf{p}}$), and evaluating the magnitude of the largest eigenvalue of the Jacobian matrix of the linearization, denoted by $\tilde{\zeta}_{AA}$ and $\tilde{\zeta}_{aa}$ for the eigenvalue of the Jacobian evaluated at the $AA$ and $aa$ boundary, respectively.
 For our model, the Jacobian, $\mbf{M}$, has an upper block triangular form, and we can therefore express the stability of $\hat{\mbf{p}}$ in terms of the leading eigenvalue of the diagonal block corresponding to changes in heterozygote frequency ($\mbf{M}_{22}$; e.g., see \citealt{OttoDay2007,deVriesCaswell2019b}).
@@
 \begin{equation}\label{eq:PolyCond}
-	1 < \frac{\rho(\mbf{M}_{22})}{\lambda_{i}}
-\end{equation}
-\end{linenomath*}
-
-\noindent is satisfied, where $\rho(\cdot)$ represents the spectral radius, and $\lambda_i$ is the intrinsic growth rate for the homozygous genotype being invaded at each boundary ($i \in (AA,aa)$; see Box.~1).
+	\rho(\mbf{M}_{22}) > \lambda_{i}
+\end{equation}
+\end{linenomath*}
+
+\noindent is satisfied, where $\rho(\cdot)$ represents the spectral radius,  $\rho(\mbf{M}_{22})$ gives the ergodic growth rate of perturbations of the boundary equilibrium according to the linear approximation, and $\lambda_i$ is the intrinsic growth rate for the homozygous genotype being invaded at each boundary ($i \in (AA,aa)$; see Box.~1).
 A full derivation of the Jacobian and details of the invasion analysis are provided in \hl{Online Appendix X.x}.
@@
 Specifically, for each boundary equilibrium we introduced the rare allele at low initial frequency, and iterated Eq(\ref{eq:Atilde}) until the population had reached demographic and genotypic equilibrium.
-In our model, the population state vector will grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see \citealt{Caswell2001}), and the intrinsic population growth rate, $\lambda$, can be calculated as $\tilde{\mbf{n}}(t)/\tilde{\mbf{n}}(t-1)$, where $t$ is the final generation of the simulation.
+In our model, the population state vector will grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see \citealt{Caswell2001}), and the intrinsic population growth rate after convergence, $\lambda$, can be calculated as $\tilde{\mbf{n}}(t)/\tilde{\mbf{n}}(t-1)$.
 We note, however, that if the ecological component of the model is non-linear, more exotic dynamics are possible.
 
-Unfortunately, closed-form solutions for the invasion conditions for each allele and the equilibrium frequencies under different selection scenarios are analytically intractable except for very restricted conditions (see \hl{Online Appendix X}).
+Unfortunately, closed-form solutions for the invasion conditions for each allele and the equilibrium frequencies under different selection scenarios are analytically intractable except for restricted conditions (see \hl{Online Appendix X}).
 We therefore used numerical techniques to identify invasion and extinction thresholds, which could then be used to define regions of demographically viable polymorphism across $s_f \times s_m$ parameter space.
@@
 The genetic basis of the fitness effects caused by Inv6 are not yet known, but the net result is a "supergene" with uniquely strong fitness effects that segregates as a single diallelic locus.
-We estimated dominance and selection coefficients for the effect of Inv6 on pollen viability and flower production from greenhouse crossing experiments using mapping poulations and field estimates.
+We estimated dominance and selection coefficients for the effect of Inv6 on pollen viability and flower production from greenhouse crossing experiments using mapping populations and field estimates.
 Across four mapping populations, the effect of Inv6 on pollen viability was partially recessive ($h_m \approx 0.35$).
@@
 It was not possible to estimate the average dominance of the effect of Inv6 on flower production, so we conservatively assume that it was additive ($h_f = 0.5$).
-Based this assumption, the selection coefficient Inv6 on flower production in the field was $s_f = 0.384$ in 2012, and $s_f = 0.232$ in 2013.
+Based on this assumption, the selection coefficient Inv6 on flower production in the field was $s_f = 0.384$ in 2012, and $s_f = 0.232$ in 2013.
 As a final proof-of-concept test of our empirically parameterized model \citep[e.g.,][]{Servedio2014}, we asked whether, given the above biologically grounded fitness effect estimates, Inv6 appears to fall in demographically viable SA polymorphic parameter space, as might be expected given its observed frequency in the Iron Mountain population.

### 2021-01-27 doc/SA-Hermaphrodites-wDemography.tex
Update Results Section and Figures
Big commit: finished up first draft of results section, minor edits
to introduction. Also added new data for fig. 3, and a new fig. 4
with Mimulus simulations.

@@
 \usepackage{pdflscape}
+\usepackage{nameref}
 \usepackage[colorlinks=true, allcolors=black]{hyperref}
-\usepackage{nameref}
 
@@
 
- Whether a population persists or goes extinct depends in large part on its ability to adapt to the environment.
+ Whether populations persist or go extinct depends heavily on their ability to adapt to the environment.
 Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, such as deleterious mutations \citep{Haldane1957}, environmental shifts \citep{Maynard-Smith1976, LandeShannon1996,OrrUnckless2008}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, or genetic constraints \citep{ConnallonHall2018}.
+Genetic constraints on adaptation can arise as a consequence of conflicting selection and gene-flow between different classes of individuals, or more generally, when there are genetic trade-offs between fitness components \citep{CharlesworthHughes2000, ConnallonHall2018}.
 
@@
 %Conflicting selection and gene-flow between different classes of individuals, or genetic trade-offs between fitness components, can impose constraints on adaptation with particularly interesting evolutionary consequences \citep{CharlesworthHughes2000, ConnallonHall2018}.
-Genetic constraints can arise as a consequence of conflicting selection and gene-flow between different classes of individuals.
-That is, genetic trade-offs between fitness components can constrain adaptation \citep{CharlesworthHughes2000, ConnallonHall2018}.
-On the one hand, such constraints prevent individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase the population's overall extinction risk \citep{holman2013consequences,harts2014demography}.
+Fitness trade-offs constraining adaptation have particularly interesting evolutionary consequences.
+On the one hand, they prevent individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase the population's overall extinction risk \citep{holman2013consequences, harts2014demography}.
 On the other hand, they provide an effective mechanism for the maintenance of genetic variation, which influences the population's capactiy for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
@@
 Classical population genetic models consider relative fitnesses of genotypes, and therefore do not consider the population dynamical consequences of evolutionary change.
-To study the population dynamical consequences of evolutionary change, a model is required that links changes in genotype distributions to population dynamics.
+To take the population dynamical consequences into account, a model is required that links changes in genotype distributions to population dynamics.
 The demographic processes of birth and death drive both changes in genotype distributions and population dynamics, and therefore provide a fundamental link between evolution and ecology \citep{metcalf2007evolutionary}.
 
-In this paper, we use such an evolutionary demographic approach (model?) to study whether the parameter space predicted to be polymorphic by population genetics models is, indeed, demographically viable.
-We apply the model to study the maintenance of polymorphisms through sexually antagonistic selection in both dioecious and hermaproditic species.
+In this paper, we use an evolutionary demographic model to study when the parameter conditions predicted to be polymorphic by population genetics models are also demographically viable.
+We apply the model to study the maintenance of polymorphisms by sexually antagonistic selection in both dioecious and hermaproditic species.
 A major strength of the model is that we can use empirically estimated demographic rates to parameterize the model, enabling us to make predictions about the scope for demographically viable SA polymorphisms that are grounded in the biology of real populations.
@@
 %Genetic trade-offs between fitness components due to pleiotropy represent a form of genetic constraint with particularly interesting evolutionary consequences because they act as a double-edged sword
-%The nature and extent of genetic variation observed in natural populations should therefore reflect the balance between the contribution of those constraints to the maintainance of genetic polymorphism and the demographic consequences of the resulting maladaptation.
 
@@
 When such genetic trade-offs occur, "sexually antagonistic" selection (abbreviated SA hereafter) can arise whereby beneficial alleles for one sex are deleterious when expressed in the other \citep{Kidwell1977, Rice1992, ConnallonClark2012}.
-Such alleles with opposing fitness effects through male and female sex functions create analogous genetic constraints on fitness in hermaphrodite populations, where both maternal and paternal reproductive success contribute jointly to each individuals' overall fitness \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
+Alleles with opposing fitness effects through male and female sex functions can cause analogous genetic constraints on fitness in hermaphrodite populations, where both maternal and paternal reproductive success contribute jointly to each individuals' overall fitness \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
 
@@
 						\sigma_j(1 - \gamma) & 0 \\
-						\sigma_j \gamma      & \sigma_A
+						\sigma_j \gamma      & \sigma_a
 					\end{array}
@@
 For simplicity, we assume survival and transition rates are constant among genotypes.
+
+Throughout our analyses, we distinguish between early- and late-acting inbreeding depression.
+We quantify early-acting inbreeding depression using $\delta$, and late-acting inbreeding depression using $\delta_i$ (where $i \in \{j,a,\gamma\}$).
+An important difference between early- and late-acting inbreeding depression in the model is that $\delta$ does not affect $\mbf{U}^S$, the survival matrix for offspring produced by selfing (see Eq.~\ref{eq:Atilde}), while the $\delta_i$ terms directly alter the stage-specific survival and transition rates in $\mbf{U}^S$ (see Eq.~\ref{eq:US}).
 
@@
  \centering
- \caption{\bf Fitness expressions for Sexually Antagonistic selection in hermaphrodites ()}
+ \caption{\bf Fitness expressions for Sexually Antagonistic selection ($w_{i}$)}
 \label{tab:Fitness}
@@
 
-Unless stated otherwise, we use the following parameter values for the demographic rates in the model: $\sigma_j = \sigma_A = 0.6$ and $\gamma = 0.05$, which are similar to those used in \citet{deVriesCaswell2019b}.
+Unless stated otherwise, we use the following parameter values for the demographic rates in the model: $\sigma_j = \sigma_a = 0.6$ and $\gamma = 0.05$, which are similar to those used in \citet{deVriesCaswell2019b}.
 Our parameters of interest include fertility, $f$, the inbreeding depression parameters, $\delta$, $\delta_j$, $\delta_a$, $\delta_{\gamma}$, and the selection parameters $h_f$, $s_f$, $h_m$, and $s_m$, which are given different values for each analyses as described in the figure captions.
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Analyses} \label{sec:analyses}
+\subsection*{Analyses} \label{subsec:analyses}
 
@@
 
+For analyses of the demographic effects of inbreeding depression, we made two main simplifying assumptions: First, we acknowledge that if inbreeding depression is caused primarily by recessive deleterious mutations (as suggested by empirical data), it should covary negatively with the population selfing rate due to purging (\citealt{Charlesworth2009}; though we note that other processes could give rise to this pattern, e.g., \citealt{CrnokrakBarrett2002, Charlesworth2009,HedrickGarcia-Dorado2016}).
+Following \citet{Olito2019}, we incorporate such negative covariance by constraining the inbreeding depression terms in the model ($\delta$ and $\delta_i$, where $i \in \{J,A,\gamma\}$) to follow a simple declining function of the selfing rate: $\delta = \delta^{\ast} (1 - b (1 - L))$, where $\delta^{\ast}$ is the hypothetical severity of inbreeding depression for a completely outcrossing population, $b$ is a shape parameter determining how far $\delta$ will decline under complete selfing (when $C = 1$), and $L$ is an expression describing the expected deleterious mutation load due to recessive deleterious mutations that includes an additional shape parameter, $a$, which determines the curvature of the overall function for $\delta$ (see Appendix E in \citealt{Olito2019} for additional details).
+We set $\delta^{\ast} = 0.8$, $b = 0.5$, and $a = 0.2$ for all analyses, values chosen to be consistent with empirical estimates of inbreeding depression (e.g., fig.~2 in \citealt{HusbandSchemske1996}).
+Second, we explore the effects of individual inbreeding depression terms in isolation.
+That is, we assume that only one of the terms $\delta$ and $\delta_i$ (where $i \in \{J,A,\gamma\}$) can be non-zero at a time.
+
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Demographic and fitness data for {\itshape M. guttatus}} \label{sec:MguttMethods}
+\subsection*{Demographic and fitness data for {\itshape M. guttatus}} \label{subsec:MguttMethods}
 
@@
 {\itshape M.~guttatus} is an herbaceous, self-compatible wildflower native to western North America that exhibits remarkable among-population variation in numerous life-history and reproductive traits including selfing rates, inbreeding depression, floral morphology, and annual-to-perennial life-history \citep[e.g.,][]{RitlandGanders1987, Ritland1990, Willis1993, Willis1999a, Willis1999b, WuWillis2008}.
-Moreover, detailed demographic studies have been conducted on multiple populations of {\itshape M.~guttatus}, with demographic matrix data made publicly available on the demographic database COMPADRE \citep{CompadreDB2020}.
+Moreover, detailed demographic studies have been conducted on multiple populations of {\itshape M.~guttatus}, with demographic matrix data available on the public demographic database COMPADRE \citep{CompadreDB2020}.
 Below, we briefly outline how we parameterized our model using the available data; full details are provided in \hl{Appendix B.x}.
 
-We used extensive demographic rate data reported for the 'Eagle Meadows' population in Stanislaus National Forest (California, USA) studied by \citet{PetersonEtAl2016}.
-This study entailed a large common garden experiment which took place at the 'Eagle Meadows' site.
-We used demographic rates estimated for this population grown in its native habitat (i.e., we did not use data for non-local populations grown at Eagle Meadows).
-\citet{PetersonEtAl2016} collected data for numerous life-history stages and transition rates; for the Eagle Meadows population these estimates were as follows: seed bank survival ($D = 0.534$), seed germination rate ($G = 0.469$), flower production ($F = 0.64$), ovules per flower ($O = 614)$, seedling recruits proportional to clonal rosette recruits ($A = 6.7e-4$), overwinter survival ($S = 0.179$), rosette production ($R = 8.71$) (see corrected Tables.~1 and S2 in \citealt{PetersonEtAl2017}).
-These data were also available through the Compadre demographic matrix database \citep{CompadreDB2020}.
-The resulting transition matrices for this population involved three life-history stages ($\omega = 3$; seed, seedling, and rosette), and individual elements of the transition matrix ($\tilde{\mbf{A}}$) were calculated as products of the above rates (see Matrix 1 in \citealt{PetersonEtAl2016}).
-
-Unfortunately, estimates of selfing rates and inbreeding depression are not also available for the Eagle Meadows population of {\itshape M.~guttatus}.
-However, selfing rate estimates are available for a variety of other populations, and vary in magnitude from near complete outcrossing to predominant selfing ($C \approx 0$ to $0.75$; \citealt{RitlandGanders1987, Ritland1990, Willis1999b}).
-Additionally, estimates of inbreeding depression at several of the life-history stages/fitness components that were included in the data of \citet{PetersonEtAl2016} are available for two intensively studied populations in the Cascade Mountains of Oregon (Iron Mountain and Cone Peak) \citet{Willis1993, Willis1999a, Willis1999b}.
+We used extensive demographic data reported in a large-scale study of local adaptation using experimental populations of {\itshape M.~guttatus} in Stanislaus National Forest (California, USA) in 2012 and 2013 \citep{PetersonEtAl2016}.
+We leverage their common-garden experimental design to focus on a comparison of demographically viable polymorphic parameter space for two experimental populations with contrasting demographic rates.
+The first was a locally adapted 'Eagle Meadows' population (data from 2012), while the second group was an experimental population composed of multiple non-locally adapted 'low-elevation perennials' (data from 2013).
+The vital rate estimates for the Eagle Meadows population are as follows: seed bank survival ($D = 0.534$), seed germination rate ($G = 0.469$), flower production ($F = 0.64$), ovules per flower ($O = 614$), seedling recruits proportional to clonal rosette recruits ($A = 6.7 \times 10^{-4}$), overwinter survival ($S = 0.179$), rosette production ($R = 8.71$).
+The corresponding estimates for the low-elevation perrenials are: $D = 0.534$, $G = 0.652$, $F = 0.4.09$, $O = 494$, $A = 6.7e-4$), $S = 0$, and $R = 0$ (see corrected Tables.~1 and S2 in \citealt{PetersonEtAl2017}).
+The same estimates for $D$ and $A$ were used for all populations.
+The resulting transition matrices for this population involved three life-history stages ($\omega = 3$; seed, seedling, and rosette), and individual elements of the transition matrix ($\tilde{\mbf{A}}$) were calculated as products of the above rates (see Matrix $1$ in \citealt{PetersonEtAl2016}).
+
+Estimates of selfing rates and inbreeding depression were not available for the same experimental populations, but are available for a variety of other western USA {\itshape M.~guttatus} populations.
+Selfing rate estimates vary in magnitude from near complete outcrossing to predominant selfing ($C \approx 0$ to $0.75$; \citealt{RitlandGanders1987, Ritland1990, Willis1999b}).
+Estimates of inbreeding depression at several of the life-history stages/fitness components that were included in the data of \citet{PetersonEtAl2016} are available for two intensively studied populations in the Cascade Mountains of Oregon (Iron Mountain and Cone Peak) \citet{Willis1993, Willis1999a, Willis1999b}.
 Using the data provided in \citet{Willis1993}, we estimated the proportional decrease due to inbreeding depression in seed germination rate ($\delta_{G} = 0.085$), flower number ($\delta_{F} = 0.2$), and overwinter survival ($\delta_{S} = 0.38$).
+The largest field-estimated selfing rate for this same Iron Mountain population was $C = 0.29$ \citep{Willis1993}.
 
 Using these combined demographic rates, selfing rate, and inbreeding depression estimates, we constructed a corresponding stage $\times$ genotype mendelian matrix model with a single SA locus affecting female and male fertility (as described above).
-Using this empirically parameterized model, we are able to make predictions about the scope for demographically viable SA polymorphism at a single locus in a hypothetical population with the same demographic rates as observed in the Eagle Meadows population, for a range of selfing and inbreeding depression rates observed in other natural populations.
-However, this should not be confused with an explicit prediction for the likelihood of SA polymorphism in the Eagle Meadows population.
-
-Interestingly, a polymorphic chromosomal inversion (Inv6) with apparently SA fitness effects has been identified in the Iron Mountain population of {\itshape M.~guttatus} \citep{LeeKelly2015}.
-Inv6 segregates at moderate frequency (about $8\%$), and carriers suffer $\approx 30 \%$ loss in pollen viability, but also increased flower production which varies among years.
-The genetic basis of the fitness effects caused by Inv6 are not yet known, but the net result is a "supergene" with uniquely strong fitness effects that segregates as a single diallelic locus.
-We estimated dominance and selection coefficients for the effect of Inv6 on pollen viability and flower production from greenhouse crossing experiments using mapping populations and field estimates.
-Across four mapping populations, the effect of Inv6 on pollen viability was partially recessive ($h_m \approx 0.35$).
+Using this empirically parameterized model, we are able to make predictions about the scope for demographically viable SA polymorphism at a single locus in hypothetical populations with the same demographic rates as observed in \citet{PetersonEtAl2016}, for a range of selfing and inbreeding depression rates observed in other natural populations.
+We stress, however, that these illustrative predictions are not explicit predictions of the likelihood of SA polymorphism in any specific population.
+
+Interestingly, a polymorphic chromosomal inversion (inv6) with apparently SA fitness effects has been identified in the Iron Mountain population of {\itshape M.~guttatus} \citep{LeeKelly2015}.
+inv6 segregates at moderate frequency (about $8\%$), and carriers suffer $\approx 30 \%$ loss in pollen viability, but also increased flower production which varies among years.
+The genetic basis of the fitness effects caused by inv6 are not yet known, but the net result is a "supergene" with strong fitness effects that segregates as a single diallelic locus.
+We estimated dominance and selection coefficients for the effect of inv6 on pollen viability and flower production from the data reported in \citet{LeeKelly2015}.
+Across four mapping populations, the effect of inv6 on pollen viability was partially recessive on average ($h_m \approx 0.35$).
 Given this average dominance, the selection coefficient in the field is approximately $s_m = 0.88$.
-It was not possible to estimate the average dominance of the effect of Inv6 on flower production, so we conservatively assume that it was additive ($h_f = 0.5$).
-Based on this assumption, the selection coefficient Inv6 on flower production in the field was $s_f = 0.384$ in 2012, and $s_f = 0.232$ in 2013.
-As a final proof-of-concept test of our empirically parameterized model \citep[e.g.,][]{Servedio2014}, we asked whether, given the above biologically grounded fitness effect estimates, Inv6 appears to fall in demographically viable SA polymorphic parameter space, as might be expected given its observed frequency in the Iron Mountain population.
+It was not possible to estimate the average dominance of the effect of inv6 on flower production (inv6 homozygotes were too rare in the field), so we conservatively assume that it was additive ($h_f = 0.5$).
+Based on this assumption, the selection coefficient inv6 on flower production in the field was $s_f = 0.384$ in 2012, and $s_f = 0.232$ in 2013.
+As a final proof-of-concept test of our empirically parameterized model \citep[e.g.,][]{Servedio2014}, we asked whether, given the above biologically grounded fitness effect estimates, inv6 appears to fall in demographically viable SA polymorphic parameter space, as might be expected given its observed frequency in the Iron Mountain population.
 
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Results}
+\section*{Results}\label{sec:Results}
 %%%%%%%%%%%%%%%%%%%%%%%%
-\ldots
 
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Polymorphism and Extinction} \label{sec:Res:PolyExt}
-
+\subsection*{Polymorphism and Extinction}\label{subsec:PolyExt}
+
+%%%%%%%%%%%%%%%%%%%%%
+% Figure 1
  \begin{figure}[htbp]
@@
 However, for some parameter conditions, populations will ultimately go extinct (red shaded regions) due to reduced female fitness resulting from the male-beneficial/female-deleterious allele that is either segregating as a balanced polymorphism (inside the funnel), or becomes fixed (area below the funnel).
-We are particularly interested in identifying and quantifying "demographically viable polymorphic parameter space", which corresponds to the area inside the funnel that is also to the left of the extinction threshold for a given fertility value.
-Results are shown for three different population selfing rates ($C = \{0,\,1/4,\,1/2\}$), and two dominance scenarios (additivity with $h = 1/2$, and dominance reversal with $h = 1/4$); extinction thresholds are illustrated for three different values of female fecundity ($f$ values annotated on each panel).}}
+"Demographically viable polymorphic parameter space" corresponds to the area inside the funnel that is also to the left of the extinction threshold for a given fertility value.
+Results are shown for three different population selfing rates ($C = \{0,\,1/4,\,1/2\}$), and two dominance scenarios (additivity, where $h = 1/2$, and dominance reversal, where $h = 1/4$); extinction thresholds are illustrated for three different values of female fecundity ($f$ values annotated on each panel).}}
  \label{fig:extThresholds}
  \end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Polymorphism in the absence of inbreeding depression} \label{sec:Res:PolyNoID}
-
+%%%%%%%%%%%%%%%%%%%%%
+
+We begin with a simple illustration of demographically viable polymorphic parameter space in the absence of inbreeding depression (i.e., $\delta_i = 0$).
+Invasion conditions for SA alleles in the evolutionary demographic model closely match the predictions from population genetic models (see \hl{Online Appendix X} \citealt{Kidwell1977, JordanConnallon2014, Olito2017}).
+In particular, the demographic model recovers the classic "funnel-shaped" region of polymorphic $s_f \times s_m$ parameter space.
+The effects of the population selfing rate ($C$) and dominance ($h$) on SA polymorphism are also similar: self-fertilization ({\itshape i}) skews the SA polymorphic space towards regions with weaker selection through female-function and reduces the total area of polymorphic parameter space; and ({\itshape ii}) dominance reversals (where deleterious SA fitness effects are partially recessive in each sex; $h = 1/4$) are much more permissive of SA polymorphism \citep{JordanConnallon2014, Olito2017}.
+
+However, a key prediction from the evolutionary demographic model is that large fractions of SA polymorphic parameter space can be demographically inviable (Fig.~\ref{fig:extThresholds}).
+The location of the extinction threshold, where the population intrinsic growth rate $\lambda = 1$, is primarily determined by the fertility parameter ($f$) but is also influenced by the population selfing rate ($C$) and dominance of the SA alleles ($h$).
+When $f$ is large, the population can sustain higher mortality caused by stronger selection through both male and female fitness components before extinction occurs (the extinction threshold shifts to higher values of $s_f$ and $s_m$).
+
+The demographic effects of SA polymorphism are driven by the fitness effects of segregating male-beneficial/female-deleterious alleles because they directly influence individuals' fecundity.
+For the fertility parameter values we explored, all extinctions occured in regions where either the female-deleterious allele segregates at intermediate frequency (i.e., where SA polymorphism is maintained), or goes to fixation (fig.~\ref{fig:extThresholds}).
+The SA polymorphisms that remained viable at lower fertilitiy values corresponded to regions where the female-deleterious allele is predicted to segregate at low frequencies, or in regions with weak selection through both sex functions.
+
+In populations with high fertility (larger $f$), the proportion of demographically viable polymorphic parameter space converges on the predictions for total SA polymorphic space in population genetic models (fig.~\ref{fig:polySpace}).
+In obligately outcrossing populations (including dioecious/gonochoristic populations; where $C = 0$), lower fertility can result in a significant reduction of demographically viable polymorphic parameter space.
+However, the effect is weaker in populations with intermediate selfing rates ($C > 0$) because self-fertilization offers reproductive assurance for a fraction ($C$) of each individual's ovules.
+This allows partially sefing populations to remain viable under selection intensities that would cause extinction in an outcrossing population.
+Interestingly, when fertility is low enough that some polymorphic parameter space is demographically inviable, populations with intermediate selfing rates having the greatest proportion of demographically viable parameter space (fig.~\ref{fig:polySpace}, med.~and low.~fertility values).
+This effect emerges from the combination of increasing population viability due to reproductive assurance and reduced total polymorphic parameter space caused by selfing.
+
+% Figure 2
  \begin{figure}[htbp]
  \centering
- \includegraphics[scale=0.75]{../output/figs/polymorphicSpaceNew.pdf}
+ \includegraphics[scale=0.75]{../output/figs/polymorphicSpaceTitrate.pdf}
  \caption{\footnotesize{Proportion of demographically viable parameter space (out of total $s_f \times s_m$ space with $\max(s) = 0.15$) in the absence of inbreeding depression (i.e., assuming $\delta = \delta_i = 0$, where $i \in \{j,a,\gamma\}$), plotted as a function of the population selfing rate.
-Results are shown for three fertility values corresponding to low, medium, and high fertility (blue, green, and red points respectively) under additive (panel A; $h = 1/2$), and partially recessive (panel B; $h = 1/4$) SA fitness effects.
-Each point was calculated by numerical integration of the corresponding SA invasion conditions and extinction threshold predicted by the mendelian matrix model (see \hyperref[sec:analyses]{Analyses} section), while solid lines were produced by numerically integrating the analytic expressions for the single-locus invasion conditions from the population genetic models of \citet{JordanConnallon2014} and \citet{Olito2017} (solid black lines).}}
+Results are shown for three fertility values corresponding to low, medium, and high fertility (blue, green, and red points respectively) under additive ($h = 1/2$; panel A), and partially recessive ($h = 1/4$; panel B) SA fitness effects.
+Each point was calculated by numerical integration of the corresponding SA invasion conditions and extinction threshold predicted by the mendelian matrix model (see \hyperref[subsec:analyses]{Analyses} section), while solid lines were produced by numerically integrating the analytic expressions for the single-locus invasion conditions from the population genetic models of \citet{JordanConnallon2014} and \citet{Olito2017} (solid black lines).}}
  \label{fig:polySpace}
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Demographic effects of inbreeding depression} \label{sec:Res:InbreedEffects}
-
+\subsection*{Demographic effects of inbreeding depression} \label{subsec:InbreedEffects}
+
+Unlike previous population genetic models which assumed constant population sizes \citep{JordanConnallon2014,Olito2017}, mortality caused by inbreeding depression can strongly influence population persistence in our evolutionary demographic model.
+Populations with high fertility rates can sustain positive population growth rates despite this higher mortality.
+This leads to a greater proportion of SA polymorphic parameter space to be demographically viable, with the demographic model predictions converging on those from population genetic models in high-fertility populations (fig.~\ref{fig:deltaPolySpace}).
+For populations with lower fertility rates, the proportion of demographically viable polymorphic parameter space matches the population genetic model predictions with the selfing rate is low; however, as the selfing rate increases, demographic viability eventually crashes when the population can no longer sustain the concomitant increase in mortality due to inbreeding depression (fig.~\ref{fig:deltaPolySpace}).
+
+Mortality from inbreeding depression reduces the effect of "reproductive assurance" individuals gain through self fertilization.
+In contrast to our earlier results (see \hyperref[subsec:PolyExt]{Polymorphism and Extinction}), predominantly outcrossing populations are predicted to have the highest proportion of demographically viable polymorphic space, when inbreeding depression is taken into acount (compare fig.~\ref{fig:polySpace} with fig.~\ref{fig:deltaPolySpace}).
+Regardless of the life-history stage at which inbreeding depression affects survival, populations with intermediate to high selfing rates are unlikely to harbour SA polymorphism unless they can "afford" the resulting loss of self-fertilized ovules/offspring.
+
+% Figure 3
  \begin{figure}[htbp]
@@
 In all plots, the strength of inbreeding depression decreases as the selfing rate goes up following a simple model of purging recessive deleterious mutations (see \hl{Appendix X.x} for details).
-Only single, $\delta_i$ terms indicated in the legend ($i \in \{j,a,\gamma\}$) are allowed to vary at once (all others are set to $0$).
-Effects of early-acting inbreeding depression are shown on the left (panels A, C; $\delta$) and for late-acting inbreeding depression on the right (panels B, D; $\delta_i$) for additive ($h = 1/2$) and partially recessive ($h = 1/4$) SA fitness effects, respectively.
-All results used the same 'high female fertility' value of $f = 6.5$ (see Fig.\ref{fig:polySpace}).
-Each point was calculated by numerically integrating the corresponding SA invasion conditions and extinction threshold predicted by the model (see \hyperref[sec:analyses]{Analyses} section), while solid lines were produced by numerically integrating the analytic expressions for the single-locus invasion conditions from the population genetic models of \citet{JordanConnallon2014} and \citet{Olito2017} (solid black lines).
-The population genetic model predictions for varying $delta$ are plotted on both left- and right-hand panels for comparison, and to highlight the different y-axes scales.}}
+Only single inbreeding depression terms ($\delta$ and $\delta_i$, where $i \in \{j,a,\gamma\}$, indicated in the legend) are allowed to vary at one time (all others are set to $0$).
+Results are shown for three fertility values ($f = \{6.5,\,7.5,\,8.5\}$) under additive ($h = 1/2$; panel A) and partially recessive ($h = 1/4$; panel B) SA fitness effects.
+Each point was calculated by numerically integrating the corresponding SA invasion conditions and extinction threshold predicted by the model (see \hyperref[subsec:analyses]{Analyses} section), while solid lines were produced by numerically integrating the single-locus invasion conditions from the population genetic models of \citet{JordanConnallon2014} and \citet{Olito2017} (solid black lines).}}
  \label{fig:deltaPolySpace}
  \end{figure}
+
+However, when during the life-cycle inbreeding depression manifests does influence the threshold selfing rate at which demographically viable polymorphic parameter space crashes.
+Our results indicate that population viability was most sensitive to inbreeding depression affecting juvenile survival rates ($\delta_j$; fig.~\ref{fig:deltaPolySpace}, dark blue points), while early-acting inbreeding depression ($\delta$, ovule abortion shortly after fertilization) had a similar effect on population viability as late-acting inbreeding depression affecting adult survial ($\delta_a$) and juvenile-to-adult transition rates ($\delta_{\gamma}$).
+Inbreeding effects on juvenile survival had the strongest effect on population viability because of the sequential nature of life-cycle stages (individuals must survive the juvenile stage in order to reproduce as adults) and the multiplicative fitness effects of the survival and transition rates in the selfed-offspring survival matrix $\mbf{U}^{S}$.
+Although early-acting inbreeding depression ($\delta$) actually manifests earlier in the life-cycle than juvenile survival, it acts only once by influencing the total number of self-fertilized zygotes that become juveniles, and does not act multiplicatively with the other terms in $\mbf{U}^{S}$.
+
+%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Polymorphism in M.~guttatus} \label{sec:Mgutt}
+
+The Eagle Meadows and Low-Elevation Perennial populations of \citet{PetersonEtAl2016} had contrasting demographic rates that strongly affected the scope for demographically viable SA polymorphism.
+The Eagle Meadows population was locally adapted with a very high intrinsic growth rate ($\lambda \approx 1.7$).
+This growth rate was high enough that all $s_f \times s_m$ selection parameter space (where $s_f, s_m \in (0,1]$) remained demographically viable, regardless of the selfing rate ($C = 0,0.29$), or the presence of inbreeding depression (fig.~\ref{fig:mimulusFig}A).
+In contrast, the Low-Elevation Perennial population had a much lower intrinsic growth rate, although it was still positive ($\lambda \approx 1.08$).
+Due to the slower growth rate, not all of the $s_f \times s_m$ selection parameter space was demographically viable.
+Extinction thresholds appeared under complete outcrossing ($C = 0$) and partial selfing with inbreeding depression ($C = 0.29$; $\mbf{U}^{S}$), but not when inbreeding depression was omitted (fig.~\ref{fig:mimulusFig}B).
+The different demographic rates from these two populations also resulted in slightly different invasion conditions when inbreeding depression was taken into account (compare dotted lines in fig/~\ref{fig:mimulusFig}A and B).
+
+Interestingly, the polymorphic inversion inv6 falls very near the lower edge of polymorphic parameter space for these hypothetical populations, but is not always predicted to be demographically viable.
+When using demographic rates for either the Eagle Meadows or Low-Elevation Perennial populations, inv6 falls within polymorphic space when the selfing rate is at the higher end of empirical estimates for the Iron Mountain population in which inv6 has been documented ($C = 0.29$), but only when inbreeding depression is not taken into account (fig.~\ref{fig:mimulusFig} dashed lines).
+When including empirical estimates of inbreeding depression, the population effective selfing rate is reduced, the polymorphic parameter space shifts upwards, and inv6 no longer falls within the polymorphic space.
+When using vital rates for the locally adapted Eagle Meadows population, the model predicts that an allele (or supergene) with the same fitness effects as inv6 will go extinct.
+In contrast, when using the demographic rates for Low-Elevation Perennials, an allele/supergene with these fitness effects falls within demographically inviable parameter space.
+That is, the model predicts that SA alleles with selection coefficients of similar magnitude to inv6 will inflict too high of a demographic cost, and are unlikely to be observed in populations with lower intrinsic growth rates like those observed in the Low-Elevation Perennial population.
+
+% Figure 4
+\begin{figure}[htbp]
+ \centering
+ \includegraphics[scale=0.75]{../output/figs/mimulusinv6Fig.pdf}
+ \caption{\footnotesize{Illustration of model predictions using empirically estimated demographic rates for {\itshape M.~guttatus}.
+Results are shown for two hypothetical populations using demographic rates for locally adapted (Eagle Meadows; panel A), and non-local (Low-Elevation Perennial; panel B) populations reported by \citet{PetersonEtAl2016}.
+Invasion conditions and extinction thresholds were calculated using the dominance coefficients estimated for inv6 ($h_f = 0.5, h_m = 0.35$; from data in \citealt{LeeKelly2015}) for three parameter conditions: obligate outcrossing (solid black lines), partial selfing using the highest field-estimate for the Iron Mountain population of \citet{Willis1993} (dashed lines), and partial selfing with inbreeding depression (dotted lines; using $\delta_i$ estimates for Iron Mountain by \citealt{Willis1993}).
+The location of inv6 is also shown on both plots, using estimated selection coefficients.
+Note that in panel A the predictions for partial selfing with inbreeding depression (dotted lines) fall extremely close to the predictions for obligate outcrossing (solid lines).}}
+ \label{fig:mimulusFig}
+ \end{figure}
+
+\hl{[SPECULATION!!!]} Finally, the location of inv6 in selection parameter space is interesting itself because in this region an SA allele with net female-beneficial fitness effects similar to inv6 are predicted to either go extinct, or segregate at relatively low frequency, depending on the population selfing-rate and severity of inbreeding depression.
+Although we cannot make specific predictions for inv6 in the Iron Mountain population in which it was observed, it is interesting that the estimated fitness effects place inv6 roughly in parameter space where segregation at low to moderate frequencies is not surprising.
+This rough concordance between theory and data is clearly speculative, but to our knowledge.
 

### 2021-01-27 R/functions-Figs.R

@@ -527,7 +527,7 @@ polySpaceFigTitrate  <-  function(df = "dataPolySpaceFig_sMax0.15_res0.003_delta
         proportionalLabel(0.03, 1.075, 'A', cex=1.2, adj=c(0.5, 0.5), xpd=NA)
 #        proportionalLabel(0.5, 1.1, substitute(paste(italic(h), " = ", hh), list(hh = pars2$hf)), cex=1.2, adj=c(0.5, 0.5), xpd=NA, srt=0)
         proportionalLabel(0.5, 1.1, expression(paste(italic(h), " = ", 1/2)), cex=1.2, adj=c(0.5, 0.5), xpd=NA, srt=0)
-        proportionalLabel(-0.3, 0.5, expression(paste("Prop. viable polymorphic space")), cex=1, adj=c(0.5, 0.5), xpd=NA, srt=90)
+        proportionalLabel(-0.3, -0.16, expression(paste("Proportion viable polymorphic parameter space")), cex=1, adj=c(0.5, 0.5), xpd=NA, srt=90)
 
         #Legend
         legend( x       =  usr[2],
@@ -579,7 +579,7 @@ polySpaceFigTitrate  <-  function(df = "dataPolySpaceFig_sMax0.15_res0.003_delta
 #        proportionalLabel(0.5, 1.1, substitute(paste(italic(h), " = ", hh), list(hh = pars5$hf)), cex=1.2, adj=c(0.5, 0.5), xpd=NA, srt=0)
         proportionalLabel(0.5, 1.1, expression(paste(italic(h), " = ", 1/4)), cex=1.2, adj=c(0.5, 0.5), xpd=NA, srt=0)
         proportionalLabel(0.5, -0.3, expression(paste("Selfing rate (",italic(C), ")")), cex=1.2, adj=c(0.5, 0.5), xpd=NA, srt=0)
-        proportionalLabel(-0.3, 0.5, expression(paste("Prop. viable polymorphic space")), cex=1, adj=c(0.5, 0.5), xpd=NA, srt=90)
+#        proportionalLabel(-0.3, 0.5, expression(paste("Prop. viable polymorphic space")), cex=1, adj=c(0.5, 0.5), xpd=NA, srt=90)
 
         #Legend
         legend( x       =  usr[2],
@@ -612,32 +612,56 @@ polySpaceFigTitrate  <-  function(df = "dataPolySpaceFig_sMax0.15_res0.003_delta
 #' Fig showing Effect of Inbreeding Depression on 
[…]

### 2021-01-27 R/functions-Simulations.R

@@ -349,11 +349,13 @@ fwdDyn2Eq  <-  function(nzero, om, g, W_prime, blkFX_prime, blkUS, blkUX, W, Iom
 			pnext  <-  n/norm(as.matrix(n), type="1")
 			pcheck <-  p
 			p      <-  pnext
+			nout[,i]  <-  nnext
 		} else{
 			n      <-  nnext
 			pnext  <-  n/norm(as.matrix(n), type="1")
 			pcheck <-  p
 			p      <-  pnext
+			nout[,i]  <-  nnext
 		}
 
 		# check to see if equilibrium has been reached
@@ -1784,7 +1786,7 @@ points(	titrateVals[,2] ~ 	titrateVals[,1])
 #' C:		Population selfing rate
 #' delta:	Early-acting inbreeding depression (proportion ovules aborted due to I.D.)
 #' delta_j:	Proportional decrease in juvenile survival rate for inbred offspring
-#' delta_j:	Proportional decrease in adult survival rate for inbred offspring
+#' delta_a:	Proportional decrease in adult survival rate for inbred offspring
 #' delta_gamma: Proportional decrease in juv. --> adult transition rate for inbred offspring
 #' tlimit:	Max # of generations for fwd simulation
 #' Ainvade:	Should A allele start off rare
@@ -2345,7 +2347,8 @@ makeDataDeltaPolyParamSpace  <-  function(sMax=0.15, res=0.0015, precision = 1e-
 
 		cat('\r', "hVal = ", hVals[i], '\n')
 
-		Cs  <- c(seq(0, 0.5, by=0.025), 0.75, 0.9)
+#		Cs  <- c(seq(0, 0.5, by=0.025), 0.75, 0.9)
+		Cs  <- seq(0, 0.9, by=0.025)
 		
 		# Define dStar ~ C
 		deltaSeq  <-  predDelta(dStar=dStar, b=1/2, a=0.2, C=Cs)
@@ -2712,10 +2715,9 @@ quantDeltaPolySpace  <-  function(data, pars) {
 #' eqThreshold:	threshold per gen. change in allele frequencies for determining demographic equilibrium
 #' Ainvade:	Should A allele start off rare
 #' intInit:	Optional initial frequency of A allele
-#' STILL NOT WORKING
[…]

### 2021-01-27 makeFigs.R

@@ -40,6 +40,11 @@ toPdf(deltaSelfingLoadPolySpaceFigTitrate(),
 			figPath(name='deltaPolymorphicSpaceTitrate.pdf'), width=5, height=7)
 embed_fonts(figPath(name='deltaPolymorphicSpaceTitrate.pdf'))
 
+# Fig. 4 - Illustration of polymorphic parameter space w/ 
+#			Inv6 plotted
+toPdf(MimulusInv6Fig(), 
+			figPath(name='MimulusInv6Fig.pdf'), width=10, height=5)
+embed_fonts(figPath(name='MimulusInv6Fig.pdf'))
 
 #################
 # Supp. Figs

### 2021-01-27 run-Simulations.R

@@ -267,6 +267,16 @@ makeDataDeltaPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
 							hVals= c(1/2, 1/4), dStar = 0.8, 
 							tlimit = 10^5, eqThreshold = 1e-8)
 
+makeDataDeltaPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
+							om = 2, g = 3, theta = c(0.6,0.6,0.05,7.5), theta_prime = 7.5, 
+							hVals= c(1/2, 1/4), dStar = 0.8, 
+							tlimit = 10^5, eqThreshold = 1e-8)
+
+makeDataDeltaPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
+							om = 2, g = 3, theta = c(0.6,0.6,0.05,8.5), theta_prime = 8.5, 
+							hVals= c(1/2, 1/4), dStar = 0.8, 
+							tlimit = 10^5, eqThreshold = 1e-8)
+
 
 ######################
 # Preliminary figures
@@ -805,6 +815,7 @@ deltaSelfingPolySpaceMakeData(sMax = 0.15, nSamples=1e+3,
 #################################################
 #' Find extinction threshold for Mimulus example
 
+# Eagle Meadows population  (Peterson et al. 2016)
 datMat  <-  matList_Mg_EM
 theta.list  <-  list(D = 0.534, 
 					 G = 0.469,
@@ -812,7 +823,8 @@ theta.list  <-  list(D = 0.534,
 					 O = 614,
 					 A = 6.7e-4,
 					 S = 0.179,
-					 R = 8.71)
+					 R = 8.71,
+					 pop = "EM")
 delta.list  <-  list(delta_D = 0,
 					 delta_G = 0,
 					 delta_F = 0,
@@ -820,28 +832,222 @@ delta.list  <-  list(delta_D = 0,
 					 delta_S = 0
 					 )
 
+
+## Obligate oucrossing
+titrateInvBoundMimulus(sMax = 0.99, res=0.01, precision=1e-4, 
+						datMat=datMat, theta.list=theta.list, delta.list=delta.list, useCompadre = FALSE,
+						hf=1/2, hm=0.35, C=0, 
+						tlimit = 10^5, eqThreshold = 1e-8, writeFile=TRUE)
 extinctThreshMimulus(sMax = 0.99, res=0.01, precision=1e-4, 
[…]

### 2021-02-16 doc/SA-Hermaphrodites-wDemography.tex
Edits plus discussion
Edited main text, added first draft of discussion, moved some material to appendix

@@
 \usepackage{nameref}
+%\usepackage[title,titletoc,toc]{appendix}
 \usepackage[colorlinks=true, allcolors=black]{hyperref}
@@
  \newcommand{\kron}{\otimes}
+
  %===========================================
@@
 
-%\section*{Outline}
+%\section{Outline}
 
@@
 
+\tableofcontents
 %%%%%%%%%%%%%%%%%%%%
 \newpage{}
-\section*{Abstract}
+\section{Abstract}
 %%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Introduction}
+\section{Introduction}
 %%%%%%%%%%%%%%%%%%%%%%%%
 
- Whether populations persist or go extinct depends heavily on their ability to adapt to the environment.
+Population persistence or extinction depends heavily on the ability of a species to adapt to the environment.
 Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, such as deleterious mutations \citep{Haldane1957}, environmental shifts \citep{Maynard-Smith1976, LandeShannon1996,OrrUnckless2008}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, or genetic constraints \citep{ConnallonHall2018}.
@@
 Fitness trade-offs constraining adaptation have particularly interesting evolutionary consequences.
-On the one hand, they prevent individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase the population's overall extinction risk \citep{holman2013consequences, harts2014demography}.
+On the one hand, they prevent individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase the population's overall extinction risk \citep{kokko2003sexy,harts2014demography}.
 On the other hand, they provide an effective mechanism for the maintenance of genetic variation, which influences the population's capactiy for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
@@
 
-Classical population genetic models consider relative fitnesses of genotypes, and therefore do not consider the population dynamical consequences of evolutionary change.
+Sexually antagonistic selection, due to genetic trade-offs between male and female fitness, is a common feature of sexually reproducing populations, and is thought to contribute not only to the maintenance of genetic variation, but also to the evolution of sexual dimorphism and gender-related traits \citep{Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, BondurianskyChenoweth2009,Olito2019}.
+When such genetic trade-offs occur, "sexually antagonistic" selection (abbreviated SA hereafter) can arise whereby beneficial alleles for one sex are deleterious when expressed in the other \citep{Kidwell1977, Rice1992, ConnallonClark2012}.
+Alleles with opposing fitness effects through male and female sex functions can cause analogous genetic constraints on fitness in hermaphrodite populations, where both maternal and paternal reproductive success contribute jointly to each individuals' overall fitness \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
+
+The evolutionary dynamics of sexually antagonistic alleles in hermaphrodites differ from those in dioecious populations because individuals may reproduce through a combination of self- and outcross fertilization \citep{Goodwillie2005, Igic2006, JarneAuld2006}.
+For example, self-fertilization is predicted to reduce the total parameter space where balanced SA polymorphisms can be maintained, while simultaneously creating a bias in selection through the female sex function (\citealt{JordanConnallon2014}; but see \citealt{Tazzyman2015}).
+Yet, other processes, such as genetic linkage to other SA loci or a sex-determining region \citep{Otto2011, JordanCharlesworth2012, Olito2017, Olito2019}, or spatial heterogeneity and the complexity of the life-cycle can expand the parameter space for SA polymorphism \citep{Olito-etal-2018,ConnallonSharmaOlito2019}.
+Overall, both theoretical predictions and current empirical data suggest there is ample scope for SA trade-offs and the maintenance of SA polymorphisms in both dioecious and hermaphrodite populations \citep{Abbott2011, WangBarrett2020}, although identifying specific SA loci from genome sequence data remains challenging \citep{RuzickaESEB2020}.
+
+Most population genetic models, including models of the maintenance of SA polymorphisms, keep track of the relative fitnesses of genotypes, and therefore do not consider the population dynamical consequences of evolutionary change.
 To take the population dynamical consequences into account, a model is required that links changes in genotype distributions to population dynamics.
-The demographic processes of birth and death drive both changes in genotype distributions and population dynamics, and therefore provide a fundamental link between evolution and ecology \citep{metcalf2007evolutionary}.
-
-In this paper, we use an evolutionary demographic model to study when the parameter conditions predicted to be polymorphic by population genetics models are also demographically viable.
-We apply the model to study the maintenance of polymorphisms by sexually antagonistic selection in both dioecious and hermaproditic species.
+\cite{harts2014demography} build such a model using an individual based approach.
+Their model tracks both the population dynamics and the spread of sexually antagonistic genes as well as dispersal genes in a species with non-overlapping generations.
+\cite{deVriesCaswell2019b} introduced a model with overlapping generations, intralocus sexual antagonism, and population dynamics but did not perform analysis of the consequences of demographic viability for the scope of sexual antagonism to maintain polymorphism.
+
+In this paper, we extend the model by \cite{deVriesCaswell2019b} to include both dioecious and hermaphroditic species, and we use this model to study when the parameter conditions predicted to be polymorphic by population genetics models are also demographically viable.
 A major strength of the model is that we can use empirically estimated demographic rates to parameterize the model, enabling us to make predictions about the scope for demographically viable SA polymorphisms that are grounded in the biology of real populations.
@@
 
-Sexually antagonstic selection, due to genetic trade-offs between male and female fitness, are a common feature of sexually reproducing populations, and are thought to contribute not only to the maintenance of genetic variation, but also the evolution of sexual dimorphism and gender-related traits \citep{Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, BondurianskyChenoweth2009,Olito2019}.
-When such genetic trade-offs occur, "sexually antagonistic" selection (abbreviated SA hereafter) can arise whereby beneficial alleles for one sex are deleterious when expressed in the other \citep{Kidwell1977, Rice1992, ConnallonClark2012}.
-Alleles with opposing fitness effects through male and female sex functions can cause analogous genetic constraints on fitness in hermaphrodite populations, where both maternal and paternal reproductive success contribute jointly to each individuals' overall fitness \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
-
-The evolutionary dynamics of SA alleles in hermaphrodites differ from those in dioecious populations because individuals may reproduce through a combination of self- and outcross fertilization \citep{Goodwillie2005, Igic2006, JarneAuld2006}.
-Such mixed-mating systems have important theoretical consequences for the maintenance of SA polymorphisms in hermaphrodite populations.
-Self-fertilization is predicted to reduce the total parameter space where balanced SA polymorphisms can be maintained, while simultaneously creating a bias in selection through the female sex function (\citealt{JordanConnallon2014}; but see \citealt{Tazzyman2015}.
-Yet, other processes, such as genetic linkage to other SA loci or a sex-determining region \citep{Otto2011, JordanCharlesworth2012, Olito2017, Olito2019}, or spatial heterogeneity and the complexity of the life-cycle \citep{Olito-etal-2018,ConnallonSharmaOlito2019} can expand the parameter space for SA polymorphism.
-Overall, both theoretical predictions and current empirical data suggest there is ample scope for SA trade-offs and the maintenance of SA polymorphisms in both dioecious and hermaphrodite populations \citep{Abbott2011, WangBarrett2020}, although identifying specific SA loci from genome sequence data remains challenging \citep{RuzickaESEB2020}.
 % \bigskip
@@
 
-In this article, we develop a stage-structured mendelian matrix model that accommodates selection through male and female sex-functions in populations of simultaneous hermaphrodites with mixed mating systems.
-Under obligate outcrossing, the model becomes mathematically identical to the two-sex model of \citet{deVriesCaswell2019b}.
-Using the model, we jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations.
-We focus our analyses on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population is also demographically viable (i.e., has a positive intrinsic growth rate).
-The model also allows for the effects of inbreeding depression to manifest in different life-history stages, enabling us to explore the fitness consequences of early- vs.~late-acting inbreeding depression on the model predictions.
-Finally, we give an illustrative example of the potential for SA polymorphism in real populations by parameterizing our model with empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus} (now {\itshape Erythranthe guttata}).
-
-A key finding of our study is that when the intrinsic population growth rate is close to $1$, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space for SA polymorphism.
-Furthermore, this demographically viable polymorphic parameter space is often biased towards alleles with stronger selection through the female- than male sex function.
-Self-fertilization can alleviate the demographic costs of balanced SA polymorphisms under some conditions, however, the concommitant effects of inbreeding depression generally exacerbate them in populations with mixed-mating systems.
-Despite these demographic pitfalls, our example using {\itshape M.~guttatus} appears to show that demographic rates observed in real populations are capable of sustaining large regions of viable SA polymorphic space.
-Overall, our findings provide a more nuanced picture of the nature of SA genetic variation that we should expect to find in natural populations, where the fate of SA alleles and the populations harboring them is determined jointly by evolutionary and demographic processes.
-
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section*{The Model}
+\section{The Model}
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 Self-fertilization is achieved when an individual's male gametes fertilize their own ovules.
-In the following any parameters, matrices or vectors relating to male sex function are denoted with a prime.
-
-In the model, individuals are classified jointly by life-cycle stage ($1, \ldots, \omega$), genotype ($1, \ldots, g$), and whether they were produced by self- or outcross fertilization (denoted by $S$ and $X$ superscripts, respectively).
+To distinguish between parameters relating to male and female function, we denote matrices or vectors relating to male sex function with a prime.
+
+Individuals in the model are jointly classified by life-cycle stage ($1, \ldots, \omega$), genotype ($1, \ldots, g$), and whether they were produced by self- or outcross fertilization (denoted by $S$ and $X$ superscripts, respectively).
 Whether individuals were produced through selfing or outcrossing is included in the individual state description because individuals produced through selfing might experience reduced survival, growth, or maturation rates as a consequence of inbreeding depression.
 To model inbreeding depression, we therefore keep track of how individuals were produced.
-
-Each genotype is characterized by a matrix of transition probabilities (including survival) and a matrix of reproductive output through male and female function.
-These matrices can include time variation or nonlinearities reflecting the environment or density dependence, but these will not be modelled here.
+Of course the level of inbreeding depression will not only be a function of whether the current individual is produced through inbreeding but also of how many generations of inbreeding have occurred (see for example, \cite{kelly1999response} and \cite{kelly2007mutation}) but keeping track of the history of inbreeding is beyond the scope of this paper.
+Therefore we use the commonly made assumption that individuals produced through selfing suffer a fixed level of inbreeding depression (\textbf{which reference(s) would work best here?}).
+
+% Each genotype is characterized by a matrix of transition probabilities (including survival), and a matrix of reproductive output through male and female function.
+These matrices can include time variation or nonlinearities reflecting the environment or density dependence (see for example, \cite{de2020matrix}), but these will not be modelled here.
 Each stage contributes offspring to genotypes at the next time step according to transition matrices that are determined by the mating system and stage-structure of the population.
@@
 
-\noindent where $\mbf{n}^{S}_{i}$ and $\mbf{n}^{X}_{i}$ ($i \in \{AA,Aa,aa\}$) are the stage $\times$ genotype distribution vectors of individuals produced by self-fertilization and outcrossing, respectively.
+\noindent where $\mbf{n}^{S}_{i}$ and $\mbf{n}^{X}_{i}$ ($i \in \{AA,Aa,aa\}$) are the stage distribution vectors of individuals of genotype $i$ produced by self-fertilization and outcrossing, respectively.
 The proportional population vector is given by
@@
 The matrices $\mbf{F}_{i}$ and $\mbf{F}^{\prime}_{i}$ contain the genotype $\times$ stage specific contribution via female and male sex-functions of genotype $i$ to the female and male gamete pools, respectively, and therefore to zygotes in the next generation.
-We assume that whether individuals were produced through selfing or outcrossing does not affect their fecundity or mating success, that is, we assume  $\mbf{F}^X_{i}=\mbf{F}^S_{i}=\mbf{F}_{i}$ and $\mbf{F}^{\prime S}_{i}=\mbf{F}^{\prime X}_{i}=\mbf{F}^{\prime}_{i}$.
+We assume that whether individuals were produced through selfing or outcrossing does not affect their fecundity or mating success, that is, we assume  $\mbf{F}^X_{i}=\mbf{F}^S_{i}=\mbf{F}_{i}$ and $\mbf{F}^{\prime X}_{i}=\mbf{F}^{\prime S}_{i}=\mbf{F}^{\prime}_{i}$.
 Deviations from this assumption are straightforward to incorporate but beyond the scope of the current paper.
-In principle, the model can admit any kind of pleiotropy among demographic traits because the transition and fertility matrices can differ among genotypes in any way.
-For simplicity, we assume linear, time-invariant demography.
-
-Notational note: The superscript on a matrix in age$\times$stage models conventionally indicates the stage or age of the subsection of the population vector that the matrix is acting on \citep{CaswellEtAl2018}.
+
+% Notational note: The superscript on a matrix in age$\times$stage models conventionally indicates the stage or age of the subsection of the population vector that the matrix is acting on (see for example \cite{CaswellEtAl2018}).
 For example, $\mbf{U}^{S}_{i}$, acts on the vector of individuals of genotype $i$ that were produced through selfing.
-In the following, we deviate from this convention by also using $S$ and $X$ as superscripts to indicate reproduction of offspring through current selfing or reproduction.
-That is,  we define the matrices $\mbf{H}^S_{j}(\tilde{\mbf{n}})$ and $\mbf{H}^X_{j}(\tilde{\mbf{n}})$, as the map of the genotypes of the parent in stage $j$ to the genotypes of their offspring produced by selfing and outcrossing, respectively.
+In the following, we deviate from this convention by also using $S$ and $X$ as superscripts to indicate production of offspring through current selfing or reproduction.
+That is,  we will write production $\mcal{F}^S$ and $\mcal{F}^X$ to denote offspring production by self-fertilization and outcrossing.
+Likewise, we define the matrices $\mbf{H}^S_{j}(\tilde{\mbf{n}})$ and $\mbf{H}^X_{j}(\tilde{\mbf{n}})$ as the map of the genotypes of the parent in stage $j$ to the genotypes of their offspring produced by selfing and outcrossing, respectively.
 The $(k, l)$ entry of $\mbf{H}^{X}_{j}$ and $\mbf{H}^{S}_{j}$ is the probability that an offspring of a genotype $l$ mother, of stage $j$, has genotype $k$.
 Since we assume that how-you-were-produced does not affect how you reproduce, these matrices will act on both $\mbf{n}^{S}$ and $\mbf{n}^{X}$.
-Likewise, we will write production $\mcal{F}^S$ and $\mcal{F}^X$ to denote offspring production by self-fertilization and outcrossing.
 
@@
 \subsubsection*{Mating and offspring production under partial selfing}
-The inclusion of simultaneous hermaphrodites that can reproduce both through selfing and through outcrossing complicate mating in our model relative to the two-sex stage-structured model of \citet{deVriesCaswell2019b}.
-Since we assume that how individuals were produced does not affect the relative success of their male gametes, the frequency distribution in the male gametes is obtained simply by summing the vectors of individuals produced through selfing and individuals produced through outcrossing, that is,
+ Outcrossing in our model proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}.
+That is, each individual's genotype determines both the number of ovules produced and the paternal "mating success", broadly defined.
+For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986, WangBarrett2020, Harder2016}.
+
+Since we assume that how individuals were produced does not affect their relative mating success, the frequency in the male mating population is obtained simply by summing the vectors of individuals produced through selfing and through outcrossing, that is,
 \begin{equation}
@@
  \end{equation} 
-
-Let $C$ denote the proportion of each individual's ovules that are self-fertilized (the remaining $1 - C$ are outcrossed), and let $\delta$ represent the proportion of self-fertilized zygotes that fail to develop due to inbreeding depression during early development \citep{Charlesworth1987}.
-In \hl{Appendix A.x}, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
-Outcrossing proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}.
-That is, each individual's genotype determines both the number of ovules produced and the paternal "mating success", broadly defined.
-For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986, WangBarrett2020, Harder2016}.
-Depending on their stage and genotype at the SA locus, individuals contribute male gametes to the overall population gamete pool according to the stage $\times$ genotype vector:
+Depending on their stage and genotype at the SA locus, individuals contribute male gametes to the overall population gamete pool according to the following equation
 \begin{linenomath*}
@@
 
+\textcolor{red}{
+I'm doubting whether to move the next paragraph to the appendices as well.
+Equation \ref{eq:maleGametePool} is all we strictly need.
+If we keep the parent-offspring maps, we can explain in the next section how the matrices
+$\mcal{F}^S$ and $\mcal{F}^X$ have the same structure. Would that be helpful? Or would it be easier to just explain in the next section what the structure of $\mcal{F}^S$ and $\mcal{F}^X$ is? }
+
 The parent-offspring map for outcrossing is a function of the allele frequencies in the male gamete pool, and is given by 
@@
 \begin{equation} \label{eq:HX}
-	\mbf{H}^X(\tilde{\mbf{n}}) = 
+	\mbf{H}^X = 
 			\left(
@@
 \end{linenomath*}
-
-\noindent From left to right, the columns of matrix $\mbf{H}^X(\tilde{\mbf{n}})$ give the genotype distribution of outcrossed offspring produced by a maternal parent of each genotype ($AA$, $Aa$, and $aa$ respectively).
+\noindent From left to right, the columns of matrix $\mbf{H}^X$ give the genotype distribution of outcrossed offspring produced by a maternal parent of each genotype ($AA$, $Aa$, and $aa$ respectively).
 
@@
 \end{linenomath*}
-
 \noindent The columns of $\mbf{H}^S(\tilde{\mbf{n}})$ again give the selfed offspring genotype distributions for parental genotypes of $AA$, $Aa$, and $aa$, respectively.
@@
 \subsubsection*{Population projection}
-
-Using the component matrices described above, the matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ that projects the eco-evolutionary dynamics is:
+Using the component matrices described above (the survival matrices, $\mathbf{U}^X_i$, $\mathbf{U}^S_i$, the fertility and mating success matrices, $\mathbf{F}_i$, $\mathbf{F}^\prime_i$, and the parent-offspring matrices $\mbf{H}^S(\tilde{\mbf{n}})$ and $\mbf{H}^X$), we construct the population projection matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ using the vec-permutation approach of \citet{CaswellEtAl2018}, see \hl{Appendix A.1} for the step by step construction of the model.
+The matrix that projects the eco-evolutionary dynamics is:
 \begin{linenomath*}
@@
 				\mcal{F}^X(\tilde{\mbf{p}}) (1 - C) & \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)\\
-			\end{array} \right)}_{\tilde{\mbf{F}}}.
-\end{equation}
-\end{linenomath*}
-
-\noindent The blocks of the component matrices in Eq(\ref{eq:Atilde}) correspond to production of offspring by self-fertilization and outcrossing ($\mcal{F}^S$ and $\mcal{F}^X$ in $\tilde{\mbf{F}}$), and survival of selfed and outcrossed offspring ($\mcal{U}^S$ and $\mcal{U}^X$ in $\tilde{\mbf{U}}$).
-
-We construct $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ using the vec-permutation approach of \citet{CaswellEtAl2018}. Briefly, we construct a set of block-diagonal matrices: e.g.,
-\begin{linenomath*}
-\begin{equation} \label{eq:BlkUS}
-	\mbb{U}^S = 
+			\end{array} \right)}_{\tilde{\mbf{F}}},
+\end{equation}
+\end{linenomath*}
+where $C$ denotes the proportion of each individual's ovules that are self-fertilized (the remaining $1 - C$ are outcrossed), and  $\delta$ represents the proportion of self-fertilized zygotes that fail to develop due to inbreeding depression during early development \citep{Charlesworth1987}.
+In \hl{Appendix A.x}, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
+
+The blocks of the component matrices in Eq(\ref{eq:Atilde}) correspond to production of offspring by self-fertilization and outcrossing ($\mcal{F}^S$ and $\mcal{F}^X$ in $\tilde{\mbf{F}}$), and survival of selfed and outcrossed offspring ($\mcal{U}^S$ and $\mcal{U}^X$ in $\tilde{\mbf{U}}$).
+The survival matrix for individuals produced through selfing and outcrossing are,
+\begin{linenomath*}
+\begin{eqnarray} \label{eq:BlkUS}
+	\mcal{U}^S  = 
 		\left(
@@
 				\mbf{0} & \mbf{0} & \mbf{U}^{S}_{aa} \\
-			\end{array} \right),
-\end{equation}
-\end{linenomath*}
-
-\noindent and corresponding matrices $\mbb{U}^X$, $\mbb{F}^S$, $\mbb{F}^X$.
-Since individuals cannot change their genotype once they are born, the survival matrices are block diagonal, and $\mcal{U}^S = \mbb{U}^S$ and $\mcal{U}^X = \mbb{U}^X$.
-
-Similarly, we also construct block-diagonal matrices
-\begin{linenomath*}
-\begin{eqnarray}
-	\mathbb{H}^X(\tilde{\bo p}) = \bo I_\om \kron \bo H^X(\tilde{\bo p}),\\
-	\mathbb{H}^S(\tilde{\bo p}) = \bo I_\om \kron \bo H^S(\tilde{\bo p}).
+			\end{array} \right),\\
+				\mcal{U}^X  = 
+		\left(
+			\begin{array}{ccc}
+				\mbf{U}^{X}_{AA} & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{U}^{X}_{Aa} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \mbf{U}^{X}_{aa} \\
+			\end{array} \right).
 \end{eqnarray}
 \end{linenomath*}
-The corresponding fertility matrices for self-fertilization and outcrossing are defined as
+Since individuals cannot change their genotype once they are born, the survival matrices are block diagonal.
+Similarly, we construct fertility matrices for individuals produced through selfing and outcrossing,
 \begin{linenomath*}
 \begin{equation} \label{eq:BlkFS}
-	\mcal{F}^S = \mbf{K}^{\intercal} \mbb{H}^S(\tilde{\mbf{p}}) \mbf{K} \mbb{F}^S = 
+	\mcal{F}^S = 
 			\left(
@@
 \begin{equation} \label{eq:BlkFX}
-	\mcal{F}^X = \mbf{K}^{\intercal} \mbb{H}^X(\tilde{\mbf{p}}) \mbf{K} \mbb{F}^X = 
+	\mcal{F}^X = 
 			\left(
@@
 \end{linenomath*}
-
-\noindent where $q^{\prime}_A$ and $q^{\prime}_a$ are given by Eq(\ref{eq:maleGametePool}), and $\mbf{K}$ is the vec-permutation matrix \citep{HendersonSearle1981}.
-Further details of the derivations of $\mcal{F}^S$ and $\mcal{F}^X$ are provided in \hl{Appendix A.x}.
+\noindent where $q^{\prime}_A$ and $q^{\prime}_a$ are given by Eq(\ref{eq:maleGametePool}).
+Further details of the derivations of $\mcal{U}^S$, $\mcal{U}^X$ ,  $\mcal{F}^S$, and $\mcal{F}^X$ are provided in \hl{Appendix A.x}.
 
@@
 
-Combining all the component matrices yields the overall eco-evolutionary projection matrix shown in \hl{Reference a Box here?  or possibly Appendix A.x?}
+Combining all the component matrices yields the overall eco-evolutionary projection matrix shown in \hl{Appendix A.x}.
 
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Sexually antagonistic selection \& inbreeding depression} \label{sec:SAsel}
+\subsection{Sexually antagonistic selection \& inbreeding depression} \label{sec:SAsel}
 
@@
 
-Following convention, we parameterize the relative fitnesses of each genotype through each sex function ($w_{i}$ and $w^{\prime}_{i}$) to be bounded by $(0,1)$, with dominance and selection coefficients $h_f, s_f$ and $h_m, s_m$ determining the decrease in fertility through each sex function relative to the most fit genotype ($AA$ has highest female fertility, $aa$ the highest male fertility; see Table \ref{tab:Fitness}).
+Following convention, we parameterize the fertility component of fitness of each genotype through each sex function, $w_{i}$ and $w^{\prime}_{i}$, to be bounded by $[0,1]$, with dominance and selection coefficients $h_f, s_f$ and $h_m, s_m$ determining the decrease in fertility through each sex function relative to the most fit genotype ($AA$ has highest female fertility, $aa$ the highest male fertility; see Table \ref{tab:Fitness}).
 The SA locus does not affect survival and transition rates.
@@
 
-\noindent where $f$ and $f^{\prime}$ represent adult fertilities, and $w_{i}$ and $w^{\prime}_{i}$ the genotypic relative fitness expressions for female and male sex-functions (see Table \ref{tab:Fitness}).
+\noindent where $f$ and $f^{\prime}$ represent adult fertilities, and $w_{i}$ and $w^{\prime}_{i}$ the genotypic relative scaling factors for female and male sex-functions (see Table \ref{tab:Fitness}).
 
@@
  \centering
- \caption{\bf Fitness expressions for Sexually Antagonistic selection ($w_{i}$)}
+ \caption{\bf Relative fertilities for Sexually Antagonistic selection ($w_{i}$)}
 \label{tab:Fitness}
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Analyses} \label{subsec:analyses}
+\subsection{Analyses} \label{subsec:analyses}
 
@@
 We identify conditions where SA polymorphism is 'protected' by evaluating the stability of populations initially fixed for either SA allele to invasion by the other (i.e., stability was assessed at the boundary equilibrium genotype frequencies of $\hat{\mbf{p}}_{AA} = 1$ and $\hat{\mbf{p}}_{aa} = 1$; \citealt{Levene1953, Prout1968, deVriesCaswell2019b}).
-The formal conditions for a protected polymorphism are determined by linearizing the model in the vicinity of the boundary equilibria ($\hat{\mbf{p}}$), and evaluating the magnitude of the largest eigenvalue of the Jacobian matrix of the linearization, denoted by $\tilde{\zeta}_{AA}$ and $\tilde{\zeta}_{aa}$ for the eigenvalue of the Jacobian evaluated at the $AA$ and $aa$ boundary, respectively.
+The formal conditions for a protected polymorphism are determined by linearizing the model in the vicinity of the boundary equilibria ($\hat{\mbf{p}}$), and evaluating the magnitude of the largest eigenvalue of the Jacobian matrix of the linearization.
+A full derivation of the Jacobian and details of the invasion analysis are provided in \hl{Online Appendix X.x}.
+
+%denoted by $\tilde{\zeta}_{AA}$ and $\tilde{\zeta}_{aa}$ for the eigenvalue of the Jacobian evaluated at the $AA$ and $aa$ boundary, respectively.
 For our model, the Jacobian, $\mbf{M}$, has an upper block triangular form, and we can therefore express the stability of $\hat{\mbf{p}}$ in terms of the leading eigenvalue of the diagonal block corresponding to changes in heterozygote frequency ($\mbf{M}_{22}$; e.g., see \citealt{OttoDay2007,deVriesCaswell2019b}).
 For a polymorphism to be 'protected', both boundary equilibria must be unstable (i.e., $\tilde{\zeta}_{AA} > 1$ and $\tilde{\zeta}_{aa} > 1$), which requires that
-\begin{linenomath*}
-\begin{equation}\label{eq:PolyCond}
-	\rho(\mbf{M}_{22}) > \lambda_{i}
-\end{equation}
-\end{linenomath*}
-
-\noindent is satisfied, where $\rho(\cdot)$ represents the spectral radius,  $\rho(\mbf{M}_{22})$ gives the ergodic growth rate of perturbations of the boundary equilibrium according to the linear approximation, and $\lambda_i$ is the intrinsic growth rate for the homozygous genotype being invaded at each boundary ($i \in (AA,aa)$; see Box.~1).
-A full derivation of the Jacobian and details of the invasion analysis are provided in \hl{Online Appendix X.x}.
+% \begin{linenomath*}
+% \begin{equation}\label{eq:PolyCond}
+% 	\rho(\mbf{M}_{22}) > \lambda_{i}
+% \end{equation}
+% \end{linenomath*}
+
+% \noindent is satisfied, where $\rho(\cdot)$ represents the spectral radius,  $\rho(\mbf{M}_{22})$ gives the ergodic growth rate of perturbations of the boundary equilibrium according to the linear approximation, and $\lambda_i$ is the intrinsic growth rate for the homozygous genotype being invaded at each boundary ($i \in (AA,aa)$; see Box.~1).
 
@@
 In our model, the population state vector will grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see \citealt{Caswell2001}), and the intrinsic population growth rate after convergence, $\lambda$, can be calculated as $\tilde{\mbf{n}}(t)/\tilde{\mbf{n}}(t-1)$.
-We note, however, that if the ecological component of the model is non-linear, more exotic dynamics are possible.
-
-Unfortunately, closed-form solutions for the invasion conditions for each allele and the equilibrium frequencies under different selection scenarios are analytically intractable except for restricted conditions (see \hl{Online Appendix X}).
+We note, however, that if the ecological component of the model is non-linear, more exotic dynamics are possible \citep{de2020matrix}.
+
+% Unfortunately, closed-form solutions for the invasion conditions for each allele and the equilibrium frequencies under different selection scenarios are analytically intractable except for restricted conditions (see \hl{Online Appendix X}).
 We therefore used numerical techniques to identify invasion and extinction thresholds, which could then be used to define regions of demographically viable polymorphism across $s_f \times s_m$ parameter space.
 Because single-locus selection coefficients are generally weak (i.e., $s < 0.1$; e.g., \citealt{Eyre-WalkerKeightly2007}) and strongly skewed, we limit our analyses to coefficients within $0 < s_f,s_m \leq 0.15$, unless stated otherwise.
-For a given parameter set, we numerically evaluated Eq(\ref{eq:PolyCond}) and $\lambda$, starting at the outer boundaries of selection parameter space, and titrated inward to identify thresholds where $\tilde{\zeta}_i \approx 1$ and $\lambda \approx 1$ (with precision of $10^{-5}$).
+For a given parameter set, we numerically evaluated the largest eigenvalue of the Jacobian evaluated at both boundaries, and the population growth rate of the equilibrium obtained after perturbation of the boundary, $\lambda$.
+To determine the viability threshold (where the population growth rate equals one) and the boundary stability thresholds (where the largest eigenvalue of the Jacobian crosses one), we start at the outer boundaries of selection parameter space, and titrated inward.
 The proportions of total $s_f \times s_m$ parameter space where SA polymorphism and/or extinction occurred were calculated by numerically integrating the appropriate regions defined by the invasion and extinction thresholds
@@
 For analyses of the demographic effects of inbreeding depression, we made two main simplifying assumptions: First, we acknowledge that if inbreeding depression is caused primarily by recessive deleterious mutations (as suggested by empirical data), it should covary negatively with the population selfing rate due to purging (\citealt{Charlesworth2009}; though we note that other processes could give rise to this pattern, e.g., \citealt{CrnokrakBarrett2002, Charlesworth2009,HedrickGarcia-Dorado2016}).
-Following \citet{Olito2019}, we incorporate such negative covariance by constraining the inbreeding depression terms in the model ($\delta$ and $\delta_i$, where $i \in \{J,A,\gamma\}$) to follow a simple declining function of the selfing rate: $\delta = \delta^{\ast} (1 - b (1 - L))$, where $\delta^{\ast}$ is the hypothetical severity of inbreeding depression for a completely outcrossing population, $b$ is a shape parameter determining how far $\delta$ will decline under complete selfing (when $C = 1$), and $L$ is an expression describing the expected deleterious mutation load due to recessive deleterious mutations that includes an additional shape parameter, $a$, which determines the curvature of the overall function for $\delta$ (see Appendix E in \citealt{Olito2019} for additional details).
+Following \citet{Olito2019}, we incorporate such negative covariance by constraining the inbreeding depression terms in the model ($\delta$ and $\delta_i$, where $i \in \{J,A,\gamma\}$) to follow a simple declining function of the selfing rate: $\delta = \delta^{\ast} (1 - b (1 - L))$, where $\delta^{\ast}$ is the hypothetical severity of inbreeding depression for a completely outcrossing population, $b$ is a shape parameter determining how far $\delta$ will decline under complete selfing (when $C = 1$), and $L$ is describes the expected deleterious mutation load as a function of the selfing rate $C$, due to recessive deleterious mutations.
+The function $L$ includes an additional shape parameter, $a$, which determines the curvature of the overall function for $\delta$ (see Appendix E in \citealt{Olito2019} for additional details).
 We set $\delta^{\ast} = 0.8$, $b = 0.5$, and $a = 0.2$ for all analyses, values chosen to be consistent with empirical estimates of inbreeding depression (e.g., fig.~2 in \citealt{HusbandSchemske1996}).
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Demographic and fitness data for {\itshape M. guttatus}} \label{subsec:MguttMethods}
-
-As an illustrative example of how our model can be used to explore whether demographic rates observed in natural populations appear likely to support balanced SA polymorphisms, we parameterized our model using empirically estimated demographic rates and fitness data for natural populations of the hermaphroditic flowering plant {\itshape Mimulus guttatus} (Scrophulariaceae; now known as {\itshape Erythranthe guttata}).
-{\itshape M.~guttatus} is an herbaceous, self-compatible wildflower native to western North America that exhibits remarkable among-population variation in numerous life-history and reproductive traits including selfing rates, inbreeding depression, floral morphology, and annual-to-perennial life-history \citep[e.g.,][]{RitlandGanders1987, Ritland1990, Willis1993, Willis1999a, Willis1999b, WuWillis2008}.
-Moreover, detailed demographic studies have been conducted on multiple populations of {\itshape M.~guttatus}, with demographic matrix data available on the public demographic database COMPADRE \citep{CompadreDB2020}.
-Below, we briefly outline how we parameterized our model using the available data; full details are provided in \hl{Appendix B.x}.
-
-We used extensive demographic data reported in a large-scale study of local adaptation using experimental populations of {\itshape M.~guttatus} in Stanislaus National Forest (California, USA) in 2012 and 2013 \citep{PetersonEtAl2016}.
-We leverage their common-garden experimental design to focus on a comparison of demographically viable polymorphic parameter space for two experimental populations with contrasting demographic rates.
-The first was a locally adapted 'Eagle Meadows' population (data from 2012), while the second group was an experimental population composed of multiple non-locally adapted 'low-elevation perennials' (data from 2013).
-The vital rate estimates for the Eagle Meadows population are as follows: seed bank survival ($D = 0.534$), seed germination rate ($G = 0.469$), flower production ($F = 0.64$), ovules per flower ($O = 614$), seedling recruits proportional to clonal rosette recruits ($A = 6.7 \times 10^{-4}$), overwinter survival ($S = 0.179$), rosette production ($R = 8.71$).
-The corresponding estimates for the low-elevation perrenials are: $D = 0.534$, $G = 0.652$, $F = 0.4.09$, $O = 494$, $A = 6.7e-4$), $S = 0$, and $R = 0$ (see corrected Tables.~1 and S2 in \citealt{PetersonEtAl2017}).
-The same estimates for $D$ and $A$ were used for all populations.
-The resulting transition matrices for this population involved three life-history stages ($\omega = 3$; seed, seedling, and rosette), and individual elements of the transition matrix ($\tilde{\mbf{A}}$) were calculated as products of the above rates (see Matrix $1$ in \citealt{PetersonEtAl2016}).
-
-Estimates of selfing rates and inbreeding depression were not available for the same experimental populations, but are available for a variety of other western USA {\itshape M.~guttatus} populations.
-Selfing rate estimates vary in magnitude from near complete outcrossing to predominant selfing ($C \approx 0$ to $0.75$; \citealt{RitlandGanders1987, Ritland1990, Willis1999b}).
-Estimates of inbreeding depression at several of the life-history stages/fitness components that were included in the data of \citet{PetersonEtAl2016} are available for two intensively studied populations in the Cascade Mountains of Oregon (Iron Mountain and Cone Peak) \citet{Willis1993, Willis1999a, Willis1999b}.
-Using the data provided in \citet{Willis1993}, we estimated the proportional decrease due to inbreeding depression in seed germination rate ($\delta_{G} = 0.085$), flower number ($\delta_{F} = 0.2$), and overwinter survival ($\delta_{S} = 0.38$).
-The largest field-estimated selfing rate for this same Iron Mountain population was $C = 0.29$ \citep{Willis1993}.
-
-Using these combined demographic rates, selfing rate, and inbreeding depression estimates, we constructed a corresponding stage $\times$ genotype mendelian matrix model with a single SA locus affecting female and male fertility (as described above).
-Using this empirically parameterized model, we are able to make predictions about the scope for demographically viable SA polymorphism at a single locus in hypothetical populations with the same demographic rates as observed in \citet{PetersonEtAl2016}, for a range of selfing and inbreeding depression rates observed in other natural populations.
-We stress, however, that these illustrative predictions are not explicit predictions of the likelihood of SA polymorphism in any specific population.
-
-Interestingly, a polymorphic chromosomal inversion (inv6) with apparently SA fitness effects has been identified in the Iron Mountain population of {\itshape M.~guttatus} \citep{LeeKelly2015}.
-inv6 segregates at moderate frequency (about $8\%$), and carriers suffer $\approx 30 \%$ loss in pollen viability, but also increased flower production which varies among years.
-The genetic basis of the fitness effects caused by inv6 are not yet known, but the net result is a "supergene" with strong fitness effects that segregates as a single diallelic locus.
-We estimated dominance and selection coefficients for the effect of inv6 on pollen viability and flower production from the data reported in \citet{LeeKelly2015}.
-Across four mapping populations, the effect of inv6 on pollen viability was partially recessive on average ($h_m \approx 0.35$).
-Given this average dominance, the selection coefficient in the field is approximately $s_m = 0.88$.
-It was not possible to estimate the average dominance of the effect of inv6 on flower production (inv6 homozygotes were too rare in the field), so we conservatively assume that it was additive ($h_f = 0.5$).
-Based on this assumption, the selection coefficient inv6 on flower production in the field was $s_f = 0.384$ in 2012, and $s_f = 0.232$ in 2013.
-As a final proof-of-concept test of our empirically parameterized model \citep[e.g.,][]{Servedio2014}, we asked whether, given the above biologically grounded fitness effect estimates, inv6 appears to fall in demographically viable SA polymorphic parameter space, as might be expected given its observed frequency in the Iron Mountain population.
 
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Results}\label{sec:Results}
+\section{Results}\label{sec:Results}
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Polymorphism and Extinction}\label{subsec:PolyExt}
+\subsection{Polymorphism and Extinction}\label{subsec:PolyExt}
 
@@
 In obligately outcrossing populations (including dioecious/gonochoristic populations; where $C = 0$), lower fertility can result in a significant reduction of demographically viable polymorphic parameter space.
-However, the effect is weaker in populations with intermediate selfing rates ($C > 0$) because self-fertilization offers reproductive assurance for a fraction ($C$) of each individual's ovules.
-This allows partially sefing populations to remain viable under selection intensities that would cause extinction in an outcrossing population.
-Interestingly, when fertility is low enough that some polymorphic parameter space is demographically inviable, populations with intermediate selfing rates having the greatest proportion of demographically viable parameter space (fig.~\ref{fig:polySpace}, med.~and low.~fertility values).
-This effect emerges from the combination of increasing population viability due to reproductive assurance and reduced total polymorphic parameter space caused by selfing.
+However, the effect is weaker in populations with intermediate selfing rates ($C > 0$) because self-fertilization protects a fraction ($C$) of the ovules of individuals homozygous for the female beneficial allele from exposure to the allele that is deleterious for females.
+In other words, self-fertilization leads to less selective death.
+This allows partially selfing populations to remain viable under selection intensities that would cause extinction in an outcrossing population.
+Interestingly, when fertility is low enough that some polymorphic parameter space is demographically inviable, populations with intermediate selfing rates have the greatest proportion of demographically viable parameter space (fig.~\ref{fig:polySpace}, med.~and low.~fertility values).
+This effect emerges from the combination of protection from selective death and reduced total polymorphic parameter space caused by selfing.
 
@@
  \centering
- \includegraphics[scale=0.75]{../output/figs/polymorphicSpaceTitrate.pdf}
+ %\includegraphics[scale=0.75]{../output/figs/polymorphicSpaceTitrate.pdf}
  \caption{\footnotesize{Proportion of demographically viable parameter space (out of total $s_f \times s_m$ space with $\max(s) = 0.15$) in the absence of inbreeding depression (i.e., assuming $\delta = \delta_i = 0$, where $i \in \{j,a,\gamma\}$), plotted as a function of the population selfing rate.
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Demographic effects of inbreeding depression} \label{subsec:InbreedEffects}
+\subsection{Demographic effects of inbreeding depression} \label{subsec:InbreedEffects}
 
@@
 
-Mortality from inbreeding depression reduces the effect of "reproductive assurance" individuals gain through self fertilization.
+Mortality from inbreeding depression reduces the effect of protection from selective death that individuals gain through self fertilization.
 In contrast to our earlier results (see \hyperref[subsec:PolyExt]{Polymorphism and Extinction}), predominantly outcrossing populations are predicted to have the highest proportion of demographically viable polymorphic space, when inbreeding depression is taken into acount (compare fig.~\ref{fig:polySpace} with fig.~\ref{fig:deltaPolySpace}).
 Regardless of the life-history stage at which inbreeding depression affects survival, populations with intermediate to high selfing rates are unlikely to harbour SA polymorphism unless they can "afford" the resulting loss of self-fertilized ovules/offspring.
+\textcolor{red}{How about the flipside: Populations with intermediate to high selfing rates and growth rates around one are vulnerable to extinction if a sexually antagonistic allele invades the population.
+}
 
@@
  \centering
- \includegraphics[scale=0.75]{../output/figs/deltaPolymorphicSpaceTitrate.pdf}
+ %\includegraphics[scale=0.75]{../output/figs/deltaPolymorphicSpaceTitrate.pdf}
  \caption{\footnotesize{Effects of early- and late-acting inbreeding depression on the proportion of demographically viable parameter space (out of total $s_f \times s_m$ space with $\max(s) = 0.15$), plotted as a function of the population selfing rate.
@@
 Our results indicate that population viability was most sensitive to inbreeding depression affecting juvenile survival rates ($\delta_j$; fig.~\ref{fig:deltaPolySpace}, dark blue points), while early-acting inbreeding depression ($\delta$, ovule abortion shortly after fertilization) had a similar effect on population viability as late-acting inbreeding depression affecting adult survial ($\delta_a$) and juvenile-to-adult transition rates ($\delta_{\gamma}$).
-Inbreeding effects on juvenile survival had the strongest effect on population viability because of the sequential nature of life-cycle stages (individuals must survive the juvenile stage in order to reproduce as adults) and the multiplicative fitness effects of the survival and transition rates in the selfed-offspring survival matrix $\mbf{U}^{S}$.
-Although early-acting inbreeding depression ($\delta$) actually manifests earlier in the life-cycle than juvenile survival, it acts only once by influencing the total number of self-fertilized zygotes that become juveniles, and does not act multiplicatively with the other terms in $\mbf{U}^{S}$.
+Inbreeding effects on juvenile survival had the strongest effect on population viability because on average individuals will spend multiple time steps in the juvenile stage before they mature.
+At each time step in the juvenile stage, individuals have a probability $\gamma$ to mature and a probability $\sigma_j(1-\delta_j)$ to survive.
+Inbreeding depression at the juvenile stage therefore makes it harder to survive long enough to mature.
+Although early-acting inbreeding depression ($\delta$) actually manifests earlier in the life-cycle than juvenile survival, it acts only once by influencing the total number of self-fertilized zygotes that become juveniles.
+
+\section{Case study: {\itshape M. guttatus}}
+
+%\subsection{Demographic and fitness data for {\itshape M. guttatus}} \label{subsec:MguttMethods}
+
+As an illustrative example of how our model can be used to explore whether demographic rates observed in natural populations appear likely to support balanced SA polymorphisms, we parameterized our model using empirically estimated demographic rates and fitness data for natural populations of the hermaphroditic flowering plant {\itshape Mimulus guttatus} (Scrophulariaceae; now known as {\itshape Erythranthe guttata}).
+{\itshape M.~guttatus} is an herbaceous, self-compatible wildflower native to western North America that exhibits remarkable among-population variation in numerous life-history and reproductive traits including selfing rates, inbreeding depression, floral morphology, and annual-to-perennial life-history \citep[e.g.,][]{RitlandGanders1987, Ritland1990, Willis1993, Willis1999a, Willis1999b, WuWillis2008}.
+Moreover, detailed demographic studies have been conducted on multiple populations of {\itshape M.~guttatus}, with demographic matrix data available on the public demographic database COMPADRE \citep{CompadreDB2020}.
+Below, we briefly outline how we parameterized our model using the available data; full details are provided in \hl{Appendix B.x}.
+
+\subsection{Demographic and fitness data for {\itshape M. guttatus}} \label{subsec:MguttMethods}
+We used extensive demographic data reported in a large-scale study of local adaptation using experimental populations of {\itshape M.~guttatus} in Stanislaus National Forest (California, USA) in 2012 and 2013 \citep{PetersonEtAl2016}.
+We leverage their common-garden experimental design to focus on a comparison of demographically viable polymorphic parameter space for two experimental populations with contrasting demographic rates.
+The first was a locally adapted 'Eagle Meadows' population (data from 2012), while the second group was an experimental population composed of multiple non-locally adapted 'low-elevation perennials' (data from 2013).
+The vital rate estimates for the Eagle Meadows population are as follows: seed bank survival ($D = 0.534$), seed germination rate ($G = 0.469$), flower production ($F = 0.64$), ovules per flower ($O = 614$), seedling recruits proportional to clonal rosette recruits ($A = 6.7 \times 10^{-4}$), overwinter survival ($S = 0.179$), rosette production ($R = 8.71$).
+The corresponding estimates for the low-elevation perrenials are: $D = 0.534$, $G = 0.652$, \textcolor{red}{$F = 0.4.09$ Should this be 0.409 or 4.09?
+}, $O = 494$, $A = 6.7e-4$), $S = 0$, and $R = 0$ (see corrected Tables.~1 and S2 in \citealt{PetersonEtAl2017}).
+The same estimates for $D$ and $A$ were used for all populations.
+The resulting transition matrices for this population involved three life-history stages ($\omega = 3$; seed, seedling, and rosette), and individual elements of the transition matrix ($\tilde{\mbf{A}}$) were calculated as products of the above rates (see Matrix $1$ in \citealt{PetersonEtAl2016}).
+
+Estimates of selfing rates and inbreeding depression were not available for the same experimental populations, but are available for a variety of other western USA {\itshape M.~guttatus} populations.
+Selfing rate estimates vary in magnitude from near complete outcrossing to predominant selfing ($C \approx 0$ to $0.75$; \citealt{RitlandGanders1987, Ritland1990, Willis1999b}).
+Estimates of inbreeding depression at several of the life-history stages/fitness components that were included in the data of \citet{PetersonEtAl2016} are available for two intensively studied populations in the Cascade Mountains of Oregon (Iron Mountain and Cone Peak) \citet{Willis1993, Willis1999a, Willis1999b}.
+Using the data provided in \citet{Willis1993}, we estimated the proportional decrease due to inbreeding depression in seed germination rate ($\delta_{G} = 0.085$), flower number ($\delta_{F} = 0.2$), and overwinter survival ($\delta_{S} = 0.38$).
+The largest field-estimated selfing rate for this same Iron Mountain population was $C = 0.29$ \citep{Willis1993}.
+
+Using these combined demographic rates, selfing rate, and inbreeding depression estimates, we constructed a corresponding stage $\times$ genotype mendelian matrix model with a single SA locus affecting female and male fertility (as described above).
+Using this empirically parameterized model, we are able to make predictions about the scope for demographically viable SA polymorphism at a single locus in hypothetical populations with the same demographic rates as observed in \citet{PetersonEtAl2016}, for a range of selfing and inbreeding depression rates observed in other natural populations.
+We stress, however, that these illustrative predictions are not explicit predictions of the likelihood of SA polymorphism in any specific population.
+
+Interestingly, a polymorphic chromosomal inversion (inv6) with apparently SA fitness effects has been identified in the Iron Mountain population of {\itshape M.~guttatus} \citep{LeeKelly2015}.
+inv6 segregates at moderate frequency (about $8\%$), and carriers suffer $\approx 30 \%$ loss in pollen viability, but also increased flower production which varies among years.
+The genetic basis of the fitness effects caused by inv6 are not yet known, but the net result is a "supergene" with strong fitness effects that segregates as a single diallelic locus.
+We estimated dominance and selection coefficients for the effect of inv6 on pollen viability and flower production from the data reported in \citet{LeeKelly2015}.
+Across four mapping populations, the effect of inv6 on pollen viability was partially recessive on average ($h_m \approx 0.35$).
+Given this average dominance, the selection coefficient in the field is approximately $s_m = 0.88$.
+It was not possible to estimate the average dominance of the effect of inv6 on flower production (inv6 homozygotes were too rare in the field), so we conservatively assume that it was additive ($h_f = 0.5$).
+Based on this assumption, the selection coefficient inv6 on flower production in the field was $s_f = 0.384$ in 2012, and $s_f = 0.232$ in 2013.
+As a final proof-of-concept test of our empirically parameterized model \citep[e.g.,][]{Servedio2014}, we asked whether, given the above biologically grounded fitness effect estimates, inv6 appears to fall in demographically viable SA polymorphic parameter space, as might be expected given its observed frequency in the Iron Mountain population.
 
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Polymorphism in M.~guttatus} \label{sec:Mgutt}
+\subsection{Polymorphism in M.~guttatus} \label{sec:Mgutt}
 
@@
 Due to the slower growth rate, not all of the $s_f \times s_m$ selection parameter space was demographically viable.
-Extinction thresholds appeared under complete outcrossing ($C = 0$) and partial selfing with inbreeding depression ($C = 0.29$; $\mbf{U}^{S}$), but not when inbreeding depression was omitted (fig.~\ref{fig:mimulusFig}B).
+Extinction thresholds appeared under complete outcrossing ($C = 0$) and partial selfing with inbreeding depression ($C = 0.29$), but not when inbreeding depression was omitted (fig.~\ref{fig:mimulusFig}B).
 The different demographic rates from these two populations also resulted in slightly different invasion conditions when inbreeding depression was taken into account (compare dotted lines in fig/~\ref{fig:mimulusFig}A and B).
@@
  \centering
- \includegraphics[scale=0.75]{../output/figs/mimulusinv6Fig.pdf}
+ %\includegraphics[scale=0.75]{../output/figs/mimulusinv6Fig.pdf}
  \caption{\footnotesize{Illustration of model predictions using empirically estimated demographic rates for {\itshape M.~guttatus}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Discussion}
+\section{Discussion}
 %%%%%%%%%%%%%%%%%%%%%%%%
-
-\ldots
+\textit{Summary} \newline
+In this paper we developed a stage-structured mendelian matrix model that accommodates selection through male and female sex-functions in populations of simultaneous hermaphrodites with mixed mating systems.
+Using the model, we jointly analyzed the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations.
+We focused our analyses on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population is also demographically viable (i.e., has a positive intrinsic growth rate).
+We also looked at the effects of inbreeding depression manifesting in different life-history stages, exploring the fitness consequences of early- vs.~late-acting inbreeding depression on the model predictions.
+Finally, we gave an illustrative example of the potential for SA polymorphism in real populations by parameterizing our model with empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus} (now {\itshape Erythranthe guttata}).
+
+A key finding of our study is that when the intrinsic population growth rate is close to $1$, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space for SA polymorphism.
+Furthermore, this demographically viable polymorphic parameter space is often biased towards alleles with stronger selection through the female- than male sex function.
+Self-fertilization can alleviate the demographic costs of balanced SA polymorphisms under some conditions, however, the concommitant effects of inbreeding depression generally exacerbate them in populations with mixed-mating systems.
+Despite these demographic pitfalls, our example using {\itshape M.~guttatus} appears to show that demographic rates observed in some real populations are capable of sustaining large regions of viable SA polymorphic space.
+Overall, our findings provide a more nuanced picture of the nature of SA genetic variation that we should expect to find in natural populations, where the fate of SA alleles and the populations harboring them is determined jointly by evolutionary and demographic processes.
+
+\textit{Extensions} \newline
+By combining the tools of demography and those of population genetics, new questions can be answered with the framework we presented here.
+For example.
+many interesting extensions of our paper towards more complicated life histories are possible.
+For the sake of simplicity, we used a basic life cycle with just two stages for most of this paper, adults and juveniles (the \textit{Mimulus} example has 3 stages).
+By including more age classes, we can ask whether the scope for SA to maintain polymorphisms is affected by whether a species exhibits positive, or negative senescence \citep{jones2014diversity}?
+That is, does the shape of the survival curve affect the area of demographically viable polymorphic parameter space, and how does this interact with the age at which the sexually antagonistic allele is expressed?
+What if the allele not only affects male and female sex functions but also affects survival of individuals?
+And last but not least, how would the scope for SA to maintain polymorphisms be affected by density-dependence acting on different life history stages?
+
+We made the simplifying assumption that only how individuals themselves were produced affects the level of inbreeding depression they experience.
+In reality, the history of inbreeding in the entire lineage will influence the level of inbreeding depression experienced by an individual.
+This could be included in our framework by expanding the individual state space from produced via outcrossing or selfing towards produced via produced via outcrossing or first generation selfing, second generation selfing, etc, as in \cite{kelly1999response} and \cite{kelly2007mutation}.
+
+\textit{Conclusion} \newline
+Despite the recent interest in eco-evolutionary dynamics, the demographic consequences of intralocus sexual antagonism have rarely been considered (but see \cite{harts2014demography}).
+In contrast, models of the population dynamical consequences of interlocus sexual conflict are more common, see for example \cite{tanaka1996sexual} and \cite{martinez2017sexual}.
+\hl{There's probably other relevant papers here, check!} In this paper we showed that including basic demography can have a significant impact on traditional population genetic results.
+For example,  Figure \ref{fig:extThresholds} shows that large fractions of polymorphic parameter space could lead to population extinction.
+This inviable fraction of parameter space is especially large for populations with growth rates already close to one (ignoring any potential compensatory density-dependent effects).
+
+Demographic models connect individual level traits to population level consequences.
+Through their focus on individuals, demographic models are ideal for linking to experimental or field data, such as capture-mark-recapture data.
+Doing so allows the field of population genetics to move from fitness as an abstract scalar metric towards the fitness of an entire life cycle obtained by combining observed rates of age- or stage-specific survival and fecundity rates.
 
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Conclusion}
+%\section{Conclusion}
 %%%%%%%%%%%%%%%%%%%%%%%%
-
-\ldots
 
@@
 
-%\section*{Acknowledgments}
+%\section{Acknowledgments}
 
@@
 
-\section*{Appendix A: Supplementary Figures}
+\section{Appendix A: Construction of the population projection matrix}
+
+\subsubsection*{Population projection}
+
+Using the component matrices described above, the matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ that projects the eco-evolutionary dynamics is:
+\begin{linenomath*}
+\begin{equation} \label{eq:Atilde_appendix}
+	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
+			\underbrace{\left(
+			\begin{array}{c|c}
+				\mcal{U}^S & \mbf{0} \\ \hline
+				\mbf{0} & \mcal{U}^X \\
+			\end{array} \right)}_{\tilde{\mbf{U}}} + 
+			\underbrace{\left(
+			\begin{array}{c|c}
+				\mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
+				\mcal{F}^X(\tilde{\mbf{p}}) (1 - C) & \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)\\
+			\end{array} \right)}_{\tilde{\mbf{F}}}.
+\end{equation}
+\end{linenomath*}
+
+\noindent The blocks of the component matrices in Eq(\ref{eq:Atilde}) correspond to production of offspring by self-fertilization and outcrossing ($\mcal{F}^S$ and $\mcal{F}^X$ in $\tilde{\mbf{F}}$), and survival of selfed and outcrossed offspring ($\mcal{U}^S$ and $\mcal{U}^X$ in $\tilde{\mbf{U}}$).
+
+We construct $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ using the vec-permutation approach of \citet{CaswellEtAl2018}, see \hl{Appendix A.1} for the step by step construction of the model.
+Briefly, we construct a set of block-diagonal matrices: e.g.,
+\begin{linenomath*}
+\begin{equation} \label{eq:BlkUS_appendix}
+	\mbb{U}^S = 
+		\left(
+			\begin{array}{ccc}
+				\mbf{U}^{S}_{AA} & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{U}^{S}_{Aa} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \mbf{U}^{S}_{aa} \\
+			\end{array} \right),
+\end{equation}
+\end{linenomath*}
+
+\noindent and corresponding matrices $\mbb{U}^X$, $\mbb{F}^S$, $\mbb{F}^X$.
+Since individuals cannot change their genotype once they are born, the survival matrices are block diagonal, and $\mcal{U}^S = \mbb{U}^S$ and $\mcal{U}^X = \mbb{U}^X$.
+
+Similarly, we also construct block-diagonal matrices
+\begin{linenomath*}
+\begin{eqnarray}
+	\mathbb{H}^X(\tilde{\bo p}) = \bo I_\om \kron \bo H^X(\tilde{\bo p}),\\
+	\mathbb{H}^S(\tilde{\bo p}) = \bo I_\om \kron \bo H^S(\tilde{\bo p}).
+\end{eqnarray}
+\end{linenomath*}
+The corresponding fertility matrices for self-fertilization and outcrossing are defined as
+\begin{linenomath*}
+\begin{equation} \label{eq:BlkFS_appendix}
+	\mcal{F}^S = \mbf{K}^{\intercal} \mbb{H}^S(\tilde{\mbf{p}}) \mbf{K} \mbb{F}^S = 
+			\left(
+			\begin{array}{ccc}
+				\mbf{F}_{AA} & \frac{1}{4} \mbf{F}_{Aa} & 0 \\
+				0 & \frac{1}{2} \mbf{F}_{Aa} & 0 \\
+				0 & \frac{1}{4} \mbf{F}_{Aa} & \mbf{F}_{aa}\\
+			\end{array} \right), 
+\end{equation}
+\end{linenomath*}
+\noindent and 
+\begin{linenomath*}
+\begin{equation} \label{eq:BlkFX_appendix}
+	\mcal{F}^X = \mbf{K}^{\intercal} \mbb{H}^X(\tilde{\mbf{p}}) \mbf{K} \mbb{F}^X = 
+			\left(
+			\begin{array}{ccc}
+				q^{\prime}_{A} \mbf{F}_{AA} & \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} & 0 \\ 
+				q^{\prime}_{a} \mbf{F}_{AA} & \frac{1}{2} \mbf{F}_{Aa} & q^{\prime}_{A} \mbf{F}_{aa}  \\ 
+				0 & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} & q^{\prime}_{a} \mbf{F}_{aa}  \\
+			\end{array} \right),
+\end{equation}
+\end{linenomath*}
+
+\noindent where $q^{\prime}_A$ and $q^{\prime}_a$ are given by Eq(\ref{eq:maleGametePool}), and $\mbf{K}$ is the vec-permutation matrix \citep{HendersonSearle1981}.
+Further details of the derivations of $\mcal{F}^S$ and $\mcal{F}^X$ are provided in \hl{Appendix A.x}.
+
+The blocks of $\mcal{F}^X(\tilde{\mbf{p}})$ can be constructed and interpreted as follows: The first row block of the first column produces $AA$ offspring by outcrossing from $AA$ maternal parents.
+This happens when the $AA$ maternal parent receives an $A$ gamete from the male gamete pool, which happens with probability $q^{\prime}_{A}$.
+The other blocks can be interpreted similarly.
+
+Combining all the component matrices yields the overall eco-evolutionary projection matrix shown in \hl{Reference a Box here?  or possibly Appendix A.x?}
 
@@
 
-\subsection*{Fox--dog encounters through the ages}
-
-The quick red fox jumps over the lazy brown dog.
-The quick red fox has always jumped over the lazy brown dog.
-The quick red fox began jumping over the lazy brown dog in the 19th century and has never ceased from so jumping, as we shall see in figure~\ref{Fig:Jumps}.
-But there can be surprises (figure~\ref{Fig:JumpsOk}).
-
-If the order and location of figures is not otherwise clear, feel free to include explanatory dummy text like this:
-
-[Figure A1 goes here.]
-
-[Figure A2 goes here.]
-
-\subsection*{Further insights}
-
-Tables in the appendices can appear in the appendix text (see table~\ref{Table:Rivers} for an example), unlike appendix figure legends which should be grouped at the end of the document together with the other figure legends.
-
-\begin{table}[h]
-\caption{Various rivers, cities, and animals}
-\label{Table:Rivers}
-\centering
-\begin{tabular}{lll}\hline
-River        & City        & Animal            \\ \hline
-Chicago      & Chicago     & Raccoon           \\
-Des Plaines  & Joliet      & Coyote            \\
-Illinois     & Peoria      & Cardinal          \\
-Kankakee     & Bourbonnais & White-tailed deer \\
-Mississippi  & Galena      & Bald eagle        \\ \hline
-\end{tabular}
-\bigskip{}
-\\
-{\footnotesize Note: See table~\ref{Table:Founders} below for further table formatting hints.}
-\end{table}
-
-Lorem ipsum dolor sit amet, as we have seen in figures~\ref{Fig:Jumps} and \ref{Fig:JumpsOk}.
-
-\newpage{}
-\renewcommand{\theequation}{B\arabic{equation}}
-% redefine the command that creates the equation number.
-\renewcommand{\thetable}{B\arabic{table}}
-\setcounter{equation}{0}  % reset counter 
-\setcounter{table}{0}
-
-\section*{Appendix B: Additional Methods}
-
-\subsection*{Measuring the height of fox jumps without a meterstick}
-
-Pellentesque ac nibh placerat, luctus lectus non, elementum mauris. 
-Morbi odio velit, eleifend ut hendrerit vitae, consequat sit amet 
-nulla. Pellentesque porttitor vitae nisl quis tempus. Pellentesque 
-habitant morbi tristique senectus et netus et malesuada fames ac 
-turpis egestas. Praesent ut nisi odio. Vivamus vel lorem gravida 
-odio molestie volutpat condimentum et arcu.
-
-\begin{equation}
-{ \frac{1}{N_k-1} \sum \limits_{t=1}^{N_k} (M_{tjk} - \bar{M}_{jk})^2}
-\end{equation}
-
-\subsection*{Quantifying the brownness of the dog}
-
-Pellentesque eu nulla odio.
-Nulla aliquam porta metus, quis malesuada orci faucibus quis.
-Suspendisse nunc magna, tristique sit amet sollicitudin nec, elementum et lacus.
-Sed vitae elementum mi.
-In hac habitasse platea dictumst.
-Etiam eu tortor elit.
-Sed ac tortor purus.
-Aliquam volutpat, odio sit amet posuere pretium, dolor ex interdum ante, sed luctus quam eros ac nulla.
-
-\begin{equation}
-{ (\sum \limits_{p=1}^P {n_{sp}})^{-1}\sum \limits_{p=1}^P {n_{sp}Q_{p}}}
-\end{equation}
-
-\newpage{}
+\section{Appendix B: Additional Methods}
 
@@
 %%%%%%%%%%%%%%%%%%%%%
-\section*{Tables}
+\section{Tables}
 \renewcommand{\thetable}{\arabic{table}}
@@
 
-\section*{Figure legends}
+\section{Figure legends}
 
@@
 
-\subsection*{Online figure legends}
+\subsection{Online figure legends}
 

### 2021-02-19 doc/SA-Hermaphrodites-wDemography.tex
Edits of discussion and added an appendix

@@
 I'm doubting whether to move the next paragraph to the appendices as well.
-Equation \ref{eq:maleGametePool} is all we strictly need.
+Equation (\ref{eq:maleGametePool}) is all we strictly need.
 If we keep the parent-offspring maps, we can explain in the next section how the matrices
@@
 \begin{linenomath*}
-\begin{eqnarray} \label{eq:BlkUS}
+\begin{eqnarray} 
 	\mcal{U}^S  = 
@@
 				\mbf{0} & \mbf{0} & \mbf{U}^{S}_{aa} \\
-			\end{array} \right),\\
+			\end{array} \right),\label{eq:BlkUS}\\
 				\mcal{U}^X  = 
@@
 				\mbf{0} & \mbf{0} & \mbf{U}^{X}_{aa} \\
-			\end{array} \right).
+			\end{array} \right).\label{eq:BlkUX}
 \end{eqnarray}
@@
 
-\textit{Extensions} \newline
-By combining the tools of demography and those of population genetics, new questions can be answered with the framework we presented here.
-For example.
-many interesting extensions of our paper towards more complicated life histories are possible.
+\textit{Extensions: demography} \newline
+By combining the tools of demography and those of population genetics, the framework we presented here enables the exploration of interactions between life cycle complexity, mating system, and sexual antagonism.
 For the sake of simplicity, we used a basic life cycle with just two stages for most of this paper, adults and juveniles (the \textit{Mimulus} example has 3 stages).
-By including more age classes, we can ask whether the scope for SA to maintain polymorphisms is affected by whether a species exhibits positive, or negative senescence \citep{jones2014diversity}?
-That is, does the shape of the survival curve affect the area of demographically viable polymorphic parameter space, and how does this interact with the age at which the sexually antagonistic allele is expressed?
+By including more age classes, we can ask if the scope for SA to maintain polymorphisms is affected by whether a species exhibits positive, or negative senescence \citep{jones2014diversity}?
+That is, does the shape of the survival curve affect the area of demographically viable polymorphic parameter space, and how does this interact with the age at which a sexually antagonistic allele is expressed?
 What if the allele not only affects male and female sex functions but also affects survival of individuals?
@@
 
+\textit{Extensions: genetics} \newline
 We made the simplifying assumption that only how individuals themselves were produced affects the level of inbreeding depression they experience.
@@
 
+\cite{grieshop2017male} found that genotypes with sexually antagonstic alleles that are male beneficial experience higher levels of inbreeding depression than genotypes with female beneficial SA alleles.
+This effect can easily be included in our model by making the inbreeding depression parameters, $\delta$ and $\delta_j$, a function of the genotype of the individual.
+Such an effect would even further reduce the demographically viable polymorphic parameter space.
+
+Other interesting extensions worth mentioning?
+
 \textit{Conclusion} \newline
-Despite the recent interest in eco-evolutionary dynamics, the demographic consequences of intralocus sexual antagonism have rarely been considered (but see \cite{harts2014demography}).
-In contrast, models of the population dynamical consequences of interlocus sexual conflict are more common, see for example \cite{tanaka1996sexual} and \cite{martinez2017sexual}.
-\hl{There's probably other relevant papers here, check!} In this paper we showed that including basic demography can have a significant impact on traditional population genetic results.
-For example,  Figure \ref{fig:extThresholds} shows that large fractions of polymorphic parameter space could lead to population extinction.
-This inviable fraction of parameter space is especially large for populations with growth rates already close to one (ignoring any potential compensatory density-dependent effects).
+Despite a surge of interest in eco-evolutionary dynamics, the demographic consequences of intralocus sexual antagonism have rarely been modeled (but see \cite{harts2014demography} and \cite{kokko2003sexy}).
+In contrast, models of the population dynamical consequences of interlocus sexual conflict are more common (e.g., \cite{tanaka1996sexual} and \cite{martinez2017sexual}).
+We found that including basic demography can have a significant impact on traditional population genetic results, as has been suggested previously both by theoretical work \citep{kokko2003sexy} and empirical work \citep{berger2016intralocus,grieshop2017male}.
+Although the potential negative consequences of sexual conflict for population viability have been known for some time, this aspect of sexual antagonism has rarely been considered in models investigating the scope for intralocus sexual antagonism to maintain polymorphisms, and hence genetic diversity.
 
 Demographic models connect individual level traits to population level consequences.
-Through their focus on individuals, demographic models are ideal for linking to experimental or field data, such as capture-mark-recapture data.
-Doing so allows the field of population genetics to move from fitness as an abstract scalar metric towards the fitness of an entire life cycle obtained by combining observed rates of age- or stage-specific survival and fecundity rates.
+As a consequence of their focus on individuals, demographic models are ideal for linking to experimental or field data, as demonstrated with our \textit{Mimulus} case study.
+Doing so allows the field of population genetics to move from fitness as an abstract scalar metric towards the fitness of an entire life cycle as calculated from observed rates of age- or stage-specific survival and fecundity rates.
 
@@
 \subsubsection*{Population projection}
-
-Using the component matrices described above, the matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ that projects the eco-evolutionary dynamics is:
+In this appendix, we will derive the expressions for the survival and fertility matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ as given by equations (\ref{eq:BlkUS})-(\ref{eq:BlkFX}) in the main text.
+The matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ are the ingredients of the population projection matrix, $\tilde{\mbf{A}}$, as given in equation (\ref{eq:Atilde}), repeated here for convenience,
 \begin{linenomath*}
@@
 			\begin{array}{c|c}
-				\mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
+				\mcal{F}^S C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
 				\mcal{F}^X(\tilde{\mbf{p}}) (1 - C) & \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)\\
@@
 
-\noindent The blocks of the component matrices in Eq(\ref{eq:Atilde}) correspond to production of offspring by self-fertilization and outcrossing ($\mcal{F}^S$ and $\mcal{F}^X$ in $\tilde{\mbf{F}}$), and survival of selfed and outcrossed offspring ($\mcal{U}^S$ and $\mcal{U}^X$ in $\tilde{\mbf{U}}$).
-
-We construct $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ using the vec-permutation approach of \citet{CaswellEtAl2018}, see \hl{Appendix A.1} for the step by step construction of the model.
-Briefly, we construct a set of block-diagonal matrices: e.g.,
-\begin{linenomath*}
-\begin{equation} \label{eq:BlkUS_appendix}
-	\mbb{U}^S = 
+Our derivation follows closely the derivations in \cite{deVriesCaswell2019a} and \cite{deVriesCaswell2019b}, but extended to model mixed mating systems.
+To construct $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ using the vec-permutation approach \citep{CaswellEtAl2018}, we create a set of block-diagonal matrices $\mathbb{U}^S$, $\mathbb{U}^X$, $\mathbb{F}^X$, $\mathbb{F}^S$, and $\mathbb{H}^S$, $\mathbb{H}^X$, by putting the corresponding matrices on the diagonal.
+These block diagonal matrices can be written as
+\begin{linenomath*}
+\begin{eqnarray}
+ \mathbb{U}^S&=&\sum_{i=1}^g \mathbf{E}_{ii} \otimes \mathbf{U}^S_{i}, \label{eq:US_app}\\
+ \mathbb{F}^S&=&\sum_{i=1}^g \mathbf{E}_{ii} \otimes \mathbf{F}^S_{i}, \label{eq:FS_app}\\
+  \mathbb{H}^S &=& \bo I_\om \kron \bo H^S , \label{eq:HS_app} \\
+      \mathbb{D} &=& \bo I_\om \kron \bo I_g, \label{eq:D_app} 
+\end{eqnarray}
+\end{linenomath*}
+where $\mathbf{E}_{ii}$ is of dimensions $g \times g$. Analogous equations can be written down for $ \mathbb{U}^X$, $ \mathbb{F}^X$, and $ \mathbb{H}^X$.  
+% \begin{linenomath*}
+% \begin{eqnarray} 
+% 	\mbb{U}^S &= &
+% 		\left(
+% 			\begin{array}{ccc}
+% 				\mbf{U}^{S}_{AA} & \mbf{0} & \mbf{0} \\
+% 				\mbf{0} & \mbf{U}^{S}_{Aa} & \mbf{0} \\
+% 				\mbf{0} & \mbf{0} & \mbf{U}^{S}_{aa} \\
+% 			\end{array} \right), \label{eq:BlkUS_appendix}\\
+%   \mbb{F}^S &= &
+% 		\left(
+% 			\begin{array}{ccc}
+% 				\mbf{F}^{S}_{AA} & \mbf{0} & \mbf{0} \\
+% 				\mbf{0} & \mbf{F}^{S}_{Aa} & \mbf{0} \\
+% 				\mbf{0} & \mbf{0} & \mbf{F}^{S}_{aa} \\
+% 			\end{array} \right), \label{eq:BlkFS_appendix}\\
+% \end{eqnarray}
+% \end{linenomath*}
+% Similarly construct block diagonal matrices from the parent-offspring genotype maps,
+% \begin{linenomath*}
+% \begin{eqnarray}
+% 	\mathbb{H}^X(\tilde{\bo p}) = \bo I_\om \kron \bo H^X(\tilde{\bo p}),\\
+% 	\mathbb{H}^S(\tilde{\bo p}) = \bo I_\om \kron \bo H^S(\tilde{\bo p}).
+% \end{eqnarray}
+% \end{linenomath*}
+
+As in other multistate models, the survival and fertility matrices are then constructed from the following four block matrices, 
+\begin{linenomath*}
+\begin{eqnarray}
+\mcal{U}^S &=&\bo K^\tr \mathbb{D} \bo K \mathbb{U}^S, \label{eq:KDKU_app}\\
+\mcal{F}^S &=&\bo K^\tr \mbb{H}^S\bo K \mathbb{F}^S, \label{eq:KHKF_app}
+\end{eqnarray}
+\end{linenomath*}
+where $\mathbf{K}=\mathbf{K}_{\omega,g}$ is the vec-permutation matrix \citep{HendersonSearle1981}.
+The vec-permutation matrix rearrangers the population vectors $\mathbf{n}^S$ and $\mathbf{n}^X$ so that $\mathbf{n}^S$ is ordered by genotype first, and then by stage, but $\mathbf{K} \mathbf{n}^S$ is organized by stage first, and then by genotype (see \cite{hunter2005use} or \cite{caswell2012matrix} for more on the vec-permutation method).
+
+The matrix $\mathbb{U}^S$ in equation (\ref{eq:KDKU_app}) generates transitions and survival of extant individuals within genotypes,  and acts on the the population vector of individuals produced through selfing, $\mathbf{n}^S$..
+The matrix $\bo K$ then permutes the resulting vector so that it is now organised by stages first, and then by genotypes within stages.
+The block diagonal matrix $ \mathbb{D} $ then accounts for changes in genotype within stage classes of extant individuals.
+Since extant individuals generally do not change their genotype $\mathbb{D}$ will be an identity matrix in our model.
+Finally, the matrix $\bo K^\tr $ permutes the population vector once again so that it returns to the original organisation by genotype first and then by stage classes.
+
+Equation (\ref{eq:KHKF_app}) describes reproduction and genotype assignment of newborn individuals.
+The block matrix $\mathbb{F}^S$ contains fecundity rates for individuals of each genotype and stage combination and acts on the the population vector of individuals produced through selfing, $\mathbf{n}^S$.
+Initially, these offspring are associated with the genotype of their parent.
+The vec-permutation matrix $\bo K$ then rearranges the resulting vector so that it is organised by stage first and then by genotype.
+The block-diagonal matrix $\mbb{H}^S$ then allocates the offspring to their genotypes, according to equation (\ref{eq:HS}).
+Finally, $\bo K^\tr $  returns the vector to its initial orientation.
+
+Substituting equations (\ref{eq:US_app}) and (\ref{eq:D_app}) into equation (\ref{eq:KDKU_app}) yields
+\begin{linenomath*}
+\begin{eqnarray}
+\mcal{U}^S &=& \sum_{i=1}^g \bo K^\tr \bo I_{\omega g}	 \bo K \mathbf{E}_{ii} \otimes \mathbf{U}^S_{i}, \\
+ &=& \sum_{i=1}^g \mathbf{E}_{ii} \otimes \mathbf{U}^S_{i},
+\end{eqnarray}
+\end{linenomath*}
+since $\bo K^\tr \bo I_{\omega g}	 \bo K=\bo I_{\omega g}$, or in block matrix form this yields equation (\ref{eq:BlkUS}), here repeated for convenience, 
+\begin{linenomath*}
+\begin{eqnarray}
+\mcal{U}^S  = 
 		\left(
@@
 				\mbf{0} & \mbf{0} & \mbf{U}^{S}_{aa} \\
-			\end{array} \right),
-\end{equation}
-\end{linenomath*}
-
-\noindent and corresponding matrices $\mbb{U}^X$, $\mbb{F}^S$, $\mbb{F}^X$.
-Since individuals cannot change their genotype once they are born, the survival matrices are block diagonal, and $\mcal{U}^S = \mbb{U}^S$ and $\mcal{U}^X = \mbb{U}^X$.
-
-Similarly, we also construct block-diagonal matrices
+			\end{array} \right).\label{eq:BlkUS}\\
+\end{eqnarray}
+\end{linenomath*}
+
+Similarly, substituting equations (\ref{eq:FS_app}) and (\ref{eq:HS_app}) into equation (\ref{eq:KHKF_app}) yields,
 \begin{linenomath*}
 \begin{eqnarray}
-	\mathbb{H}^X(\tilde{\bo p}) = \bo I_\om \kron \bo H^X(\tilde{\bo p}),\\
-	\mathbb{H}^S(\tilde{\bo p}) = \bo I_\om \kron \bo H^S(\tilde{\bo p}).
+\mcal{F}^S &=&\bo K^\tr \mbb{H}^S\bo K \mathbb{F}^S, \label{eq:KHKF_app}\\
+ &=& \sum_{i=1}^g \bo K^\tr \left( \bo I_\om \kron \bo H^S \right)  \bo K \left( \mathbf{E}_{ii} \otimes \mathbf{F}^S_{i} \right).
 \end{eqnarray}
 \end{linenomath*}
-The corresponding fertility matrices for self-fertilization and outcrossing are defined as
-\begin{linenomath*}
-\begin{equation} \label{eq:BlkFS_appendix}
-	\mcal{F}^S = \mbf{K}^{\intercal} \mbb{H}^S(\tilde{\mbf{p}}) \mbf{K} \mbb{F}^S = 
+Using $\bo K^\tr \left(\bo A \otimes \bo B  \right) \; \bo K=\bo B \otimes \bo A$ \citep{magnus1979commutation}, we obtain
+\begin{linenomath*}
+\begin{eqnarray}
+\mcal{F}^S  &=& \sum_{i=1}^g  \left(  \bo H^S  \kron \bo I_\om\right) \left( \mathbf{E}_{ii} \otimes \mathbf{F}^S_{i} \right).
+\end{eqnarray}
+\end{linenomath*}
+Finally note that $(\mathbf{A} \otimes \mathbf{B}) (\mathbf{C} \otimes \mathbf{D}) = \mathbf{A} \mathbf{C}\otimes \mathbf{B} \mathbf{D}$  to write
+\begin{linenomath*}
+\begin{eqnarray}
+\mcal{F}^S  &=& \sum_{i=1}^g  \left(  \bo H^S \mathbf{E}_{ii} \right) \kron  \mathbf{F}^S_{i} .
+\end{eqnarray}
+\end{linenomath*}
+Using the exact same sequence of equations, we can derive
+\begin{linenomath*}
+\begin{eqnarray}
+\mcal{F}^X  &=& \sum_{i=1}^g  \left(  \bo H^X(\tilde{\mbf{p}}) \mathbf{E}_{ii} \right) \kron  \mathbf{F}^X_{i} .
+\end{eqnarray}
+\end{linenomath*}
+
+Finally, substitute the expressions derived for the parent-offpspring genotype maps, $\bo H^X(\tilde{\mbf{p}}) $, and $\bo H^S$, in the main text (equations (\ref{eq:HS}) and (\ref{eq:HX})) into the above equations to obtain
+\begin{linenomath*}
+\begin{eqnarray}
+	\mcal{F}^S = \mbf{K}^{\intercal} \mbb{H}^S\mbf{K} \mbb{F}^S = 
 			\left(
@@
 				0 & \frac{1}{4} \mbf{F}_{Aa} & \mbf{F}_{aa}\\
-			\end{array} \right), 
-\end{equation}
+			\end{array} \right),  \label{eq:BlkFS_appendix}
+\end{eqnarray}
 \end{linenomath*}
@@
 \end{linenomath*}
-
-\noindent where $q^{\prime}_A$ and $q^{\prime}_a$ are given by Eq(\ref{eq:maleGametePool}), and $\mbf{K}$ is the vec-permutation matrix \citep{HendersonSearle1981}.
-Further details of the derivations of $\mcal{F}^S$ and $\mcal{F}^X$ are provided in \hl{Appendix A.x}.
-
-The blocks of $\mcal{F}^X(\tilde{\mbf{p}})$ can be constructed and interpreted as follows: The first row block of the first column produces $AA$ offspring by outcrossing from $AA$ maternal parents.
-This happens when the $AA$ maternal parent receives an $A$ gamete from the male gamete pool, which happens with probability $q^{\prime}_{A}$.
-The other blocks can be interpreted similarly.
-
-Combining all the component matrices yields the overall eco-evolutionary projection matrix shown in \hl{Reference a Box here?  or possibly Appendix A.x?}
 

### 2021-03-11 doc/SA-Hermaphrodites-wDemography.tex
Expanded Discussion, New Abstract, Table, etc.
Expanded Discussion, added a couple alternative abstracts, finished
(hopefully) table of model parameters, and made minor edits
throughout the ms.

Still have not updated appendices. Need to chat about these I think.

@@
 
-\tableofcontents
+%\tableofcontents
 %%%%%%%%%%%%%%%%%%%%
 \newpage{}
-\section{Abstract}
+\section*{Abstract -- currently 215/200 word max.}
 %%%%%%%%%%%%%%%%%%%%
-\ldots
+Genetic trade-offs between male and female fitness impose fascinating genetic constraints on adaptation.
+On one hand they can maintain genetic variation for a variety of life-history and gender-related traits, while on the other hand they cause unavoidable maladaptation as individuals are prevented from reaching their sex-specific fitness optima.
+In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of sexually antagonistic (SA) selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations.
+We focus on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population intrinsic growth rate remains positive.
+We also analyze the effects of inbreeding depression manifesting in different life-history stages.
+Finally, we gave an illustrative example of the potential for SA polymorphism in real populations using empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus}.
+Our results indicate that when population growth rates are near one, large swathes of polymorphic parameter space become demographically inviable, and that inbreeding depression poses a significant challenge to the maintenance of SA polymorphism in partially selfing populations.
+Despite these demographic challenges, our example with {\itshape M.~guttatus} appears to show that demographic rates observed in some real populations are capable of sustaining large regions of viable SA polymorphic space.
 
 \newpage{}
+\section*{ALTERNATIVE Abstract -- currently 200/200 word max.}
+% Alternative opening sentence to the abstract:sentence
+Now classic population genetics theory has been fundamental to understanding the evolution of sex-differences and the maintenance of sexually antagonistic (SA) genetic variation, but these models have rarely considered the demographic consequences of intralocus sexual antagonism.
+In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations.
+We focus on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection and the population intrinsic growth rate remains positive.
+We also analyze the effects of inbreeding depression manifesting in different life-history stages, and give an illustrative example of the potential for SA polymorphism in real populations using empirically estimated demographic rates for the hermaphroditic flowering plant Mimulus guttatus.
+Our results indicate that when population growth rates are near one, large swathes of polymorphic parameter space become demographically inviable, and that inbreeding depression poses a significant challenge to the maintenance of SA polymorphism in partially selfing populations.
+Despite these demographic challenges, our example with M.
+guttatus appears to show that demographic rates observed in some real populations are capable of sustaining large regions of viable SA polymorphic space.
+
+% GOOD STUFF HERE...
+KEEPING THIS IN CASE WE WANT TO WORK MORE OF THIS INTO THE ABSTRACT...In this paper we developed a stage-structured mendelian matrix model that accommodates selection through male and female sex-functions in populations of simultaneous hermaphrodites with mixed mating systems.
+Using the model, we jointly analyzed the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations.
+We focused our analyses on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population is also demographically viable (i.e., has a positive intrinsic growth rate).
+We also looked at the effects of inbreeding depression manifesting in different life-history stages, exploring the fitness consequences of early- vs.~late-acting inbreeding depression on the model predictions.
+Finally, we gave an illustrative example of the potential for SA polymorphism in real populations by parameterizing our model with empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus} (now {\itshape Erythranthe guttata}).
+
+\newpage{}
 
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
+\section*{Introduction}
 %%%%%%%%%%%%%%%%%%%%%%%%
 
-Population persistence or extinction depends heavily on the ability of a species to adapt to the environment.
-Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, such as deleterious mutations \citep{Haldane1957}, environmental shifts \citep{Maynard-Smith1976, LandeShannon1996,OrrUnckless2008}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, or genetic constraints \citep{ConnallonHall2018}.
+In order to persist in the long term, a population must be able to adapt to its environment.
+Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, such as deleterious mutations \citep{Haldane1957}, changing environmental conditions \citep{Maynard-Smith1976, LandeShannon1996,OrrUnckless2008}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, and genetic constraints \citep{ConnallonHall2018, MatthewsConnallon2019}.
 Genetic constraints on adaptation can arise as a consequence of conflicting selection and gene-flow between different classes of individuals, or more generally, when there are genetic trade-offs between fitness components \citep{CharlesworthHughes2000, ConnallonHall2018}.
 
- % Whether a population persists or goes extinct depends in large part on its ability to adapt to the environment.
-Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, including deleterious mutations \citep{Haldane1957}, environmental shifts causing a population to chase a new phenotypic optimum \citep{Maynard-Smith1976, LandeShannon1996,OrrUnckless2008}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, as well as genetic constraints arising from intrinsic features of the inheritance system \citep{ConnallonHall2018}.
-
-%Conflicting selection and gene-flow between different classes of individuals, or genetic trade-offs between fitness components, can impose constraints on adaptation with particularly interesting evolutionary consequences \citep{CharlesworthHughes2000, ConnallonHall2018}.
-Fitness trade-offs constraining adaptation have particularly interesting evolutionary consequences.
-On the one hand, they prevent individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase the population's overall extinction risk \citep{kokko2003sexy,harts2014demography}.
-On the other hand, they provide an effective mechanism for the maintenance of genetic variation, which influences the population's capactiy for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
-Hence, for traits under conflicting selection, the nature and extent of genetic variation observed in natural populations should reflect a balance between the contribution of genetic constraints to the maintainance of genetic polymorphisms, and the population dynamical consequences of the resulting maladaptation.
-
-Sexually antagonistic selection, due to genetic trade-offs between male and female fitness, is a common feature of sexually reproducing populations, and is thought to contribute not only to the maintenance of genetic variation, but also to the evolution of sexual dimorphism and gender-related traits \citep{Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, BondurianskyChenoweth2009,Olito2019}.
-When such genetic trade-offs occur, "sexually antagonistic" selection (abbreviated SA hereafter) can arise whereby beneficial alleles for one sex are deleterious when expressed in the other \citep{Kidwell1977, Rice1992, ConnallonClark2012}.
+Fitness trade-offs that impose genetic constraints on adaptation have particularly interesting evolutionary consequences.
+On the one hand, they prevent individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase a population's overall extinction risk \citep{kokko2003sexy,harts2014demography}.
+On the other hand, they provide an effective mechanism for the maintenance of genetic variation, which influences a population's capactiy for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
+Hence, for traits under conflicting selection, the nature and extent of genetic variation observed in natural populations should reflect a balance between the maintainance of genetic polymorphisms, and the population dynamical consequences of the resulting maladaptation.
+
+Sexually antagonistic selection (abbreviated SA hereafter) arising from genetic trade-offs between male and female fitness, is a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and also the evolution of sexual dimorphism and gender-related traits \citep{Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, BondurianskyChenoweth2009,Olito2019}.
+When such genetic trade-offs occur, SA selection can arise when beneficial alleles for one sex are deleterious when expressed in the other \citep{Kidwell1977, Rice1992, ConnallonClark2012}.
 Alleles with opposing fitness effects through male and female sex functions can cause analogous genetic constraints on fitness in hermaphrodite populations, where both maternal and paternal reproductive success contribute jointly to each individuals' overall fitness \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
@@
 For example, self-fertilization is predicted to reduce the total parameter space where balanced SA polymorphisms can be maintained, while simultaneously creating a bias in selection through the female sex function (\citealt{JordanConnallon2014}; but see \citealt{Tazzyman2015}).
-Yet, other processes, such as genetic linkage to other SA loci or a sex-determining region \citep{Otto2011, JordanCharlesworth2012, Olito2017, Olito2019}, or spatial heterogeneity and the complexity of the life-cycle can expand the parameter space for SA polymorphism \citep{Olito-etal-2018,ConnallonSharmaOlito2019}.
+Yet, other factors, such as genetic linkage to other SA loci or a sex-determining region \citep{Otto2011, JordanCharlesworth2012, Olito2017, Olito2019}, or spatial heterogeneity and complexity of the life-cycle can expand the parameter space for SA polymorphism \citep{Olito-etal-2018,ConnallonSharmaOlito2019}.
 Overall, both theoretical predictions and current empirical data suggest there is ample scope for SA trade-offs and the maintenance of SA polymorphisms in both dioecious and hermaphrodite populations \citep{Abbott2011, WangBarrett2020}, although identifying specific SA loci from genome sequence data remains challenging \citep{RuzickaESEB2020}.
 
-Most population genetic models, including models of the maintenance of SA polymorphisms, keep track of the relative fitnesses of genotypes, and therefore do not consider the population dynamical consequences of evolutionary change.
-To take the population dynamical consequences into account, a model is required that links changes in genotype distributions to population dynamics.
-\cite{harts2014demography} build such a model using an individual based approach.
-Their model tracks both the population dynamics and the spread of sexually antagonistic genes as well as dispersal genes in a species with non-overlapping generations.
-\cite{deVriesCaswell2019b} introduced a model with overlapping generations, intralocus sexual antagonism, and population dynamics but did not perform analysis of the consequences of demographic viability for the scope of sexual antagonism to maintain polymorphism.
-
-In this paper, we extend the model by \cite{deVriesCaswell2019b} to include both dioecious and hermaphroditic species, and we use this model to study when the parameter conditions predicted to be polymorphic by population genetics models are also demographically viable.
-A major strength of the model is that we can use empirically estimated demographic rates to parameterize the model, enabling us to make predictions about the scope for demographically viable SA polymorphisms that are grounded in the biology of real populations.
+Most population genetic models, including models of the maintenance of SA polymorphisms, keep track of the relative fitnesses of genotypes in populations of constant size, and therefore do not consider the population dynamical consequences of evolutionary change.
+In fact, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals.
+To take the population dynamical consequences of polymorphism into account, a model is required that links changes in genotype distributions to population dynamics.
+For example, \cite{harts2014demography} developed an individual-based modelling approach that tracks population dynamics and the spread of both SA genes and dispersal genes in a species with non-overlapping generations.
+\cite{deVriesCaswell2019b} introduced a mendelian matrix model with overlapping generations, intralocus sexual antagonism, and population dynamics but did not perform analysis of the consequences of demographic viability for the scope of sexual antagonism to maintain polymorphism.
+
+In this paper, we extend the model of \cite{deVriesCaswell2019b} to include both dioecious and hermaphroditic species in order to study the parameter conditions under which balanced SA polymorphisms are also demographically viable.
+A major strength of our approach is that the model can be parameterized using empirically estimated demographic rates, enabling us to make predictions about the scope for demographically viable SA polymorphisms that are grounded in the biology of real populations.
 We demonstrate this with a case study of {\itshape Mimulus guttatus} (now {\itshape Erythranthe guttata})
-
-%Such an eco-evolutionary model requires a map from genotype to phenotype, from phenotype to individual demographic rates (birth and death rates), and finally from demographic rates to population level characteristics \citep{coulson2006putting}.
-%Genetic trade-offs between fitness components due to pleiotropy represent a form of genetic constraint with particularly interesting evolutionary consequences because they act as a double-edged sword
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section{The Model}
+\section*{The Model}
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 Here, we briefly describe a matrix model incorporating multiple life-cycle stages and a single diallelic locus under SA selection on male and female fertility for populations of simultaneous hermaphrodites with outcrossing or mixed-mating systems.
-The derivation and some of the key results follow closely those presented in \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b}; and in fact, the model presented here reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019b} under obligate outcrossing.
-A full derivation of the model and analyses is presented in \hl{Appendix A}, and all simulation code necessary to reproduce the results are available at \url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
-
-Simultaneous hermaphrodites express both male and female sex functions (i.e., can be both a mother and father), and have the potential to reproduce by a combination of self- and outcross-fertilizations.
+The derivation and some of the key results follow closely those presented in \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b}.
+In fact, the present model reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019b} under obligate outcrossing.
+A full derivation of the model and analyses is presented in \hl{Appendix A}, and all computer code necessary to reproduce the results are available at \url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
+
+Simultaneous hermaphrodites express both male and female sex functions (i.e., can transmit genes to the next generation via both sperm and ovules), and have the potential to reproduce by a combination of self- and outcross-fertilizations.
 Maternal outcrossing involves receiving male gametes from another individual in the population, while paternal outcrossing involves exporting male gametes to another individual and fertilizing their ovules.
@@
 Whether individuals were produced through selfing or outcrossing is included in the individual state description because individuals produced through selfing might experience reduced survival, growth, or maturation rates as a consequence of inbreeding depression.
-To model inbreeding depression, we therefore keep track of how individuals were produced.
-Of course the level of inbreeding depression will not only be a function of whether the current individual is produced through inbreeding but also of how many generations of inbreeding have occurred (see for example, \cite{kelly1999response} and \cite{kelly2007mutation}) but keeping track of the history of inbreeding is beyond the scope of this paper.
-Therefore we use the commonly made assumption that individuals produced through selfing suffer a fixed level of inbreeding depression (\textbf{which reference(s) would work best here?}).
-
-% Each genotype is characterized by a matrix of transition probabilities (including survival), and a matrix of reproductive output through male and female function.
-These matrices can include time variation or nonlinearities reflecting the environment or density dependence (see for example, \cite{de2020matrix}), but these will not be modelled here.
-Each stage contributes offspring to genotypes at the next time step according to transition matrices that are determined by the mating system and stage-structure of the population.
+In reality, the severity of inbreeding depression will not only be a function of whether an individual is produced by selfing but also of how many consecutive generations of inbreeding have occurred in their lineage (e.g., \citealt{kelly1999response, kelly2007mutation}).
+However, tracking lineage-specific inbreeding histories is beyond the scope of this paper, and we therefore make the common assumption that the severity of inbreeding depression is the same for all individuals (e.g., \citealt{Charlesworth1987, Charlesworth2009, Charlesworth2010,JordanConnallon2014}).
 
@@
 \noindent where $\| \cdot \|$ is the one-norm.
-The population vector $\tilde{\mbf{n}}(t)$ is projected forward from time $t$ to $t + 1$ by the $projection$ matrix $\tilde{\mbf{A}}(\tilde{\mbf{n}})$ such that
+The population vector $\tilde{\mbf{n}}(t)$ is projected forward from time $t$ to $t + 1$ by the projection matrix $\tilde{\mbf{A}}(\tilde{\mbf{n}})$ such that
 \begin{linenomath*}
@@
 We assume that whether individuals were produced through selfing or outcrossing does not affect their fecundity or mating success, that is, we assume  $\mbf{F}^X_{i}=\mbf{F}^S_{i}=\mbf{F}_{i}$ and $\mbf{F}^{\prime X}_{i}=\mbf{F}^{\prime S}_{i}=\mbf{F}^{\prime}_{i}$.
-Deviations from this assumption are straightforward to incorporate but beyond the scope of the current paper.
+Deviations from this assumption are straightforward to incorporate but beyond the scope of this paper.
 
@@
 
-Since we assume that how individuals were produced does not affect their relative mating success, the frequency in the male mating population is obtained simply by summing the vectors of individuals produced through selfing and through outcrossing, that is,
+Because we assume that how individuals were produced (either by selfing or outcrossing) does not affect their relative mating success, the frequency in the male mating population is obtained by simply summing the vectors of individuals produced through selfing and outcrossing:
+\begin{linenomath*}
 \begin{equation}
@@
  \end{equation}
+\end{linenomath*}
 Depending on their stage and genotype at the SA locus, individuals contribute male gametes to the overall population gamete pool according to the following equation
@@
 \end{linenomath*}
-
-\textcolor{red}{
-I'm doubting whether to move the next paragraph to the appendices as well.
+The key difference between selfing and outcrossing reproduction can be seen in the parent-offspring maps, which reflect the joint effects of meiosis and mating on the distribution of offspring genotypes.
+% \textcolor{red}{
+% I'm doubting whether to move the next paragraph to the appendices as well.
 Equation (\ref{eq:maleGametePool}) is all we strictly need.
-If we keep the parent-offspring maps, we can explain in the next section how the matrices
-$\mcal{F}^S$ and $\mcal{F}^X$ have the same structure. Would that be helpful? Or would it be easier to just explain in the next section what the structure of $\mcal{F}^S$ and $\mcal{F}^X$ is? }
+If we keep the parent-offspring maps, we can explain in the next section how the matrices $\mcal{F}^S$ and $\mcal{F}^X$ have the same structure.
+Would that be helpful?
+Or would it be easier to just explain in the next section what the structure of $\mcal{F}^S$ and $\mcal{F}^X$ is?
+}
 
@@
 \noindent The columns of $\mbf{H}^S(\tilde{\mbf{n}})$ again give the selfed offspring genotype distributions for parental genotypes of $AA$, $Aa$, and $aa$, respectively.
+As described in the next section, these parent-offspring maps influence the projection matrix by altering the fertilitiy matrices for individuals produced by selfing and outcrossing (see Eq(\ref{eq:BlkFS}) and Eq(\ref{eq:BlkFX}) below).
 
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection{Sexually antagonistic selection \& inbreeding depression} \label{sec:SAsel}
+\subsection*{Sexually antagonistic selection \& inbreeding depression} \label{sec:SAsel}
 
@@
 To keep the model analytically tractable, we assume our hypothetical species has two life stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active.
-Suppose that there is a genetic trade-off between the sex-functions at a single bi-allelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect.
+Suppose that there is a genetic trade-off between the sex-functions at a single diallelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection{Analyses} \label{subsec:analyses}
+\subsection*{Analyses} \label{subsec:analyses}
 
@@
 
-We identify conditions where SA polymorphism is 'protected' by evaluating the stability of populations initially fixed for either SA allele to invasion by the other (i.e., stability was assessed at the boundary equilibrium genotype frequencies of $\hat{\mbf{p}}_{AA} = 1$ and $\hat{\mbf{p}}_{aa} = 1$; \citealt{Levene1953, Prout1968, deVriesCaswell2019b}).
+We identify conditions where SA polymorphism is 'protected' by evaluating the stability of populations initially fixed for either SA allele to invasion by the other (i.e., we assessed stability at the boundary equilibrium genotype frequencies of $\hat{\mbf{p}}_{AA} = 1$ and $\hat{\mbf{p}}_{aa} = 1$; \citealt{Levene1953, Prout1968, deVriesCaswell2019b}).
 The formal conditions for a protected polymorphism are determined by linearizing the model in the vicinity of the boundary equilibria ($\hat{\mbf{p}}$), and evaluating the magnitude of the largest eigenvalue of the Jacobian matrix of the linearization.
@@
 Because single-locus selection coefficients are generally weak (i.e., $s < 0.1$; e.g., \citealt{Eyre-WalkerKeightly2007}) and strongly skewed, we limit our analyses to coefficients within $0 < s_f,s_m \leq 0.15$, unless stated otherwise.
+For simplicity, we also assume equal dominance in all analyses (i.e., $h_f = h_m = h$).
 For a given parameter set, we numerically evaluated the largest eigenvalue of the Jacobian evaluated at both boundaries, and the population growth rate of the equilibrium obtained after perturbation of the boundary, $\lambda$.
@@
 
-For analyses of the demographic effects of inbreeding depression, we made two main simplifying assumptions: First, we acknowledge that if inbreeding depression is caused primarily by recessive deleterious mutations (as suggested by empirical data), it should covary negatively with the population selfing rate due to purging (\citealt{Charlesworth2009}; though we note that other processes could give rise to this pattern, e.g., \citealt{CrnokrakBarrett2002, Charlesworth2009,HedrickGarcia-Dorado2016}).
-Following \citet{Olito2019}, we incorporate such negative covariance by constraining the inbreeding depression terms in the model ($\delta$ and $\delta_i$, where $i \in \{J,A,\gamma\}$) to follow a simple declining function of the selfing rate: $\delta = \delta^{\ast} (1 - b (1 - L))$, where $\delta^{\ast}$ is the hypothetical severity of inbreeding depression for a completely outcrossing population, $b$ is a shape parameter determining how far $\delta$ will decline under complete selfing (when $C = 1$), and $L$ is describes the expected deleterious mutation load as a function of the selfing rate $C$, due to recessive deleterious mutations.
+For analyses of the demographic effects of inbreeding depression, we made two main simplifying assumptions: First, we assume that if inbreeding depression is caused primarily by recessive deleterious mutations (as suggested by empirical data), it should covary negatively with the population selfing rate due to purging (\citealt{Charlesworth2009}; though we note that other processes could give rise to this pattern, e.g., \citealt{CrnokrakBarrett2002, Charlesworth2009,HedrickGarcia-Dorado2016}).
+Following \citet{Olito2019}, we incorporate such negative covariance by constraining the inbreeding depression terms in the model ($\delta$ and $\delta_i$, where $i \in \{j,a,\gamma\}$) to follow a simple declining function of the selfing rate: $\delta = \delta^{\ast} (1 - b (1 - L))$, where $\delta^{\ast}$ is the hypothetical severity of inbreeding depression for a completely outcrossing population, $b$ is a shape parameter determining how far $\delta$ will decline under complete selfing (when $C = 1$), and $L$ describes the expected deleterious mutation load as a function of the selfing rate $C$.
 The function $L$ includes an additional shape parameter, $a$, which determines the curvature of the overall function for $\delta$ (see Appendix E in \citealt{Olito2019} for additional details).
 We set $\delta^{\ast} = 0.8$, $b = 0.5$, and $a = 0.2$ for all analyses, values chosen to be consistent with empirical estimates of inbreeding depression (e.g., fig.~2 in \citealt{HusbandSchemske1996}).
-Second, we explore the effects of individual inbreeding depression terms in isolation.
-That is, we assume that only one of the terms $\delta$ and $\delta_i$ (where $i \in \{J,A,\gamma\}$) can be non-zero at a time.
+Second, to keep our analyses tractable we explore the effects of individual inbreeding depression terms in isolation.
+That is, we assume that only one of the $\delta$ and $\delta_i$ terms (where $i \in \{j,a,\gamma\}$) can be non-zero at a time.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}\label{sec:Results}
+\section*{Results}\label{sec:Results}
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection{Polymorphism and Extinction}\label{subsec:PolyExt}
+\subsection*{Polymorphism and Extinction}\label{subsec:PolyExt}
 
@@
  \includegraphics[width=\linewidth]{../output/figs/extinctionThresholdsFig.pdf}
- \caption{\footnotesize{Illustration of parameter space for SA polymorphism and extinction thresholds predicted by the mendelian stage-structured matrix model.
+ \caption{\footnotesize{Illustration of parameter space for SA polymorphism and extinction thresholds predicted by the model.
 Balanced SA polymorphisms can be maintained in the funnel-shaped region between the invasion conditions for each SA allele (dark solid lines).
@@
 
-We begin with a simple illustration of demographically viable polymorphic parameter space in the absence of inbreeding depression (i.e., $\delta_i = 0$).
-Invasion conditions for SA alleles in the evolutionary demographic model closely match the predictions from population genetic models (see \hl{Online Appendix X} \citealt{Kidwell1977, JordanConnallon2014, Olito2017}).
+We begin with a simple illustration of demographically viable polymorphic parameter space in the absence of inbreeding depression (i.e., $\delta = \delta_i = 0$).
+Invasion conditions for SA alleles in the evolutionary demographic model closely match the predictions from population genetic models (see \hl{Online Appendix X}; \citealt{Kidwell1977, JordanConnallon2014, Olito2017}).
 In particular, the demographic model recovers the classic "funnel-shaped" region of polymorphic $s_f \times s_m$ parameter space.
@@
 
-However, a key prediction from the evolutionary demographic model is that large fractions of SA polymorphic parameter space can be demographically inviable (Fig.~\ref{fig:extThresholds}).
+However, a key prediction from the evolutionary demographic model is that large fractions of SA polymorphic parameter space can be demographically inviable (fig.~\ref{fig:extThresholds}).
 The location of the extinction threshold, where the population intrinsic growth rate $\lambda = 1$, is primarily determined by the fertility parameter ($f$) but is also influenced by the population selfing rate ($C$) and dominance of the SA alleles ($h$).
@@
 In obligately outcrossing populations (including dioecious/gonochoristic populations; where $C = 0$), lower fertility can result in a significant reduction of demographically viable polymorphic parameter space.
-However, the effect is weaker in populations with intermediate selfing rates ($C > 0$) because self-fertilization protects a fraction ($C$) of the ovules of individuals homozygous for the female beneficial allele from exposure to the allele that is deleterious for females.
+However, the effect is weaker in populations with intermediate selfing rates ($C > 0$) because self-fertilization protects a fraction, $C$, of the ovules of individuals homozygous for the female beneficial allele from exposure to the allele that is deleterious for females.
 In other words, self-fertilization leads to less selective death.
@@
  \centering
- %\includegraphics[scale=0.75]{../output/figs/polymorphicSpaceTitrate.pdf}
+ \includegraphics[scale=0.75]{../output/figs/polymorphicSpaceTitrate.pdf}
  \caption{\footnotesize{Proportion of demographically viable parameter space (out of total $s_f \times s_m$ space with $\max(s) = 0.15$) in the absence of inbreeding depression (i.e., assuming $\delta = \delta_i = 0$, where $i \in \{j,a,\gamma\}$), plotted as a function of the population selfing rate.
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection{Demographic effects of inbreeding depression} \label{subsec:InbreedEffects}
-
-Unlike previous population genetic models which assumed constant population sizes \citep{JordanConnallon2014,Olito2017}, mortality caused by inbreeding depression can strongly influence population persistence in our evolutionary demographic model.
+\subsection*{Demographic effects of inbreeding depression} \label{subsec:InbreedEffects}
+
+Unlike previous population genetic models, which assume constant population sizes \citep{JordanConnallon2014,Olito2017}, mortality caused by inbreeding depression can strongly influence population persistence in our evolutionary demographic model.
 Populations with high fertility rates can sustain positive population growth rates despite this higher mortality.
-This leads to a greater proportion of SA polymorphic parameter space to be demographically viable, with the demographic model predictions converging on those from population genetic models in high-fertility populations (fig.~\ref{fig:deltaPolySpace}).
-For populations with lower fertility rates, the proportion of demographically viable polymorphic parameter space matches the population genetic model predictions with the selfing rate is low; however, as the selfing rate increases, demographic viability eventually crashes when the population can no longer sustain the concomitant increase in mortality due to inbreeding depression (fig.~\ref{fig:deltaPolySpace}).
+This causes a greater proportion of SA polymorphic parameter space to be demographically viable, with the demographic model predictions converging on those from population genetic models in high-fertility populations (fig.~\ref{fig:deltaPolySpace}).
+For populations with lower fertility rates, the proportion of demographically viable polymorphic parameter space matches the population genetic model predictions under complete outcrossing; however, as the selfing rate increases, demographic viability eventually crashes when the population can no longer sustain the concomitant increase in mortality due to inbreeding depression (fig.~\ref{fig:deltaPolySpace}).
 
@@
 In contrast to our earlier results (see \hyperref[subsec:PolyExt]{Polymorphism and Extinction}), predominantly outcrossing populations are predicted to have the highest proportion of demographically viable polymorphic space, when inbreeding depression is taken into acount (compare fig.~\ref{fig:polySpace} with fig.~\ref{fig:deltaPolySpace}).
-Regardless of the life-history stage at which inbreeding depression affects survival, populations with intermediate to high selfing rates are unlikely to harbour SA polymorphism unless they can "afford" the resulting loss of self-fertilized ovules/offspring.
-\textcolor{red}{How about the flipside: Populations with intermediate to high selfing rates and growth rates around one are vulnerable to extinction if a sexually antagonistic allele invades the population.
-}
+Regardless of the life-history stage at which inbreeding depression affects survival, populations with intermediate to high selfing rates are unlikely to harbour SA polymorphism unless they can afford the resulting loss of self-fertilized ovules/offspring.
+An alterntaive view of these predictions is that populations with intermediate to high selfing rates and growth rates near one are vulnerable to extinction if a sexually antagonistic allele invades the population (e.g., \citealt{harts2014demography}).
+\textcolor{red}{CITE ostracod examples?
+one issue: sexual dimorphism and sexual conflict doesn't imply SA selection, it depends on the inheritance system.}.
 
@@
  \centering
- %\includegraphics[scale=0.75]{../output/figs/deltaPolymorphicSpaceTitrate.pdf}
+ \includegraphics[scale=0.75]{../output/figs/deltaPolymorphicSpaceTitrate.pdf}
  \caption{\footnotesize{Effects of early- and late-acting inbreeding depression on the proportion of demographically viable parameter space (out of total $s_f \times s_m$ space with $\max(s) = 0.15$), plotted as a function of the population selfing rate.
@@
 
-However, when during the life-cycle inbreeding depression manifests does influence the threshold selfing rate at which demographically viable polymorphic parameter space crashes.
+When during the life-cycle inbreeding depression manifests influences the threshold selfing rate at which demographically viable polymorphic parameter space crashes.
 Our results indicate that population viability was most sensitive to inbreeding depression affecting juvenile survival rates ($\delta_j$; fig.~\ref{fig:deltaPolySpace}, dark blue points), while early-acting inbreeding depression ($\delta$, ovule abortion shortly after fertilization) had a similar effect on population viability as late-acting inbreeding depression affecting adult survial ($\delta_a$) and juvenile-to-adult transition rates ($\delta_{\gamma}$).
@@
 
-\section{Case study: {\itshape M. guttatus}}
-
-%\subsection{Demographic and fitness data for {\itshape M. guttatus}} \label{subsec:MguttMethods}
-
-As an illustrative example of how our model can be used to explore whether demographic rates observed in natural populations appear likely to support balanced SA polymorphisms, we parameterized our model using empirically estimated demographic rates and fitness data for natural populations of the hermaphroditic flowering plant {\itshape Mimulus guttatus} (Scrophulariaceae; now known as {\itshape Erythranthe guttata}).
+\section*{Case study: {\itshape M. guttatus}}
+
+%\subsection*{Demographic and fitness data for {\itshape M. guttatus}} \label{subsec:MguttMethods}
+
+As an illustrative example of how our model can be used to explore whether demographic rates observed in natural populations appear likely to support balanced SA polymorphisms, we parameterized the model using empirically estimated demographic rates and fitness data for natural populations of the hermaphroditic flowering plant {\itshape Mimulus guttatus} (Scrophulariaceae; now known as {\itshape Erythranthe guttata}).
 {\itshape M.~guttatus} is an herbaceous, self-compatible wildflower native to western North America that exhibits remarkable among-population variation in numerous life-history and reproductive traits including selfing rates, inbreeding depression, floral morphology, and annual-to-perennial life-history \citep[e.g.,][]{RitlandGanders1987, Ritland1990, Willis1993, Willis1999a, Willis1999b, WuWillis2008}.
@@
 
-\subsection{Demographic and fitness data for {\itshape M. guttatus}} \label{subsec:MguttMethods}
+\subsection*{Demographic and fitness data for {\itshape M. guttatus}} \label{subsec:MguttMethods}
 We used extensive demographic data reported in a large-scale study of local adaptation using experimental populations of {\itshape M.~guttatus} in Stanislaus National Forest (California, USA) in 2012 and 2013 \citep{PetersonEtAl2016}.
@@
 The vital rate estimates for the Eagle Meadows population are as follows: seed bank survival ($D = 0.534$), seed germination rate ($G = 0.469$), flower production ($F = 0.64$), ovules per flower ($O = 614$), seedling recruits proportional to clonal rosette recruits ($A = 6.7 \times 10^{-4}$), overwinter survival ($S = 0.179$), rosette production ($R = 8.71$).
-The corresponding estimates for the low-elevation perrenials are: $D = 0.534$, $G = 0.652$, \textcolor{red}{$F = 0.4.09$ Should this be 0.409 or 4.09?
-}, $O = 494$, $A = 6.7e-4$), $S = 0$, and $R = 0$ (see corrected Tables.~1 and S2 in \citealt{PetersonEtAl2017}).
+The corresponding estimates for the low-elevation perrenials are: $D = 0.534$, $G = 0.652$, $F = 4.09$, $O = 494$, $A = 6.7e-4$), $S = 0$, and $R = 0$ (see corrected Tables.~1 and S2 in \citealt{PetersonEtAl2017}).
 The same estimates for $D$ and $A$ were used for all populations.
@@
 
-Using these combined demographic rates, selfing rate, and inbreeding depression estimates, we constructed a corresponding stage $\times$ genotype mendelian matrix model with a single SA locus affecting female and male fertility (as described above).
+Using these combined demographic rates, selfing rates, and inbreeding depression estimates, we constructed a corresponding stage $\times$ genotype mendelian matrix model with a single SA locus affecting female and male fertility (as described above).
 Using this empirically parameterized model, we are able to make predictions about the scope for demographically viable SA polymorphism at a single locus in hypothetical populations with the same demographic rates as observed in \citet{PetersonEtAl2016}, for a range of selfing and inbreeding depression rates observed in other natural populations.
@@
 inv6 segregates at moderate frequency (about $8\%$), and carriers suffer $\approx 30 \%$ loss in pollen viability, but also increased flower production which varies among years.
-The genetic basis of the fitness effects caused by inv6 are not yet known, but the net result is a "supergene" with strong fitness effects that segregates as a single diallelic locus.
+The genetic basis of the fitness effects caused by inv6 are not yet known, but the net result is a "supergene" with strong fitness effects on both female and male fertility that segregates as a single diallelic locus.
 We estimated dominance and selection coefficients for the effect of inv6 on pollen viability and flower production from the data reported in \citet{LeeKelly2015}.
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection{Polymorphism in M.~guttatus} \label{sec:Mgutt}
+\subsection*{Polymorphism in M.~guttatus} \label{sec:Mgutt}
 
@@
 The Eagle Meadows population was locally adapted with a very high intrinsic growth rate ($\lambda \approx 1.7$).
-This growth rate was high enough that all $s_f \times s_m$ selection parameter space (where $s_f, s_m \in (0,1]$) remained demographically viable, regardless of the selfing rate ($C = 0,0.29$), or the presence of inbreeding depression (fig.~\ref{fig:mimulusFig}A).
-In contrast, the Low-Elevation Perennial population had a much lower intrinsic growth rate, although it was still positive ($\lambda \approx 1.08$).
+This growth rate was high enough that all $s_f \times s_m$ selection parameter space (where $s_f, s_m \in (0,1]$) remained demographically viable, regardless of the selfing rate ($C = 0$ or $0.29$), or the presence of inbreeding depression (fig.~\ref{fig:mimulusFig}A).
+In contrast, the Low-Elevation Perennial population had a much lower, but still positive, intrinsic growth rate ($\lambda \approx 1.08$).
 Due to the slower growth rate, not all of the $s_f \times s_m$ selection parameter space was demographically viable.
 Extinction thresholds appeared under complete outcrossing ($C = 0$) and partial selfing with inbreeding depression ($C = 0.29$), but not when inbreeding depression was omitted (fig.~\ref{fig:mimulusFig}B).
-The different demographic rates from these two populations also resulted in slightly different invasion conditions when inbreeding depression was taken into account (compare dotted lines in fig/~\ref{fig:mimulusFig}A and B).
-
-Interestingly, the polymorphic inversion inv6 falls very near the lower edge of polymorphic parameter space for these hypothetical populations, but is not always predicted to be demographically viable.
+The different demographic rates from these two populations also resulted in slightly different invasion conditions when inbreeding depression was taken into account (compare dotted lines in fig.~\ref{fig:mimulusFig}A and B).
+
+The polymorphic inversion inv6 falls very near the lower edge of polymorphic parameter space for these hypothetical populations, but is not always predicted to be demographically viable.
 When using demographic rates for either the Eagle Meadows or Low-Elevation Perennial populations, inv6 falls within polymorphic space when the selfing rate is at the higher end of empirical estimates for the Iron Mountain population in which inv6 has been documented ($C = 0.29$), but only when inbreeding depression is not taken into account (fig.~\ref{fig:mimulusFig} dashed lines).
@@
 In contrast, when using the demographic rates for Low-Elevation Perennials, an allele/supergene with these fitness effects falls within demographically inviable parameter space.
-That is, the model predicts that SA alleles with selection coefficients of similar magnitude to inv6 will inflict too high of a demographic cost, and are unlikely to be observed in populations with lower intrinsic growth rates like those observed in the Low-Elevation Perennial population.
+That is, populations with lower intrinsic growth rates like those observed in the Low-Elevation Perennials will be unable to support the demographic costs associated with segregating SA alleles with selection coefficients of similar magnitude to inv6.
 
@@
  \centering
- %\includegraphics[scale=0.75]{../output/figs/mimulusinv6Fig.pdf}
+ \includegraphics[scale=0.75]{../output/figs/mimulusinv6Fig.pdf}
  \caption{\footnotesize{Illustration of model predictions using empirically estimated demographic rates for {\itshape M.~guttatus}.
@@
 
-\hl{[SPECULATION!!!]} Finally, the location of inv6 in selection parameter space is interesting itself because in this region an SA allele with net female-beneficial fitness effects similar to inv6 are predicted to either go extinct, or segregate at relatively low frequency, depending on the population selfing-rate and severity of inbreeding depression.
-Although we cannot make specific predictions for inv6 in the Iron Mountain population in which it was observed, it is interesting that the estimated fitness effects place inv6 roughly in parameter space where segregation at low to moderate frequencies is not surprising.
-This rough concordance between theory and data is clearly speculative, but to our knowledge.
-
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
+\section*{Discussion}
 %%%%%%%%%%%%%%%%%%%%%%%%
-\textit{Summary} \newline
-In this paper we developed a stage-structured mendelian matrix model that accommodates selection through male and female sex-functions in populations of simultaneous hermaphrodites with mixed mating systems.
-Using the model, we jointly analyzed the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations.
-We focused our analyses on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population is also demographically viable (i.e., has a positive intrinsic growth rate).
-We also looked at the effects of inbreeding depression manifesting in different life-history stages, exploring the fitness consequences of early- vs.~late-acting inbreeding depression on the model predictions.
-Finally, we gave an illustrative example of the potential for SA polymorphism in real populations by parameterizing our model with empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus} (now {\itshape Erythranthe guttata}).
-
-A key finding of our study is that when the intrinsic population growth rate is close to $1$, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space for SA polymorphism.
-Furthermore, this demographically viable polymorphic parameter space is often biased towards alleles with stronger selection through the female- than male sex function.
-Self-fertilization can alleviate the demographic costs of balanced SA polymorphisms under some conditions, however, the concommitant effects of inbreeding depression generally exacerbate them in populations with mixed-mating systems.
-Despite these demographic pitfalls, our example using {\itshape M.~guttatus} appears to show that demographic rates observed in some real populations are capable of sustaining large regions of viable SA polymorphic space.
+
+Classic population genetics theory predicts that SA selection is unlikely to maintain genetic variation except under narrow conditions, with polymorphism requiring either finely balanced or unusually strong selection, or partially recessive fitness effects through each sex \citep{Kidwell1977,Pamilo1979,Prout2000, ConnallonClark2014}.
+Extensions of the theory have identified numerous ways in which the conditions for polymorphism become more permissive in both dioecious and hermaphroditic organisms, including genetic linkage of SA loci, population subdivision, and life-cycle complexity \citep[e.g.,][]{Patten2010,JordanConnallon2014,JordanCharlesworth2012, ConnallonSharmaOlito2019, Olito-etal-2018}.
+However, by ignoring the demographic consequences of SA genetic variation, these population genetic models have missed the possibility that SA polymorphisms may be inviable under realistic parameter conditions, and therefore unlikely to be observed in natural populations.
+By linking the individual-level fitness consequences of SA selection to population level demographic rates, our theoretical framework provides several key insights into the processes shaping SA genetic variation in natural populations.
+
+The first and central finding of our study is that when intrinsic population growth rates are close to one, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space where SA polymorphism is maintained.
+\textcolor{red}{Importantly, evolutionary demography theory predicts that many populations' intrinsic growth rates will lie near $\lambda \approx 1$ (Lotte, do you have citation suggestions?), suggesting that our model predictions may be highly relevant for SA polymorphism in many real-world populations}.
+Moreover, these findings compliment recent theoretical and empirical studies indicating that SA selection is likely to be both condition dependent, and stronger in locally-adapted populations near the center of a species' range \citep{BergerEtAl2014, Connallon2015}.
+
+We also find that demographically viable polymorphic parameter space is often biased towards alleles with stronger selection through the female- than male sex function; conditions where female-deleterious alleles are predicted to segregate at lower frequencies, lessening the demographic cost of polymorphism.
+Given natural variation in population growth rates, our model suggests that, on average, SA polymorphisms involving low-frequency female-deleterious SA alleles are more likely to be demographically viable, and therefore observable in real populations.
+Once again, this prediction is supports experimental results indicating that male-beneficial SA genotypes are more susceptible to extinction under environmental stress or inbreeding \citep{BergerEtAl2014, berger2016intralocus, grieshop2017male}.
+Furthermore, our prediction that demographically viable SA variation should often favour female-beneficial genotypes is broadly consistent with the well established pattern of increased allocation to female reproduction in mixed-mating and selfing species (i.e., "selfing syndromes"; reviewed in \citealt{HarderBarrett2006, Sicard2011}).
+
+In hermaphroditic populations, self-fertilization can alleviate the demographic costs of balanced SA polymorphisms under some conditions, however, the concommitant effects of inbreeding depression generally exacerbate them in populations with mixed-mating systems.
+This prediction is in stark contrast to previous population genetics models of SA selection in hermaphrodites, where the sole effect of inbreeding depression is to reduce the population effective selfing rate through the loss of selfed-zygotes, thereby expanding polymorphic parameter space \citep{JordanConnallon2014, Olito2017}.
+In our model, this reduction in effective selfing is accompanied by significant mortality due to inbreeding depression (whether early- or late-acting), which can quickly tip partially-selfing populations over the brink to extinction.
+Beyond the maintenance of SA polymorphisms, this finding underscores a simple but important point that is often overlooked in studies of the evolution of self-fertilization and selfing-syndromes, which tend to emphasize the coevolution of the deleterious mutation load and mating system \citep[e.g.,][]{Charlesworth1987, LandeSchemske1985, Goodwillie2005}: populations with high fertility rates are better able to sustain positive growth rates in the face of severe demographic costs of inbreeding.
+This suggests that traits related to female fecundity, such as ovule and flower production, in ancestral populations likely play an important role in determining the distribution of transitions to self-fertilization among hermaphroditic taxa as well as a variety of ecological correlates of selfing and mixed-mating \citep{Goodwillie2005,Igic2006,Grossenbacher2015}.
+
+Despite the demographic pitfalls associated with SA alleles, our example using {\itshape M.~guttatus} appears to show that demographic rates observed in some real populations are capable of sustaining large regions of viable SA polymorphic space.
+The example also appears to provide some empirical support for the conjecure that locally-adapted populations are more likely to harbor SA polymorphisms than marginal or non-locally adapted ones; inviable polymorphic parameter space only occured when using the demographic data for non-local high-elevation perennial populations.
+The location of inv6 in selection parameter space is also interesting because it corresponds to a region where an SA allele with net female-beneficial fitness effects similar to inv6 are predicted to either go extinct, or segregate at relatively low frequency, depending on the population selfing-rate and severity of inbreeding depression.
+Although we cannot make specific predictions for inv6 in the Iron Mountain population in which it was observed, it is interesting that the estimated fitness effects place inv6 roughly in parameter space where the empirically observed frequencies at Iron Mountain are not surprising.
+
 Overall, our findings provide a more nuanced picture of the nature of SA genetic variation that we should expect to find in natural populations, where the fate of SA alleles and the populations harboring them is determined jointly by evolutionary and demographic processes.
 
-\textit{Extensions: demography} \newline
+\subsubsection*{Extensions and future directions}
+
 By combining the tools of demography and those of population genetics, the framework we presented here enables the exploration of interactions between life cycle complexity, mating system, and sexual antagonism.
-For the sake of simplicity, we used a basic life cycle with just two stages for most of this paper, adults and juveniles (the \textit{Mimulus} example has 3 stages).
-By including more age classes, we can ask if the scope for SA to maintain polymorphisms is affected by whether a species exhibits positive, or negative senescence \citep{jones2014diversity}?
+For the sake of simplicity, we used a basic life cycle with just two stages for most of this paper, adults and juveniles (the \textit{Mimulus} example has $3$ stages).
+However, it is possible to include additional age classes, allowing us to explore whether the scope for SA selection to maintain polymorphisms is affected by whether a species exhibits positive, or negative senescence \citep{jones2014diversity}?
 That is, does the shape of the survival curve affect the area of demographically viable polymorphic parameter space, and how does this interact with the age at which a sexually antagonistic allele is expressed?
 What if the allele not only affects male and female sex functions but also affects survival of individuals?
-And last but not least, how would the scope for SA to maintain polymorphisms be affected by density-dependence acting on different life history stages?
-
-\textit{Extensions: genetics} \newline
-We made the simplifying assumption that only how individuals themselves were produced affects the level of inbreeding depression they experience.
-In reality, the history of inbreeding in the entire lineage will influence the level of inbreeding depression experienced by an individual.
-This could be included in our framework by expanding the individual state space from produced via outcrossing or selfing towards produced via produced via outcrossing or first generation selfing, second generation selfing, etc, as in \cite{kelly1999response} and \cite{kelly2007mutation}.
-
+And last but not least, how would the scope for SA selection to maintain polymorphisms be affected by density-dependence acting on different life history stages?
+
+One major simplifying assumption in our model was that how individuals themselves were produced (i.e., by selfing or outcrossing) affects the level of inbreeding depression they suffer.
+In reality, the history of consecutive generations of inbreeding in each individual's lineage will influence the severity of inbreeding depression they experience, particularly when inbreeding depression is caused primarily by recessive deleterious mutations.
+This more precise approach to modelling inbreeding could be included in our framework by expanding the individual state space to include selfing cohorts (i.e., first generation selfing, second generation selfing, etc.), as in the models of \citet{kelly1999response,kelly2007mutation}.
+
+\textcolor{red}{This needs some polishing, but I though you may want to embellish.}.
 \cite{grieshop2017male} found that genotypes with sexually antagonstic alleles that are male beneficial experience higher levels of inbreeding depression than genotypes with female beneficial SA alleles.
-This effect can easily be included in our model by making the inbreeding depression parameters, $\delta$ and $\delta_j$, a function of the genotype of the individual.
-Such an effect would even further reduce the demographically viable polymorphic parameter space.
-
-Other interesting extensions worth mentioning?
-
-\textit{Conclusion} \newline
-Despite a surge of interest in eco-evolutionary dynamics, the demographic consequences of intralocus sexual antagonism have rarely been modeled (but see \cite{harts2014demography} and \cite{kokko2003sexy}).
-In contrast, models of the population dynamical consequences of interlocus sexual conflict are more common (e.g., \cite{tanaka1996sexual} and \cite{martinez2017sexual}).
-We found that including basic demography can have a significant impact on traditional population genetic results, as has been suggested previously both by theoretical work \citep{kokko2003sexy} and empirical work \citep{berger2016intralocus,grieshop2017male}.
-Although the potential negative consequences of sexual conflict for population viability have been known for some time, this aspect of sexual antagonism has rarely been considered in models investigating the scope for intralocus sexual antagonism to maintain polymorphisms, and hence genetic diversity.
+This effect can easily be included in our model by making the inbreeding depression parameters, $\delta$ and $\delta_i$, a function of the genotype of the individual.
+Such an effect would further reduce the demographically viable polymorphic parameter space, and increase the bias in viable SA polymorphisms towards alleles with weaker selection in females.
+
+\subsubsection*{Conclusion}
+
+Despite a surge of interest in eco-evolutionary dynamics, the demographic consequences of intralocus sexual antagonism have rarely been modeled (but see \citealt{harts2014demography,kokko2003sexy, MatthewsConnallon2019}).
+In contrast, models of the population dynamical consequences of interlocus sexual conflict are more common \citep[e.g.,][]{tanaka1996sexual, martinez2017sexual}.
+We found that including basic demography can have a significant impact on traditional population genetic results, as has been suggested previously both by theoretical and empirical work \citep{kokko2003sexy, berger2016intralocus, grieshop2017male}.
+Although the potential negative consequences of sexual conflict for population viability have been known for some time, this aspect of sexual antagonism has rarely been considered in models investigating the scope for intralocus sexual antagonism to maintain genetic variation.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Conclusion}
+%\section*{Conclusion}
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newpage{}
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+% Appendix A 
+%%%%%%%%%%%%%%%%%%%%%%%%
+\renewcommand{\theequation}{A\arabic{equation}}
+\setcounter{equation}{0}  % reset counter
 
@@
 \end{linenomath*}
-since $\bo K^\tr \bo I_{\omega g}	 \bo K=\bo I_{\omega g}$, or in block matrix form this yields equation (\ref{eq:BlkUS}), here repeated for convenience, 
+since $\bo K^\tr \bo I_{\omega g}	 \bo K=\bo I_{\omega g}$, or in block matrix form this yields equation (\ref{eq:BlkUS_app}), here repeated for convenience, 
 \begin{linenomath*}
@@
 				\mbf{0} & \mbf{0} & \mbf{U}^{S}_{aa} \\
-			\end{array} \right).\label{eq:BlkUS}\\
+			\end{array} \right).\label{eq:BlkUS_app}\\
 \end{eqnarray}
@@
 \begin{eqnarray}
-\mcal{F}^S &=&\bo K^\tr \mbb{H}^S\bo K \mathbb{F}^S, \label{eq:KHKF_app}\\
+\mcal{F}^S &=&\bo K^\tr \mbb{H}^S\bo K \mathbb{F}^S, \label{eq:KHKF_app2}\\
  &=& \sum_{i=1}^g \bo K^\tr \left( \bo I_\om \kron \bo H^S \right)  \bo K \left( \mathbf{E}_{ii} \otimes \mathbf{F}^S_{i} \right).
@@
 \renewcommand{\thetable}{A\arabic{table}}
-\setcounter{equation}{0}  % reset counter 
 \setcounter{figure}{0}
@@
 
-\section{Appendix B: Additional Methods}
+%\section{Appendix B: Additional Methods}
 
@@
 % Dryad references.
+\newpage{}
 \bibliography{Refs2.bib}
-
-\newpage{}
 
@@
 \label{tab:Terms}
-\begin{tabular}{ l l r }
+\begin{tabular}{ l p{0.7\linewidth} r }
  \toprule
@@
 \hline
-$g$      & number of genotypes ($3$; $AA$, $Aa$, and $aa$) & \\
-$\omega$ & number of stages ($2$; juvenile and adult) & \\
+$g$      & Number of genotypes ($3$; $AA$, $Aa$, and $aa$) & \\
+$\omega$ & Number of stages ($2$; juvenile and adult) & \\
 $\tilde{\mbf{n}}$ & Joint stage $\times$ genotype vector & $2 \omega g \times 1$ \\
 $\tilde{\mbf{p}}$ & Joint stage $\times$ genotype frequency vector & $2 \omega g \times 1$ \\
-
-$\mbf{U}^{S}_{i}$, $\mbf{U}^{X}_{i}$ & Genotype-specific transition and survival matrices & $\omega \times \omega$ \\
-$\mbf{F}_{i}$, $\mbf{F}^{\prime}_{i}$ & Genotype-specific fertility matrices & $\omega \times \omega$ \\
-$\mbf{H}^{S}_{i}$, $\mbf{H}^{X}_{i}$ & Parent-offspring genotype maps & $g \times g$ \\
+$\mbf{U}^{S}_{i}$,\, $\mbf{U}^{X}_{i}$ & Genotype-specific transition and survival matrices & $\omega \times \omega$ \\
+$\mcal{U}^{S},\, \mcal{U}^{X}$ & Block diagonal selfed/outcrossed survival matrices & $\omega g \times \omega$ \\
+$\mbf{F}_{i}$,\, $\mbf{F}^{\prime}_{i}$ & Genotype-specific fertility matrices & $\omega \times \omega$ \\
+$\mbb{F}^{\prime}$ & Block diagonal fertility matrix & $\omega g \times \omega$ \\
+$\mcal{F}^{S},\, \mcal{F}^{X}$ & Block diagonal selfed/outcrossed fertility matrix & $\omega g \times \omega$ \\
+$\mbf{H}^{S}_{i}$,\, $\mbf{H}^{X}_{i}$ & Parent-offspring genotype maps & $g \times g$ \\
 $\mbf{W}$   & Allele segregation matrix & $\omega \times g$ \\
-$\mbf{K}$   & Vec-permutation matrix & $g \times g$ \\
-$\mbf{I}$   & Identity matrix & given in subscript \\
+% $\mbf{K}$   & Vec-permutation matrix & $g \times g$ \\
+% $\mbf{I}$   & Identity matrix & given in subscript \\
+$q^{\prime}_{A}$,\, $q^{\prime}_{a}$ & Allele frequencies in male gamete pool & $1$ \\
+$\tilde{\mbf{A}}(\tilde{\mbf{p}})$ & Population projection matrix & $2 \omega g \times g$ \\
+
+$s_f,\,s_m$ & Selection coefficients through female and male reproductive function & \\
+$h_f,\,h_m$ & Dominance coefficients through female and male reproductive function & \\
+$\sigma_j,\, \sigma_a$ & Survival rates for juvenile and adult stages & \\
+$\gamma$ & Transition rate from juvenile to adult stages \\
+$C$ & The population rate of self-fertilization & \\
+$\delta,\, \delta_j,\,\delta_a,\,\delta_{\gamma}$ & Inbreeding depression terms for ovule viability, survival, and transition rates respectively & \\
+$f_i,\,f^{\prime}_i$ & Adult fertilities through male and female sex functions & \\
+
 \hline
@@
 
-\begin{table}[h]
-\caption{Founders of \textit{The~American Naturalist}}
-\label{Table:Founders}
-\centering
-\begin{tabular}{lll}\hline
-Early editor            & Years with the journal \\ \hline
-Alpheus S. Packard Jr.  & 1867--1886 \\
-Frederick W. Putnam     & 1867--1874 \\ 
-Edward S. Morse         & 1867--1871 \\ 
-Alpheus Hyatt           & 1867--1871 \\
-Edward Drinker Cope$^a$ & 1878--1897 \\
-J.~S. Kingsley          & 1887--1896 \\ \hline 
-\end{tabular}
-\bigskip{}
-\\
-{\footnotesize Note: Table titles should be short. Further details should go in a `notes' area after the tabular environment, like this. $^a$ Published the first description of \textit{Dimetrodon}.}
-\end{table}
-
-\newpage{}
-
 \section{Figure legends}
@@
 \caption{Figure legends can be longer than the titles of tables. However, they should not be excessively long.}
-\label{Fig:OkapiHorn}
+%\label{Fig:OkapiHorn}
+\end{figure}
+
+\begin{figure}[h!]
+%\includegraphics{elegance}
+\caption{In this way, figure legends can be listed at the end of the document, with references that work, even though the graphic itself should be included for final files after acceptance.
+Instead, upload the relevant figure files separately to Editorial Manager; Editorial Manager should insert them at the end of the PDF automatically.}
+%\label{Fig:AnotherFigure}
 \end{figure}
@@
 
-\renewcommand{\figurename}{Video} 
-\setcounter{figure}{0}
+%\renewcommand{\figurename}{Video} 
+%\setcounter{figure}{0}
 % Thanks to Flo Debarre for the pro tip of putting
@@
 
-\begin{figure}[h!]
+%\begin{figure}[h!]
 %\includegraphics{VideoScreengrab.jpg}
-\caption{Video legends can follow the same principles as figure legends. Counters should be set and reset so that videos and figures are enumerated separately.}
-\label{VideoExample}
-\end{figure}
-
-\renewcommand{\figurename}{Figure}
-\setcounter{figure}{1}
+%\caption{Video legends can follow the same principles as figure legends. Counters should be set and reset so that videos and figures are enumerated separately.}
+%\label{VideoExample}
+%\end{figure}
 
@@
 
-\begin{figure}[h!]
-%\includegraphics{elegance}
-\caption{In this way, figure legends can be listed at the end of the document, with references that work, even though the graphic itself should be included for final files after acceptance.
-Instead, upload the relevant figure files separately to Editorial Manager; Editorial Manager should insert them at the end of the PDF automatically.}
-\label{Fig:AnotherFigure}
-\end{figure}
-
-\subsection{Online figure legends}
-
-\renewcommand{\thefigure}{A\arabic{figure}}
-\setcounter{figure}{0}
-
-\begin{figure}[h!]
+%\renewcommand{\figurename}{Figure}
+%\setcounter{figure}{1}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Online figure legends
+%\subsection*{Online figure legends}
+
+%\renewcommand{\thefigure}{A\arabic{figure}}
+%\setcounter{figure}{0}
+
+%\begin{figure}[h!]
 %\includegraphics{jumps20m}
-\caption{\textit{A}, the quick red fox proceeding to jump 20~m straight into the air over not one, but several lazy dogs.
+%\caption{\textit{A}, the quick red fox proceeding to jump 20~m straight into the air over not one, but several lazy dogs.
 \textit{B}, the quick red fox landing gracefully despite the skepticism of naysayers.}
-\label{Fig:Jumps}
-\end{figure}
-
-\begin{figure}[h!]
+%\label{Fig:Jumps}
+%\end{figure}
+
+%\begin{figure}[h!]
 %\includegraphics{jumps20m}
-\caption{The quicker the red fox jumps, the likelier it is to land near an okapi. For further details,.}
-\label{Fig:JumpsOk}
-\end{figure}
-
-\renewcommand{\thefigure}{B\arabic{figure}}
-\setcounter{figure}{0}
+%\caption{The quicker the red fox jumps, the likelier it is to land near an okapi. For further details,.}
+%\label{Fig:JumpsOk}
+%\end{figure}
+
+%\renewcommand{\thefigure}{B\arabic{figure}}
+%\setcounter{figure}{0}
 

### 2021-03-11 doc/Supplements/AmNat_supp_template.tex
Expanded Discussion, New Abstract, Table, etc.
Expanded Discussion, added a couple alternative abstracts, finished
(hopefully) table of model parameters, and made minor edits
throughout the ms.

Still have not updated appendices. Need to chat about these I think.

\documentclass[11pt]{article}
\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
\usepackage{fullpage}
\usepackage[authoryear,sectionbib,sort]{natbib}
\linespread{1.7}
\usepackage[utf8]{inputenc}
\usepackage{lineno}
\usepackage{titlesec}
\titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
\titleformat{\subsection}[block]{\Large\itshape\filcenter}{\thesubsection}{1em}{}
\titleformat{\subsubsection}[block]{\large\itshape}{\thesubsubsection}{1em}{}
\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1em}{}[. ]
\usepackage{fancyhdr}
\pagestyle{fancy}

%%%%%%%%%%%%%%%%%%%%%
% Header
%%%%%%%%%%%%%%%%%%%%%
%
% Customize the line below with the last name of your first author and
% the short title of your MS. You can comment authorship information out
% while your MS is undergoing double-blind review.
%
\rhead{Supplement to AUTHOR-NAME et al., ``SHORT TITLE,'' \textit{Am.~Nat.}}
\setlength{\headsep}{0.3in}  
\lhead{}

%%%%%%%%%%%%%%%%%%%%%
% Line numbering
%%%%%%%%%%%%%%%%%%%%%
%
% Please use line numbering with your initial submission and
% subsequent revisions. After acceptance, please turn line numbering
% off by adding percent signs to the lines %\usepackage{lineno} and
% to %\linenumbers{} and %\modulolinenumbers[3] below.
%
% To avoid line numbering being thrown off around math environments,
% the math environments have to be wrapped using
% \begin{linenomath*} and \end{linenomath*}
%
% (Thanks to Vlastimil Krivan for pointing this out to us!)

\title{Online Supplement: \\ 
\LaTeX{} Template for Author-Supplied Supplementary PDFs, \\ 
\textit{The~American~Naturalist} }

% This version of the LaTeX supplementary template was last updated on
% November 8, 2019.

%%%%%%%%%%%%%%%%%%%%%
% Authorship
%%%%%%%%%%%%%%%%%%%%%
% Please remove authorship information while your paper is under review,
% unless you wish to waive your anonymity under double-blind review. You
% will need to add this information back in to your final files after
% acceptance.
%
% Once accepted for publication, author-supplied PDFs should have a 
% title page that includes (at least) the authors' names, the title of 
% the MS, and the name of the journal. It should also have a header and
% page numbers.

\author{Owen E. Cook$^{1,\ast}$ \\ 
Generic H. Collaborator$^{2,\dag}$ \\ 
Additional Q. Expert$^{3}$}

\date{}

\begin{document}

\maketitle

\noindent{} 1. University of Chicago, Chicago, Illinois 60637;

\noindent{} 2. University of Toronto, Toronto, Ontario M5S 1A5, Canada;

\noindent{} 3. Middle Eastern Technical University, Çankaya, Ankara 06800, Turkey.

\noindent{} $\ast$ Corresponding author; e-mail: amnat@uchicago.edu.

\noindent{} $\dag$ Deceased.

%\linenumbers{}
%\modulolinenumbers[3]

% In many cases, The American Naturalist allows authors to typeset their
% own supplementary material in an author-supplied PDF. This template
% applies to such cases. 
% 
% For appendices that will be typeset by the AmNat editorial staff, 
% please see the main LaTeX template, available from
% https://www.journals.uchicago.edu/journals/an/instruct
% Such appendices typically include descriptions of methods and tables
% defining parameters.
%
% In general, you have wide discretion for how you want to format an
% author-supplied PDF. They should in any case have a title page, 
% page numbers, and a header identifying the MS's (short) title.
%
% Counters for the online supplement should normally begin with an S
% (thus normally figure S1, figure S2, table S1, equation S1, etc.).

\renewcommand{\theequation}{S\arabic{equation}}
% redefine the command that creates the equation number.
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{equation}{0}  % reset counter 
\setcounter{figure}{0}
\setcounter{table}{0}

% In online supplementary PDFs, sections can be numbered or not
% (at your discretion). If they are numbered, sections should usually
% begin with an S.

\newpage{}

\section*{Supplementary Figures and Tables}

Lorem ipsum dolor sit amet (fig.~\ref{Fig:OkapiStripe}). Curabitur pretium tincidunt lacus. Nulla gravida orci a odio. Nullam varius, turpis et commodo pharetra, est eros bibendum elit, nec luctus magna felis sollicitudin mauris. Integer in mauris eu nibh euismod gravida. Duis ac tellus et risus vulputate vehicula. Donec lobortis risus a elit. Etiam tempor. Ut ullamcorper, ligula eu tempor congue, eros est euismod turpis, id tincidunt sapien risus a quam. Maecenas fermentum consequat mi. Donec fermentum. Pellentesque malesuada nulla a mi. Duis sapien sem, aliquet nec, commodo eget, consequat quis, neque.

\begin{figure}[h!]
%\includegraphics{okapi-stripes}
\caption{Figure legend.}
\label{Fig:OkapiStripe}
\end{figure}

Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

Tables can exist too (see table~\ref{Table:Rivers} for an example).

\begin{table}[htbp]
\caption{Various rivers, cities, and animals}
\label{Table:Rivers}
\centering
\begin{tabular}{lll}\hline
River        & City        & Animal            \\ \hline
Chicago      & Chicago     & Raccoon           \\
Des Plaines  & Joliet      & Coyote            \\
Illinois     & Peoria      & Cardinal          \\
Kankakee     & Bourbonnais & White-tailed deer \\
Mississippi  & Galena      & Bald eagle        \\ \hline
\end{tabular}
\bigskip{}
\\
{\footnotesize Note: Notes can go here.}
\end{table}

\section*{Additional Methods}

\subsection*{Measuring the height of fox jumps}

The quick red fox jumps over the lazy brown dog. Morbi odio velit, eleifend ut hendrerit vitae, consequat sit amet nulla. Pellentesque porttitor vitae nisl quis tempus. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Praesent ut nisi odio.

\begin{equation}
{ \frac{1}{N_k-1} \sum \limits_{t=1}^{N_k} (M_{tjk} - \bar{M}_{jk})^2}
\end{equation}

\subsection*{Quantifying the brownness of the dog}

Pellentesque eu nulla odio (\citealt{Xiao2015,CookEtAl2015}). Nulla aliquam porta metus, quis malesuada orci faucibus quis. Suspendisse nunc magna, tristique sit amet sollicitudin nec, elementum et lacus. Sed vitae elementum mi. In hac habitasse platea dictumst. Etiam eu tortor elit. Sed ac tortor purus. Aliquam volutpat, odio sit amet posuere pretium, dolor ex interdum ante, sed luctus quam eros ac nulla.

\begin{equation}
{ (\sum \limits_{p=1}^P {n_{sp}})^{-1}\sum \limits_{p=1}^P {n_{sp}Q_{p}}}
\end{equation}

\newpage{}

%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%
%
% Any references in this section should also be copied into the main 
% template, under the heading
% \section*{References Cited Only in the Online Enhancements}
% except when they are already cited in the main text.

\begin{thebibliography}{}

\bibitem[{Cook et~al.(2015)Cook, Collaborator, and Expert}]{CookEtAl2015}
Cook, O.~E., G.~H. Collaborator, and A.~Q. Expert. 2015.
\newblock Data from: Template and guidelines for using \LaTeX{} in \textit{The American Naturalist}.
\newblock American Naturalist, Dryad Digital Repository, https://dx.doi.org/10.5061/dryad.XYZAB123.

\bibitem[{Darwin(1859)Darwin}]{OriginSpec}
Darwin, C. 1859.
\newblock On the origin of species by means of natural selection, or the preservation of favoured races in the struggle for life.
\newblock J.~Murray, London.

\bibitem[{Davis et~al.(2011)Davis, Brakora, and Lee}]{DavisEtAl2011}
Davis, E.~B., K.~A. Brakora, and A.~H. Lee. 2011.
\newblock Evolution of ruminant headgear: a review.
\newblock Proceedings of the Royal Society~B 278:2857--2865.

\bibitem[{Inglis et~al.(2011)Inglis, Roberts, Gardner, and Buckling}]{Ing11}
Inglis, R.~F., P.~G. Roberts, A.~Gardner, and A.~Buckling. 2011.
\newblock Spite and the scale of competition in \textit{Pseudomonas
  aeruginosa}.
\newblock American Naturalist 178:276--285.

\bibitem[{Fastovsky(2009)Fastovsky}]{LemKapEx07}
Fastovsky, D.~E. 2009.
\newblock Ideas in dinosaur paleontology: resonating to social and political context.
\newblock Pages 239--253 \emph{in} D. Sepkoski and M. Ruse, eds. The Paleobiological Revolution. University of Chicago Press, Chicago~IL.

\bibitem[{Xiao et~al.(2015)Xiao, McGlinn, and White}]{Xiao2015}
Xiao, X., D.~J. McGlinn, and E.~P. White. 2015.
\newblock A strong test of the maximum entropy theory of ecology.
\newblock American Naturalist 185:E705--E80.

\end{thebibliography}

\end{document}

### 2021-03-11 makeFigs.R

@@ -46,6 +46,7 @@ toPdf(MimulusInv6Fig(),
 			figPath(name='MimulusInv6Fig.pdf'), width=10, height=5)
 embed_fonts(figPath(name='MimulusInv6Fig.pdf'))
 
+
 #################
 # Supp. Figs
 #################

### 2021-03-25 doc/SA-Hermaphrodites-wDemography.tex
Minor edits to the main text

@@
-\documentclass[11pt]{article}
+\documentclass[11pt,draft]{article}
 % Preamble
@@
 \title{Evolutionary demography and the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites}
+
 \author{Colin Olito$^{1,\ast}$ \\ 
@@
 \newpage{}
+\section{Title options? \title{Population viability affects/impacts the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites}}
+
+ \section{Demographic viability of antagonistic alleles impacts conditions for the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites}
+
+ \section{Conditions for the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites impacted by considering population viability}
+
 \section*{Abstract -- currently 215/200 word max.}
@@
 In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of sexually antagonistic (SA) selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations.
-We focus on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population intrinsic growth rate remains positive.
+We focus on identifying the parameter conditions under which SA polymorphisms are maintained {\itshape and} the population intrinsic growth rate remains positive.
 We also analyze the effects of inbreeding depression manifesting in different life-history stages.
-Finally, we gave an illustrative example of the potential for SA polymorphism in real populations using empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus}.
+Finally, we show an illustrative example of the potential for SA polymorphism in real populations using empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus}.
 Our results indicate that when population growth rates are near one, large swathes of polymorphic parameter space become demographically inviable, and that inbreeding depression poses a significant challenge to the maintenance of SA polymorphism in partially selfing populations.
@@
 \newpage{}
-\section*{ALTERNATIVE Abstract -- currently 200/200 word max.}
+\section*{ALTERNATIVE Abstract -- currently 200/200 word max.\textcolor{green}{ I like this abstract best}}
+
 % Alternative opening sentence to the abstract:sentence
-Now classic population genetics theory has been fundamental to understanding the evolution of sex-differences and the maintenance of sexually antagonistic (SA) genetic variation, but these models have rarely considered the demographic consequences of intralocus sexual antagonism.
+Classic population genetics theory has been fundamental to understanding the evolution of sex-differences and the maintenance of sexually antagonistic (SA) genetic variation, but these models have rarely considered the demographic consequences of intralocus sexual antagonism.
 In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations.
-We focus on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection and the population intrinsic growth rate remains positive.
-We also analyze the effects of inbreeding depression manifesting in different life-history stages, and give an illustrative example of the potential for SA polymorphism in real populations using empirically estimated demographic rates for the hermaphroditic flowering plant Mimulus guttatus.
+We focus on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection \textit{and} the population intrinsic growth rate remains positive.
+In addition, we analyze the effects of inbreeding depression manifesting in different life-history stages, and give an illustrative example of the potential for SA polymorphism in real populations using empirically estimated demographic rates for the hermaphroditic flowering plant Mimulus guttatus.
 Our results indicate that when population growth rates are near one, large swathes of polymorphic parameter space become demographically inviable, and that inbreeding depression poses a significant challenge to the maintenance of SA polymorphism in partially selfing populations.
@@
 
-In order to persist in the long term, a population must be able to adapt to its environment.
+To persist in the long term, a population must be able to adapt to its environment.
 Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, such as deleterious mutations \citep{Haldane1957}, changing environmental conditions \citep{Maynard-Smith1976, LandeShannon1996,OrrUnckless2008}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, and genetic constraints \citep{ConnallonHall2018, MatthewsConnallon2019}.
@@
 
-Sexually antagonistic selection (abbreviated SA hereafter) arising from genetic trade-offs between male and female fitness, is a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and also the evolution of sexual dimorphism and gender-related traits \citep{Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, BondurianskyChenoweth2009,Olito2019}.
+Sexually antagonistic selection (abbreviated SA hereafter) arising from genetic trade-offs between male and female fitness, is a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and the evolution of sexual dimorphism and gender-related traits \citep{Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, BondurianskyChenoweth2009,Olito2019}.
 When such genetic trade-offs occur, SA selection can arise when beneficial alleles for one sex are deleterious when expressed in the other \citep{Kidwell1977, Rice1992, ConnallonClark2012}.
+
 Alleles with opposing fitness effects through male and female sex functions can cause analogous genetic constraints on fitness in hermaphrodite populations, where both maternal and paternal reproductive success contribute jointly to each individuals' overall fitness \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
-
-The evolutionary dynamics of sexually antagonistic alleles in hermaphrodites differ from those in dioecious populations because individuals may reproduce through a combination of self- and outcross fertilization \citep{Goodwillie2005, Igic2006, JarneAuld2006}.
-For example, self-fertilization is predicted to reduce the total parameter space where balanced SA polymorphisms can be maintained, while simultaneously creating a bias in selection through the female sex function (\citealt{JordanConnallon2014}; but see \citealt{Tazzyman2015}).
+Self-fertilization in hermaphrodites is predicted to reduce the total parameter space where balanced SA polymorphisms can be maintained, while simultaneously creating a bias in selection through the female sex function (\citealt{JordanConnallon2014}; but see \citealt{Tazzyman2015}).
 Yet, other factors, such as genetic linkage to other SA loci or a sex-determining region \citep{Otto2011, JordanCharlesworth2012, Olito2017, Olito2019}, or spatial heterogeneity and complexity of the life-cycle can expand the parameter space for SA polymorphism \citep{Olito-etal-2018,ConnallonSharmaOlito2019}.
@@
 Most population genetic models, including models of the maintenance of SA polymorphisms, keep track of the relative fitnesses of genotypes in populations of constant size, and therefore do not consider the population dynamical consequences of evolutionary change.
-In fact, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals.
+In fact, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals (e.g., \cite{johnston2013life},\cite{merot2020balancing}).
 To take the population dynamical consequences of polymorphism into account, a model is required that links changes in genotype distributions to population dynamics.
-For example, \cite{harts2014demography} developed an individual-based modelling approach that tracks population dynamics and the spread of both SA genes and dispersal genes in a species with non-overlapping generations.
-\cite{deVriesCaswell2019b} introduced a mendelian matrix model with overlapping generations, intralocus sexual antagonism, and population dynamics but did not perform analysis of the consequences of demographic viability for the scope of sexual antagonism to maintain polymorphism.
+The potential demographic costs of sexual antagonism were pointed out by \cite{kokko2003sexy}, but few papers since then have taken this into account (with the notable exception of \cite{harts2014demography}).
+\cite{deVriesCaswell2019b} introduced a mendelian matrix model with intralocus sexual antagonism, and population dynamics but did not perform an analysis of the consequences of demographic viability for the scope of sexual antagonism to maintain polymorphism.
 
@@
 			\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime}}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime} \|} = 
-				\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)\|}
+				\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)\|}.
 \end{equation}
@@
 				0 & \frac{1}{2} q^{\prime}_{a} & q^{\prime}_{a} \\
-			\end{array} \right)
+			\end{array} \right).
 \end{equation}
@@
 
-Reproduction via self-fertilization differs from outcrossing.
+The parent-offpsring map for reproduction via self-fertilization differs from the outcrossing parent-offspring map.
 As in previous models of partial selfing, we assume that individuals produce enough male gametes to easily self-fertilize all of their ovules, and that self-fertilization involves little or no selection from external factors relative to outcrossing (e.g., \citealt{Charlesworth1978a,JordanConnallon2014,Olito2017}, but see \citealt{Tazzyman2015}).
@@
 
-The blocks of the component matrices in Eq(\ref{eq:Atilde}) correspond to production of offspring by self-fertilization and outcrossing ($\mcal{F}^S$ and $\mcal{F}^X$ in $\tilde{\mbf{F}}$), and survival of selfed and outcrossed offspring ($\mcal{U}^S$ and $\mcal{U}^X$ in $\tilde{\mbf{U}}$).
+The blocks of the component matrices in Eq(\ref{eq:Atilde}) correspond to production of offspring by self-fertilization and outcrossing ($\mcal{F}^S$ and $\mcal{F}^X$ in $\tilde{\mbf{F}}$), and survival of extant individuals produced by selfing or outcrossing ($\mcal{U}^S$ and $\mcal{U}^X$ in $\tilde{\mbf{U}}$).
 The survival matrix for individuals produced through selfing and outcrossing are,
@@
 We now construct and analyze a genotype $\times$ stage-classified model for a hypothetical species with intralocus sexual conflict via the two sex functions.
-To keep the model analytically tractable, we assume our hypothetical species has two life stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active.
+For the sake of simplicity, we assume our hypothetical species has a basic life cycle with only two life stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active.
 Suppose that there is a genetic trade-off between the sex-functions at a single diallelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect.
@@
 
-Iterating the projection matrix, Eq(\ref{eq:Atilde}) with the above demographic matrices, given an initial population state vector, Eq(\ref{eq:PopStateVec}), allows numerical simulation of the eco-evolutionary dynamics for selection operating on any of the sex-function, or stage-specific demographic parameters.
+Iterating the projection matrix, Eq(\ref{eq:Atilde}) with the above demographic matrices, given an initial population state vector, allows numerical simulation of the eco-evolutionary dynamics for selection operating on any of the sex-function, or stage-specific demographic parameters.
 As we outline below, we use numerical techniques together with mathematical analyses to study the conditions for the maintenance of SA polymorphisms, and the demographic fate of the populations (i.e., positive growth, or extinction).
@@
 
-Diverse eco-evolutionary outcomes are possible in the model, including fixation of either allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and suicide (e.g., see \citealt{CaswellEtAl2018,deVriesCaswell2019a,deVriesCaswell2019b}).
+Diverse eco-evolutionary outcomes are possible in the model, including fixation of either allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and suicide (e.g., see \citealt{deVriesCaswell2019a,deVriesCaswell2019b}).
 We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection and ({\itshape ii}) the intrinsic population growth rate at equilibrium is positive; a situation that we refer to as a 'demographically viable SA polymorphism'.
@@
 Specifically, for each boundary equilibrium we introduced the rare allele at low initial frequency, and iterated Eq(\ref{eq:Atilde}) until the population had reached demographic and genotypic equilibrium.
-In our model, the population state vector will grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see \citealt{Caswell2001}), and the intrinsic population growth rate after convergence, $\lambda$, can be calculated as $\tilde{\mbf{n}}(t)/\tilde{\mbf{n}}(t-1)$.
+In our model with density independent demographic rates, the population state vector will grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see chapter 17 in \citealt{Caswell2001}), and the intrinsic population growth rate after convergence, $\lambda$, can be calculated as $\tilde{\mbf{n}}(t)/\tilde{\mbf{n}}(t-1)$.
 We note, however, that if the ecological component of the model is non-linear, more exotic dynamics are possible \citep{de2020matrix}.
@@
 For simplicity, we also assume equal dominance in all analyses (i.e., $h_f = h_m = h$).
-For a given parameter set, we numerically evaluated the largest eigenvalue of the Jacobian evaluated at both boundaries, and the population growth rate of the equilibrium obtained after perturbation of the boundary, $\lambda$.
-To determine the viability threshold (where the population growth rate equals one) and the boundary stability thresholds (where the largest eigenvalue of the Jacobian crosses one), we start at the outer boundaries of selection parameter space, and titrated inward.
+For a given parameter set, we numerically evaluate the largest eigenvalue of the Jacobian evaluated at both boundaries, and the population growth rate of the equilibrium obtained after perturbation of the boundary, $\lambda$.
+To determine the viability threshold (where the population growth rate equals one) and the boundary stability thresholds (where the largest eigenvalue of the Jacobian crosses one), we start at the outer boundaries of selection parameter space, and titrate inward.
 The proportions of total $s_f \times s_m$ parameter space where SA polymorphism and/or extinction occurred were calculated by numerically integrating the appropriate regions defined by the invasion and extinction thresholds
@@
 
-We begin with a simple illustration of demographically viable polymorphic parameter space in the absence of inbreeding depression (i.e., $\delta = \delta_i = 0$).
+We begin with an illustration of demographically viable polymorphic parameter space in the absence of inbreeding depression in Figure \ref{fig:extThresholds} (i.e., $\delta = \delta_i = 0$).
 Invasion conditions for SA alleles in the evolutionary demographic model closely match the predictions from population genetic models (see \hl{Online Appendix X}; \citealt{Kidwell1977, JordanConnallon2014, Olito2017}).
 In particular, the demographic model recovers the classic "funnel-shaped" region of polymorphic $s_f \times s_m$ parameter space.
-The effects of the population selfing rate ($C$) and dominance ($h$) on SA polymorphism are also similar: self-fertilization ({\itshape i}) skews the SA polymorphic space towards regions with weaker selection through female-function and reduces the total area of polymorphic parameter space; and ({\itshape ii}) dominance reversals (where deleterious SA fitness effects are partially recessive in each sex; $h = 1/4$) are much more permissive of SA polymorphism \citep{JordanConnallon2014, Olito2017}.
+The effects of the population selfing rate ($C$) and dominance ($h$) on SA polymorphism are also similar: self-fertilization ({\itshape i}) skews the SA polymorphic space towards regions with weaker selection through female-function and reduces the total area of polymorphic parameter space (e.g., contrast figure 1A with figure 1C); and ({\itshape ii}) dominance reversals (where deleterious SA fitness effects are partially recessive in each sex; $h = 1/4$) are much more permissive of SA polymorphism (e.g., contrast Figure 1A with Figure 1D) \citep{JordanConnallon2014, Olito2017}.
 
@@
 In obligately outcrossing populations (including dioecious/gonochoristic populations; where $C = 0$), lower fertility can result in a significant reduction of demographically viable polymorphic parameter space.
-However, the effect is weaker in populations with intermediate selfing rates ($C > 0$) because self-fertilization protects a fraction, $C$, of the ovules of individuals homozygous for the female beneficial allele from exposure to the allele that is deleterious for females.
+However, the effect is weaker in populations with intermediate selfing rates (compare $C > 0$ vs $C=0$) because self-fertilization protects a fraction, $C$, of the ovules of individuals homozygous for the female beneficial allele from exposure to the allele that is deleterious for females.
 In other words, self-fertilization leads to less selective death.
 This allows partially selfing populations to remain viable under selection intensities that would cause extinction in an outcrossing population.
-Interestingly, when fertility is low enough that some polymorphic parameter space is demographically inviable, populations with intermediate selfing rates have the greatest proportion of demographically viable parameter space (fig.~\ref{fig:polySpace}, med.~and low.~fertility values).
-This effect emerges from the combination of protection from selective death and reduced total polymorphic parameter space caused by selfing.
+The combination of protection from selective death and reduced total polymorphic parameter space caused by selfing results in populations with intermediate selfing rates having the greatest proportion of demographically viable parameter space at medium and low fertilities (fig.~\ref{fig:polySpace}, med.~and low.~fertility values).
 
@@
 Mortality from inbreeding depression reduces the effect of protection from selective death that individuals gain through self fertilization.
-In contrast to our earlier results (see \hyperref[subsec:PolyExt]{Polymorphism and Extinction}), predominantly outcrossing populations are predicted to have the highest proportion of demographically viable polymorphic space, when inbreeding depression is taken into acount (compare fig.~\ref{fig:polySpace} with fig.~\ref{fig:deltaPolySpace}).
+In contrast to our earlier results (see Figure \ref{fig:polySpace} in \hyperref[subsec:PolyExt]{Polymorphism and Extinction}), predominantly outcrossing populations are predicted to have the highest proportion of demographically viable polymorphic space, when inbreeding depression is taken into acount (compare fig.~\ref{fig:polySpace} with fig.~\ref{fig:deltaPolySpace}).
 Regardless of the life-history stage at which inbreeding depression affects survival, populations with intermediate to high selfing rates are unlikely to harbour SA polymorphism unless they can afford the resulting loss of self-fertilized ovules/offspring.
-An alterntaive view of these predictions is that populations with intermediate to high selfing rates and growth rates near one are vulnerable to extinction if a sexually antagonistic allele invades the population (e.g., \citealt{harts2014demography}).
-\textcolor{red}{CITE ostracod examples?
-one issue: sexual dimorphism and sexual conflict doesn't imply SA selection, it depends on the inheritance system.}.
+An alternative view of these predictions is that populations with intermediate to high selfing rates and growth rates near one are vulnerable to extinction if a sexually antagonistic allele invades the population.
 
@@
 
-When during the life-cycle inbreeding depression manifests influences the threshold selfing rate at which demographically viable polymorphic parameter space crashes.
+The point in the life-cycle where inbreeding depression manifests influences the threshold selfing rate at which demographically viable polymorphic parameter space crashes.
 Our results indicate that population viability was most sensitive to inbreeding depression affecting juvenile survival rates ($\delta_j$; fig.~\ref{fig:deltaPolySpace}, dark blue points), while early-acting inbreeding depression ($\delta$, ovule abortion shortly after fertilization) had a similar effect on population viability as late-acting inbreeding depression affecting adult survial ($\delta_a$) and juvenile-to-adult transition rates ($\delta_{\gamma}$).
@@
 Based on this assumption, the selection coefficient inv6 on flower production in the field was $s_f = 0.384$ in 2012, and $s_f = 0.232$ in 2013.
-As a final proof-of-concept test of our empirically parameterized model \citep[e.g.,][]{Servedio2014}, we asked whether, given the above biologically grounded fitness effect estimates, inv6 appears to fall in demographically viable SA polymorphic parameter space, as might be expected given its observed frequency in the Iron Mountain population.
+As a final proof-of-concept test of our empirically parameterized model \citep[e.g.,][]{Servedio2014}, we asked whether, given the above biologically grounded fitness effect estimates, inv6 appears to fall in demographically viable SA polymorphic parameter space (see Figure \ref{fig:mimulusFig}), as might be expected given its observed frequency in the Iron Mountain population.
 
@@
 
-Classic population genetics theory predicts that SA selection is unlikely to maintain genetic variation except under narrow conditions, with polymorphism requiring either finely balanced or unusually strong selection, or partially recessive fitness effects through each sex \citep{Kidwell1977,Pamilo1979,Prout2000, ConnallonClark2014}.
+Classic population genetics theory predicts that sexually antagonistic selection is unlikely to maintain genetic variation except under narrow conditions, with polymorphism requiring either finely balanced or unusually strong selection, or partially recessive fitness effects through each sex \citep{Kidwell1977,Pamilo1979,Prout2000, ConnallonClark2014}.
 Extensions of the theory have identified numerous ways in which the conditions for polymorphism become more permissive in both dioecious and hermaphroditic organisms, including genetic linkage of SA loci, population subdivision, and life-cycle complexity \citep[e.g.,][]{Patten2010,JordanConnallon2014,JordanCharlesworth2012, ConnallonSharmaOlito2019, Olito-etal-2018}.
@@
 The first and central finding of our study is that when intrinsic population growth rates are close to one, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space where SA polymorphism is maintained.
-\textcolor{red}{Importantly, evolutionary demography theory predicts that many populations' intrinsic growth rates will lie near $\lambda \approx 1$ (Lotte, do you have citation suggestions?), suggesting that our model predictions may be highly relevant for SA polymorphism in many real-world populations}.
+Since growth rates far exceeding one suggest rapid exponential growth, they are generally rare (with the notable exception of recently introduced, invasive populations), suggesting that our model predictions may be highly relevant for SA polymorphism in many real-world populations.
 Moreover, these findings compliment recent theoretical and empirical studies indicating that SA selection is likely to be both condition dependent, and stronger in locally-adapted populations near the center of a species' range \citep{BergerEtAl2014, Connallon2015}.
@@
 Given natural variation in population growth rates, our model suggests that, on average, SA polymorphisms involving low-frequency female-deleterious SA alleles are more likely to be demographically viable, and therefore observable in real populations.
-Once again, this prediction is supports experimental results indicating that male-beneficial SA genotypes are more susceptible to extinction under environmental stress or inbreeding \citep{BergerEtAl2014, berger2016intralocus, grieshop2017male}.
+Once again, this prediction is supported by experimental results indicating that male-beneficial SA genotypes are more susceptible to extinction under environmental stress or inbreeding \citep{BergerEtAl2014, berger2016intralocus, grieshop2017male}.
 Furthermore, our prediction that demographically viable SA variation should often favour female-beneficial genotypes is broadly consistent with the well established pattern of increased allocation to female reproduction in mixed-mating and selfing species (i.e., "selfing syndromes"; reviewed in \citealt{HarderBarrett2006, Sicard2011}).
@@
 The location of inv6 in selection parameter space is also interesting because it corresponds to a region where an SA allele with net female-beneficial fitness effects similar to inv6 are predicted to either go extinct, or segregate at relatively low frequency, depending on the population selfing-rate and severity of inbreeding depression.
-Although we cannot make specific predictions for inv6 in the Iron Mountain population in which it was observed, it is interesting that the estimated fitness effects place inv6 roughly in parameter space where the empirically observed frequencies at Iron Mountain are not surprising.
+Although we cannot make specific predictions for inv6 in the Iron Mountain population in which it was observed, it is interesting that the estimated fitness effects place inv6 roughly in parameter space where the low frequencies observed at Iron Mountain are not surprising.
 
@@
 
-\textcolor{red}{This needs some polishing, but I though you may want to embellish.}.
 \cite{grieshop2017male} found that genotypes with sexually antagonstic alleles that are male beneficial experience higher levels of inbreeding depression than genotypes with female beneficial SA alleles.

### 2021-03-26 doc/SA-Hermaphrodites-wDemography.tex
Draft ready for friendly review
Made minor edits, new title, added Supplementary material .tex file.

@@
-\documentclass[11pt,draft]{article}
+\documentclass[11pt]{article}
 % Preamble
@@
 \begin{document}
-\title{Evolutionary demography and the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites}
+\title{Demographic costs of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites}
+
+% ALTERNATIVE TITLES:
+%Evolutionary demography and the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites
+
+% Population viability affects/impacts the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites
+
+ % Demographic viability of antagonistic alleles impacts conditions for the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites
+
+ %Conditions for the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites impacted by considering population viability
 
@@
 
-\textit{Manuscript elements}: Figure~1, figure~2, table~1, online appendices~A and B (including figure~A1 and figure~A2). Figure~2 is to print in color.
+\textit{Manuscript elements}: figure~1, figure~2, figure~3, figure~4, table~1, table~2, appendices~A, B, and C, Online Supplementary Material. All figures are to print in color.
 
@@
 
-\textit{Keywords}: Intralocus sexual conflict, Evolutionary Demography, Balancing selection, Hermaphrodite, mixed mating systems, inbreeding depression
+\textit{Keywords}: Intralocus sexual conflict, Evolutionary Demography, Balancing selection, Hermaphrodite, Mixed mating systems, Inbreeding depression, polymorphism
 
@@
 \newpage{}
-\section{Title options? \title{Population viability affects/impacts the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites}}
-
- \section{Demographic viability of antagonistic alleles impacts conditions for the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites}
-
- \section{Conditions for the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites impacted by considering population viability}
-
-\section*{Abstract -- currently 215/200 word max.}
+\section*{Abstract -- currently 200/200 word max.}%\textcolor{green}{ I like this abstract best (CdV)} \textcolor{blue}{Agreed (CJO)}}
+
+% Alternative opening sentence to the abstract:sentence
+Classic population genetics theory has been fundamental to understanding the evolution of sex-differences and the maintenance of sexually antagonistic (SA) genetic variation, but these models have rarely considered the demographic consequences of intralocus sexual antagonism.
+In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations.
+We focus on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population intrinsic growth rate remains positive.
+In addition, we analyze the effects of inbreeding depression manifesting in different life-history stages, and give an illustrative example of the potential for SA polymorphism in real populations using empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus}.
+Our results indicate that when population growth rates are near one, large swathes of polymorphic parameter space become demographically inviable, and that inbreeding depression poses a significant challenge to the maintenance of SA polymorphism in partially selfing populations.
+Despite these demographic challenges, our example with {\itshape M.
+guttatus} appears to show that demographic rates observed in some real populations are capable of sustaining large regions of viable SA polymorphic space.
+
+%\section*{Alternative Abstract -- currently 215/200 word max.}
 %%%%%%%%%%%%%%%%%%%%
-Genetic trade-offs between male and female fitness impose fascinating genetic constraints on adaptation.
+%Genetic trade-offs between male and female fitness impose fascinating genetic constraints on adaptation.
 On one hand they can maintain genetic variation for a variety of life-history and gender-related traits, while on the other hand they cause unavoidable maladaptation as individuals are prevented from reaching their sex-specific fitness optima.
@@
 
-\newpage{}
-\section*{ALTERNATIVE Abstract -- currently 200/200 word max.\textcolor{green}{ I like this abstract best}}
-
-% Alternative opening sentence to the abstract:sentence
-Classic population genetics theory has been fundamental to understanding the evolution of sex-differences and the maintenance of sexually antagonistic (SA) genetic variation, but these models have rarely considered the demographic consequences of intralocus sexual antagonism.
-In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations.
-We focus on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection \textit{and} the population intrinsic growth rate remains positive.
-In addition, we analyze the effects of inbreeding depression manifesting in different life-history stages, and give an illustrative example of the potential for SA polymorphism in real populations using empirically estimated demographic rates for the hermaphroditic flowering plant Mimulus guttatus.
-Our results indicate that when population growth rates are near one, large swathes of polymorphic parameter space become demographically inviable, and that inbreeding depression poses a significant challenge to the maintenance of SA polymorphism in partially selfing populations.
-Despite these demographic challenges, our example with M.
-guttatus appears to show that demographic rates observed in some real populations are capable of sustaining large regions of viable SA polymorphic space.
-
 % GOOD STUFF HERE...
@@
 Yet, other factors, such as genetic linkage to other SA loci or a sex-determining region \citep{Otto2011, JordanCharlesworth2012, Olito2017, Olito2019}, or spatial heterogeneity and complexity of the life-cycle can expand the parameter space for SA polymorphism \citep{Olito-etal-2018,ConnallonSharmaOlito2019}.
-Overall, both theoretical predictions and current empirical data suggest there is ample scope for SA trade-offs and the maintenance of SA polymorphisms in both dioecious and hermaphrodite populations \citep{Abbott2011, WangBarrett2020}, although identifying specific SA loci from genome sequence data remains challenging \citep{RuzickaESEB2020}.
+Overall, both theoretical predictions and current empirical data suggest that there is ample scope for SA trade-offs and the maintenance of SA polymorphisms in both dioecious and hermaphrodite populations \citep{Abbott2011, WangBarrett2020}, although identifying specific SA loci from genome sequence data remains challenging \citep{RuzickaESEB2020}.
 
 Most population genetic models, including models of the maintenance of SA polymorphisms, keep track of the relative fitnesses of genotypes in populations of constant size, and therefore do not consider the population dynamical consequences of evolutionary change.
-In fact, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals (e.g., \cite{johnston2013life},\cite{merot2020balancing}).
+In fact, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals (e.g., \citealt {johnston2013life, merot2020balancing}).
 To take the population dynamical consequences of polymorphism into account, a model is required that links changes in genotype distributions to population dynamics.
-The potential demographic costs of sexual antagonism were pointed out by \cite{kokko2003sexy}, but few papers since then have taken this into account (with the notable exception of \cite{harts2014demography}).
-\cite{deVriesCaswell2019b} introduced a mendelian matrix model with intralocus sexual antagonism, and population dynamics but did not perform an analysis of the consequences of demographic viability for the scope of sexual antagonism to maintain polymorphism.
+The potential demographic costs of sexual antagonism were pointed out by \cite{kokko2003sexy}, but few papers since then have taken this into account (with the notable exception of \citealt{harts2014demography}).
+\citet{deVriesCaswell2019b} introduced a mendelian matrix model with intralocus sexual antagonism, and population dynamics but did not perform an analysis of the consequences of demographic viability for the scope of sexual antagonism to maintain polymorphism.
 
@@
 In fact, the present model reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019b} under obligate outcrossing.
-A full derivation of the model and analyses is presented in \hl{Appendix A}, and all computer code necessary to reproduce the results are available at \url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
+A full derivation of the model and analyses is presented in the \hl{Online Supplementary Material}, and all computer code necessary to reproduce the results are available at \url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
 
@@
 
-Individuals in the model are jointly classified by life-cycle stage ($1, \ldots, \omega$), genotype ($1, \ldots, g$), and whether they were produced by self- or outcross fertilization (denoted by $S$ and $X$ superscripts, respectively).
+Individuals in the model are jointly classified by life-cycle stage ($1, \ldots, \omega$), genotype ($1, \ldots, g$), and whether they were produced by self- or outcross fertilization (denoted by $S$ and $X$ superscripts, respectively; see Table \ref{tab:Terms} for a full description of terms included in the model).
 Whether individuals were produced through selfing or outcrossing is included in the individual state description because individuals produced through selfing might experience reduced survival, growth, or maturation rates as a consequence of inbreeding depression.
@@
 \subsubsection*{Population projection}
-Using the component matrices described above (the survival matrices, $\mathbf{U}^X_i$, $\mathbf{U}^S_i$, the fertility and mating success matrices, $\mathbf{F}_i$, $\mathbf{F}^\prime_i$, and the parent-offspring matrices $\mbf{H}^S(\tilde{\mbf{n}})$ and $\mbf{H}^X$), we construct the population projection matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ using the vec-permutation approach of \citet{CaswellEtAl2018}, see \hl{Appendix A.1} for the step by step construction of the model.
+Using the component matrices described above (the survival matrices, $\mathbf{U}^X_i$, $\mathbf{U}^S_i$, the fertility and mating success matrices, $\mathbf{F}_i$, $\mathbf{F}^\prime_i$, and the parent-offspring matrices $\mbf{H}^S(\tilde{\mbf{n}})$ and $\mbf{H}^X$), we construct the population projection matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ using the vec-permutation approach of \citet{CaswellEtAl2018}, see \hl{Online Supplementary Material} for the step by step construction of the model.
 The matrix that projects the eco-evolutionary dynamics is:
@@
 where $C$ denotes the proportion of each individual's ovules that are self-fertilized (the remaining $1 - C$ are outcrossed), and  $\delta$ represents the proportion of self-fertilized zygotes that fail to develop due to inbreeding depression during early development \citep{Charlesworth1987}.
-In \hl{Appendix A.x}, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
+In the \hl{Online Supplement}, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
 
@@
 \noindent where $q^{\prime}_A$ and $q^{\prime}_a$ are given by Eq(\ref{eq:maleGametePool}).
-Further details of the derivations of $\mcal{U}^S$, $\mcal{U}^X$ ,  $\mcal{F}^S$, and $\mcal{F}^X$ are provided in \hl{Appendix A.x}.
+Further details of the derivations of $\mcal{U}^S$, $\mcal{U}^X$ ,  $\mcal{F}^S$, and $\mcal{F}^X$ are provided in the \hl{Online Supplement}.
 
@@
 
-Combining all the component matrices yields the overall eco-evolutionary projection matrix shown in \hl{Appendix A.x}.
+Combining all the component matrices yields the overall eco-evolutionary projection matrix shown in \hl{Appendix A}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}[htbp]
+\centering
+\caption{\bf Definition of terms.}
+\label{tab:Terms}
+\begin{tabular}{ l p{0.7\linewidth} r }
+ \toprule
+Symbol & Definition & Dimension \\
+\hline
+$g$      & Number of genotypes ($3$; $AA$, $Aa$, and $aa$) & \\
+$\omega$ & Number of stages ($2$; juvenile and adult) & \\
+$\tilde{\mbf{n}}$ & Joint stage $\times$ genotype vector & $2 \omega g \times 1$ \\
+$\tilde{\mbf{p}}$ & Joint stage $\times$ genotype frequency vector & $2 \omega g \times 1$ \\
+$\mbf{U}^{S}_{i}$,\, $\mbf{U}^{X}_{i}$ & Genotype-specific transition and survival matrices & $\omega \times \omega$ \\
+$\mcal{U}^{S},\, \mcal{U}^{X}$ & Block diagonal selfed/outcrossed survival matrices & $\omega g \times \omega$ \\
+$\mbf{F}_{i}$,\, $\mbf{F}^{\prime}_{i}$ & Genotype-specific fertility matrices & $\omega \times \omega$ \\
+$\mbb{F}^{\prime}$ & Block diagonal fertility matrix & $\omega g \times \omega$ \\
+$\mcal{F}^{S},\, \mcal{F}^{X}$ & Block diagonal selfed/outcrossed fertility matrix & $\omega g \times \omega$ \\
+$\mbf{H}^{S}_{i}$,\, $\mbf{H}^{X}_{i}$ & Parent-offspring genotype maps & $g \times g$ \\
+$\mbf{W}$   & Allele segregation matrix & $\omega \times g$ \\
+% $\mbf{K}$   & Vec-permutation matrix & $g \times g$ \\
+% $\mbf{I}$   & Identity matrix & given in subscript \\
+$q^{\prime}_{A}$,\, $q^{\prime}_{a}$ & Allele frequencies in male gamete pool & $1$ \\
+$\tilde{\mbf{A}}(\tilde{\mbf{p}})$ & Population projection matrix & $2 \omega g \times g$ \\
+
+$s_f,\,s_m$ & Selection coefficients through female and male reproductive function & \\
+$h_f,\,h_m$ & Dominance coefficients through female and male reproductive function & \\
+$\sigma_j,\, \sigma_a$ & Survival rates for juvenile and adult stages & \\
+$\gamma$ & Transition rate from juvenile to adult stages \\
+$C$ & The population rate of self-fertilization & \\
+$\delta,\, \delta_j,\,\delta_a,\,\delta_{\gamma}$ & Inbreeding depression terms for ovule viability, survival, and transition rates respectively & \\
+$f_i,\,f^{\prime}_i$ & Adult fertilities through male and female sex functions & \\
+
+\hline
+\end{tabular}
+\end{table}
+\newpage{}
 
@@
 Suppose that there is a genetic trade-off between the sex-functions at a single diallelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect.
-
 Following convention, we parameterize the fertility component of fitness of each genotype through each sex function, $w_{i}$ and $w^{\prime}_{i}$, to be bounded by $[0,1]$, with dominance and selection coefficients $h_f, s_f$ and $h_m, s_m$ determining the decrease in fertility through each sex function relative to the most fit genotype ($AA$ has highest female fertility, $aa$ the highest male fertility; see Table \ref{tab:Fitness}).
+
 The SA locus does not affect survival and transition rates.
@@
 The formal conditions for a protected polymorphism are determined by linearizing the model in the vicinity of the boundary equilibria ($\hat{\mbf{p}}$), and evaluating the magnitude of the largest eigenvalue of the Jacobian matrix of the linearization.
-A full derivation of the Jacobian and details of the invasion analysis are provided in \hl{Online Appendix X.x}.
-
-%denoted by $\tilde{\zeta}_{AA}$ and $\tilde{\zeta}_{aa}$ for the eigenvalue of the Jacobian evaluated at the $AA$ and $aa$ boundary, respectively.
-For our model, the Jacobian, $\mbf{M}$, has an upper block triangular form, and we can therefore express the stability of $\hat{\mbf{p}}$ in terms of the leading eigenvalue of the diagonal block corresponding to changes in heterozygote frequency ($\mbf{M}_{22}$; e.g., see \citealt{OttoDay2007,deVriesCaswell2019b}).
-For a polymorphism to be 'protected', both boundary equilibria must be unstable (i.e., $\tilde{\zeta}_{AA} > 1$ and $\tilde{\zeta}_{aa} > 1$), which requires that
-% \begin{linenomath*}
-% \begin{equation}\label{eq:PolyCond}
-% 	\rho(\mbf{M}_{22}) > \lambda_{i}
-% \end{equation}
-% \end{linenomath*}
-
-% \noindent is satisfied, where $\rho(\cdot)$ represents the spectral radius,  $\rho(\mbf{M}_{22})$ gives the ergodic growth rate of perturbations of the boundary equilibrium according to the linear approximation, and $\lambda_i$ is the intrinsic growth rate for the homozygous genotype being invaded at each boundary ($i \in (AA,aa)$; see Box.~1).
+A full derivation of the Jacobian and details of the invasion analysis are provided in the \hl{Online Supplementary Material}, and the relevant leading eigenvalues are presented in \hl{Appendix B}.
 
@@
  \centering
- \includegraphics[width=\linewidth]{../output/figs/extinctionThresholdsFig.pdf}
+ \includegraphics[width=\linewidth]{extinctionThresholdsFig}
  \caption{\footnotesize{Illustration of parameter space for SA polymorphism and extinction thresholds predicted by the model.
@@
 We begin with an illustration of demographically viable polymorphic parameter space in the absence of inbreeding depression in Figure \ref{fig:extThresholds} (i.e., $\delta = \delta_i = 0$).
-Invasion conditions for SA alleles in the evolutionary demographic model closely match the predictions from population genetic models (see \hl{Online Appendix X}; \citealt{Kidwell1977, JordanConnallon2014, Olito2017}).
+Invasion conditions for SA alleles in the evolutionary demographic model closely match the predictions from population genetic models \cite{Kidwell1977, JordanConnallon2014, Olito2017}.
 In particular, the demographic model recovers the classic "funnel-shaped" region of polymorphic $s_f \times s_m$ parameter space.
-The effects of the population selfing rate ($C$) and dominance ($h$) on SA polymorphism are also similar: self-fertilization ({\itshape i}) skews the SA polymorphic space towards regions with weaker selection through female-function and reduces the total area of polymorphic parameter space (e.g., contrast figure 1A with figure 1C); and ({\itshape ii}) dominance reversals (where deleterious SA fitness effects are partially recessive in each sex; $h = 1/4$) are much more permissive of SA polymorphism (e.g., contrast Figure 1A with Figure 1D) \citep{JordanConnallon2014, Olito2017}.
+The effects of the population selfing rate ($C$) and dominance ($h$) on SA polymorphism are also similar: self-fertilization ({\itshape i}) skews the SA polymorphic space towards regions with weaker selection through female-function and reduces the total area of polymorphic parameter space (e.g., contrast figure 1A with figure 1C); and ({\itshape ii}) dominance reversals (where deleterious SA fitness effects are partially recessive in each sex; $h = 1/4$) are much more permissive of SA polymorphism (e.g., contrast figures 1A-C with 1D-F) \citep{JordanConnallon2014, Olito2017}.
 
@@
 In obligately outcrossing populations (including dioecious/gonochoristic populations; where $C = 0$), lower fertility can result in a significant reduction of demographically viable polymorphic parameter space.
-However, the effect is weaker in populations with intermediate selfing rates (compare $C > 0$ vs $C=0$) because self-fertilization protects a fraction, $C$, of the ovules of individuals homozygous for the female beneficial allele from exposure to the allele that is deleterious for females.
+However, the effect is weaker in populations with intermediate selfing rates (compare $C=0$ vs.~$C > 0$) because self-fertilization protects a fraction, $C$, of the ovules of individuals homozygous for the female beneficial allele from exposure to the allele that is deleterious for females.
 In other words, self-fertilization leads to less selective death.
@@
 Mortality from inbreeding depression reduces the effect of protection from selective death that individuals gain through self fertilization.
-In contrast to our earlier results (see Figure \ref{fig:polySpace} in \hyperref[subsec:PolyExt]{Polymorphism and Extinction}), predominantly outcrossing populations are predicted to have the highest proportion of demographically viable polymorphic space, when inbreeding depression is taken into acount (compare fig.~\ref{fig:polySpace} with fig.~\ref{fig:deltaPolySpace}).
+In contrast to our earlier results (see Figure \ref{fig:polySpace} in \hyperref[subsec:PolyExt]{Polymorphism and Extinction}), predominantly outcrossing populations are predicted to have the highest proportion of demographically viable polymorphic space when inbreeding depression is taken into acount (compare fig.~\ref{fig:polySpace} with fig.~\ref{fig:deltaPolySpace}).
 Regardless of the life-history stage at which inbreeding depression affects survival, populations with intermediate to high selfing rates are unlikely to harbour SA polymorphism unless they can afford the resulting loss of self-fertilized ovules/offspring.
@@
  \caption{\footnotesize{Effects of early- and late-acting inbreeding depression on the proportion of demographically viable parameter space (out of total $s_f \times s_m$ space with $\max(s) = 0.15$), plotted as a function of the population selfing rate.
-In all plots, the strength of inbreeding depression decreases as the selfing rate goes up following a simple model of purging recessive deleterious mutations (see \hl{Appendix X.x} for details).
+In all plots, the strength of inbreeding depression decreases as the selfing rate goes up following a simple model of purging recessive deleterious mutations (see \hyperref[subsec:analyses]{Analyses}).
 Only single inbreeding depression terms ($\delta$ and $\delta_i$, where $i \in \{j,a,\gamma\}$, indicated in the legend) are allowed to vary at one time (all others are set to $0$).
@@
 Moreover, detailed demographic studies have been conducted on multiple populations of {\itshape M.~guttatus}, with demographic matrix data available on the public demographic database COMPADRE \citep{CompadreDB2020}.
-Below, we briefly outline how we parameterized our model using the available data; full details are provided in \hl{Appendix B.x}.
+Below, we briefly outline how we parameterized our model using the available data; full details are provided in \hl{Appendix C}.
 
@@
 Selfing rate estimates vary in magnitude from near complete outcrossing to predominant selfing ($C \approx 0$ to $0.75$; \citealt{RitlandGanders1987, Ritland1990, Willis1999b}).
-Estimates of inbreeding depression at several of the life-history stages/fitness components that were included in the data of \citet{PetersonEtAl2016} are available for two intensively studied populations in the Cascade Mountains of Oregon (Iron Mountain and Cone Peak) \citet{Willis1993, Willis1999a, Willis1999b}.
+Estimates of inbreeding depression at several of the life-history stages/fitness components that were included in the data of \citet{PetersonEtAl2016} are available for two intensively studied populations in the Cascade Mountains of Oregon (Iron Mountain and Cone Peak; \citealt{Willis1993, Willis1999a, Willis1999b}).
 Using the data provided in \citet{Willis1993}, we estimated the proportional decrease due to inbreeding depression in seed germination rate ($\delta_{G} = 0.085$), flower number ($\delta_{F} = 0.2$), and overwinter survival ($\delta_{S} = 0.38$).
@@
 When including empirical estimates of inbreeding depression, the population effective selfing rate is reduced, the polymorphic parameter space shifts upwards, and inv6 no longer falls within the polymorphic space.
-When using vital rates for the locally adapted Eagle Meadows population, the model predicts that an allele (or supergene) with the same fitness effects as inv6 will go extinct.
+Although a population with vital rates like those of the locally adapted Eagle Meadows population can support the demographic costs of inv6, our model predicts that an allele (or supergene) with the same fitness effects as inv6 will not be maintained as a polymorphism.
 In contrast, when using the demographic rates for Low-Elevation Perennials, an allele/supergene with these fitness effects falls within demographically inviable parameter space.
@@
 Classic population genetics theory predicts that sexually antagonistic selection is unlikely to maintain genetic variation except under narrow conditions, with polymorphism requiring either finely balanced or unusually strong selection, or partially recessive fitness effects through each sex \citep{Kidwell1977,Pamilo1979,Prout2000, ConnallonClark2014}.
-Extensions of the theory have identified numerous ways in which the conditions for polymorphism become more permissive in both dioecious and hermaphroditic organisms, including genetic linkage of SA loci, population subdivision, and life-cycle complexity \citep[e.g.,][]{Patten2010,JordanConnallon2014,JordanCharlesworth2012, ConnallonSharmaOlito2019, Olito-etal-2018}.
-However, by ignoring the demographic consequences of SA genetic variation, these population genetic models have missed the possibility that SA polymorphisms may be inviable under realistic parameter conditions, and therefore unlikely to be observed in natural populations.
+Extensions of the theory have identified numerous ways in which the conditions for polymorphism become more permissive in both dioecious and hermaphroditic organisms, including genetic linkage of SA loci, the evolutioin of sex-specific dominance, population subdivision, and life-cycle complexity \citep[e.g.,][]{Patten2010,JordanConnallon2014,JordanCharlesworth2012,SpencerPriest2016, ConnallonSharmaOlito2019, Olito-etal-2018}.
+However, by ignoring the demographic consequences of SA genetic variation, these population genetic models have missed the possibility that SA polymorphisms may not be viable under realistic parameter conditions, and therefore unlikely to be observed in natural populations.
 By linking the individual-level fitness consequences of SA selection to population level demographic rates, our theoretical framework provides several key insights into the processes shaping SA genetic variation in natural populations.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Appendices
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% In many cases, The American Naturalist allows authors to typeset 
+% their own supplementary material in an author-supplied PDF. For author-
+% supplied PDFs, please consult the AmNat_supp_template.tex document,
+% available from https://www.journals.uchicago.edu/journals/an/instruct 
+%
+% By contrast, the Appendix instructions below apply to cases in which
+% supplementary material is to be typeset by the AmNat editorial staff.
+% That notably includes descriptions of methods, tables defining parameters,
+% and other material necessary for reproducing the MS's results.
+%
+% Please reset counters for the appendix (thus normally figure A1, 
+% figure A2, table A1, etc.).
+%
+% In certain cases, it may be appropriate to have a PRINT appendix in
+% addition to (or instead of) an online appendix. In this case, please 
+% name the print appendix Appendix A, and any subsequent appendixes (if 
+% there are any) should be named Online Appendix B, Online Appendix C,
+% etc.
+%
+% Counters for each appendix should match the letter of that appendix.
+% For example, tables in Appendix C should be numbered table C1, table C2,
+% etc. This applies to tables, equations, and figures.
+%
+% It's better not to use the \appendix command, because we have some
+% formatting peculiarities that \appendix conflicts with.
+
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
+\section{Appendix A: Population projection matrix}
 \renewcommand{\theequation}{A\arabic{equation}}
@@
 
-\section{Appendix A: Construction of the population projection matrix}
+\textcolor{red}{NEED TO CHANGE THIS TO A MINIMAL PRESENTATION OF THE PROJECTION MATRIX... Can move the material provided below to the Online Supplement}
 
@@
 
-% In many cases, The American Naturalist allows authors to typeset 
-% their own supplementary material in an author-supplied PDF. For author-
-% supplied PDFs, please consult the AmNat_supp_template.tex document,
-% available from https://www.journals.uchicago.edu/journals/an/instruct 
-%
-% By contrast, the Appendix instructions below apply to cases in which
-% supplementary material is to be typeset by the AmNat editorial staff.
-% That notably includes descriptions of methods, tables defining parameters,
-% and other material necessary for reproducing the MS's results.
-%
-% Please reset counters for the appendix (thus normally figure A1, 
-% figure A2, table A1, etc.).
-%
-% In certain cases, it may be appropriate to have a PRINT appendix in
-% addition to (or instead of) an online appendix. In this case, please 
-% name the print appendix Appendix A, and any subsequent appendixes (if 
-% there are any) should be named Online Appendix B, Online Appendix C,
-% etc.
-%
-% Counters for each appendix should match the letter of that appendix.
-% For example, tables in Appendix C should be numbered table C1, table C2,
-% etc. This applies to tables, equations, and figures.
-%
-% It's better not to use the \appendix command, because we have some
-% formatting peculiarities that \appendix conflicts with.
-
-\renewcommand{\theequation}{A\arabic{equation}}
-\renewcommand{\thetable}{A\arabic{table}}
-\setcounter{figure}{0}
-\setcounter{table}{0}
-
-%\section{Appendix B: Additional Methods}
+%%%%%%%%%%%%%%%%%%%%%%%%
+% Appendix B
+%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Appendix B: Eigenvalues for invasion analysis}
+\renewcommand{\theequation}{B\arabic{equation}}
+\setcounter{equation}{0}  % reset counter
+
+\textcolor{red}{STILL NEED TO FILL THIS OUT WITH EIGENVALUES ETC.}
+
+denoted by $\tilde{\zeta}_{AA}$ and $\tilde{\zeta}_{aa}$ for the eigenvalue of the Jacobian evaluated at the $AA$ and $aa$ boundary, respectively.
+For our model, the Jacobian, $\mbf{M}$, has an upper block triangular form, and we can therefore express the stability of $\hat{\mbf{p}}$ in terms of the leading eigenvalue of the diagonal block corresponding to changes in heterozygote frequency ($\mbf{M}_{22}$; e.g., see \citealt{OttoDay2007,deVriesCaswell2019b}).
+For a polymorphism to be 'protected', both boundary equilibria must be unstable (i.e., $\tilde{\zeta}_{AA} > 1$ and $\tilde{\zeta}_{aa} > 1$), which requires that
+ \begin{linenomath*}
+ \begin{equation}\label{eq:PolyCond}
+ 	\rho(\mbf{M}_{22}) > \lambda_{i}
+ \end{equation}
+ \end{linenomath*}
+
+ \noindent is satisfied, where $\rho(\cdot)$ represents the spectral radius,  $\rho(\mbf{M}_{22})$ gives the ergodic growth rate of perturbations of the boundary equilibrium according to the linear approximation, and $\lambda_i$ is the intrinsic growth rate for the homozygous genotype being invaded at each boundary ($i \in (AA,aa)$; see Box.~1).
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+% Appendix C
+%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Appendix C: Demographic and empirical data for {\itshape M.~guttatus} case study}
+\renewcommand{\theequation}{C\arabic{equation}}
+\renewcommand{\thetable}{C\arabic{table}}
+\setcounter{equation}{0}  % reset counter
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}[htbp]
-\centering
-\caption{\bf Definition of terms.}
-\label{tab:Terms}
-\begin{tabular}{ l p{0.7\linewidth} r }
- \toprule
-Symbol & Definition & Dimension \\
-\hline
-$g$      & Number of genotypes ($3$; $AA$, $Aa$, and $aa$) & \\
-$\omega$ & Number of stages ($2$; juvenile and adult) & \\
-$\tilde{\mbf{n}}$ & Joint stage $\times$ genotype vector & $2 \omega g \times 1$ \\
-$\tilde{\mbf{p}}$ & Joint stage $\times$ genotype frequency vector & $2 \omega g \times 1$ \\
-$\mbf{U}^{S}_{i}$,\, $\mbf{U}^{X}_{i}$ & Genotype-specific transition and survival matrices & $\omega \times \omega$ \\
-$\mcal{U}^{S},\, \mcal{U}^{X}$ & Block diagonal selfed/outcrossed survival matrices & $\omega g \times \omega$ \\
-$\mbf{F}_{i}$,\, $\mbf{F}^{\prime}_{i}$ & Genotype-specific fertility matrices & $\omega \times \omega$ \\
-$\mbb{F}^{\prime}$ & Block diagonal fertility matrix & $\omega g \times \omega$ \\
-$\mcal{F}^{S},\, \mcal{F}^{X}$ & Block diagonal selfed/outcrossed fertility matrix & $\omega g \times \omega$ \\
-$\mbf{H}^{S}_{i}$,\, $\mbf{H}^{X}_{i}$ & Parent-offspring genotype maps & $g \times g$ \\
-$\mbf{W}$   & Allele segregation matrix & $\omega \times g$ \\
-% $\mbf{K}$   & Vec-permutation matrix & $g \times g$ \\
-% $\mbf{I}$   & Identity matrix & given in subscript \\
-$q^{\prime}_{A}$,\, $q^{\prime}_{a}$ & Allele frequencies in male gamete pool & $1$ \\
-$\tilde{\mbf{A}}(\tilde{\mbf{p}})$ & Population projection matrix & $2 \omega g \times g$ \\
-
-$s_f,\,s_m$ & Selection coefficients through female and male reproductive function & \\
-$h_f,\,h_m$ & Dominance coefficients through female and male reproductive function & \\
-$\sigma_j,\, \sigma_a$ & Survival rates for juvenile and adult stages & \\
-$\gamma$ & Transition rate from juvenile to adult stages \\
-$C$ & The population rate of self-fertilization & \\
-$\delta,\, \delta_j,\,\delta_a,\,\delta_{\gamma}$ & Inbreeding depression terms for ovule viability, survival, and transition rates respectively & \\
-$f_i,\,f^{\prime}_i$ & Adult fertilities through male and female sex functions & \\
-
-\hline
-\end{tabular}
-\end{table}
-\newpage{}
+Tables moved up to Model section in the main text.
 
@@
 
-\begin{figure}[h!]
+Figure legends provided beneath each figure in the main text.
+%\begin{figure}[h!]
 %\includegraphics{horn-of-okapi}
-\caption{Figure legends can be longer than the titles of tables. However, they should not be excessively long.}
+%\caption{Figure legends can be longer than the titles of tables. However, they should not be excessively long.}
 %\label{Fig:OkapiHorn}
-\end{figure}
-
-\begin{figure}[h!]
+%\end{figure}
+
+%\begin{figure}[h!]
 %\includegraphics{elegance}
-\caption{In this way, figure legends can be listed at the end of the document, with references that work, even though the graphic itself should be included for final files after acceptance.
+%\caption{In this way, figure legends can be listed at the end of the document, with references that work, even though the graphic itself should be included for final files after acceptance.
 Instead, upload the relevant figure files separately to Editorial Manager; Editorial Manager should insert them at the end of the PDF automatically.}
 %\label{Fig:AnotherFigure}
-\end{figure}
+%\end{figure}
 

### 2021-03-26 doc/Supplements/SupplementaryMaterial.tex
Draft ready for friendly review
Made minor edits, new title, added Supplementary material .tex file.

--- 2021-03-11 doc/Supplements/AmNat_supp_template.tex
@@
 \pagestyle{fancy}
-
+\usepackage[colorlinks=true, allcolors=black]{hyperref}
+\setlength{\headheight}{13.6pt}
 %%%%%%%%%%%%%%%%%%%%%
@@
 %
-\rhead{Supplement to AUTHOR-NAME et al., ``SHORT TITLE,'' \textit{Am.~Nat.}}
+\rhead{Supplement to Olito \& de Vries., "Costs of SA polymorphism" \textit{Am.~Nat.}}
 \setlength{\headsep}{0.3in}  
@@
 
-\title{Online Supplement: \\ 
-\LaTeX{} Template for Author-Supplied Supplementary PDFs, \\ 
+\title{Online Supplement: Demographic costs of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites \\ 
+%\LaTeX{} Template for Author-Supplied Supplementary PDFs, \\ 
 \textit{The~American~Naturalist} }
@@
 % page numbers.
+\author{Colin Olito$^{1,\ast}$ \\ 
+Charlotte de Vries$^{2}$}
 
-\author{Owen E. Cook$^{1,\ast}$ \\ 
-Generic H. Collaborator$^{2,\dag}$ \\ 
-Additional Q. Expert$^{3}$}
-
-\date{}
+\date{\today}
 
@@
 
-\noindent{} 1. University of Chicago, Chicago, Illinois 60637;
+\noindent{} 1. Department of Biology, Lund University, Lund 223 62, Sweden;
 
-\noindent{} 2. University of Toronto, Toronto, Ontario M5S 1A5, Canada;
+\noindent{} 2.  Department of Evolutionary Biology and Environmental Studies, University of Zurich, Zurich CH-8057, Switzerland;
 
-\noindent{} 3. Middle Eastern Technical University, Çankaya, Ankara 06800, Turkey.
-
-\noindent{} $\ast$ Corresponding author; e-mail: amnat@uchicago.edu.
-
-\noindent{} $\dag$ Deceased.
+\noindent{} $\ast$ Corresponding author; e-mail: \url{colin.olito@gmail.com}
 

### 2021-03-29 doc/SA-Hermaphrodites-wDemography.tex
Fixed inv6 sel coeffs, Updated appendices
In this commit, I've fixed how we were estimating the selection
coefficients for inv6, updated Fig. 4, altered our results discussion
to reflect these changes. I have also added in 3 appendices (A,B,C)
with the overall projection matrix (appendix A), the eigenvalues used
in the invasion analysis (appendix B), and a summary of demographic
data for Mimulus guttatus and the new methods for calculating
selection coefficients for inv6.

@@
 \usepackage[colorlinks=true, allcolors=black]{hyperref}
+\usepackage{pdflscape}
 
@@
     \node (X) [draw, shape=circle, inner sep=0] {\strut #1};}
+
+% Equation numbering
+\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
 
@@
 The Eagle Meadows population was locally adapted with a very high intrinsic growth rate ($\lambda \approx 1.7$).
-This growth rate was high enough that all $s_f \times s_m$ selection parameter space (where $s_f, s_m \in (0,1]$) remained demographically viable, regardless of the selfing rate ($C = 0$ or $0.29$), or the presence of inbreeding depression (fig.~\ref{fig:mimulusFig}A).
+This growth rate was sufficiently high enough that all $s_f \times s_m$ selection parameter space (where $s_f, s_m \in (0,1]$) remained demographically viable, regardless of the selfing rate and effects of inbreeding depression ($C = 0$ or $0.29$; fig.~\ref{fig:mimulusFig}A).
 In contrast, the Low-Elevation Perennial population had a much lower, but still positive, intrinsic growth rate ($\lambda \approx 1.08$).
-Due to the slower growth rate, not all of the $s_f \times s_m$ selection parameter space was demographically viable.
-Extinction thresholds appeared under complete outcrossing ($C = 0$) and partial selfing with inbreeding depression ($C = 0.29$), but not when inbreeding depression was omitted (fig.~\ref{fig:mimulusFig}B).
-The different demographic rates from these two populations also resulted in slightly different invasion conditions when inbreeding depression was taken into account (compare dotted lines in fig.~\ref{fig:mimulusFig}A and B).
-
-The polymorphic inversion inv6 falls very near the lower edge of polymorphic parameter space for these hypothetical populations, but is not always predicted to be demographically viable.
-When using demographic rates for either the Eagle Meadows or Low-Elevation Perennial populations, inv6 falls within polymorphic space when the selfing rate is at the higher end of empirical estimates for the Iron Mountain population in which inv6 has been documented ($C = 0.29$), but only when inbreeding depression is not taken into account (fig.~\ref{fig:mimulusFig} dashed lines).
-When including empirical estimates of inbreeding depression, the population effective selfing rate is reduced, the polymorphic parameter space shifts upwards, and inv6 no longer falls within the polymorphic space.
-Although a population with vital rates like those of the locally adapted Eagle Meadows population can support the demographic costs of inv6, our model predicts that an allele (or supergene) with the same fitness effects as inv6 will not be maintained as a polymorphism.
-In contrast, when using the demographic rates for Low-Elevation Perennials, an allele/supergene with these fitness effects falls within demographically inviable parameter space.
-That is, populations with lower intrinsic growth rates like those observed in the Low-Elevation Perennials will be unable to support the demographic costs associated with segregating SA alleles with selection coefficients of similar magnitude to inv6.
+Due to the slower growth rate, not all of the $s_f \times s_m$ selection parameter space was demographically viable: extinction thresholds appear under both complete outcrossing and partial selfing with inbreeding depression (fig.~\ref{fig:mimulusFig}B).
+The different demographic rates from these two populations also resulted in slightly different invasion conditions when inbreeding depression was taken into account (compare dashed with solid lines in fig.~\ref{fig:mimulusFig}A and B).
+
+When using demographic rates for either the Eagle Meadows or Low-Elevation Perennial populations, inv6 falls within demographically viable polymorphic parameter space (fig.~\ref{fig:mimulusFig}; estimated selection coefficients for inv6 were $(s_f,\,s_m) = (0.30,\, 0.31)$).
+Whether outcrossing or partial selfing ($C=0.29$), inv6 always falls squarely in the middle of SA polymorphic space when using the locally adapted Eagle Meadows demographic rates.
+In contrast, when the selfing rate is at the higher end of empirical estimates for the Iron Mountain population in which inv6 has been documented ($C = 0.29$), inv6 falls nearer to the upper boundary of SA polymorhic space when using the Low-Elevation Perennial demographic rates.
+This happens because, when including empirical estimates of selfing rates and inbreeding depression, the polymorphic parameter space shifts downwards.
+Additionally, inv6 falls much closer to the extinction threshold under partial selfing.
+Interesting, this suggests that even relatively small perturbations to demgraphic rates or selection coefficients could result in non-locally adapted populations being unable to support the demographic costs associated with segregating SA alleles with selection coefficients of similar magnitude to inv6.
 
@@
  \centering
- \includegraphics[scale=0.75]{../output/figs/mimulusinv6Fig.pdf}
+ \includegraphics[scale=0.75]{../output/figs/mimulusinv6FigNewSel.pdf}
  \caption{\footnotesize{Illustration of model predictions using empirically estimated demographic rates for {\itshape M.~guttatus}.
 Results are shown for two hypothetical populations using demographic rates for locally adapted (Eagle Meadows; panel A), and non-local (Low-Elevation Perennial; panel B) populations reported by \citet{PetersonEtAl2016}.
-Invasion conditions and extinction thresholds were calculated using the dominance coefficients estimated for inv6 ($h_f = 0.5, h_m = 0.35$; from data in \citealt{LeeKelly2015}) for three parameter conditions: obligate outcrossing (solid black lines), partial selfing using the highest field-estimate for the Iron Mountain population of \citet{Willis1993} (dashed lines), and partial selfing with inbreeding depression (dotted lines; using $\delta_i$ estimates for Iron Mountain by \citealt{Willis1993}).
-The location of inv6 is also shown on both plots, using estimated selection coefficients.
-Note that in panel A the predictions for partial selfing with inbreeding depression (dotted lines) fall extremely close to the predictions for obligate outcrossing (solid lines).}}
+Invasion conditions and extinction thresholds were calculated using the conservative assumption of additive fitness effects in both sexes ($h_f = h_m = 0.5$) for two parameter conditions: obligate outcrossing (solid black lines)  and partial selfing with inbreeding depression using the highest field-estimate of selfing, and inbreeding depression parameters ($\delta_i$) calculated for the Iron Mountain population of \citet{Willis1993} (dashed lines).
+The location of inv6 is also shown on both plots, using selection coefficients calculated from field estimates of male and female fitness components from \citet{LeeKelly2015} under the relatively conservative assumption of additive fitness effects in both sexes, where $(s_f,\,s_m) = (0.30,\, 0.31)$.
+Note that in panel A the invasion conditions for partial selfing with inbreeding depression (dotted lines) are nearly indistinguishable from those for obligate outcrossing (solid lines).}}
  \label{fig:mimulusFig}
@@
 The example also appears to provide some empirical support for the conjecure that locally-adapted populations are more likely to harbor SA polymorphisms than marginal or non-locally adapted ones; inviable polymorphic parameter space only occured when using the demographic data for non-local high-elevation perennial populations.
-The location of inv6 in selection parameter space is also interesting because it corresponds to a region where an SA allele with net female-beneficial fitness effects similar to inv6 are predicted to either go extinct, or segregate at relatively low frequency, depending on the population selfing-rate and severity of inbreeding depression.
-Although we cannot make specific predictions for inv6 in the Iron Mountain population in which it was observed, it is interesting that the estimated fitness effects place inv6 roughly in parameter space where the low frequencies observed at Iron Mountain are not surprising.
+Although we cannot make concrete predictions for inv6 in the Iron Mountain population in which it was observed, it is interesting that the estimated SA fitness effects place this polymophic inversion squarely in demographically viable polymorphic parameter space predicted by our model.
+The available data do not allow for confident estimation of selection coefficients for inv6 (and even require making assumptions about dominance), yet our theoretical predictions are encouragingly consistent with the available data that inv6 is segregating at intermediate frequencies in the large and locally adapted Iron Mountain population.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section{Appendix A: Population projection matrix}
+\section*{Appendix A: Population projection matrix}
 \renewcommand{\theequation}{A\arabic{equation}}
 \setcounter{equation}{0}  % reset counter 
-
-\textcolor{red}{NEED TO CHANGE THIS TO A MINIMAL PRESENTATION OF THE PROJECTION MATRIX... Can move the material provided below to the Online Supplement}
-
-\subsubsection*{Population projection}
-In this appendix, we will derive the expressions for the survival and fertility matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ as given by equations (\ref{eq:BlkUS})-(\ref{eq:BlkFX}) in the main text.
-The matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ are the ingredients of the population projection matrix, $\tilde{\mbf{A}}$, as given in equation (\ref{eq:Atilde}), repeated here for convenience,
-\begin{linenomath*}
-\begin{equation} \label{eq:Atilde_appendix}
-	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
+\setcounter{table}{0}  % reset counter
+
+The complete population projection matrix $\tilde{\mbf{A}}$ consists of $3 \times 3$ blocks, which act on the genotype specific population vectors:
+
+{
+\small
+
+\begin{align*} \label{eq:AtildeFull}
+	&\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
 		 \underbrace{\left(
-			\begin{array}{c|c}
-				\mcal{U}^S & \mbf{0} \\ \hline
-				\mbf{0} & \mcal{U}^X \\
-			\end{array} \right)}_{\tilde{\mbf{U}}} + 
-			\underbrace{\left(
-			\begin{array}{c|c}
-				\mcal{F}^S C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
-				\mcal{F}^X(\tilde{\mbf{p}}) (1 - C) & \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)\\
-			\end{array} \right)}_{\tilde{\mbf{F}}}.
-\end{equation}
-\end{linenomath*}
-
-Our derivation follows closely the derivations in \cite{deVriesCaswell2019a} and \cite{deVriesCaswell2019b}, but extended to model mixed mating systems.
-To construct $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ using the vec-permutation approach \citep{CaswellEtAl2018}, we create a set of block-diagonal matrices $\mathbb{U}^S$, $\mathbb{U}^X$, $\mathbb{F}^X$, $\mathbb{F}^S$, and $\mathbb{H}^S$, $\mathbb{H}^X$, by putting the corresponding matrices on the diagonal.
-These block diagonal matrices can be written as
-\begin{linenomath*}
-\begin{eqnarray}
- \mathbb{U}^S&=&\sum_{i=1}^g \mathbf{E}_{ii} \otimes \mathbf{U}^S_{i}, \label{eq:US_app}\\
- \mathbb{F}^S&=&\sum_{i=1}^g \mathbf{E}_{ii} \otimes \mathbf{F}^S_{i}, \label{eq:FS_app}\\
-  \mathbb{H}^S &=& \bo I_\om \kron \bo H^S , \label{eq:HS_app} \\
-      \mathbb{D} &=& \bo I_\om \kron \bo I_g, \label{eq:D_app} 
-\end{eqnarray}
-\end{linenomath*}
-where $\mathbf{E}_{ii}$ is of dimensions $g \times g$. Analogous equations can be written down for $ \mathbb{U}^X$, $ \mathbb{F}^X$, and $ \mathbb{H}^X$.  
-% \begin{linenomath*}
-% \begin{eqnarray} 
-% 	\mbb{U}^S &= &
-% 		\left(
-% 			\begin{array}{ccc}
-% 				\mbf{U}^{S}_{AA} & \mbf{0} & \mbf{0} \\
-% 				\mbf{0} & \mbf{U}^{S}_{Aa} & \mbf{0} \\
-% 				\mbf{0} & \mbf{0} & \mbf{U}^{S}_{aa} \\
-% 			\end{array} \right), \label{eq:BlkUS_appendix}\\
-%   \mbb{F}^S &= &
-% 		\left(
-% 			\begin{array}{ccc}
-% 				\mbf{F}^{S}_{AA} & \mbf{0} & \mbf{0} \\
-% 				\mbf{0} & \mbf{F}^{S}_{Aa} & \mbf{0} \\
-% 				\mbf{0} & \mbf{0} & \mbf{F}^{S}_{aa} \\
-% 			\end{array} \right), \label{eq:BlkFS_appendix}\\
-% \end{eqnarray}
-% \end{linenomath*}
-% Similarly construct block diagonal matrices from the parent-offspring genotype maps,
-% \begin{linenomath*}
-% \begin{eqnarray}
-% 	\mathbb{H}^X(\tilde{\bo p}) = \bo I_\om \kron \bo H^X(\tilde{\bo p}),\\
-% 	\mathbb{H}^S(\tilde{\bo p}) = \bo I_\om \kron \bo H^S(\tilde{\bo p}).
-% \end{eqnarray}
-% \end{linenomath*}
-
-As in other multistate models, the survival and fertility matrices are then constructed from the following four block matrices, 
-\begin{linenomath*}
-\begin{eqnarray}
-\mcal{U}^S &=&\bo K^\tr \mathbb{D} \bo K \mathbb{U}^S, \label{eq:KDKU_app}\\
-\mcal{F}^S &=&\bo K^\tr \mbb{H}^S\bo K \mathbb{F}^S, \label{eq:KHKF_app}
-\end{eqnarray}
-\end{linenomath*}
-where $\mathbf{K}=\mathbf{K}_{\omega,g}$ is the vec-permutation matrix \citep{HendersonSearle1981}.
-The vec-permutation matrix rearrangers the population vectors $\mathbf{n}^S$ and $\mathbf{n}^X$ so that $\mathbf{n}^S$ is ordered by genotype first, and then by stage, but $\mathbf{K} \mathbf{n}^S$ is organized by stage first, and then by genotype (see \cite{hunter2005use} or \cite{caswell2012matrix} for more on the vec-permutation method).
-
-The matrix $\mathbb{U}^S$ in equation (\ref{eq:KDKU_app}) generates transitions and survival of extant individuals within genotypes,  and acts on the the population vector of individuals produced through selfing, $\mathbf{n}^S$..
-The matrix $\bo K$ then permutes the resulting vector so that it is now organised by stages first, and then by genotypes within stages.
-The block diagonal matrix $ \mathbb{D} $ then accounts for changes in genotype within stage classes of extant individuals.
-Since extant individuals generally do not change their genotype $\mathbb{D}$ will be an identity matrix in our model.
-Finally, the matrix $\bo K^\tr $ permutes the population vector once again so that it returns to the original organisation by genotype first and then by stage classes.
-
-Equation (\ref{eq:KHKF_app}) describes reproduction and genotype assignment of newborn individuals.
-The block matrix $\mathbb{F}^S$ contains fecundity rates for individuals of each genotype and stage combination and acts on the the population vector of individuals produced through selfing, $\mathbf{n}^S$.
-Initially, these offspring are associated with the genotype of their parent.
-The vec-permutation matrix $\bo K$ then rearranges the resulting vector so that it is organised by stage first and then by genotype.
-The block-diagonal matrix $\mbb{H}^S$ then allocates the offspring to their genotypes, according to equation (\ref{eq:HS}).
-Finally, $\bo K^\tr $  returns the vector to its initial orientation.
-
-Substituting equations (\ref{eq:US_app}) and (\ref{eq:D_app}) into equation (\ref{eq:KDKU_app}) yields
-\begin{linenomath*}
-\begin{eqnarray}
-\mcal{U}^S &=& \sum_{i=1}^g \bo K^\tr \bo I_{\omega g}	 \bo K \mathbf{E}_{ii} \otimes \mathbf{U}^S_{i}, \\
- &=& \sum_{i=1}^g \mathbf{E}_{ii} \otimes \mathbf{U}^S_{i},
-\end{eqnarray}
-\end{linenomath*}
-since $\bo K^\tr \bo I_{\omega g}	 \bo K=\bo I_{\omega g}$, or in block matrix form this yields equation (\ref{eq:BlkUS_app}), here repeated for convenience, 
-\begin{linenomath*}
-\begin{eqnarray}
-\mcal{U}^S  = 
-		\left(
-			\begin{array}{ccc}
-				\mbf{U}^{S}_{AA} & \mbf{0} & \mbf{0} \\
-				\mbf{0} & \mbf{U}^{S}_{Aa} & \mbf{0} \\
-				\mbf{0} & \mbf{0} & \mbf{U}^{S}_{aa} \\
-			\end{array} \right).\label{eq:BlkUS_app}\\
-\end{eqnarray}
-\end{linenomath*}
-
-Similarly, substituting equations (\ref{eq:FS_app}) and (\ref{eq:HS_app}) into equation (\ref{eq:KHKF_app}) yields,
-\begin{linenomath*}
-\begin{eqnarray}
-\mcal{F}^S &=&\bo K^\tr \mbb{H}^S\bo K \mathbb{F}^S, \label{eq:KHKF_app2}\\
- &=& \sum_{i=1}^g \bo K^\tr \left( \bo I_\om \kron \bo H^S \right)  \bo K \left( \mathbf{E}_{ii} \otimes \mathbf{F}^S_{i} \right).
-\end{eqnarray}
-\end{linenomath*}
-Using $\bo K^\tr \left(\bo A \otimes \bo B  \right) \; \bo K=\bo B \otimes \bo A$ \citep{magnus1979commutation}, we obtain
-\begin{linenomath*}
-\begin{eqnarray}
-\mcal{F}^S  &=& \sum_{i=1}^g  \left(  \bo H^S  \kron \bo I_\om\right) \left( \mathbf{E}_{ii} \otimes \mathbf{F}^S_{i} \right).
-\end{eqnarray}
-\end{linenomath*}
-Finally note that $(\mathbf{A} \otimes \mathbf{B}) (\mathbf{C} \otimes \mathbf{D}) = \mathbf{A} \mathbf{C}\otimes \mathbf{B} \mathbf{D}$  to write
-\begin{linenomath*}
-\begin{eqnarray}
-\mcal{F}^S  &=& \sum_{i=1}^g  \left(  \bo H^S \mathbf{E}_{ii} \right) \kron  \mathbf{F}^S_{i} .
-\end{eqnarray}
-\end{linenomath*}
-Using the exact same sequence of equations, we can derive
-\begin{linenomath*}
-\begin{eqnarray}
-\mcal{F}^X  &=& \sum_{i=1}^g  \left(  \bo H^X(\tilde{\mbf{p}}) \mathbf{E}_{ii} \right) \kron  \mathbf{F}^X_{i} .
-\end{eqnarray}
-\end{linenomath*}
-
-Finally, substitute the expressions derived for the parent-offpspring genotype maps, $\bo H^X(\tilde{\mbf{p}}) $, and $\bo H^S$, in the main text (equations (\ref{eq:HS}) and (\ref{eq:HX})) into the above equations to obtain
-\begin{linenomath*}
-\begin{eqnarray}
-	\mcal{F}^S = \mbf{K}^{\intercal} \mbb{H}^S\mbf{K} \mbb{F}^S = 
-			\left(
-			\begin{array}{ccc}
-				\mbf{F}_{AA} & \frac{1}{4} \mbf{F}_{Aa} & 0 \\
-				0 & \frac{1}{2} \mbf{F}_{Aa} & 0 \\
-				0 & \frac{1}{4} \mbf{F}_{Aa} & \mbf{F}_{aa}\\
-			\end{array} \right),  \label{eq:BlkFS_appendix}
-\end{eqnarray}
-\end{linenomath*}
-\noindent and 
-\begin{linenomath*}
-\begin{equation} \label{eq:BlkFX_appendix}
-	\mcal{F}^X = \mbf{K}^{\intercal} \mbb{H}^X(\tilde{\mbf{p}}) \mbf{K} \mbb{F}^X = 
-			\left(
-			\begin{array}{ccc}
-				q^{\prime}_{A} \mbf{F}_{AA} & \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} & 0 \\ 
-				q^{\prime}_{a} \mbf{F}_{AA} & \frac{1}{2} \mbf{F}_{Aa} & q^{\prime}_{A} \mbf{F}_{aa}  \\ 
-				0 & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} & q^{\prime}_{a} \mbf{F}_{aa}  \\
-			\end{array} \right),
-\end{equation}
-\end{linenomath*}
+			\begin{array}{ccc|ccc}
+				\mbf{U}^S_{AA} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
+				\mbf{0} & \mbf{U}^S_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \mbf{U}^S_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{U}^X_{AA} & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{U}^X_{aa} & \mbf{0}\\
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{U}^X_{aa}
+			\end{array} \right)}_{\tilde{\mbf{U}}} + \\ 
+	&\underbrace{\left(
+			\begin{array}{ccc|ccc}
+				C (1 - \delta) \mbf{F}_{AA} & \frac{1}{4} C (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & C (1 - \delta) \mbf{F}_{AA} & \frac{1}{4} C (1 - \delta) \mbf{F}_{Aa} & \mbf{0} \\ 
+				\mbf{0} & \frac{1}{2} C (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} & \frac{1}{2} C (1 - \delta) \mbf{F}_{Aa} & \mbf{0}  \\
+				\mbf{0} & \frac{1}{4} C (1 - \delta) \mbf{F}_{Aa} & C (1 - \delta) \mbf{F}_{aa} & \mbf{0} & \frac{1}{4} C (1 - \delta) \mbf{F}_{Aa} & C (1 - \delta) \mbf{F}_{aa} \\ \hline
+				q^{\prime}_{A} (1 - C)\mbf{F}_{AA} & \frac{1}{2} q^{\prime}_{A} (1 - C)\mbf{F}_{Aa}  & \mbf{0} & q^{\prime}_{A} (1 - C) \mbf{F}_{AA}& \frac{1}{2} q^{\prime}_{A} (1 - C) \mbf{F}_{Aa} & \mbf{0} \\
+				q^{\prime}_{a} (1 - C) \mbf{F}_{AA} & \frac{1}{2} (1 - C)\mbf{F}_{Aa}  & q^{\prime}_{A} (1 - C)\mbf{F}_{aa}  & q^{\prime}_{a} (1 - C) \mbf{F}_{AA} & \frac{1}{2}(1 - C) \mbf{F}_{Aa}  & q^{\prime}_{A} (1 - C)\mbf{F}_{aa}  \\
+				\mbf{0} & \frac{1}{2} q^{\prime}_{a} (1 - C)\mbf{F}_{Aa} & q^{\prime}_{a} (1 - C) \mbf{F}_{aa} & \mbf{0} & \frac{1}{2} q^{\prime}_{a} (1 - C)\mbf{F}_{Aa} & q^{\prime}_{a} (1 - C)\mbf{F}_{aa} 
+			\end{array} \right)}_{\tilde{\mbf{F}}} \numberthis
+\end{align*}
+
+}
+
+\noindent with symbols as defined in the main text.
+The survival matrices appear on the diagonal because individuals do not change their genotype once they are born.
+The fertility matrix incorporates the process of Mendelian inheritance and is an extension of the fertility matrix derived by \citet{deVriesCaswell2019a}.
+
+The first block column of $\tilde{\mbf{A}}$ describes the production of offspring by an $AA$ female with stage-specific fertility rates $\mbf{F}_{AA}$ by both selfing and outcrossing.
+The probability of picking an $A$ allele out of the pool of available male gametes.
+When reproducing by selfing, this is entirely determined by the probability of sampling an $A$ allele after Mendelian segregation.
+For outcross reproduction the probability, and hence the probability of this $AA$ female producing an $AA$ offspring, is $q^{\prime}_{A}$, as derived above.
+Conversely, the probability of picking an $a$ allele and producing an $Aa$ offspringis zero for an $AA$ female when selfing, but $q^{\prime}_a$ for outcrossing.
+Similarly, the middle column of block matrices are offspring produced by $Aa$ females, which can produce offspring of all $3$ genotypes.
+
+A full derivation of the model, including all component matrices, is provided in the Online Supplementary Material.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section{Appendix B: Eigenvalues for invasion analysis}
+\begin{landscape}
+\section*{Appendix B: Eigenvalues for invasion analysis}
 \renewcommand{\theequation}{B\arabic{equation}}
 \setcounter{equation}{0}  % reset counter 
-
-\textcolor{red}{STILL NEED TO FILL THIS OUT WITH EIGENVALUES ETC.}
-
-denoted by $\tilde{\zeta}_{AA}$ and $\tilde{\zeta}_{aa}$ for the eigenvalue of the Jacobian evaluated at the $AA$ and $aa$ boundary, respectively.
-For our model, the Jacobian, $\mbf{M}$, has an upper block triangular form, and we can therefore express the stability of $\hat{\mbf{p}}$ in terms of the leading eigenvalue of the diagonal block corresponding to changes in heterozygote frequency ($\mbf{M}_{22}$; e.g., see \citealt{OttoDay2007,deVriesCaswell2019b}).
-For a polymorphism to be 'protected', both boundary equilibria must be unstable (i.e., $\tilde{\zeta}_{AA} > 1$ and $\tilde{\zeta}_{aa} > 1$), which requires that
- \begin{linenomath*}
- \begin{equation}\label{eq:PolyCond}
- 	\rho(\mbf{M}_{22}) > \lambda_{i}
- \end{equation}
- \end{linenomath*}
-
- \noindent is satisfied, where $\rho(\cdot)$ represents the spectral radius,  $\rho(\mbf{M}_{22})$ gives the ergodic growth rate of perturbations of the boundary equilibrium according to the linear approximation, and $\lambda_i$ is the intrinsic growth rate for the homozygous genotype being invaded at each boundary ($i \in (AA,aa)$; see Box.~1).
+\setcounter{table}{0}  % reset counter
+
+The invasion analysis is made easier if we first reorder the population vector by genotype, then by how individuals were produced (selfing vs.~outcrossing), and finally by stage.
+Ordering by genotype first facilitates the invasion analysis in two ways.
+First, the resulting Jacobian matrix is upper block triangular, and the eigenvalues of $\mbf{M}$ are therefore the eigenvalues of the diagonal blocks.
+Second, the blocks along the diagonal correspond to perturbations in each of the the three genotype 'directions' at the boundary equilibria.
+As outlined in the Online Supplementary Material, the stability of both boundary equilibria $\hat{\mbf{p}}_{AA} = 1$ and $\hat{\mbf{p}}_{aa} = 1$ are determined by the central block of the Jacobian, $\mbf{M}_{22}$, which corresponds to perturbations in the direction of heterozygote $Aa$ genotypes.
+A boundary equilibrium is unstable to invasion by the rare allele if the largest absolute value of the eigenvalues of the Jacobian matrix, the leading eigenvalue, evaluated at the equilibrium is greater than $1$.
+The resulting conditions for a 'protected polymorphism' require that
+
+{\footnotesize
+\begin{linenomath*}
+ \begin{equation} \label{eq:coexist_AA}
+	\lambda_{AA} < 
+			\rho\left(\begin{array}{ccc}
+\mathbf{U}^S_{Aa} +\frac{1}{2} C (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C (1 - \delta) \mbf{F}_{Aa}  & \mbf{0} \\
+ \frac{1}{2} (1 - C) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C)\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C)\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & (1-C)\mbf{F}_{aa} + \Theta\\
+  \frac{1}{4} C (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C (1 - \delta) \mbf{F}_{aa}  \\ 
+			\end{array} \right), \\ 
+\end{equation} 
+\end{linenomath*}
+
+\noindent and 
+\begin{linenomath*}
+ \begin{equation} \label{eq:coexist_aa}
+	\lambda_{aa} < 
+			\rho\left(\begin{array}{ccc}
+\mathbf{U}^S_{Aa} +\frac{1}{2} C (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C (1 - \delta) \mbf{F}_{Aa}  & \mbf{0} \\
+ \frac{1}{2} (1 - C) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C)\mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C)\mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & (1-C)\mbf{F}_{AA} + \Phi\\
+  \frac{1}{4} C (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{AA} + C (1 - \delta) \mbf{F}_{AA}  \\ 
+			\end{array} \right), \\ 
+\end{equation} 
+\end{linenomath*}
+ }
+
+\noindent is satisfied, where where $\rho(\cdot)$ represents the spectral radius, and $\Phi = \frac{(1 - C)}{p_{n}} \mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime$.
+$\rho(\mbf{M}_{22})$ gives the ergodic growth rate of perturbations of the boundary equilibrium according to the linear approximation, and $\lambda_i$ is the intrinsic growth rate for the homozygous genotype being invaded at each boundary ($i \in (AA,aa)$).
+
+In our numerical simulations, equations Eq(\ref{eq:coexist_AA}) and Eq(\ref{eq:coexist_aa}) were evaluated to determine the invasion thresholds for the relvant parameter conditions described in each figure.
+\end{landscape}
+
+Under obligate outcrossing (when $C=0$), the leading eigenvalue evaluated at the $\hat{\mbf{p}}_{AA} = 1$ and $\hat{\mbf{p}}_{aa} = 1$ boundaries is equal to
+\begin{align*} \label{eq:eigAAOut}
+	\tilde{\zeta}_{AA} = 
+		\frac{1}{\lambda_{AA}}\rho\left(\begin{array}{cc}
+\mathbf{U}^S_{Aa}  & \mathbf{0} \\
+ \frac{1}{2} \mbf{F}_{Aa}+\frac{1}{2\mbox{p}_{n}}\mbf{F}_{AA} \hat{\mbf{p}}_{X,AA} \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} \mbf{F}_{Aa} +\frac{1}{2\mbox{p}_{n}} \mbf{F}_{AA} \hat{\mbf{p}}_{X,AA} \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime
+			\end{array} \right), \numberthis
+\end{align*}
+
+\noindent and
+
+\begin{align*} \label{eq:eigaaOut}
+	\tilde{\zeta}_{aa} = 
+		\frac{1}{\lambda_{aa}}\rho\left(\begin{array}{cc}
+\mathbf{U}^S_{Aa}  & \mathbf{0} \\
+ \frac{1}{2} \mbf{F}_{Aa}+\frac{1}{2\mbox{p}_{n}}\mbf{F}_{aa} \hat{\mbf{p}}_{X,aa} \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} \mbf{F}_{Aa} +\frac{1}{2\mbox{p}_{n}} \mbf{F}_{aa} \hat{\mbf{p}}_{X,aa} \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime
+			\end{array} \right), \numberthis
+\end{align*}
+
+\noindent which reduce to 
+\begin{align*} \label{eq:eigAATwoSex}
+	\tilde{\zeta}_{AA} = 
+		\frac{1}{\lambda_{AA}}\rho\left(\begin{array}{c}
+\mathbf{U}^X_{Aa} + \frac{1}{2} \mbf{F}_{Aa} +\frac{1}{2\mbox{p}_{n}} \mbf{F}_{AA} \hat{\mbf{p}}_{X,AA} \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime
+			\end{array} \right), \numberthis
+\end{align*}
+
+\noindent and 
+\begin{align*} \label{eq:eigaaTwoSex}
+	\tilde{\zeta}_{aa} = 
+		\frac{1}{\lambda_{aa}}\rho\left(\begin{array}{c}
+\mathbf{U}^X_{Aa} + \frac{1}{2} \mbf{F}_{Aa} +\frac{1}{2\mbox{p}_{n}} \mbf{F}_{aa} \hat{\mbf{p}}_{X,aa} \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime
+			\end{array} \right), \numberthis
+\end{align*}
+
+\noindent when $\mathbf{U}^S_{Aa}=\mathbf{U}^X_{Aa}$.
+These eigenvalues are identical to those derived by \citet{deVriesCaswell2019a}, confirming that our model predictions reduce to the two-sex model under obligate outcrossing, as expected.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section{Appendix C: Demographic and empirical data for {\itshape M.~guttatus} case study}
+\section*{Appendix C: Demographic and empirical data for {\itshape M.~guttatus} case study}
 \renewcommand{\theequation}{C\arabic{equation}}
@@
 \setcounter{equation}{0}  % reset counter 
+\setcounter{table}{0}  % reset counter
+
+\subsection*{Demographic data}
+
+We used demographic data for multiple experimental populations of {\itshape M.~guttatus} from a large-scale common garden experiment conducted in Stanislaus National Forest (California, USA) in 2012 and 2013 \cite{PetersonEtAl2016}.
+Specifically, we used data from two experimental populations with contrasting demographic rates: the locally adapted 'Eagle Meadows' population (data from 2012), and a non-local 'low-elevation perennials' population (data from 2013).
+The vital rates used in theirs, and our, calculations are summarized below in Table(\ref{tab:MimDemData}; see corrected Tables.~1 and S2 in \citet{PetersonEtAl2016, PetersonEtAl2017}.
+Note that the same estimates for $D$ and $A$ were used for all populations in the study.
+
+The resulting transition matrices for these populations involve three life-history stages ($\omega = 3$; seed, seedling, and rosette; see Matrix 1 \citealt{PetersonEtAl2016}), and can represented as the sum of survival and fertilitiy matrices as follows:
+
+\begin{linenomath*}
+\begin{equation} \label{eq:AtildeMimulus}
+	\mathbf{\tilde{A}}[\tilde{\mbf{n}}] = 
+			\underbrace{\left(
+			\begin{array}{ccc}
+				D(1 - G) & 0 & 0 \\
+				DG & FOAG & FOAG \\
+				0 & SR & SR \\
+			\end{array} \right)}_{\tilde{\mbf{U}}} +
+			\underbrace{\left(
+			\begin{array}{ccc}
+				0 & FOA(1 - G) & FOA(1 - G) \\
+				0 & 0 & 0 \\
+				0 & 0 & 0 \\
+			\end{array} \right)}_{\tilde{\mbf{F}}}.
+\end{equation}
+\end{linenomath*}
+
+Overall transition matrix data for the same populations is also available on the public demographic database COMPADRE \citep{CompadreDB2020}, but are not decomposed into the product of terms described in Eq(\ref{eq:AtildeMimulus}) above.
+
+As described in the main text, we incorporated empirical estimates of the population selfing rate and inbreeding depression for two intensively studied populations in the Cascade Mountains of Oregon into our parameterized model (Iron Mountain and Cone Peak; \citealt{Willis1993,Willis1999a,Willis1999b}).
+Using the data provided in Table.~2 of \citet{Willis1993}, we calculated the proportional decrease due to inbreeding depression in three fitness components that were also included in vital rate estimates of \citet{PetersonEtAl2016}: seed germination rate ($G$), flower number ($F$), and overwinter survival ($S$).
+In each case, we calculated the inbreeding depression terms as $\delta_i = 1 - w_{\text{inbred}}/w_{\text{outcross}}$.
+To incorporate these terms into our evolutionary demographic model, we multiplied each of the three terms $G$, $F$, and $S$ in Eq(\ref{eq:AtildeMimulus}) by a corresponding proportional inbreeding depression term $(1 - \delta_i)$, where $i \{\in G,F,S\}$.
+The largest field-estimated selfing rate for this same Iron Mountain population was $C = 0.29$ \cite{Willis1993}, which we use in our calculations for fig.~\ref{fig:mimulusFig}.
+
+\begin{table}[htbp]
+ \centering
+ \caption{\bf Vital rates estimated for {\itshape M.~gutattus} experimental populations by \citet{PetersonEtAl2016}}
+\label{tab:MimDemData}
+\begin{tabular}{lcc}
+ \toprule
+ Vital rate estimated &  Eagle Meadows (2012) & Low-Elevation Perrenials (2013)\\ \hline
+ Seed bank survival (D) & $0.534$ & $0.534$ \\
+ Seed germination rate (G) & $0.469$ & $0.652$ \\
+ Flower production (F) & $0.64$ & $4.09$ \\
+ Ovules per flower (O) & $614$ & $494$ \\
+ Seedling recruits (A) & $6.7 \times 10^{-4}$ & $6.7 \times 10^{-4}$ \\
+ Overwinter survival (S) & $0.179$ & $0$ \\
+ Rosette production (R) & $8.71$ & $0$ \\
+\hline
+ Population intrinsic growth rate ($\lambda$) & $1.71$ & $1.08$ \\
+ \hline
+\end{tabular}
+\end{table}
+
+\subsection*{Selection on inv6}
+
+\citet{LeeKelly2015} describe a polymorphic chromosomal inversion segregating in the Iron Mountain population of {\itshape M.~guttatus}.
+Importantly, the study reports fitness effects of inv6 on both male and female fitness components.
+inv6 appears to cause a significant decrease in pollen viability, while simultaneously causing an increase in flower number which varies among years.
+Unfortunately, it was not possible to estimate the dominance of the fitness effects of inv6 in the field because individuals homozygous for the inversion were quite rare (although they were viable in greenhouse conditions).
+We therefore made the relatively conservative assumption that the fitness effects of inv6 on pollen viability and flower number were both additive.
+Under this assumption, it is possible to estimate selection coefficients for proxies of female and male fertilities.
+
+We used flower number as a proxy for female fertility.
+To calculate selection on flower number, we first calculated the geometric mean flower number in the field for wild-type and inv6 heterozygotes across the two years for which data was available (2012 and 2013), extrapolated the expected flower number for inv6 homozygotes, and calculated the relative fertility of each genotype given that inv6 homozygotes have the highest flower number ($w_i$, where $i \in \{inv6/inv6,\, inv6/w+,\, w+/w+\}$).
+Using the relative fertility expressions outlined in Table \ref{tab:inv6Fitness}, we then solved for $s_f$.
+
+We used pollen production as a proxy for male fertility.
+We calculated average pollen production as the product of the geometric mean flower number and field estimated pollen viability for wild-type and inv6 heterozygotes, and extrapolated the expected pollen production for inv6 homozygotes.
+As before, we calculated relative male fertility ($w^{\prime}_i$) for each genotype given that wild-type homozygotes had the highest pollen viability, and used the expressions for $w^{\prime}_{i}$ from Table \ref{tab:inv6Fitness} to solve for $s_m$.
+
+It is worth noting that \citet{LeeKelly2015} also report greenhouse pollen viabilty data for several mapping populations of {\itshape M.~guttatus}, which includes data for inv6 homozygotes, from which relative fertilities can be estimated.
+We chose to use the field estimated pollen viability and flower number data, despite the lack of inv6 homozygotes, because it probably gives a more accurate reflection of the fitness effects of inv6 in the field.
+Moreover, both approaches require making assumptions regarding the dominance effects of inv6.
+Given the obvious pitfalls of estimating selection on inv6 from the available data, it goes without saying that the location of inv6 in demographically viable polymorphic parameter space is a highly speculative, but nevertheless interesting, proof of concept for our model predictions.
+
+\begin{table}[htbp]
+ \centering
+ \caption{\bf Relative fertilities for inv6 genotypes ($w_{i}$)}
+\label{tab:inv6Fitness}
+\begin{tabular}{lccc}
+ \toprule
+					&  \multicolumn{3}{c}{{\textit{Genotype}}} \\ 
+\cline{2-4}
+									& $inv6/inv6$	& $inv6/w+$		& $w+/w+$ 	\\ \hline
+Female function ($w_{i}$):			& $1$			& $1 - s_f/2$	& $1 - s_f$ \\	
+Male function ($w^{\prime}_{i}$):	& $1 - s_m$		& $1 - s_m/2$	& $1$ 		\\	
+\hline
+\end{tabular}
+\end{table}
 

### 2021-03-29 R/functions-Figs.R

@@ -830,30 +830,18 @@ MimulusInv6Fig  <-  function() {
 
     # import data
     path  <-  './output/simData/'
-    inv1  <-  read.csv(paste(path, "invBoundMimulus_sMax0.99_res0.01_hf0.5_hm0.35_C0_deltaF_useCompadreFALSE", '.csv', sep=""), header=TRUE)
-    inv2  <-  read.csv(paste(path, "invBoundMimulus_sMax0.99_res0.01_hf0.5_hm0.35_C0.29_deltaF_useCompadreFALSE", '.csv', sep=""), header=TRUE)
-    inv3  <-  read.csv(paste(path, "invBoundMimulus_sMax0.99_res0.01_hf0.5_hm0.35_C0.29_deltaT_useCompadreFALSE", '.csv', sep=""), header=TRUE)
-    inv4  <-  read.csv(paste(path, "invBoundMimulus_sMax0.99_res0.01_hf0.5_hm0.35_C0_deltaF_useCompadreFALSE_LEP", '.csv', sep=""), header=TRUE)
-    inv5  <-  read.csv(paste(path, "invBoundMimulus_sMax0.99_res0.01_hf0.5_hm0.35_C0.29_deltaF_useCompadreFALSE_LEP", '.csv', sep=""), header=TRUE)
-    inv6  <-  read.csv(paste(path, "invBoundMimulus_sMax0.99_res0.01_hf0.5_hm0.35_C0.29_deltaT_useCompadreFALSE_LEP", '.csv', sep=""), header=TRUE)
-#    ext1  <-  read.csv(paste(path, "extThresholdMimulus_SfxSm_sMax0.99_res0.01_hf0.5_hm0.35_C0_useCompadreFALSE_IDFALSE", '.csv', sep=""), header=TRUE)
-#    ext2  <-  read.csv(paste(path, "extThresholdMimulus_SfxSm_sMax0.99_res0.01_hf0.5_hm0.35_C0.29_useCompadreFALSE_IDFALSE", '.csv', sep=""), header=TRUE)
-#    ext3  <-  read.csv(paste(path, "extThresholdMimulus_SfxSm_sMax0.99_res0.01_hf0.5_hm0.35_C0.75_useCompadreFALSE_IDFALSE", '.csv', sep=""), header=TRUE)
[…]

### 2021-03-29 makeFigs.R

@@ -43,8 +43,8 @@ embed_fonts(figPath(name='deltaPolymorphicSpaceTitrate.pdf'))
 # Fig. 4 - Illustration of polymorphic parameter space w/ 
 #			Inv6 plotted
 toPdf(MimulusInv6Fig(), 
-			figPath(name='MimulusInv6Fig.pdf'), width=10, height=5)
-embed_fonts(figPath(name='MimulusInv6Fig.pdf'))
+			figPath(name='MimulusInv6FigNewSel.pdf'), width=10, height=5)
+embed_fonts(figPath(name='MimulusInv6FigNewSel.pdf'))
 
 
 #################

### 2021-03-29 run-Simulations.R

@@ -896,9 +896,6 @@ extinctThreshMimulus(sMax = 0.99, res=0.01, precision=1e-4,
 					 tlimit = 10^5, intInit = TRUE, Ainvade=FALSE, eqThreshold=1e-9, 
 					 writeFile=TRUE, verbose=TRUE) 
 
-test  <-  fwdSimMimulusDat(datMat=datMat, theta.list=theta.list, delta.list=delta.list, useCompadre = FALSE,
-				 hf = hf, hm = hm, sf = 0.308, sm = 0.88, C = 0.24,
-				 tlimit = 10^5, eqThreshold=1e-9, Ainvade = FALSE, intInit = TRUE)
 
 ##########################
 # Low-elevation perenials (Peterson et al. 2016; Correction Table 1)
@@ -963,17 +960,244 @@ extinctThreshMimulus(sMax = 0.99, res=0.01, precision=1e-4,
 	
 
 
-test  <-  fwdSimMimulusDat(datMat=datMat, theta.list=theta.list, delta.list=delta.list, useCompadre = FALSE,
-				 hf = 1/2, hm = 0.35, sf = 0.308, sm = 0, C = 0,
-				 tlimit = 10^5, eqThreshold=1e-9, Ainvade = FALSE, intInit = TRUE)
+
+
+
+
+#################################################
+#' Find extinction threshold for Mimulus example
+#' TAKING AVERAGE (2012 & 2013) FLOWER # EFFECT 
+#' INTO ACCOUNT WHEN CALCULATING MALE SELECTION
+#' COEFFICIENTS AND DOMINANCE FOR INV6
+# Eagle Meadows population  (Peterson et al. 2016)
+datMat  <-  matList_Mg_EM
+theta.list  <-  list(D = 0.534, 
+					 G = 0.469,
+					 F = 0.64,
+					 O = 614,
+					 A = 6.7e-4,
+					 S = 0.179,
+					 R = 8.71,
+					 pop = "EM")
+delta.list  <-  list(delta_D = 0,
+					 delta_G = 0,
+					 delta_F = 0,
+					 delta_O = 0,
+					 delta_S = 0
+					 )
+
+
+## Obligate oucrossing
+titrateInvBoundMimulus(sMax = 0.99, res=0.01, precision=1e-4, 
+						datMat=datMat, theta.list=theta.list, delta.list=delta.list, useCompadre = FALSE,
+						hf=1/2, hm = 1/2, C=0, 
[…]

### 2021-05-03 doc/SA-Hermaphrodites-wDemography.tex
Minor Edits
Honestly, don't remember what this commit does. Minor changes to
the MS.

@@
 On the one hand, they prevent individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase a population's overall extinction risk \citep{kokko2003sexy,harts2014demography}.
-On the other hand, they provide an effective mechanism for the maintenance of genetic variation, which influences a population's capactiy for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
+On the other hand, they provide an effective mechanism for the maintenance of genetic variation, which influences a population's capacity for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
 Hence, for traits under conflicting selection, the nature and extent of genetic variation observed in natural populations should reflect a balance between the maintainance of genetic polymorphisms, and the population dynamical consequences of the resulting maladaptation.

### 2021-05-06 doc/SA-Hermaphrodites-wDemography.tex
Changes Re: comments from Tim and Crispin
This commit makes changes to the manuscript in response to comments
from Tim and Crispin on our earlier draft. Minor changes throughout,
but most notable changes include a slight change to the title, a
couple new (short) paragraphs, and a couple new references.

@@
 \begin{document}
-\title{Demographic costs of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites}
+\title{The demographic costs of sexually antagonistic selection in simultaneous hermaphrodites}
 
@@
 Classic population genetics theory has been fundamental to understanding the evolution of sex-differences and the maintenance of sexually antagonistic (SA) genetic variation, but these models have rarely considered the demographic consequences of intralocus sexual antagonism.
-In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations.
-We focus on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population intrinsic growth rate remains positive.
+In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious/gonochorous) and partially selfing hermaphrodite populations.
+We focus on identifying parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population intrinsic growth rate remains positive.
 In addition, we analyze the effects of inbreeding depression manifesting in different life-history stages, and give an illustrative example of the potential for SA polymorphism in real populations using empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus}.
-Our results indicate that when population growth rates are near one, large swathes of polymorphic parameter space become demographically inviable, and that inbreeding depression poses a significant challenge to the maintenance of SA polymorphism in partially selfing populations.
+Our results indicate that as population intrinsic growth rates approach one, large swathes of polymorphic parameter space become demographically inviable, and that inbreeding depression poses a significant challenge to the maintenance of SA polymorphism in partially selfing populations.
 Despite these demographic challenges, our example with {\itshape M.
@@
 To persist in the long term, a population must be able to adapt to its environment.
-Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, such as deleterious mutations \citep{Haldane1957}, changing environmental conditions \citep{Maynard-Smith1976, LandeShannon1996,OrrUnckless2008}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, and genetic constraints \citep{ConnallonHall2018, MatthewsConnallon2019}.
+Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, such as deleterious mutations \citep{Haldane1937}, changing environmental conditions \citep{Maynard-Smith1976, LandeShannon1996,OrrUnckless2008}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, and genetic constraints \citep{ConnallonHall2018, MatthewsConnallon2019}.
 Genetic constraints on adaptation can arise as a consequence of conflicting selection and gene-flow between different classes of individuals, or more generally, when there are genetic trade-offs between fitness components \citep{CharlesworthHughes2000, ConnallonHall2018}.
@@
 On the one hand, they prevent individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase a population's overall extinction risk \citep{kokko2003sexy,harts2014demography}.
-On the other hand, they provide an effective mechanism for the maintenance of genetic variation, which influences a population's capacity for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
+On the other hand, they provide an effective mechanism for the maintenance of genetic variation, which can increase a population's capacity for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
 Hence, for traits under conflicting selection, the nature and extent of genetic variation observed in natural populations should reflect a balance between the maintainance of genetic polymorphisms, and the population dynamical consequences of the resulting maladaptation.
 
-Sexually antagonistic selection (abbreviated SA hereafter) arising from genetic trade-offs between male and female fitness, is a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and the evolution of sexual dimorphism and gender-related traits \citep{Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, BondurianskyChenoweth2009,Olito2019}.
+Sexually antagonistic selection (abbreviated SA hereafter) arising from genetic trade-offs between male and female fitness, is a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and the evolution of sexual dimorphism and sex-related traits \citep{Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, BondurianskyChenoweth2009,Olito2019}.
 When such genetic trade-offs occur, SA selection can arise when beneficial alleles for one sex are deleterious when expressed in the other \citep{Kidwell1977, Rice1992, ConnallonClark2012}.
@@
 Alleles with opposing fitness effects through male and female sex functions can cause analogous genetic constraints on fitness in hermaphrodite populations, where both maternal and paternal reproductive success contribute jointly to each individuals' overall fitness \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
-Self-fertilization in hermaphrodites is predicted to reduce the total parameter space where balanced SA polymorphisms can be maintained, while simultaneously creating a bias in selection through the female sex function (\citealt{JordanConnallon2014}; but see \citealt{Tazzyman2015}).
-Yet, other factors, such as genetic linkage to other SA loci or a sex-determining region \citep{Otto2011, JordanCharlesworth2012, Olito2017, Olito2019}, or spatial heterogeneity and complexity of the life-cycle can expand the parameter space for SA polymorphism \citep{Olito-etal-2018,ConnallonSharmaOlito2019}.
+Self-fertilization in hermaphrodites is predicted to reduce the total parameter space where balanced SA polymorphisms can be maintained, while simultaneously creating a bias in selection through the female sex function (\citealt{JordanConnallon2014,Glemin2021}; but see \citealt{Tazzyman2015}).
+Yet, other factors, such as genetic linkage to other SA loci or a sex-determining region \citep{Otto2011, JordanCharlesworth2012, Olito2017, Olito2019}, antagonism between viability and fecundity selection \citep{Glemin2021}, or spatial heterogeneity and complexity of the life-cycle can expand the parameter space for SA polymorphism \citep{Olito-etal-2018,ConnallonSharmaOlito2019, Glemin2021}.
 Overall, both theoretical predictions and current empirical data suggest that there is ample scope for SA trade-offs and the maintenance of SA polymorphisms in both dioecious and hermaphrodite populations \citep{Abbott2011, WangBarrett2020}, although identifying specific SA loci from genome sequence data remains challenging \citep{RuzickaESEB2020}.
@@
 In fact, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals (e.g., \citealt {johnston2013life, merot2020balancing}).
-To take the population dynamical consequences of polymorphism into account, a model is required that links changes in genotype distributions to population dynamics.
-The potential demographic costs of sexual antagonism were pointed out by \cite{kokko2003sexy}, but few papers since then have taken this into account (with the notable exception of \citealt{harts2014demography}).
+To take the population dynamical consequences of SA selection into account, a model is required that links changes in genotype distributions to population dynamics.
+The potential demographic costs of sexual antagonism were pointed out by \citep{kokko2003sexy}, but few papers since then explicitly incorporated demography into models of sexual antagonism (with the notable exception of \citealt{harts2014demography}).
 \citet{deVriesCaswell2019b} introduced a mendelian matrix model with intralocus sexual antagonism, and population dynamics but did not perform an analysis of the consequences of demographic viability for the scope of sexual antagonism to maintain polymorphism.
 
-In this paper, we extend the model of \cite{deVriesCaswell2019b} to include both dioecious and hermaphroditic species in order to study the parameter conditions under which balanced SA polymorphisms are also demographically viable.
+In this paper, we extend the model of \citep{deVriesCaswell2019b} to include both dioecious and hermaphroditic species in order to study the parameter conditions under which balanced SA polymorphisms are also demographically viable.
 A major strength of our approach is that the model can be parameterized using empirically estimated demographic rates, enabling us to make predictions about the scope for demographically viable SA polymorphisms that are grounded in the biology of real populations.
@@
 
-Here, we briefly describe a matrix model incorporating multiple life-cycle stages and a single diallelic locus under SA selection on male and female fertility for populations of simultaneous hermaphrodites with outcrossing or mixed-mating systems.
+Here, we briefly describe a matrix model incorporating multiple life-cycle stages and a single diallelic locus under SA selection on male and female fertility for populations of partially-selfing simultaneous hermaphrodites.
 The derivation and some of the key results follow closely those presented in \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b}.
-In fact, the present model reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019b} under obligate outcrossing.
+In fact, the model reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019b} under obligate outcrossing.
 A full derivation of the model and analyses is presented in the \hl{Online Supplementary Material}, and all computer code necessary to reproduce the results are available at \url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
@@
 
-% Notational note: The superscript on a matrix in age$\times$stage models conventionally indicates the stage or age of the subsection of the population vector that the matrix is acting on (see for example \cite{CaswellEtAl2018}).
+% Notational note: The superscript on a matrix in age$\times$stage models conventionally indicates the stage or age of the subsection of the population vector that the matrix is acting on (see for example \citep{CaswellEtAl2018}).
 For example, $\mbf{U}^{S}_{i}$, acts on the vector of individuals of genotype $i$ that were produced through selfing.
@@
 
-For the purpose of this article, we assume that mating is random with respect to stage and hence that the parent-offspring map is the same for all stages (i.e., $\mbf{H}^S_{j}(\tilde{\mbf{n}}) = \mbf{H}^S(\tilde{\mbf{n}})$, and $\mbf{H}^X_{j}(\tilde{\mbf{n}}) = \mbf{H}^X(\tilde{\mbf{n}})$).
+For the purpose of this article, we assume that mating is random with respect to stage and hence that the parent-offspring map is the same for all stages (i.e., $\mbf{H}^S_{j}(\tilde{\mbf{n}}) = \mbf{H}^S(\tilde{\mbf{n}})$, and $\mbf{H}^X_{j}(\tilde{\mbf{n}}) = \mbf{H}^X(\tilde{\mbf{n}})$; where $j \in \{1,\omega\}$).
 The matrices $\mbf{H}^S(\tilde{\mbf{n}})$ and $\mbf{H}^X(\tilde{\mbf{n}})$ contain the population genetic processes and are presented in the next section.
@@
 For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986, WangBarrett2020, Harder2016}.
-
-Because we assume that how individuals were produced (either by selfing or outcrossing) does not affect their relative mating success, the frequency in the male mating population is obtained by simply summing the vectors of individuals produced through selfing and outcrossing:
+Note, however, that by modeling paternal relative mating success rather than pollen/sperm production, we implicitly assume female demographic dominance (i.e., production/transport of male gametes does not limit ovule fertilization and hence population growth).
+
+Because we assume that inbreeding depression only affects offspring viability, the allele frequency in the male mating population is obtained by simply summing the vectors of individuals produced through selfing and outcrossing:
 \begin{linenomath*}
@@
 
-The parent-offpsring map for reproduction via self-fertilization differs from the outcrossing parent-offspring map.
-As in previous models of partial selfing, we assume that individuals produce enough male gametes to easily self-fertilize all of their ovules, and that self-fertilization involves little or no selection from external factors relative to outcrossing (e.g., \citealt{Charlesworth1978a,JordanConnallon2014,Olito2017}, but see \citealt{Tazzyman2015}).
+The parent-offspring map for reproduction via self-fertilization differs from the outcrossing parent-offspring map.
+As in previous models of partial selfing, we assume that individuals produce enough male gametes to easily self-fertilize all of their ovules, and that self-fertilization involves little or no selection via the male sex function from external factors (e.g., pollinator visitation) relative to outcrossing (\citealt{Charlesworth1978a,JordanConnallon2014,Olito2017}, but see \citealt{Tazzyman2015}).
 Under these assumptions, the genotype distributions of selfed offspring are determined entirely by the parental genotype and the probabilities of segregation and fertilization during and after meiosis:
@@
 As described in the next section, these parent-offspring maps influence the projection matrix by altering the fertilitiy matrices for individuals produced by selfing and outcrossing (see Eq(\ref{eq:BlkFS}) and Eq(\ref{eq:BlkFX}) below).
+
+Following previous theory for partially selfing populations \cite[e.g.,][]{Charlesworth2010,JordanConnallon2014,Glemin2021}, we use a 'fixed prior selfing' model where all individuals are assumed to self-fertilize at a constant rate, $C$, prior to receiving outcross pollen, for reasons of analytic tractability.
+In the \hl{Online Supplement}, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
 
@@
 where $C$ denotes the proportion of each individual's ovules that are self-fertilized (the remaining $1 - C$ are outcrossed), and  $\delta$ represents the proportion of self-fertilized zygotes that fail to develop due to inbreeding depression during early development \citep{Charlesworth1987}.
-In the \hl{Online Supplement}, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
 
@@
 Suppose that there is a genetic trade-off between the sex-functions at a single diallelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect.
-Following convention, we parameterize the fertility component of fitness of each genotype through each sex function, $w_{i}$ and $w^{\prime}_{i}$, to be bounded by $[0,1]$, with dominance and selection coefficients $h_f, s_f$ and $h_m, s_m$ determining the decrease in fertility through each sex function relative to the most fit genotype ($AA$ has highest female fertility, $aa$ the highest male fertility; see Table \ref{tab:Fitness}).
+Following convention, we parameterize the fertility component of fitness for each genotype through each sex function, $w_{i}$ and $w^{\prime}_{i}$, to be bounded by $[0,1]$, with dominance and selection coefficients $h_f, s_f$ and $h_m, s_m$ determining the decrease in fertility through each sex function relative to the most fit genotype ($AA$ has highest female fertility, $aa$ the highest male fertility; see Table \ref{tab:Fitness}).
 
@@
 We quantify early-acting inbreeding depression using $\delta$, and late-acting inbreeding depression using $\delta_i$ (where $i \in \{j,a,\gamma\}$).
-An important difference between early- and late-acting inbreeding depression in the model is that $\delta$ does not affect $\mbf{U}^S$, the survival matrix for offspring produced by selfing (see Eq.~\ref{eq:Atilde}), while the $\delta_i$ terms directly alter the stage-specific survival and transition rates in $\mbf{U}^S$ (see Eq.~\ref{eq:US}).
+$\delta$ denotes the fraction of self-fertilized ovules that do not develop into juveniles due to inbreeding depression.
+An important difference between early- and late-acting inbreeding depression in the model is that $\delta$ does not affect $\mbf{U}^S$, the survival matrix for offspring produced by selfing (see Eq.~\ref{eq:Atilde}), while the $\delta_i$ terms directly alter the stage-specific survival and transition rates of offspring produced by selfing (see Eq.~\ref{eq:US}).
 
@@
 
-Unless stated otherwise, we use the following parameter values for the demographic rates in the model: $\sigma_j = \sigma_a = 0.6$ and $\gamma = 0.05$, which are similar to those used in \citet{deVriesCaswell2019b}.
+Unless stated otherwise, we use the following parameter values for the demographic rates in the model: $\sigma_j = \sigma_a = 0.6$ and $\gamma = 0.05$.
+These chosen values are similar to those used in \citet{deVriesCaswell2019b}, facilitating comparison between models, and to correspond to a life-history in which individuals spend multiple timesteps in the juvenile phase prior to maturing into reproductively active adults, but are otherwise arbitrary.
 Our parameters of interest include fertility, $f$, the inbreeding depression parameters, $\delta$, $\delta_j$, $\delta_a$, $\delta_{\gamma}$, and the selection parameters $h_f$, $s_f$, $h_m$, and $s_m$, which are given different values for each analyses as described in the figure captions.
@@
 Specifically, for each boundary equilibrium we introduced the rare allele at low initial frequency, and iterated Eq(\ref{eq:Atilde}) until the population had reached demographic and genotypic equilibrium.
-In our model with density independent demographic rates, the population state vector will grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see chapter 17 in \citealt{Caswell2001}), and the intrinsic population growth rate after convergence, $\lambda$, can be calculated as $\tilde{\mbf{n}}(t)/\tilde{\mbf{n}}(t-1)$.
+Our model implements density independent demographic rates, and so the population state vector will grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see chapter 17 in \citealt{Caswell2001}).
+The intrinsic population growth rate after convergence, $\lambda$, can be calculated as $\tilde{\mbf{n}}(t)/\tilde{\mbf{n}}(t-1)$.
 We note, however, that if the ecological component of the model is non-linear, more exotic dynamics are possible \citep{de2020matrix}.
@@
 We therefore used numerical techniques to identify invasion and extinction thresholds, which could then be used to define regions of demographically viable polymorphism across $s_f \times s_m$ parameter space.
-Because single-locus selection coefficients are generally weak (i.e., $s < 0.1$; e.g., \citealt{Eyre-WalkerKeightly2007}) and strongly skewed, we limit our analyses to coefficients within $0 < s_f,s_m \leq 0.15$, unless stated otherwise.
-For simplicity, we also assume equal dominance in all analyses (i.e., $h_f = h_m = h$).
+Because single-locus selection coefficients are generally weak (e.g., \citealt{Eyre-WalkerKeightly2007}) and strongly skewed, we limit our analyses to coefficients within $0 < s_f,s_m \leq 0.15$, unless stated otherwise.
+For simplicity, we present results for scenarios of equal dominance in the main text (i.e., $h_f = h_m = h$).
+Specifically, we examine scenarios of additive SA fitness effects ($h = 1/2$) and dominance reversals, where the deleterious fitness effect of each SA allele is partially recessive in each sex ($h = 1/4$).
+We relax these assumptions to explore the consequences of sex-specific dominance, which may evolve under asymmetric SA selection \citep{SpencerPriest2016}, on our model predictions in \hl{Appendix X}).
+
 For a given parameter set, we numerically evaluate the largest eigenvalue of the Jacobian evaluated at both boundaries, and the population growth rate of the equilibrium obtained after perturbation of the boundary, $\lambda$.
@@
 
-For analyses of the demographic effects of inbreeding depression, we made two main simplifying assumptions: First, we assume that if inbreeding depression is caused primarily by recessive deleterious mutations (as suggested by empirical data), it should covary negatively with the population selfing rate due to purging (\citealt{Charlesworth2009}; though we note that other processes could give rise to this pattern, e.g., \citealt{CrnokrakBarrett2002, Charlesworth2009,HedrickGarcia-Dorado2016}).
+For analyses of the demographic effects of inbreeding depression, we make two additional simplifying assumptions: First, we assume that if inbreeding depression is caused primarily by recessive deleterious mutations (as suggested by empirical data), it should covary negatively with the population selfing rate due to purging (\citealt{Charlesworth2009}; though we note that other processes could give rise to this pattern, e.g., \citealt{CrnokrakBarrett2002, Charlesworth2009,HedrickGarcia-Dorado2016}).
 Following \citet{Olito2019}, we incorporate such negative covariance by constraining the inbreeding depression terms in the model ($\delta$ and $\delta_i$, where $i \in \{j,a,\gamma\}$) to follow a simple declining function of the selfing rate: $\delta = \delta^{\ast} (1 - b (1 - L))$, where $\delta^{\ast}$ is the hypothetical severity of inbreeding depression for a completely outcrossing population, $b$ is a shape parameter determining how far $\delta$ will decline under complete selfing (when $C = 1$), and $L$ describes the expected deleterious mutation load as a function of the selfing rate $C$.
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Polymorphism and Extinction}\label{subsec:PolyExt}
+\subsection*{Fixation, Polymorphism, and Extinction}\label{subsec:PolyExt}
 
@@
 We begin with an illustration of demographically viable polymorphic parameter space in the absence of inbreeding depression in Figure \ref{fig:extThresholds} (i.e., $\delta = \delta_i = 0$).
-Invasion conditions for SA alleles in the evolutionary demographic model closely match the predictions from population genetic models \cite{Kidwell1977, JordanConnallon2014, Olito2017}.
+Invasion conditions for SA alleles in the evolutionary demographic model closely match the predictions from population genetic models \citep{Kidwell1977, JordanConnallon2014, Olito2017}.
 In particular, the demographic model recovers the classic "funnel-shaped" region of polymorphic $s_f \times s_m$ parameter space.
-The effects of the population selfing rate ($C$) and dominance ($h$) on SA polymorphism are also similar: self-fertilization ({\itshape i}) skews the SA polymorphic space towards regions with weaker selection through female-function and reduces the total area of polymorphic parameter space (e.g., contrast figure 1A with figure 1C); and ({\itshape ii}) dominance reversals (where deleterious SA fitness effects are partially recessive in each sex; $h = 1/4$) are much more permissive of SA polymorphism (e.g., contrast figures 1A-C with 1D-F) \citep{JordanConnallon2014, Olito2017}.
+The effects of the population selfing rate ($C$) and dominance ($h$) on SA polymorphism are also similar: self-fertilization ({\itshape i}) reduces the parameter space in which male-benefit alleles can invade, thereby increasing opportunity for spread of female-benefit alleles (e.g., contrast figure 1A with figure 1C); and ({\itshape ii}) dominance reversals (where deleterious SA fitness effects are partially recessive in each sex; $h = 1/4$) are much more permissive of SA polymorphism (e.g., contrast figures 1A-C with 1D-F) \citep{JordanConnallon2014, Olito2017}.
 
@@
 The location of the extinction threshold, where the population intrinsic growth rate $\lambda = 1$, is primarily determined by the fertility parameter ($f$) but is also influenced by the population selfing rate ($C$) and dominance of the SA alleles ($h$).
-When $f$ is large, the population can sustain higher mortality caused by stronger selection through both male and female fitness components before extinction occurs (the extinction threshold shifts to higher values of $s_f$ and $s_m$).
-
-The demographic effects of SA polymorphism are driven by the fitness effects of segregating male-beneficial/female-deleterious alleles because they directly influence individuals' fecundity.
+When $f$ is large, the population can experience stronger selection on fertility through both male and female fitness components before extinction occurs (the extinction threshold shifts to higher values of $s_f$ and $s_m$).
+
+The demographic consequences of SA selection are driven by the fitness effects of male-beneficial/female-deleterious alleles because they directly influence individuals' female fecundity, and therefore population growth rates.
 For the fertility parameter values we explored, all extinctions occured in regions where either the female-deleterious allele segregates at intermediate frequency (i.e., where SA polymorphism is maintained), or goes to fixation (fig.~\ref{fig:extThresholds}).
-The SA polymorphisms that remained viable at lower fertilitiy values corresponded to regions where the female-deleterious allele is predicted to segregate at low frequencies, or in regions with weak selection through both sex functions.
+The SA polymorphisms that remained viable at lower fertilitiy values corresponded to regions where the female-deleterious allele is predicted to segregate at low frequencies, or in regions with weak selection through both sex functions, an asymmetry that reflects the assumption of female demographic dominance.
 
@@
 In obligately outcrossing populations (including dioecious/gonochoristic populations; where $C = 0$), lower fertility can result in a significant reduction of demographically viable polymorphic parameter space.
-However, the effect is weaker in populations with intermediate selfing rates (compare $C=0$ vs.~$C > 0$) because self-fertilization protects a fraction, $C$, of the ovules of individuals homozygous for the female beneficial allele from exposure to the allele that is deleterious for females.
-In other words, self-fertilization leads to less selective death.
-This allows partially selfing populations to remain viable under selection intensities that would cause extinction in an outcrossing population.
-The combination of protection from selective death and reduced total polymorphic parameter space caused by selfing results in populations with intermediate selfing rates having the greatest proportion of demographically viable parameter space at medium and low fertilities (fig.~\ref{fig:polySpace}, med.~and low.~fertility values).
+However, the effect is weaker in populations with intermediate selfing rates (compare $C=0$ vs.~$C > 0$) because self-fertilization protects a fraction, $C$, of the ovules of individuals homozygous for the female-beneficial allele from exposure to the male-beneficial allele.
+In other words, self-fertilization reduces the opportunity for selection favouring the female-deleterious allele, thereby reducing the equilibrium load on female fecundity and allowing partially selfing populations to remain viable under selection intensities that would cause extinction in an outcrossing population.
+The combination of protection from reduced female fertility and reduced total polymorphic parameter space caused by selfing results in populations with intermediate selfing rates having the greatest proportion of demographically viable parameter space at medium and low fertilities (fig.~\ref{fig:polySpace}, med.~and low.~fertility values).
 
@@
 Unlike previous population genetic models, which assume constant population sizes \citep{JordanConnallon2014,Olito2017}, mortality caused by inbreeding depression can strongly influence population persistence in our evolutionary demographic model.
-Populations with high fertility rates can sustain positive population growth rates despite this higher mortality.
+Populations with high fertility rates can maintain positive population growth rates despite this additional mortality.
 This causes a greater proportion of SA polymorphic parameter space to be demographically viable, with the demographic model predictions converging on those from population genetic models in high-fertility populations (fig.~\ref{fig:deltaPolySpace}).
@@
 
-Mortality from inbreeding depression reduces the effect of protection from selective death that individuals gain through self fertilization.
-In contrast to our earlier results (see Figure \ref{fig:polySpace} in \hyperref[subsec:PolyExt]{Polymorphism and Extinction}), predominantly outcrossing populations are predicted to have the highest proportion of demographically viable polymorphic space when inbreeding depression is taken into acount (compare fig.~\ref{fig:polySpace} with fig.~\ref{fig:deltaPolySpace}).
+Mortality from inbreeding depression quickly outweighs the beneficial effect of selfing on the equilbirium female load that was apparent in the previous section.
+In contrast to our earlier results, predominantly outcrossing populations are predicted to have the highest proportion of demographically viable polymorphic space when inbreeding depression is taken into acount (compare fig.~\ref{fig:polySpace} with fig.~\ref{fig:deltaPolySpace}).
 Regardless of the life-history stage at which inbreeding depression affects survival, populations with intermediate to high selfing rates are unlikely to harbour SA polymorphism unless they can afford the resulting loss of self-fertilized ovules/offspring.
-An alternative view of these predictions is that populations with intermediate to high selfing rates and growth rates near one are vulnerable to extinction if a sexually antagonistic allele invades the population.
+An interesting alternative interpetation of these results is that populations with intermediate to high selfing rates and population intrinsic growth rates near one are vulnerable to extinction if a sexually antagonistic allele invades the population.
 
@@
 
-The point in the life-cycle where inbreeding depression manifests influences the threshold selfing rate at which demographically viable polymorphic parameter space crashes.
+The point in the life-cycle where inbreeding depression manifests can influence the threshold selfing rate at which demographically viable polymorphic parameter space crashes.
 Our results indicate that population viability was most sensitive to inbreeding depression affecting juvenile survival rates ($\delta_j$; fig.~\ref{fig:deltaPolySpace}, dark blue points), while early-acting inbreeding depression ($\delta$, ovule abortion shortly after fertilization) had a similar effect on population viability as late-acting inbreeding depression affecting adult survial ($\delta_a$) and juvenile-to-adult transition rates ($\delta_{\gamma}$).
@@
 \subsection*{Demographic and fitness data for {\itshape M. guttatus}} \label{subsec:MguttMethods}
-We used extensive demographic data reported in a large-scale study of local adaptation using experimental populations of {\itshape M.~guttatus} in Stanislaus National Forest (California, USA) in 2012 and 2013 \citep{PetersonEtAl2016}.
-We leverage their common-garden experimental design to focus on a comparison of demographically viable polymorphic parameter space for two experimental populations with contrasting demographic rates.
-The first was a locally adapted 'Eagle Meadows' population (data from 2012), while the second group was an experimental population composed of multiple non-locally adapted 'low-elevation perennials' (data from 2013).
+To parameterize our model we utilize extensive demographic data reported in a large-scale study of local adaptation using experimental populations of {\itshape M.~guttatus} in Stanislaus National Forest (California, USA) in 2012 and 2013 \citep{PetersonEtAl2016}.
+We leverage their common-garden experimental design to focus on a comparison of the consequences of SA selection for two experimental populations with contrasting demographic rates.
+The first was a locally adapted 'Eagle Meadows' population (data from 2012), while the second was an experimental population composed of multiple non-locally adapted 'low-elevation perennials' (data from 2013).
 The vital rate estimates for the Eagle Meadows population are as follows: seed bank survival ($D = 0.534$), seed germination rate ($G = 0.469$), flower production ($F = 0.64$), ovules per flower ($O = 614$), seedling recruits proportional to clonal rosette recruits ($A = 6.7 \times 10^{-4}$), overwinter survival ($S = 0.179$), rosette production ($R = 8.71$).
-The corresponding estimates for the low-elevation perrenials are: $D = 0.534$, $G = 0.652$, $F = 4.09$, $O = 494$, $A = 6.7e-4$), $S = 0$, and $R = 0$ (see corrected Tables.~1 and S2 in \citealt{PetersonEtAl2017}).
+The corresponding estimates for the low-elevation perrenials are: $G = 0.652$, $F = 4.09$, $O = 494$, $S = 0$, and $R = 0$ (see corrected Tables.~1 and S2 in \citealt{PetersonEtAl2017}).
 The same estimates for $D$ and $A$ were used for all populations.
@@
 Using the data provided in \citet{Willis1993}, we estimated the proportional decrease due to inbreeding depression in seed germination rate ($\delta_{G} = 0.085$), flower number ($\delta_{F} = 0.2$), and overwinter survival ($\delta_{S} = 0.38$).
-The largest field-estimated selfing rate for this same Iron Mountain population was $C = 0.29$ \citep{Willis1993}.
+The largest field-estimated selfing rate for the Iron Mountain population was $C = 0.29$ \citep{Willis1993}.
 
 Using these combined demographic rates, selfing rates, and inbreeding depression estimates, we constructed a corresponding stage $\times$ genotype mendelian matrix model with a single SA locus affecting female and male fertility (as described above).
-Using this empirically parameterized model, we are able to make predictions about the scope for demographically viable SA polymorphism at a single locus in hypothetical populations with the same demographic rates as observed in \citet{PetersonEtAl2016}, for a range of selfing and inbreeding depression rates observed in other natural populations.
-We stress, however, that these illustrative predictions are not explicit predictions of the likelihood of SA polymorphism in any specific population.
+With the empirically parameterized model, we are able to make predictions about the scope for demographically viable SA polymorphism at a single locus in hypothetical populations with the same demographic rates as observed in \citet{PetersonEtAl2016}, for a range of selfing and inbreeding depression rates observed in other natural populations.
+We stress, however, that these are illustrative rather than explicit predictions of the likelihood of SA polymorphism in any specific population, and ignore measurement error for the estimated demographic rates.
 
 Interestingly, a polymorphic chromosomal inversion (inv6) with apparently SA fitness effects has been identified in the Iron Mountain population of {\itshape M.~guttatus} \citep{LeeKelly2015}.
-inv6 segregates at moderate frequency (about $8\%$), and carriers suffer $\approx 30 \%$ loss in pollen viability, but also increased flower production which varies among years.
-The genetic basis of the fitness effects caused by inv6 are not yet known, but the net result is a "supergene" with strong fitness effects on both female and male fertility that segregates as a single diallelic locus.
-We estimated dominance and selection coefficients for the effect of inv6 on pollen viability and flower production from the data reported in \citet{LeeKelly2015}.
-Across four mapping populations, the effect of inv6 on pollen viability was partially recessive on average ($h_m \approx 0.35$).
-Given this average dominance, the selection coefficient in the field is approximately $s_m = 0.88$.
-It was not possible to estimate the average dominance of the effect of inv6 on flower production (inv6 homozygotes were too rare in the field), so we conservatively assume that it was additive ($h_f = 0.5$).
-Based on this assumption, the selection coefficient inv6 on flower production in the field was $s_f = 0.384$ in 2012, and $s_f = 0.232$ in 2013.
-As a final proof-of-concept test of our empirically parameterized model \citep[e.g.,][]{Servedio2014}, we asked whether, given the above biologically grounded fitness effect estimates, inv6 appears to fall in demographically viable SA polymorphic parameter space (see Figure \ref{fig:mimulusFig}), as might be expected given its observed frequency in the Iron Mountain population.
+inv6 segregates at moderate frequency (about $8\%$), and carriers suffer $\approx 30 \%$ loss in pollen viability, but also increased flower production.
+The genetic basis of these effects remain unclear, but the net result is a "supergene" with strong fitness effects on both female and male fertility that segregates as a single diallelic locus.
+We estimated selection coefficients for the effect of inv6 on pollen production (i.e., taking into account the simultaneous effect on flower number) and ovule production from the data reported in \citet{LeeKelly2015} under the relatively conservative assumptions that pollen/ovule production is proportional to flower number, and  additive SA fitness effects (dominance coefficients could not be estimated from field data; see Appendix \hl{C} for details of estimating selection coefficients).
+Under these assumptions, the average selection coefficient for inv6 on flower production across two years in the field was $(s_f,s_m) = (0.3,0.31)$.
+As a final proof-of-concept for our empirically parameterized model \citep[e.g.,][]{Servedio2014}, we asked whether, given the above biologically grounded fitness effect estimates, inv6 appears to fall in demographically viable SA polymorphic parameter space (see Figure \ref{fig:mimulusFig}), as might be expected given its observed frequency in the Iron Mountain population.
 
@@
 The Eagle Meadows population was locally adapted with a very high intrinsic growth rate ($\lambda \approx 1.7$).
-This growth rate was sufficiently high enough that all $s_f \times s_m$ selection parameter space (where $s_f, s_m \in (0,1]$) remained demographically viable, regardless of the selfing rate and effects of inbreeding depression ($C = 0$ or $0.29$; fig.~\ref{fig:mimulusFig}A).
+This growth rate was sufficiently high that all $s_f \times s_m$ selection parameter space (where $s_f, s_m \in (0,1]$) remained demographically viable, regardless of the selfing rate and effects of inbreeding depression ($C = 0$ or $0.29$; fig.~\ref{fig:mimulusFig}A).
 In contrast, the Low-Elevation Perennial population had a much lower, but still positive, intrinsic growth rate ($\lambda \approx 1.08$).
@@
 When using demographic rates for either the Eagle Meadows or Low-Elevation Perennial populations, inv6 falls within demographically viable polymorphic parameter space (fig.~\ref{fig:mimulusFig}; estimated selection coefficients for inv6 were $(s_f,\,s_m) = (0.30,\, 0.31)$).
-Whether outcrossing or partial selfing ($C=0.29$), inv6 always falls squarely in the middle of SA polymorphic space when using the locally adapted Eagle Meadows demographic rates.
-In contrast, when the selfing rate is at the higher end of empirical estimates for the Iron Mountain population in which inv6 has been documented ($C = 0.29$), inv6 falls nearer to the upper boundary of SA polymorhic space when using the Low-Elevation Perennial demographic rates.
-This happens because, when including empirical estimates of selfing rates and inbreeding depression, the polymorphic parameter space shifts downwards.
+Whether under obligate outcrossing or partial selfing, inv6 always falls squarely in the middle of SA polymorphic space when using the locally adapted Eagle Meadows demographic rates.
+In contrast, when the selfing rate is at the higher end of empirical estimates for the Iron Mountain population in which inv6 has been documented ($C = 0.29$), inv6 falls near the upper boundary but still within SA polymorhic space when using the Low-Elevation Perennial demographic rates.
+This happens because under the empirical estimates of selfing and inbreeding depression, the polymorphic parameter space shifts downwards.
 Additionally, inv6 falls much closer to the extinction threshold under partial selfing.
-Interesting, this suggests that even relatively small perturbations to demgraphic rates or selection coefficients could result in non-locally adapted populations being unable to support the demographic costs associated with segregating SA alleles with selection coefficients of similar magnitude to inv6.
+This suggests that even relatively small perturbations to demgraphic rates or selection coefficients could result in non-locally adapted populations being unable to support the demographic costs associated with segregating SA alleles with selection coefficients of similar magnitude to inv6.
 
@@
 
-The first and central finding of our study is that when intrinsic population growth rates are close to one, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space where SA polymorphism is maintained.
+The first and central finding of our study is that when intrinsic population growth rates approach one, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space where SA polymorphism is maintained.
 Since growth rates far exceeding one suggest rapid exponential growth, they are generally rare (with the notable exception of recently introduced, invasive populations), suggesting that our model predictions may be highly relevant for SA polymorphism in many real-world populations.
-Moreover, these findings compliment recent theoretical and empirical studies indicating that SA selection is likely to be both condition dependent, and stronger in locally-adapted populations near the center of a species' range \citep{BergerEtAl2014, Connallon2015}.
-
-We also find that demographically viable polymorphic parameter space is often biased towards alleles with stronger selection through the female- than male sex function; conditions where female-deleterious alleles are predicted to segregate at lower frequencies, lessening the demographic cost of polymorphism.
-Given natural variation in population growth rates, our model suggests that, on average, SA polymorphisms involving low-frequency female-deleterious SA alleles are more likely to be demographically viable, and therefore observable in real populations.
-Once again, this prediction is supported by experimental results indicating that male-beneficial SA genotypes are more susceptible to extinction under environmental stress or inbreeding \citep{BergerEtAl2014, berger2016intralocus, grieshop2017male}.
-Furthermore, our prediction that demographically viable SA variation should often favour female-beneficial genotypes is broadly consistent with the well established pattern of increased allocation to female reproduction in mixed-mating and selfing species (i.e., "selfing syndromes"; reviewed in \citealt{HarderBarrett2006, Sicard2011}).
+Moreover, these findings compliment recent theoretical and empirical studies indicating that SA selection is likely to be both condition dependent, and stronger in locally-adapted populations near the center of a species' range, where population growth rates are expected to be high \citep{BergerEtAl2014, Connallon2015}.
+
+We also find that demographically viable parameter space is often biased towards alleles with stronger selection through the female than male sex function.
+That is, given natural variation in population growth rates, the most demographically viable (and therefore observable in natural populations) outcomes of SA selection are either fixation of a female-beneficial allele, or polymorphisms involving low-frequency female-deleterious alleles.
+This key prediction is supported by a series of experimental results in seed beetles ({\itshape Callosobruchus maculatus}) which show that male-beneficial SA genotypes are less likely to contribute positively to population growth rates, and are more susceptible to extinction under environmental stress or inbreeding \citep{BergerEtAl2014, berger2016intralocus, grieshop2017male}.
+An interesting corrollary is that, since strong SA fitness effects, especially those affecting female fertility, often lead to extinction, the SA genetic variation we are able to observe in natural populations may often be under weak selection, and therefore strongly susceptible to genetic drift whether or not selection favours the fixation of one allele or balancing selection \citep{ConnallonClark2012}.
 
@@
 
-By combining the tools of demography and those of population genetics, the framework we presented here enables the exploration of interactions between life cycle complexity, mating system, and sexual antagonism.
-For the sake of simplicity, we used a basic life cycle with just two stages for most of this paper, adults and juveniles (the \textit{Mimulus} example has $3$ stages).
-However, it is possible to include additional age classes, allowing us to explore whether the scope for SA selection to maintain polymorphisms is affected by whether a species exhibits positive, or negative senescence \citep{jones2014diversity}?
-That is, does the shape of the survival curve affect the area of demographically viable polymorphic parameter space, and how does this interact with the age at which a sexually antagonistic allele is expressed?
-What if the allele not only affects male and female sex functions but also affects survival of individuals?
-And last but not least, how would the scope for SA selection to maintain polymorphisms be affected by density-dependence acting on different life history stages?
-
-One major simplifying assumption in our model was that how individuals themselves were produced (i.e., by selfing or outcrossing) affects the level of inbreeding depression they suffer.
+By combining the tools of demography and those of population genetics, the framework we present here enables the exploration of interactions between life cycle complexity, mating system, and sexual antagonism.
+For the sake of simplicity, we made a variety of simplifying assumptions in our analyses, and a variety of extensions are possible.
+For example, we used a basic life cycle with just two stages for most of this paper, adults and juveniles (the \textit{Mimulus} example has $3$ stages).
+However, it is possible to include additional age classes, allowing us to explore whether the scope for SA selection to maintain polymorphisms is affected by whether a species exhibits positive or negative senescence \citep{jones2014diversity}.
+Put another way, does the shape of the survival curve affect the demographic and population genetic consequences of SA selection, and how does this interact with the age at which a sexually antagonistic allele is expressed?
+What if the allele not only affects male and female fertility but also affects survival of individuals?
+Additionally, how would the scope for SA polymorphism be affected by density-dependence acting on different life history stages?
+
+Like most demographic matrix models, ours also assumes female demographic dominance, where population growth rates are determined entirely by female fecundity \citep{Caswell2001,CaswellEtAl2018,deVriesCaswell2019b}.
+Yet, many hermaphroditic populations experience limitation of reproductive success due to both quantity and quality of male gametes \citep[e.g.,][]{Yund2000,AizenHarder2007, Harder2016}.
+Explicitly modeling male gamete production is a natural extension to our modeling framework, and would enable us to analyze how tension between the demographic consequences of SA fitness variation through both female and male function alters or abolishes asymmetries in the extinction thresholds and polymorphic parameter space \citep[e.g.,][]{Tazzyman2015}.
+Interestingly, different forms of self-fertilization can aid in reproductive assurance, and should therefore have demographic consequences -- for example, under pollen limitation and delayed selfing (where only ovules that fail to receive outcross pollen are selfed), the selfing rate will be a function of genotype frequencies because these will directly influence pollen production \citep{HarderBarrett2006}.
+
+Another major simplifying assumption in our model was that how individuals themselves were produced (i.e., by selfing or outcrossing) affects the level of inbreeding depression they suffer.
 In reality, the history of consecutive generations of inbreeding in each individual's lineage will influence the severity of inbreeding depression they experience, particularly when inbreeding depression is caused primarily by recessive deleterious mutations.
-This more precise approach to modelling inbreeding could be included in our framework by expanding the individual state space to include selfing cohorts (i.e., first generation selfing, second generation selfing, etc.), as in the models of \citet{kelly1999response,kelly2007mutation}.
+It would be an interesting and feasible extension of our modeling framework to expand the individual state space to include selfing cohorts (i.e., first generation selfing, second generation selfing, etc.), as in the models of \citet{kelly1999response,kelly2007mutation}, thereby enabling a more biologically plausible approach to modelling self-fertilization.
 
@@
 For outcross reproduction the probability, and hence the probability of this $AA$ female producing an $AA$ offspring, is $q^{\prime}_{A}$, as derived above.
-Conversely, the probability of picking an $a$ allele and producing an $Aa$ offspringis zero for an $AA$ female when selfing, but $q^{\prime}_a$ for outcrossing.
+Conversely, the probability of picking an $a$ allele and producing an $Aa$ offspring is zero for an $AA$ female when selfing, but $q^{\prime}_a$ for outcrossing.
 Similarly, the middle column of block matrices are offspring produced by $Aa$ females, which can produce offspring of all $3$ genotypes.
@@
 
-We used demographic data for multiple experimental populations of {\itshape M.~guttatus} from a large-scale common garden experiment conducted in Stanislaus National Forest (California, USA) in 2012 and 2013 \cite{PetersonEtAl2016}.
+We used demographic data for multiple experimental populations of {\itshape M.~guttatus} from a large-scale common garden experiment conducted in Stanislaus National Forest (California, USA) in 2012 and 2013 \citep{PetersonEtAl2016}.
 Specifically, we used data from two experimental populations with contrasting demographic rates: the locally adapted 'Eagle Meadows' population (data from 2012), and a non-local 'low-elevation perennials' population (data from 2013).
@@
 To incorporate these terms into our evolutionary demographic model, we multiplied each of the three terms $G$, $F$, and $S$ in Eq(\ref{eq:AtildeMimulus}) by a corresponding proportional inbreeding depression term $(1 - \delta_i)$, where $i \{\in G,F,S\}$.
-The largest field-estimated selfing rate for this same Iron Mountain population was $C = 0.29$ \cite{Willis1993}, which we use in our calculations for fig.~\ref{fig:mimulusFig}.
+The largest field-estimated selfing rate for this same Iron Mountain population was $C = 0.29$ \citep{Willis1993}, which we use in our calculations for fig.~\ref{fig:mimulusFig}.
 

### 2021-05-07 doc/Supplements/SupplementaryMaterial.tex

@@
 \usepackage[authoryear,sectionbib,sort]{natbib}
+
+\bibliographystyle{amnatnat.bst}
+\setlength{\bibsep}{0.0pt}
 \linespread{1.7}
@@
 \usepackage{titlesec}
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{amssymb}
 \titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
@@
 \usepackage{fancyhdr}
+\usepackage{color,soul}
 \pagestyle{fancy}
@@
 \setlength{\headheight}{13.6pt}
+\renewcommand{\refname}{Literature Cited}
+
+\usepackage{tikz}
+
+% \newcommand\encircle[1]{%
+%   \tikz[baseline=(X.base)] 
+%     \node (X) [draw, shape=circle, inner sep=0] {\strut #1};}
+\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
+            \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
+
+% Equation numbering
+\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
+
+% Graphics package
+\usepackage{graphicx}
+\graphicspath{{../output/figs/}.pdf}
+
+% Change default margins
+\usepackage[top=0.75in, bottom=0.75in, left=0.75in, right=0.75in]{geometry}
+
+% Definitions
+\def\mathbi#1{\textbf{\em #1}}
+\def\mbf#1{\mathbf{#1}}
+\def\mbb#1{\mathbb{#1}}
+\def\mcal#1{\mathcal{#1}}
+\newcommand{\bo}[1]{{\bf #1}}
+\newcommand{\tr}{{\mbox{\tiny \sf T}}}
+\newcommand{\bm}[1]{\mbox{\boldmath $#1$}}
+\newcommand{\om}{\omega}
+\DeclareRobustCommand{\firstsecond}[2]{#2}
+ \def\linenumberfont{\normalfont\scriptsize}
+ \newcommand{\kron}{\otimes}
+\usepackage{pdflscape}
+
 %%%%%%%%%%%%%%%%%%%%%
@@
 \newpage{}
+\section{Construction of the population projection matrix}
+We will now derive the expressions for the survival and fertility matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ as given by equations \textcolor{red}{(10)-(13)} in the main text.
+The matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ are the ingredients of the population projection matrix, $\tilde{\mbf{A}}$, as given in equation \textcolor{red}{(9)}, repeated here for convenience,
+\begin{linenomath*}
+\begin{equation} \label{eq:Atilde_appendix}
+	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
+			\underbrace{\left(
+			\begin{array}{c|c}
+				\mcal{U}^S & \mbf{0} \\ \hline
+				\mbf{0} & \mcal{U}^X \\
+			\end{array} \right)}_{\tilde{\mbf{U}}} + 
+			\underbrace{\left(
+			\begin{array}{c|c}
+				\mcal{F}^S C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
+				\mcal{F}^X(\tilde{\mbf{p}}) (1 - C) & \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)\\
+			\end{array} \right)}_{\tilde{\mbf{F}}}.
+\end{equation}
+\end{linenomath*}
 
-\section*{Supplementary Figures and Tables}
+Our derivation follows closely the derivations in \cite{deVriesCaswell2019a} and \cite{deVriesCaswell2019b}, but extended to model mixed mating systems.
+To construct $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ using the vec-permutation approach \citep{CaswellEtAl2018}, we create a set of block-diagonal matrices $\mathbb{U}^S$, $\mathbb{U}^X$, $\mathbb{F}^X$, $\mathbb{F}^S$, and $\mathbb{H}^S$, $\mathbb{H}^X$, by putting the corresponding matrices on the diagonal.
+These block diagonal matrices can be written as
+\begin{linenomath*}
+\begin{eqnarray}
+ \mathbb{U}^S&=&\sum_{i=1}^g \mathbf{E}_{ii} \otimes \mathbf{U}^S_{i}, \label{eq:US_app}\\
+ \mathbb{F}^S&=&\sum_{i=1}^g \mathbf{E}_{ii} \otimes \mathbf{F}^S_{i}, \label{eq:FS_app}\\
+  \mathbb{H}^S &=& \bo I_\om \kron \bo H^S , \label{eq:HS_app} \\
+      \mathbb{D} &=& \bo I_\om \kron \bo I_g, \label{eq:D_app} 
+\end{eqnarray}
+\end{linenomath*}
+where $\mathbf{E}_{ii}$ is of dimensions $g \times g$. Analogous equations can be written down for $ \mathbb{U}^X$, $ \mathbb{F}^X$, and $ \mathbb{H}^X$.  
+% \begin{linenomath*}
+% \begin{eqnarray} 
+% 	\mbb{U}^S &= &
+% 		\left(
+% 			\begin{array}{ccc}
+% 				\mbf{U}^{S}_{AA} & \mbf{0} & \mbf{0} \\
+% 				\mbf{0} & \mbf{U}^{S}_{Aa} & \mbf{0} \\
+% 				\mbf{0} & \mbf{0} & \mbf{U}^{S}_{aa} \\
+% 			\end{array} \right), \label{eq:BlkUS_appendix}\\
+%   \mbb{F}^S &= &
+% 		\left(
+% 			\begin{array}{ccc}
+% 				\mbf{F}^{S}_{AA} & \mbf{0} & \mbf{0} \\
+% 				\mbf{0} & \mbf{F}^{S}_{Aa} & \mbf{0} \\
+% 				\mbf{0} & \mbf{0} & \mbf{F}^{S}_{aa} \\
+% 			\end{array} \right), \label{eq:BlkFS_appendix}\\
+% \end{eqnarray}
+% \end{linenomath*}
+% Similarly construct block diagonal matrices from the parent-offspring genotype maps,
+% \begin{linenomath*}
+% \begin{eqnarray}
+% 	\mathbb{H}^X(\tilde{\bo p}) = \bo I_\om \kron \bo H^X(\tilde{\bo p}),\\
+% 	\mathbb{H}^S(\tilde{\bo p}) = \bo I_\om \kron \bo H^S(\tilde{\bo p}).
+% \end{eqnarray}
+% \end{linenomath*}
 
-Lorem ipsum dolor sit amet (fig.~\ref{Fig:OkapiStripe}).
-Curabitur pretium tincidunt lacus.
-Nulla gravida orci a odio.
-Nullam varius, turpis et commodo pharetra, est eros bibendum elit, nec luctus magna felis sollicitudin mauris.
-Integer in mauris eu nibh euismod gravida.
-Duis ac tellus et risus vulputate vehicula.
-Donec lobortis risus a elit.
-Etiam tempor.
-Ut ullamcorper, ligula eu tempor congue, eros est euismod turpis, id tincidunt sapien risus a quam.
-Maecenas fermentum consequat mi.
-Donec fermentum.
-Pellentesque malesuada nulla a mi.
-Duis sapien sem, aliquet nec, commodo eget, consequat quis, neque.
+As in other multistate models, the survival and fertility matrices are then constructed from the following four block matrices, 
+\begin{linenomath*}
+\begin{eqnarray}
+\mcal{U}^S &=&\bo K^\tr \mathbb{D} \bo K \mathbb{U}^S, \label{eq:KDKU_app}\\
+\mcal{F}^S &=&\bo K^\tr \mbb{H}^S\bo K \mathbb{F}^S, \label{eq:KHKF_app}
+\end{eqnarray}
+\end{linenomath*}
+where $\mathbf{K}=\mathbf{K}_{\omega,g}$ is the vec-permutation matrix \citep{HendersonSearle1981}.
+The vec-permutation matrix rearrangers the population vectors $\mathbf{n}^S$ and $\mathbf{n}^X$ so that $\mathbf{n}^S$ is ordered by genotype first, and then by stage, but $\mathbf{K} \mathbf{n}^S$ is organized by stage first, and then by genotype (see \cite{hunter2005use} or \cite{caswell2012matrix} for more on the vec-permutation method).
 
-\begin{figure}[h!]
-%\includegraphics{okapi-stripes}
-\caption{Figure legend.}
-\label{Fig:OkapiStripe}
-\end{figure}
+The matrix $\mathbb{U}^S$ in equation (\ref{eq:KDKU_app}) generates transitions and survival of extant individuals within genotypes,  and acts on the the population vector of individuals produced through selfing, $\mathbf{n}^S$..
+The matrix $\bo K$ then permutes the resulting vector so that it is now organised by stages first, and then by genotypes within stages.
+The block diagonal matrix $ \mathbb{D} $ then accounts for changes in genotype within stage classes of extant individuals.
+Since extant individuals generally do not change their genotype $\mathbb{D}$ will be an identity matrix in our model.
+Finally, the matrix $\bo K^\tr $ permutes the population vector once again so that it returns to the original organisation by genotype first and then by stage classes.
 
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+Equation (\ref{eq:KHKF_app}) describes reproduction and genotype assignment of newborn individuals.
+The block matrix $\mathbb{F}^S$ contains fecundity rates for individuals of each genotype and stage combination and acts on the the population vector of individuals produced through selfing, $\mathbf{n}^S$.
+Initially, these offspring are associated with the genotype of their parent.
+The vec-permutation matrix $\bo K$ then rearranges the resulting vector so that it is organised by stage first and then by genotype.
+The block-diagonal matrix $\mbb{H}^S$ then allocates the offspring to their genotypes, according to equation (\ref{eq:HS}).
+Finally, $\bo K^\tr $  returns the vector to its initial orientation.
 
-Tables can exist too (see table~\ref{Table:Rivers} for an example).
+Substituting equations (\ref{eq:US_app}) and (\ref{eq:D_app}) into equation (\ref{eq:KDKU_app}) yields
+\begin{linenomath*}
+\begin{eqnarray}
+\mcal{U}^S &=& \sum_{i=1}^g \bo K^\tr \bo I_{\omega g}	 \bo K \mathbf{E}_{ii} \otimes \mathbf{U}^S_{i}, \\
+ &=& \sum_{i=1}^g \mathbf{E}_{ii} \otimes \mathbf{U}^S_{i},
+\end{eqnarray}
+\end{linenomath*}
+since $\bo K^\tr \bo I_{\omega g}	 \bo K=\bo I_{\omega g}$, or in block matrix form this yields equation (\ref{eq:BlkUS}), here repeated for convenience, 
+\begin{linenomath*}
+\begin{eqnarray}
+\mcal{U}^S  = 
+		\left(
+			\begin{array}{ccc}
+				\mbf{U}^{S}_{AA} & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{U}^{S}_{Aa} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \mbf{U}^{S}_{aa} \\
+			\end{array} \right).\label{eq:BlkUS}\\
+\end{eqnarray}
+\end{linenomath*}
 
-\begin{table}[htbp]
-\caption{Various rivers, cities, and animals}
-\label{Table:Rivers}
-\centering
-\begin{tabular}{lll}\hline
-River        & City        & Animal            \\ \hline
-Chicago      & Chicago     & Raccoon           \\
-Des Plaines  & Joliet      & Coyote            \\
-Illinois     & Peoria      & Cardinal          \\
-Kankakee     & Bourbonnais & White-tailed deer \\
-Mississippi  & Galena      & Bald eagle        \\ \hline
-\end{tabular}
-\bigskip{}
-\\
-{\footnotesize Note: Notes can go here.}
-\end{table}
+Similarly, substituting equations (\ref{eq:FS_app}) and (\ref{eq:HS_app}) into equation (\ref{eq:KHKF_app}) yields,
+\begin{linenomath*}
+\begin{eqnarray}
+\mcal{F}^S &=&\bo K^\tr \mbb{H}^S\bo K \mathbb{F}^S, \label{eq:KHKF_app}\\
+ &=& \sum_{i=1}^g \bo K^\tr \left( \bo I_\om \kron \bo H^S \right)  \bo K \left( \mathbf{E}_{ii} \otimes \mathbf{F}^S_{i} \right).
+\end{eqnarray}
+\end{linenomath*}
+Using $\bo K^\tr \left(\bo A \otimes \bo B  \right) \; \bo K=\bo B \otimes \bo A$ \citep{magnus1979commutation}, we obtain
+\begin{linenomath*}
+\begin{eqnarray}
+\mcal{F}^S  &=& \sum_{i=1}^g  \left(  \bo H^S  \kron \bo I_\om\right) \left( \mathbf{E}_{ii} \otimes \mathbf{F}^S_{i} \right).
+\end{eqnarray}
+\end{linenomath*}
+Finally note that $(\mathbf{A} \otimes \mathbf{B}) (\mathbf{C} \otimes \mathbf{D}) = \mathbf{A} \mathbf{C}\otimes \mathbf{B} \mathbf{D}$  to write
+\begin{linenomath*}
+\begin{eqnarray}
+\mcal{F}^S  &=& \sum_{i=1}^g  \left(  \bo H^S \mathbf{E}_{ii} \right) \kron  \mathbf{F}^S_{i} .
+\end{eqnarray}
+\end{linenomath*}
+Using the exact same sequence of equations, we can derive
+\begin{linenomath*}
+\begin{eqnarray}
+\mcal{F}^X  &=& \sum_{i=1}^g  \left(  \bo H^X(\tilde{\mbf{p}}) \mathbf{E}_{ii} \right) \kron  \mathbf{F}^X_{i} .
+\end{eqnarray}
+\end{linenomath*}
 
-\section*{Additional Methods}
+Finally, substitute the expressions derived for the parent-offpspring genotype maps, $\bo H^X(\tilde{\mbf{p}}) $, and $\bo H^S$, in the main text (\textcolor{red}{equations (12) and (13))} into the above equations to obtain
+\begin{linenomath*}
+\begin{eqnarray}
+	\mcal{F}^S = \mbf{K}^{\intercal} \mbb{H}^S\mbf{K} \mbb{F}^S = 
+			\left(
+			\begin{array}{ccc}
+				\mbf{F}_{AA} & \frac{1}{4} \mbf{F}_{Aa} & 0 \\
+				0 & \frac{1}{2} \mbf{F}_{Aa} & 0 \\
+				0 & \frac{1}{4} \mbf{F}_{Aa} & \mbf{F}_{aa}\\
+			\end{array} \right),  \label{eq:BlkFS_appendix}
+\end{eqnarray}
+\end{linenomath*}
+\noindent and 
+\begin{linenomath*}
+\begin{equation} \label{eq:BlkFX_appendix}
+	\mcal{F}^X = \mbf{K}^{\intercal} \mbb{H}^X(\tilde{\mbf{p}}) \mbf{K} \mbb{F}^X = 
+			\left(
+			\begin{array}{ccc}
+				q^{\prime}_{A} \mbf{F}_{AA} & \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} & 0 \\ 
+				q^{\prime}_{a} \mbf{F}_{AA} & \frac{1}{2} \mbf{F}_{Aa} & q^{\prime}_{A} \mbf{F}_{aa}  \\ 
+				0 & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} & q^{\prime}_{a} \mbf{F}_{aa}  \\
+			\end{array} \right),
+\end{equation}
+\end{linenomath*}
+\newpage
 
-\subsection*{Measuring the height of fox jumps}
+\section{Boundary stability analysis}
 
-The quick red fox jumps over the lazy brown dog.
-Morbi odio velit, eleifend ut hendrerit vitae, consequat sit amet nulla.
-Pellentesque porttitor vitae nisl quis tempus.
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Praesent ut nisi odio.
+In this section we follow closely the derivations presented in \cite{deVriesCaswell2019a} but we extended the model to include selfing which complicates the boundary stability analysis.
+
+We will derive the conditions for a protected polymorphism.
+The derivation is simplest when the population vector is ordered by genotype first, and then by whether individuals were produced through selfing or outcrossing, and finally by stage, in contrast to the ordering used in the main text (produced by selfing vs outcrossing, then genotype, then stage).
+
+The population vector is then given by 
+\begin{equation}
+	\tilde{\mbf{n}} = \left(
+			\begin{array}{c}
+							\mbf{n}^S_{AA} \\
+							\mbf{n}^X_{AA} \\ \hline
+							\mbf{n}^S_{Aa} \\
+							\mbf{n}^X_{Aa} \\ \hline
+							\mbf{n}^S_{aa} \\
+							\mbf{n}^X_{aa} \\
+			\end{array} \right).
+\end{equation}
+The population projection matrix $\tilde{\mbf{A}}$ consists of $3 \times 3$ blocks, which act on the genotype specific population vectors:
+{
+\small
+\begin{align*} 
+	\tilde{\mbf{A}}[\tilde{\mbf{n}}] &= \tilde{\mbf{U}} + \tilde{\mbf{F}}\\
+		&= \left(
+			\begin{array}{cc|cc|cc}
+				\mathbf{U}^S_{AA} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
+				\mbf{0} & \mathbf{U}^X_{AA} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\\hline
+				\mbf{0} & \mbf{0} & \mathbf{U}^S_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \mbf{0} & \mathbf{U}^X_{Aa} & \mbf{0} &  \mbf{0}\\\hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mathbf{U}^S_{aa} & \mbf{0} \\ 
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mathbf{U}^X_{aa}
+			\end{array} \right) \\ 
+		&+ \left(
+			\begin{array}{cc|cc|cc}
+				C_{AA} (1 - \delta) \mbf{F}_{AA} & C_{AA} (1 - \delta) \mbf{F}_{AA}  & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} \\ 
+				q^{\prime}_{A}  (1 - C_{AA}) \mbf{F}_{AA}& q^{\prime}_{A} (1 - C_{AA}) \mbf{F}_{AA} & \frac{1}{2} q^{\prime}_{A} (1 - C_{Aa})\mbf{F}_{Aa}  & \frac{1}{2} q^{\prime}_{A} (1 - C_{Aa})\mbf{F}_{Aa}  & \mbf{0} & \mbf{0} \\ \hline
+				\mbf{0} & \mbf{0} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}& \mbf{0} & \mbf{0} \\
+				q^{\prime}_{a} (1 - C_{AA}) \mbf{F}_{AA} & q^{\prime}_{a}(1 - C_{AA})  \mbf{F}_{AA} & \frac{1}{2}  (1 - C_{Aa}) \mbf{F}_{Aa}& \frac{1}{2} (1 - C_{Aa})\mbf{F}_{Aa} & q^{\prime}_{A} (1 - C_{aa})\mbf{F}_{aa}  & q^{\prime}_{A}  (1 - C_{aa})\mbf{F}_{aa}\\ \hline
+				\mbf{0} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & C_{aa} (1 - \delta) \mbf{F}_{aa} &  C_{aa} (1 - \delta) \mbf{F}_{aa}\\ 
+				\mbf{0} & \mbf{0} & \frac{1}{2} q^{\prime}_{a}  (1 - C_{Aa}) \mbf{F}_{Aa} & \frac{1}{2} q^{\prime}_{a}  (1 - C_{Aa})\mbf{F}_{Aa}& q^{\prime}_{a}(1 - C_{aa})  \mbf{F}_{aa} & q^{\prime}_{a} (1 - C_{aa})\mbf{F}_{aa}
+			\end{array} \right) \numberthis \label{eq:AtildeCoexistence}
+\end{align*}
+}
+
+The Mendelian matrix model defined in equations \textcolor{red}{(9)-(13)}, or (\ref{eq:AtildeCoexistence}), reduces to a linear matrix model on the boundary where the population is initially fixed for the $A$ allele (since $q^{\prime}_{A} = 1$ and $q^{\prime}_{a} = 0$).
+Provided the initial population contains a nonzero number of females, the population will grow or shrink exponentially after converging to a stable population structure (see Caswell (2001), section 4.5.2.1).
+We can take advantage of the homogeneity of $\tilde{\mbf{F}}$ by rewriting the model in terms of the normalized population vector (the frequency vector):
+\begin{equation} 
+	\tilde{\mbf{p}}(t + 1) = \frac{ \tilde{\mbf{A}}[\tilde{\mbf{p}}(t)] \tilde{\mbf{p}}(t) }{ \| \tilde{\mbf{A}}[\tilde{\mbf{p}}(t)] \tilde{\mbf{p}}(t) \|},\label{eq:freq_model}
+\end{equation}
+
+\noindent where $\| \mbf{a} \|$ denotes the $1$-norm of the vector $\mbf{a}$, defined as the sum of the absolute values of the entries of the vector $\mbf{a}$.
+Equilibrium solutions, denoted by $\hat{\mbf{p}}$, satisfy
 
 \begin{equation} 
-{ \frac{1}{N_k-1} \sum \limits_{t=1}^{N_k} (M_{tjk} - \bar{M}_{jk})^2}
+	\hat{\mbf{p}} = \frac{ \tilde{\mbf{A}}[\hat{\mbf{p}}] \hat{\mbf{p}} }{ \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}}[\hat{\mbf{p}}] \hat{\mbf{p}} },\label{eq:freq_equilibrium}
 \end{equation}
 
-\subsection*{Quantifying the brownness of the dog}
+\noindent where the one norm can be replaced by $\mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}}[\hat{\mbf{p}}] \hat{\mbf{p}}$ because $\hat{\mbf{p}}$ is nonnegative.
 
-Pellentesque eu nulla odio (\citealt{Xiao2015,CookEtAl2015}).
-Nulla aliquam porta metus, quis malesuada orci faucibus quis.
-Suspendisse nunc magna, tristique sit amet sollicitudin nec, elementum et lacus.
-Sed vitae elementum mi.
-In hac habitasse platea dictumst.
-Etiam eu tortor elit.
-Sed ac tortor purus.
-Aliquam volutpat, odio sit amet posuere pretium, dolor ex interdum ante, sed luctus quam eros ac nulla.
+\subsection{Linearization at the boundary equilibrium}
+
+The stability of the boundary equilibrium to invasions by the other allele is determined by the magnitude of the largest eigenvalue of the Jacobian matrix of the frequency model (equation (\ref{eq:freq_model})) evaluated at the equilibrium defined by equation (\ref{eq:freq_equilibrium}).
+If the magnitude of this eigenvalue is larger than one, then the equilibrium is unstable.
+The Jacobian matrix,
+
+\begin{equation} \label{eq:genJacobian}
+	\mbf{M} =  \frac{ \text{d} \tilde{\mbf{p}}(t + 1) }{ \text{d} \tilde{\mbf{p}}(t) } \bigg\rvert_{\hat{\mbf{p}}},
+\end{equation}
+
+\noindent is obtained by differentiating equation (\ref{eq:genJacobian}) and evaluating the resulting derivatives at the boundary equilibrium.
+This requires a long series of matrix calculus operations, and repeatedly takes advantage of the fact that $\hat{\mbf{p}}$ at the boundary contains zeros for the blocks corresponding to $Aa$ and $aa$ genotypes.
+\newpage
+For notational convenience, first define a matrix $\mbf{B}$ as
 
 \begin{equation}
-{ (\sum \limits_{p=1}^P {n_{sp}})^{-1}\sum \limits_{p=1}^P {n_{sp}Q_{p}}}
+	\mbf{B}[\tilde{\mbf{p}}] = \frac{ \tilde{\mbf{A}}[\tilde{\mbf{p}}] }{ \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}}[\tilde{\mbf{p}}] \tilde{\mbf{p}}}, 
 \end{equation}
 
-\newpage{}
+\noindent such that equation (\ref{eq:freq_model}) can be rewritten as
+
+\begin{equation} 
+	\tilde{\mbf{p}}(t + 1) = \mbf{B}[\tilde{\mbf{p}}(t)]\tilde{\mbf{p}}(t).
+\end{equation}
+
+\noindent We differentiate to obtain
+
+\begin{equation} 
+	\text{d} \tilde{\mbf{p}}(t + 1) = \mbf{B}\text{d} \tilde{\mbf{p}}(t) + \left(\text{d} \tilde{\mbf{B}} \right) \tilde{\mbf{p}}(t),
+\end{equation}
+
+\noindent where the explicit dependence of $\mbf{B}$ on $\tilde{\mbf{p}}$ has been omitted to avoid a cluttering of brackets.
+Multiply the second term by a $2 \omega g \times 2 \omega g$ identity matrix,
+
+\begin{equation} 
+	\text{d} \tilde{\mbf{p}}(t + 1) = \mbf{B}\text{d} \tilde{\mbf{p}}(t) + \mbf{1}_{2 \omega g} \left(\text{d} \tilde{\mbf{B}} \right) \tilde{\mbf{p}}(t).
+\end{equation}
+
+\noindent and apply the vector operator to both sides, remembering that as $\tilde{\mbf{p}}$ is a vector, $\text{vec}\tilde{\mbf{p}} = \tilde{\mbf{p}}$,
+
+\begin{equation} 
+	\text{d} \tilde{\mbf{p}}(t + 1) = \mbf{B}\text{d} \tilde{\mbf{p}}(t) + \text{vec} \left[ \mbf{1}_{2 \omega g} \left(\text{d} \tilde{\mbf{B}} \right) \tilde{\mbf{p}}(t) \right].
+\end{equation}
+
+\noindent Next, we can apply Roth's theorem (Roth 1934), $\text{vec}\mbf{ABC} = \left( \mbf{C}^{\intercal} \otimes \mbf{A} \right) \text{vec}\mbf{B}$, to replace the the $\text{vec}$ operator with the Kronecer product:
+
+\begin{equation} 
+	\text{d} \tilde{\mbf{p}}(t + 1) = \mbf{B}\text{d} \tilde{\mbf{p}}(t) + \left( \tilde{\mbf{p}}^{\intercal}(t) \otimes \mbf{1}_{2 \omega g} \right) \text{d}\mbox{vec} \mbf{B}.
+\end{equation}
+
+\noindent Then the first identification theorem and the chain rule together give the following formula for the Jacobian (Magnus and Neudecker, 1985; Caswell, 2007),
+
+\begin{align*} \label{eq:genJacobian}
+	\mbf{M} &=  \frac{ \text{d} \tilde{\mbf{p}}(t + 1) }{ \text{d} \tilde{\mbf{p}}(t) } \bigg\rvert_{\hat{\mbf{p}}}, \\
+			&= \mbf{B}[\tilde{\mbf{p}}] + \left( \tilde{\mbf{p}}^{\intercal} \otimes \mbf{1}_{2 \omega g} \right) \frac{ \partial \text{vec}{\mbf{B}[\tilde{\mbf{p}}]} }{ \partial \tilde{\mbf{p}}^{\intercal} } \bigg\rvert_{\hat{\mbf{p}}},
+\end{align*}
+\newpage
+Our aim is to express the Jacobian matrix $\mbf{M}$ in terms of the genotype specific matrices, $\mbf{U}^S_i$, $\mbf{U}^X_i$, $\mbf{F}^S_i$, $\mbf{F}^X_i$.
+We analyze the Jacobian at the $AA$ boundary; the expression at the $aa$ boundary can be derived afterwards using symmetry arguments.
+First it will be convenient to define the scalar $f(\tilde{\mbf{p}})$ as
+
+\begin{equation}
+	f(\tilde{\mbf{p}}) = \frac{1}{ \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}}[\tilde{\mbf{p}}] \tilde{\mbf{p}}},
+\end{equation}
+
+\noindent so that
+
+\begin{equation}
+	\mbf{B}[\tilde{\mbf{p}}] = f(\tilde{\mbf{p}}) \tilde{\mbf{A}}[\tilde{\mbf{p}}]. \label{eq:BfandA}
+\end{equation}
+
+\noindent Note that where it does not create confusion, we omit the explicit dependence on $\tilde{\mbf{A}}$, $\mbf{B}$, and $f$ on $\tilde{\mbf{p}}$.
+Next, we take the vec of both sides of equation (\ref{eq:BfandA}) and differentiate to obtain
+
+\begin{equation} \label{eq:dvecB}
+	\text{dvec} \mbf{B} = \text{vec} \tilde{\mbf{A}} \text{d} f + f \text{dvec} \tilde{\mbf{A}},
+\end{equation}
+
+\noindent or
+
+\begin{equation}
+	\frac{\partial \text{vec} \mbf{B}} {\partial \tilde{\mbf{p}}^{\intercal}} = \text{vec} \tilde{\mbf{A}} \frac{\partial f}{\partial \tilde{\mbf{p}}^{\intercal} } + f(\tilde{\mbf{p}}) \frac{ \partial \text{vec} \tilde{\mbf{A}} }{ \partial \tilde{\mbf{p}}^{\intercal} }.
+\end{equation}
+
+\noindent Next differentiate $f$ to obtain
+
+\begin{equation}
+	\text{d} f = \frac{ -1 }{ \left( \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}}[\tilde{\mbf{p}}] \tilde{\mbf{p}} \right)^2 } \left[ \mbf{1}^{\intercal}_{2 \omega g} \left( \text{d} \tilde{\mbf{A}} \right) \tilde{\mbf{p}} + \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}} \text{d} \tilde{\mbf{p}} \right].
+\end{equation}
+
+\noindent All the terms in the Jacobian are evaluated at the $AA$ boundary, which simplifies some of the equations, e.g., $\tilde{\mbf{A}}[ \hat{\mbf{p}} ] \hat{\mbf{p}} = \lambda_{AA} \hat{\mbf{p}}$, and therefore
+
+\begin{equation} \label{eq:lambdaAA}
+	\mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}}[\hat{\mbf{p}}] \hat{\mbf{p}} = \lambda_{AA}.
+\end{equation}
+
+Evaluate the differential of $f$ at the boundary and use equation (\ref{eq:lambdaAA}) to obtain
+
+\begin{equation} \label{eq:df}
+	\text{d} f(\tilde{\mbf{p}}) = \frac{ -1 }{\lambda_{AA}^2 } \left[ \mbf{1}^{\intercal}_{2 \omega g} \left( \text{d} \tilde{\mbf{A}} \right) \hat{\mbf{p}} + \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}} \text{d} \tilde{\mbf{p}} \right] \bigg\rvert_{\hat{\mbf{p}}}.
+\end{equation}
+
+\newpage
+
+The first term in this sum, $\mbf{1}^{\intercal}_{2 \omega g} \left( \text{d} \tilde{\mbf{A}} \right) \hat{\mbf{p}}$, is equal to zero because
+\begin{align*}
+	q^{\prime}_A + q^{\prime}_a &= 1 \\
+	\text{d}q^{\prime}_A + \text{d}q^{\prime}_a &= 0, 
+\end{align*}
+and therefore every column in $\left( \text{d} \tilde{\mbf{A}} \right)$ sums to zero, see equation (\ref{eq:AtildeCoexistence}).
+
+Substituting equation (\ref{eq:df}) into equation (\ref{eq:dvecB}) and evaluating at the boundary yields
+\begin{equation} \label{eq:dvecBsubs}
+	\text{dvec} \mbf{B} = \frac{ -1 }{\lambda_{AA}^2 } \text{vec} \tilde{\mbf{A}} \left[ \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}} \text{d} \tilde{\mbf{p}} \right] + \frac{ 1 }{\lambda_{AA} } \text{dvec} \tilde{\mbf{A}},
+\end{equation}
+\noindent or 
+\begin{equation}
+	\frac{\partial \text{vec} \mbf{B}} {\partial \tilde{\mbf{p}}^{\intercal}} \bigg\rvert_{\hat{\mbf{p}}} = \frac{ -1 }{\lambda_{AA}^2 } \left( \text{vec} \tilde{\mbf{A}} \right) \left( \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}} \text{d} \tilde{\mbf{p}} \right) + \frac{ 1 }{\lambda_{AA} } \left.
+\frac{\partial \text{vec} \mbf{A}} {\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}
+\end{equation}
+
+\noindent Finally, substituting the expression above into equation (\ref{eq:genJacobian}) yields the Jacobian matrix:
+
+\begin{align*} \label{eq:subJacobian}
+	\mbf{M} &=  \mbf{B}[\tilde{\mbf{p}}] + \left( \tilde{\mbf{p}}^{\intercal} \otimes \mbf{1}_{2 \omega g} \right) \frac{ \partial \text{vec}{\mbf{B}} }{ \partial \tilde{\mbf{p}}^{\intercal} } \bigg\rvert_{\hat{\mbf{p}}}, \\
+			&= \underbrace{ \mbf{B}[\hat{\mbf{p}}]}_{\circled{A}} - 
+			   \underbrace{ \frac{ 1 }{\lambda_{AA}^2 } (\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{\omega g}) \left( \text{vec} \tilde{\mbf{A}} \right) \left( \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}} \right)}_{\circled{B}} \\
+			&~~~~~~~~~~~+ \underbrace{\frac{ 1 }{\lambda_{AA} } (\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{\omega g}) \left.
+\frac{\partial \text{vec} \tilde{\mbf{A}}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}}_{\circled{C}},
+\end{align*}
+
+\newpage
+%%%%%%%%%%%%%%%%
+\subsection{Components of the Jacobian}
+
+The next task is to work out each of the three terms \circled{A}, \circled{B}, and \circled{C}, in $\mbf{M}$. Beginning with \circled{A}, we have
+
+\begin{align*}
+	\circled{A}&=\mbf{B}[\hat{\mbf{p}}] = \frac{ \tilde{\mbf{A}}[\hat{\mbf{p}}] }{ \mbf{1}^{\intercal}_{\omega g} \tilde{\mbf{A}}[\hat{\mbf{p}}] \hat{\mbf{p}} } = \\
+						   \frac{1}{\lambda_{AA}} &
+						   \left(
+			\begin{array}{cc|cc|cc}
+				\mathbf{U}^S_{AA} & \mbf{0}& \mbf{0} & \mbf{0} & \mbf{0} \\ 
+			\mbf{0} & \mathbf{U}^X_{AA} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
+				\mbf{0} & \mbf{0} & \mathbf{U}^S_{Aa}& \mbf{0} & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \mbf{0}& \mcal{U}^X_{Aa} & \mbf{0} & \mbf{0}\\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mathbf{U}^S_{aa}  & \mbf{0} \\ 
+				\mbf{0} & \mbf{0} & \mbf{0}  & \mbf{0} & \mbf{0} & \mathbf{U}^X_{aa}
+			\end{array} \right) \\
+			 +\frac{1}{\lambda_{AA}} &
+						   \left(
+			\begin{array}{cc|cc|cc}
+		 C_{AA} (1 - \delta) \mbf{F}_{AA} & C_{AA} (1 - \delta) \mbf{F}_{AA}& \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}& \mbf{0} & \mbf{0} \\ 
+				(1 - C_{AA}) \mbf{F}_{AA} & (1 - C_{AA}) \mbf{F}_{AA}  & \frac{1}{2}(1 - C_{Aa}) \mbf{F}_{Aa}  & \frac{1}{2} (1 - C_{Aa})\mbf{F}_{Aa} & \mbf{0} & \mbf{0} \\ \hline
+				\mbf{0} & \mbf{0} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} & \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} & (1 - C_{aa}) \mbf{F}_{aa} & (1 - C_{aa}) \mbf{F}_{aa}\\ \hline
+				\mbf{0} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &C_{aa} (1 - \delta) \mbf{F}_{aa} &C_{aa} (1 - \delta) \mbf{F}_{aa} \\ 
+				\mbf{0} & \mbf{0} & \mbf{0}  & \mbf{0} & \mbf{0} &\mbf{0}
+			\end{array} \right) 
+			\numberthis
+\end{align*}
+Next, we have \circled{B},
+\begin{equation}
+	\circled{B} = - \frac{1}{\lambda_{AA}^2} \left( \hat{\mbf{p}}^{\intercal} \otimes \mbf{1}_{2 \omega g} \right) \left( \text{vec} \tilde{\mbf{A}} \right) \left( \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}} \right).
+\end{equation}
+Again using Roth's theorem (Roth 1934), $\left( \mbf{C}^{\intercal} \otimes \mbf{A} \right) \text{vec}\mbf{B} = \text{vec}\mbf{ABC}$, we can simplify as follows:
+\begin{align*}
+	\left( \hat{\mbf{p}}^{\intercal} \otimes \mbf{1}_{2 \omega g} \right) \left( \text{vec} \tilde{\mbf{A}}[\hat{\mbf{p}}] \right) &= \text{vec} \left( \mbf{I}_{2 \omega g}\, \tilde{\mbf{A}}[\hat{\mbf{p}}]\, \hat{\mbf{p}} \right) \\
+					   &= \lambda_{AA} \hat{\mbf{p}}, \numberthis
+\end{align*}
+so that 
+\begin{equation}
+	\circled{B} = - \frac{1}{\lambda_{AA}} \hat{\mbf{p}} \left( \mbf{1}^{\intercal}_{2 \omega g} \tilde{\mbf{A}}[\hat{\mbf{p}}] \right).
+\end{equation}
+
+\newpage
+Substitute the population vector at the $AA$ boundary,
+\begin{equation}
+	\hat{\mbf{p}} = \left(
+			\begin{array}{c}
+							\hat{\mbf{p}}^S_{AA} \\
+							\hat{\mbf{p}}^X_{AA} \\ \hline
+							0                  \\
+							0                  \\ \hline
+							0                  \\
+							0                  \\
+			\end{array} \right)
+\end{equation}
+and rewrite in terms of the block matrices to obtain the big expression on the next page 
+\begin{landscape}
+{\footnotesize
+\begin{align*}
+	\circled{B} =& \\
+	&-\frac{1}{\lambda_{AA}} \left(
+			\begin{array}{cc|cc|cc}
+				\hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{aa} \\
+				\hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{aa} \\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} 
+			\end{array} \right) \\
+	&-\frac{1}{\lambda_{AA}} \left(
+			\begin{array}{cc|cc|cc}
+				(1 - \delta C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & (1 - \delta C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & (1 - \delta C_{Aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & (1 - \delta C_{Aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa}  & (1 - \delta C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} & (1 - \delta C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} \\
+				 	(1 - \delta C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & 	(1 - \delta C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & (1 - \delta C_{Aa})  \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & (1 - \delta C_{Aa})  \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & (1 - \delta C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} & (1 - \delta C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa}\\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0}
+			\end{array} \right) \\
+			 \numberthis			
+\end{align*}
+}
+
+\end{landscape}
+
+Finally, we turn to term \circled{C} of the Jacobian. 
+To derive term $\circled{C}$ in the Jacobian, we first derive a useful expression for $\text{vec} \tilde{\bo A}$ in terms of its component block matrices.
+The matrix $\tilde{\mathbf{A}}$ can be decomposed into 36 $\omega \times \omega$ block matrices, as in equation (\ref{eq:AtildeCoexistence}), so that for example
+\begin{eqnarray}
+\mathbf{A}_{11}&=&  \mathbf{U}^S_{AA} + C_{AA} (1 - \delta) \mbf{F}_{AA} ,\\
+\mathbf{A}_{22}&=&\mathbf{U}^X_{AA} +q^{\prime}_{A} (1 - C_{AA}) \mbf{F}_{AA},
+\end{eqnarray}
+and
+\begin{eqnarray}
+  \mathbf{A}_{21}&=&q^{\prime}_{a} (1 - C_{AA}) \mbf{F}_{AA},\\
+  \mathbf{A}_{12}&=& C_{AA} (1 - \delta) \mbf{F}_{AA} .
+\end{eqnarray}
+The matrix $\tilde{\mathbf{A}}$ can then be written as 
+\begin{eqnarray}
+\tilde{\mathbf{A}}&=&\sum_{i,j=1}^6 \mathbf{E}_{ij} \otimes \mathbf{A}_{ij}, \\
+&=&\sum_{i,j=1}^6 \left(\mathbf{e}_i \mathbf{e}^\tr_j\right) \otimes \left(\mathbf{A}_{ij} \mathbf{I}_\omega \right),
+\label{eq:A_sum_appendix}
+\end{eqnarray}
+where we have used the definition of the matrix $\mathbf{E}_{ij}=\mathbf{e}_i \mathbf{e}^\tr_j$.
+Using $\mathbf{A} \mathbf{C}\otimes \mathbf{B} \mathbf{D}=(\mathbf{A} \otimes \mathbf{B}) (\mathbf{C} \otimes \mathbf{D})$, equation (\ref{eq:A_sum_appendix}) can be rewritten as
+\begin{equation}
+  \tilde{\mathbf{A}}=\sum_{i,j=1}^6 \left(\mathbf{e}_i \otimes \mathbf{A}_{ij} \right) \left( \mathbf{e}^\tr_j \otimes \mathbf{I}_\omega \right) ,
+\end{equation}
+next use the identity $ \sum_i \left( \mathbf{e}_i \otimes \mathbf{I}_\omega \right) \mathbf{A}_{ij} = \sum_i \mathbf{e}_i \otimes \mathbf{A}_{ij}$ to rewrite this again 
+\begin{eqnarray}
+\tilde{\mathbf{A}}=\sum_{i,j=1}^6 \left(\mathbf{e}_i \otimes  \mathbf{I}_\omega \right) \mathbf{A}_{ij} \left( \mathbf{e}^\tr_j \otimes \mathbf{I}_\omega \right).
+  \end{eqnarray}
+Use Roth's theorem, $\text{vec}\mathbf{A}\mathbf{B}\mathbf{C}=\left(\mathbf{C}^\tr \otimes \mathbf{A}\right) \text{vec} \mathbf{B}$, where $\mathbf{A}=\left(\mathbf{e}_i \otimes  \mathbf{I}_\omega \right)$, $\mathbf{B}=\mathbf{A}_{ij}$ and $\mathbf{C}=\left( \mathbf{e}^\tr_j \otimes \mathbf{I}_\omega \right)$, to obtain the following formula for  $\mbox{vec}\mathbf{A}$:
+\begin{equation}
+\mbox{vec}\tilde{\mathbf{A}}=\sum_{i,j}^6 \left(\mathbf{e}_j \otimes \mathbf{I}_\omega\right) \otimes \left(\mathbf{e}_i \otimes \mathbf{I}_\omega\right)\mbox{vec}\mathbf{A}_{ij}.
+\label{eq:vecAinAijs}
+\end{equation}
+
+Now we are ready to analyze term $\circled{C}$ in the Jacobian. Replace the derivative of $\mbox{vec} \tilde{\mathbf{A}}$ with equation (\ref{eq:vecAinAijs}), such that 
+\begin{eqnarray}
+  \frac{1}{\lambda_{AA}}(\hat{\bo p}^\tr \otimes  \mathbf{I}_{2\omega g})\frac{\partial \mbox{vec} \mathbf{A}}{\partial \tilde{\bo p}^\tr}\bigg\rvert_{\hat{\bo p}}&=&\frac{1}{\lambda_{AA}}\sum_{i,j=1}^{6} \left(\hat{\bo p}^\tr \otimes \mathbf{I}_{2\omega g}\right)\left(\left(\mathbf{e}_{j} \otimes \mathbf{I}_\omega\right) \otimes \left(\mathbf{e}_{i} \otimes \mathbf{I}_\omega\right) \right)\frac{\partial \mbox{vec} \mathbf{A}_{ij}}{\partial \tilde{\bo p}^\tr}\bigg\rvert_{\hat{\bo p}}.\nonumber\\ \label{eq:third_term_J}
+\end{eqnarray}
+Use $(\mathbf{A} \otimes \mathbf{B}) (\mathbf{C} \otimes \mathbf{D})= \mathbf{A} \mathbf{C}\otimes \mathbf{B} \mathbf{D}$ to rewrite  
+\begin{equation}
+  \left(\hat{\bo p}^\tr \otimes  \mathbf{I}_{2\omega g}\right) \left(\left(\mathbf{e}_{j} \otimes \mathbf{I}_\omega\right) \otimes \left(\mathbf{e}_{i} \otimes \mathbf{I}_\omega\right)\right)= \left(\hat{\bo p}^\tr \left(\mathbf{e}_{j} \otimes \mathbf{I}_\omega\right) \right) \otimes  \left(\mathbf{I}_{2\omega g} (\mathbf{e}_{i} \otimes \mathbf{I}_\omega) \right), \label{eq:rewrite_kron_four_matrices}
+\end{equation}
+substituting this expression into the right hand side of equation (\ref{eq:third_term_J}) yields
+\begin{eqnarray}
+\frac{1}{\lambda_{AA}}(\hat{\mathbf{p}}^\tr \otimes \mathbf{I}_{2\omega g})\frac{\partial \mbox{vec} \mathbf{A}}{\partial \tilde{\bo p}^\tr}\bigg\rvert_{\hat{\bo p}}&=&\frac{1}{\lambda_{AA}}\sum_{i,j=1}^{6} \left(\hat{\bo p}^\tr \left(\mathbf{e}_{j} \otimes \mathbf{I}_\omega\right)\right)\otimes \left( \mathbf{I}_{2\omega g} \left(\mathbf{e}_{i} \otimes \mathbf{I}_\omega\right)\right) \frac{\partial \mbox{vec} \mathbf{A}_{ij}}{\partial \tilde{\bo p}^\tr}\bigg\rvert_{\hat{\bo p}}.\nonumber\\ \label{eq:45_app}
+\end{eqnarray}
+
+Next substitute $\hat{\mbf{p}}^{\intercal} = \left( \hat{\mbf{p}}^{\intercal}_{S,AA},\hat{\mbf{p}}^{\intercal}_{X,AA},\mbf{0},\mbf{0},\mbf{0},\mbf{0} \right)$ into the right side of equation (\ref{eq:45_app}) so that only terms with $j=1$ and $j=2$ are nonzero, yielding
+
+\begin{align*}
+	\frac{1}{\lambda_{AA}}(\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{2 \omega g}) \left.
+\frac{ \partial \text{vec}\mbf{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} &= \frac{1}{\lambda_{AA}} \sum^{6}_{i=1} \big( \hat{\mbf{p}}^{\intercal}_{S,AA} \otimes (\mbf{e}_{i} \otimes \mbf{I}_{\omega}) \big) \left.
+\frac{ \partial \text{vec}\mbf{A}_{i,1}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} \\
+		&+ \frac{1}{\lambda_{AA}} \sum^{6}_{i=1} \big( \hat{\mbf{p}}^{\intercal}_{X,AA} \otimes (\mbf{e}_{i} \otimes \mbf{I}_{\omega}) \big) \left.
+\frac{ \partial \text{vec}\mbf{A}_{i,2}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} \numberthis
+\end{align*}
+
+\noindent Selfing does not depend on the frequency vector, and therefore none of the $\mbf{A}_{1,i}$ are a function of the frequency vector $\mbf{p}$,
+
+\begin{equation}
+	\left. \frac{ \partial \text{vec}\mbf{A}_{1,i}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} = 0,\,\text{for all }i. 
+\end{equation}
+
+\noindent (see Eq{\ref{eq:AtildeCoexistence}}). Next, write down each term in the sum over $i$ and take the derivative  of the $\text{vec}\mbf{A}_{i,1}$'s to obtain
+\begin{align*}
+	\frac{1}{\lambda_{AA}}(\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{2 \omega g}) \left.
+\frac{ \partial \text{vec}\mbf{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} &= \frac{1}{\lambda_{AA}} \big( \hat{\mbf{p}}^{\intercal}_{S,AA} \otimes (\mbf{e}_{2} \otimes \mbf{I}_{\omega}) \big) \left.
+\frac{ \partial \text{vec}\mbf{A}_{21}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}, \\
+		&+ \frac{1}{\lambda_{AA}} \big( \hat{\mbf{p}}^{\intercal}_{X,AA} \otimes (\mbf{e}_{2} \otimes \mbf{I}_{\omega}) \big) \left.
+\frac{ \partial \text{vec}\mbf{A}_{22}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}, \\
+		&+\frac{1}{\lambda_{AA}} \big( \hat{\mbf{p}}^{\intercal}_{S,AA} \otimes (\mbf{e}_{4} \otimes \mbf{I}_{\omega}) \big) \left.
+\frac{ \partial \text{vec}\mbf{A}_{41}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}, \\
+		&+ \frac{1}{\lambda_{AA}} \big( \hat{\mbf{p}}^{\intercal}_{X,AA} \otimes (\mbf{e}_{4} \otimes \mbf{I}_{\omega}) \big) \left.
+\frac{ \partial \text{vec}\mbf{A}_{42}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}.
+\numberthis
+\end{align*}
+Substituting in the expressions for the block matrices yields
+\begin{align*}
+	\frac{1}{\lambda_{AA}}(\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{2 \omega g}) \left. \frac{ \partial \text{vec}\mbf{A}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} = 
+		&\frac{(1 - C_{AA})}{\lambda_{AA}} \big[ \hat{\mbf{p}}^{\intercal}_{X,AA}+ \hat{\mbf{p}}^{\intercal}_{S,AA} \big]\otimes (\mbf{e}_{2} \otimes \mbf{I}_{\omega}) \text{vec}\mbf{F}_{AA} \left.\frac{\partial q^{\prime}_{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} \\
+		&-\frac{(1 - C_{AA})}{\lambda_{AA}} \big[ \hat{\mbf{p}}^{\intercal}_{X,AA}+ \hat{\mbf{p}}^{\intercal}_{S,AA} \big]\otimes (\mbf{e}_{4} \otimes \mbf{I}_{\omega}) \text{vec}\mbf{F}_{AA} \left.\frac{\partial q^{\prime}_{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} \\
+	   \numberthis
+\end{align*}
+
+\begin{landscape}
+Finally, apply Roth's theorem (Roth 1934), $\left( \mbf{C}^{\intercal} \otimes \mbf{A} \right) \text{vec}\mbf{B} = \text{vec}\mbf{ABC}$ (e.g., where $\mbf{C}^{\intercal} = \hat{\mbf{p}}^{\intercal}_{AA}$, $\mbf{A} = (\mbf{e}_1 \otimes \mbf{I}_{\omega})$, and $\text{vec}\mbf{B} = \text{vec}\mbf{F}_{AA}$) to rewrite as
+
+\begin{align*}
+	\frac{1}{\lambda_{AA}}(\hat{\mbf{p}}^{\intercal} \otimes \mbf{I}_{2 \omega g}) \left. \frac{ \partial \text{vec}\mbf{A}}{\partial \hat{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}} = 
+		&\frac{(1 - C_{AA})}{\lambda_{AA}}  \text{vec}\big[(\mbf{e}_{2} \otimes \mbf{I}_{\omega}) \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big)\big]  \left.\frac{\partial q^{\prime}_{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}  \\
+		&- \frac{(1 - C_{AA})}{\lambda_{AA}}  \text{vec}\big[(\mbf{e}_{4} \otimes \mbf{I}_{\omega}) \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big)\big]  \left.\frac{\partial q^{\prime}_{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{p}}}  \label{eq:partCwithpartials_app} \numberthis
+\end{align*}
+
+\noindent written in terms of the block matrices, this yields
+{
+\scriptsize
+\begin{align*}
+	\circled{C} =& \frac{(1 - C_{AA})}{\lambda_{AA}} \,* \\
+	&\left(
+			\begin{array}{cc|cc|cc}
+					\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0}  \\
+			\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{ \intercal}_{S,AA}}  &  \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{ \intercal}_{X,AA}} &  \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\intercal}_{S,Aa}}& \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\intercal}_{X,Aa}} & \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\intercal}_{S,aa}}&  \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\intercal}_{X,aa}} \\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
+				-\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{ \intercal}_{S,AA}}  &  -\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{ \intercal}_{X,AA}} & - \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\intercal}_{S,Aa}}& -\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\intercal}_{X,Aa}} &- \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\intercal}_{S,aa}}& - \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \frac{\partial q^{\prime}_{A}}{\partial \mbf{p}^{\intercal}_{X,aa}} \\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} 
+			\end{array} \right)  \numberthis			
+\end{align*}
+}
+The above expression requires the derivative of the frequency of allele $A$ in the gamete pool with respect to the population frequency vector:
+
+\begin{equation}
+	\left. \frac{\partial q^{\prime}_{A}}{\partial \tilde{\mbf{p}}^{\intercal}} \right|_{\hat{\mbf{P}}}
+\end{equation}
+
+ \end{landscape}
+
+ Start with equation \textcolor{red}{(4)} from the main text:
+\begin{linenomath*}
+\begin{equation} \label{eq:maleGametePool}
+	\left(
+		\begin{array}{c}
+			q^{\prime}_{A} \\
+			q^{\prime}_{a} \\
+		\end{array} \right) = 
+			\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime}}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime} \|} = 
+				\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)\|},
+\end{equation}
+\end{linenomath*}
+therefore
+\begin{equation}
+  q^\prime_A=\frac{\mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime  \mathbf{p}^\prime}{\mathbf{1}^\tr_{2}\mathbf{W}^\prime\mathbf{F}^\prime\mathbf{p}^\prime}, \label{eq:male_gamete_pool}
+\end{equation}
+ where the one norm can be replaced by $\mathbf{1}^\tr_{2}\mathbf{W}^\prime\mathbf{F}^\prime \mathbf{p}^\prime$ because $\mathbf{p}^\prime$ is nonnegative.
+For convenience, we will denote the normalizing factor in the denominator with $ \mbox{p}_{n}$,
+ \begin{equation}
+   \mbox{p}_{n}=\mathbf{1}^\tr_{2}\mathbf{W}^\prime\mathbf{F}^\prime \mathbf{p}^\prime
+ \end{equation}
+ Taking the derivative of $ q^\prime_A$ yields
+\begin{eqnarray}
+  \frac{\partial   q^\prime_A }{\partial \tilde{\bo p}^\tr}&=& \frac{1}{ \mbox{p}_{n}}\mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime   \frac{\partial  \mathbf{p}^\prime}{\partial \tilde{\bo p}^\tr} - \frac{\mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime \mathbf{p}^\prime}{ \mbox{p}_{n}^2} \left( \mathbf{1}^\tr_{2}\mathbf{W}^\prime\mathbf{F}^\prime \frac{\partial  \mathbf{p}^\prime}{\partial \tilde{\bo p}^\tr} \right).
+\label{eq:56_app}
+\end{eqnarray}
+Recall 
+\begin{equation}
+  \mathbf{p}^\prime=\left(\begin{array} {c}
+\mathbf{p}^S_{AA}+\mathbf{p}^X_{AA}\\
+\mathbf{p}^S_{Aa}+\mathbf{p}^X_{Aa}\\
+\mathbf{p}^S_{aa}+\mathbf{p}^X_{aa}
+  \end{array}\right),
+\end{equation}
+and 
+\begin{equation}
+  \tilde{\bo p}=\left(\begin{array} {c}
+\mathbf{p}_{AA}^S\\
+\mathbf{p}_{AA}^X\\
+\hline
+\mathbf{p}_{Aa}^S\\
+\mathbf{p}_{Aa}^X\\
+\hline
+\mathbf{p}_{aa}^S\\
+\mathbf{p}_{aa}^X
+  \end{array}\right),
+\end{equation}
+to calculate 
+\begin{equation}
+\frac{\partial  \mathbf{p}^\prime}{\partial \tilde{\bo p}^\tr} =  \left(\begin{array} {cccccc}
+\mathbf{I}& \mathbf{I}& \mathbf{0}  &\mathbf{0} &\mathbf{0}  &\mathbf{0} \\
+\mathbf{0} &\mathbf{0}  &\mathbf{I}& \mathbf{I}& \mathbf{0} &\mathbf{0} \\
+\mathbf{0} &\mathbf{0}  &\mathbf{0} &\mathbf{0}  &  \mathbf{I}& \mathbf{I}
+\end{array}\right).
+\end{equation}
+First we will evaluate the first term in the sum in equation (\ref{eq:56_app}),
+\begin{eqnarray}
+	 \frac{1}{ \mbox{p}_{n}}\mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime   \frac{\partial  \mathbf{p}^\prime}{\partial \tilde{\bo p}^\tr} &=&  \frac{1}{ \mbox{p}_{n}} (1,0) \left(\begin{array} {ccc}
+\mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime& \frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{0} \\
+\mathbf{0} &\frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime
+\end{array}\right) \left(\begin{array} {cccccc}
+\mathbf{I}& \mathbf{I}& \mathbf{0}  &\mathbf{0} &\mathbf{0}  &\mathbf{0} \\
+\mathbf{0} &\mathbf{0}  &\mathbf{I}& \mathbf{I}& \mathbf{0} &\mathbf{0} \\
+\mathbf{0} &\mathbf{0}  &\mathbf{0} &\mathbf{0}  &  \mathbf{I}& \mathbf{I}
+\end{array}\right) \nonumber \\
+&=& \frac{1}{ \mbox{p}_{n}} \left( \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime, \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime , \frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime, \frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime, \mathbf{0}, \mathbf{0} \right)\label{eq:first_term_app_reviewer}
+\end{eqnarray}
+Similarly for the second term in the sum in equation (\ref{eq:56_app}),
+\begin{eqnarray}
+	- \frac{\mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime \mathbf{p}^\prime}{ \mbox{p}_{n}^2} \left( \mathbf{1}^\tr_{2}\mathbf{W}^\prime\mathbf{F}^\prime \frac{\partial  \mathbf{p}^\prime}{\partial \tilde{\bo p}^\tr} \right) &=& -\frac{1}{ \mbox{p}_{n}}(1,1) \left(\begin{array} {ccc}
+\mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime& \frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{0} \\
+\mathbf{0} &\frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime
+\end{array}\right) \left(\begin{array} {cccccc}
+\mathbf{I}& \mathbf{I}& \mathbf{0}  &\mathbf{0} &\mathbf{0}  &\mathbf{0} \\
+\mathbf{0} &\mathbf{0}  &\mathbf{I}& \mathbf{I}& \mathbf{0} &\mathbf{0} \\
+\mathbf{0} &\mathbf{0}  &\mathbf{0} &\mathbf{0}  &  \mathbf{I}& \mathbf{I}
+\end{array}\right) \nonumber \\
+ &=& -\frac{1}{ \mbox{p}_{n}} \left( \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime, \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime , \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime, \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime, \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime , \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime  \right) \label{eq:second_term_app_reviewer}
+\end{eqnarray}
+Finally add equations (\ref{eq:first_term_app_reviewer}) and (\ref{eq:second_term_app_reviewer}) to obtain
+\begin{eqnarray}
+  \frac{\partial   q^\prime_A }{\partial \tilde{\bo p}^\tr}\bigg\rvert_{\hat{\bo p}}&=& \frac{1}{ \mbox{p}_{n}} \left( \mathbf{0}, \mathbf{0}, -\frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime, -\frac{1}{2}\mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime, -\mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime, -\mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime  \right) \label{eq:79_app},
+\end{eqnarray}
+where at the boundary
+\begin{eqnarray}
+  \mbox{p}_{n}&=&\mathbf{1}^\tr_\omega \mathbf{F}_{AA}^\prime \left( \hat{\mathbf{p}}^S_{AA}+\hat{\mathbf{p}}^X_{AA} \right), \\
+  \mathbf{e}_1^\tr \mathbf{W}^\prime \mathbf{F}^\prime \mathbf{p}^\prime&=&\mathbf{1}^\tr_\omega \mathbf{F}_{AA}^\prime \left( \hat{\mathbf{p}}^S_{AA}+\hat{\mathbf{p}}^X_{AA} \right)=  \mbox{p}_{n}
+\end{eqnarray}
+
+\begin{landscape}
+Finally, plugging equation (\ref{eq:79_app}) into (\ref{eq:partCwithpartials_app}) yields
+
+{
+\scriptsize
+\begin{align*}
+	\circled{C} =& \\
+	\frac{(1 - C_{AA})}{\mbox{p}_{n}\lambda_{AA}} &\left(
+			\begin{array}{cc|cc|cc}
+					\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0}  \\
+				\mbf{0}   &  	\mbf{0}  & -\frac{1}{2} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime& -\frac{1}{2} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & - \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime&  - \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime \\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
+					\mbf{0}   &  	\mbf{0}  & \frac{1}{2} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime& \frac{1}{2} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &  \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime&  \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime \\  \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} 
+			\end{array} \right)  \numberthis			
+\end{align*}
+}
+
+\subsection{The Jacobian}
+
+Putting all the pieces together into $M$, we get the Jacobian:
+
+{
+\footnotesize
+\begin{align*} \label{eq:combinedM}
+	\mbf{M} = 
+			 +\frac{1}{\lambda_{AA}} &
+						   \left(
+			\begin{array}{cc|cc|cc}
+		 \mathbf{U}^S_{AA} +C_{AA} (1 - \delta) \mbf{F}_{AA} & C_{AA} (1 - \delta) \mbf{F}_{AA}& \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}& \mbf{0} & \mbf{0} \\
+				(1 - C_{AA}) \mbf{F}_{AA} &\mathbf{U}^X_{AA} + (1 - C_{AA}) \mbf{F}_{AA}  & \frac{1}{2}(1 - C_{Aa}) \mbf{F}_{Aa}  & \frac{1}{2} (1 - C_{Aa})\mbf{F}_{Aa} & \mbf{0} & \mbf{0} \\ \hline
+				\mbf{0} & \mbf{0} & \mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} & (1 - C_{aa}) \mbf{F}_{aa} & (1 - C_{aa}) \mbf{F}_{aa}\\ \hline
+				\mbf{0} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa} &C_{aa} (1 - \delta) \mbf{F}_{aa} \\
+				\mbf{0} & \mbf{0} & \mbf{0}  & \mbf{0} & \mbf{0} &\mathbf{U}^X_{aa} 
+			\end{array} \right) \\
+		-\frac{1}{\lambda_{AA}} &\left(
+			\begin{array}{cc|cc|cc}
+				\hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{aa} \\
+				\hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{AA} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{Aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^S_{aa} & \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{U}^X_{aa} \\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} 
+			\end{array} \right) \\
+		-\frac{1}{\lambda_{AA}}& \left(
+			\begin{array}{cc|cc|cc}
+				(1 - \delta C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & (1 - \delta C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & (1 - \delta C_{Aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & (1 - \delta C_{Aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa}  & (1 - \delta C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} & (1 - \delta C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} \\
+				 	(1 - \delta C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & 	(1 - \delta C_{AA}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{AA} & (1 - \delta C_{Aa})  \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & (1 - \delta C_{Aa})  \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{Aa} & (1 - \delta C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa} & (1 - \delta C_{aa}) \hat{\mbf{p}}_{AA} \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}_{aa}\\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0}
+			\end{array} \right)\\
+		\frac{(1 - C_{AA})}{\mbox{p}_{n}\lambda_{AA}} &\left(
+			\begin{array}{cc|cc|cc}
+					\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0}  \\
+				\mbf{0}   &  	\mbf{0}  & -\frac{1}{2} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime& -\frac{1}{2} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & - \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime&  - \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime \\ \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
+					\mbf{0}   &  	\mbf{0}  & \frac{1}{2} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime& \frac{1}{2} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &  \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime&  \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime \\  \hline
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
+				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} 
+			\end{array} \right) 
+			\numberthis
+\end{align*}
+}
+
+\end{landscape}
+
+%%%%%%%%%%%%%%%
+\subsection{Eigenvalues of the Jacobian}
+
+The Jacobian matrix given in equation (\ref{eq:combinedM}) is upper block triangular, and the eigenvalues of $\mbf{M}$ are therefore the eigenvalues of the diagonal blocks.
+Because of the way that self-fertilization alters the projection matrix, our blocks are different than those in both  \cite{deVriesCaswell2019a} and \cite{deVriesCaswell2019b}.
+The largest absolute eigenvalue of the Jacobian, i.e.
+the spectral radius $\rho(\mathbf{M})$, determines the stability of the boundary equilibrium.
+We will denote the three blocks along the diagonal with $\mathbf{M}_{11}$, $\mathbf{M}_{22}$, and $\mathbf{M}_{33}$, such that
+\begin{equation}
+\mathbf{M}_{33}=\frac{1}{\lambda_{AA}}\mathbf{U}^X_{aa} .
+\end{equation}
+Block $\mathbf{M}_{33}$ projects perturbations in the $aa$ outcrossing $aa$ direction.
+In the neighbourhood of the $AA$ equilibrium, $aa$ homozygotes are rare, and thus $\mathbf{M}_{33}$ normally does not determine the stability of $\mathbf{M}$.
+An exception occurs when
+
+\begin{equation}
+  \lambda_{AA}<\rho\left(\mathbf{U}^X_{aa} \right)<1.
+\end{equation}
+ That is, if the $AA$ population is declining sufficiently rapidly, the $aa$ homozygote may increase in frequency simply by declining to extinction more slowly.
+If the homozygous $AA$ population is stable or increasing, so that $\lambda_{AA} \geq 1$, this cannot happen.
+Similarly, if $\mathbf{U}^X_{aa}$ is age-classified with a maximum age, $\rho(\mathbf{U}^X_{aa})=0$, and the phenomenon can not happen.
+We neglect this pathological case in our discussions.
+The first 2$\omega\times$2$\omega$ block on the diagonal, $\mathbf{M}_{11}$, projects perturbations in the $AA$ boundary, and because $\hat{\mathbf{p}}$ is stable to perturbations in that boundary, $\rho(\mathbf{M}_{11})<1$.
+
+The stability of $\hat{\mathbf{p}}$ is thus determined by the middle diagonal block, 
+{
+\footnotesize
+ \begin{align*} \label{eq:M22}
+	\mbf{M}_{22} = 
+		\frac{1}{\lambda_{AA}}\left(\begin{array}{ccc}
+\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & 0 \\
+ \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2}  \Theta_{Aa} &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2}\Theta_{Aa} & (1-C_{aa})\mbf{F}_{aa} + \Theta_{aa}\\
+  \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa}  \\ 
+			\end{array} \right), \numberthis
+\end{align*}
+}
+where $\Theta_{aa} = \frac{(1 - C_{AA})}{p_{n}} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{aa}^\prime$, and  $\Theta_{Aa} = \frac{(1 - C_{AA})}{p_{n}} \mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime$
+
+\begin{equation}
+  \mbox{p}_{n}=\mathbf{1}^\tr_\omega \mathbf{F}_{AA}^\prime \left( \hat{\mathbf{p}}^S_{AA}+\hat{\mathbf{p}}^X_{AA} \right)	
+\end{equation}
+Note that the outcrossing terms are the sum of a term due to heterozygote ovules being fertilized by pollen of the resident ($\frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}$), and heterozygote pollen fertilizing resident females ($\frac{1}{2 }\Theta_{Aa}=\frac{1}{2  \mbox{p}_{n}} (1 - C_{AA})\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime$.
+
+ The largest absolute value of the eigenvalues of the Jacobian matrix, the leading eigenvalue, evaluated at the $AA$ boundary, denoted by $\tilde{\zeta}_{AA}$, is therefore
+{
+\footnotesize
+\begin{align*} \label{eq:eigAAFull}
+	\tilde{\zeta}_{AA} = 
+		\frac{1}{\lambda_{AA}}\rho\left(\begin{array}{ccc}
+\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & 0 \\
+ \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2}  \Theta_{Aa} &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2}\Theta_{Aa} & (1-C_{aa})\mbf{F}_{aa} + \Theta_{aa}\\
+  \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa}  \\ 
+			\end{array} \right). \numberthis
+\end{align*}
+}
+
+% \noindent Rearranging the projection matrix so that outcrossing comes on top of selfing in each block simplifies things, but in the end the leading eigenvalue evaluated at the $aa$ boundary is
+
+% \begin{align*} \label{eq:eigaa}
+% 	\tilde{\zeta}_{aa} = 
+% 		\frac{1}{\lambda_{AA}} \rho&\left(\begin{array}{c|cc}
+% 				\mcal{U}^S_{AA} + C_{AA} (1 - \delta) \mbf{F}_{AA} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}\\ \hline
+% 				\mbf{F}_{AA} (1 - C_{AA}) & \mcal{U}^X_{Aa} + \frac{(1 - C_{Aa})}{2 p_n} (\mbf{F}_{aa} \hat{\mbf{p}}_{aa}) \otimes \mbf{1}^{\intercal}_{\omega} \mbf{F}^{\prime}_{Aa} & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) \\
+% 				\mbf{0} & -C_{Aa}(1 - \delta) (\mbf{F}_{aa} \hat{\mbf{p}}_{aa}) & \mcal{U}^S_{Aa} + \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} \\
+% 			\end{array} \right). \numberthis
+% \end{align*}
+
+A boundary equilibrium is unstable to invasion by the rare allele if the leading eigenvalue of the Jacobian evaluated at the equilibrium is greater than $1$.
+The conditions for a protected polymorphism is that both boundaries are unstable (i.e., both $\tilde{\zeta}_{AA} > 1$ and $\tilde{\zeta}_{aa} > 1$).
+The conditions for coexistence are, therefore, as given in equation \textcolor{red}{(B1)} and \textcolor{red}{(B2)},
+
+{\footnotesize
+\begin{equation} \label{eq:coexist_AA}
+	\lambda_{AA} < 
+			\rho\left(\begin{array}{ccc}
+\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & 0 \\
+ \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2}  \Theta_{Aa} &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2}\Theta_{Aa} & (1-C_{aa})\mbf{F}_{aa} + \Theta_{aa}\\
+  \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa}  \\ 
+			\end{array} \right), \\ 
+\end{equation}
+
+\begin{equation} \label{eq:coexist_aa}
+	\lambda_{aa} < 
+			\rho\left(\begin{array}{ccc}
+\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & 0 \\
+ \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2 } \Phi_{Aa} &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2} \Phi_{Aa} & (1-C_{AA})\mbf{F}_{AA} + \Phi_{AA}\\
+  \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{AA} + C_{AA} (1 - \delta) \mbf{F}_{AA}  \\ 
+			\end{array} \right), \\ 
+\end{equation} 
+}
+
+where $\Phi_{AA} = \frac{(1 - C_{aa})}{p_{n}} \mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime$  and $\Phi_{Aa}=\frac{(1 - C_{aa})}{p_{n}} \mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime$.
 
@@
 % except when they are already cited in the main text.
-
-\begin{thebibliography}{}
-
-\bibitem[{Cook et~al.(2015)Cook, Collaborator, and Expert}]{CookEtAl2015}
-Cook, O.~E., G.~H. Collaborator, and A.~Q. Expert. 2015.
-\newblock Data from: Template and guidelines for using \LaTeX{} in \textit{The American Naturalist}.
-\newblock American Naturalist, Dryad Digital Repository, https://dx.doi.org/10.5061/dryad.XYZAB123.
-
-\bibitem[{Darwin(1859)Darwin}]{OriginSpec}
-Darwin, C. 1859.
-\newblock On the origin of species by means of natural selection, or the preservation of favoured races in the struggle for life.
-\newblock J.~Murray, London.
-
-\bibitem[{Davis et~al.(2011)Davis, Brakora, and Lee}]{DavisEtAl2011}
-Davis, E.~B., K.~A. Brakora, and A.~H. Lee. 2011.
-\newblock Evolution of ruminant headgear: a review.
-\newblock Proceedings of the Royal Society~B 278:2857--2865.
-
-\bibitem[{Inglis et~al.(2011)Inglis, Roberts, Gardner, and Buckling}]{Ing11}
-Inglis, R.~F., P.~G. Roberts, A.~Gardner, and A.~Buckling. 2011.
-\newblock Spite and the scale of competition in \textit{Pseudomonas
-  aeruginosa}.
-\newblock American Naturalist 178:276--285.
-
-\bibitem[{Fastovsky(2009)Fastovsky}]{LemKapEx07}
-Fastovsky, D.~E. 2009.
-\newblock Ideas in dinosaur paleontology: resonating to social and political context.
-\newblock Pages 239--253 \emph{in} D. Sepkoski and M. Ruse, eds. The Paleobiological Revolution. University of Chicago Press, Chicago~IL.
-
-\bibitem[{Xiao et~al.(2015)Xiao, McGlinn, and White}]{Xiao2015}
-Xiao, X., D.~J. McGlinn, and E.~P. White. 2015.
-\newblock A strong test of the maximum entropy theory of ecology.
-\newblock American Naturalist 185:E705--E80.
-
-\end{thebibliography}
+\bibliography{Refs2.bib}
 

### 2021-05-07 doc/SA-Hermaphrodites-wDemography.tex
Updated README.md
Self explanatory commit.
empty commit for fixMaster
Again. empty commit to fix erroneous merge
Minor edits main text and suppl

@@
 \begin{document}
-\title{The demographic costs of sexually antagonistic selection in simultaneous hermaphrodites}
+\title{The demographic costs of sexually antagonistic selection in partially selfing populations}
 
@@
 In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious/gonochorous) and partially selfing hermaphrodite populations.
-We focus on identifying parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population intrinsic growth rate remains positive.
+We focus on identifying parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the intrinsic population growth rate remains positive.
 In addition, we analyze the effects of inbreeding depression manifesting in different life-history stages, and give an illustrative example of the potential for SA polymorphism in real populations using empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus}.
@@
 To take the population dynamical consequences of SA selection into account, a model is required that links changes in genotype distributions to population dynamics.
-The potential demographic costs of sexual antagonism were pointed out by \citep{kokko2003sexy}, but few papers since then explicitly incorporated demography into models of sexual antagonism (with the notable exception of \citealt{harts2014demography}).
-\citet{deVriesCaswell2019b} introduced a mendelian matrix model with intralocus sexual antagonism, and population dynamics but did not perform an analysis of the consequences of demographic viability for the scope of sexual antagonism to maintain polymorphism.
-
-In this paper, we extend the model of \citep{deVriesCaswell2019b} to include both dioecious and hermaphroditic species in order to study the parameter conditions under which balanced SA polymorphisms are also demographically viable.
+The potential demographic costs of sexual antagonism were pointed out by \cite{kokko2003sexy}, but few papers since then explicitly incorporated demography into models of sexual antagonism (with the notable exception of \citealt{harts2014demography}).
+\citet{deVriesCaswell2019a} introduced a mendelian matrix model with intralocus sexual antagonism, and population dynamics but did not perform an analysis of the consequences of demographic viability for the scope of sexual antagonism to maintain polymorphism.
+
+In this paper, we extend the model of \citep{deVriesCaswell2019a,deVriesCaswell2019b} to include both dioecious and hermaphroditic species in order to study the parameter conditions under which balanced SA polymorphisms are also demographically viable.
 A major strength of our approach is that the model can be parameterized using empirically estimated demographic rates, enabling us to make predictions about the scope for demographically viable SA polymorphisms that are grounded in the biology of real populations.
@@
 The derivation and some of the key results follow closely those presented in \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b}.
-In fact, the model reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019b} under obligate outcrossing.
+In fact, the model reduces to the stage-structured model of \citet{deVriesCaswell2019a} under obligate outcrossing.
 A full derivation of the model and analyses is presented in the \hl{Online Supplementary Material}, and all computer code necessary to reproduce the results are available at \url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
@@
 Following previous theory for partially selfing populations \cite[e.g.,][]{Charlesworth2010,JordanConnallon2014,Glemin2021}, we use a 'fixed prior selfing' model where all individuals are assumed to self-fertilize at a constant rate, $C$, prior to receiving outcross pollen, for reasons of analytic tractability.
-In the \hl{Online Supplement}, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
+\textcolor{red}{In the \hl{Online Supplement}, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
+Do we???}
 
@@
 The blocks of the component matrices in Eq(\ref{eq:Atilde}) correspond to production of offspring by self-fertilization and outcrossing ($\mcal{F}^S$ and $\mcal{F}^X$ in $\tilde{\mbf{F}}$), and survival of extant individuals produced by selfing or outcrossing ($\mcal{U}^S$ and $\mcal{U}^X$ in $\tilde{\mbf{U}}$).
-The survival matrix for individuals produced through selfing and outcrossing are,
+The survival matrices for individuals produced through selfing and outcrossing are,
 \begin{linenomath*}
@@
 $\mbf{U}^{S}_{i}$,\, $\mbf{U}^{X}_{i}$ & Genotype-specific transition and survival matrices & $\omega \times \omega$ \\
-$\mcal{U}^{S},\, \mcal{U}^{X}$ & Block diagonal selfed/outcrossed survival matrices & $\omega g \times \omega$ \\
+$\mcal{U}^{S},\, \mcal{U}^{X}$ & Block diagonal selfed/outcrossed survival matrices & $\omega g \times \omega$ g\\
 $\mbf{F}_{i}$,\, $\mbf{F}^{\prime}_{i}$ & Genotype-specific fertility matrices & $\omega \times \omega$ \\
-$\mbb{F}^{\prime}$ & Block diagonal fertility matrix & $\omega g \times \omega$ \\
-$\mcal{F}^{S},\, \mcal{F}^{X}$ & Block diagonal selfed/outcrossed fertility matrix & $\omega g \times \omega$ \\
+$\mcal{F}^{S},\, \mcal{F}^{X}$ & Block diagonal selfed/outcrossed fertility matrix & $\omega g \times \omega$g \\
 $\mbf{H}^{S}_{i}$,\, $\mbf{H}^{X}_{i}$ & Parent-offspring genotype maps & $g \times g$ \\
-$\mbf{W}$   & Allele segregation matrix & $\omega \times g$ \\
+$\mbf{W}$   & Allele segregation matrix & $2 \times \omega $g \\
 % $\mbf{K}$   & Vec-permutation matrix & $g \times g$ \\
@@
 $q^{\prime}_{A}$,\, $q^{\prime}_{a}$ & Allele frequencies in male gamete pool & $1$ \\
-$\tilde{\mbf{A}}(\tilde{\mbf{p}})$ & Population projection matrix & $2 \omega g \times g$ \\
+$\tilde{\mbf{A}}(\tilde{\mbf{p}})$ & Population projection matrix & $2 \omega g \times 2\omega g$ \\
 
@@
 However, the survival matrices can be used to model the fitness effects of inbreeding depression at later stages of development by allowing different stage-specific survival and transition rates for individuals produced by self-fertilization vs.~outcrossing.
-By contrast, the parameter $\sigma$ only affects inbreeding depression through viability of selfed ovules.
+By contrast, the parameter $\delta$ only affects inbreeding depression through viability of selfed ovules.
 With this in mind, we define survival matrices for individuals produced by selfing and outcrossing as follows:
@@
 $\delta$ denotes the fraction of self-fertilized ovules that do not develop into juveniles due to inbreeding depression.
-An important difference between early- and late-acting inbreeding depression in the model is that $\delta$ does not affect $\mbf{U}^S$, the survival matrix for offspring produced by selfing (see Eq.~\ref{eq:Atilde}), while the $\delta_i$ terms directly alter the stage-specific survival and transition rates of offspring produced by selfing (see Eq.~\ref{eq:US}).
+An important difference between early- and late-acting inbreeding depression in the model is that $\delta$ affects the production of new individuals, whereas the $\delta_i$ affect the demographic rates of extant individuals, contained in $\mbf{U}^S$ (see Eq.~\ref{eq:US}).
 
@@
 Specifically, we examine scenarios of additive SA fitness effects ($h = 1/2$) and dominance reversals, where the deleterious fitness effect of each SA allele is partially recessive in each sex ($h = 1/4$).
-We relax these assumptions to explore the consequences of sex-specific dominance, which may evolve under asymmetric SA selection \citep{SpencerPriest2016}, on our model predictions in \hl{Appendix X}).
+\textcolor{red}{We relax these assumptions to explore the consequences of sex-specific dominance, which may evolve under asymmetric SA selection \citep{SpencerPriest2016}, on our model predictions in \hl{Appendix X}).
+}
 
@@
 {\itshape M.~guttatus} is an herbaceous, self-compatible wildflower native to western North America that exhibits remarkable among-population variation in numerous life-history and reproductive traits including selfing rates, inbreeding depression, floral morphology, and annual-to-perennial life-history \citep[e.g.,][]{RitlandGanders1987, Ritland1990, Willis1993, Willis1999a, Willis1999b, WuWillis2008}.
-Moreover, detailed demographic studies have been conducted on multiple populations of {\itshape M.~guttatus}, with demographic matrix data available on the public demographic database COMPADRE \citep{CompadreDB2020}.
+Moreover, detailed demographic studies have been conducted on multiple populations of {\itshape M.~guttatus}, with demographic data available on the public demographic database COMPADRE \citep{CompadreDB2020}.
 Below, we briefly outline how we parameterized our model using the available data; full details are provided in \hl{Appendix C}.
@@
 With the empirically parameterized model, we are able to make predictions about the scope for demographically viable SA polymorphism at a single locus in hypothetical populations with the same demographic rates as observed in \citet{PetersonEtAl2016}, for a range of selfing and inbreeding depression rates observed in other natural populations.
-We stress, however, that these are illustrative rather than explicit predictions of the likelihood of SA polymorphism in any specific population, and ignore measurement error for the estimated demographic rates.
+We stress, however, that these are illustrative rather than explicit predictions of the likelihood of SA polymorphism in any specific population, and they ignore measurement error for the estimated demographic rates.
 
@@
 When using demographic rates for either the Eagle Meadows or Low-Elevation Perennial populations, inv6 falls within demographically viable polymorphic parameter space (fig.~\ref{fig:mimulusFig}; estimated selection coefficients for inv6 were $(s_f,\,s_m) = (0.30,\, 0.31)$).
-Whether under obligate outcrossing or partial selfing, inv6 always falls squarely in the middle of SA polymorphic space when using the locally adapted Eagle Meadows demographic rates.
+\hl{I like Crispins idea of reporting predicted inv6 frequencies in our model, it nicely illustrates another useful output of the model} Whether under obligate outcrossing or partial selfing, inv6 always falls squarely in the middle of SA polymorphic space when using the locally adapted Eagle Meadows demographic rates.
 In contrast, when the selfing rate is at the higher end of empirical estimates for the Iron Mountain population in which inv6 has been documented ($C = 0.29$), inv6 falls near the upper boundary but still within SA polymorhic space when using the Low-Elevation Perennial demographic rates.
-This happens because under the empirical estimates of selfing and inbreeding depression, the polymorphic parameter space shifts downwards.
+\hl{Another place where we could report inv6 frequencies?} This happens because under the empirical estimates of selfing and inbreeding depression, the polymorphic parameter space shifts downwards.
 Additionally, inv6 falls much closer to the extinction threshold under partial selfing.
-This suggests that even relatively small perturbations to demgraphic rates or selection coefficients could result in non-locally adapted populations being unable to support the demographic costs associated with segregating SA alleles with selection coefficients of similar magnitude to inv6.
+This suggests that even relatively small perturbations to demographic rates or selection coefficients could result in non-locally adapted populations being unable to support the demographic costs associated with segregating SA alleles with selection coefficients of similar magnitude to inv6.
 
@@
 However, by ignoring the demographic consequences of SA genetic variation, these population genetic models have missed the possibility that SA polymorphisms may not be viable under realistic parameter conditions, and therefore unlikely to be observed in natural populations.
-By linking the individual-level fitness consequences of SA selection to population level demographic rates, our theoretical framework provides several key insights into the processes shaping SA genetic variation in natural populations.
+By linking the individual-level fitness consequences of SA selection to population level consequences, our theoretical framework provides several key insights into the processes shaping SA genetic variation in natural populations.
 
@@
 
-Like most demographic matrix models, ours also assumes female demographic dominance, where population growth rates are determined entirely by female fecundity \citep{Caswell2001,CaswellEtAl2018,deVriesCaswell2019b}.
+Like most demographic matrix models, ours also assumes female demographic dominance, where population growth rates are determined entirely by female fecundity \citep{pollard1975mathematical,Caswell2001,iannelli2005gender}.
 Yet, many hermaphroditic populations experience limitation of reproductive success due to both quantity and quality of male gametes \citep[e.g.,][]{Yund2000,AizenHarder2007, Harder2016}.
@@
 Interestingly, different forms of self-fertilization can aid in reproductive assurance, and should therefore have demographic consequences -- for example, under pollen limitation and delayed selfing (where only ovules that fail to receive outcross pollen are selfed), the selfing rate will be a function of genotype frequencies because these will directly influence pollen production \citep{HarderBarrett2006}.
-
+,
 Another major simplifying assumption in our model was that how individuals themselves were produced (i.e., by selfing or outcrossing) affects the level of inbreeding depression they suffer.

### 2021-05-07 doc/Supplements/SupplementaryMaterial.tex
Copied over newSupp changes
Minor edits main text and suppl

@@
 %
-\rhead{Supplement to Olito \& de Vries., "Costs of SA polymorphism" \textit{Am.~Nat.}}
+\rhead{Supplement to Olito \& de Vries., "Costs of SA selection" \textit{Am.~Nat.}}
 \setlength{\headsep}{0.3in}  
@@
 }
+where $\Phi_{AA} = \frac{(1 - C_{aa})}{p_{n}} \mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime$  and $\Phi_{Aa}=\frac{(1 - C_{aa})}{p_{n}} \mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime$.
 
-where $\Phi_{AA} = \frac{(1 - C_{aa})}{p_{n}} \mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{AA}^\prime$  and $\Phi_{Aa}=\frac{(1 - C_{aa})}{p_{n}} \mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime$.
+Because these coexistence conditions contain the stable stage specific frequencies on the boundaries, $\hat{\mbf{p}}_{X,AA}$, and $\hat{\mbf{p}}_{S,aa}$, finding an analytical expression in terms of the model parameters (i.e., a closed form solution),  is in general a non trivial task.
+We therefore used numerical techniques as described in the main text to identify invasion and extinction thresholds, which could then be used to define regions of demographically viable polymorphism across $s_f \times s_m$ parameter space.
+All computer code necessary to reproduce the results are available at \url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
 

### 2021-05-25 doc/SA-Hermaphrodites-wDemography.tex
Code for sex-specific dominance simulations
This commit adds functionality to several simulation and plotting
functions to explore the consequences of sex-specific dominance

@@
 For simplicity, we present results for scenarios of equal dominance in the main text (i.e., $h_f = h_m = h$).
-Specifically, we examine scenarios of additive SA fitness effects ($h = 1/2$) and dominance reversals, where the deleterious fitness effect of each SA allele is partially recessive in each sex ($h = 1/4$).
-\textcolor{red}{We relax these assumptions to explore the consequences of sex-specific dominance, which may evolve under asymmetric SA selection \citep{SpencerPriest2016}, on our model predictions in \hl{Appendix X}).
-}
+Specifically, we examine scenarios of ({\itshape i}) additive SA fitness effects ($h = 1/2$); and ({\itshape ii}) dominance reversals, where the deleterious fitness effect of each SA allele is partially recessive in each sex ($h = 1/4$).
+These scenarios are of particular interest because additive effects are commonly observed for small-effect alleles or quantitative traits \citep{AgrawalWhitlock2011}, while dominance reversals are predicted under fitness landscape models provided the population is not too far from the phenotypic optimum \citep{Manna2011, ConnallonClark2014}.
+\textcolor{red}{We explore the consequences of sex-specific dominance, which may evolve under asymmetric SA selection \citep{SpencerPriest2016} in \hl{Appendix X}).}
 
@@
 We used flower number as a proxy for female fertility.
-To calculate selection on flower number, we first calculated the geometric mean flower number in the field for wild-type and inv6 heterozygotes across the two years for which data was available (2012 and 2013), extrapolated the expected flower number for inv6 homozygotes, and calculated the relative fertility of each genotype given that inv6 homozygotes have the highest flower number ($w_i$, where $i \in \{inv6/inv6,\, inv6/w+,\, w+/w+\}$).
+To calculate selection on flower number, we first calculated the geometric mean flower number in the field for wild-type and inv6 heterozygotes across the two years for which data was available (2012 and 2013), extrapolated the expected flower number for inv6 homozygotes (table \ref{tab:MimReproEst}), and calculated the relative fertility of each genotype given that inv6 homozygotes have the highest flower number ($w_i$, where $i \in \{inv6/inv6,\, inv6/w+,\, w+/w+\}$).
 Using the relative fertility expressions outlined in Table \ref{tab:inv6Fitness}, we then solved for $s_f$.
 
+\begin{table}[htbp]
+ \centering
+ \caption{\bf Selection coefficient estimates for inv6}
+\label{tab:MimReproEst}
+\begin{tabular}{lccc}
+ \toprule
+ 	&  \multicolumn{3}{c}{{Genotype}} \\ 
+\cline{2-4}
+Geom.~Mean (2012 \& 2013)			& $inv6/inv6$	& $inv6/w+$		& $w+/w+$ \\ \hline
+Flower \#:								& $4.63$	& $3.91$	& $3.19$ \\	
+Flower \# $\times$ Pollen viab.:		& $1.89$	& $2.29$	& $2.70$ \\ \hline
+Rel.~Flower \#:							& $1$		& $0.85$	& $0.69$ \\	
+Rel.~Flower \# $\times$ Pollen viab.:	& $0.70$	& $0.85$	& $1$ \\	
+\hline
+\end{tabular}
+\caption{\footnotesize{Note that Values for $inv6/inv6$ homozygotes are extrapolated based on the assumption of additive effects of the inversion on flower and pollen production.}}
+\end{table}
+
 We used pollen production as a proxy for male fertility.
-We calculated average pollen production as the product of the geometric mean flower number and field estimated pollen viability for wild-type and inv6 heterozygotes, and extrapolated the expected pollen production for inv6 homozygotes.
+We calculated average pollen production as the product of the geometric mean flower number and field estimated pollen viability for wild-type and inv6 heterozygotes, and extrapolated the expected pollen production for inv6 homozygotes (table \ref{tab:MimReproEst}).
 As before, we calculated relative male fertility ($w^{\prime}_i$) for each genotype given that wild-type homozygotes had the highest pollen viability, and used the expressions for $w^{\prime}_{i}$ from Table \ref{tab:inv6Fitness} to solve for $s_m$.
-
-It is worth noting that \citet{LeeKelly2015} also report greenhouse pollen viabilty data for several mapping populations of {\itshape M.~guttatus}, which includes data for inv6 homozygotes, from which relative fertilities can be estimated.
-We chose to use the field estimated pollen viability and flower number data, despite the lack of inv6 homozygotes, because it probably gives a more accurate reflection of the fitness effects of inv6 in the field.
-Moreover, both approaches require making assumptions regarding the dominance effects of inv6.
-Given the obvious pitfalls of estimating selection on inv6 from the available data, it goes without saying that the location of inv6 in demographically viable polymorphic parameter space is a highly speculative, but nevertheless interesting, proof of concept for our model predictions.
 
@@
  \toprule
-					&  \multicolumn{3}{c}{{\textit{Genotype}}} \\ 
+					&  \multicolumn{3}{c}{{Genotype}} \\ 
 \cline{2-4}
@@
 \end{table}
+
+\citet{LeeKelly2015} also report greenhouse pollen viabilty data for several mapping populations of {\itshape M.~guttatus}, which includes data for inv6 homozygotes, from which relative fertilities can be estimated.
+We chose to use the field estimated pollen viability and flower number data, despite the lack of inv6 homozygotes, because it probably gives a more accurate reflection of the fitness effects of inv6 in the field.
+Moreover, both approaches require making assumptions regarding the dominance effects of inv6.
+Given the obvious pitfalls of estimating selection on inv6 from the available data, it goes without saying that the location of inv6 in demographically viable polymorphic parameter space is a highly speculative, but nevertheless interesting, proof of concept for our model predictions.
 

### 2021-05-25 doc/Supplements/SupplementaryMaterial.tex
Code for sex-specific dominance simulations
This commit adds functionality to several simulation and plotting
functions to explore the consequences of sex-specific dominance

@@
 \setlength{\bibsep}{0.0pt}
-\linespread{1.7}
+\linespread{1.5}
 \usepackage[utf8]{inputenc}
@@
 \setlength{\headheight}{13.6pt}
-\renewcommand{\refname}{Literature Cited}
+\renewcommand{\refname}{References}
 
@@
 %
-\rhead{Supplement to Olito \& de Vries., "Costs of SA selection" \textit{Am.~Nat.}}
+\rhead{Supplement to Olito \& de Vries, "Costs of SA selection" \textit{Am.~Nat.}}
 \setlength{\headsep}{0.3in}  
@@
 
-\title{Online Supplement: Demographic costs of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites \\ 
+\title{Online Supplement to "Demographic costs of sexually antagonistic selection in partially selfing populations \\ 
 %\LaTeX{} Template for Author-Supplied Supplementary PDFs, \\ 
@@
 % except when they are already cited in the main text.
-\bibliography{Refs2.bib}
+\bibliography{suppRefs.bib}
 

### 2021-05-25 R/functions-Figs.R

@@ -944,6 +944,366 @@ MimulusInv6Fig  <-  function() {
 ##############################################################
 ##############################################################
 
+
+demViablePolySpace_SexSpecDominance_Fig  <-  function() {
+
+    # Import data sets
+    invA       <-  read.csv("./output/simData/invasionBoundaries_sMax0.15_res0.0015_hf0.25_hm0.75_C0_delta0_dj0_da0_dg0_f5.8.csv", head=TRUE)
+    extA_f5.8  <-  read.csv("./output/simData/extThreshold_SfxSm_sMax0.15_res0.0015_hf0.25_hm0.75_C0_delta0_dj0_da0_dg0_f5.8.csv", head=TRUE)
+    extA_f6.0  <-  read.csv("./output/simData/extThreshold_SfxSm_sMax0.15_res0.0015_hf0.25_hm0.75_C0_delta0_dj0_da0_dg0_f6.csv", head=TRUE)
+    extA_f6.2  <-  read.csv("./output/simData/extThreshold_SfxSm_sMax0.15_res0.0015_hf0.25_hm0.75_C0_delta0_dj0_da0_dg0_f6.2.csv", head=TRUE)
+
+    invB       <-  read.csv("./output/simData/invasionBoundaries_sMax0.15_res0.0015_hf0.25_hm0.75_C0.25_delta0_dj0_da0_dg0_f5.8.csv", head=TRUE)
+    extB_f5.8  <-  read.csv("./output/simData/extThreshold_SfxSm_sMax0.15_res0.0015_hf0.25_hm0.75_C0.25_delta0_dj0_da0_dg0_f5.8.csv", head=TRUE)
+    extB_f6.0  <-  read.csv("./output/simData/extThreshold_SfxSm_sMax0.15_res0.0015_hf0.25_hm0.75_C0.25_delta0_dj0_da0_dg0_f6.csv", head=TRUE)
+    extB_f6.2  <-  read.csv("./output/simData/extThreshold_SfxSm_sMax0.15_res0.0015_hf0.25_hm0.75_C0.25_delta0_dj0_da0_dg0_f6.2.csv", head=TRUE)
+
+    invC       <-  read.csv("./output/simData/invasionBoundaries_sMax0.15_res0.0015_hf0.25_hm0.75_C0.5_delta0_dj0_da0_dg0_f5.8.csv", head=TRUE)
[…]

### 2021-05-25 R/functions-Simulations.R

@@ -2110,7 +2110,7 @@ titrateInvBoundaries  <-  function(sMax=0.15, res=0.0015, precision=1e-4,
 #' selfing rates, and fertility values.
 makeDataPolyParamSpace  <-  function(sMax=0.1, res=0.01, precision = 1e-4,
 									 om = 2, g = 3, theta = c(0.6,0.6,0.05,NA), theta_prime = NA, 
-									 hVals= c(1/2, 1/4), fVals = c(5.8, 5.9, 6.0, 6.5), 
+									 hfVals= c(1/2, 1/4), hmVals= c(1/2, 1/4), fVals = c(5.8, 5.9, 6.0, 6.5), 
 									 delta = 0, delta_j = 0, delta_a = 0, delta_gamma = 0,
 									 tlimit = 10^5, eqThreshold = 1e-8) {
 
@@ -2120,11 +2120,11 @@ makeDataPolyParamSpace  <-  function(sMax=0.1, res=0.01, precision = 1e-4,
 
 	# Setup progress bar
 	print('Making Data For Fig.2')
-	pb   <-  txtProgressBar(min=0, max=(length(hVals)*length(fVals)*length(Cs)), style=3)
+	pb   <-  txtProgressBar(min=0, max=(length(hfVals)*length(fVals)*length(Cs)), style=3)
 	setTxtProgressBar(pb, 0)
 
 	# Loop over dominance, fertility, selfing
-	for(i in 1:length(hVals)) {
+	for(i in 1:length(hfVals)) {
 		for(j in 1:length(fVals)) {
 
 			# assign fertility value to theta & theta_prime
@@ -2136,20 +2136,21 @@ makeDataPolyParamSpace  <-  function(sMax=0.1, res=0.01, precision = 1e-4,
 				# Find Invasion Boundaries
 				invData  <-  titrateInvBoundaries(sMax=sMax, res=res, precision=precision,
 												  om = om, g = g, theta = theta, theta_prime = theta_prime, 
-												  hf = hVals[i], hm = hVals[i], C = Cs[k], 
+												  hf = hfVals[i], hm = hmVals[i], C = Cs[k], 
 												  delta = delta, delta_j = delta_j, delta_a = delta_a, delta_gamma = delta_gamma,
 												  tlimit = tlimit, eqThreshold = eqThreshold, verbose=FALSE, writeFile=FALSE)
 
 				# Find Extinction Threshold
 				extData  <-  extinctThreshTitrate(sMax=sMax, res=res, precision=precision,
[…]

### 2021-05-25 makeFigs.R

@@ -51,6 +51,18 @@ embed_fonts(figPath(name='MimulusInv6FigNewSel.pdf'))
 # Supp. Figs
 #################
 
+# 
+toPdf(demViablePolySpace_SexSpecDominance_Fig(), 
+			figPath(name='extinctionThresholdsSexSpecDomFig.pdf'), width=10, height=5)
+embed_fonts(figPath(name='extinctionThresholdsSexSpecDomFig.pdf'))
+
+
+
+
+
+
+
+
 toPdf(suppPolySpaceThresholdFigs(), 
 			figPath(name='SuppFig-polySpace.pdf'), width=20, height=8)
 embed_fonts(figPath(name='SuppFig-polySpace.pdf'))

### 2021-05-25 run-Simulations.R

@@ -247,7 +247,7 @@ extinctThreshTitrate(sMax=0.15, res=0.0015, precision=1e-4,
 
 makeDataPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
 						om = 2, g = 3, theta = c(0.6,0.6,0.05,NA), theta_prime = NA, 
-						hVals= c(1/2, 1/4), fVals = c(5.8, 5.9, 6.0, 6.5), 
+						hfVals= c(1/2, 1/4), hmVals= c(1/2, 1/4), fVals = c(5.8, 5.9, 6.0, 6.5), 
 						delta = 0, delta_j = 0, delta_a = 0, delta_gamma = 0,
 						tlimit = 10^5, eqThreshold = 1e-8)
 
@@ -1275,3 +1275,266 @@ test  <-  fwdSimMimulusDat(datMat=datMat, theta.list=theta.list, delta.list=delt
 				 hf = 1/2, hm = 0.35, sf = 0.308, sm = 0, C = 0.24,
 				 tlimit = 10^5, eqThreshold=1e-9, Ainvade = FALSE, intInit = TRUE)
 
+
+
+
+
+##################################
+
+##########################################################
+# Supplementary Figures
+
+## Illustrating Effects of sex-specific dominance
+# Assume complementary dominance (hf = 1 - hm)
+# Fig. SX - 
+
+## hf 1/4
+## hm = 3/4 
+
+	# Obligate Outcrossing
+	# C = 0
+	pars  <-  list(
+					"hf"     =  1/4,
+					"hm"     =  3/4,
+					"C"      =  0,
+					"delta"  =  0,
+					"dj"     =  0,
+					"da"     =  0,
+					"dg"     =  0
+					)
+
+titrateInvBoundaries(sMax=0.15, res=0.0015, precision=1e-4,
+					 om = 2, g = 3, theta = c(0.6,0.6,0.05,5.8), theta_prime = 5.8, 
+					 hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, 
+					 delta_j = pars$dj, delta_a = pars$da, delta_gamma = pars$dg,
+					 tlimit = 10^3, eqThreshold = 1e-8, verbose=TRUE, writeFile=TRUE)
+
+extinctThreshTitrate(sMax=0.15, res=0.0015, precision=1e-4,
[…]

### 2021-05-31 doc/SA-Hermaphrodites-wDemography.tex
Functions for  sex-specific dominance figs
This commit adds some new functions to plot supplementary figures
illustrating how the model behaves udner sex-specific dominance.

@@
 Classic population genetics theory has been fundamental to understanding the evolution of sex-differences and the maintenance of sexually antagonistic (SA) genetic variation, but these models have rarely considered the demographic consequences of intralocus sexual antagonism.
-In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious/gonochorous) and partially selfing hermaphrodite populations.
-We focus on identifying parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the intrinsic population growth rate remains positive.
-In addition, we analyze the effects of inbreeding depression manifesting in different life-history stages, and give an illustrative example of the potential for SA polymorphism in real populations using empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus}.
-Our results indicate that as population intrinsic growth rates approach one, large swathes of polymorphic parameter space become demographically inviable, and that inbreeding depression poses a significant challenge to the maintenance of SA polymorphism in partially selfing populations.
-Despite these demographic challenges, our example with {\itshape M.
-guttatus} appears to show that demographic rates observed in some real populations are capable of sustaining large regions of viable SA polymorphic space.
+In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (i.e., dioecious/gonochorous) and partially selfing hermaphrodite populations.
+We focus on identifying parameter conditions under which SA polymorphism is maintained {\itshape and} the population growth rate remains positive.
+Additionally, we analyze the effects of inbreeding depression manifesting at different life-history stages and give an illustrative example of the potential for SA polymorphism in real populations using empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus}.
+Our results show that when population intrinsic growth rates approach one, extinction occurs across large swathes of parameter space favoring SA polymorphism or the fixation of male-beneficial alleles, and that inbreeding depression is a significant problem for maintaining SA polymorphism in partially selfing populations.
+Despite these demographic challenges, our example with {\itshape M.~guttatus} appears to show that demographic rates observed in some real populations are capable of sustaining large regions of viable SA polymorphic space.
 
@@
 
-Most population genetic models, including models of the maintenance of SA polymorphisms, keep track of the relative fitnesses of genotypes in populations of constant size, and therefore do not consider the population dynamical consequences of evolutionary change.
+Most population genetic models, including models of the maintenance of SA polymorphisms, keep track of the relative fitnesses of genotypes in populations of constant (and often infinite) size, and therefore do not consider the population dynamical consequences of evolutionary change.
 In fact, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals (e.g., \citealt {johnston2013life, merot2020balancing}).
-To take the population dynamical consequences of SA selection into account, a model is required that links changes in genotype distributions to population dynamics.
+To take into account the consequences of population dynamics under SA selection, a model is required that links changes in genotype distributions to population dynamics.
 The potential demographic costs of sexual antagonism were pointed out by \cite{kokko2003sexy}, but few papers since then explicitly incorporated demography into models of sexual antagonism (with the notable exception of \citealt{harts2014demography}).
@@
 
-In this paper, we extend the model of \citep{deVriesCaswell2019a,deVriesCaswell2019b} to include both dioecious and hermaphroditic species in order to study the parameter conditions under which balanced SA polymorphisms are also demographically viable.
+In this paper, we extend the model of \citep{deVriesCaswell2019a,deVriesCaswell2019b} to include both dioecious and hermaphroditic species in order to study the parameter conditions under which different outcomes of SA selection, including fixation of male- or female-beneficial alleles as well as balanced polymorphisms, are also demographically viable.
 A major strength of our approach is that the model can be parameterized using empirically estimated demographic rates, enabling us to make predictions about the scope for demographically viable SA polymorphisms that are grounded in the biology of real populations.
@@
 
-Simultaneous hermaphrodites express both male and female sex functions (i.e., can transmit genes to the next generation via both sperm and ovules), and have the potential to reproduce by a combination of self- and outcross-fertilizations.
+Simultaneous hermaphrodites express both male and female sex functions (i.e., can transmit genes to the next generation via both sperm/pollen and eggs/ovules), and have the potential to reproduce by a combination of self- and outcross-fertilizations.
 Maternal outcrossing involves receiving male gametes from another individual in the population, while paternal outcrossing involves exporting male gametes to another individual and fertilizing their ovules.
@@
 Following previous theory for partially selfing populations \cite[e.g.,][]{Charlesworth2010,JordanConnallon2014,Glemin2021}, we use a 'fixed prior selfing' model where all individuals are assumed to self-fertilize at a constant rate, $C$, prior to receiving outcross pollen, for reasons of analytic tractability.
-\textcolor{red}{In the \hl{Online Supplement}, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
-Do we???}
+In the \hl{Online Supplement}, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
 
@@
 Because single-locus selection coefficients are generally weak (e.g., \citealt{Eyre-WalkerKeightly2007}) and strongly skewed, we limit our analyses to coefficients within $0 < s_f,s_m \leq 0.15$, unless stated otherwise.
-For simplicity, we present results for scenarios of equal dominance in the main text (i.e., $h_f = h_m = h$).
-Specifically, we examine scenarios of ({\itshape i}) additive SA fitness effects ($h = 1/2$); and ({\itshape ii}) dominance reversals, where the deleterious fitness effect of each SA allele is partially recessive in each sex ($h = 1/4$).
-These scenarios are of particular interest because additive effects are commonly observed for small-effect alleles or quantitative traits \citep{AgrawalWhitlock2011}, while dominance reversals are predicted under fitness landscape models provided the population is not too far from the phenotypic optimum \citep{Manna2011, ConnallonClark2014}.
-\textcolor{red}{We explore the consequences of sex-specific dominance, which may evolve under asymmetric SA selection \citep{SpencerPriest2016} in \hl{Appendix X}).}
+We present scenarios of equal dominance in the main text (i.e., $h_f = h_m = h$).
+Specifically, we examine scenarios of ({\itshape i}) additive SA fitness effects ($h = 1/2$), which are commonly observed for small-effect alleles or quantitative traits \citep{AgrawalWhitlock2011}; and ({\itshape ii}) dominance reversals, where the deleterious fitness effect of each SA allele is partially recessive in each sex ($h = 1/4$), which are predicted under fitness landscape models provided the population is not too far from the phenotypic optimum \citep{Manna2011, ConnallonClark2014}.
+We briefly explore the consequences of sex-specific dominance, which may evolve under asymmetric SA selection \citep{SpencerPriest2016} in \hl{Appendix X}).
 
@@
 To determine the viability threshold (where the population growth rate equals one) and the boundary stability thresholds (where the largest eigenvalue of the Jacobian crosses one), we start at the outer boundaries of selection parameter space, and titrate inward.
-The proportions of total $s_f \times s_m$ parameter space where SA polymorphism and/or extinction occurred were calculated by numerically integrating the appropriate regions defined by the invasion and extinction thresholds
-
-For analyses of the demographic effects of inbreeding depression, we make two additional simplifying assumptions: First, we assume that if inbreeding depression is caused primarily by recessive deleterious mutations (as suggested by empirical data), it should covary negatively with the population selfing rate due to purging (\citealt{Charlesworth2009}; though we note that other processes could give rise to this pattern, e.g., \citealt{CrnokrakBarrett2002, Charlesworth2009,HedrickGarcia-Dorado2016}).
+The proportions of total $s_f \times s_m$ parameter space where SA polymorphism and/or extinction occurred were calculated by numerically integrating the appropriate regions defined by the invasion and extinction thresholds.
+
+For analyses of the demographic effects of inbreeding depression, we make two additional simplifying assumptions: First, we assume that if inbreeding depression is caused primarily by recessive deleterious mutations (as suggested by empirical data), it should covary negatively with the population selfing rate due to purging, provided the population selfing rate has been relatively constant in recent evolutionary time (\citealt{Charlesworth2009}; though we note that other processes could give rise to this pattern, e.g., \citealt{CrnokrakBarrett2002, Charlesworth2009,HedrickGarcia-Dorado2016}).
 Following \citet{Olito2019}, we incorporate such negative covariance by constraining the inbreeding depression terms in the model ($\delta$ and $\delta_i$, where $i \in \{j,a,\gamma\}$) to follow a simple declining function of the selfing rate: $\delta = \delta^{\ast} (1 - b (1 - L))$, where $\delta^{\ast}$ is the hypothetical severity of inbreeding depression for a completely outcrossing population, $b$ is a shape parameter determining how far $\delta$ will decline under complete selfing (when $C = 1$), and $L$ describes the expected deleterious mutation load as a function of the selfing rate $C$.
@@
 The location of the extinction threshold, where the population intrinsic growth rate $\lambda = 1$, is primarily determined by the fertility parameter ($f$) but is also influenced by the population selfing rate ($C$) and dominance of the SA alleles ($h$).
-When $f$ is large, the population can experience stronger selection on fertility through both male and female fitness components before extinction occurs (the extinction threshold shifts to higher values of $s_f$ and $s_m$).
+As $f$ increases, the population can survive in the face of stronger fertility selection through both male and female sex-functions before extinction occurs (the extinction threshold shifts to higher values of $s_f$ and $s_m$).
 
@@
 Inbreeding effects on juvenile survival had the strongest effect on population viability because on average individuals will spend multiple time steps in the juvenile stage before they mature.
-At each time step in the juvenile stage, individuals have a probability $\gamma$ to mature and a probability $\sigma_j(1-\delta_j)$ to survive.
+At each time, juveniles have a probability $\gamma$ to mature and a probability $\sigma_j(1-\delta_j)$ to survive.
 Inbreeding depression at the juvenile stage therefore makes it harder to survive long enough to mature.
 Although early-acting inbreeding depression ($\delta$) actually manifests earlier in the life-cycle than juvenile survival, it acts only once by influencing the total number of self-fertilized zygotes that become juveniles.
+Note, however, that the relative strength of inbreeding depression at different stages of the life history will depend on the interaction between the various inbreeding depression parameters, which we have precluded from our analyses.
 
@@
 inv6 segregates at moderate frequency (about $8\%$), and carriers suffer $\approx 30 \%$ loss in pollen viability, but also increased flower production.
-The genetic basis of these effects remain unclear, but the net result is a "supergene" with strong fitness effects on both female and male fertility that segregates as a single diallelic locus.
+The genetic basis of these effects remain unclear -- though they likely polygenic -- but the net result is a "supergene" with strong fitness effects on both female and male fertility that segregates as a single diallelic locus.
 We estimated selection coefficients for the effect of inv6 on pollen production (i.e., taking into account the simultaneous effect on flower number) and ovule production from the data reported in \citet{LeeKelly2015} under the relatively conservative assumptions that pollen/ovule production is proportional to flower number, and  additive SA fitness effects (dominance coefficients could not be estimated from field data; see Appendix \hl{C} for details of estimating selection coefficients).
@@
 Results are shown for two hypothetical populations using demographic rates for locally adapted (Eagle Meadows; panel A), and non-local (Low-Elevation Perennial; panel B) populations reported by \citet{PetersonEtAl2016}.
-Invasion conditions and extinction thresholds were calculated using the conservative assumption of additive fitness effects in both sexes ($h_f = h_m = 0.5$) for two parameter conditions: obligate outcrossing (solid black lines)  and partial selfing with inbreeding depression using the highest field-estimate of selfing, and inbreeding depression parameters ($\delta_i$) calculated for the Iron Mountain population of \citet{Willis1993} (dashed lines).
+Invasion conditions and extinction thresholds were calculated using the conservative assumption of additive fitness effects in both sexes ($h_f = h_m = 0.5$) for two parameter conditions: obligate outcrossing (solid black lines, dark red shaded region)  and partial selfing with inbreeding depression using the highest field-estimate of selfing, and inbreeding depression parameters ($\delta_i$) calculated for the Iron Mountain population of \citet{Willis1993} (dashed lines, light red shaded region).
 The location of inv6 is also shown on both plots, using selection coefficients calculated from field estimates of male and female fitness components from \citet{LeeKelly2015} under the relatively conservative assumption of additive fitness effects in both sexes, where $(s_f,\,s_m) = (0.30,\, 0.31)$.
@@
 Classic population genetics theory predicts that sexually antagonistic selection is unlikely to maintain genetic variation except under narrow conditions, with polymorphism requiring either finely balanced or unusually strong selection, or partially recessive fitness effects through each sex \citep{Kidwell1977,Pamilo1979,Prout2000, ConnallonClark2014}.
-Extensions of the theory have identified numerous ways in which the conditions for polymorphism become more permissive in both dioecious and hermaphroditic organisms, including genetic linkage of SA loci, the evolutioin of sex-specific dominance, population subdivision, and life-cycle complexity \citep[e.g.,][]{Patten2010,JordanConnallon2014,JordanCharlesworth2012,SpencerPriest2016, ConnallonSharmaOlito2019, Olito-etal-2018}.
+Extensions of the theory have identified numerous ways in which the conditions for polymorphism become more permissive in both dioecious and hermaphroditic organisms, including genetic linkage of SA loci, the evolution of sex-specific dominance, population subdivision, and life-cycle complexity \citep[e.g.,][]{Patten2010,JordanConnallon2014,JordanCharlesworth2012,SpencerPriest2016, ConnallonSharmaOlito2019, Olito-etal-2018}.
 However, by ignoring the demographic consequences of SA genetic variation, these population genetic models have missed the possibility that SA polymorphisms may not be viable under realistic parameter conditions, and therefore unlikely to be observed in natural populations.
@@
 Since growth rates far exceeding one suggest rapid exponential growth, they are generally rare (with the notable exception of recently introduced, invasive populations), suggesting that our model predictions may be highly relevant for SA polymorphism in many real-world populations.
-Moreover, these findings compliment recent theoretical and empirical studies indicating that SA selection is likely to be both condition dependent, and stronger in locally-adapted populations near the center of a species' range, where population growth rates are expected to be high \citep{BergerEtAl2014, Connallon2015}.
+Moreover, these findings complement recent theoretical and empirical studies indicating that SA selection is likely to be both condition dependent, and stronger in locally-adapted populations near the center of a species' range, where population growth rates are expected to be high \citep{BergerEtAl2014, Connallon2015}.
 
@@
 This key prediction is supported by a series of experimental results in seed beetles ({\itshape Callosobruchus maculatus}) which show that male-beneficial SA genotypes are less likely to contribute positively to population growth rates, and are more susceptible to extinction under environmental stress or inbreeding \citep{BergerEtAl2014, berger2016intralocus, grieshop2017male}.
-An interesting corrollary is that, since strong SA fitness effects, especially those affecting female fertility, often lead to extinction, the SA genetic variation we are able to observe in natural populations may often be under weak selection, and therefore strongly susceptible to genetic drift whether or not selection favours the fixation of one allele or balancing selection \citep{ConnallonClark2012}.
+\textcolor{red}{Jessica's comment Re: sex-limited exp.
+evol.}.
+An interesting corrollary of our findings is that, since strong SA fitness effects often lead to extinction, the observable SA genetic variation in natural populations may often be under weak selection, and therefore strongly susceptible to genetic drift whether or not selection favours the fixation of one allele or balancing selection \citep{ConnallonClark2012}.
 
@@
 In our model, this reduction in effective selfing is accompanied by significant mortality due to inbreeding depression (whether early- or late-acting), which can quickly tip partially-selfing populations over the brink to extinction.
-Beyond the maintenance of SA polymorphisms, this finding underscores a simple but important point that is often overlooked in studies of the evolution of self-fertilization and selfing-syndromes, which tend to emphasize the coevolution of the deleterious mutation load and mating system \citep[e.g.,][]{Charlesworth1987, LandeSchemske1985, Goodwillie2005}: populations with high fertility rates are better able to sustain positive growth rates in the face of severe demographic costs of inbreeding.
-This suggests that traits related to female fecundity, such as ovule and flower production, in ancestral populations likely play an important role in determining the distribution of transitions to self-fertilization among hermaphroditic taxa as well as a variety of ecological correlates of selfing and mixed-mating \citep{Goodwillie2005,Igic2006,Grossenbacher2015}.
+Beyond the maintenance of SA polymorphisms, this finding underscores a simple but important point that is often overlooked in studies of the evolution of self-fertilization and selfing-syndromes, which tend to emphasize the coevolution of the deleterious mutation load and mating system \citep[e.g.,][]{Charlesworth1987, LandeSchemske1985, Goodwillie2005}: highly fertile populations can better afford the severe demographic costs of inbreeding.
+This suggests that traits related to female fecundity, such as ovule and flower production, may strongly influence the distribution of successful transitions to self-fertilization among hermaphroditic taxa, as well as a variety of ecological correlates of selfing and mixed-mating \citep{Goodwillie2005, Igic2006, Grossenbacher2015}.
 
@@
 Interestingly, different forms of self-fertilization can aid in reproductive assurance, and should therefore have demographic consequences -- for example, under pollen limitation and delayed selfing (where only ovules that fail to receive outcross pollen are selfed), the selfing rate will be a function of genotype frequencies because these will directly influence pollen production \citep{HarderBarrett2006}.
-,
+
 Another major simplifying assumption in our model was that how individuals themselves were produced (i.e., by selfing or outcrossing) affects the level of inbreeding depression they suffer.
@@
 In each case, we calculated the inbreeding depression terms as $\delta_i = 1 - w_{\text{inbred}}/w_{\text{outcross}}$.
-To incorporate these terms into our evolutionary demographic model, we multiplied each of the three terms $G$, $F$, and $S$ in Eq(\ref{eq:AtildeMimulus}) by a corresponding proportional inbreeding depression term $(1 - \delta_i)$, where $i \{\in G,F,S\}$.
+To incorporate these terms into our evolutionary demographic model, we multiplied each of the three terms $G$, $F$, and $S$ in Eq(\ref{eq:AtildeMimulus}) by a corresponding proportional inbreeding depression term $(1 - \delta_i)$, where $i \in \{ G,F,S\}$.
 The largest field-estimated selfing rate for this same Iron Mountain population was $C = 0.29$ \citep{Willis1993}, which we use in our calculations for fig.~\ref{fig:mimulusFig}.

### 2021-05-31 R/functions-Figs.R

@@ -1089,7 +1089,7 @@ layout     <- layout(layout.mat,respect=TRUE)
         # axes        
         axis(1, las=1, labels=NA)
         axis(2, las=1)
-        proportionalLabel(-0.5, 0.5, expression(paste(italic(h[f]), " = 3/4, ", italic(h[m]), " = 1/4")), cex=1.5, adj=c(0.5, 0.5), xpd=NA, srt=90)
+        proportionalLabel(-0.5, 0.5, expression(paste(italic(h[f]), " = 1/4, ", italic(h[m]), " = 3/4")), cex=1.5, adj=c(0.5, 0.5), xpd=NA, srt=90)
         proportionalLabel(-0.3, 0.5, expression(paste(italic(s[f]))), cex=1.2, adj=c(0.5, 0.5), xpd=NA, srt=90)
         proportionalLabel(0.5, 1.15, expression(paste(italic(C), ' = 0')), cex=1.5, adj=c(0.5, 0.5), xpd=NA)
         proportionalLabel(0.04, 1.075, 'A', cex=1.2, adj=c(0.5, 0.5), xpd=NA)
@@ -1292,6 +1292,360 @@ layout     <- layout(layout.mat,respect=TRUE)
 
 
 
+#' Fig showing Proportion of polymorphic parameter space
+#' (w/out inbreeding depression)
+polySpaceFigTitrateSexSpec  <-  function(df = "dataPolySpaceFig_hf0.25_0.75_hm0.75_0.25_sMax0.15_res0.003_delta0_dj0_da0_dg0") {
+
+    # Make filenames for import from df names
+    fName  <-  paste('./output/simData/', df, '.csv', sep="")
+
+    # Extract plotting parameter values from df names
+    df1  <-  strsplit(df, '_')[[1]][c(2:11)]
+    pars1  <-  list(
+                    "hf"    =  as.numeric(strsplit(df1[1],'f')[[1]][2]),
+                    "hm"    =  as.numeric(strsplit(df1[3],'m')[[1]][2]),
+                    "sMax"  =  as.numeric(strsplit(df1[5],'x')[[1]][2]),
+                    "res"   =  as.numeric(strsplit(df1[6],'s')[[1]][2]),
+                    "d"     =  as.numeric(strsplit(df1[7],'a')[[1]][2]),
+                    "dj"    =  as.numeric(strsplit(df1[8],'j')[[1]][2]),
[…]

### 2021-05-31 R/functions-Simulations.R

@@ -2169,7 +2169,7 @@ makeDataPolyParamSpace  <-  function(sMax=0.1, res=0.01, precision = 1e-4,
 	colnames(dataSet)  <-  c("hf", "hm", "f", "C", "smInv", "aInv", "AInv", "sf", "smExt", "sm", "sfExt")
 
 	# Export data as .csv to ./output/data
-	filename <-  paste("./output/simData/dataPolySpaceFig", "_hf", hfVals[1], hfVals[2], "_hm", hmVals[1], hmVals[2],, "_sMax", sMax, "_res", res, "_delta", delta, "_dj", delta_j, "_da", delta_a, "_dg", delta_gamma, ".csv", sep="")
+	filename <-  paste("./output/simData/dataPolySpaceFig", "_hf", hfVals[1], "_", hfVals[2], "_hm", hmVals[1], "_", hmVals[2], "_sMax", sMax, "_res", res, "_delta", delta, "_dj", delta_j, "_da", delta_a, "_dg", delta_gamma, ".csv", sep="")
 	write.csv(dataSet, file=filename, row.names = FALSE)
 
 }
@@ -2182,10 +2182,13 @@ makeDataPolyParamSpace  <-  function(sMax=0.1, res=0.01, precision = 1e-4,
 #' 
 #' Parameters
 #' df:	polySpace dataframe
-cleanPolySpaceData  <-  function(data) {
+cleanPolySpaceData  <-  function(data, sexSpecDom = FALSE) {
 
     # Get plotting levels from df
-    hLev  <-  unique(data$h)
+	if(sexSpecDom) {
+	    hLev  <-  unique(data$hf)
+	}
+    else{hLev  <-  unique(data$h)}
     fLev  <-  unique(data$f)
     CLev  <-  unique(data$C)
     nHs   <-  length(hLev)
@@ -2198,7 +2201,12 @@ cleanPolySpaceData  <-  function(data) {
         for(j in 1:nFs) {
 			for(k in 1:nCs) {
 				# Subset data by Dominance
-				d  <-  data[data$h == hLev[i],]            
+				if(sexSpecDom) {
+					d  <-  data[data$hf == hLev[i],]
+				}
+				else{
+					d  <-  data[data$h == hLev[i],]
+				}
 				# Subset data by Fertility Value
 				d  <-  d[d$f == fLev[j],]     
 				# Subset data by Selfing Rate
@@ -2324,6 +2332,92 @@ quantPolySpace  <-  function(data, pars) {
 }
 
 
+
+
+##############################
[…]

### 2021-05-31 makeFigs.R

@@ -51,14 +51,16 @@ embed_fonts(figPath(name='MimulusInv6FigNewSel.pdf'))
 # Supp. Figs
 #################
 
-# 
+# SuppFig. -- Illustration of Invaions & Extinction thresholds for sex-specific dominance
 toPdf(demViablePolySpace_SexSpecDominance_Fig(), 
 			figPath(name='extinctionThresholdsSexSpecDomFig.pdf'), width=10, height=5)
 embed_fonts(figPath(name='extinctionThresholdsSexSpecDomFig.pdf'))
 
+# SuppFig. -- Quantification of demographically viable parameter space, no inbreeding depression
 
-
-
+toPdf(polySpaceFigTitrateSexSpec(), 
+			figPath(name='polymorphicSpaceTitrateSexSpec.pdf'), width=5, height=7)
+embed_fonts(figPath(name='polymorphicSpaceTitrateSexSpec.pdf'))
 
 
 

### 2021-05-31 run-Simulations.R

@@ -1505,7 +1505,8 @@ extinctThreshTitrate(sMax=0.15, res=0.0015, precision=1e-4,
 
 
 ##########################################################
-# Fig. 2 - Dem. viable polymorphic param. space (no I.D.)
+# Fig. SX - Dem. viable polymorphic param. space (no I.D.)
+#			Sex-specific dominance
 
 makeDataPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
 						om = 2, g = 3, theta = c(0.6,0.6,0.05,NA), theta_prime = NA, 
@@ -1520,21 +1521,15 @@ makeDataPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
 
 makeDataDeltaPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
 							om = 2, g = 3, theta = c(0.6,0.6,0.05,6.5), theta_prime = 6.5, 
-							hVals= c(1/2, 1/4), dStar = 0.8, 
-							tlimit = 10^5, eqThreshold = 1e-8)
-
-
-makeDataDeltaPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
-							om = 2, g = 3, theta = c(0.6,0.6,0.05,7), theta_prime = 7, 
-							hVals= c(1/2, 1/4), dStar = 0.8, 
+							hfVals= c(1/4, 3/4), hmVals= c(3/4, 1/4), dStar = 0.8, 
 							tlimit = 10^5, eqThreshold = 1e-8)
 
 makeDataDeltaPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
 							om = 2, g = 3, theta = c(0.6,0.6,0.05,7.5), theta_prime = 7.5, 
-							hVals= c(1/2, 1/4), dStar = 0.8, 
+							hfVals= c(1/4, 3/4), hmVals= c(3/4, 1/4), dStar = 0.8, 
 							tlimit = 10^5, eqThreshold = 1e-8)
 
 makeDataDeltaPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
 							om = 2, g = 3, theta = c(0.6,0.6,0.05,8.5), theta_prime = 8.5, 
-							hVals= c(1/2, 1/4), dStar = 0.8, 
+							hfVals= c(1/4, 3/4), hmVals= c(3/4, 1/4), dStar = 0.8, 
 							tlimit = 10^5, eqThreshold = 1e-8)

### 2021-06-03 doc/SA-Hermaphrodites-wDemography.tex
Minor edits
Edited first paragraph of Results to hopefully hold demographers by the hand a bit more when looking at the funnel plots, and as a consequence moved some info around in the results.

Shortened methods by one paragraph.

Added few sentences to discussion.

@@
 
-For a given parameter set, we numerically evaluate the largest eigenvalue of the Jacobian evaluated at both boundaries, and the population growth rate of the equilibrium obtained after perturbation of the boundary, $\lambda$.
+% For a given parameter set, we numerically evaluate the largest eigenvalue of the Jacobian evaluated at both boundaries, and the population growth rate of the equilibrium obtained after perturbation of the boundary, $\lambda$.
 To determine the viability threshold (where the population growth rate equals one) and the boundary stability thresholds (where the largest eigenvalue of the Jacobian crosses one), we start at the outer boundaries of selection parameter space, and titrate inward.
@@
 
-For analyses of the demographic effects of inbreeding depression, we make two additional simplifying assumptions: First, we assume that if inbreeding depression is caused primarily by recessive deleterious mutations (as suggested by empirical data), it should covary negatively with the population selfing rate due to purging, provided the population selfing rate has been relatively constant in recent evolutionary time (\citealt{Charlesworth2009}; though we note that other processes could give rise to this pattern, e.g., \citealt{CrnokrakBarrett2002, Charlesworth2009,HedrickGarcia-Dorado2016}).
+%%% LdV: We say more or less the same thing two paragraphs earlier, so I think we don't need to explain this here anymroe?
+
+For analyses of the demographic effects of inbreeding depression, we make two additional simplifying assumptions: First, to keep our analyses tractable we explore the effects of individual inbreeding depression terms in isolation.
+That is, we assume that only one of the $\delta$ and $\delta_i$ terms (where $i \in \{j,a,\gamma\}$) can be non-zero at a time.
+Second, we assume that if inbreeding depression is caused primarily by recessive deleterious mutations (as suggested by empirical data), it should covary negatively with the population selfing rate due to purging, provided the population selfing rate has been relatively constant in recent evolutionary time (\citealt{Charlesworth2009}; though we note that other processes could give rise to this pattern, e.g., \citealt{CrnokrakBarrett2002, Charlesworth2009,HedrickGarcia-Dorado2016}).
 Following \citet{Olito2019}, we incorporate such negative covariance by constraining the inbreeding depression terms in the model ($\delta$ and $\delta_i$, where $i \in \{j,a,\gamma\}$) to follow a simple declining function of the selfing rate: $\delta = \delta^{\ast} (1 - b (1 - L))$, where $\delta^{\ast}$ is the hypothetical severity of inbreeding depression for a completely outcrossing population, $b$ is a shape parameter determining how far $\delta$ will decline under complete selfing (when $C = 1$), and $L$ describes the expected deleterious mutation load as a function of the selfing rate $C$.
@@
 We set $\delta^{\ast} = 0.8$, $b = 0.5$, and $a = 0.2$ for all analyses, values chosen to be consistent with empirical estimates of inbreeding depression (e.g., fig.~2 in \citealt{HusbandSchemske1996}).
-Second, to keep our analyses tractable we explore the effects of individual inbreeding depression terms in isolation.
-That is, we assume that only one of the $\delta$ and $\delta_i$ terms (where $i \in \{j,a,\gamma\}$) can be non-zero at a time.
 
@@
 We begin with an illustration of demographically viable polymorphic parameter space in the absence of inbreeding depression in Figure \ref{fig:extThresholds} (i.e., $\delta = \delta_i = 0$).
+The red areas in Figure \ref{fig:extThresholds} indicate areas where the population goes extinct (population growth rate is less than one) for different values of the fertility parameter.
+Lower female fertility values correspond to a larger demographically inviable area.
+The funnel between the thick black lines indicates the area where the sexually antagostic alleles coexist.
+Above the funnel, the female-benefit and male-detriment allele fixates; below the funnel, the male-benefit and female-detriment allele fixates.
+Much of the region where the male-benefit and female-detriment allele fixates is demographically inviable due to the demographic consequences of reduced female fitness in this region.
+The SA polymorphisms that remain viable at lower fertility values correspond to regions where the female-deleterious allele is predicted to segregate at low frequencies, or in regions with weak selection through both sex functions, an asymmetry that reflects the assumption of female demographic dominance.
+
 Invasion conditions for SA alleles in the evolutionary demographic model closely match the predictions from population genetic models \citep{Kidwell1977, JordanConnallon2014, Olito2017}.
@@
 The location of the extinction threshold, where the population intrinsic growth rate $\lambda = 1$, is primarily determined by the fertility parameter ($f$) but is also influenced by the population selfing rate ($C$) and dominance of the SA alleles ($h$).
-As $f$ increases, the population can survive in the face of stronger fertility selection through both male and female sex-functions before extinction occurs (the extinction threshold shifts to higher values of $s_f$ and $s_m$).
-
-The demographic consequences of SA selection are driven by the fitness effects of male-beneficial/female-deleterious alleles because they directly influence individuals' female fecundity, and therefore population growth rates.
+
+% I moved this information up. 
+%The demographic consequences of SA selection are driven by the fitness effects of male-beneficial/female-deleterious alleles because they directly influence individuals' female fecundity, and therefore population growth rates.
 For the fertility parameter values we explored, all extinctions occured in regions where either the female-deleterious allele segregates at intermediate frequency (i.e., where SA polymorphism is maintained), or goes to fixation (fig.~\ref{fig:extThresholds}).
@@
 Since growth rates far exceeding one suggest rapid exponential growth, they are generally rare (with the notable exception of recently introduced, invasive populations), suggesting that our model predictions may be highly relevant for SA polymorphism in many real-world populations.
+In addition, we find that much of the parameter space where a male-beneficial allele fixates is demographically inviable for populations whose growth rates are close to one prior to invasion.
+This demographic consequence of masculinization is rarely considered \citep{hitchcock2020gene}.
 Moreover, these findings complement recent theoretical and empirical studies indicating that SA selection is likely to be both condition dependent, and stronger in locally-adapted populations near the center of a species' range, where population growth rates are expected to be high \citep{BergerEtAl2014, Connallon2015}.
@@
 This key prediction is supported by a series of experimental results in seed beetles ({\itshape Callosobruchus maculatus}) which show that male-beneficial SA genotypes are less likely to contribute positively to population growth rates, and are more susceptible to extinction under environmental stress or inbreeding \citep{BergerEtAl2014, berger2016intralocus, grieshop2017male}.
-\textcolor{red}{Jessica's comment Re: sex-limited exp.
-evol.}.
 An interesting corrollary of our findings is that, since strong SA fitness effects often lead to extinction, the observable SA genetic variation in natural populations may often be under weak selection, and therefore strongly susceptible to genetic drift whether or not selection favours the fixation of one allele or balancing selection \citep{ConnallonClark2012}.
+
+%Comment Jess:  Many sex-limited experimental evolution studies have found a stronger response to male-limited selection compared to female-limited selection.
+I think people often think that this is because sexual selection is expected to be stronger in males.
+But it seems like these results support other explanations, e.g.
+that fixation of male-benefit alleles is only possible in a lab setting where population growth rates are controlled, or maybe that female-benefit alleles should fix more easily and therefore be less likely to segregate in the population.
 

### 2021-06-04 doc/SA-Hermaphrodites-wDemography.tex
Pre-Submission Edits
This commit adds final edits to the main manuscript and online Supp.
Mat., includes new results for complementary sex-specific
dominance in the Supp. Mat., and removes author information for
double-blind peer review.

@@
 
-\author{Colin Olito$^{1,\ast}$ \\ 
-Charlotte de Vries$^{2}$}
+%\author{Colin Olito$^{1,\ast}$ \\ 
+%Charlotte de Vries$^{2}$}
 \date{\today}
@@
 
-\noindent{} 1. Department of Biology, Lund University, Lund 223 62, Sweden;
-
-\noindent{} 2.  Department of Evolutionary Biology and Environmental Studies, University of Zurich, Zurich CH-8057, Switzerland;
-
-\noindent{} $\ast$ Corresponding author; e-mail: colin.olito@gmail.com
+%\noindent{} 1. Department of Biology, Lund University, Lund 223 62, Sweden;
+
+%\noindent{} 2.  Department of Evolutionary Biology and Environmental Studies, University of Zurich, Zurich CH-8057, Switzerland;
+
+%\noindent{} $\ast$ Corresponding author; e-mail: colin.olito@gmail.com
 
@@
 \newpage{}
-\section*{Abstract -- currently 200/200 word max.}%\textcolor{green}{ I like this abstract best (CdV)} \textcolor{blue}{Agreed (CJO)}}
+\section*{Abstract}%\textcolor{green}{ I like this abstract best (CdV)} \textcolor{blue}{Agreed (CJO)}}
 
@@
 Most population genetic models, including models of the maintenance of SA polymorphisms, keep track of the relative fitnesses of genotypes in populations of constant (and often infinite) size, and therefore do not consider the population dynamical consequences of evolutionary change.
-In fact, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals (e.g., \citealt {johnston2013life, merot2020balancing}).
+In reality, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals (e.g., \citealt {johnston2013life, merot2020balancing}).
 To take into account the consequences of population dynamics under SA selection, a model is required that links changes in genotype distributions to population dynamics.
-The potential demographic costs of sexual antagonism were pointed out by \cite{kokko2003sexy}, but few papers since then explicitly incorporated demography into models of sexual antagonism (with the notable exception of \citealt{harts2014demography}).
+The potential demographic costs of sexual antagonism were pointed out by \cite{kokko2003sexy}, but few papers since then have explicitly incorporated demography into models of sexual antagonism (with the notable exception of \citealt{harts2014demography}).
 \citet{deVriesCaswell2019a} introduced a mendelian matrix model with intralocus sexual antagonism, and population dynamics but did not perform an analysis of the consequences of demographic viability for the scope of sexual antagonism to maintain polymorphism.
@@
 In fact, the model reduces to the stage-structured model of \citet{deVriesCaswell2019a} under obligate outcrossing.
-A full derivation of the model and analyses is presented in the \hl{Online Supplementary Material}, and all computer code necessary to reproduce the results are available at \url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
-
-Simultaneous hermaphrodites express both male and female sex functions (i.e., can transmit genes to the next generation via both sperm/pollen and eggs/ovules), and have the potential to reproduce by a combination of self- and outcross-fertilizations.
+A full derivation of the model and analyses is presented in the Online Supplementary Material, and all computer code necessary to reproduce the results are available at \url{GitHub Link omitted for anonymity}.%\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
+
+Simultaneous hermaphrodites can transmit genes to the next generation via both sperm/pollen and eggs/ovules, and have the potential to reproduce by a combination of self- and outcross-fertilizations.
 Maternal outcrossing involves receiving male gametes from another individual in the population, while paternal outcrossing involves exporting male gametes to another individual and fertilizing their ovules.
 Self-fertilization is achieved when an individual's male gametes fertilize their own ovules.
-To distinguish between parameters relating to male and female function, we denote matrices or vectors relating to male sex function with a prime.
+To distinguish between parameters relating to male and female function, we denote matrices or vectors relating to the male sex function with a prime.
 
@@
 The population projection matrix $\tilde{\mbf{A}}$ is constructed from four sets of matrices representing the demographic and genetic processes: The matrices $\mbf{U}^{S}_{i}$ and $\mbf{U}^{X}_{i}$ contain transition and survival probabilities for each genotype, produced by selfing and outcrossing respectively.
-The matrices $\mbf{F}_{i}$ and $\mbf{F}^{\prime}_{i}$ contain the genotype $\times$ stage specific contribution via female and male sex-functions of genotype $i$ to the female and male gamete pools, respectively, and therefore to zygotes in the next generation.
+The matrices $\mbf{F}_{i}$ and $\mbf{F}^{\prime}_{i}$ contain the genotype $\times$ stage specific contribution of genotype $i$ to the female and male gamete pools, respectively, and therefore to zygotes in the next generation.
 We assume that whether individuals were produced through selfing or outcrossing does not affect their fecundity or mating success, that is, we assume  $\mbf{F}^X_{i}=\mbf{F}^S_{i}=\mbf{F}_{i}$ and $\mbf{F}^{\prime X}_{i}=\mbf{F}^{\prime S}_{i}=\mbf{F}^{\prime}_{i}$.
@@
  Outcrossing in our model proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}.
-That is, each individual's genotype determines both the number of ovules produced and the paternal "mating success", broadly defined.
+That is, each individual's genotype determines both the number of ovules produced and the paternal mating success, broadly defined.
 For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986, WangBarrett2020, Harder2016}.
-Note, however, that by modeling paternal relative mating success rather than pollen/sperm production, we implicitly assume female demographic dominance (i.e., production/transport of male gametes does not limit ovule fertilization and hence population growth).
+Note, however, that by modeling paternal relative mating success rather than pollen/sperm production, we implicitly assume female demographic dominance (i.e., production/transport of male gametes does not limit ovule fertilization and hence population growth), a point we return to in the Discussion.
 
@@
 Following previous theory for partially selfing populations \cite[e.g.,][]{Charlesworth2010,JordanConnallon2014,Glemin2021}, we use a 'fixed prior selfing' model where all individuals are assumed to self-fertilize at a constant rate, $C$, prior to receiving outcross pollen, for reasons of analytic tractability.
-In the \hl{Online Supplement}, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
+In the Online Supplement, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
 
@@
 \subsubsection*{Population projection}
-Using the component matrices described above (the survival matrices, $\mathbf{U}^X_i$, $\mathbf{U}^S_i$, the fertility and mating success matrices, $\mathbf{F}_i$, $\mathbf{F}^\prime_i$, and the parent-offspring matrices $\mbf{H}^S(\tilde{\mbf{n}})$ and $\mbf{H}^X$), we construct the population projection matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ using the vec-permutation approach of \citet{CaswellEtAl2018}, see \hl{Online Supplementary Material} for the step by step construction of the model.
+Using the component matrices described above (the survival matrices, $\mathbf{U}^X_i$, $\mathbf{U}^S_i$, the fertility and mating success matrices, $\mathbf{F}_i$, $\mathbf{F}^\prime_i$, and the parent-offspring matrices $\mbf{H}^S(\tilde{\mbf{n}})$ and $\mbf{H}^X$), we construct the population projection matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ using the vec-permutation approach of \citet{CaswellEtAl2018}, see Online Supplementary Material for the step by step construction of the model.
 The matrix that projects the eco-evolutionary dynamics is:
@@
 \noindent where $q^{\prime}_A$ and $q^{\prime}_a$ are given by Eq(\ref{eq:maleGametePool}).
-Further details of the derivations of $\mcal{U}^S$, $\mcal{U}^X$ ,  $\mcal{F}^S$, and $\mcal{F}^X$ are provided in the \hl{Online Supplement}.
 
@@
 
-Combining all the component matrices yields the overall eco-evolutionary projection matrix shown in \hl{Appendix A}.
+Combining all the component matrices yields the overall eco-evolutionary projection matrix shown in Appendix A.
 
@@
 $\delta,\, \delta_j,\,\delta_a,\,\delta_{\gamma}$ & Inbreeding depression terms for ovule viability, survival, and transition rates respectively & \\
-$f_i,\,f^{\prime}_i$ & Adult fertilities through male and female sex functions & \\
+$f_i,\,f^{\prime}_i$ & Adult fertilities through female and male sex functions respectively & \\
 
@@
 We now construct and analyze a genotype $\times$ stage-classified model for a hypothetical species with intralocus sexual conflict via the two sex functions.
-For the sake of simplicity, we assume our hypothetical species has a basic life cycle with only two life stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active.
+For the sake of simplicity, we assume our hypothetical species has a life cycle with only two stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active.
 Suppose that there is a genetic trade-off between the sex-functions at a single diallelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect.
@@
 
-The SA locus does not affect survival and transition rates.
+The SA locus does not affect survival or transition rates.
 However, the survival matrices can be used to model the fitness effects of inbreeding depression at later stages of development by allowing different stage-specific survival and transition rates for individuals produced by self-fertilization vs.~outcrossing.
@@
 Unless stated otherwise, we use the following parameter values for the demographic rates in the model: $\sigma_j = \sigma_a = 0.6$ and $\gamma = 0.05$.
-These chosen values are similar to those used in \citet{deVriesCaswell2019b}, facilitating comparison between models, and to correspond to a life-history in which individuals spend multiple timesteps in the juvenile phase prior to maturing into reproductively active adults, but are otherwise arbitrary.
+These chosen values are similar to those used in \citet{deVriesCaswell2019b}, facilitating comparison between models, and correspond to a life-history in which individuals spend multiple timesteps in the juvenile phase prior to maturing into reproductively active adults, but are otherwise arbitrary.
 Our parameters of interest include fertility, $f$, the inbreeding depression parameters, $\delta$, $\delta_j$, $\delta_a$, $\delta_{\gamma}$, and the selection parameters $h_f$, $s_f$, $h_m$, and $s_m$, which are given different values for each analyses as described in the figure captions.
@@
 The formal conditions for a protected polymorphism are determined by linearizing the model in the vicinity of the boundary equilibria ($\hat{\mbf{p}}$), and evaluating the magnitude of the largest eigenvalue of the Jacobian matrix of the linearization.
-A full derivation of the Jacobian and details of the invasion analysis are provided in the \hl{Online Supplementary Material}, and the relevant leading eigenvalues are presented in \hl{Appendix B}.
+A full derivation of the Jacobian and details of the invasion analysis are provided in the Online Supplementary Material, and the relevant leading eigenvalues are presented in Appendix B.
 
@@
 Specifically, we examine scenarios of ({\itshape i}) additive SA fitness effects ($h = 1/2$), which are commonly observed for small-effect alleles or quantitative traits \citep{AgrawalWhitlock2011}; and ({\itshape ii}) dominance reversals, where the deleterious fitness effect of each SA allele is partially recessive in each sex ($h = 1/4$), which are predicted under fitness landscape models provided the population is not too far from the phenotypic optimum \citep{Manna2011, ConnallonClark2014}.
-We briefly explore the consequences of sex-specific dominance, which may evolve under asymmetric SA selection \citep{SpencerPriest2016} in \hl{Appendix X}).
+We briefly explore the consequences of sex-specific dominance in the Online Supplementary Material.
 
@@
 We begin with an illustration of demographically viable polymorphic parameter space in the absence of inbreeding depression in Figure \ref{fig:extThresholds} (i.e., $\delta = \delta_i = 0$).
-The red areas in Figure \ref{fig:extThresholds} indicate areas where the population goes extinct (population growth rate is less than one) for different values of the fertility parameter.
+The red regions in Figure \ref{fig:extThresholds} indicate areas where the population goes extinct (population growth rate is less than one) for different values of the fertility parameter.
 Lower female fertility values correspond to a larger demographically inviable area.
 The funnel between the thick black lines indicates the area where the sexually antagostic alleles coexist.
-Above the funnel, the female-benefit and male-detriment allele fixates; below the funnel, the male-benefit and female-detriment allele fixates.
+Above the funnel, the female-benefit and male-detriment allele goes to fixation; below the funnel, the male-benefit and female-detriment allele fixates.
 Much of the region where the male-benefit and female-detriment allele fixates is demographically inviable due to the demographic consequences of reduced female fitness in this region.
@@
 The SA polymorphisms that remained viable at lower fertilitiy values corresponded to regions where the female-deleterious allele is predicted to segregate at low frequencies, or in regions with weak selection through both sex functions, an asymmetry that reflects the assumption of female demographic dominance.
-
-In populations with high fertility (larger $f$), the proportion of demographically viable polymorphic parameter space converges on the predictions for total SA polymorphic space in population genetic models (fig.~\ref{fig:polySpace}).
-In obligately outcrossing populations (including dioecious/gonochoristic populations; where $C = 0$), lower fertility can result in a significant reduction of demographically viable polymorphic parameter space.
-However, the effect is weaker in populations with intermediate selfing rates (compare $C=0$ vs.~$C > 0$) because self-fertilization protects a fraction, $C$, of the ovules of individuals homozygous for the female-beneficial allele from exposure to the male-beneficial allele.
-In other words, self-fertilization reduces the opportunity for selection favouring the female-deleterious allele, thereby reducing the equilibrium load on female fecundity and allowing partially selfing populations to remain viable under selection intensities that would cause extinction in an outcrossing population.
-The combination of protection from reduced female fertility and reduced total polymorphic parameter space caused by selfing results in populations with intermediate selfing rates having the greatest proportion of demographically viable parameter space at medium and low fertilities (fig.~\ref{fig:polySpace}, med.~and low.~fertility values).
 
@@
  \end{figure}
+
+In populations with high fertility (larger $f$), the proportion of demographically viable polymorphic parameter space converges on the predictions for total SA polymorphic space in population genetic models (fig.~\ref{fig:polySpace}).
+In obligately outcrossing populations (including dioecious/gonochoristic populations; where $C = 0$), lower fertility can result in a significant reduction of demographically viable polymorphic parameter space.
+The effect is weaker in populations with intermediate selfing rates (compare $C=0$ vs.~$C > 0$) because self-fertilization generates a greater proportion of offspring that are homozygous for the female-beneficial allele relative to heterozygotes.
+In other words, self-fertilization reduces the opportunity for selection favouring the female-deleterious allele, thereby reducing the equilibrium load on female fecundity and allowing partially selfing populations to remain viable under selection intensities that would cause extinction in an outcrossing population.
+The combination of protection from reduced female fertility and reduced total polymorphic parameter space caused by selfing results in populations with intermediate selfing rates having the greatest proportion of demographically viable parameter space at medium and low fertilities (fig.~\ref{fig:polySpace}, med.~and low.~fertility values).
 
@@
 Moreover, detailed demographic studies have been conducted on multiple populations of {\itshape M.~guttatus}, with demographic data available on the public demographic database COMPADRE \citep{CompadreDB2020}.
-Below, we briefly outline how we parameterized our model using the available data; full details are provided in \hl{Appendix C}.
+Below, we briefly outline how we parameterized our model using the available data; full details are provided in Appendix C.
 
@@
 The first was a locally adapted 'Eagle Meadows' population (data from 2012), while the second was an experimental population composed of multiple non-locally adapted 'low-elevation perennials' (data from 2013).
-The vital rate estimates for the Eagle Meadows population are as follows: seed bank survival ($D = 0.534$), seed germination rate ($G = 0.469$), flower production ($F = 0.64$), ovules per flower ($O = 614$), seedling recruits proportional to clonal rosette recruits ($A = 6.7 \times 10^{-4}$), overwinter survival ($S = 0.179$), rosette production ($R = 8.71$).
+The vital rate estimates for the Eagle Meadows population are as follows: seed bank survival ($D = 0.534$), seed germination rate ($G = 0.469$), flower production ($F = 0.64$), ovules per flower ($O = 614$), seedling recruits proportional to clonal rosette recruits ($A = 6.7 \times 10^{-4}$), overwinter survival ($S = 0.179$), and rosette production ($R = 8.71$).
 The corresponding estimates for the low-elevation perrenials are: $G = 0.652$, $F = 4.09$, $O = 494$, $S = 0$, and $R = 0$ (see corrected Tables.~1 and S2 in \citealt{PetersonEtAl2017}).
 The same estimates for $D$ and $A$ were used for all populations.
-The resulting transition matrices for this population involved three life-history stages ($\omega = 3$; seed, seedling, and rosette), and individual elements of the transition matrix ($\tilde{\mbf{A}}$) were calculated as products of the above rates (see Matrix $1$ in \citealt{PetersonEtAl2016}).
+The resulting transition matrices for this population involved three life-history stages ($\omega = 3$; seed, seedling, and rosette), and individual elements of the transition matrix ($\tilde{\mbf{A}}$) were calculated as products of the above rates (see Matrix $1$ in \citealt{PetersonEtAl2016}, also Eq(\ref{eq:AtildeMimulus}) in our Appendix C).
 
@@
 Using these combined demographic rates, selfing rates, and inbreeding depression estimates, we constructed a corresponding stage $\times$ genotype mendelian matrix model with a single SA locus affecting female and male fertility (as described above).
-With the empirically parameterized model, we are able to make predictions about the scope for demographically viable SA polymorphism at a single locus in hypothetical populations with the same demographic rates as observed in \citet{PetersonEtAl2016}, for a range of selfing and inbreeding depression rates observed in other natural populations.
+With the empirically parameterized model, we are able to make predictions about the genetic and demographic outcome of SA selection a single locus in hypothetical populations with the same demographic rates as observed in \citet{PetersonEtAl2016}, for a range of selfing and inbreeding depression rates observed in other natural populations.
 We stress, however, that these are illustrative rather than explicit predictions of the likelihood of SA polymorphism in any specific population, and they ignore measurement error for the estimated demographic rates.
@@
 Interestingly, a polymorphic chromosomal inversion (inv6) with apparently SA fitness effects has been identified in the Iron Mountain population of {\itshape M.~guttatus} \citep{LeeKelly2015}.
-inv6 segregates at moderate frequency (about $8\%$), and carriers suffer $\approx 30 \%$ loss in pollen viability, but also increased flower production.
-The genetic basis of these effects remain unclear -- though they likely polygenic -- but the net result is a "supergene" with strong fitness effects on both female and male fertility that segregates as a single diallelic locus.
-We estimated selection coefficients for the effect of inv6 on pollen production (i.e., taking into account the simultaneous effect on flower number) and ovule production from the data reported in \citet{LeeKelly2015} under the relatively conservative assumptions that pollen/ovule production is proportional to flower number, and  additive SA fitness effects (dominance coefficients could not be estimated from field data; see Appendix \hl{C} for details of estimating selection coefficients).
+inv6 segregates at moderate frequency (about $8\%$), and carriers suffer an approximately $30 \%$ loss in pollen viability, but also increased flower (and therefore ovule and pollen) production.
+The genetic basis of these effects remain unclear -- though they likely polygenic -- but the net result is a "supergene" with remarkably strong effects on both female and male fertility that segregates as a single diallelic locus.
+We estimated selection coefficients for the effect of inv6 on pollen production (i.e., taking into account the simultaneous effect on flower number) and ovule production from the data reported in \citet{LeeKelly2015} under the relatively conservative assumptions that pollen/ovule production is proportional to flower number, and  additive SA fitness effects (dominance coefficients could not be estimated from field data; see Appendix C for details of estimating selection coefficients).
 Under these assumptions, the average selection coefficient for inv6 on flower production across two years in the field was $(s_f,s_m) = (0.3,0.31)$.
-As a final proof-of-concept for our empirically parameterized model \citep[e.g.,][]{Servedio2014}, we asked whether, given the above biologically grounded fitness effect estimates, inv6 appears to fall in demographically viable SA polymorphic parameter space (see Figure \ref{fig:mimulusFig}), as might be expected given its observed frequency in the Iron Mountain population.
-
-%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Polymorphism in M.~guttatus} \label{sec:Mgutt}
-
-The Eagle Meadows and Low-Elevation Perennial populations of \citet{PetersonEtAl2016} had contrasting demographic rates that strongly affected the scope for demographically viable SA polymorphism.
-The Eagle Meadows population was locally adapted with a very high intrinsic growth rate ($\lambda \approx 1.7$).
-This growth rate was sufficiently high that all $s_f \times s_m$ selection parameter space (where $s_f, s_m \in (0,1]$) remained demographically viable, regardless of the selfing rate and effects of inbreeding depression ($C = 0$ or $0.29$; fig.~\ref{fig:mimulusFig}A).
-In contrast, the Low-Elevation Perennial population had a much lower, but still positive, intrinsic growth rate ($\lambda \approx 1.08$).
-Due to the slower growth rate, not all of the $s_f \times s_m$ selection parameter space was demographically viable: extinction thresholds appear under both complete outcrossing and partial selfing with inbreeding depression (fig.~\ref{fig:mimulusFig}B).
-The different demographic rates from these two populations also resulted in slightly different invasion conditions when inbreeding depression was taken into account (compare dashed with solid lines in fig.~\ref{fig:mimulusFig}A and B).
-
-When using demographic rates for either the Eagle Meadows or Low-Elevation Perennial populations, inv6 falls within demographically viable polymorphic parameter space (fig.~\ref{fig:mimulusFig}; estimated selection coefficients for inv6 were $(s_f,\,s_m) = (0.30,\, 0.31)$).
-\hl{I like Crispins idea of reporting predicted inv6 frequencies in our model, it nicely illustrates another useful output of the model} Whether under obligate outcrossing or partial selfing, inv6 always falls squarely in the middle of SA polymorphic space when using the locally adapted Eagle Meadows demographic rates.
-In contrast, when the selfing rate is at the higher end of empirical estimates for the Iron Mountain population in which inv6 has been documented ($C = 0.29$), inv6 falls near the upper boundary but still within SA polymorhic space when using the Low-Elevation Perennial demographic rates.
-\hl{Another place where we could report inv6 frequencies?} This happens because under the empirical estimates of selfing and inbreeding depression, the polymorphic parameter space shifts downwards.
-Additionally, inv6 falls much closer to the extinction threshold under partial selfing.
-This suggests that even relatively small perturbations to demographic rates or selection coefficients could result in non-locally adapted populations being unable to support the demographic costs associated with segregating SA alleles with selection coefficients of similar magnitude to inv6.
+As a final proof-of-concept for our empirically parameterized model \citep[e.g.,][]{Servedio2014}, we asked whether, given the above biologically grounded fitness effect estimates, inv6 appears to fall in demographically viable SA polymorphic parameter space (see fig.~\ref{fig:mimulusFig}), as might be expected given its observed frequency in the Iron Mountain population.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Polymorphism in M.~guttatus} \label{sec:Mgutt}
+
+The Eagle Meadows and Low-Elevation Perennial populations of \citet{PetersonEtAl2016} had contrasting demographic rates that strongly affected the scope for demographically viable SA polymorphism.
+The Eagle Meadows population was locally adapted with a very high intrinsic growth rate ($\lambda \approx 1.7$).
+This growth rate was sufficiently high that all $s_f \times s_m$ selection parameter space (where $s_f, s_m \in (0,1]$) remained demographically viable, regardless of the selfing rate and effects of inbreeding depression ($C = 0$ or $0.29$; fig.~\ref{fig:mimulusFig}A).
+In contrast, the Low-Elevation Perennial population had a much lower, but still positive, intrinsic growth rate ($\lambda \approx 1.08$).
+Due to the slower growth rate, not all of the $s_f \times s_m$ selection parameter space was demographically viable: extinction thresholds appear under both complete outcrossing and partial selfing with inbreeding depression (fig.~\ref{fig:mimulusFig}B).
+The different demographic rates from these two populations also resulted in slightly different invasion conditions when inbreeding depression was taken into account (compare dashed with solid lines in fig.~\ref{fig:mimulusFig}A and B).
+
+When using demographic rates for either the Eagle Meadows or Low-Elevation Perennial populations, inv6 falls within demographically viable polymorphic parameter space (fig.~\ref{fig:mimulusFig}).
+Whether under obligate outcrossing or partial selfing, inv6 always falls squarely in the middle of SA polymorphic space when using the locally adapted Eagle Meadows demographic rates.
+In contrast, when the selfing rate is at the higher end of empirical estimates for the Iron Mountain population in which inv6 has been documented ($C = 0.29$), inv6 falls nearer the upper boundary but still within SA polymorhic space when using the Low-Elevation Perennial demographic rates.
+This happens because under the empirical estimates of selfing and inbreeding depression, the polymorphic parameter space shifts downwards.
+Additionally, inv6 falls much closer to the extinction threshold under partial selfing,  suggesting that even relatively small perturbations to demographic rates or selection coefficients could result in non-locally adapted populations being unable to support the demographic costs associated with segregating SA alleles with selection coefficients of similar magnitude to inv6.
+
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 The first and central finding of our study is that when intrinsic population growth rates approach one, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space where SA polymorphism is maintained.
-Since growth rates far exceeding one suggest rapid exponential growth, they are generally rare (with the notable exception of recently introduced, invasive populations), suggesting that our model predictions may be highly relevant for SA polymorphism in many real-world populations.
-In addition, we find that much of the parameter space where a male-beneficial allele fixates is demographically inviable for populations whose growth rates are close to one prior to invasion.
+Since intrinsic growth rates far exceeding one suggest rapid exponential growth, they are generally rare (with the notable exception of recently introduced, invasive populations), suggesting that our model predictions may be highly relevant for SA polymorphism in many real-world populations.
+In addition, we find that much of the parameter space where a male-beneficial allele fixates is demographically inviable for populations whose intrinsic growth rates are close to one prior to invasion.
 This demographic consequence of masculinization is rarely considered \citep{hitchcock2020gene}.
@@
 
-By combining the tools of demography and those of population genetics, the framework we present here enables the exploration of interactions between life cycle complexity, mating system, and sexual antagonism.
+By combining the tools of demography and population genetics, the framework we present here enables the exploration of interactions between life cycle complexity, mating system, and sexual antagonism.
 For the sake of simplicity, we made a variety of simplifying assumptions in our analyses, and a variety of extensions are possible.
-For example, we used a basic life cycle with just two stages for most of this paper, adults and juveniles (the \textit{Mimulus} example has $3$ stages).
+For example, we used a simple life cycle with just two stages for most of this paper, adults and juveniles (the \textit{Mimulus} example has $3$ stages).
 However, it is possible to include additional age classes, allowing us to explore whether the scope for SA selection to maintain polymorphisms is affected by whether a species exhibits positive or negative senescence \citep{jones2014diversity}.
@@
 We are particularly grateful to the European Society for Evolutionary Biology (ESEB) for funding the Special Topics Network workshops “Linking local adaptation with the evolution of sex differences”, and to the participants, all of whom made this study possible.
-We would also like to thank ....
+We would also like to thank J.K.~ Abbott, T.~Connallon, H.~Caswell, S.F.~van Daalen, C.~Venables, the Genetics of Sex Research Group at Lund University, and XXX reviewers for constructive feedback on earlier versions of the manuscript.
 This work was supported by a Wenner-Gren Foundation postdoctoral stipend to C.O., and [Lotte's funding] to C.dV..
+
+%%%%%%%%%%%%%%%%%%%%%
+% Author Contributions
+%%%%%%%%%%%%%%%%%%%%%
+
+%\section{Author Contributions}
+%The authors jointly conceived the ideas, refined the concepts, constructed the model, performed the analyses, and wrote/edited the manuscript.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\begin{landscape}
 \section*{Appendix B: Eigenvalues for invasion analysis}
@@
 A boundary equilibrium is unstable to invasion by the rare allele if the largest absolute value of the eigenvalues of the Jacobian matrix, the leading eigenvalue, evaluated at the equilibrium is greater than $1$.
-The resulting conditions for a 'protected polymorphism' require that
+The resulting conditions for a protected polymorphism require that
 
 {\footnotesize
-\begin{linenomath*}
 \begin{equation} \label{eq:coexist_AA}
@@
 			\rho\left(\begin{array}{ccc}
-\mathbf{U}^S_{Aa} +\frac{1}{2} C (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C (1 - \delta) \mbf{F}_{Aa}  & \mbf{0} \\
- \frac{1}{2} (1 - C) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C)\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C)\mbf{F}_{AA} \big(\hat{\mbf{p}}_{X,AA}+ \hat{\mbf{p}}_{S,AA}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & (1-C)\mbf{F}_{aa} + \Theta\\
-  \frac{1}{4} C (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C (1 - \delta) \mbf{F}_{aa}  \\ 
+\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & 0 \\
+ \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2}  \Theta_{Aa} &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2}\Theta_{Aa} & (1-C_{aa})\mbf{F}_{aa} + \Theta_{aa}\\
+  \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{aa} +C_{aa} (1 - \delta) \mbf{F}_{aa}  \\ 
 			\end{array} \right), \\ 
 \end{equation}
-\end{linenomath*}
 
 \noindent and
-\begin{linenomath*}
 \begin{equation} \label{eq:coexist_aa}
@@
 			\rho\left(\begin{array}{ccc}
-\mathbf{U}^S_{Aa} +\frac{1}{2} C (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C (1 - \delta) \mbf{F}_{Aa}  & \mbf{0} \\
- \frac{1}{2} (1 - C) \mbf{F}_{Aa}+\frac{1}{2  \mbox{p}_{n}} (1 - C)\mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C) \mbf{F}_{Aa} +\frac{1}{2  \mbox{p}_{n}} (1 - C)\mbf{F}_{aa} \big(\hat{\mbf{p}}_{X,aa}+ \hat{\mbf{p}}_{S,aa}  \big) \otimes \mathbf{1}^\tr_{\omega} \mathbf{F}_{Aa}^\prime & (1-C)\mbf{F}_{AA} + \Phi\\
-  \frac{1}{4} C (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{AA} + C (1 - \delta) \mbf{F}_{AA}  \\ 
+\mathbf{U}^S_{Aa} +\frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  & 0 \\
+ \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa}+\frac{1}{2 } \Phi_{Aa} &\mathbf{U}^X_{Aa} + \frac{1}{2} (1 - C_{Aa}) \mbf{F}_{Aa} +\frac{1}{2} \Phi_{Aa} & (1-C_{AA})\mbf{F}_{AA} + \Phi_{AA}\\
+  \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa}  &\mathbf{U}^S_{AA} + C_{AA} (1 - \delta) \mbf{F}_{AA}  \\ 
 			\end{array} \right), \\ 
 \end{equation} 
-\end{linenomath*}
 }
@@
 In our numerical simulations, equations Eq(\ref{eq:coexist_AA}) and Eq(\ref{eq:coexist_aa}) were evaluated to determine the invasion thresholds for the relvant parameter conditions described in each figure.
-\end{landscape}
+%\renewcommand{\theequation}{B\arabic{equation}}
 
@@
 Specifically, we used data from two experimental populations with contrasting demographic rates: the locally adapted 'Eagle Meadows' population (data from 2012), and a non-local 'low-elevation perennials' population (data from 2013).
-The vital rates used in theirs, and our, calculations are summarized below in Table(\ref{tab:MimDemData}; see corrected Tables.~1 and S2 in \citet{PetersonEtAl2016, PetersonEtAl2017}.
-Note that the same estimates for $D$ and $A$ were used for all populations in the study.
+The vital rates used in theirs, and our, calculations are summarized below in Table \ref{tab:MimDemData} (see corrected Tables.~1 and S2 in \citet{PetersonEtAl2016, PetersonEtAl2017}).
+Note that the same estimates for seed bank survival ($D$) and Seedling recruitment ($A$) were used for all populations in the study.
 
@@
 Importantly, the study reports fitness effects of inv6 on both male and female fitness components.
-inv6 appears to cause a significant decrease in pollen viability, while simultaneously causing an increase in flower number which varies among years.
+inv6 appears to cause a significant decrease in pollen viability, while simultaneously causing an increase in flower number (and therefore ovule and pollen production) which varies among years.
 Unfortunately, it was not possible to estimate the dominance of the fitness effects of inv6 in the field because individuals homozygous for the inversion were quite rare (although they were viable in greenhouse conditions).
@@
 
-We used flower number as a proxy for female fertility.
+We used flower number as a proxy for ovule production and hence female fertility.
 To calculate selection on flower number, we first calculated the geometric mean flower number in the field for wild-type and inv6 heterozygotes across the two years for which data was available (2012 and 2013), extrapolated the expected flower number for inv6 homozygotes (table \ref{tab:MimReproEst}), and calculated the relative fertility of each genotype given that inv6 homozygotes have the highest flower number ($w_i$, where $i \in \{inv6/inv6,\, inv6/w+,\, w+/w+\}$).

### 2021-06-04 doc/Supplements/SupplementaryMaterial.tex
Pre-Submission Edits
This commit adds final edits to the main manuscript and online Supp.
Mat., includes new results for complementary sex-specific
dominance in the Supp. Mat., and removes author information for
double-blind peer review.

@@
 \usepackage{graphicx}
-\graphicspath{{../output/figs/}.pdf}
+\graphicspath{{../../output/figs/}.pdf}
 
@@
 %
-\rhead{Supplement to Olito \& de Vries, "Costs of SA selection" \textit{Am.~Nat.}}
+%\rhead{Supplement to Olito \& de Vries, "Costs of SA selection" \textit{Am.~Nat.}}
+\rhead{Supplement to "Costs of SA selection" \textit{Am.~Nat.}}
 \setlength{\headsep}{0.3in}  
@@
 
-\title{Online Supplement to "Demographic costs of sexually antagonistic selection in partially selfing populations \\ 
+\title{Online Supplement to "Demographic costs of sexually antagonistic selection in partially selfing populations" \\ 
 %\LaTeX{} Template for Author-Supplied Supplementary PDFs, \\ 
@@
 % page numbers.
-\author{Colin Olito$^{1,\ast}$ \\ 
-Charlotte de Vries$^{2}$}
+%\author{Colin Olito$^{1,\ast}$ \\ 
+%Charlotte de Vries$^{2}$}
 
@@
 
-\noindent{} 1. Department of Biology, Lund University, Lund 223 62, Sweden;
+%\noindent{} 1. Department of Biology, Lund University, Lund 223 62, Sweden;
 
-\noindent{} 2.  Department of Evolutionary Biology and Environmental Studies, University of Zurich, Zurich CH-8057, Switzerland;
+%\noindent{} 2.  Department of Evolutionary Biology and Environmental Studies, University of Zurich, Zurich CH-8057, Switzerland;
 
-\noindent{} $\ast$ Corresponding author; e-mail: \url{colin.olito@gmail.com}
+%\noindent{} $\ast$ Corresponding author; e-mail: \url{colin.olito@gmail.com}
 
@@
 \renewcommand{\thetable}{S\arabic{table}}
-\renewcommand{\thesection}{S\arabic{section}}
+\renewcommand{\thesection}{Supplement~\arabic{section}}
 \renewcommand{\thefigure}{S\arabic{figure}}
@@
 \section{Construction of the population projection matrix}
-We will now derive the expressions for the survival and fertility matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ as given by equations \textcolor{red}{(10)-(13)} in the main text.
-The matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ are the ingredients of the population projection matrix, $\tilde{\mbf{A}}$, as given in equation \textcolor{red}{(9)}, repeated here for convenience,
+We will now derive the expressions for the survival and fertility matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ as given by Eq(10)-(13) in the main text.
+The matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ are the ingredients of the population projection matrix, $\tilde{\mbf{A}}$, as given in Eq(9), repeated here for convenience,
 \begin{linenomath*}
@@
 
-Finally, substitute the expressions derived for the parent-offpspring genotype maps, $\bo H^X(\tilde{\mbf{p}}) $, and $\bo H^S$, in the main text (\textcolor{red}{equations (12) and (13))} into the above equations to obtain
+Finally, substitute the expressions derived for the parent-offpspring genotype maps, $\bo H^X(\tilde{\mbf{p}}) $, and $\bo H^S$ into the above equations to obtain Eq(12) and Eq (13) in the main text:
 \begin{linenomath*}
@@
 
-The Mendelian matrix model defined in equations \textcolor{red}{(9)-(13)}, or (\ref{eq:AtildeCoexistence}), reduces to a linear matrix model on the boundary where the population is initially fixed for the $A$ allele (since $q^{\prime}_{A} = 1$ and $q^{\prime}_{a} = 0$).
+The Mendelian matrix model defined in Eq(9)-(13), Eq(A1), or in reorderd form as (\ref{eq:AtildeCoexistence}), reduces to a linear matrix model on the boundary where the population is initially fixed for the $A$ allele (since $q^{\prime}_{A} = 1$ and $q^{\prime}_{a} = 0$).
 Provided the initial population contains a nonzero number of females, the population will grow or shrink exponentially after converging to a stable population structure (see Caswell (2001), section 4.5.2.1).
@@
 
- Start with equation \textcolor{red}{(4)} from the main text:
+ Start with Eq(6) from the main text:
 \begin{linenomath*}
@@
 The conditions for a protected polymorphism is that both boundaries are unstable (i.e., both $\tilde{\zeta}_{AA} > 1$ and $\tilde{\zeta}_{aa} > 1$).
-The conditions for coexistence are, therefore, as given in equation \textcolor{red}{(B1)} and \textcolor{red}{(B2)},
+The conditions for coexistence are, therefore, as given in Eq(B1) and Eq(B2),
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage{}
+\section{Sex-specific dominance}
+\renewcommand{\thefigure}{S\arabic{equation}}
+\setcounter{figure}{1}  % reset counter 
+\setcounter{equation}{1}  % reset counter 
+\setcounter{table}{1}  % reset counter
+
+A variety of biologically interesting dominance scenarios are possible under sex-specific selection.
+For reasons explained in the main text, we focused on two scenarios of equal dominance (where $h_f = h_m = h$): additivity ($h = 1/2$), and dominance reversal ($h = 1/4$).
+In fact, these scenarios are special cases of two broader classes of sex-specific dominance scenarios.
+Additivity represents a case where sex-specific dominance is complementary ($h_f + h_m = 1$), and dominance reversals represent one where $h_f + h_m < 1$.
+Complementarity means there is a trade-off in the levels of dominance between male and female function.
+Furthermore, it is of particular biological interest because it represents the case where phenotypes under sex-specific selection are identical in males and females (or for hermaphrodites, investment in eggs/ovules and sperm/pollen is equivalent) and fitness is a linear function of the phenotype.
+Complementarity therefore seems plausible when sex-specific selection first arises \citep{SpencerPriest2016}.
+However, it is not intuitively obvious what the demographic consequences of sex-specific dominance will be in our model, especially given the assumption of female demographic dominance.
+Below we explore two scenarios of complementary sex-specific dominance, one where SA fitness effects are recessive in females and dominant in males ($h_f = 1/4$, $h_m = 3/4$), and the reverse ($h_f = 3/4$, $h_m = 1/4$).
+Overall, our supplementary results show that under complementary sex-specific dominance, the demographic consequences of SA selection are broadly similar to those under additive SA fitness effects.
+
 %%%%%%%%%%%%%%%%%%%%%
+% Figure S1
+ \begin{figure}[htbp]
+ \centering
+ \includegraphics[width=\linewidth]{extinctionThresholdsSexSpecDomFig}
+ \caption{\footnotesize{Illustration of parameter space for SA polymorphism and extinction thresholds predicted by the model.
+Balanced SA polymorphisms can be maintained in the funnel-shaped region between the invasion conditions for each SA allele (dark solid lines).
+However, for some parameter conditions, populations will ultimately go extinct (red shaded regions) due to reduced female fitness resulting from the male-beneficial/female-deleterious allele that is either segregating as a balanced polymorphism (inside the funnel), or becomes fixed (area below the funnel).
+"Demographically viable polymorphic parameter space" corresponds to the area inside the funnel that is also to the left of the extinction threshold for a given fertility value.
+Results are shown for three different population selfing rates ($C = \{0,\,1/4,\,1/2\}$), and two cases of complementary dominance ($h_f = 1/4$, $h_m = 3/4$, top row), and ($h_f = 3/4$, $h_m = 1/4$, bottom row), and extinction thresholds are illustrated for the same three values of female fecundity illustrated in fig.~1 of the main text ($f$ values annotated on each panel).}}
+ \label{fig:extThresholdsCompSexSpec}
+ \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%
+% Figure S2
+ \begin{figure}[htbp]
+ \centering
+ \includegraphics[scale=0.75]{polymorphicSpaceTitrateSexSpec}
+ \caption{\footnotesize{Proportion of demographically viable parameter space (out of total $s_f \times s_m$ space with $\max(s) = 0.15$) in the absence of inbreeding depression (i.e., assuming $\delta = \delta_i = 0$, where $i \in \{j,a,\gamma\}$), plotted as a function of the population selfing rate.
+Results are shown for three fertility values corresponding to low, medium, and high fertility (blue, green, and red points respectively) under two cases of complementary dominance for  SA fitness effects ($h_f = 1/4$, $h_m = 3/4$, panel A; and $h_f = 3/4$, $h_m = 1/4$, panel B).
+Each point was calculated by numerical integration of the corresponding SA invasion conditions and extinction threshold predicted by the mendelian matrix model (see Analyses section in the main text), while solid lines were produced by numerically integrating the analytic expressions for the single-locus invasion conditions from the population genetic models of \citet{JordanConnallon2014} and \citet{Olito2017} (solid black lines).}}
+ \label{fig:polySpaceCompSexSPec}
+ \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%
+% Figure 3
+ \begin{figure}[htbp]
+ \centering
+ \includegraphics[scale=0.75]{deltaPolymorphicSpaceTitrateCompSexSpec}
+ \caption{\footnotesize{Effects of early- and late-acting inbreeding depression on the proportion of demographically viable parameter space (out of total $s_f \times s_m$ space with $\max(s) = 0.15$), plotted as a function of the population selfing rate.
+In all plots, the strength of inbreeding depression decreases as the selfing rate goes up following a simple model of purging recessive deleterious mutations.
+Only single inbreeding depression terms ($\delta$ and $\delta_i$, where $i \in \{j,a,\gamma\}$, indicated in the legend) are allowed to vary at one time (all others are set to $0$).
+Results are shown for the same three fertility values as used in the main text ($f = \{6.5,\,7.5,\,8.5\}$) under two cases of complementary dominance for  SA fitness effects ($h_f = 1/4$, $h_m = 3/4$, panel A; and $h_f = 3/4$, $h_m = 1/4$, panel B).
+Each point was calculated by numerically integrating the corresponding SA invasion conditions and extinction threshold predicted by the model (see Analyses section in the main text), while solid lines were produced by numerically integrating the single-locus invasion conditions from the population genetic models of \citet{JordanConnallon2014} and \citet{Olito2017} (solid black lines).}}
+ \label{fig:deltaPolySpaceCompSexSpec}
+ \end{figure}
+
+%%%%%%%%%%%%%%%%%%
 % Bibliography
@@
 % except when they are already cited in the main text.
+\newpage{}
 \bibliography{suppRefs.bib}

### 2021-06-04 R/functions-Figs.R

@@ -1438,9 +1438,328 @@ polySpaceFigTitrateSexSpec  <-  function(df = "dataPolySpaceFig_hf0.25_0.75_hm0.
 #' Fig showing Effect of Inbreeding Depression on 
 #' Proportion of polymorphic parameter space
 
-deltaSelfingLoadPolySpaceFigTitrate  <-  function(df1 = "dataDeltaPolySpaceFigSexSpecDom_sMax0.15_res0.003_dStar0.8_f6.5",
-                                                  df2 = "dataDeltaPolySpaceFig_sMax0.15_res0.003_dStar0.8_f7.5",
-                                                  df3 = "dataDeltaPolySpaceFig_sMax0.15_res0.003_dStar0.8_f8.5") {
+deltaSelfingLoadPolySpaceFigTitrateSexSpec  <-  function(df1 = "dataDeltaPolySpaceFigSexSpecDom_sMax0.15_res0.003_dStar0.8_f6.5",
+                                                         df2 = "dataDeltaPolySpaceFigSexSpecDom_sMax0.15_res0.003_dStar0.8_f7.5",
+                                                         df3 = "dataDeltaPolySpaceFigSexSpecDom_sMax0.15_res0.003_dStar0.8_f8.5") {
+
+    # Make filenames for import from df names
+    fName1  <-  paste('./output/simData/', df1, '.csv', sep="")
+    fName2  <-  paste('./output/simData/', df2, '.csv', sep="")
+    fName3  <-  paste('./output/simData/', df3, '.csv', sep="")
+
+    # import data
+    data1  <-  read.csv(file=fName1, header=TRUE)
+    data2  <-  read.csv(file=fName2, header=TRUE)
+    data3  <-  read.csv(file=fName3, header=TRUE)
+
+    # Extract plotting parameter values from df names
+    d1   <-  strsplit(df1, '_')[[1]][c(2:5)]
+    pars1  <-  list(
+                    "sMax"  =  as.numeric(strsplit(d1[1],'x')[[1]][2]),
+                    "res"   =  as.numeric(strsplit(d1[2],'s')[[1]][2]),
+                    "dStar" =  as.numeric(strsplit(d1[3],'r')[[1]][2]),
+                    "f"     =  as.numeric(strsplit(d1[4],'f')[[1]][2])
+                    )
[…]

### 2021-06-04 R/functions-Simulations.R

@@ -2186,10 +2186,13 @@ cleanPolySpaceData  <-  function(data, sexSpecDom = FALSE) {
 
     # Get plotting levels from df
 	if(sexSpecDom) {
-	    hLev  <-  unique(data$hf)
+		hLev  <-  unique(data$hf)
 	}
     else{hLev  <-  unique(data$h)}
     fLev  <-  unique(data$f)
+    if(is.null(fLev)){
+		fLev  <-  1
+    }
     CLev  <-  unique(data$C)
     nHs   <-  length(hLev)
     nCs   <-  length(CLev)
@@ -2208,7 +2211,9 @@ cleanPolySpaceData  <-  function(data, sexSpecDom = FALSE) {
 					d  <-  data[data$h == hLev[i],]
 				}
 				# Subset data by Fertility Value
-				d  <-  d[d$f == fLev[j],]     
+				if(!is.null(d$f)){
+					d  <-  d[d$f == fLev[j],]
+				}
 				# Subset data by Selfing Rate
 				d  <-  d[d$C == CLev[k],]     
 
@@ -2352,8 +2357,11 @@ quantPolySpaceSexSpec  <-  function(data, pars) {
     # get levels for loops
     res      <-  pars$res
     sMax     <-  pars$sMax
-    hfLev     <-  unique(data$hf)
+    hfLev    <-  unique(data$hf)
     fLev     <-  unique(data$f)
+    if(is.null(fLev)) {
+		fLev  <-  1
+    }
     CLev     <-  unique(data$C)
     nHs      <-  length(hfLev)
     nFs      <-  length(fLev)
@@ -2371,7 +2379,9 @@ quantPolySpaceSexSpec  <-  function(data, pars) {
 				# Subset data by Dominance
 				d  <-  data[data$hf == hfLev[i],]
 				# Subset data by Fertility Value
-				d  <-  d[d$f == fLev[j],]
+				if(!is.null(d$f)) {
+					d  <-  d[d$f == fLev[j],]
+				}
 				# Subset data by Selfing Rate
 				d  <-  d[d$C == CLev[k],]
 
@@ -2401,13 +2411,16 @@ quantPolySpaceSexSpec  <-  function(data, pars) {
 						PrViaPoly  <-  viaPoly/totSpace
 					}
 
+					if(fLev == 1){
+						fLev  <-  pars$f
+					}
 				# append results
[…]

### 2021-06-04 makeFigs.R

@@ -51,17 +51,22 @@ embed_fonts(figPath(name='MimulusInv6FigNewSel.pdf'))
 # Supp. Figs
 #################
 
-# SuppFig. -- Illustration of Invaions & Extinction thresholds for sex-specific dominance
+# Fig. S1 -- Illustration of Invaions & Extinction thresholds for sex-specific dominance
 toPdf(demViablePolySpace_SexSpecDominance_Fig(), 
 			figPath(name='extinctionThresholdsSexSpecDomFig.pdf'), width=10, height=5)
 embed_fonts(figPath(name='extinctionThresholdsSexSpecDomFig.pdf'))
 
-# SuppFig. -- Quantification of demographically viable parameter space, no inbreeding depression
+# Fig. S2 -- Quantification of demographically viable parameter space, no inbreeding depression
 
 toPdf(polySpaceFigTitrateSexSpec(), 
 			figPath(name='polymorphicSpaceTitrateSexSpec.pdf'), width=5, height=7)
 embed_fonts(figPath(name='polymorphicSpaceTitrateSexSpec.pdf'))
 
+# Fig. S3 -- Quantification of demographically viable parameter space, 
+#			showing effects of inbreeding depression
+toPdf(deltaSelfingLoadPolySpaceFigTitrateSexSpec(), 
+			figPath(name='deltaPolymorphicSpaceTitrateCompSexSpec.pdf'), width=5, height=7)
+embed_fonts(figPath(name='deltaPolymorphicSpaceTitrateCompSexSpec.pdf'))
 
 
 

### 2021-06-04 run-Simulations.R

@@ -1517,7 +1517,7 @@ makeDataPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
 
 
 ##########################################################
-# Fig. 3 - Dem. viable polymorphic param. space (w/ I.D.)
+# Fig. SX- Dem. viable polymorphic param. space (w/ I.D.)
 
 makeDataDeltaPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
 							om = 2, g = 3, theta = c(0.6,0.6,0.05,6.5), theta_prime = 6.5, 

### 2021-06-04 doc/Supplements/SupplementaryMaterial.tex
Anonymized github links
This commit adds anonymized GitHub links to the main and supp.
documents.

@@
 We therefore used numerical techniques as described in the main text to identify invasion and extinction thresholds, which could then be used to define regions of demographically viable polymorphism across $s_f \times s_m$ parameter space.
-All computer code necessary to reproduce the results are available at \url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
+All computer code necessary to reproduce the results are available at \url{https://anonymous.4open.science/r/SA-Hermaphrodites-wDemography-BF66}.%\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
 

### 2021-11-21 doc/ResponseLetter-R1.tex
Setup for revisions
This commit restructures the repo directory tree to set up a legacy
directory for the original submission, and creates new .tex files for
the revised ms *_R1.tex etc, and a response letter to the editor and
reviewer's comments.

%===========================================
% Preamble

\documentclass[11pt]{article}

%**********
%Dependencies

%**********
%Dependencies
%\usepackage[left]{lineno}
\usepackage{titlesec}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
%\usepackage[utf8]{inputenc}
\usepackage{color,soul}
\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
\usepackage{fullpage}
\usepackage[authoryear,sectionbib,sort]{natbib}
\linespread{1}
\usepackage[utf8]{inputenc}
\usepackage{lineno}
\usepackage[hidelinks]{hyperref}

% New commands: fonts
\newcommand{\code}{\fontfamily{pcr}\selectfont}
\newcommand*\chem[1]{\ensuremath{\mathrm{#1}}}

\renewcommand\textbullet{\ensuremath{\bullet}}
%===========================================
% Title page

% \title{Cover Letter}
% \author{Colin Olito}
% \date{\today}

\begin{document}
%\maketitle
%\newpage{}

%===========================================
% Document

\section*{}
\noindent To the AE, Ed., \& Editorial Board
\bigskip

\noindent Please consider this revision of our manuscript (MS \#60734), titled "The demographic costs of sexually antagonistic selection in partially selfing populations" for publication in \textit{The American Naturalist}. 
\bigskip

\noindent First of all, we want to thank the AE, Ed., and EiC for their thoughtful and forthright responses (and apologies) regarding the editorial delay on our manuscript. We understand completely that pandemic-related personal issues have been and continue to be extremely disruptive in these difficult times, and that editorial duties for our journals is a volunteer service. To our AE, Dr.~Orive, we sincerely hope that you and your family are doing ok, and we truly appreciate the time and effort you took to evaluate our paper while you clearly had more imporant issues to deal with.
\bigskip

\noindent We received very helpful and constructive comments on our first submission from Associate Editor Dr.~Maria Orive, Editor Erol Ak\c{c}ay, and two reviewers, and have now made a variety of major revisions to the MS in order to address their concerns. In particular, we have ... (1) made significant changes to our introduction and discussion in an attempt to engage with a broader readership by developing conceptual links between ours and previous work on evolutionarily explicit structured models and sexually antagonistic selection; (2) attempted to lessen the burden on non-theoretical readers by moving most of the development of the general 2-stage model to the Supplementary Material and following R2's many helpful suggestions regarding the presentation of the model; and (3) tried to more fully address the impacts of density dependence in three ways:  ... \textcolor{blue}{({\itshape i}) by revising our figures to illustrate predicted absolute population intrinsic growth rates rather than simply the boundary between demographically viable \& inviable regions} ; ({\itshape ii}) by including a brief supplement with a direct comparison between our current density-independent predictions and a new version of our model which includes simple DD; and ({\itshape iii}) by expanding our discussion about the implicaitons of DD to address these additions. Overall, our changes have shortened the manuscript by \textcolor{blue}{some impressive amount}, but we feel that they have greatly improved the paper. 
\bigskip

\noindent We have provided detailed responses to each of the editors' and reviewers' comments below. For clarity, we use bold face for all line and figure references to the revised ms in our responses. We have also used blue text in the revised manuscript file to help identify the key additions we have made to the revised text.
\bigskip

\noindent Thank you again for efforts and consideration.
\bigskip

\noindent Sincerely,

\noindent The Authors \\
%\noindent Colin Olito \& Charlotte DeVries \\
\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}
\smallskip

%%%%%%%%%%%%%%%%%%%%
\section*{Decision letter with our responses}

\subsection*{Responses to Editor Erol Ak\c{c}ay}

I must first apologize for the time it took to return this decision, and thank you for your patience. Even in the best of times, unexpected events and circumstances cause delays for reviewers and editors, and many of us are still operating under the continuing strain of the pandemic. In this case, I must also apologize to have added to the delay personally, which is because of a backlog that built up due to disruptions that happened in the last couple of months that ordinarily might not have led to huge issues, but after a year and a half of the pandemic seem to snowball out of control.

\begin{quote}
	{\itshape Thank you, again, for your candid and forthright apology and explanation. We truly appreciate it, and sympathize with the growing challenges that seem to creep up on us all during these difficult times.}
\end{quote}

The two reviewers, Dr. Orive, and myself all read your paper with great interest. We are all agreed that your model is well-constructed, and well-explained, and your conclusions are clearly presented. The insight that sexually antagonistic genetic variation might depress population demographic rates and affect population viability is a valuable one. We also appreciate that you use empirical data from a case study to parameterize your model, and agree that this is very useful to judge the real world relevance of your calculations. As such, there is good potential for this manuscript to be a great American Naturalist paper.

\begin{quote}
	{\itshape This is very encouraging indeed, and we certainly hope that our revisions can successfully fulfill the potential you see in our paper.}
\end{quote}

As you will see in Dr. Orive’s excellent summary of the reviews, one issue that we feel you need to address is to make the paper more suitable for the general audience of the American Naturalist. To my mind, this has two parts: First, I think you can reduce the amount of space you devote to the general model in the main text (this can go into the SI as a generalization and also as a reference for the later three-stage model for Mimulus), as most of the biological conclusions actually come from the two-stage model which requires less notation and the construction of which is more easily explained. Although I personally appreciated the construction of the general model, I think for our non-theoretical readership, the cost/benefit ratio of reading through the general model (which is not used for the numerical calculations) is too great.

\begin{quote}
	{\itshape ...}
\end{quote}

Second, like Reviewer 1 and Dr. Orive, I do think you should better connect your model to the broader topics of evolution in structured populations and sexual antagonism. That would give your paper a broader impact in terms of readership.

\begin{quote}
	{\itshape We agree entirely. Please see our detailed responses to Dr.~Orive and R1 on this count.}
\end{quote}

For my own part, I do have one additional major comment, which concerns how you decide on demographic viability. As you are admirably explicit in the model construction, this is a linear projection model with no density dependence in demographic rates. As such, the population growth or decline you infer from a given parameter set is necessarily "local," i.e., it applies around a particular population density where the base demographic rates are measured or modeled. To my mind, this is a potentially major gap between model calculations and what one can say about viability of actual populations in the wild, which will have density dependence and therefore will face varying demographic rates with density. For example you can have fertility rates that give a growth rate less than 1 at some densities, but as density becomes low, mortality might decrease (or fertility increase) to bring growth rates back to one. (That is of course the assumption of most eco-evolutionary theory such as adaptive dynamics: some sort of density dependence is assumed to keep populations at demographic equilibrium while new traits invade.) I understand that incorporating such non-linear, density dependent feedbacks would make the model complicated, but I also think that their lack in the current model does dampen your conclusions, in the sense that they don’t necessarily mean that populations would not be able to persist if their growth rate at some undefined density becomes less than one. At the very least, this is an issue that deserves acknowledgment and fairly extensive discussion because it directly affects your major conclusion. I was disappointed that the issue only came up once in passing in the discussion. Ideally, you could present some numerical simulations with density dependence. I sorta suspect that what you might get (at least for some cases) is depressed population densities rather than outright population inviability. I don’t know how difficult these simulations would be, but at least some numerical exploration might not be too much out of the way.

\begin{quote}
	{\itshape This is a very good and totally valid point. To be honest, we believe that a separate paper is probably necessary to deal adequately with the many consequences of density dependence for our model predictions. However, we also agree that sidestepping this issue completely isn't good enough. In the end, we have made three main changes to address density dependence. First, we have taken your advice in the following comment and revised our figures 1 \& 4 to show absolute intrinsic growth rates across selection parameter space rather than just the boundary between demographically viable \& inviable parameter space (\textcolor{blue}{caveat about Fig.1 vs. 4?}). Second, we have included a new supplementary appendix to the paper where we briefly explain how density dependence can be incorporated into the model, and provide a direct comparison between our current density-independent predictions and a new version of our model which includes DD logistic growth. Finally, we have expanded our results and discussion to more fully address the implications of DD for our model, with reference to our modified figures and the new supplement.}
\end{quote}

Related to that, and without doing more modeling, one quantity of interest you can present is the actual growth rate, and not just the region of viability/inviability. In a world where we acknowledge density dependence as a possibility, how far above or below one growth rate is can be informative about whether you actually expect extinction or depressed densities.

\begin{quote}
	{\itshape This is great suggestion. As mentioned above, we have changed figures 1 \& 4 to include a heatmap of absolute values for predicted $\lambda$ instead of the extinction boundaries. \textcolor{blue}{However, one unfortunate consequence of presenting $\lambda$ instead of the extinction boundaries is that it is no longer possible to illustrate multiple values of maximum fertility ($f$) on each figure panel. In particular, Figure 1, now requires 18 panels illustrating $\lambda$ for the same parameter combinations. We felt that this resulted in an unnecessarily complicated figure, especially since Figure 1's main purpose was to briefly illustrate the general behaviour of the model. For simplicity, we have decided to include this modified Figure 1 as a supplementary figure. However, we now present the modified version of Figure 4 in the main text}.}
\end{quote}

In addition to these general points, both reviewers and Dr. Orive have numerous helpful suggestions and corrections that will no doubt be useful while you revise the manuscript, and I urge you to pay close attention to them.
\bigskip

\noindent xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx

\section*{Responses to Associate Editor Maria Orive}

Both reviewers found the work interesting and potentially to be of strong interest to the readership of AmNat, and I agree. However, Reviewer 1 brings up an important point regarding the framing of the work and placing it within a broader context of the literature on evolutionarily explicit structured models, which I think is important for the authors to do in order to make the paper accessible to and of interest for the overall readership of AmNat. There are other points that the reviewers bring up that are important to consider, but I think this one needs to be addressed carefully.

\begin{quote}
	{\itshape ...}
\end{quote}

Reviewer 2 has some very nice suggestions regarding ideas for simplifying and clarifying the model description.

\begin{quote}
	{\itshape ...}
\end{quote}

In addition to the clarifications/edits and other points raised by the reviewers, I include a handful of minor edits/typos here of my own:

\noindent (1) Abstract, perhaps add a very general and short explanation of the term sexually antagonistic polymorphism.
\begin{quote}
	{\itshape ...}
\end{quote}

\noindent (2) Line 165; "fertility"
\begin{quote}
	{\itshape Fixed.}
\end{quote}

\noindent (3) Table 1, definition of inbreeding depression terms; it would be clearer to define each of the 4 terms separately, "Inbreeding depression terms for ovule viability, juvenile survival, adult survival, and transition rates respectively"
\begin{quote}
	{\itshape ...}
\end{quote}

\noindent (4) Figure 1; I generally prefer axes labels to have clear meanings, rather than just symbols – here, the vertical and horizontal axes give the female (vertical) vs. male (horizontal) selection on reproductive function.
\begin{quote}
	{\itshape ...}
\end{quote}

\noindent (5) Lines 305, 306, and 478; the use of the term "fixates" rather than "fixes" is not standard, I would change to "fixes" in all cases.
\begin{quote}
	{\itshape ...}
\end{quote}

\noindent (6) Figure 4; same comment as for Figure 1.
\begin{quote}
	{\itshape ...}
\end{quote}

Finally, I thank the authors for their patience; the delay in receiving these reviews lies not with the reviewers, but with myself, due to some personal and family issues that have recently arisen.

\begin{quote}
	{\itshape Thank you again for your efforts and consideration in spite of these challenges. We really appreciate it.}
\end{quote}

\noindent xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx

\subsection*{Reviewer \#1:}

The manuscript reports the derivation, analysis, and example application of an extended frequency-dependent matrix model to explore links between sexual antagonistic selection, population persistence, and the maintenance of heritable variation. The paper is very well written and is straightforward to follow. The derivations are correct, although I did not try to derive all the equations in the appendices. My one comment is how might environmental stochasticity in the rate of selfing and outcrossing impact general conclusions? Other than that, I have no concerns about the science that is reported. My concerns rate the very narrow focus of the paper as written.

The manuscript adds to a growing and important literature where eco-evolutionary feedbacks are examined in structured models, some of which incorporate evolution explicitly. It also contributes nicely to a body of empirical and theoretical work on sexually antagonistic selection.

\begin{quote}
	{\itshape Thank you for the encouraging words.}
\end{quote}

As written the manuscript will not appeal to the broad readership of the American Naturalist because the weakness of the manuscript is the author makes little attempt to fit the work into the broader relevant literature. Instead, the paper has been written for a small audience of experts interested in the demographic consequences of sexually antagonistic selection. As written, the work is suited to a journal such as Journal of Theoretical Biology. This is disappointing, and the authors do themselves no favors by writing for such a niche audience. The manuscript could be well-suited to the American Naturalist if the authors were to engage with the broader literature. This can be done by extending the introduction to link this work to the wider literature on evolutionarily explicit structured models, and to the large empirical literature on sexually antagonistic selection. If they are able to do this appropriately, I believe this paper would make a strong contribution to the American Naturalist.

\begin{quote}
	{\itshape This is a really important comment and we thank you pointing out this major shortcoming. One of our main goals while crafting the first draft of the manuscript was to find a middle ground that would appeal to both demographers and people studying sexual conflict. It seems clear now that we didn't quite acheive that goal. Rather, by making compromises in the introduction we ended up lacking scholarship on both sides. To address your main concern, we have restructured our Introduction to more directly connect with the broader literature on evolutionarily explicit structured models (see {\bf L.XXX--XXX}) and now try to link our overview of sexually antagonistic selection theory with the existing empirical literature (see {\bf L.XXX--XXX}). We hope these changes succeed in broadening the appeal of our manuscript to the wider readership of AmNat.}
\end{quote}

\noindent xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx

\subsection*{Reviewer \#2:}

I have read, with great interest, this manuscript analyzing the effects of sexually antagonistic selection in hermaphrodites at both the population genetics and population dynamics levels. As a representative (I guess) of the general readership of AmNat and also as a population dynamics specialist, I appreciated the description of the model, its parametrization, the way it is constructed and analyzed. Not being a population genetics specialist, it is harder for me to comment on that side of the interpretations, apart from generic comments and in the light of the model’s characteristics. Overall, my take is that this paper is interesting, well written, the model well built and the ideas quite novel; and that it therefore corresponds to what readers except to find in AmNat both in terms of content and quality.

I have one main comment:
-- Discussion: Conclusions are drawn before limitations of the model are mentioned. However, it seems to me that it would be very important to discuss the effect of those limitations on the results themselves and therefore discuss the generality of the results. For instance, what is the potential effect of female dominance in the marriage function on the results and in particular on conclusion of lines 477-480

\begin{quote}
	{\itshape ...}
\end{quote}

\noindent Apart from that, I have only minor comments, that I list below (roughly in decreasing order of importance):
\bigskip

\noindent -- Model: lines 114-195
\begin{itemize}
	\item The matrices H are mentioned before they are introduced. (Therefore, the parenthesis in lines 123-124 cannot be understood by the reader). And then are only defined in the following section	
	\begin{itemize}
		\item It would make sense to reorder the model presentation by presenting the building blocks (arguably U, F and H matrices, but I'll come back to H in the next point), what they do, what they are for; then provide some properties (F independent from selfing/crossing, H from stage, etc ...) and then go into the details of their construction
	\end{itemize}
	\item The main issue I have here, is that H is introduced, but then disappears from the output (see equations 12 and 13 for instance that use H but it does not appear, or the general output eq 9)
	\begin{itemize}
		\item This suggests combining F and H directly into an object.
	\end{itemize}
\end{itemize}

\begin{quote}
	{\itshape ... our response}
\end{quote}

\begin{itemize}
	\item More generally, the matrix formulations of the model may be a bit hard to read for the general reader of AmNat. Getting rid of H (or to the contrary, using it all the way, thus simplifying eq 12 and 13 for instance) and not giving a name to matrix W (which is just a bunch of 0s 1s and 1/2s) are steps that can make it easier for the reader to follow.
	\item It is quite confusing that HX is written as a constant (eq 7) whilst it depends on the frequencies and that HS is written as dependent (eq6) whilst it is constant
	\item As the building blocks are given expressions (eq 10 and 11 for instance) it would be nice to mention the fact that the elements in these will be characterize later (otherwise the reader thinks she/he should already know what the expression of, say, UsAA is)
\end{itemize}

\begin{quote}
	{\itshape ... our response}
\end{quote}
\noindent -- Line 19: I find it weird to write that the "process of adaptation can be impeded by ...changing environmental conditions".

\begin{itemize}
	\item One organism can be adapted to an environment (as in a series of environmental parameters) but also to changes in that environment (plasticity, genetic variance etc ...). It is actually from this consideration that a lot of studies on adaptation and adaptive dynamics.
	\item I think this sentence should be rewritten to make it clear that it is not changes in environmental conditions per se that impede the adaptive process but some of their characteristics (speed, range etc.)
\end{itemize}

\begin{quote}
	{\itshape ... our response}
\end{quote}

\noindent -- Line 23: The link between genetic constraints and genetic trade-offs should be clarified. Or rather a definition of each term provided, for the reader to understand this sentence. I suspect that, for many a reader of AmNat, genetic constraints and genetic trade-offs are equivalent concepts.

\begin{quote}
	{\itshape ... our response}
\end{quote}

\noindent -- Line 30: "fitness trade-offs ... provide ...mechanism for the maintenance of genetic variation". This requires a few words of explanation. Is that a general statement about (fitness) trade-offs? or is it actually particular to SA trade-offs or to trade-offs between fixed heterogeneity components?
\begin{itemize}
	\item For instance, a fitness trade off between fertility at 3 years and fertility at 6 years will have no positive effect on genetic variance. To the contrary, it may lead the organism to evolve an optimal strategy accounting for overall fitness and the trade-offs, vis-à-vis investments in reproduction at those different stages in the life cycle. And as such this may lead to the selection of individuals with such a strategy and therefore to low genetic variance.
\end{itemize}

\begin{quote}
	{\itshape ... our response}
\end{quote}

\noindent -- Line 264: this is an important point that is a bit unclear here, according to me. Could you clarify that what you call demographic rates are the transitions in the overall model and so include both demography and genetics? Clarify also that the model is not density dependent because it only depends on frequencies ... and maybe relate it to a general equation where A does not depend on n per se but on the q frequencies.

\begin{quote}
	{\itshape ... our response}
\end{quote}

- Line 35: SA selection arises also from non-genetic trade-offs, which makes the formulation of that sentence a bit weird.

\begin{quote}
	{\itshape ... our response}
\end{quote}

- Lines 289-292: inbreeding depression is expressed as a function of C. However, C is not written in the equation but L which depends on C. C is the introduced parameter and so should appear. If the expression is too complicated then it should be described literally, but there is no need to introduce L here, in any case.

\begin{quote}
	{\itshape ... our response}
\end{quote}

- Line 42: Define analogous here

\begin{quote}
	{\itshape ... our response}
\end{quote}

- Line 54: maybe a few words to remind the reader about what dioecious and hermaphrodite organisms are.

\begin{quote}
	{\itshape ... our response}
\end{quote}

- Line 80: just a few words to clarify "obligate outcrossing" (something like (no selfing allowed) maybe)

\begin{quote}
	{\itshape ... our response}
\end{quote}

- Line 233: f and f’ are called "adult fertilities". However, they correspond to that rate only for certain specific genotypes. Maybe clarify that by calling them "adult maximum fertilities" ?

\begin{quote}
	{\itshape ... our response}
\end{quote}

- Results: it took me a while to figure out Fig 1 (mainly because I was looking at it in B and W). I am just wondering whether it would be easier to read with only one f per chart. Maybe as a separate chart that illustrates the general shape of the parameter space for polym * extinction for only one C, one h and one f... then followed by these 6 graphs as they are now.

\begin{quote}
	{\itshape ... our response}
\end{quote}

\noindent-- Maybe mention once, in a few words, what is meant by "segregate as a diallelic locus".
\begin{quote}
	{\itshape ... our response}
\end{quote}

\noindent-- Fig 4 : they are supposed to be dashed lines and dotted lines but all I see are dashed lines with various dashes. I found 4B in particular, hard to understand at first sight.

\begin{quote}
	{\itshape ... our response}
\end{quote}
\bigskip

\noindent Typos:
\smallskip

\noindent-- The 0s representing 0 matrices in eq 5 are not bold
\begin{quote}
	{\itshape ... our response}
\end{quote}

\noindent-- In same equation 1w is not defined
\begin{quote}
	{\itshape ... our response}
\end{quote}

\noindent-- Line 142, 2 "the"s
\begin{quote}
	{\itshape ... our response}
\end{quote}

\noindent-- Line 553 : not a typo, but the "was that how" in the middle of the sentence makes it hard to read
\begin{quote}
	{\itshape ... our response}
\end{quote}

\end{document}

### 2021-11-21 doc/SA-Hermaphrodites-wDemography-R1.tex
Setup for revisions
This commit restructures the repo directory tree to set up a legacy
directory for the original submission, and creates new .tex files for
the revised ms *_R1.tex etc, and a response letter to the editor and
reviewer's comments.

--- 2021-06-04 doc/SA-Hermaphrodites-wDemography.tex
@@
 In fact, the model reduces to the stage-structured model of \citet{deVriesCaswell2019a} under obligate outcrossing.
-A full derivation of the model and analyses is presented in the Online Supplementary Material, and all computer code necessary to reproduce the results are available at \url{GitHub Link omitted for anonymity}.%\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
+A full derivation of the model and analyses is presented in the Online Supplementary Material, and all computer code necessary to reproduce the results are available at \url{https://anonymous.4open.science/r/SA-Hermaphrodites-wDemography-BF66}.%\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
 

### 2021-12-10 doc/ResponseLetter-R1.tex
Reviews1_part1
Edit methods,  paragraph intro about models combined population structure&genetics, fixed some minor things from the minor edits list (see response letter)

@@
 \begin{quote}
-	{\itshape ...}
+	{\itshape Fixed}
 \end{quote}
 
-\noindent (4) Figure 1; I generally prefer axes labels to have clear meanings, rather than just symbols – here, the vertical and horizontal axes give the female (vertical) vs.
-male (horizontal) selection on reproductive function.
+\noindent (4) \textcolor{red}{Figure 1; I generally prefer axes labels to have clear meanings, rather than just symbols – here, the vertical and horizontal axes give the female (vertical) vs.
+male (horizontal) selection on reproductive function.}
 \begin{quote}
@@
 \begin{quote}
-	{\itshape ...}
+	{\itshape Fixed \textcolor{red}{Note to Colin: I find that fixes sounds a bit odd? I changed it to "goes to fixation" in one instance. I trust your final judgement on this as our native.}}
 \end{quote}
 
-\noindent (6) Figure 4; same comment as for Figure 1.
+\noindent (6) \textcolor{red}{Figure 4; same comment as for Figure 1.}
 \begin{quote}
@@
 One of our main goals while crafting the first draft of the manuscript was to find a middle ground that would appeal to both demographers and people studying sexual conflict.
-It seems clear now that we didn't quite acheive that goal.
+It seems clear now that we didn't quite achieve that goal.
 Rather, by making compromises in the introduction we ended up lacking scholarship on both sides.
@@
 	\item It is quite confusing that HX is written as a constant (eq 7) whilst it depends on the frequencies and that HS is written as dependent (eq6) whilst it is constant
+	\begin{quote}
+	{\itshape That's a mistake on our part, apologies.}
+\end{quote}
 	\item As the building blocks are given expressions (eq 10 and 11 for instance) it would be nice to mention the fact that the elements in these will be characterize later (otherwise the reader thinks she/he should already know what the expression of, say, UsAA is)
@@
 \begin{quote}
-	{\itshape ... our response}
+	{\itshape Good point, fixed.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape ... our response}
+	{\itshape Fixed}
 \end{quote}

### 2021-12-10 doc/SA-Hermaphrodites-wDemography-R1.tex
Reviews1_part1
Edit methods,  paragraph intro about models combined population structure&genetics, fixed some minor things from the minor edits list (see response letter)

--- 2021-06-04 doc/SA-Hermaphrodites-wDemography.tex
@@
 To take into account the consequences of population dynamics under SA selection, a model is required that links changes in genotype distributions to population dynamics.
+\textcolor{blue}{One of the earliest such models \citep{roughgarden1971density} combined a population genetic model with a density-dependent population model with no intraspecific demographic structure.
+Charlesworth then combined an age-structured population model with population genetics in a series of papers, consolidated in \cite{BCharlesworth1991}.
+\cite{orive1995senescence} extended this work to study life-history evolution in organisms with both clonal and sexual reproduction (see also \cite{orive2001somatic}).
+\cite{ronce2000kin} added spatial structure and stochasticity to age-structured genetic models to model kin selection, a body of work consolidated in \cite{rousset2004genetic}.
+Of course there are many other approaches to studying eco-evolutionary dynamics, e.g.
+combining quantitative genetics models with demographic structure (e.g.
+\cite{coulson2006putting,barfield2011evolution,childs2016evolution,orive2017effects}) or adaptive dynamics approaches (e.g.
+\cite{dieckmann2002adaptive,durinx2008adaptive,ten2020habitat}).
+For a comprehensive review of the eco-evolutionary literature, see \cite{fussmann2007eco}, or more recently, \cite{govaert2019eco}.}
+
 The potential demographic costs of sexual antagonism were pointed out by \cite{kokko2003sexy}, but few papers since then have explicitly incorporated demography into models of sexual antagonism (with the notable exception of \citealt{harts2014demography}).
@@
 Here, we briefly describe a matrix model incorporating multiple life-cycle stages and a single diallelic locus under SA selection on male and female fertility for populations of partially-selfing simultaneous hermaphrodites.
-The derivation and some of the key results follow closely those presented in \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b}.
+\textcolor{blue}{Rather than go through the full mathematical derivation of the general model, we showcase a two-stage example of our model and discuss the biological assumptions that go into the model.
+The derivation of the general model for any number of age or stage classes is in the Online Supplementary Material.} The derivation follows closely the logic presented in \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b}.
 In fact, the model reduces to the stage-structured model of \citet{deVriesCaswell2019a} under obligate outcrossing.
-A full derivation of the model and analyses is presented in the Online Supplementary Material, and all computer code necessary to reproduce the results are available at \url{GitHub Link omitted for anonymity}.%\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
+All computer code necessary to reproduce the results are available at \url{https://anonymous.4open.science/r/SA-Hermaphrodites-wDemography-BF66}.%\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
 
@@
 The matrices $\mbf{F}_{i}$ and $\mbf{F}^{\prime}_{i}$ contain the genotype $\times$ stage specific contribution of genotype $i$ to the female and male gamete pools, respectively, and therefore to zygotes in the next generation.
-We assume that whether individuals were produced through selfing or outcrossing does not affect their fecundity or mating success, that is, we assume  $\mbf{F}^X_{i}=\mbf{F}^S_{i}=\mbf{F}_{i}$ and $\mbf{F}^{\prime X}_{i}=\mbf{F}^{\prime S}_{i}=\mbf{F}^{\prime}_{i}$.
+We assume that whether individuals were produced through selfing or outcrossing does not affect their fecundity or mating success, that is, we assume  $\mbf{F}^X_{i}=\mbf{F}^S_{i}=\mbf{F}_{i}$ and $\mbf{F}^{\prime X}_{i}=\mbf{F}^{\prime S}_{i}=\mbf{F}^{\prime}_{i}$, respectively.
 Deviations from this assumption are straightforward to incorporate but beyond the scope of this paper.
 
-% Notational note: The superscript on a matrix in age$\times$stage models conventionally indicates the stage or age of the subsection of the population vector that the matrix is acting on (see for example \citep{CaswellEtAl2018}).
-For example, $\mbf{U}^{S}_{i}$, acts on the vector of individuals of genotype $i$ that were produced through selfing.
-In the following, we deviate from this convention by also using $S$ and $X$ as superscripts to indicate production of offspring through current selfing or reproduction.
-That is,  we will write production $\mcal{F}^S$ and $\mcal{F}^X$ to denote offspring production by self-fertilization and outcrossing.
-Likewise, we define the matrices $\mbf{H}^S_{j}(\tilde{\mbf{n}})$ and $\mbf{H}^X_{j}(\tilde{\mbf{n}})$ as the map of the genotypes of the parent in stage $j$ to the genotypes of their offspring produced by selfing and outcrossing, respectively.
-The $(k, l)$ entry of $\mbf{H}^{X}_{j}$ and $\mbf{H}^{S}_{j}$ is the probability that an offspring of a genotype $l$ mother, of stage $j$, has genotype $k$.
-Since we assume that how-you-were-produced does not affect how you reproduce, these matrices will act on both $\mbf{n}^{S}$ and $\mbf{n}^{X}$.
-
-For the purpose of this article, we assume that mating is random with respect to stage and hence that the parent-offspring map is the same for all stages (i.e., $\mbf{H}^S_{j}(\tilde{\mbf{n}}) = \mbf{H}^S(\tilde{\mbf{n}})$, and $\mbf{H}^X_{j}(\tilde{\mbf{n}}) = \mbf{H}^X(\tilde{\mbf{n}})$; where $j \in \{1,\omega\}$).
-The matrices $\mbf{H}^S(\tilde{\mbf{n}})$ and $\mbf{H}^X(\tilde{\mbf{n}})$ contain the population genetic processes and are presented in the next section.
-
-%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection*{Sexually antagonistic selection \& inbreeding depression} \label{sec:SAsel}
+
+\textcolor{blue}{To demonstrate how to construct and analyze a genotype $\times$ stage-classified model, we will now construct an example model for a hypothetical species with intralocus sexual conflict via the two sex functions.
+}For the sake of simplicity, we assume our hypothetical species has a life cycle with only two stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active.
+Suppose that there is a genetic trade-off between the sex-functions at a single diallelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect.
+Following convention, we parameterize the fertility component of fitness for each genotype through each sex function, $w_{i}$ and $w^{\prime}_{i}$, to be bounded by $[0,1]$, with dominance and selection coefficients $h_f, s_f$ and $h_m, s_m$ determining the decrease in fertility through each sex function relative to the most fit genotype ($AA$ has highest female fertility, $aa$ the highest male fertility; see Table \ref{tab:Fitness}).
+
+The SA locus does not affect survival or transition rates.
+However, the survival matrices can be used to model the fitness effects of inbreeding depression at later stages of development by allowing different stage-specific survival and transition rates for individuals produced by self-fertilization vs.~outcrossing.
+By contrast, the parameter $\delta$ only affects inbreeding depression through viability of selfed ovules.
+With this in mind, we define survival matrices for individuals produced by selfing and outcrossing as follows:
+\begin{linenomath*}
+\begin{equation} \label{eq:US}
+	\mbf{U}^S = \left(
+					\begin{array}{cc}
+						\sigma_j (1 - \delta_j) \big(1 - \gamma (1 - \delta_{\gamma}) \big) & 0 \\
+						\sigma_j (1 - \delta_j) \gamma (1 - \delta_{\gamma})      & \sigma_a (1 - \delta_a)
+					\end{array}
+				\right) \\
+\end{equation}
+\end{linenomath*}
+\noindent and 
+\begin{linenomath*}
+\begin{equation}\label{eq:UX}
+	\mbf{U}^X = \left(
+					\begin{array}{cc}
+						\sigma_j(1 - \gamma) & 0 \\
+						\sigma_j \gamma      & \sigma_a
+					\end{array}
+				\right)
+\end{equation}
+\end{linenomath*}
+
+\noindent where $\sigma_j$ and $\sigma_a$ are the juvenile and adult stage survival rates, $\gamma$ is the maturation rate from juvenile to adult stages, and the corresponding $\delta_j$, $\delta_a$, and $\delta_{\gamma}$ terms denote the proportional decreases in stage-specific survival and transition rates due to inbreeding depression (i.e., deleterious effects of inbreeding at later life-history stages; e.g., \citealt{HarderRoutely2006}).
+For simplicity, we assume survival and transition rates are constant among genotypes.
+
+Throughout our analyses, we distinguish between early- and late-acting inbreeding depression.
+We quantify early-acting inbreeding depression using $\delta$, and late-acting inbreeding depression using $\delta_i$ (where $i \in \{j,a,\gamma\}$).
+$\delta$ denotes the fraction of self-fertilized ovules that do not develop into juveniles due to inbreeding depression.
+An important difference between early- and late-acting inbreeding depression in the model is that $\delta$ affects the production of new individuals, whereas the $\delta_i$ affect the demographic rates of extant individuals, contained in $\mbf{U}^S$ (see Eq.~\ref{eq:US}).
+
+The fertility matrices through female and male function are
+\begin{linenomath*}
+\begin{equation}\label{eq:F_female}
+	\mbf{F}_{i} = \left(
+					\begin{array}{cc}
+						0 & f w_{i} \\
+						0 & 0
+					\end{array}
+				\right),
+\end{equation}
+\end{linenomath*}
+\noindent and
+\begin{linenomath*}
+\begin{equation}\label{eq:F_male}
+	\mbf{F}^{\prime}_{i} = \left(
+					\begin{array}{cc}
+						0 & f^{\prime} w^{\prime}_{i} \\
+						0 & 0
+					\end{array}
+				\right),
+\end{equation}
+\end{linenomath*}
+
+\noindent where $f$ and $f^{\prime}$ represent maximum adult fertilities, and $w_{i}$ and $w^{\prime}_{i}$ the genotypic relative scaling factors for female and male sex-functions (see Table \ref{tab:Fitness}).
+
+\begin{table}[htbp]
+ \centering
+ \caption{\bf Relative fertilities for Sexually Antagonistic selection ($w_{i}$)}
+\label{tab:Fitness}
+\begin{tabular}{lccc}
+ \toprule
+					&  \multicolumn{3}{c}{{\textit{Genotype}}} \\ 
+\cline{2-4}
+					& $AA$			& $Aa$ 					& $aa$ 		\\ \hline
+Female function ($w_{i}$):	& $1$		& $1 - h_f s_f$	& $1 - s_f$ \\	
+Male function ($w^{\prime}_{i}$):		& $1 - s_m$& $1 - h_m s_m$	& $1$ 		\\	
+\hline
+\end{tabular}
+\end{table}
+
 \subsubsection*{Mating and offspring production under partial selfing}
@@
 
-Because we assume that inbreeding depression only affects offspring viability, the allele frequency in the male mating population is obtained by simply summing the vectors of individuals produced through selfing and outcrossing:
+\textcolor{blue}{ For the purpose of this article, we also assume that females in the population all have the same mating preferences regardless of female age, stage, or genotype.
+The model can be extended to consider assortative mating, but that is beyond the scope of the current paper.
+In addition, we assume that males produced through selfing or through outcrossing do not differ in their mating success.
+Given these assumptions, the allele frequencies in the male mating population can obtained by simply summing the vectors of individuals produced through selfing and outcrossing:
 \begin{linenomath*}
@@
  \end{equation}
-\end{linenomath*}
-Depending on their stage and genotype at the SA locus, individuals contribute male gametes to the overall population gamete pool according to the following equation
-\begin{linenomath*}
-\begin{equation}
-		\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime}= 
-		\left(
-			\begin{array}{ccc}
-				\mbf{1}^{\intercal}_{\omega} & \frac{1}{2} \mbf{1}^{\intercal}_{\omega} & 0 \\
-				0 & \frac{1}{2} \mbf{1}^{\intercal}_{\omega} & \mbf{1}^{\intercal}_{\omega} \\
-			\end{array} \right)
-		\left(
-			\begin{array}{ccc}
-				\mbf{F}^{\prime}_{AA} & 0 & 0 \\
-				0 & \mbf{F}^{\prime}_{Aa} & 0 \\
-				0 & 0 & \mbf{F}^{\prime}_{aa} \\
-			\end{array} \right)
-		\left(
-			\begin{array}{c}
-				\mbf{p}^{\prime}_{AA} \\
-				\mbf{p}^{\prime}_{Aa} \\
-				\mbf{p}^{\prime}_{aa} \\
-			\end{array} \right).
-\end{equation}
-\end{linenomath*}
-
-\noindent The matrix $\mbb{F}^{\prime}$ is composed of genotype-specific fertility matrices, and operates on the the vector of genotype frequencies to give their relative contribution of each genotype to the gamete pool.
-The matrix $\mbf{W}^{\prime}$ converts these contributions to allele numbers.
-Normalizing the resulting vector gives the allele frequencies in the male gamete pool:
+\end{linenomath*}}
+
+\textcolor{blue}{Males contribute  gametes to the population gamete pool according to their mating success, which depends on their stage and genotype at the SA locus.
+The allele frequencies in the male gamete pool are then calculated as follows (see Online Supplementary Materials for a derivation of this equation),
 \begin{linenomath*}
@@
 			q^{\prime}_{a} \\
+		\end{array} \right) \propto 
+				\left(
+		\begin{array}{c}
+			\mbf{1}^{\intercal}_{\omega}\mbf{F}^{\prime}_{AA}\mbf{p}^{\prime}_{AA}+ \frac{1}{2} \mbf{1}^{\intercal}_{\omega}  \mbf{F}^{\prime}_{Aa} \mbf{p}^{\prime}_{Aa}  \\
+		\frac{1}{2} \mbf{1}^{\intercal}_{\omega}  \mbf{F}^{\prime}_{Aa} \mbf{p}^{\prime}_{Aa} + \mbf{1}^{\intercal}_{\omega}\mbf{F}^{\prime}_{aa}\mbf{p}^{\prime}_{aa}\\
+		\end{array} \right),
+\end{equation}
+\end{linenomath*}
+where equality is obtained by normalizing the frequencies such that they add to one.
+The first entry is simply summing the stage-specific contributions to the gamete pool of heterozygotes and $AA$ homozygotes; the second entry sums the stage-specific contributions to the gamete pool of heterozygotes and $aa$ homozygotes.
+}
+
+\textcolor{blue}{ FWe now have all the ingredients to calculate the allele frequencies in the (male) gamete pool for the two stage example model in equations \ref{eq:F_male}, \ref{eq:maleGametePool}, and Table \ref{tab:Fitness}.
+Denoting the proportion of adults of genotype $i$ with $p^A_i$, the allele frequencies in the male gamete pool are given by
+\begin{linenomath*}
+\begin{equation} \label{eq:maleGametePool_example}
+\left(
+		\begin{array}{c}
+			q^{\prime}_{A} \\
+			q^{\prime}_{a} \\
 		\end{array} \right) = 
-			\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime}}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime} \|} = 
-				\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)\|}.
+				\frac{1}{(1-s_m)p^A_{aa}+ (1-hs_m)p^A_{Aa}+p_{aa}}\left(
+		\begin{array}{c}
+(1-s_m)p^A_{AA}	+\frac{1}{2}(1-hs_m)p^A_{Aa}\\
+p^A_{aa}+\frac{1}{2}(1-hs_m)p^A_{Aa}
+		\end{array} \right).
 \end{equation}
 \end{linenomath*}
-The key difference between selfing and outcrossing reproduction can be seen in the parent-offspring maps, which reflect the joint effects of meiosis and mating on the distribution of offspring genotypes.
-% \textcolor{red}{
-% I'm doubting whether to move the next paragraph to the appendices as well.
-Equation (\ref{eq:maleGametePool}) is all we strictly need.
-If we keep the parent-offspring maps, we can explain in the next section how the matrices $\mcal{F}^S$ and $\mcal{F}^X$ have the same structure.
-Would that be helpful?
-Or would it be easier to just explain in the next section what the structure of $\mcal{F}^S$ and $\mcal{F}^X$ is?
+Note that this equation is identical to the classical equation for frequencies of male gametes in a sexually antagonistic model except that now we have to specify that only adults contribute to the gamete pool whereas in a classical population genetic model with no population structure, we do not distinguish between reproducing adults and non-reproductive juveniles.
+Of course equation (\ref{eq:maleGametePool_example}) starts deviating from the classical population genetics case more drastically if we include multiple reproducing adult age or stage classes (see equation (\ref{eq:maleGametePool})).
 }
 
-The parent-offspring map for outcrossing is a function of the allele frequencies in the male gamete pool, and is given by 
-\begin{linenomath*}
-\begin{equation} \label{eq:HX}
-	\mbf{H}^X = 
-			\left(
-			\begin{array}{ccc}
-				q^{\prime}_{A} & \frac{1}{2} q^{\prime}_{A} & 0 \\
-				q^{\prime}_{a} & \frac{1}{2} & q^{\prime}_{A}  \\
-				0 & \frac{1}{2} q^{\prime}_{a} & q^{\prime}_{a} \\
-			\end{array} \right).
-\end{equation}
-\end{linenomath*}
-\noindent From left to right, the columns of matrix $\mbf{H}^X$ give the genotype distribution of outcrossed offspring produced by a maternal parent of each genotype ($AA$, $Aa$, and $aa$ respectively).
-
-The parent-offspring map for reproduction via self-fertilization differs from the outcrossing parent-offspring map.
-As in previous models of partial selfing, we assume that individuals produce enough male gametes to easily self-fertilize all of their ovules, and that self-fertilization involves little or no selection via the male sex function from external factors (e.g., pollinator visitation) relative to outcrossing (\citealt{Charlesworth1978a,JordanConnallon2014,Olito2017}, but see \citealt{Tazzyman2015}).
-Under these assumptions, the genotype distributions of selfed offspring are determined entirely by the parental genotype and the probabilities of segregation and fertilization during and after meiosis:
-\begin{linenomath*}
-\begin{equation} \label{eq:HS}
-	\mbf{H}^S(\tilde{\mbf{n}}) = 
-			\left(
-			\begin{array}{ccc}
-				1 & 1/4 & 0 \\
-				0 & 1/2 & 0 \\
-				0 & 1/4 & 1 \\
-			\end{array} \right).
-\end{equation}
-\end{linenomath*}
-\noindent The columns of $\mbf{H}^S(\tilde{\mbf{n}})$ again give the selfed offspring genotype distributions for parental genotypes of $AA$, $Aa$, and $aa$, respectively.
-As described in the next section, these parent-offspring maps influence the projection matrix by altering the fertilitiy matrices for individuals produced by selfing and outcrossing (see Eq(\ref{eq:BlkFS}) and Eq(\ref{eq:BlkFX}) below).
-
-Following previous theory for partially selfing populations \cite[e.g.,][]{Charlesworth2010,JordanConnallon2014,Glemin2021}, we use a 'fixed prior selfing' model where all individuals are assumed to self-fertilize at a constant rate, $C$, prior to receiving outcross pollen, for reasons of analytic tractability.
-In the Online Supplement, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
-
-%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection*{Population projection}
-Using the component matrices described above (the survival matrices, $\mathbf{U}^X_i$, $\mathbf{U}^S_i$, the fertility and mating success matrices, $\mathbf{F}_i$, $\mathbf{F}^\prime_i$, and the parent-offspring matrices $\mbf{H}^S(\tilde{\mbf{n}})$ and $\mbf{H}^X$), we construct the population projection matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ using the vec-permutation approach of \citet{CaswellEtAl2018}, see Online Supplementary Material for the step by step construction of the model.
-The matrix that projects the eco-evolutionary dynamics is:
+Using the component matrices described above (the survival matrices, $\mathbf{U}^X_i$, $\mathbf{U}^S_i$, the fertility and mating success matrices, $\mathbf{F}_i$, $\mathbf{F}^\prime_i$) and the allele frequencies in the gamete pool, we can construct the population projection matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$, see Online Supplementary Material for the step by step construction of the model.
+The resulting matrix that projects the eco-evolutionary dynamics is:
 \begin{linenomath*}
@@
 Since individuals cannot change their genotype once they are born, the survival matrices are block diagonal.
-Similarly, we construct fertility matrices for individuals produced through selfing and outcrossing,
+
+The fertility matrices for individuals produced through selfing and outcrossing are given by, 
 \begin{linenomath*}
@@
 			\begin{array}{ccc}
-				\mbf{F}_{AA} & \frac{1}{4} \mbf{F}_{Aa} & 0 \\
-				0 & \frac{1}{2} \mbf{F}_{Aa} & 0 \\
-				0 & \frac{1}{4} \mbf{F}_{Aa} & \mbf{F}_{aa}\\
+				\mbf{F}_{AA} & \frac{1}{4} \mbf{F}_{Aa} & \mbf{0}  \\
+				\mbf{0}  & \frac{1}{2} \mbf{F}_{Aa} & \mbf{0}  \\
+				\mbf{0} & \frac{1}{4} \mbf{F}_{Aa} & \mbf{F}_{aa}\\
 			\end{array} \right), 
@@
 			\begin{array}{ccc}
-				q^{\prime}_{A} \mbf{F}_{AA} & \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} & 0 \\ 
+				q^{\prime}_{A} \mbf{F}_{AA} & \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} & \mbf{0}  \\ 
 				q^{\prime}_{a} \mbf{F}_{AA} & \frac{1}{2} \mbf{F}_{Aa} & q^{\prime}_{A} \mbf{F}_{aa}  \\ 
-				0 & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} & q^{\prime}_{a} \mbf{F}_{aa}  \\
+				\mbf{0}  & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} & q^{\prime}_{a} \mbf{F}_{aa}  \\
 			\end{array} \right),
@@
 \end{linenomath*}
-\noindent where $q^{\prime}_A$ and $q^{\prime}_a$ are given by Eq(\ref{eq:maleGametePool}).
+\noindent where $q^{\prime}_A$ and $q^{\prime}_a$ are the frequencies of alleles $A$ and $a$ in the gamete pool, given by Eq(\ref{eq:maleGametePool}).
 
@@
 $\mcal{F}^{S},\, \mcal{F}^{X}$ & Block diagonal selfed/outcrossed fertility matrix & $\omega g \times \omega$g \\
-$\mbf{H}^{S}_{i}$,\, $\mbf{H}^{X}_{i}$ & Parent-offspring genotype maps & $g \times g$ \\
-$\mbf{W}$   & Allele segregation matrix & $2 \times \omega $g \\
 % $\mbf{K}$   & Vec-permutation matrix & $g \times g$ \\
@@
 $C$ & The population rate of self-fertilization & \\
-$\delta,\, \delta_j,\,\delta_a,\,\delta_{\gamma}$ & Inbreeding depression terms for ovule viability, survival, and transition rates respectively & \\
+$\delta$ &\textcolor{blue}{ Inbreeding depression effect on ovule viability}&\\
+$ \delta_j$ & \textcolor{blue}{ Inbreeding depression effect on juvenile survival}& \\
+$\delta_a$ & \textcolor{blue}{ Inbreeding depression effect on adult survival}& \\
+$\delta_{\gamma}$ &\textcolor{blue}{  Inbreeding depression effect on maturation rate} & \\
 $f_i,\,f^{\prime}_i$ & Adult fertilities through female and male sex functions respectively & \\
@@
 
-%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Sexually antagonistic selection \& inbreeding depression} \label{sec:SAsel}
-
-We now construct and analyze a genotype $\times$ stage-classified model for a hypothetical species with intralocus sexual conflict via the two sex functions.
-For the sake of simplicity, we assume our hypothetical species has a life cycle with only two stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active.
-Suppose that there is a genetic trade-off between the sex-functions at a single diallelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect.
-Following convention, we parameterize the fertility component of fitness for each genotype through each sex function, $w_{i}$ and $w^{\prime}_{i}$, to be bounded by $[0,1]$, with dominance and selection coefficients $h_f, s_f$ and $h_m, s_m$ determining the decrease in fertility through each sex function relative to the most fit genotype ($AA$ has highest female fertility, $aa$ the highest male fertility; see Table \ref{tab:Fitness}).
-
-The SA locus does not affect survival or transition rates.
-However, the survival matrices can be used to model the fitness effects of inbreeding depression at later stages of development by allowing different stage-specific survival and transition rates for individuals produced by self-fertilization vs.~outcrossing.
-By contrast, the parameter $\delta$ only affects inbreeding depression through viability of selfed ovules.
-With this in mind, we define survival matrices for individuals produced by selfing and outcrossing as follows:
-\begin{linenomath*}
-\begin{equation} \label{eq:US}
-	\mbf{U}^S = \left(
-					\begin{array}{cc}
-						\sigma_j (1 - \delta_j) \big(1 - \gamma (1 - \delta_{\gamma}) \big) & 0 \\
-						\sigma_j (1 - \delta_j) \gamma (1 - \delta_{\gamma})      & \sigma_a (1 - \delta_a)
-					\end{array}
-				\right) \\
-\end{equation}
-\end{linenomath*}
-\noindent and 
-\begin{linenomath*}
-\begin{equation}\label{eq:UX}
-	\mbf{U}^X = \left(
-					\begin{array}{cc}
-						\sigma_j(1 - \gamma) & 0 \\
-						\sigma_j \gamma      & \sigma_a
-					\end{array}
-				\right)
-\end{equation}
-\end{linenomath*}
-
-\noindent where $\sigma_j$ and $\sigma_a$ are the juvenile and adult stage survival rates, $\gamma$ is the maturation rate from juvenile to adult stages, and the corresponding $\delta_j$, $\delta_a$, and $\delta_{\gamma}$ terms denote the proportional decreases in stage-specific survival and transition rates due to inbreeding depression (i.e., deleterious effects of inbreeding at later life-history stages; e.g., \citealt{HarderRoutely2006}).
-For simplicity, we assume survival and transition rates are constant among genotypes.
-
-Throughout our analyses, we distinguish between early- and late-acting inbreeding depression.
-We quantify early-acting inbreeding depression using $\delta$, and late-acting inbreeding depression using $\delta_i$ (where $i \in \{j,a,\gamma\}$).
-$\delta$ denotes the fraction of self-fertilized ovules that do not develop into juveniles due to inbreeding depression.
-An important difference between early- and late-acting inbreeding depression in the model is that $\delta$ affects the production of new individuals, whereas the $\delta_i$ affect the demographic rates of extant individuals, contained in $\mbf{U}^S$ (see Eq.~\ref{eq:US}).
-
-The fertility matrices through female and male function are
-\begin{linenomath*}
-\begin{equation}\label{eq:FS}
-	\mbf{F}_{i} = \left(
-					\begin{array}{cc}
-						0 & f w_{i} \\
-						0 & 0
-					\end{array}
-				\right),
-\end{equation}
-\end{linenomath*}
-\noindent and
-\begin{linenomath*}
-\begin{equation}\label{eq:FX}
-	\mbf{F}^{\prime}_{i} = \left(
-					\begin{array}{cc}
-						0 & f^{\prime} w^{\prime}_{i} \\
-						0 & 0
-					\end{array}
-				\right),
-\end{equation}
-\end{linenomath*}
-
-\noindent where $f$ and $f^{\prime}$ represent adult fertilities, and $w_{i}$ and $w^{\prime}_{i}$ the genotypic relative scaling factors for female and male sex-functions (see Table \ref{tab:Fitness}).
-
-\begin{table}[htbp]
- \centering
- \caption{\bf Relative fertilities for Sexually Antagonistic selection ($w_{i}$)}
-\label{tab:Fitness}
-\begin{tabular}{lccc}
- \toprule
-					&  \multicolumn{3}{c}{{\textit{Genotype}}} \\ 
-\cline{2-4}
-					& $AA$			& $Aa$ 					& $aa$ 		\\ \hline
-Female function ($w_{i}$):	& $1$		& $1 - h_f s_f$	& $1 - s_f$ \\	
-Male function ($w^{\prime}_{i}$):		& $1 - s_m$& $1 - h_m s_m$	& $1$ 		\\	
-\hline
-\end{tabular}
-\end{table}
-
 Iterating the projection matrix, Eq(\ref{eq:Atilde}) with the above demographic matrices, given an initial population state vector, allows numerical simulation of the eco-evolutionary dynamics for selection operating on any of the sex-function, or stage-specific demographic parameters.
@@
 
+% Notational note: The superscript on a matrix in age$\times$stage models conventionally indicates the stage or age of the subsection of the population vector that the matrix is acting on (see for example \citep{CaswellEtAl2018}).
+For example, $\mbf{U}^{S}_{i}$, acts on the vector of individuals of genotype $i$ that were produced through selfing.
+In the following, we deviate from this convention by also using $S$ and $X$ as superscripts to indicate production of offspring through current selfing or reproduction.
+That is,  we will write production $\mcal{F}^S$ and $\mcal{F}^X$ to denote offspring production by self-fertilization and outcrossing.
+Likewise, we define the matrices $\mbf{H}^S_{j}(\tilde{\mbf{n}})$ and $\mbf{H}^X_{j}(\tilde{\mbf{n}})$ as the map of the genotypes of the parent in stage $j$ to the genotypes of their offspring produced by selfing and outcrossing, respectively.
+The $(k, l)$ entry of $\mbf{H}^{X}_{j}$ and $\mbf{H}^{S}_{j}$ is the probability that an offspring of a genotype $l$ mother, of stage $j$, has genotype $k$.
+Since we assume that how-you-were-produced does not affect how you reproduce, these matrices will act on both $\mbf{n}^{S}$ and $\mbf{n}^{X}$.
+
+%%%%%%%%%%%%%%%%%%%%%%
+% \subsubsection*{Mating and offspring production under partial selfing}
+%  Outcrossing in our model proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}.
+That is, each individual's genotype determines both the number of ovules produced and the paternal mating success, broadly defined.
+For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986, WangBarrett2020, Harder2016}.
+Note, however, that by modeling paternal relative mating success rather than pollen/sperm production, we implicitly assume female demographic dominance (i.e., production/transport of male gametes does not limit ovule fertilization and hence population growth), a point we return to in the Discussion.
+For the purpose of this article, we also assume that mating is random with respect to stage.
+
+%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%
@@
 The funnel between the thick black lines indicates the area where the sexually antagostic alleles coexist.
-Above the funnel, the female-benefit and male-detriment allele goes to fixation; below the funnel, the male-benefit and female-detriment allele fixates.
-Much of the region where the male-benefit and female-detriment allele fixates is demographically inviable due to the demographic consequences of reduced female fitness in this region.
+Above the funnel, the female-benefit and male-detriment allele goes to fixation; below the funnel, the male-benefit and female-detriment allele \textcolor{blue}{goes to fixation}.
+Much of the region where the male-benefit and female-detriment allele \textcolor{blue}{fixes} is demographically inviable due to the demographic consequences of reduced female fitness in this region.
 The SA polymorphisms that remain viable at lower fertility values correspond to regions where the female-deleterious allele is predicted to segregate at low frequencies, or in regions with weak selection through both sex functions, an asymmetry that reflects the assumption of female demographic dominance.
@@
 Since intrinsic growth rates far exceeding one suggest rapid exponential growth, they are generally rare (with the notable exception of recently introduced, invasive populations), suggesting that our model predictions may be highly relevant for SA polymorphism in many real-world populations.
-In addition, we find that much of the parameter space where a male-beneficial allele fixates is demographically inviable for populations whose intrinsic growth rates are close to one prior to invasion.
+In addition, we find that much of the parameter space where a male-beneficial allele \textcolor{blue}{fixes} is demographically inviable for populations whose intrinsic growth rates are close to one prior to invasion.
 This demographic consequence of masculinization is rarely considered \citep{hitchcock2020gene}.
@@
 {\footnotesize
-\begin{equation} \label{eq:coexist_AA}
+\begin{equation} \label{eq:coexist_bigAA}
 	\lambda_{AA} < 
@@
 
-In our numerical simulations, equations Eq(\ref{eq:coexist_AA}) and Eq(\ref{eq:coexist_aa}) were evaluated to determine the invasion thresholds for the relvant parameter conditions described in each figure.
+In our numerical simulations, equations Eq(\ref{eq:coexist_bigAA}) and Eq(\ref{eq:coexist_aa}) were evaluated to determine the invasion thresholds for the relvant parameter conditions described in each figure.
 %\renewcommand{\theequation}{B\arabic{equation}}

### 2021-12-10 doc/Supplements/SupplementaryMaterial.tex
Edits to supp. mat.
Reviews1_part1
Edit methods,  paragraph intro about models combined population structure&genetics, fixed some minor things from the minor edits list (see response letter)

@@
 %\rhead{Supplement to Olito \& de Vries, "Costs of SA selection" \textit{Am.~Nat.}}
-\rhead{Supplement to "Costs of SA selection" \textit{Am.~Nat.}}
+\rhead{Supplement to "Demographic costs of SA selection" \textit{Am.~Nat.}}
 \setlength{\headsep}{0.3in}  
@@
 \section{Construction of the population projection matrix}
+\subsubsection*{Mating and offspring production under partial selfing}
+ Outcrossing in our model proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}.
+That is, each individual's genotype determines both the number of ovules produced and the paternal mating success, broadly defined.
+For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986, WangBarrett2020, Harder2016}.
+Note, however, that by modeling paternal relative mating success rather than pollen/sperm production, we implicitly assume female demographic dominance (i.e., production/transport of male gametes does not limit ovule fertilization and hence population growth), a point we return to in the Discussion.
+For the purpose of this article, we also assume that mating is random with respect to stage.
+
+Because we assume that inbreeding depression only affects offspring viability, the allele frequency in the male mating population is obtained by simply summing the vectors of individuals produced through selfing and outcrossing:
+\begin{linenomath*}
+\begin{equation}
+ 	\mbf{p}^{\prime}=\frac{(\mbf{n}^X + \mbf{n}^S)}{\| (\mbf{n}^X + \mbf{n}^S)\|}.
+ \end{equation}
+\end{linenomath*}
+Depending on their stage and genotype at the SA locus, individuals contribute male gametes to the overall population gamete pool according to the following equation
+\begin{linenomath*}
+\begin{equation}
+		\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime}= 
+		\left(
+			\begin{array}{ccc}
+				\mbf{1}^{\intercal}_{\omega} & \frac{1}{2} \mbf{1}^{\intercal}_{\omega} & \mbf{0} \\
+				\mbf{0}  & \frac{1}{2} \mbf{1}^{\intercal}_{\omega} & \mbf{1}^{\intercal}_{\omega} \\
+			\end{array} \right)
+		\left(
+			\begin{array}{ccc}
+				\mbf{F}^{\prime}_{AA} & \mbf{0}  &\mbf{0}  \\
+				\mbf{0}  & \mbf{F}^{\prime}_{Aa} & \mbf{0}  \\
+				\mbf{0}  & \mbf{0}  & \mbf{F}^{\prime}_{aa} \\
+			\end{array} \right)
+		\left(
+			\begin{array}{c}
+				\mbf{p}^{\prime}_{AA} \\
+				\mbf{p}^{\prime}_{Aa} \\
+				\mbf{p}^{\prime}_{aa} \\
+			\end{array} \right),
+\end{equation}
+\end{linenomath*}
+
+\noindent where $\mbf{1}^{\intercal}_{\omega} $ is a vector of ones of size $1 \times \omega$.
+The matrix $\mbb{F}^{\prime}$ is composed of genotype-specific fertility matrices, and operates on the vector of genotype frequencies to give their relative contribution of each genotype to the gamete pool.
+The matrix $\mbf{W}^{\prime}$ converts these contributions to allele numbers.
+Normalizing the resulting vector gives the allele frequencies in the male gamete pool:
+\begin{linenomath*}
+\begin{equation} \label{eq:maleGametePool}
+	\left(
+		\begin{array}{c}
+			q^{\prime}_{A} \\
+			q^{\prime}_{a} \\
+		\end{array} \right) = 
+			\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime}}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime} \|} = 
+				\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)\|}.
+\end{equation}
+\end{linenomath*}
+The key difference between selfing and outcrossing reproduction can be seen in the parent-offspring maps, which reflect the joint effects of meiosis and mating on the distribution of offspring genotypes.
+% \textcolor{red}{
+% I'm doubting whether to move the next paragraph to the appendices as well.
+Equation (\ref{eq:maleGametePool}) is all we strictly need.
+If we keep the parent-offspring maps, we can explain in the next section how the matrices $\mcal{F}^S$ and $\mcal{F}^X$ have the same structure.
+Would that be helpful?
+Or would it be easier to just explain in the next section what the structure of $\mcal{F}^S$ and $\mcal{F}^X$ is?
+}
+
+The parent-offspring map for outcrossing is a function of the allele frequencies in the male gamete pool, and is given by 
+\begin{linenomath*}
+\begin{equation} \label{eq:HX}
+	\mbf{H}^X = 
+			\left(
+			\begin{array}{ccc}
+				q^{\prime}_{A} & \frac{1}{2} q^{\prime}_{A} & 0 \\
+				q^{\prime}_{a} & \frac{1}{2} & q^{\prime}_{A}  \\
+				0 & \frac{1}{2} q^{\prime}_{a} & q^{\prime}_{a} \\
+			\end{array} \right).
+\end{equation}
+\end{linenomath*}
+\noindent From left to right, the columns of matrix $\mbf{H}^X$ give the genotype distribution of outcrossed offspring produced by a maternal parent of each genotype ($AA$, $Aa$, and $aa$ respectively).
+
+The parent-offspring map for reproduction via self-fertilization differs from the outcrossing parent-offspring map.
+As in previous models of partial selfing, we assume that individuals produce enough male gametes to easily self-fertilize all of their ovules, and that self-fertilization involves little or no selection via the male sex function from external factors (e.g., pollinator visitation) relative to outcrossing (\citealt{Charlesworth1978a,JordanConnallon2014,Olito2017}, but see \citealt{Tazzyman2015}).
+Under these assumptions, the genotype distributions of selfed offspring are determined entirely by the parental genotype and the probabilities of segregation and fertilization during and after meiosis:
+\begin{linenomath*}
+\begin{equation} \label{eq:HS}
+	\mbf{H}^S(\tilde{\mbf{n}}) = 
+			\left(
+			\begin{array}{ccc}
+				1 & 1/4 & 0 \\
+				0 & 1/2 & 0 \\
+				0 & 1/4 & 1 \\
+			\end{array} \right).
+\end{equation}
+\end{linenomath*}
+\noindent The columns of $\mbf{H}^S(\tilde{\mbf{n}})$ again give the selfed offspring genotype distributions for parental genotypes of $AA$, $Aa$, and $aa$, respectively.
+As described in the next section, these parent-offspring maps influence the projection matrix by altering the fertility matrices for individuals produced by selfing and outcrossing (see Eq(\ref{eq:BlkFS}) and Eq(\ref{eq:BlkFX}) below).
+
+Following previous theory for partially selfing populations \cite[e.g.,][]{Charlesworth2010,JordanConnallon2014,Glemin2021}, we use a 'fixed prior selfing' model where all individuals are assumed to self-fertilize at a constant rate, $C$, prior to receiving outcross pollen, for reasons of analytic tractability.
+In the Online Supplement, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
+
 We will now derive the expressions for the survival and fertility matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ as given by Eq(10)-(13) in the main text.
@@
 % Figure S1
- \begin{figure}[htbp]
- \centering
- \includegraphics[width=\linewidth]{extinctionThresholdsSexSpecDomFig}
- \caption{\footnotesize{Illustration of parameter space for SA polymorphism and extinction thresholds predicted by the model.
+%  \begin{figure}[htbp]
+%  \centering
+%  \includegraphics[width=\linewidth]{extinctionThresholdsSexSpecDomFig}
+%  \caption{\footnotesize{Illustration of parameter space for SA polymorphism and extinction thresholds predicted by the model.
 Balanced SA polymorphisms can be maintained in the funnel-shaped region between the invasion conditions for each SA allele (dark solid lines).
@@
 Results are shown for three different population selfing rates ($C = \{0,\,1/4,\,1/2\}$), and two cases of complementary dominance ($h_f = 1/4$, $h_m = 3/4$, top row), and ($h_f = 3/4$, $h_m = 1/4$, bottom row), and extinction thresholds are illustrated for the same three values of female fecundity illustrated in fig.~1 of the main text ($f$ values annotated on each panel).}}
- \label{fig:extThresholdsCompSexSpec}
- \end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%
-% Figure S2
- \begin{figure}[htbp]
- \centering
- \includegraphics[scale=0.75]{polymorphicSpaceTitrateSexSpec}
- \caption{\footnotesize{Proportion of demographically viable parameter space (out of total $s_f \times s_m$ space with $\max(s) = 0.15$) in the absence of inbreeding depression (i.e., assuming $\delta = \delta_i = 0$, where $i \in \{j,a,\gamma\}$), plotted as a function of the population selfing rate.
+%  \label{fig:extThresholdsCompSexSpec}
+%  \end{figure}
+
+% %%%%%%%%%%%%%%%%%%%%%
+% % Figure S2
+%  \begin{figure}[htbp]
+%  \centering
+%  \includegraphics[scale=0.75]{polymorphicSpaceTitrateSexSpec}
+%  \caption{\footnotesize{Proportion of demographically viable parameter space (out of total $s_f \times s_m$ space with $\max(s) = 0.15$) in the absence of inbreeding depression (i.e., assuming $\delta = \delta_i = 0$, where $i \in \{j,a,\gamma\}$), plotted as a function of the population selfing rate.
 Results are shown for three fertility values corresponding to low, medium, and high fertility (blue, green, and red points respectively) under two cases of complementary dominance for  SA fitness effects ($h_f = 1/4$, $h_m = 3/4$, panel A; and $h_f = 3/4$, $h_m = 1/4$, panel B).
 Each point was calculated by numerical integration of the corresponding SA invasion conditions and extinction threshold predicted by the mendelian matrix model (see Analyses section in the main text), while solid lines were produced by numerically integrating the analytic expressions for the single-locus invasion conditions from the population genetic models of \citet{JordanConnallon2014} and \citet{Olito2017} (solid black lines).}}
- \label{fig:polySpaceCompSexSPec}
- \end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%
-% Figure 3
- \begin{figure}[htbp]
- \centering
- \includegraphics[scale=0.75]{deltaPolymorphicSpaceTitrateCompSexSpec}
- \caption{\footnotesize{Effects of early- and late-acting inbreeding depression on the proportion of demographically viable parameter space (out of total $s_f \times s_m$ space with $\max(s) = 0.15$), plotted as a function of the population selfing rate.
+%  \label{fig:polySpaceCompSexSPec}
+%  \end{figure}
+
+% %%%%%%%%%%%%%%%%%%%%%
+% % Figure 3
+%  \begin{figure}[htbp]
+%  \centering
+%  \includegraphics[scale=0.75]{deltaPolymorphicSpaceTitrateCompSexSpec}
+%  \caption{\footnotesize{Effects of early- and late-acting inbreeding depression on the proportion of demographically viable parameter space (out of total $s_f \times s_m$ space with $\max(s) = 0.15$), plotted as a function of the population selfing rate.
 In all plots, the strength of inbreeding depression decreases as the selfing rate goes up following a simple model of purging recessive deleterious mutations.
@@
 Each point was calculated by numerically integrating the corresponding SA invasion conditions and extinction threshold predicted by the model (see Analyses section in the main text), while solid lines were produced by numerically integrating the single-locus invasion conditions from the population genetic models of \citet{JordanConnallon2014} and \citet{Olito2017} (solid black lines).}}
- \label{fig:deltaPolySpaceCompSexSpec}
- \end{figure}
+%  \label{fig:deltaPolySpaceCompSexSpec}
+%  \end{figure}
 

### 2021-12-10 doc/ResponseLetter-R1.tex
Supplementaries
Edited supplementaries and added a few more answers to the response letter

@@
 
-\begin{quote}
-	{\itshape ... our response}
-\end{quote}
-
 \begin{itemize}
@@
 Getting rid of H (or to the contrary, using it all the way, thus simplifying eq 12 and 13 for instance) and not giving a name to matrix W (which is just a bunch of 0s 1s and 1/2s) are steps that can make it easier for the reader to follow.
+\begin{quote}
+	{\itshape Response to the above 5 points: As per the reviewers (and the editors) suggestion, we have moved a lot of the derivations into the supplementary materials, and now no longer use H or W in the main text.
+In doing so, we hope to shift the focus from mathematical derivations, which are probably unhelpful to many readers, towards explaining the end result instead.
+}
+\end{quote}
+
 	\item It is quite confusing that HX is written as a constant (eq 7) whilst it depends on the frequencies and that HS is written as dependent (eq6) whilst it is constant
@@
 \begin{quote}
-	{\itshape ... our response}
+	{\itshape As suggested by this reviewer and by the editor, we have now moved up the two stage example, so that we show the reader what these matrices can look like immediately after introducing them.
+Again shifting the focus from mathematical derivations to an explicit example, thereby hopefully improving the manuscripts' accessibility/readability.
+}
 \end{quote}
@@
 \begin{quote}
-	{\itshape ... our response}
+	{\itshape ... our response \textcolor{red}{I cannot figure out where in the manuscript this is, I don't think I have the exact R0 version as it was submitted?}}
 \end{quote}

### 2021-12-10 doc/Supplements/SupplementaryMaterial.tex
Supplementaries
Edited supplementaries and added a few more answers to the response letter

@@
 \section{Construction of the population projection matrix}
+
+In this section we will go through the derivation of the matrix model incorporating multiple life-cycle stages and a single diallelic locus under SA selection on male and female fertility for populations of partially-selfing simultaneous hermaphrodites model presented in the main text.
+We start by calculating the frequencies of alleles in the gamete pool.
+These frequencies are the ingredients for the map of parent genotype to offspring genotype, which itself is a necessary ingredient to create the fertility matrices.
+Finally putting the fertility components together with the (slightly simpler) survival matrices, we obtain all the ingredients to build the population projection matrix.
+
 \subsubsection*{Mating and offspring production under partial selfing}
@@
 \noindent The columns of $\mbf{H}^S(\tilde{\mbf{n}})$ again give the selfed offspring genotype distributions for parental genotypes of $AA$, $Aa$, and $aa$, respectively.
-As described in the next section, these parent-offspring maps influence the projection matrix by altering the fertility matrices for individuals produced by selfing and outcrossing (see Eq(\ref{eq:BlkFS}) and Eq(\ref{eq:BlkFX}) below).
+As described in the next section, these parent-offspring maps influence the projection matrix by altering the fertility matrices for individuals produced by selfing and outcrossing (see Eq(\ref{eq:BlkFS_appendix}) and Eq(\ref{eq:BlkFX_appendix}) below).
 
 Following previous theory for partially selfing populations \cite[e.g.,][]{Charlesworth2010,JordanConnallon2014,Glemin2021}, we use a 'fixed prior selfing' model where all individuals are assumed to self-fertilize at a constant rate, $C$, prior to receiving outcross pollen, for reasons of analytic tractability.
-In the Online Supplement, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
-
-We will now derive the expressions for the survival and fertility matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ as given by Eq(10)-(13) in the main text.
+Further down in the Online Supplement, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
+
+\subsubsection*{The Population Projection Matrix}
+We will now derive the expressions for the survival and fertility matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ as given by Eq(12)-(15) in the main text.
 The matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ are the ingredients of the population projection matrix, $\tilde{\mbf{A}}$, as given in Eq(9), repeated here for convenience,

### 2021-12-14 doc/ResponseLetter-R1.tex
Added new figure 1 and few more responses to reviewers

@@
 \begin{quote}
-	{\itshape ...}
+	{\itshape We agree that less focus on the (derivation of) general model and more focus on the specific two stage example will hopefully make the paper more accessible for non-theoretical readers, and we have  restructured the methods accordingly.
+}
 \end{quote}
@@
 \begin{quote}
-	{\itshape ...}
+	{\itshape We agree that in our efforts to appeal to both demographers and evolutionary biologists studying sexual conflict, we ended up insufficiently discussing the literature on both accounts.
+We have attempted to improve this, see also our response to reviewer 1.
+}
 \end{quote}
@@
 Rather, by making compromises in the introduction we ended up lacking scholarship on both sides.
-To address your main concern, we have restructured our Introduction to more directly connect with the broader literature on evolutionarily explicit structured models (see {\bf L.XXX--XXX}) and now try to link our overview of sexually antagonistic selection theory with the existing empirical literature (see {\bf L.XXX--XXX}).
-We hope these changes succeed in broadening the appeal of our manuscript to the wider readership of AmNat.}
+To address your main concern, we have restructured our Introduction to more directly connect with the broader literature on evolutionarily explicit structured models (see {\bf L.61-73}) and now try to link our overview of sexually antagonistic selection theory with the existing empirical literature (see {\bf L.XXX--XXX}).
+We hope these changes succeed in broadening the appeal of our manuscript to the wider readership of AmNat.
+If there are still important references missing, we are of course open to specific suggestions.
+}
 \end{quote}
@@
 \begin{quote}
-	{\itshape ... our response \textcolor{red}{I cannot figure out where in the manuscript this is, I don't think I have the exact R0 version as it was submitted?}}
+	{\itshape We have added a few lines to discuss frequency vs density dependence when we first introduce the population projection matrix in equation 3 (CHECK LINE NUMBERS AFTER WE"RE DONE EDITING ).
+In addition, we have tried to clarify to improve this section by discussing that the survival and fertility rates ("demographic rates") are density-indepedent, and that the allele frequency dynamics make the model frequency dependent.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape ... our response}
+	{\itshape We want to thank the reviewer for this suggestion.
+We now start with a figure that only has one f, and that separately explains what happens to the allele frequencies (Population genetic outcome) and to the population growth rate (Demographic Outcome), see new Figure 1.
+We hope that this will make it easier to interpret the old figure 1, which is now Figure 2.}
 \end{quote}

### 2021-12-14 doc/SA-Hermaphrodites-wDemography-R1.tex
Supplementaries
Edited supplementaries and added a few more answers to the response letter
Added new figure 1 and few more responses to reviewers

@@
 In reality, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals (e.g., \citealt {johnston2013life, merot2020balancing}).
+
 To take into account the consequences of population dynamics under SA selection, a model is required that links changes in genotype distributions to population dynamics.
@@
 The derivation of the general model for any number of age or stage classes is in the Online Supplementary Material.} The derivation follows closely the logic presented in \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b}.
-In fact, the model reduces to the stage-structured model of \citet{deVriesCaswell2019a} under obligate outcrossing.
+In fact, the model reduces to the stage-structured model of \citet{deVriesCaswell2019a} under obligate outcrossing, \textcolor{blue}{that is, when selfing is not allowed}.
 All computer code necessary to reproduce the results are available at \url{https://anonymous.4open.science/r/SA-Hermaphrodites-wDemography-BF66}.%\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
@@
 \end{linenomath*}
+If the survival or fertility rates are a function of the population density, $\tilde{\mbf{n}}(t)$, then the model will be a density dependent model and the population projection matrix is a function of the full population vector, $\tilde{\mbf{A}}[\tilde{\mbf{n}}(t)]$.
+In the online supplementary materials we show results for a density-dependent version of the model.
+In the main text of this paper, we will consider models with density-independent survival and fertility rates.
+The population projection matrix is then only a function of the proportional population vector, $\tilde{\mbf{A}}[\tilde{\mbf{p}}(t)]$, and the model is referred to as a frequency-dependent model.
 
@@
 
-Diverse eco-evolutionary outcomes are possible in the model, including fixation of either allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and suicide (e.g., see \citealt{deVriesCaswell2019a,deVriesCaswell2019b}).
-We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection and ({\itshape ii}) the intrinsic population growth rate at equilibrium is positive; a situation that we refer to as a 'demographically viable SA polymorphism'.
+Diverse eco-evolutionary outcomes are possible in the model, including fixation of either allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and suicide (e.g., see \citealt{deVriesCaswell2019a,deVriesCaswell2019b}), see Figure 1 for an example of the possible outcomes in different regions of parameter space.
+We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection (see left-hand side of Figure \ref{fig:didactic}) and ({\itshape ii}) the intrinsic population growth rate at equilibrium is positive (see right-hand side of Figure \ref{fig:didactic}); a situation that we refer to as a 'demographically viable SA polymorphism'.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%
+% Figure 1
+ \begin{figure}[htbp]
+ \centering
+ \includegraphics[width=\linewidth]{didacticFig}
+ \caption{\footnotesize{Illustration of the possible population genetic outcomes of the model on the left (fixation of either allele, or a polymorphism), and of our demographic outcome of interest on the right:  population growth rates.}}
+ \label{fig:didactic}
+ \end{figure}
+%%%%%%%%%%%%%%%%%%%%%
+
 We used numerical simulation to determine whether a protected SA polymorphism was also demographically viable.
 Specifically, for each boundary equilibrium we introduced the rare allele at low initial frequency, and iterated Eq(\ref{eq:Atilde}) until the population had reached demographic and genotypic equilibrium.
-Our model implements density independent demographic rates, and so the population state vector will grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see chapter 17 in \citealt{Caswell2001}).
+Although we use density-independent survival and fertility rates, the allele frequency dynamics introduce frequency-dependence into the model.
+As a consequence, the population state vector will grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see chapter 17 in \citealt{Caswell2001}).
 The intrinsic population growth rate after convergence, $\lambda$, can be calculated as $\tilde{\mbf{n}}(t)/\tilde{\mbf{n}}(t-1)$.
@@
 %%%%%%%%%%%%%%%%%%%%%
-% Figure 1
+% Figure 2
  \begin{figure}[htbp]

### 2021-12-15 doc/SA-Hermaphrodites-wDemography-R1.tex
blueified yesterdays changes
Forgot to put my changes in blue yesterday, sorry, fixed it now

@@
 \end{linenomath*}
-If the survival or fertility rates are a function of the population density, $\tilde{\mbf{n}}(t)$, then the model will be a density dependent model and the population projection matrix is a function of the full population vector, $\tilde{\mbf{A}}[\tilde{\mbf{n}}(t)]$.
+\textcolor{blue}{If the survival or fertility rates are a function of the population density, $\tilde{\mbf{n}}(t)$, then the model will be a density dependent model and the population projection matrix is a function of the full population vector, $\tilde{\mbf{A}}[\tilde{\mbf{n}}(t)]$.
 In the online supplementary materials we show results for a density-dependent version of the model.
 In the main text of this paper, we will consider models with density-independent survival and fertility rates.
-The population projection matrix is then only a function of the proportional population vector, $\tilde{\mbf{A}}[\tilde{\mbf{p}}(t)]$, and the model is referred to as a frequency-dependent model.
+The population projection matrix is then only a function of the proportional population vector, $\tilde{\mbf{A}}[\tilde{\mbf{p}}(t)]$, and the model is referred to as a frequency-dependent model.}
 
@@
 Diverse eco-evolutionary outcomes are possible in the model, including fixation of either allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and suicide (e.g., see \citealt{deVriesCaswell2019a,deVriesCaswell2019b}), see Figure 1 for an example of the possible outcomes in different regions of parameter space.
-We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection (see left-hand side of Figure \ref{fig:didactic}) and ({\itshape ii}) the intrinsic population growth rate at equilibrium is positive (see right-hand side of Figure \ref{fig:didactic}); a situation that we refer to as a 'demographically viable SA polymorphism'.
+We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection \textcolor{blue}{(see left-hand side of Figure \ref{fig:didactic})} and ({\itshape ii}) the intrinsic population growth rate at equilibrium is positive \textcolor{blue}{(see right-hand side of Figure \ref{fig:didactic})}; a situation that we refer to as a 'demographically viable SA polymorphism'.
 
@@
 Specifically, for each boundary equilibrium we introduced the rare allele at low initial frequency, and iterated Eq(\ref{eq:Atilde}) until the population had reached demographic and genotypic equilibrium.
-Although we use density-independent survival and fertility rates, the allele frequency dynamics introduce frequency-dependence into the model.
-As a consequence, the population state vector will grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see chapter 17 in \citealt{Caswell2001}).
+\textcolor{blue}{Although we use density-independent survival and fertility rates, the allele frequency dynamics introduce frequency-dependence into the model.
+As a consequence, }the population state vector will grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see chapter 17 in \citealt{Caswell2001}).
 The intrinsic population growth rate after convergence, $\lambda$, can be calculated as $\tilde{\mbf{n}}(t)/\tilde{\mbf{n}}(t-1)$.

### 2021-12-16 doc/ResponseLetter-R1.tex
Edits to introduction, SA
A first pass at editing the introduction to include more specific
references to empirical SA selection literature. Also some edits to
the other passages.

@@
 We now start with a figure that only has one f, and that separately explains what happens to the allele frequencies (Population genetic outcome) and to the population growth rate (Demographic Outcome), see new Figure 1.
-We hope that this will make it easier to interpret the old figure 1, which is now Figure 2.}
+We hope that this will make it easier to interpret the old figure 1, which is now Figure 2.
+We have also included a new supplementary figure (CHECK FIG NUMBERING IN SUPPLEMENTS) with detailed heatmaps for each of the parameter combinations presented in the this figure (i.e., one f per panel).
+However, we felt that this revised figure was to large and complicated to present in the main text (18 vs.
+6 panels).}
 \end{quote}

### 2021-12-16 doc/SA-Hermaphrodites-wDemography-R1.tex
Edits to introduction, SA
A first pass at editing the introduction to include more specific
references to empirical SA selection literature. Also some edits to
the other passages.

@@
 \usepackage{pdflscape}
+\usepackage{textgreek}
 
@@
 
-Fitness trade-offs that impose genetic constraints on adaptation have particularly interesting evolutionary consequences.
-On the one hand, they prevent individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase a population's overall extinction risk \citep{kokko2003sexy,harts2014demography}.
-On the other hand, they provide an effective mechanism for the maintenance of genetic variation, which can increase a population's capacity for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
+\textcolor{blue}{Genetic constraints on adaptation arising from conflicting selection} have particularly interesting evolutionary consequences.
+On the one hand, they provide an effective mechanism for the maintenance of genetic variation, which can increase a population's capacity for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
+On the other hand, they prevent individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase a population's overall extinction risk \citep{kokko2003sexy,harts2014demography}.
 Hence, for traits under conflicting selection, the nature and extent of genetic variation observed in natural populations should reflect a balance between the maintainance of genetic polymorphisms, and the population dynamical consequences of the resulting maladaptation.
 
-Sexually antagonistic selection (abbreviated SA hereafter) arising from genetic trade-offs between male and female fitness, is a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and the evolution of sexual dimorphism and sex-related traits \citep{Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, BondurianskyChenoweth2009,Olito2019}.
-When such genetic trade-offs occur, SA selection can arise when beneficial alleles for one sex are deleterious when expressed in the other \citep{Kidwell1977, Rice1992, ConnallonClark2012}.
-
-Alleles with opposing fitness effects through male and female sex functions can cause analogous genetic constraints on fitness in hermaphrodite populations, where both maternal and paternal reproductive success contribute jointly to each individuals' overall fitness \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
-Self-fertilization in hermaphrodites is predicted to reduce the total parameter space where balanced SA polymorphisms can be maintained, while simultaneously creating a bias in selection through the female sex function (\citealt{JordanConnallon2014,Glemin2021}; but see \citealt{Tazzyman2015}).
-Yet, other factors, such as genetic linkage to other SA loci or a sex-determining region \citep{Otto2011, JordanCharlesworth2012, Olito2017, Olito2019}, antagonism between viability and fecundity selection \citep{Glemin2021}, or spatial heterogeneity and complexity of the life-cycle can expand the parameter space for SA polymorphism \citep{Olito-etal-2018,ConnallonSharmaOlito2019, Glemin2021}.
-Overall, both theoretical predictions and current empirical data suggest that there is ample scope for SA trade-offs and the maintenance of SA polymorphisms in both dioecious and hermaphrodite populations \citep{Abbott2011, WangBarrett2020}, although identifying specific SA loci from genome sequence data remains challenging \citep{RuzickaESEB2020}.
-
-Most population genetic models, including models of the maintenance of SA polymorphisms, keep track of the relative fitnesses of genotypes in populations of constant (and often infinite) size, and therefore do not consider the population dynamical consequences of evolutionary change.
-In reality, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals (e.g., \citealt {johnston2013life, merot2020balancing}).
-
-To take into account the consequences of population dynamics under SA selection, a model is required that links changes in genotype distributions to population dynamics.
-\textcolor{blue}{One of the earliest such models \citep{roughgarden1971density} combined a population genetic model with a density-dependent population model with no intraspecific demographic structure.
-Charlesworth then combined an age-structured population model with population genetics in a series of papers, consolidated in \cite{BCharlesworth1991}.
+Sexually antagonistic selection (abbreviated SA hereafter), \textcolor{blue}{is an important form of conflicting selection that arises when beneficial alleles for one sex are deleterious when expressed in the other.
+SA selection appears to be} a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and the evolution of sexual dimorphism and sex-related traits \citep{Kidwell1977, Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, ConnallonClark2012, BondurianskyChenoweth2009, Olito2019}.
+\textcolor{blue}{Sex-specific estimates of phenotypic selection indicate that sex-differences in directional selection are common \citep[e.g.,][]{DeLisle-etal-2018,Lewis-etal-2011,CoxCalsbeek2009,SinghPunzalan2018}, implying that many genetic variants underlying quantitative traits have SA fitness effects.
+Similarly, quantitative genetic studies frequently find negative cross-sex genetic correlations for fitness, indicating SA fitness effects for multilocus genotypes (\citealt{Chippendale-etal-2001,Delph-etal-2011}; reviewed in \citealt{ConnallonMatthews2019}).
+Moreover, polygenic signals of SA selection are evident in recent population genetic analyses of large genomic datasets \citep{RuzickaESEB2020, RuzickaConnallon2022}.}
+
+Alleles with opposing fitness effects through male and female sex functions can also cause SA selection in hermaphrodite populations, where individuals' overall fitness is determined jointy by maternal and paternal reproductive success \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
+\textcolor{blue}{In obligately outcrossing hermaphrodite populations, SA selection operates similarly as in seperate-sexed ones (dioecious species).
+However, the possibility of self-fertilization has important consequences for the maintenance of SA genetic variation in hermaphrodites}.
+Self-fertilization reduces \textcolor{blue}{opportunities for} balanced SA polymorphisms, while simultaneously creating a bias in selection through the female sex function (\citealt{JordanConnallon2014,Glemin2021}; but see \citealt{Tazzyman2015}).
+Yet, other factors, such as genetic linkage to other SA loci or a sex-determining region \citep{Otto2011, JordanCharlesworth2012, Olito2017, Olito2019}, antagonism between viability and fecundity selection \citep{Glemin2021}, or spatial heterogeneity and complexity of the life-cycle \textcolor{blue}{can all facilitate the maintenance of} SA polymorphisms \citep{Olito-etal-2018,ConnallonSharmaOlito2019, Glemin2021}.
+Overall, both theoretical predictions and current empirical data suggest that there is ample scope for SA trade-offs and the maintenance of SA polymorphisms in both dioecious and hermaphrodite populations \citep{Abbott2011, DelphHerlihy2012, WangBarrett2020, RuzickaESEB2020, RuzickaConnallon2022}.
+
+\textcolor{blue}{While significant efforts have been made to predict when SA polymophisms can be maintained and to detect SA genetic variation from empirical data, much less consideration has been given to the demographic consequences of the maladaptation caused by SA selection}.
+Most population genetic models, including models of the maintenance of SA polymorphisms, keep track of the relative fitnesses of genotypes in populations of constant (and often infinite) size, and do not consider the population dynamical consequences of evolutionary change.
+In reality, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals \textcolor{blue}{which link genotype distributions to population dynamics (e.g., \citealt {johnston2013life, merot2020balancing}).
+Over the past 50 years, a rich field of eco-evolutionary dynamics has developed which seeks to jointly model the genetic and demographic processes involved in evolutionary change.
+In particular, theoretical models combining population genetics with density-dependent population growth \citep{roughgarden1971density} and stage-structured populations (consolidated in \citealt{Charlesworth_1994}) paved the way toward understanding the demographic changes accompanying allele frequency changes due to selection.
+Extensions of these theories to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and quantitative traits \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects} all highlight the often counterintuitive demographic consequences of adaptive evolution that are orten missed by classical population genetics models (comprehensively reviewed by \citealt{fussmann2007eco} and \citealt{govaert2019eco}).}
+
+%\textcolor{blue}{One of the earliest such models \citep{roughgarden1971density} combined a population genetic model with a density-dependent population model with no intraspecific demographic structure.
+Charlesworth then combined an age-structured population model with population genetics in a series of papers, consolidated in \cite{Charlesworth_1994}.
 \cite{orive1995senescence} extended this work to study life-history evolution in organisms with both clonal and sexual reproduction (see also \cite{orive2001somatic}).
@@
 
-The potential demographic costs of sexual antagonism were pointed out by \cite{kokko2003sexy}, but few papers since then have explicitly incorporated demography into models of sexual antagonism (with the notable exception of \citealt{harts2014demography}).
-\citet{deVriesCaswell2019a} introduced a mendelian matrix model with intralocus sexual antagonism, and population dynamics but did not perform an analysis of the consequences of demographic viability for the scope of sexual antagonism to maintain polymorphism.
-
-In this paper, we extend the model of \citep{deVriesCaswell2019a,deVriesCaswell2019b} to include both dioecious and hermaphroditic species in order to study the parameter conditions under which different outcomes of SA selection, including fixation of male- or female-beneficial alleles as well as balanced polymorphisms, are also demographically viable.
+The potential demographic costs of sexual antagonism were pointed out by \cite{kokko2003sexy}, but few studies since have explicitly incorporated demography into models of sexual antagonism (with the notable exception of \citealt{harts2014demography}).
+\citet{deVriesCaswell2019a} introduced a mendelian matrix model with intralocus sexual antagonism and population dynamics, but did not analyze the consequences of demographic viability for the scope of sexual antagonism to maintain polymorphism.
+In this paper, we extend the model of \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b} to include both dioecious and hermaphroditic \textcolor{blue}{populations} in order to study the parameter conditions under which different outcomes of SA selection, including fixation of male- or female-beneficial alleles as well as balanced polymorphisms, are also demographically viable.
 A major strength of our approach is that the model can be parameterized using empirically estimated demographic rates, enabling us to make predictions about the scope for demographically viable SA polymorphisms that are grounded in the biology of real populations.
-We demonstrate this with a case study of {\itshape Mimulus guttatus} (now {\itshape Erythranthe guttata})
+We demonstrate this with a case study of {\itshape Mimulus guttatus} (now {\itshape Erythranthe guttata}).
 
@@
  \includegraphics[width=\linewidth]{didacticFig}
- \caption{\footnotesize{Illustration of the possible population genetic outcomes of the model on the left (fixation of either allele, or a polymorphism), and of our demographic outcome of interest on the right:  population growth rates.}}
+ \caption{\footnotesize{Illustration of the possible population genetic outcomes of the model on the left (fixation of either allele, or a polymorphism), and of our demographic outcome of interest on the right:  population growth rates.
+For clarity, results are shown for the case of a dominance reversal, where $h_f = h_m = 1/4$.}}
  \label{fig:didactic}
@@
 
-% Figure 4
+% Figure 5
 \begin{figure}[htbp]
  \centering
- \includegraphics[scale=0.75]{../output/figs/mimulusinv6FigNewSel.pdf}
+ \includegraphics[scale=0.75]{../output/figs/MimulusInv6LambdaFig.pdf}
  \caption{\footnotesize{Illustration of model predictions using empirically estimated demographic rates for {\itshape M.~guttatus}.
@@
 Despite the demographic pitfalls associated with SA alleles, our example using {\itshape M.~guttatus} appears to show that demographic rates observed in some real populations are capable of sustaining large regions of viable SA polymorphic space.
-The example also appears to provide some empirical support for the conjecure that locally-adapted populations are more likely to harbor SA polymorphisms than marginal or non-locally adapted ones; inviable polymorphic parameter space only occured when using the demographic data for non-local high-elevation perennial populations.
+The example also appears to provide some empirical support for the conjecure that locally-adapted populations are more likely to harbor SA polymorphisms than marginal or non-locally adapted ones; inviable polymorphic parameter space only occured when using the demographic data for non-local low-elevation perennial populations.
 Although we cannot make concrete predictions for inv6 in the Iron Mountain population in which it was observed, it is interesting that the estimated SA fitness effects place this polymophic inversion squarely in demographically viable polymorphic parameter space predicted by our model.
@@
 By combining the tools of demography and population genetics, the framework we present here enables the exploration of interactions between life cycle complexity, mating system, and sexual antagonism.
-For the sake of simplicity, we made a variety of simplifying assumptions in our analyses, and a variety of extensions are possible.
+For the sake of simplicity, we made a variety of simplifying assumptions in our analyses, and \textcolor{blue}{many extensions of the model are} possible.
 For example, we used a simple life cycle with just two stages for most of this paper, adults and juveniles (the \textit{Mimulus} example has $3$ stages).

### 2021-12-17 doc/SA-Hermaphrodites-wDemography-R1.tex
Update SA-Hermaphrodites-wDemography-R1.tex

@@
 \textcolor{blue}{If the survival or fertility rates are a function of the population density, $\tilde{\mbf{n}}(t)$, then the model will be a density dependent model and the population projection matrix is a function of the full population vector, $\tilde{\mbf{A}}[\tilde{\mbf{n}}(t)]$.
-In the online supplementary materials we show results for a density-dependent version of the model.
+In appendix B we construct and analyze a density-dependent version of the model.
 In the main text of this paper, we will consider models with density-independent survival and fertility rates.
@@
 
+\textcolor{blue}{Of course in natural populations processes like resource competition or competition for breeding sites will often prevent populations from exponentially growing (or shrinking).
+That is,  natural populations experience population regulation, or density-dependence.
+In appendix B we extend the model presented in the main paper by including density-dependence that acts on the survival of all life history stages equally.
+We reanalyse the results for \textit{Mimulus guttatus} using this density-dependent version of the model in Figure \ref{fig:mimulusFig_dd}, whereby the heatmap in Figure \ref{fig:mimulusFig_dd} now shows equilibrium population densities rather than population growth rates that are shown in Figure \ref{fig:mimulusFig}.
+The shape of the coexistence region is unchanged by the density-dependence, and the shape of the population density heatmap in Figure \ref{fig:mimulusFig_dd} is qualitatively the same as the shape of the population growth rate heatmap in Figure \ref{fig:mimulusFig} because density-dependence that acts age-independently does not change the selection gradients, see also discussions in cite\textbf{mylius1995evolutionarily, caswell20174}.
+}
+
+\textcolor{blue}{In many natural populations density-dependence does not have the exact same effect on survival of individuals of all ages or developmental stages.
+For example, competition for breeding sites will lead to a reduction in fecundity rather than survival.
+In addition, experienced females might be better at securing a breeding site than newcomers.
+Similarly, very young or very old individuals might be worse competitors for resources than adults, so resource competition might affect juvenile and old-age survival more severely.
+This type of age or stage specific density-dependence will change the selection gradients, and therefore the shape of the coexistence boundaries and the extinction boundaries in Figure \ref{fig:mimulusFig}, and warrants further research.}
+
+% @article{mylius1995evolutionarily,
+%   title={On evolutionarily stable life histories, optimization and the need to be specific about density dependence},
+%   author={Mylius, Sido D and Diekmann, Odo},
+%   journal={Oikos},
+%   pages={218--224},
+%   year={1995},
+%   publisher={JSTOR}
+% }
+
+% @article{caswell20174,
+%   title={4 Senescence, Selection Gradients and Mortality},
+%   author={Caswell, Hal and Shyu, Esther},
+%   journal={The Evolution of Senescence in the Tree of Life},
+%   pages={56},
+%   year={2017},
+%   publisher={Cambridge University Press}
+% }
+
 \subsubsection*{Extensions and future directions}
@@
 What if the allele not only affects male and female fertility but also affects survival of individuals?
-Additionally, how would the scope for SA polymorphism be affected by density-dependence acting on different life history stages?
 
@@
 
-\cite{grieshop2017male} found that genotypes with sexually antagonstic alleles that are male beneficial experience higher levels of inbreeding depression than genotypes with female beneficial SA alleles.
+\cite{grieshop2017male} found that genotypes with sexually antagonistic alleles that are male beneficial experience higher levels of inbreeding depression than genotypes with female beneficial SA alleles.
 This effect can easily be included in our model by making the inbreeding depression parameters, $\delta$ and $\delta_i$, a function of the genotype of the individual.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Appendix B: Eigenvalues for invasion analysis}
+\section*{Appendix B: Example with density-dependence}
 \renewcommand{\theequation}{B\arabic{equation}}
+\renewcommand{\thetable}{B\arabic{table}}
+\setcounter{equation}{0}  % reset counter 
+\setcounter{table}{0}  % reset counter
+
+\textcolor{blue}{Here we extend the model in the main text by introducing population regulation into the model.
+We chose the simplest population regulation possible in a structured population model: population regulation affecting all life history stages in the same way.
+Note that this is the only type of population regulation that is possible in unstructured population models, like the classical Lotka-Volterra model, or the Rosenzweig-MacArthur model.
+We model survival of all life-history stages as decreasing exponentially with increasing population density, such that the population projection becomes
+\begin{linenomath*}
+\begin{equation}
+	\tilde{\mbf{n}}(t + 1) = \exp\left(-\alpha \sum_i n_i \right) \tilde{\mbf{A}}[\tilde{\mbf{n}}(t)] \, \tilde{\mbf{n}}(t),
+\end{equation}
+\end{linenomath*}
+where $\sum_i n_i$ simply sums over all entries of the vector, so it sums over all genotypes and all life-history stages. Equivalently, in matrix algebra notation,
+\begin{linenomath*}
+\begin{equation}
+	\tilde{\mbf{n}}(t + 1) = \exp\left(-\alpha \| \tilde{\mbf{n}}(t) \|\right) \tilde{\mbf{A}}[\tilde{\mbf{n}}(t)] \, \tilde{\mbf{n}}(t),
+\end{equation}
+\end{linenomath*}
+where $\| \tilde{\mbf{n}}(t) \|$ is the one-norm which is identical to the sum of the entries of $\tilde{\mbf{n}}(t)$ because the population vector only has nonnegative entries. }
+
+\textcolor{blue}{We reanalyzed the {\itshape Mimulus guttatus} example from the main text using this density-dependent version of the model.
+Figure \ref{fig:mimulusFig_dd} has the same structure as Figure \ref{fig:mimulusFig}  except that now the heatmap indicates the equilibrium population size, $\sum_i n_i $, rather than the population growth rate.
+}
+
+\textcolor{blue}{Note that with this simplest kind of density-dependence, very little changes about the results.
+By introducing a density-dependent factor that acts on survival of all life-history stages in the same way, we did not change any of the selection gradients.
+That is, because the density-dependent factor essentially scales the entire population projection matrix, it does not change the stable stage distribution vector, or the normalized reproductive value vector.
+As a consequence, adding this density-dependence did not change the relative importance of any of the life history stages, or any of the demographic rates in the population projection matrix.
+}
+
+\textcolor{blue}{Of course, more realistic and more interesting density-dependent scenarios that will change the results in Figure \ref{fig:mimulusFig_dd} significantly are possible.
+For example, density-dependence could instead mostly impact recruitment of juveniles into the population, or density-dependence could affect only survival of older individuals.
+In this case we would expect significant qualitative as well as quantitative changes to the resuls in Figures \ref{fig:mimulusFig} and \ref{fig:mimulusFig_dd}.
+}
+
+\begin{figure}[htbp]
+ \centering
+ \includegraphics[scale=0.8]{../output/figs/MimulusInv6NstarFig.pdf}
+ \caption{\footnotesize{Illustration of density-dependent model predictions, using empirically estimated demographic rates for {\itshape M.~guttatus} as in the main text.
+Results are shown for two hypothetical populations using demographic rates for locally adapted (Eagle Meadows; panel A), and non-local (Low-Elevation Perennial; panel B) populations reported by \citet{PetersonEtAl2016}.
+Invasion conditions and equilibrium population densities were calculated using the conservative assumption of additive fitness effects in both sexes ($h_f = h_m = 0.5$) for two parameter conditions: obligate outcrossing (solid black lines, dark red shaded region)  and partial selfing with inbreeding depression using the highest field-estimate of selfing, and inbreeding depression parameters ($\delta_i$) calculated for the Iron Mountain population of \citet{Willis1993} (dashed lines, light red shaded region).
+The location of inv6 is also shown on both plots, using selection coefficients calculated from field estimates of male and female fitness components from \citet{LeeKelly2015} under the relatively conservative assumption of additive fitness effects in both sexes, where $(s_f,\,s_m) = (0.30,\, 0.31)$.
+Note that in panel A the invasion conditions for partial selfing with inbreeding depression (dotted lines) are nearly indistinguishable from those for obligate outcrossing (solid lines).}}
+ \label{fig:mimulusFig_dd}
+ \end{figure}
+
+\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%
+% Appendix C
+%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Appendix C: Eigenvalues for invasion analysis}
+\renewcommand{\theequation}{C\arabic{equation}}
 \setcounter{equation}{0}  % reset counter 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-% Appendix C
-%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Appendix C: Demographic and empirical data for {\itshape M.~guttatus} case study}
-\renewcommand{\theequation}{C\arabic{equation}}
-\renewcommand{\thetable}{C\arabic{table}}
+% Appendix D
+%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Appendix D: Demographic and empirical data for {\itshape M.~guttatus} case study}
+\renewcommand{\theequation}{D\arabic{equation}}
+\renewcommand{\thetable}{D\arabic{table}}
 \setcounter{equation}{0}  % reset counter 

### 2021-12-17 doc/SA-Hermaphrodites-wDemography-R1.tex
Minor additional edits

--- 2021-12-16 doc/SA-Hermaphrodites-wDemography-R1.tex
@@
 In particular, theoretical models combining population genetics with density-dependent population growth \citep{roughgarden1971density} and stage-structured populations (consolidated in \citealt{Charlesworth_1994}) paved the way toward understanding the demographic changes accompanying allele frequency changes due to selection.
-Extensions of these theories to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and quantitative traits \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects} all highlight the often counterintuitive demographic consequences of adaptive evolution that are orten missed by classical population genetics models (comprehensively reviewed by \citealt{fussmann2007eco} and \citealt{govaert2019eco}).}
+Extensions of these theories to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and quantitative traits \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects} all highlight the often counterintuitive demographic consequences of adaptive evolution that are often missed by classical population genetics models (comprehensively reviewed by \citealt{fussmann2007eco} and \citealt{govaert2019eco}).}
 

### 2021-12-20 doc/ResponseLetter-R1.tex
Full edits first pass
This commit includes a first pass at all major edits/revisions to the
manuscript. Ready to hand it over.

--- 2021-12-10 doc/ResponseLetter-R1.tex
@@
 \section*{}
-\noindent To the AE, Ed., \& Editorial Board
+\noindent To the Ed., AE, \& Editorial Board
 \bigskip
@@
 
-\noindent First of all, we want to thank the AE, Ed., and EiC for their thoughtful and forthright responses (and apologies) regarding the editorial delay on our manuscript.
+\noindent First of all, we want to thank the Ed., AE, and EiC for their thoughtful and forthright responses (and apologies) regarding the editorial delay on our manuscript.
 We understand completely that pandemic-related personal issues have been and continue to be extremely disruptive in these difficult times, and that editorial duties for our journals is a volunteer service.
-To our AE, Dr.~Orive, we sincerely hope that you and your family are doing ok, and we truly appreciate the time and effort you took to evaluate our paper while you clearly had more imporant issues to deal with.
+To our AE, Dr.~Orive, we sincerely hope that you and your family are well, and we truly appreciate the time and effort you took to evaluate our paper when you clearly had more imporant issues to deal with.
 \bigskip
 
-\noindent We received very helpful and constructive comments on our first submission from Associate Editor Dr.~Maria Orive, Editor Erol Ak\c{c}ay, and two reviewers, and have now made a variety of major revisions to the MS in order to address their concerns.
-In particular, we have ...
-(1) made significant changes to our introduction and discussion in an attempt to engage with a broader readership by developing conceptual links between ours and previous work on evolutionarily explicit structured models and sexually antagonistic selection; (2) attempted to lessen the burden on non-theoretical readers by moving most of the development of the general 2-stage model to the Supplementary Material and following R2's many helpful suggestions regarding the presentation of the model; and (3) tried to more fully address the impacts of density dependence in three ways:  ...
-\textcolor{blue}{({\itshape i}) by revising our figures to illustrate predicted absolute population intrinsic growth rates rather than simply the boundary between demographically viable \& inviable regions} ; ({\itshape ii}) by including a brief supplement with a direct comparison between our current density-independent predictions and a new version of our model which includes simple DD; and ({\itshape iii}) by expanding our discussion about the implicaitons of DD to address these additions.
-Overall, our changes have shortened the manuscript by \textcolor{blue}{some impressive amount}, but we feel that they have greatly improved the paper.
+\noindent We received very helpful and constructive comments on our first submission and have now made a variety of major revisions to the MS in order to address them.
+First, we have made significant changes to our introduction to develop conceptual links between our model and previous work on evolutionarily explicit structured models and empirical research on sexually antagonistic selection.
+Second, we have tried to lessen the burden on non-theoretical readers by shifting development of the general model to the Supplementary Material, by following R2's many helpful suggestions regarding the presentation of the 2-stage model, and by including a new explanatory figure to help clarify the genetic and demographic outcomes from our model.
+Third, we have revised several figures to show population intrinsic growth rates ($\lambda$) rather than extinction boundaries.
+Finally, we have addressed the impacts of density dependence by ({\itshape i}) developing a new version of our model which includes simple density-dependence and reanalyzing the {\itshape M.
+guttatus} example; and ({\itshape ii}) expanding our discussion of density-dependence to address these additions.
+Overall, these various additions have lengthened the manuscript \textcolor{blue}{by about a half-page}, but we feel that they have greatly improved the paper.
 \bigskip
 
-\noindent We have provided detailed responses to each of the editors' and reviewers' comments below.
+\noindent We give detailed responses to each of the editors' and reviewers' comments below.
 For clarity, we use bold face for all line and figure references to the revised ms in our responses.
-We have also used blue text in the revised manuscript file to help identify the key additions we have made to the revised text.
+We have also used blue text in the revised manuscript file to help reviewers identify the key additions we have made.
 \bigskip
 
-\noindent Thank you again for efforts and consideration.
+\noindent Thank you again for your efforts and consideration.
 \bigskip
@@
 \begin{quote}
-	{\itshape ...}
+	{\itshape We agree that shifting the focus from deriving the general model to the specific two stage example will make the paper more accessible and shorten the Methods.
+We have now moved the development of the general model to the Online Supplementary Material (Supplement 1), and restructured the Methods in the main paper accordingly.
+}
 \end{quote}
@@
 	{\itshape This is a very good and totally valid point.
-To be honest, we believe that a separate paper is probably necessary to deal adequately with the many consequences of density dependence for our model predictions.
+To be honest, we believe that a separate paper is necessary to deal adequately with the many consequences of density dependence for our model predictions.
 However, we also agree that sidestepping this issue completely isn't good enough.
-In the end, we have made three main changes to address density dependence.
-First, we have taken your advice in the following comment and revised our figures 1 \& 4 to show absolute intrinsic growth rates across selection parameter space rather than just the boundary between demographically viable \& inviable parameter space (\textcolor{blue}{caveat about Fig.1 vs.
-4?}).
-Second, we have included a new supplementary appendix to the paper where we briefly explain how density dependence can be incorporated into the model, and provide a direct comparison between our current density-independent predictions and a new version of our model which includes DD logistic growth.
-Finally, we have expanded our results and discussion to more fully address the implications of DD for our model, with reference to our modified figures and the new supplement.}
+In the end, we have made two main changes to address density dependence.
+First, have added a new {\bf Appendix B} where we briefly explain how density dependence can be incorporated into the model, and reanalyze our {\itshape M.
+guttatus} example with a simple DD version of the model.
+Second, we have expanded our results and discussion to more fully address the implications of DD, with reference to our modified figures and the new Appendix ({\bf L.XXX--XXX}).}
 \end{quote}
@@
 	{\itshape This is great suggestion.
-As mentioned above, we have changed figures 1 \& 4 to include a heatmap of absolute values for predicted $\lambda$ instead of the extinction boundaries.
-\textcolor{blue}{However, one unfortunate consequence of presenting $\lambda$ instead of the extinction boundaries is that it is no longer possible to illustrate multiple values of maximum fertility ($f$) on each figure panel.
-In particular, Figure 1, now requires 18 panels illustrating $\lambda$ for the same parameter combinations.
-We felt that this resulted in an unnecessarily complicated figure, especially since Figure 1's main purpose was to briefly illustrate the general behaviour of the model.
-For simplicity, we have decided to include this modified Figure 1 as a supplementary figure.
-However, we now present the modified version of Figure 4 in the main text}.}
+We have now included a new illustrative figure ({\bf fig.~1}), and revised versions of figs.
+1 \& 4 (now {\bf figs.~S1 \& 5}) to show heatmaps of absolute values for predicted $\lambda$ instead of the extinction boundaries.
+However, one unfortunate consequence of presenting $\lambda$ instead of the extinction boundaries is that it is no longer possible to illustrate multiple values of maximum fertility ($f$) on each figure panel.
+In particular, our old fig.
+1 now requires 18 panels to illustrate $\lambda$ for the same set of parameter combinations.
+We felt that this resulted in an unnecessarily complicated figure, especially since fig.
+1's main purpose was to illustrate the general behaviour of the model for different fertility values.
+For simplicity, we have decided to retain our original version of fig.1 (now {\bf fig.
+2}) in the main text and include the modified version showing $\lambda$ as a supplementary figure (now {\bf fig.~XXX}); but will swap this figure if the editor/reviewers feel differently.
+However, we agree that the {\itshape M.
+guttatus} example was greatly improved by showing $\lambda$, and we now present the modified version of fig.
+4 (now {\bf fig.
+5}) in the main text}.
 \end{quote}
@@
 \begin{quote}
-	{\itshape ...}
+	{\itshape We agree that in our efforts to appeal to both demographers and evolutionary biologists studying sexual conflict, we ended up insufficiently discussing the relevant literature on both counts (but especially regarding evolutionarily explicit structured models).
+We have made a variety of changes to our introduction to improve this ({\bf L.XX--XX, L.XX-XX}).
+See also our specific responses to Reviewer 1.
+}
 \end{quote}
@@
 \begin{quote}
-	{\itshape ...}
+	{\itshape Agreed. Please see our sepecific responses to Reviewer 2 below.}
 \end{quote}
@@
 In addition to the clarifications/edits and other points raised by the reviewers, I include a handful of minor edits/typos here of my own:
+\bigskip
 
@@
 
-\noindent (4) \textcolor{red}{Figure 1; I generally prefer axes labels to have clear meanings, rather than just symbols – here, the vertical and horizontal axes give the female (vertical) vs.
-male (horizontal) selection on reproductive function.}
+\noindent (4) Figure 1; I generally prefer axes labels to have clear meanings, rather than just symbols – here, the vertical and horizontal axes give the female (vertical) vs.
+male (horizontal) selection on reproductive function.
 \begin{quote}
-	{\itshape ...}
+	{\itshape We have changed our axes labels to be more descriptive }
 \end{quote}
@@
 \begin{quote}
-	{\itshape Fixed \textcolor{red}{Note to Colin: I find that fixes sounds a bit odd? I changed it to "goes to fixation" in one instance. I trust your final judgement on this as our native.}}
+	{\itshape Fixed, or rephrased as "goes to fixation".}
 \end{quote}
 
-\noindent (6) \textcolor{red}{Figure 4; same comment as for Figure 1.}
+\noindent (6) Figure 4; same comment as for Figure 1.
 \begin{quote}
-	{\itshape ...}
+	{\itshape Fixed.}
 \end{quote}
@@
 My one comment is how might environmental stochasticity in the rate of selfing and outcrossing impact general conclusions?
-Other than that, I have no concerns about the science that is reported.
-My concerns rate the very narrow focus of the paper as written.
+\begin{quote}
+	{\itshape We now briefly discuss this possibility in the Discussion ({\bf L.586--593}).}
+\end{quote}
+
+Other than that, I have no concerns about the science that is reported. My concerns rate the very narrow focus of the paper as written.
 
@@
 \begin{quote}
-	{\itshape Thank you for the encouraging words.}
+	{\itshape Thank you for these encouraging words.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape This is a really important comment and we thank you pointing out this major shortcoming.
+	{\itshape This is a really important comment and we thank you for pointing out this major shortcoming.
 One of our main goals while crafting the first draft of the manuscript was to find a middle ground that would appeal to both demographers and people studying sexual conflict.
-It seems clear now that we didn't quite achieve that goal.
+It is clear that we didn't quite achieve that goal.
 Rather, by making compromises in the introduction we ended up lacking scholarship on both sides.
-To address your main concern, we have restructured our Introduction to more directly connect with the broader literature on evolutionarily explicit structured models (see {\bf L.XXX--XXX}) and now try to link our overview of sexually antagonistic selection theory with the existing empirical literature (see {\bf L.XXX--XXX}).
+To address your main concern, we have extended the Introduction to more directly connect with the broader literature on evolutionarily explicit structured models (see {\bf L.66-83}) and now also provide a more thorough link with the empirical sexually antagonistic selection literature (see {\bf L.42--49}).
 We hope these changes succeed in broadening the appeal of our manuscript to the wider readership of AmNat.}
@@
 
+\begin{quote}
+	{\itshape Thank you for the encouraging assessment.}
+\end{quote}
+
 I have one main comment:
@@
 However, it seems to me that it would be very important to discuss the effect of those limitations on the results themselves and therefore discuss the generality of the results.
-For instance, what is the potential effect of female dominance in the marriage function on the results and in particular on conclusion of lines 477-480
+For instance, what is the potential effect of female dominance in the marriage function on the results and in particular on conclusion of lines 477-480.
 
 \begin{quote}
-	{\itshape ...}
+	{\itshape While we feel that there is strong precedence for discussing the key results of a study before diving into the various caveats that come with the analysis, the point is well taken that we need to clearly direct readers to the places in our paper where these limitations are discussed.
+In our Methods section, we are now careful to point readers to the various Appendixes, Supplements, and places in the Discussion where these limitations are dealt with (e.g., see {\bf 100-105; 112--117; 149--155; 210; fig.~2 caption}), and we have added more explicit discussion of several of our key assumptions in the discussion ({\bf e.g., L.556--574; 594--601})}
 \end{quote}
@@
 	\begin{quote}
-	{\itshape That's a mistake on our part, apologies.}
+	{\itshape That's a mistake on our part, apologies. We have corrected this in {\bf Appendix A}, where it now appears.}
 \end{quote}
 	\item As the building blocks are given expressions (eq 10 and 11 for instance) it would be nice to mention the fact that the elements in these will be characterize later (otherwise the reader thinks she/he should already know what the expression of, say, UsAA is)
-\end{itemize}
-
 \begin{quote}
@@
 \end{quote}
+
+\end{itemize}
+
 \noindent -- Line 19: I find it weird to write that the "process of adaptation can be impeded by ...changing environmental conditions".
@@
 \begin{quote}
-	{\itshape ... our response}
+	{\itshape Very good point, that was awkward wording on our part.
+We have rephrased to more clearly explain that it is primarily the speed/magnitude of environmental change that we are referring to ({\bf L.20}).
+We have also included a new citation of a recent AmNat paper that explores some of these issues explicitly.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape ... our response}
+	{\itshape Agreed. We have rephrased to avoid the term "trade-off" which tends to over-simplify the concept in or opinion ({\bf L.25}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape ... our response}
+	{\itshape We have rephrased the first sentence of this paragraph to be more specific ({\bf L.27}), which we hope also addresses this comment.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape ... our response \textcolor{red}{I cannot figure out where in the manuscript this is, I don't think I have the exact R0 version as it was submitted?}}
+	{\itshape \textcolor{red}{Lotte, double check here?} We have added a few lines to discuss frequency vs density dependence when we first introduce the population projection matrix in equation 3 (CHECK LINE NUMBERS AFTER WE"RE DONE EDITING ).
+In addition, we have tried to clarify to improve this section by discussing that the survival and fertility rates ("demographic rates") are density-indepedent, and that the allele frequency dynamics make the model frequency dependent.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape ... our response}
+	{\itshape We have rephrased this sentence to clarify and be more consistent with our language earlier in the preceding paragraphs ({\bf L.37--39}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape ... our response}
+	{\itshape We have removed $L$ and written the full expression involving $C$.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape ... our response}
+	{\itshape This sentence has been rephrased.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape ... our response}
+	{\itshape We have added a sentence earlier in this paragraph where we remind readers what dioecy is ({\bf L.54}).
+Hermaphroditism is defined earlier in the paragraph, so an additional reference seemed redundant.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape Fixed}
+	{\itshape Fixed ({\bf L.103-104}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape Good point, fixed.}
+	{\itshape Good point, fixed ({\bf L.161, 185}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape ... our response}
+	{\itshape We want to thank the reviewer for this suggestion.
+We now start with a figure that only has one f, and that separately explains what happens to the allele frequencies (Population genetic outcome) and to the population growth rate (Demographic Outcome), see new {\bf fig.
+1}.
+We hope that this will make it easier to interpret the old fig.
+1, which is now {\bf fig.
+2}.
+We have also included a new supplementary figure, {\bf fig.
+S1}, with detailed heatmaps for each of the parameter combinations presented in the this figure (i.e., one $f$ per panel).
+However, we felt that this revised figure was to large and complicated to present in the main text (18 vs.
+6 panels).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape ... our response}
+	{\itshape We have added a brief parenthesis to clarify ({\bf L.434--435}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape ... our response}
+	{\itshape This figure now appears as four separate heatmaps, each with one set of invasion conditions ({\bf fig.~5}). We hope this eases interpretation of the figure.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape ... our response}
+	{\itshape We have re-written this sentence to avoid this awkward phrasing ({\bf L.579--580}).}
 \end{quote}

### 2021-12-20 doc/SA-Hermaphrodites-wDemography-R1.tex
Full edits first pass
This commit includes a first pass at all major edits/revisions to the
manuscript. Ready to hand it over.

--- 2021-12-17 doc/SA-Hermaphrodites-wDemography-R1.tex
@@
 %\section{Outline}
-
-%OK, just putting some ideas out here based roughly on our conversations about an {\itshape AmNat}-style article, the results we have so far, and what I think we should be able to do with data from COMPADRE.
-
-%\begin{enumerate}
-	%\item {\bf Introduction}
-	%\begin{enumerate}
-%		\item Briefly introduce sexual antagonism and intra-locus sexual conflict. (Colin)
-%		\item Simultaneous hermaphrodites, mixed mating systems, unique predictions for SA polymorphism. (Colin)
-%		\item Introduce Evolutionary Demography, pivot to focus on the consequences of taking demography into account explicitly for the maintenance of SA polymorphism. (Lotte)
-%		\item Introduce key difference between existing population genetic models of SA polymorphism and our models: possibility of diverse population and evolutionary dynamics. (Lotte then Colin)
-%		\item Explain and highlight "demographically viable" parameter space, indicate that we are interested in whether parameter space predicted to be polymorphic by Pop.~gen.~models is, indeed, "demographically viable" (Lotte then Colin)
-%		\begin{enumerate}
-%			\item Pop.
-Gen.
-models tacitly assume stable population size, parameterize genotypic fitnesses relative to the 'best performing' genotype.
-Non-overlapping generations.
-Casually define fitness as lifetime reproductive fitness of a given genotype.
-%			\item "In practice, these genotypic fitnesses emerge from myriad process acting throughout the life-history of individuals"
-%			\item these processes contribute to key demographic rates such as the intrinsic population growth rate, 
-%			\item 
-%		\end{enumerate}
-%		\item Building upon the two-sex stage-structured matrix models developed by \citet{deVriesCaswell2019a,deVriesCaswell2019b}, we develop stage-structured matrix models that can accommodate selection through male and female function in simultaneous hermaphrodites with mixed mating systems.
-The models also allow for the effects of inbreeding depression to manifest in several demographic parameters, allowing us to explore the fitness consequences of early- vs.~late-acting inbreeding depression.
-(Colin then Lotte)
-	%\end{enumerate}
-	%\item {\bf Methods}
-%		\begin{enumerate}
-%			\item Citation heavy review of de Vries \& Caswell AmNat (2018) \& de Vries \& Caswell AmNat (2019) TheorPopBiol
-%			\item Briefly describe extension to simultaneous hermaphrodites, citing Jordan \& Connallon (2014), Abbot \& Tazzyman (2014), Olito (2017). 
-%			\item Breakdown into individuals produced by selfing vs.~outcrossing.
-%			\item Brief anatomy of $\mathbf{\tilde{A}}$.
-%			\item Analysis. Invasion analysis \& abridged derivation of invasion conditions \& identification of extinction threholds.
-%			\item Cite Appendix with full derivation distilled from notes file.
-%		\end{enumerate}
-	%\item {\bf When is polymorphism demographically viable?}
-	%\begin{enumerate}
-%		\item Review previous predictions for outcrossers \& partial selfing
-%		\begin{enumerate}
-%			\item Greater scope for SA polymorphism in outcrossers vs.~selfers
-%			\item Greater scope for SA polymorphism when SA fitness effects are recessive (dominance reversal).
-%			\item but our model allows us to see where/when invasion of SA alleles can drive the population extinct
-%			\item \textbf{Fig.1 - illustration of funnel plots w/ extinction isoclines}
-%		\end{enumerate}
-%		\item Key results summarized in \textbf{Fig.~2}: 
-%		\begin{enumerate}
-%			\item For simplicity, temporarily assume no inbreeding depression ($\delta = 0$).
-%			\item When populations do not have high growth-rates (i.e., when $\lambda \approx 1$), much of of polymorphic parameter space becomes demographically inviable.
-%			\item Specifically, would-be polymorphic paramter space becomes inviable when segregating male-beneficial/female-deleterious alleles can reduce fecundity enough to drive the population extinct.
-%			\item This effect is most dramatic for separate-sexed species and predominant outcrossers.
-Selfing shelters the population from the female-deleterious fitness effects because there is no opportunity for selection via male function.
-This can also be thought of as reproductive assurance in our model....
-%			\item Convergence to pop.~gen.~predictions when $f$ is big.
-%		\end{enumerate}
-	%\end{enumerate}
-	%\item {\bf Effects of inbreeding depression}
-	%\begin{enumerate}
-%		\item Review effect of I.D.~in population genetic models: causes selfing populations to look more like outcrossing ones (at least in terms of where SA polymorphism is maintained).
-This is because a fraction of selfed ovules are aborted, meaning the overall ratio of outcrossed vs.~selfed progeny increases.
-%		\item Matrix model allows us to explore effects of I.D.
-in different parts of the life-cycle (early- vs.~late-acting I.D.).
-Point out that previous pop.~gen.~models only address early-acting I.D.
-via ovule viability.
-%		\item MAKE CLEAR that purging of deleterious recessives thought to cause I.D. is already included in the model results presented in Fig.~3
-%		\item Key results summarized in Fig.~3
-%		\begin{enumerate}
-%			\item Pop.~Gen.~models predict that higher early-acting ID should lead to greater proportion of polymorphic parameter space, while our models predict that we get a critical threshold of selfing at which population viability plummets.
-%			\item When and how I.D. influences fitness during the life-cyclecan change the location and steepness of the decline in polymorphic parameter space.
-%			\item Critically, largest deviations from pop.~gen.~predictions happen for predominantly selfing populations.
-%			\item A perhaps obvious, but sometimes overlooked implication of our model predictions: selfing can offer reproductive assurance, and SA polymorphisms can be maintained in a non-trivial proportion of parameter space in partially selfing populations...
-BUT demographically, the population must be able to "afford" the loss of selfed ovules/offspring.
-%			\item HIGHLIGHT that this result holds even though we incorporated decrease of the mutation load due to inbreeding (and corresponding decrease in I.D.)...	
-%		\end{enumerate}
-	%\end{enumerate}
-	%\item {\bf Mimulus as a potential real world example}
-	%\begin{enumerate}
-%		\item We can parameterize our model with real populations' demographic rates, and make explicit predictions regarding the opportunity for SA polymorphism given the species' demography.
-We provide two illustrative examples:
-%		\item Opportunities for SA polymorphism in {\itshape Mimulus guttatus} (now {\itshape Erythranthe guttata})
-%		\item Possibly use data from \citet{PetersonEtAl2016} Eagle Meadows population, which is the 'local' population in their big common garden local adaptation expeirment.
-%		\begin{enumerate}
-%			\item Illustrate funnel plot using {\itshape E.~gutatta} demographic rates.
-%			\item \textbf{Fig.~4}: PLOT inversion from \citet{LeeKelly2015}, on that funnel!
-%		\end{enumerate}
-	%\end{enumerate}
-	%\item Discussion
-%		\begin{enumerate}
-%			\item Classic prediction from demography that most species should have a $\lambda \approx 1$
-%			\item If true, the demographic impacts on the opportunities for the maintenance of sexually antagonistic polymorphisms may be pervasive...
-%			\item Condition dependence, fecundity, scope for sexual confict/antagonistic polymorphism?
-%			\item Do sexually dimorphic species exhibit demographic rates that increase opportunities for SA polymorphism?
-%		\end{enumerate}
-%\end{enumerate}
-
 %\tableofcontents
@@
 \newpage{}
-\section*{Abstract}%\textcolor{green}{ I like this abstract best (CdV)} \textcolor{blue}{Agreed (CJO)}}
+\section*{Abstract}
 
@@
 To persist in the long term, a population must be able to adapt to its environment.
-Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, such as deleterious mutations \citep{Haldane1937}, changing environmental conditions \citep{Maynard-Smith1976, LandeShannon1996,OrrUnckless2008}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, and genetic constraints \citep{ConnallonHall2018, MatthewsConnallon2019}.
-Genetic constraints on adaptation can arise as a consequence of conflicting selection and gene-flow between different classes of individuals, or more generally, when there are genetic trade-offs between fitness components \citep{CharlesworthHughes2000, ConnallonHall2018}.
-
-\textcolor{blue}{Genetic constraints on adaptation arising from conflicting selection} have particularly interesting evolutionary consequences.
+Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, such as deleterious mutations \citep{Haldane1937}, \textcolor{blue}{rapid, highly unpredictable, or extreme changes} in environmental conditions \citep{Maynard-Smith1976, LandeShannon1996, OrrUnckless2008,Lyberger-etal-2021}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, and genetic constraints \citep{ConnallonHall2018, MatthewsConnallon2019}.
+Genetic constraints on adaptation can arise as a consequence of conflicting selection and gene-flow between different classes of individuals, or more generally, when \textcolor{blue}{there is negative correlation between different fitness components} \citep{CharlesworthHughes2000, ConnallonHall2018}.
+
+\textcolor{blue}{Genetic constraints on adaptation arising from conflicting selection and gene flow} have particularly interesting evolutionary consequences \citep{ConnallonDebarreLi2018}.
 On the one hand, they provide an effective mechanism for the maintenance of genetic variation, which can increase a population's capacity for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
@@
 SA selection appears to be} a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and the evolution of sexual dimorphism and sex-related traits \citep{Kidwell1977, Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, ConnallonClark2012, BondurianskyChenoweth2009, Olito2019}.
-\textcolor{blue}{Sex-specific estimates of phenotypic selection indicate that sex-differences in directional selection are common \citep[e.g.,][]{DeLisle-etal-2018,Lewis-etal-2011,CoxCalsbeek2009,SinghPunzalan2018}, implying that many genetic variants underlying quantitative traits have SA fitness effects.
-Similarly, quantitative genetic studies frequently find negative cross-sex genetic correlations for fitness, indicating SA fitness effects for multilocus genotypes (\citealt{Chippendale-etal-2001,Delph-etal-2011}; reviewed in \citealt{ConnallonMatthews2019}).
+\textcolor{blue}{For instance, phenotypic selection studies commonly find sex-differences in directional selection \citep[e.g.,][]{DeLisle-etal-2018,Lewis-etal-2011,CoxCalsbeek2009,SinghPunzalan2018}, implying that many genetic variants underlying quantitative traits have SA fitness effects.
+Likewise, quantitative genetic studies frequently find negative cross-sex genetic correlations for fitness, indicating SA fitness effects for multilocus genotypes (\citealt{Chippendale-etal-2001,Delph-etal-2011}; reviewed in \citealt{ConnallonMatthews2019}).
 Moreover, polygenic signals of SA selection are evident in recent population genetic analyses of large genomic datasets \citep{RuzickaESEB2020, RuzickaConnallon2022}.}
 
-Alleles with opposing fitness effects through male and female sex functions can also cause SA selection in hermaphrodite populations, where individuals' overall fitness is determined jointy by maternal and paternal reproductive success \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
+Alleles with opposing fitness effects through male and female sex functions can also cause SA selection in hermaphrodite populations, where individuals' overall fitness is determined jointly by their maternal and paternal reproductive success \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
 \textcolor{blue}{In obligately outcrossing hermaphrodite populations, SA selection operates similarly as in seperate-sexed ones (dioecious species).
@@
 Yet, other factors, such as genetic linkage to other SA loci or a sex-determining region \citep{Otto2011, JordanCharlesworth2012, Olito2017, Olito2019}, antagonism between viability and fecundity selection \citep{Glemin2021}, or spatial heterogeneity and complexity of the life-cycle \textcolor{blue}{can all facilitate the maintenance of} SA polymorphisms \citep{Olito-etal-2018,ConnallonSharmaOlito2019, Glemin2021}.
-Overall, both theoretical predictions and current empirical data suggest that there is ample scope for SA trade-offs and the maintenance of SA polymorphisms in both dioecious and hermaphrodite populations \citep{Abbott2011, DelphHerlihy2012, WangBarrett2020, RuzickaESEB2020, RuzickaConnallon2022}.
-
-\textcolor{blue}{While significant efforts have been made to predict when SA polymophisms can be maintained and to detect SA genetic variation from empirical data, much less consideration has been given to the demographic consequences of the maladaptation caused by SA selection}.
-Most population genetic models, including models of the maintenance of SA polymorphisms, keep track of the relative fitnesses of genotypes in populations of constant (and often infinite) size, and do not consider the population dynamical consequences of evolutionary change.
-In reality, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals \textcolor{blue}{which link genotype distributions to population dynamics (e.g., \citealt {johnston2013life, merot2020balancing}).
-Over the past 50 years, a rich field of eco-evolutionary dynamics has developed which seeks to jointly model the genetic and demographic processes involved in evolutionary change.
+Overall, both theoretical predictions and current empirical data suggest that there is ample scope for SA \textcolor{blue}{fitness variation} and the maintenance of SA polymorphisms in both dioecious and hermaphrodite populations \citep{Abbott2011, DelphHerlihy2012, WangBarrett2020, RuzickaESEB2020, RuzickaConnallon2022}.
+
+\textcolor{blue}{While significant efforts have been made to understand when balanced SA polymophisms can be maintained and to detect SA genetic variation from empirical data, much less consideration has been given to the demographic consequences of the maladaptation caused by SA selection}.
+Most population genetic models, including models of SA selection, model the relative fitness of genotypes in populations of constant (and often infinite) size, and do not consider the population dynamical consequences of evolutionary change.
+In reality, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals \textcolor{blue}{that link genotype distributions to population dynamics (e.g., \citealt {johnston2013life, merot2020balancing}).
+Over the past 50 years, the field of eco-evolutionary dynamics has sought to jointly model the genetic and demographic processes involved in evolutionary change.
 In particular, theoretical models combining population genetics with density-dependent population growth \citep{roughgarden1971density} and stage-structured populations (consolidated in \citealt{Charlesworth_1994}) paved the way toward understanding the demographic changes accompanying allele frequency changes due to selection.
-Extensions of these theories to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and quantitative traits \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects} all highlight the often counterintuitive demographic consequences of adaptive evolution that are often missed by classical population genetics models (comprehensively reviewed by \citealt{fussmann2007eco} and \citealt{govaert2019eco}).}
+Extensions of these theories to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and quantitative traits \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects} all highlight the sometimes counterintuitive demographic consequences of adaptive evolution that are missed by classical population genetics models (see \citealt{fussmann2007eco} and \citealt{govaert2019eco} for comprehensive reviews).
+More generally, eco-evolutionary approaches contribute a }
 
@@
 \citet{deVriesCaswell2019a} introduced a mendelian matrix model with intralocus sexual antagonism and population dynamics, but did not analyze the consequences of demographic viability for the scope of sexual antagonism to maintain polymorphism.
-In this paper, we extend the model of \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b} to include both dioecious and hermaphroditic \textcolor{blue}{populations} in order to study the parameter conditions under which different outcomes of SA selection, including fixation of male- or female-beneficial alleles as well as balanced polymorphisms, are also demographically viable.
+In this paper, we extend the model of \citet{deVriesCaswell2019a,deVriesCaswell2019b} to include both dioecious and hermaphroditic \textcolor{blue}{populations} in order to study when different genetic outcomes of SA selection, including fixation of male- or female-beneficial alleles as well as balanced polymorphisms, are also demographically viable.
 A major strength of our approach is that the model can be parameterized using empirically estimated demographic rates, enabling us to make predictions about the scope for demographically viable SA polymorphisms that are grounded in the biology of real populations.
@@
 Here, we briefly describe a matrix model incorporating multiple life-cycle stages and a single diallelic locus under SA selection on male and female fertility for populations of partially-selfing simultaneous hermaphrodites.
-\textcolor{blue}{Rather than go through the full mathematical derivation of the general model, we showcase a two-stage example of our model and discuss the biological assumptions that go into the model.
-The derivation of the general model for any number of age or stage classes is in the Online Supplementary Material.} The derivation follows closely the logic presented in \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b}.
-In fact, the model reduces to the stage-structured model of \citet{deVriesCaswell2019a} under obligate outcrossing, \textcolor{blue}{that is, when selfing is not allowed}.
+\textcolor{blue}{We focus our presentation on a two-stage example of the model and discuss its' underlying biological assumptions.
+The derivation of the general model for any number of age or stage classes is provided in the Online Supplementary Material.} The derivation follows closely the logic presented in \citet{deVriesCaswell2019a, deVriesCaswell2019b}, and in fact, it reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019a} under obligate outcrossing \textcolor{blue}{(when selfing is not allowed)}.
 All computer code necessary to reproduce the results are available at \url{https://anonymous.4open.science/r/SA-Hermaphrodites-wDemography-BF66}.%\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
@@
 To distinguish between parameters relating to male and female function, we denote matrices or vectors relating to the male sex function with a prime.
+\textcolor{blue}{Following previous theory for partially selfing populations \cite[e.g.,][]{Charlesworth2010,JordanConnallon2014,Glemin2021}, we use a 'fixed prior selfing' model where all individuals are assumed to self-fertilize at a constant rate, $C$, prior to receiving outcross pollen, for reasons of analytic tractability.
+We present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}) in the Online Supplement, and briefly discuss implications of this approach in the Discussion.}
 
@@
 The population state at time $t$ is described by a population state vector, $\tilde{\mbf{n}}(t)$, which is ordered by how individuals were produced (selfing vs.~outcrossing), then by genotype, and finally by stage.
-For a single diallelic locus with alleles $A$ and $a$, we have three genotypes ($AA,\, Aa,\, aa$; $g = 3$), giving the population state vector:
+For a single diallelic locus with alleles $A$ and $a$, we have three genotypes ($AA,\, Aa,\, aa$; $g = 3$), giving the population state vector is:
 \begin{linenomath*}
@@
 \end{linenomath*}
-\textcolor{blue}{If the survival or fertility rates are a function of the population density, $\tilde{\mbf{n}}(t)$, then the model will be a density dependent model and the population projection matrix is a function of the full population vector, $\tilde{\mbf{A}}[\tilde{\mbf{n}}(t)]$.
-In the online supplementary materials we show results for a density-dependent version of the model.
-In the main text of this paper, we will consider models with density-independent survival and fertility rates.
-The population projection matrix is then only a function of the proportional population vector, $\tilde{\mbf{A}}[\tilde{\mbf{p}}(t)]$, and the model is referred to as a frequency-dependent model.}
 
@@
 
-\subsubsection*{Sexually antagonistic selection \& inbreeding depression} \label{sec:SAsel}
-
-\textcolor{blue}{To demonstrate how to construct and analyze a genotype $\times$ stage-classified model, we will now construct an example model for a hypothetical species with intralocus sexual conflict via the two sex functions.
-}For the sake of simplicity, we assume our hypothetical species has a life cycle with only two stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active.
-Suppose that there is a genetic trade-off between the sex-functions at a single diallelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect.
-Following convention, we parameterize the fertility component of fitness for each genotype through each sex function, $w_{i}$ and $w^{\prime}_{i}$, to be bounded by $[0,1]$, with dominance and selection coefficients $h_f, s_f$ and $h_m, s_m$ determining the decrease in fertility through each sex function relative to the most fit genotype ($AA$ has highest female fertility, $aa$ the highest male fertility; see Table \ref{tab:Fitness}).
-
-The SA locus does not affect survival or transition rates.
-However, the survival matrices can be used to model the fitness effects of inbreeding depression at later stages of development by allowing different stage-specific survival and transition rates for individuals produced by self-fertilization vs.~outcrossing.
-By contrast, the parameter $\delta$ only affects inbreeding depression through viability of selfed ovules.
-With this in mind, we define survival matrices for individuals produced by selfing and outcrossing as follows:
-\begin{linenomath*}
-\begin{equation} \label{eq:US}
-	\mbf{U}^S = \left(
-					\begin{array}{cc}
-						\sigma_j (1 - \delta_j) \big(1 - \gamma (1 - \delta_{\gamma}) \big) & 0 \\
-						\sigma_j (1 - \delta_j) \gamma (1 - \delta_{\gamma})      & \sigma_a (1 - \delta_a)
-					\end{array}
-				\right) \\
-\end{equation}
-\end{linenomath*}
-\noindent and 
-\begin{linenomath*}
-\begin{equation}\label{eq:UX}
-	\mbf{U}^X = \left(
-					\begin{array}{cc}
-						\sigma_j(1 - \gamma) & 0 \\
-						\sigma_j \gamma      & \sigma_a
-					\end{array}
-				\right)
-\end{equation}
-\end{linenomath*}
-
-\noindent where $\sigma_j$ and $\sigma_a$ are the juvenile and adult stage survival rates, $\gamma$ is the maturation rate from juvenile to adult stages, and the corresponding $\delta_j$, $\delta_a$, and $\delta_{\gamma}$ terms denote the proportional decreases in stage-specific survival and transition rates due to inbreeding depression (i.e., deleterious effects of inbreeding at later life-history stages; e.g., \citealt{HarderRoutely2006}).
-For simplicity, we assume survival and transition rates are constant among genotypes.
-
-Throughout our analyses, we distinguish between early- and late-acting inbreeding depression.
-We quantify early-acting inbreeding depression using $\delta$, and late-acting inbreeding depression using $\delta_i$ (where $i \in \{j,a,\gamma\}$).
-$\delta$ denotes the fraction of self-fertilized ovules that do not develop into juveniles due to inbreeding depression.
-An important difference between early- and late-acting inbreeding depression in the model is that $\delta$ affects the production of new individuals, whereas the $\delta_i$ affect the demographic rates of extant individuals, contained in $\mbf{U}^S$ (see Eq.~\ref{eq:US}).
-
-The fertility matrices through female and male function are
-\begin{linenomath*}
-\begin{equation}\label{eq:F_female}
-	\mbf{F}_{i} = \left(
-					\begin{array}{cc}
-						0 & f w_{i} \\
-						0 & 0
-					\end{array}
-				\right),
-\end{equation}
-\end{linenomath*}
-\noindent and
-\begin{linenomath*}
-\begin{equation}\label{eq:F_male}
-	\mbf{F}^{\prime}_{i} = \left(
-					\begin{array}{cc}
-						0 & f^{\prime} w^{\prime}_{i} \\
-						0 & 0
-					\end{array}
-				\right),
-\end{equation}
-\end{linenomath*}
-
-\noindent where $f$ and $f^{\prime}$ represent maximum adult fertilities, and $w_{i}$ and $w^{\prime}_{i}$ the genotypic relative scaling factors for female and male sex-functions (see Table \ref{tab:Fitness}).
-
-\begin{table}[htbp]
- \centering
- \caption{\bf Relative fertilities for Sexually Antagonistic selection ($w_{i}$)}
-\label{tab:Fitness}
-\begin{tabular}{lccc}
- \toprule
-					&  \multicolumn{3}{c}{{\textit{Genotype}}} \\ 
-\cline{2-4}
-					& $AA$			& $Aa$ 					& $aa$ 		\\ \hline
-Female function ($w_{i}$):	& $1$		& $1 - h_f s_f$	& $1 - s_f$ \\	
-Male function ($w^{\prime}_{i}$):		& $1 - s_m$& $1 - h_m s_m$	& $1$ 		\\	
-\hline
-\end{tabular}
-\end{table}
-
-\subsubsection*{Mating and offspring production under partial selfing}
- Outcrossing in our model proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}.
-That is, each individual's genotype determines both the number of ovules produced and the paternal mating success, broadly defined.
-For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986, WangBarrett2020, Harder2016}.
-Note, however, that by modeling paternal relative mating success rather than pollen/sperm production, we implicitly assume female demographic dominance (i.e., production/transport of male gametes does not limit ovule fertilization and hence population growth), a point we return to in the Discussion.
-
-\textcolor{blue}{ For the purpose of this article, we also assume that females in the population all have the same mating preferences regardless of female age, stage, or genotype.
-The model can be extended to consider assortative mating, but that is beyond the scope of the current paper.
-In addition, we assume that males produced through selfing or through outcrossing do not differ in their mating success.
-Given these assumptions, the allele frequencies in the male mating population can obtained by simply summing the vectors of individuals produced through selfing and outcrossing:
-\begin{linenomath*}
-\begin{equation}
- 	\mbf{p}^{\prime}=\frac{(\mbf{n}^X + \mbf{n}^S)}{\| (\mbf{n}^X + \mbf{n}^S)\|}.
- \end{equation}
-\end{linenomath*}}
-
-\textcolor{blue}{Males contribute  gametes to the population gamete pool according to their mating success, which depends on their stage and genotype at the SA locus.
-The allele frequencies in the male gamete pool are then calculated as follows (see Online Supplementary Materials for a derivation of this equation),
-\begin{linenomath*}
-\begin{equation} \label{eq:maleGametePool}
-	\left(
-		\begin{array}{c}
-			q^{\prime}_{A} \\
-			q^{\prime}_{a} \\
-		\end{array} \right) \propto 
-				\left(
-		\begin{array}{c}
-			\mbf{1}^{\intercal}_{\omega}\mbf{F}^{\prime}_{AA}\mbf{p}^{\prime}_{AA}+ \frac{1}{2} \mbf{1}^{\intercal}_{\omega}  \mbf{F}^{\prime}_{Aa} \mbf{p}^{\prime}_{Aa}  \\
-		\frac{1}{2} \mbf{1}^{\intercal}_{\omega}  \mbf{F}^{\prime}_{Aa} \mbf{p}^{\prime}_{Aa} + \mbf{1}^{\intercal}_{\omega}\mbf{F}^{\prime}_{aa}\mbf{p}^{\prime}_{aa}\\
-		\end{array} \right),
-\end{equation}
-\end{linenomath*}
-where equality is obtained by normalizing the frequencies such that they add to one.
-The first entry is simply summing the stage-specific contributions to the gamete pool of heterozygotes and $AA$ homozygotes; the second entry sums the stage-specific contributions to the gamete pool of heterozygotes and $aa$ homozygotes.
-}
-
-\textcolor{blue}{ FWe now have all the ingredients to calculate the allele frequencies in the (male) gamete pool for the two stage example model in equations \ref{eq:F_male}, \ref{eq:maleGametePool}, and Table \ref{tab:Fitness}.
-Denoting the proportion of adults of genotype $i$ with $p^A_i$, the allele frequencies in the male gamete pool are given by
-\begin{linenomath*}
-\begin{equation} \label{eq:maleGametePool_example}
-\left(
-		\begin{array}{c}
-			q^{\prime}_{A} \\
-			q^{\prime}_{a} \\
-		\end{array} \right) = 
-				\frac{1}{(1-s_m)p^A_{aa}+ (1-hs_m)p^A_{Aa}+p_{aa}}\left(
-		\begin{array}{c}
-(1-s_m)p^A_{AA}	+\frac{1}{2}(1-hs_m)p^A_{Aa}\\
-p^A_{aa}+\frac{1}{2}(1-hs_m)p^A_{Aa}
-		\end{array} \right).
-\end{equation}
-\end{linenomath*}
-Note that this equation is identical to the classical equation for frequencies of male gametes in a sexually antagonistic model except that now we have to specify that only adults contribute to the gamete pool whereas in a classical population genetic model with no population structure, we do not distinguish between reproducing adults and non-reproductive juveniles.
-Of course equation (\ref{eq:maleGametePool_example}) starts deviating from the classical population genetics case more drastically if we include multiple reproducing adult age or stage classes (see equation (\ref{eq:maleGametePool})).
-}
-
-\subsubsection*{Population projection}
-Using the component matrices described above (the survival matrices, $\mathbf{U}^X_i$, $\mathbf{U}^S_i$, the fertility and mating success matrices, $\mathbf{F}_i$, $\mathbf{F}^\prime_i$) and the allele frequencies in the gamete pool, we can construct the population projection matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$, see Online Supplementary Material for the step by step construction of the model.
-The resulting matrix that projects the eco-evolutionary dynamics is:
-\begin{linenomath*}
-\begin{equation} \label{eq:Atilde}
-	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
-			\underbrace{\left(
-			\begin{array}{c|c}
-				\mcal{U}^S & \mbf{0} \\ \hline
-				\mbf{0} & \mcal{U}^X \\
-			\end{array} \right)}_{\tilde{\mbf{U}}} + 
-			\underbrace{\left(
-			\begin{array}{c|c}
-				\mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
-				\mcal{F}^X(\tilde{\mbf{p}}) (1 - C) & \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)\\
-			\end{array} \right)}_{\tilde{\mbf{F}}},
-\end{equation}
-\end{linenomath*}
-where $C$ denotes the proportion of each individual's ovules that are self-fertilized (the remaining $1 - C$ are outcrossed), and  $\delta$ represents the proportion of self-fertilized zygotes that fail to develop due to inbreeding depression during early development \citep{Charlesworth1987}.
-
-The blocks of the component matrices in Eq(\ref{eq:Atilde}) correspond to production of offspring by self-fertilization and outcrossing ($\mcal{F}^S$ and $\mcal{F}^X$ in $\tilde{\mbf{F}}$), and survival of extant individuals produced by selfing or outcrossing ($\mcal{U}^S$ and $\mcal{U}^X$ in $\tilde{\mbf{U}}$).
-The survival matrices for individuals produced through selfing and outcrossing are,
-\begin{linenomath*}
-\begin{eqnarray} 
-	\mcal{U}^S  = 
-		\left(
-			\begin{array}{ccc}
-				\mbf{U}^{S}_{AA} & \mbf{0} & \mbf{0} \\
-				\mbf{0} & \mbf{U}^{S}_{Aa} & \mbf{0} \\
-				\mbf{0} & \mbf{0} & \mbf{U}^{S}_{aa} \\
-			\end{array} \right),\label{eq:BlkUS}\\
-				\mcal{U}^X  = 
-		\left(
-			\begin{array}{ccc}
-				\mbf{U}^{X}_{AA} & \mbf{0} & \mbf{0} \\
-				\mbf{0} & \mbf{U}^{X}_{Aa} & \mbf{0} \\
-				\mbf{0} & \mbf{0} & \mbf{U}^{X}_{aa} \\
-			\end{array} \right).\label{eq:BlkUX}
-\end{eqnarray}
-\end{linenomath*}
-Since individuals cannot change their genotype once they are born, the survival matrices are block diagonal.
-
-The fertility matrices for individuals produced through selfing and outcrossing are given by, 
-\begin{linenomath*}
-\begin{equation} \label{eq:BlkFS}
-	\mcal{F}^S = 
-			\left(
-			\begin{array}{ccc}
-				\mbf{F}_{AA} & \frac{1}{4} \mbf{F}_{Aa} & \mbf{0}  \\
-				\mbf{0}  & \frac{1}{2} \mbf{F}_{Aa} & \mbf{0}  \\
-				\mbf{0} & \frac{1}{4} \mbf{F}_{Aa} & \mbf{F}_{aa}\\
-			\end{array} \right), 
-\end{equation}
-\end{linenomath*}
-\noindent and 
-\begin{linenomath*}
-\begin{equation} \label{eq:BlkFX}
-	\mcal{F}^X = 
-			\left(
-			\begin{array}{ccc}
-				q^{\prime}_{A} \mbf{F}_{AA} & \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} & \mbf{0}  \\ 
-				q^{\prime}_{a} \mbf{F}_{AA} & \frac{1}{2} \mbf{F}_{Aa} & q^{\prime}_{A} \mbf{F}_{aa}  \\ 
-				\mbf{0}  & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} & q^{\prime}_{a} \mbf{F}_{aa}  \\
-			\end{array} \right),
-\end{equation}
-\end{linenomath*}
-\noindent where $q^{\prime}_A$ and $q^{\prime}_a$ are the frequencies of alleles $A$ and $a$ in the gamete pool, given by Eq(\ref{eq:maleGametePool}).
-
-The blocks of $\mcal{F}^X(\tilde{\mbf{p}})$ can be constructed and interpreted as follows: The first row block of the first column produces $AA$ offspring by outcrossing from $AA$ maternal parents.
-This happens when the $AA$ maternal parent receives an $A$ gamete from the male gamete pool, which happens with probability $q^{\prime}_{A}$.
-The other blocks can be interpreted similarly.
-
-Combining all the component matrices yields the overall eco-evolutionary projection matrix shown in Appendix A.
+\textcolor{blue}{If the survival or fertility rates are a function of the population density, $\tilde{\mbf{n}}(t)$, then the model will be a density dependent model and the population projection matrix becomes a function of the full population vector, $\tilde{\mbf{A}}[\tilde{\mbf{n}}(t)]$.
+Below, we limit our analyses to consider density-independent survival and fertility rates.
+The population projection matrix is then only a function of the proportional population vector, $\tilde{\mbf{A}}[\tilde{\mbf{p}}(t)]$, and the model is referred to as a frequency-dependent model.
+In \ref{App:DensDep} we construct and analyze a density-dependent version of the model and show that with stage-independent density-dependent growth, our main conclusions remain intact.}
 
@@
 $\delta_a$ & \textcolor{blue}{ Inbreeding depression effect on adult survival}& \\
-$\delta_{\gamma}$ &\textcolor{blue}{  Inbreeding depression effect on maturation rate} & \\
+$\delta_{\gamma}$ &\textcolor{blue}{Inbreeding depression effect on maturation rate} & \\
 $f_i,\,f^{\prime}_i$ & Adult fertilities through female and male sex functions respectively & \\
@@
 \newpage{}
+
+\subsubsection*{Sexually antagonistic selection \& inbreeding depression} \label{sec:SAsel}
+
+\textcolor{blue}{To demonstrate how to construct and analyze a genotype $\times$ stage-classified model, we will now construct an example model for a hypothetical species with intralocus sexual conflict via the two sex functions.
+}For the sake of simplicity, we assume our hypothetical species has a life cycle with only two stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active.
+Suppose that there is a genetic trade-off between the sex-functions at a single diallelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect.
+Following convention, we parameterize the fertility component of fitness for each genotype through each sex function, $w_{i}$ and $w^{\prime}_{i}$, to be bounded by $[0,1]$, with dominance and selection coefficients $h_f, s_f$ and $h_m, s_m$ determining the decrease in fertility through each sex function relative to the most fit genotype ($AA$ has highest female fertility, $aa$ the highest male fertility; see Table \ref{tab:Fitness}).
+
+The SA locus does not affect survival or transition rates.
+However, the survival matrices can be used to model the fitness effects of inbreeding depression at later stages of development by allowing different stage-specific survival and transition rates for individuals produced by self-fertilization vs.~outcrossing.
+By contrast, the parameter $\delta$ only affects inbreeding depression through viability of selfed ovules.
+With this in mind, we define survival matrices for individuals produced by selfing and outcrossing as follows:
+\begin{linenomath*}
+\begin{equation} \label{eq:US}
+	\mbf{U}^S = \left(
+					\begin{array}{cc}
+						\sigma_j (1 - \delta_j) \big(1 - \gamma (1 - \delta_{\gamma}) \big) & 0 \\
+						\sigma_j (1 - \delta_j) \gamma (1 - \delta_{\gamma})      & \sigma_a (1 - \delta_a)
+					\end{array}
+				\right) \\
+\end{equation}
+\end{linenomath*}
+\noindent and 
+\begin{linenomath*}
+\begin{equation}\label{eq:UX}
+	\mbf{U}^X = \left(
+					\begin{array}{cc}
+						\sigma_j(1 - \gamma) & 0 \\
+						\sigma_j \gamma      & \sigma_a
+					\end{array}
+				\right)
+\end{equation}
+\end{linenomath*}
+
+\noindent where $\sigma_j$ and $\sigma_a$ are the juvenile and adult stage survival rates, $\gamma$ is the maturation rate from juvenile to adult stages, and the corresponding $\delta_j$, $\delta_a$, and $\delta_{\gamma}$ terms denote the proportional decreases in stage-specific survival and transition rates due to inbreeding depression (i.e., deleterious effects of inbreeding at later life-history stages; e.g., \citealt{HarderRoutely2006}).
+For simplicity, we assume survival and transition rates are constant among genotypes.
+
+Throughout our analyses, we distinguish between early- and late-acting inbreeding depression.
+We quantify early-acting inbreeding depression using $\delta$, and late-acting inbreeding depression using $\delta_i$ (where $i \in \{j,a,\gamma\}$).
+$\delta$ denotes the fraction of self-fertilized ovules that do not develop into juveniles due to inbreeding depression.
+An important difference between early- and late-acting inbreeding depression in the model is that $\delta$ affects the production of new individuals, whereas the $\delta_i$ affect the demographic rates of extant individuals, contained in $\mbf{U}^S$ (see Eq.~\ref{eq:US}).
+
+The fertility matrices through female and male function are
+\begin{linenomath*}
+\begin{equation}\label{eq:F_female}
+	\mbf{F}_{i} = \left(
+					\begin{array}{cc}
+						0 & f w_{i} \\
+						0 & 0
+					\end{array}
+				\right),
+\end{equation}
+\end{linenomath*}
+\noindent and
+\begin{linenomath*}
+\begin{equation}\label{eq:F_male}
+	\mbf{F}^{\prime}_{i} = \left(
+					\begin{array}{cc}
+						0 & f^{\prime} w^{\prime}_{i} \\
+						0 & 0
+					\end{array}
+				\right),
+\end{equation}
+\end{linenomath*}
+
+\noindent where $f$ and $f^{\prime}$ represent \textcolor{blue}{maximum adult fertilities}, and $w_{i}$ and $w^{\prime}_{i}$ the genotypic relative scaling factors for female and male sex-functions (see Table \ref{tab:Fitness}).
+
+\begin{table}[htbp]
+ \centering
+ \caption{\bf Relative fertilities for Sexually Antagonistic selection ($w_{i}$)}
+\label{tab:Fitness}
+\begin{tabular}{lccc}
+ \toprule
+					&  \multicolumn{3}{c}{{\textit{Genotype}}} \\ 
+\cline{2-4}
+					& $AA$			& $Aa$ 					& $aa$ 		\\ \hline
+Female function ($w_{i}$):	& $1$		& $1 - h_f s_f$	& $1 - s_f$ \\	
+Male function ($w^{\prime}_{i}$):		& $1 - s_m$& $1 - h_m s_m$	& $1$ 		\\	
+\hline
+\end{tabular}
+\end{table}
+
+\subsubsection*{Mating and offspring production under partial selfing}
+ Outcrossing in our model proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}.
+That is, each individual's genotype determines both the number of ovules produced and the paternal mating success, broadly defined.
+For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986, WangBarrett2020, Harder2016}.
+Note, however, that by modeling paternal relative mating success rather than pollen/sperm production, we implicitly assume female demographic dominance (i.e., production/transport of male gametes does not limit ovule fertilization and hence population growth), a point we return to in the Discussion.
+
+\textcolor{red}{Lotte, I've reworded this to avoid dioecious language (males/females) and to emphasize that what really matters is that outcross matings are random, and selfing rates are genotype-independent...
+please check to make sure you agree with my edits} \textcolor{blue}{ For the purpose of this article, we assume that all outcross matings are random.
+The model can be extended to consider various forms of assortative mating beyond genotype-dependent selfing (see Online Supplement).
+In addition, we assume that individuals produced by selfing or by outcrossing do not differ in their mating success, though they can differ in their surival (see below).
+Given these assumptions, the allele frequencies in the male mating population can obtained by simply summing the vectors of individuals produced through selfing and outcrossing:
+\begin{linenomath*}
+\begin{equation}
+ 	\mbf{p}^{\prime}=\frac{(\mbf{n}^X + \mbf{n}^S)}{\| (\mbf{n}^X + \mbf{n}^S)\|}.
+ \end{equation}
+\end{linenomath*}}
+
+\textcolor{blue}{Individuals contribute  gametes to the population outcrossing gamete pool according to their fertility, which depends on their stage and genotype at the SA locus.
+The allele frequencies in the male gamete pool are then calculated as follows (see Online Supplementary Materials for a derivation of this equation),
+\begin{linenomath*}
+\begin{equation} \label{eq:maleGametePool}
+	\left(
+		\begin{array}{c}
+			q^{\prime}_{A} \\
+			q^{\prime}_{a} \\
+		\end{array} \right) \propto 
+				\left(
+		\begin{array}{c}
+			\mbf{1}^{\intercal}_{\omega}\mbf{F}^{\prime}_{AA}\mbf{p}^{\prime}_{AA}+ \frac{1}{2} \mbf{1}^{\intercal}_{\omega}  \mbf{F}^{\prime}_{Aa} \mbf{p}^{\prime}_{Aa}  \\
+		\frac{1}{2} \mbf{1}^{\intercal}_{\omega}  \mbf{F}^{\prime}_{Aa} \mbf{p}^{\prime}_{Aa} + \mbf{1}^{\intercal}_{\omega}\mbf{F}^{\prime}_{aa}\mbf{p}^{\prime}_{aa}\\
+		\end{array} \right),
+\end{equation}
+\end{linenomath*}
+where equality is obtained by normalizing the frequencies such that they add to one.
+The first entry sums the the stage-specific contributions to the gamete pool of heterozygotes and $AA$ homozygotes; the second entry sums the stage-specific contributions to the gamete pool of heterozygotes and $aa$ homozygotes.
+}
+
+\textcolor{blue}{ We now have all the ingredients to calculate the allele frequencies in the male gamete pool for the two-stage model in Eq(\ref{eq:F_male}), Eq(\ref{eq:maleGametePool}), and Table \ref{tab:Fitness}.
+Denoting the proportion of adults of genotype $i$ with $p^A_i$, the allele frequencies in the male gamete pool are given by
+\begin{linenomath*}
+\begin{equation} \label{eq:maleGametePool_example}
+\left(
+		\begin{array}{c}
+			q^{\prime}_{A} \\
+			q^{\prime}_{a} \\
+		\end{array} \right) = 
+				\frac{1}{(1-s_m)p^A_{aa}+ (1-hs_m)p^A_{Aa}+p_{aa}}\left(
+		\begin{array}{c}
+(1-s_m)p^A_{AA}	+\frac{1}{2}(1-hs_m)p^A_{Aa}\\
+p^A_{aa}+\frac{1}{2}(1-hs_m)p^A_{Aa}
+		\end{array} \right).
+\end{equation}
+\end{linenomath*}
+Note that this equation is identical to the classical population genetic results for frequencies of male gametes in a sexually antagonistic model, with the caveat that only reproductively active adults contribute to the gamete pool.
+Eq(\ref{eq:maleGametePool_example}) increasingly deviates from the classical population genetics result if additional reproductive adult age or stage classes are included (see Eq(\ref{eq:maleGametePool})).
+}
+
+\subsubsection*{Population projection}
+Using the component matrices described above (the survival matrices, $\mathbf{U}^X_i$, $\mathbf{U}^S_i$, the fertility and mating success matrices, $\mathbf{F}_i$, $\mathbf{F}^\prime_i$) and the allele frequencies in the gamete pool, we can construct the population projection matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$.
+The resulting matrix that projects the eco-evolutionary dynamics is:
+\begin{linenomath*}
+\begin{equation} \label{eq:Atilde}
+	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
+			\underbrace{\left(
+			\begin{array}{c|c}
+				\mcal{U}^S & \mbf{0} \\ \hline
+				\mbf{0} & \mcal{U}^X \\
+			\end{array} \right)}_{\tilde{\mbf{U}}} + 
+			\underbrace{\left(
+			\begin{array}{c|c}
+				\mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
+				\mcal{F}^X(\tilde{\mbf{p}}) (1 - C) & \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)\\
+			\end{array} \right)}_{\tilde{\mbf{F}}},
+\end{equation}
+\end{linenomath*}
+where $C$ denotes the proportion of each individual's ovules that are self-fertilized (the remaining $1 - C$ are outcrossed), and  $\delta$ represents the proportion of self-fertilized zygotes that fail to develop due to inbreeding depression during early development \citep{Charlesworth1987}.
+
+The blocks of the component matrices in Eq(\ref{eq:Atilde}) correspond to production of offspring by self-fertilization and outcrossing ($\mcal{F}^S$ and $\mcal{F}^X$ in $\tilde{\mbf{F}}$), and survival of extant individuals produced by selfing or outcrossing ($\mcal{U}^S$ and $\mcal{U}^X$ in $\tilde{\mbf{U}}$).
+The survival matrices for individuals produced through selfing and outcrossing are,
+\begin{linenomath*}
+\begin{eqnarray} 
+	\mcal{U}^S  = 
+		\left(
+			\begin{array}{ccc}
+				\mbf{U}^{S}_{AA} & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{U}^{S}_{Aa} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \mbf{U}^{S}_{aa} \\
+			\end{array} \right),\label{eq:BlkUS}\\
+				\mcal{U}^X  = 
+		\left(
+			\begin{array}{ccc}
+				\mbf{U}^{X}_{AA} & \mbf{0} & \mbf{0} \\
+				\mbf{0} & \mbf{U}^{X}_{Aa} & \mbf{0} \\
+				\mbf{0} & \mbf{0} & \mbf{U}^{X}_{aa} \\
+			\end{array} \right).\label{eq:BlkUX}
+\end{eqnarray}
+\end{linenomath*}
+The survival matrices are block diagonal because individuals cannot change their genotype once they are born.
+
+The fertility matrices for individuals produced through selfing and outcrossing are given by, 
+\begin{linenomath*}
+\begin{equation} \label{eq:BlkFS}
+	\mcal{F}^S = 
+			\left(
+			\begin{array}{ccc}
+				\mbf{F}_{AA} & \frac{1}{4} \mbf{F}_{Aa} & \mbf{0}  \\
+				\mbf{0}  & \frac{1}{2} \mbf{F}_{Aa} & \mbf{0}  \\
+				\mbf{0} & \frac{1}{4} \mbf{F}_{Aa} & \mbf{F}_{aa}\\
+			\end{array} \right), 
+\end{equation}
+\end{linenomath*}
+\noindent and 
+\begin{linenomath*}
+\begin{equation} \label{eq:BlkFX}
+	\mcal{F}^X = 
+			\left(
+			\begin{array}{ccc}
+				q^{\prime}_{A} \mbf{F}_{AA} & \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} & \mbf{0}  \\ 
+				q^{\prime}_{a} \mbf{F}_{AA} & \frac{1}{2} \mbf{F}_{Aa} & q^{\prime}_{A} \mbf{F}_{aa}  \\ 
+				\mbf{0}  & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} & q^{\prime}_{a} \mbf{F}_{aa}  \\
+			\end{array} \right),
+\end{equation}
+\end{linenomath*}
+\noindent where $q^{\prime}_A$ and $q^{\prime}_a$ are the frequencies of alleles $A$ and $a$ in the gamete pool, given by Eq(\ref{eq:maleGametePool}).
+
+The blocks of $\mcal{F}^X(\tilde{\mbf{p}})$ can be constructed and interpreted as follows: The first row block of the first column produces $AA$ offspring by outcrossing from $AA$ maternal parents.
+This happens when the $AA$ maternal parent receives an $A$ gamete from the male gamete pool, which happens with probability $q^{\prime}_{A}$.
+The other blocks can be interpreted similarly.
+Combining all the component matrices yields the overall eco-evolutionary projection matrix shown in \ref{App:Projection}.
 
@@
 
-Diverse eco-evolutionary outcomes are possible in the model, including fixation of either allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and suicide (e.g., see \citealt{deVriesCaswell2019a,deVriesCaswell2019b}), see Figure 1 for an example of the possible outcomes in different regions of parameter space.
-We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection \textcolor{blue}{(see left-hand side of Figure \ref{fig:didactic})} and ({\itshape ii}) the intrinsic population growth rate at equilibrium is positive \textcolor{blue}{(see right-hand side of Figure \ref{fig:didactic})}; a situation that we refer to as a 'demographically viable SA polymorphism'.
+Diverse eco-evolutionary outcomes are possible in the model, including fixation of either allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and suicide (e.g., see Figure 1 of \citealt{deVriesCaswell2019a, deVriesCaswell2019b} for examples of the possible outcomes).
+We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection \textcolor{blue}{(Figure \ref{fig:didactic}, left-hand panel)} and ({\itshape ii}) the intrinsic population growth rate at equilibrium is positive \textcolor{blue}{(Figure \ref{fig:didactic}, right-hand panel)}; a situation that we refer to as a 'demographically viable SA polymorphism'.
 
@@
 The formal conditions for a protected polymorphism are determined by linearizing the model in the vicinity of the boundary equilibria ($\hat{\mbf{p}}$), and evaluating the magnitude of the largest eigenvalue of the Jacobian matrix of the linearization.
-A full derivation of the Jacobian and details of the invasion analysis are provided in the Online Supplementary Material, and the relevant leading eigenvalues are presented in Appendix B.
+A full derivation of the Jacobian and details of the invasion analysis are provided in the Online Supplementary Material, and the relevant leading eigenvalues are presented in \ref{App:Eigen}.
 
@@
 We used numerical simulation to determine whether a protected SA polymorphism was also demographically viable.
-Specifically, for each boundary equilibrium we introduced the rare allele at low initial frequency, and iterated Eq(\ref{eq:Atilde}) until the population had reached demographic and genotypic equilibrium.
+Specifically, for each boundary equilibrium we introduced the rare allele at low initial frequency and iterated Eq(\ref{eq:Atilde}) until the population had reached demographic and genotypic equilibrium.
 \textcolor{blue}{Although we use density-independent survival and fertility rates, the allele frequency dynamics introduce frequency-dependence into the model.
@@
 
-For analyses of the demographic effects of inbreeding depression, we make two additional simplifying assumptions: First, to keep our analyses tractable we explore the effects of individual inbreeding depression terms in isolation.
+To analyze the demographic effects of inbreeding depression, we make two additional simplifying assumptions: First, to keep our analyses tractable we explore the effects of individual inbreeding depression terms in isolation.
 That is, we assume that only one of the $\delta$ and $\delta_i$ terms (where $i \in \{j,a,\gamma\}$) can be non-zero at a time.
 Second, we assume that if inbreeding depression is caused primarily by recessive deleterious mutations (as suggested by empirical data), it should covary negatively with the population selfing rate due to purging, provided the population selfing rate has been relatively constant in recent evolutionary time (\citealt{Charlesworth2009}; though we note that other processes could give rise to this pattern, e.g., \citealt{CrnokrakBarrett2002, Charlesworth2009,HedrickGarcia-Dorado2016}).
-Following \citet{Olito2019}, we incorporate such negative covariance by constraining the inbreeding depression terms in the model ($\delta$ and $\delta_i$, where $i \in \{j,a,\gamma\}$) to follow a simple declining function of the selfing rate: $\delta = \delta^{\ast} (1 - b (1 - L))$, where $\delta^{\ast}$ is the hypothetical severity of inbreeding depression for a completely outcrossing population, $b$ is a shape parameter determining how far $\delta$ will decline under complete selfing (when $C = 1$), and $L$ describes the expected deleterious mutation load as a function of the selfing rate $C$.
-The function $L$ includes an additional shape parameter, $a$, which determines the curvature of the overall function for $\delta$ (see Appendix E in \citealt{Olito2019} for additional details).
+Following \citet{Olito2019}, we incorporate such negative covariance by constraining the inbreeding depression terms in the model ($\delta$ and $\delta_i$, where $i \in \{j,a,\gamma\}$) to follow a simple declining function of the selfing rate: \textcolor{blue}{$\delta = \delta^{\ast} (1 - b (1 - a(1 - C)/(C + a(1 - C))))$}, where $\delta^{\ast}$ is the hypothetical severity of inbreeding depression for a completely outcrossing population, $b$ is a shape parameter determining how far $\delta$ will decline under complete selfing (when $C = 1$), \textcolor{blue}{and $a$ is a shape parameter} controlling the curvature of the overall function for $\delta$ (see Appendix E in \citealt{Olito2019} for additional details).
 We set $\delta^{\ast} = 0.8$, $b = 0.5$, and $a = 0.2$ for all analyses, values chosen to be consistent with empirical estimates of inbreeding depression (e.g., fig.~2 in \citealt{HusbandSchemske1996}).
@@
 "Demographically viable polymorphic parameter space" corresponds to the area inside the funnel that is also to the left of the extinction threshold for a given fertility value.
-Results are shown for three different population selfing rates ($C = \{0,\,1/4,\,1/2\}$), and two dominance scenarios (additivity, where $h = 1/2$, and dominance reversal, where $h = 1/4$); extinction thresholds are illustrated for three different values of female fecundity ($f$ values annotated on each panel).}}
+Results are shown for three different population selfing rates ($C = \{0,\,1/4,\,1/2\}$), and two dominance scenarios (additivity, where $h = 1/2$, and dominance reversal, where $h = 1/4$); extinction thresholds are illustrated for three different values of female fecundity ($f$ values annotated on each panel).
+\textcolor{blue}{Predicted population intrinsic growth rates ($\lambda$) for the same parameter conditions are presented in fig.~S1, in the Online Supplementary Material.}}}
  \label{fig:extThresholds}
@@
 We begin with an illustration of demographically viable polymorphic parameter space in the absence of inbreeding depression in Figure \ref{fig:extThresholds} (i.e., $\delta = \delta_i = 0$).
-The red regions in Figure \ref{fig:extThresholds} indicate areas where the population goes extinct (population growth rate is less than one) for different values of the fertility parameter.
+The red regions in Figure \ref{fig:extThresholds} indicate areas where the population goes extinct (\textcolor{blue}{i.e., where $\lambda < 1$}) for different values of the fertility parameter.
 Lower female fertility values correspond to a larger demographically inviable area.
-The funnel between the thick black lines indicates the area where the sexually antagostic alleles coexist.
-Above the funnel, the female-benefit and male-detriment allele goes to fixation; below the funnel, the male-benefit and female-detriment allele \textcolor{blue}{goes to fixation}.
-Much of the region where the male-benefit and female-detriment allele \textcolor{blue}{fixes} is demographically inviable due to the demographic consequences of reduced female fitness in this region.
+The funnel between the thick black lines indicates the area where the sexually antagonistic alleles coexist.
+Above the funnel, the female-benefit/male-detriment allele goes to fixation; below the funnel, the male-benefit/female-detriment allele \textcolor{blue}{goes to fixation}.
+Much of the region where the male-benefit/female-detriment allele \textcolor{blue}{fixes} is demographically inviable due to the demographic consequences of reduced female fertility in this region.
 The SA polymorphisms that remain viable at lower fertility values correspond to regions where the female-deleterious allele is predicted to segregate at low frequencies, or in regions with weak selection through both sex functions, an asymmetry that reflects the assumption of female demographic dominance.
@@
 Moreover, detailed demographic studies have been conducted on multiple populations of {\itshape M.~guttatus}, with demographic data available on the public demographic database COMPADRE \citep{CompadreDB2020}.
-Below, we briefly outline how we parameterized our model using the available data; full details are provided in Appendix C.
+Below, we briefly outline how we parameterized our model using the available data; full details are provided in \ref{App:Mimulus}.
 
 \subsection*{Demographic and fitness data for {\itshape M. guttatus}} \label{subsec:MguttMethods}
+
 To parameterize our model we utilize extensive demographic data reported in a large-scale study of local adaptation using experimental populations of {\itshape M.~guttatus} in Stanislaus National Forest (California, USA) in 2012 and 2013 \citep{PetersonEtAl2016}.
@@
 The same estimates for $D$ and $A$ were used for all populations.
-The resulting transition matrices for this population involved three life-history stages ($\omega = 3$; seed, seedling, and rosette), and individual elements of the transition matrix ($\tilde{\mbf{A}}$) were calculated as products of the above rates (see Matrix $1$ in \citealt{PetersonEtAl2016}, also Eq(\ref{eq:AtildeMimulus}) in our Appendix C).
+The resulting transition matrices for this population involved three life-history stages ($\omega = 3$; seed, seedling, and rosette), and individual elements of the transition matrix ($\tilde{\mbf{A}}$) were calculated as products of the above rates (see Matrix $1$ in \citealt{PetersonEtAl2016}, also Eq(\ref{eq:AtildeMimulus}) in our \ref{App:Mimulus}).
 
@@
 Using these combined demographic rates, selfing rates, and inbreeding depression estimates, we constructed a corresponding stage $\times$ genotype mendelian matrix model with a single SA locus affecting female and male fertility (as described above).
-With the empirically parameterized model, we are able to make predictions about the genetic and demographic outcome of SA selection a single locus in hypothetical populations with the same demographic rates as observed in \citet{PetersonEtAl2016}, for a range of selfing and inbreeding depression rates observed in other natural populations.
+With the empirically parameterized model, we are able to make predictions about the genetic and demographic outcome of SA selection at a single locus in hypothetical populations with the same demographic rates as observed in \citet{PetersonEtAl2016}, for a range of selfing and inbreeding depression rates observed in other natural populations.
 We stress, however, that these are illustrative rather than explicit predictions of the likelihood of SA polymorphism in any specific population, and they ignore measurement error for the estimated demographic rates.
@@
 inv6 segregates at moderate frequency (about $8\%$), and carriers suffer an approximately $30 \%$ loss in pollen viability, but also increased flower (and therefore ovule and pollen) production.
-The genetic basis of these effects remain unclear -- though they likely polygenic -- but the net result is a "supergene" with remarkably strong effects on both female and male fertility that segregates as a single diallelic locus.
-We estimated selection coefficients for the effect of inv6 on pollen production (i.e., taking into account the simultaneous effect on flower number) and ovule production from the data reported in \citet{LeeKelly2015} under the relatively conservative assumptions that pollen/ovule production is proportional to flower number, and  additive SA fitness effects (dominance coefficients could not be estimated from field data; see Appendix C for details of estimating selection coefficients).
+The genetic basis of these effects remain unclear -- though they likely polygenic -- but the net result is a "supergene" with remarkably strong effects on both female and male fertility that segregates as a single diallelic locus \textcolor{blue}{(i.e., with two alleles, wild-type and inverted)}.
+We estimated selection coefficients for the effect of inv6 on pollen production (i.e., taking into account the simultaneous effect on flower number) and ovule production from the data reported in \citet{LeeKelly2015} under the relatively conservative assumptions that pollen/ovule production is proportional to flower number, and  additive SA fitness effects (dominance coefficients could not be estimated from field data; see \ref{App:Mimulus} for details of estimating selection coefficients).
 Under these assumptions, the average selection coefficient for inv6 on flower production across two years in the field was $(s_f,s_m) = (0.3,0.31)$.
@@
  \caption{\footnotesize{Illustration of model predictions using empirically estimated demographic rates for {\itshape M.~guttatus}.
-Results are shown for two hypothetical populations using demographic rates for locally adapted (Eagle Meadows; panel A), and non-local (Low-Elevation Perennial; panel B) populations reported by \citet{PetersonEtAl2016}.
-Invasion conditions and extinction thresholds were calculated using the conservative assumption of additive fitness effects in both sexes ($h_f = h_m = 0.5$) for two parameter conditions: obligate outcrossing (solid black lines, dark red shaded region)  and partial selfing with inbreeding depression using the highest field-estimate of selfing, and inbreeding depression parameters ($\delta_i$) calculated for the Iron Mountain population of \citet{Willis1993} (dashed lines, light red shaded region).
-The location of inv6 is also shown on both plots, using selection coefficients calculated from field estimates of male and female fitness components from \citet{LeeKelly2015} under the relatively conservative assumption of additive fitness effects in both sexes, where $(s_f,\,s_m) = (0.30,\, 0.31)$.
-Note that in panel A the invasion conditions for partial selfing with inbreeding depression (dotted lines) are nearly indistinguishable from those for obligate outcrossing (solid lines).}}
+Results are shown for two hypothetical populations using demographic rates for locally adapted (Eagle Meadows; panels A,C), and non-local (Low-Elevation Perennial; panels B,D) populations reported by \citet{PetersonEtAl2016}.
+Invasion conditions and population intrinsic growth rates ($\lambda$) were calculated using the conservative assumption of additive fitness effects in both sexes ($h_f = h_m = 0.5$) for two parameter conditions: obligate outcrossing (panels A,B) and partial selfing with inbreeding depression using the highest field-estimate of selfing and inbreeding depression parameters ($\delta_i$) calculated for the Iron Mountain population of \citet{Willis1993} (panels C,D).
+The location of inv6 is also shown on both plots, using selection coefficients calculated from field estimates of male and female fitness components from \citet{LeeKelly2015} under the relatively conservative assumption of additive fitness effects in both sexes, where $(s_f,\,s_m) = (0.30,\, 0.31)$.}}
  \label{fig:mimulusFig}
@@
 The Eagle Meadows population was locally adapted with a very high intrinsic growth rate ($\lambda \approx 1.7$).
-This growth rate was sufficiently high that all $s_f \times s_m$ selection parameter space (where $s_f, s_m \in (0,1]$) remained demographically viable, regardless of the selfing rate and effects of inbreeding depression ($C = 0$ or $0.29$; fig.~\ref{fig:mimulusFig}A).
+This growth rate was sufficiently high that all $s_f \times s_m$ selection parameter space (where $s_f, s_m \in (0,1]$) remained demographically viable, regardless of the selfing rate and effects of inbreeding depression ($C = 0$ or $0.29$; fig.~\ref{fig:mimulusFig}A,C).
 In contrast, the Low-Elevation Perennial population had a much lower, but still positive, intrinsic growth rate ($\lambda \approx 1.08$).
-Due to the slower growth rate, not all of the $s_f \times s_m$ selection parameter space was demographically viable: extinction thresholds appear under both complete outcrossing and partial selfing with inbreeding depression (fig.~\ref{fig:mimulusFig}B).
-The different demographic rates from these two populations also resulted in slightly different invasion conditions when inbreeding depression was taken into account (compare dashed with solid lines in fig.~\ref{fig:mimulusFig}A and B).
-
-When using demographic rates for either the Eagle Meadows or Low-Elevation Perennial populations, inv6 falls within demographically viable polymorphic parameter space (fig.~\ref{fig:mimulusFig}).
-Whether under obligate outcrossing or partial selfing, inv6 always falls squarely in the middle of SA polymorphic space when using the locally adapted Eagle Meadows demographic rates.
-In contrast, when the selfing rate is at the higher end of empirical estimates for the Iron Mountain population in which inv6 has been documented ($C = 0.29$), inv6 falls nearer the upper boundary but still within SA polymorhic space when using the Low-Elevation Perennial demographic rates.
-This happens because under the empirical estimates of selfing and inbreeding depression, the polymorphic parameter space shifts downwards.
+Due to the slower growth rate, not all of the $s_f \times s_m$ selection parameter space was demographically viable: extinction thresholds appear under both complete outcrossing and partial selfing with inbreeding depression (fig.~\ref{fig:mimulusFig}B,D).
+
+Whether under obligate outcrossing or partial selfing, inv6 always falls squarely in the middle of SA polymorphic space when using the locally adapted Eagle Meadows demographic rates (fig.~\ref{fig:mimulusFig}A,C).
+In contrast, when the selfing rate is at the higher end of empirical estimates for the Iron Mountain population in which inv6 has been documented ($C = 0.29$), inv6 falls nearer the upper boundary but still within SA polymorhic space when using the Low-Elevation Perennial demographic rates (fig.~\ref{fig:mimulusFig}B,D).
+This happens because under the empirical estimates of selfing and inbreeding depression, the polymorphic parameter space shifts downwards \textcolor{blue}{towards the x-axis}.
 Additionally, inv6 falls much closer to the extinction threshold under partial selfing, suggesting that even relatively small perturbations to demographic rates or selection coefficients could result in non-locally adapted populations being unable to support the demographic costs associated with segregating SA alleles with selection coefficients of similar magnitude to inv6.
@@
 
-The first and central finding of our study is that when intrinsic population growth rates approach one, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space where SA polymorphism is maintained.
+The first and central finding of our study is that when population intrinsic growth rates approach one, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space where SA polymorphism is maintained.
 Since intrinsic growth rates far exceeding one suggest rapid exponential growth, they are generally rare (with the notable exception of recently introduced, invasive populations), suggesting that our model predictions may be highly relevant for SA polymorphism in many real-world populations.
@@
 This prediction is in stark contrast to previous population genetics models of SA selection in hermaphrodites, where the sole effect of inbreeding depression is to reduce the population effective selfing rate through the loss of selfed-zygotes, thereby expanding polymorphic parameter space \citep{JordanConnallon2014, Olito2017}.
-In our model, this reduction in effective selfing is accompanied by significant mortality due to inbreeding depression (whether early- or late-acting), which can quickly tip partially-selfing populations over the brink to extinction.
+In our model, this reduction in the effective selfing rate is accompanied by significant mortality due to inbreeding depression (whether early- or late-acting), which can quickly tip partially-selfing populations over the brink to extinction.
 Beyond the maintenance of SA polymorphisms, this finding underscores a simple but important point that is often overlooked in studies of the evolution of self-fertilization and selfing-syndromes, which tend to emphasize the coevolution of the deleterious mutation load and mating system \citep[e.g.,][]{Charlesworth1987, LandeSchemske1985, Goodwillie2005}: highly fertile populations can better afford the severe demographic costs of inbreeding.
@@
 
-Despite the demographic pitfalls associated with SA alleles, our example using {\itshape M.~guttatus} appears to show that demographic rates observed in some real populations are capable of sustaining large regions of viable SA polymorphic space.
+Despite the demographic pitfalls associated with SA alleles, our case study using {\itshape M.~guttatus} appears to show that demographic rates observed in some real populations are capable of sustaining large regions of viable SA polymorphic space.
 The example also appears to provide some empirical support for the conjecure that locally-adapted populations are more likely to harbor SA polymorphisms than marginal or non-locally adapted ones; inviable polymorphic parameter space only occured when using the demographic data for non-local low-elevation perennial populations.
@@
 
-\subsubsection*{Extensions and future directions}
+\subsubsection*{Extensions and future directions}\label{subsubsec:extensions}
 
@@
 What if the allele not only affects male and female fertility but also affects survival of individuals?
-Additionally, how would the scope for SA polymorphism be affected by density-dependence acting on different life history stages?
+
+\textcolor{blue}{One particularly important limitation of our results is that we modeled density-independent growth.
+Of course, in natural populations density-dependent processes like resource competition or competition for breeding sites will often prevent populations from growing (or shrinking) exponentially.
+In \ref{App:DensDep} we extend our model to include a simple form of density-dependence that acts on the survival of all life history stages equally, and reanalyse our case study of \textit{M.
+guttatus}.
+The density-dependent version of the model predicts equilibrium population densities rather than population intrinsic growth rates (see fig.~ \ref{fig:mimulusFig_dd}).
+This brief example shows that the shape of the polymorphic region is unchanged by density-dependence, and that the shape of the population density surface in fig.~\ref{fig:mimulusFig_dd} is very similar that for population growth rate in fig.~\ref{fig:mimulusFig}.
+The similarity between the model predictions underscores an important point: density-dependence that acts age-independently does not alter the selection gradients (see also \citealt{mylius1995evolutionarily, caswell2017}).}
+
+\textcolor{blue}{In many natural populations, however, density-dependence does not have the same effect on the survival of individuals of all ages or developmental stages.
+For example, competition for breeding sites will cause a reduction in fecundity rather than survival.
+Likewise, in plants density-dependent competition for soil nutrients will likely influence seed germination rates differently than seedling growth rates or flower/fruit production in adults (CITATION).
+These types of age or stage specific density-dependence will change the selection gradients and therefore the shape of polymorphic parameter space as well as the equilibrium population densities predicted by the model.
+A thorough exploration of these more complicated forms of density dependence remains an important avenue for future work.}
 
@@
 
-Another major simplifying assumption in our model was that how individuals themselves were produced (i.e., by selfing or outcrossing) affects the level of inbreeding depression they suffer.
+\textcolor{blue}{Our assumption that how individuals themselves were produced (i.e., by selfing or outcrossing) affects the level of inbreeding depression they suffer also represents a major simplification}.
 In reality, the history of consecutive generations of inbreeding in each individual's lineage will influence the severity of inbreeding depression they experience, particularly when inbreeding depression is caused primarily by recessive deleterious mutations.
 It would be an interesting and feasible extension of our modeling framework to expand the individual state space to include selfing cohorts (i.e., first generation selfing, second generation selfing, etc.), as in the models of \citet{kelly1999response,kelly2007mutation}, thereby enabling a more biologically plausible approach to modelling self-fertilization.
-
-\cite{grieshop2017male} found that genotypes with sexually antagonstic alleles that are male beneficial experience higher levels of inbreeding depression than genotypes with female beneficial SA alleles.
+\textcolor{blue}{Likewise, we implemented a simple "fixed selfing" model, where all individuals in the population self-fertilize at the same rate.
+In reality, individuals will vary in rate of self-fertilization, which will influence both the expression of inbreeding depression and the scope for SA polymorphism \citep{JordanConnallon2014}.
+A naive expectation might be that among-individual variation in selfing might reduce the effective selfing rate in a population, expanding the scope for balanced SA polymorphisms, as it does in genotype-dependent mass-action selfing models \citep{JordanConnallon2014}.
+However, it is difficult to speculate how the associated demographic effects of such variation via reproductive assurance and inbreeding depression  would play out.}
+
+\cite{grieshop2017male} found that genotypes with sexually antagonistic alleles that are male-beneficial experience higher levels of inbreeding depression than genotypes with female beneficial SA alleles.
 This effect can easily be included in our model by making the inbreeding depression parameters, $\delta$ and $\delta_i$, a function of the genotype of the individual.
@@
 
-\subsubsection*{Conclusion}
+\subsubsection*{Conclusions}
 
@@
 Demographic models connect individual level traits to population level consequences.
-As a consequence of their focus on individuals, demographic models are ideal for linking to experimental or field data, as demonstrated with our \textit{Mimulus} case study.
+As a consequence of their focus on individuals, demographic models are ideal for linking theory to experimental or field data, as demonstrated with our \textit{Mimulus} case study.
 Doing so allows the field of population genetics to move from fitness as an abstract scalar metric towards the fitness of an entire life cycle as calculated from observed rates of age- or stage-specific survival and fecundity rates.
@@
 % formatting peculiarities that \appendix conflicts with.
+\renewcommand\thesection{Appendix~\Alph{section}}
+\renewcommand\thesubsection{\Alph{section}.\arabic{subsection}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Appendix A: Population projection matrix}
+\section{Population projection matrix}\label{App:Projection}
 \renewcommand{\theequation}{A\arabic{equation}}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Appendix B: Eigenvalues for invasion analysis}
+\section{Example with density-dependence}\label{App:DensDep}
 \renewcommand{\theequation}{B\arabic{equation}}
+\renewcommand{\thetable}{B\arabic{table}}
+\renewcommand{\thefigure}{B\arabic{figure}}
+\setcounter{figure}{0}  % reset counter 
+\setcounter{equation}{0}  % reset counter 
+\setcounter{table}{0}  % reset counter
+
+\textcolor{blue}{Here we extend the model presented in the main text by incorporating density-dependent population regulation.
+We chose the simplest population regulation possible in a structured population model: population regulation affecting all life history stages in the same way.
+Note that this is the only type of population regulation that is possible in unstructured population models, like the classical Lotka-Volterra or Rosenzweig-MacArthur models.
+We model survival of all life-history stages as a exponentially decreasing function of population density, such that the population projection becomes
+\begin{linenomath*}
+\begin{equation}
+	\tilde{\mbf{n}}(t + 1) = \exp\left(-\alpha \sum_i n_i \right) \tilde{\mbf{A}}[\tilde{\mbf{n}}(t)] \, \tilde{\mbf{n}}(t),
+\end{equation}
+\end{linenomath*}
+where $\sum_i n_i$ sums over all entries of the population state vector, and hence over all genotypes in all life-history stages. Equivalently, in matrix algebra notation,
+\begin{linenomath*}
+\begin{equation}
+	\tilde{\mbf{n}}(t + 1) = \exp\left(-\alpha \| \tilde{\mbf{n}}(t) \|\right) \tilde{\mbf{A}}[\tilde{\mbf{n}}(t)] \, \tilde{\mbf{n}}(t),
+\end{equation}
+\end{linenomath*}
+where $\| \tilde{\mbf{n}}(t) \|$ is the one-norm which is identical to the sum of the entries of $\tilde{\mbf{n}}(t)$ because the population vector only has nonnegative entries. }
+
+\textcolor{blue}{Using this density-dependent version of the model, we reanalyzed the {\itshape M.
+guttatus} example from the main text.
+Fig.~\ref{fig:mimulusFig_dd} has the same structure as fig.~\ref{fig:mimulusFig}, but now shows the equilibrium population size, $N^{\ast} = \sum_i n_i $, rather than the population growth rate.
+}
+
+\textcolor{blue}{With this simplest form of density-dependence, our our main conclusions remain intact.
+By introducing a density-dependent factor acting on survival of all life-history stages in the same way, we did not change any of the selection gradients.
+That is, the density-dependent factor essentially scales the entire population projection matrix, and so it does not change the stable stage distribution vector, or the normalized reproductive value vector.
+Adding this simple form of density-dependence does not change the relative importance of any of the life history stages, nor any of the demographic rates in the population projection matrix.
+}
+
+\textcolor{blue}{Of course, many other more realistic (and more interesting) density-dependent scenarios are possible that will change the results presented in fig.~\ref{fig:mimulusFig_dd} (see Discussion).
+In these more complicated scenarios, we would expect significant qualitative as well as quantitative changes to our model predictions.
+}
+
+\begin{figure}[htbp]
+ \centering
+ \includegraphics[scale=0.8]{../output/figs/MimulusInv6NstarFig2.pdf}
+ \caption{\footnotesize{Illustration of density-dependent model predictions using empirically estimated demographic rates for {\itshape M.~guttatus}.
+Results are shown for two hypothetical populations using demographic rates for locally adapted (Eagle Meadows; panels A,C), and non-local (Low-Elevation Perennial; panels B,D) populations reported by \citet{PetersonEtAl2016}.
+Invasion conditions and equilibrium densities ($N^{\ast}$) were calculated using the conservative assumption of additive fitness effects in both sexes ($h_f = h_m = 0.5$) for two parameter conditions: obligate outcrossing (panels A,B) and partial selfing with inbreeding depression using the highest field-estimate of selfing and inbreeding depression parameters ($\delta_i$) calculated for the Iron Mountain population of \citet{Willis1993} (panels C,D).
+The location of inv6 is also shown on both plots, using selection coefficients calculated from field estimates of male and female fitness components from \citet{LeeKelly2015} under the relatively conservative assumption of additive fitness effects in both sexes, where $(s_f,\,s_m) = (0.30,\, 0.31)$.}}
+ \label{fig:mimulusFig_dd}
+ \end{figure}
+
+\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%
+% Appendix C
+%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Eigenvalues for invasion analysis}\label{App:Eigen}
+\renewcommand{\theequation}{C\arabic{equation}}
 \setcounter{equation}{0}  % reset counter 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-% Appendix C
-%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Appendix C: Demographic and empirical data for {\itshape M.~guttatus} case study}
-\renewcommand{\theequation}{C\arabic{equation}}
-\renewcommand{\thetable}{C\arabic{table}}
+% Appendix D
+%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Demographic and empirical data for {\itshape M.~guttatus} case study}\label{App:Mimulus}
+\renewcommand{\theequation}{D\arabic{equation}}
+\renewcommand{\thetable}{D\arabic{table}}
 \setcounter{equation}{0}  % reset counter 
@@
 
-\subsection*{Demographic data}
+\subsection{Demographic data}
 
@@
 
-\subsection*{Selection on inv6}
+\subsection{Selection on inv6}
 

### 2021-12-20 doc/Supplements/SupplementaryMaterial.tex
Full edits first pass
This commit includes a first pass at all major edits/revisions to the
manuscript. Ready to hand it over.

--- 2021-06-04 doc/Supplements/SupplementaryMaterial.tex
@@
 \section{Construction of the population projection matrix}
-We will now derive the expressions for the survival and fertility matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ as given by Eq(10)-(13) in the main text.
+
+In this section we will go through the derivation of the matrix model incorporating multiple life-cycle stages and a single diallelic locus under SA selection on male and female fertility for populations of partially-selfing simultaneous hermaphrodites model presented in the main text.
+We start by calculating the frequencies of alleles in the gamete pool.
+These frequencies are the ingredients for the map of parent genotype to offspring genotype, which itself is a necessary ingredient to create the fertility matrices.
+Finally putting the fertility components together with the (slightly simpler) survival matrices, we obtain all the ingredients to build the population projection matrix.
+
+\subsubsection*{Mating and offspring production under partial selfing}
+ Outcrossing in our model proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}.
+That is, each individual's genotype determines both the number of ovules produced and the paternal mating success, broadly defined.
+For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986, WangBarrett2020, Harder2016}.
+Note, however, that by modeling paternal relative mating success rather than pollen/sperm production, we implicitly assume female demographic dominance (i.e., production/transport of male gametes does not limit ovule fertilization and hence population growth), a point we return to in the Discussion.
+For the purpose of this article, we also assume that mating is random with respect to stage.
+
+Because we assume that inbreeding depression only affects offspring viability, the allele frequency in the male mating population is obtained by simply summing the vectors of individuals produced through selfing and outcrossing:
+\begin{linenomath*}
+\begin{equation}
+ 	\mbf{p}^{\prime}=\frac{(\mbf{n}^X + \mbf{n}^S)}{\| (\mbf{n}^X + \mbf{n}^S)\|}.
+ \end{equation}
+\end{linenomath*}
+Depending on their stage and genotype at the SA locus, individuals contribute male gametes to the overall population gamete pool according to the following equation
+\begin{linenomath*}
+\begin{equation}
+		\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime}= 
+		\left(
+			\begin{array}{ccc}
+				\mbf{1}^{\intercal}_{\omega} & \frac{1}{2} \mbf{1}^{\intercal}_{\omega} & \mbf{0} \\
+				\mbf{0}  & \frac{1}{2} \mbf{1}^{\intercal}_{\omega} & \mbf{1}^{\intercal}_{\omega} \\
+			\end{array} \right)
+		\left(
+			\begin{array}{ccc}
+				\mbf{F}^{\prime}_{AA} & \mbf{0}  &\mbf{0}  \\
+				\mbf{0}  & \mbf{F}^{\prime}_{Aa} & \mbf{0}  \\
+				\mbf{0}  & \mbf{0}  & \mbf{F}^{\prime}_{aa} \\
+			\end{array} \right)
+		\left(
+			\begin{array}{c}
+				\mbf{p}^{\prime}_{AA} \\
+				\mbf{p}^{\prime}_{Aa} \\
+				\mbf{p}^{\prime}_{aa} \\
+			\end{array} \right),
+\end{equation}
+\end{linenomath*}
+
+\noindent where $\mbf{1}^{\intercal}_{\omega} $ is a vector of ones of size $1 \times \omega$.
+The matrix $\mbb{F}^{\prime}$ is composed of genotype-specific fertility matrices, and operates on the vector of genotype frequencies to give their relative contribution of each genotype to the gamete pool.
+The matrix $\mbf{W}^{\prime}$ converts these contributions to allele numbers.
+Normalizing the resulting vector gives the allele frequencies in the male gamete pool:
+\begin{linenomath*}
+\begin{equation} \label{eq:maleGametePool}
+	\left(
+		\begin{array}{c}
+			q^{\prime}_{A} \\
+			q^{\prime}_{a} \\
+		\end{array} \right) = 
+			\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime}}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p}^{\prime} \|} = 
+				\frac{\mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)}{\| \mbf{W}^{\prime} \mbb{F}^{\prime} (\mbf{n}^X + \mbf{n}^S)\|}.
+\end{equation}
+\end{linenomath*}
+The key difference between selfing and outcrossing reproduction can be seen in the parent-offspring maps, which reflect the joint effects of meiosis and mating on the distribution of offspring genotypes.
+% \textcolor{red}{
+% I'm doubting whether to move the next paragraph to the appendices as well.
+Equation (\ref{eq:maleGametePool}) is all we strictly need.
+If we keep the parent-offspring maps, we can explain in the next section how the matrices $\mcal{F}^S$ and $\mcal{F}^X$ have the same structure.
+Would that be helpful?
+Or would it be easier to just explain in the next section what the structure of $\mcal{F}^S$ and $\mcal{F}^X$ is?
+}
+
+The parent-offspring map for outcrossing is a function of the allele frequencies in the male gamete pool, and is given by 
+\begin{linenomath*}
+\begin{equation} \label{eq:HX}
+	\mbf{H}^X = 
+			\left(
+			\begin{array}{ccc}
+				q^{\prime}_{A} & \frac{1}{2} q^{\prime}_{A} & 0 \\
+				q^{\prime}_{a} & \frac{1}{2} & q^{\prime}_{A}  \\
+				0 & \frac{1}{2} q^{\prime}_{a} & q^{\prime}_{a} \\
+			\end{array} \right).
+\end{equation}
+\end{linenomath*}
+\noindent From left to right, the columns of matrix $\mbf{H}^X$ give the genotype distribution of outcrossed offspring produced by a maternal parent of each genotype ($AA$, $Aa$, and $aa$ respectively).
+
+The parent-offspring map for reproduction via self-fertilization differs from the outcrossing parent-offspring map.
+As in previous models of partial selfing, we assume that individuals produce enough male gametes to easily self-fertilize all of their ovules, and that self-fertilization involves little or no selection via the male sex function from external factors (e.g., pollinator visitation) relative to outcrossing (\citealt{Charlesworth1978a,JordanConnallon2014,Olito2017}, but see \citealt{Tazzyman2015}).
+Under these assumptions, the genotype distributions of selfed offspring are determined entirely by the parental genotype and the probabilities of segregation and fertilization during and after meiosis:
+\begin{linenomath*}
+\begin{equation} \label{eq:HS}
+	\mbf{H}^S(\tilde{\mbf{n}}) = 
+			\left(
+			\begin{array}{ccc}
+				1 & 1/4 & 0 \\
+				0 & 1/2 & 0 \\
+				0 & 1/4 & 1 \\
+			\end{array} \right).
+\end{equation}
+\end{linenomath*}
+\noindent The columns of $\mbf{H}^S(\tilde{\mbf{n}})$ again give the selfed offspring genotype distributions for parental genotypes of $AA$, $Aa$, and $aa$, respectively.
+As described in the next section, these parent-offspring maps influence the projection matrix by altering the fertility matrices for individuals produced by selfing and outcrossing (see Eq(\ref{eq:BlkFS_appendix}) and Eq(\ref{eq:BlkFX_appendix}) below).
+
+Following previous theory for partially selfing populations \cite[e.g.,][]{Charlesworth2010,JordanConnallon2014,Glemin2021}, we use a 'fixed prior selfing' model where all individuals are assumed to self-fertilize at a constant rate, $C$, prior to receiving outcross pollen, for reasons of analytic tractability.
+Further down in the Online Supplement, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).
+
+\subsubsection*{The Population Projection Matrix}
+We will now derive the expressions for the survival and fertility matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ as given by Eq(12)-(15) in the main text.
 The matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ are the ingredients of the population projection matrix, $\tilde{\mbf{A}}$, as given in Eq(9), repeated here for convenience,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Supplementary Figures}
+\renewcommand{\thefigure}{S\arabic{equation}}
+\setcounter{figure}{1}  % reset counter 
+\setcounter{equation}{1}  % reset counter 
+\setcounter{table}{1}  % reset counter
+
+%%%%%%%%%%%%%%%%%%%%%
+% Lambda Heatmap Version of Fig. 2
+ \begin{figure}[htbp]
+ \centering
+ \includegraphics[scale=0.45]{lambdaFig1_2}
+ \caption{\footnotesize{Illustration of parameter space for SA polymorphism and population intrinsic growth rates ($\lambda$) predicted by the model for the same combination of parameter values presented in fig.~2 in the main text.
+Balanced SA polymorphisms can be maintained in the funnel-shaped region between the invasion conditions for each SA allele (dark solid lines).
+However, for some parameter conditions, population growth rates are less than one, and will ultimately go extinct (yellow-red shaded regions) due to reduced female fitness resulting from the male-beneficial/female-deleterious allele that is either segregating as a balanced polymorphism (inside the funnel), or becomes fixed (area below the funnel).
+Results are shown for three different population selfing rates ($C = \{0,\,1/4,\,1/2\}$), and two dominance scenarios (additivity, where $h = 1/2$, and dominance reversal, where $h = 1/4$); extinction thresholds are illustrated for three different values of female fecundity ($f$ values annotated on each panel).}}
+ \label{fig:LambdaOverviewFig}
+ \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
 \newpage{}
@@
 \renewcommand{\thefigure}{S\arabic{equation}}
-\setcounter{figure}{1}  % reset counter 
 \setcounter{equation}{1}  % reset counter 

### 2021-12-20 R/functions-Figs.R

@@ -502,8 +502,8 @@ lambdaPolySpaceFig  <-  function() {
     # Import data sets
     invA        <-  read.csv("./output/simData/invasionBoundaries_sMax0.15_res0.0015_hf0.5_hm0.5_C0_delta0_dj0_da0_dg0_f5.8.csv", head=TRUE)
     heatA_f5.8  <-  read.csv("./output/simData/lambdaHeatMapData_sMax0.15_len100_alpha0_hf0.5_hm0.5_C0_delta0_dj0_da0_dg0_f5.8.csv", head=TRUE)
-    heatA_f6.0  <-  read.csv("./output/simData/lambdaHeatMapData_sMax0.15_len100_hf0.5_hm0.5_C0_delta0_dj0_da0_dg0_f6.csv", head=TRUE)
-    heatA_f6.2  <-  read.csv("./output/simData/lambdaHeatMapData_sMax0.15_len100_hf0.5_hm0.5_C0_delta0_dj0_da0_dg0_f6.2.csv", head=TRUE)
+    heatA_f6.0  <-  read.csv("./output/simData/lambdaHeatMapData_sMax0.15_len100_alpha0_hf0.5_hm0.5_C0_delta0_dj0_da0_dg0_f6.csv", head=TRUE)
+    heatA_f6.2  <-  read.csv("./output/simData/lambdaHeatMapData_sMax0.15_len100_alpha0_hf0.5_hm0.5_C0_delta0_dj0_da0_dg0_f6.2.csv", head=TRUE)
 
     invB        <-  read.csv("./output/simData/invasionBoundaries_sMax0.15_res0.0015_hf0.5_hm0.5_C0.25_delta0_dj0_da0_dg0_f5.8.csv", head=TRUE)
     heatB_f5.8  <-  read.csv("./output/simData/lambdaHeatMapData_sMax0.15_len100_hf0.5_hm0.5_C0.25_delta0_dj0_da0_dg0_f5.8.csv", head=TRUE)

### 2021-12-20 makeFigs.R

@@ -71,8 +71,8 @@ embed_fonts(figPath(name='MimulusInv6LambdaFig.pdf'))
 # DENSITY DEPENDENT DYNAMICS
 # Mimulus Case study: Nstar heatmap w/ Inv6
 toPdf(MimulusInv6NstarFig(), 
-			figPath(name='MimulusInv6NstarFig.pdf'), width=8, height=8)
-embed_fonts(figPath(name='MimulusInv6NstarFig.pdf'))
+			figPath(name='MimulusInv6NstarFig2.pdf'), width=8, height=8)
+embed_fonts(figPath(name='MimulusInv6NstarFig2.pdf'))
 
 
 

### 2021-12-20 run-Simulations.R

@@ -753,7 +753,7 @@ delta.list  <-  list(delta_D = 0,
 					 delta_O = 0,
 					 delta_S = 0
 					 )
-alpha  <-  10^-3
+alpha  <-  1e-4
 
 ## Obligate oucrossing
 makeLambdaHeatMapMimulusData(sMax = 0.99, len = 100,

### 2021-12-23 doc/SA-Hermaphrodites-wDemography-R1.tex
minor edits

@@
 We focus on identifying parameter conditions under which SA polymorphism is maintained {\itshape and} the population growth rate remains positive.
-Additionally, we analyze the effects of inbreeding depression manifesting at different life-history stages and give an illustrative example of the potential for SA polymorphism in real populations using empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus}.
+\textcolor{blue}{A sexually antagonistic polymorphism is a polymorphism of two alleles with antagonistic effects on the two sexes, that is, one of the two coexisting allele confers a fitness benefit to females and reduces male fitness, and the other allele confers a fitness benefit to males and reduces female fitness.
+} Additionally, we analyze the effects of inbreeding depression manifesting at different life-history stages and give an illustrative example of the potential for SA polymorphism in real populations using empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus}.
 Our results show that when population intrinsic growth rates approach one, extinction occurs across large swathes of parameter space favoring SA polymorphism or the fixation of male-beneficial alleles, and that inbreeding depression is a significant problem for maintaining SA polymorphism in partially selfing populations.
@@
 To persist in the long term, a population must be able to adapt to its environment.
-Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, such as deleterious mutations \citep{Haldane1937}, \textcolor{blue}{rapid, highly unpredictable, or extreme changes} in environmental conditions \citep{Maynard-Smith1976, LandeShannon1996, OrrUnckless2008,Lyberger-etal-2021}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, and genetic constraints \citep{ConnallonHall2018, MatthewsConnallon2019}.
-Genetic constraints on adaptation can arise as a consequence of conflicting selection and gene-flow between different classes of individuals, or more generally, when \textcolor{blue}{there is negative correlation between different fitness components} \citep{CharlesworthHughes2000, ConnallonHall2018}.
-
-\textcolor{blue}{Genetic constraints on adaptation arising from conflicting selection and gene flow} have particularly interesting evolutionary consequences \citep{ConnallonDebarreLi2018}.
+Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, such as deleterious mutations \citep{Haldane1937}, \textcolor{blue}{rapid, unpredictable, or extreme changes} in environmental conditions \citep{Maynard-Smith1976, LandeShannon1996, OrrUnckless2008,Lyberger-etal-2021}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, and genetic constraints \citep{ConnallonHall2018, MatthewsConnallon2019}.
+Genetic constraints on adaptation can arise as a consequence of conflicting selection and gene-flow between different classes of individuals, or more generally, when \textcolor{blue}{there is a negative correlation between different fitness components} \citep{CharlesworthHughes2000, ConnallonHall2018}.
+
+\textcolor{blue}{Genetic constraints on adaptation arising from conflicting selection and gene flow between different classes of individuals} have particularly interesting evolutionary consequences \citep{ConnallonDebarreLi2018}.
 On the one hand, they provide an effective mechanism for the maintenance of genetic variation, which can increase a population's capacity for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
@@
 
-Sexually antagonistic selection (abbreviated SA hereafter), \textcolor{blue}{is an important form of conflicting selection that arises when beneficial alleles for one sex are deleterious when expressed in the other.
+Sexually antagonistic selection (abbreviated SA hereafter) \textcolor{blue}{is an important form of conflicting selection that arises when beneficial alleles for one sex are deleterious when expressed in the other.
 SA selection appears to be} a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and the evolution of sexual dimorphism and sex-related traits \citep{Kidwell1977, Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, ConnallonClark2012, BondurianskyChenoweth2009, Olito2019}.
@@
 In particular, theoretical models combining population genetics with density-dependent population growth \citep{roughgarden1971density} and stage-structured populations (consolidated in \citealt{Charlesworth_1994}) paved the way toward understanding the demographic changes accompanying allele frequency changes due to selection.
-Extensions of these theories to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and quantitative traits \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects} all highlight the sometimes counterintuitive demographic consequences of adaptive evolution that are missed by classical population genetics models (see \citealt{fussmann2007eco} and \citealt{govaert2019eco} for comprehensive reviews).
-More generally, eco-evolutionary approaches contribute a }
+Extensions of these theories to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and quantitative traits \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects} all highlight the sometimes counterintuitive demographic consequences of adaptive evolution that are missed by classical population genetics models.
+}
 
@@
 The population state at time $t$ is described by a population state vector, $\tilde{\mbf{n}}(t)$, which is ordered by how individuals were produced (selfing vs.~outcrossing), then by genotype, and finally by stage.
-For a single diallelic locus with alleles $A$ and $a$, we have three genotypes ($AA,\, Aa,\, aa$; $g = 3$), giving the population state vector is:
+For a single diallelic locus with alleles $A$ and $a$, we have three genotypes ($AA,\, Aa,\, aa$; $g = 3$), giving the population state vector:
 \begin{linenomath*}
@@
 The population projection matrix $\tilde{\mbf{A}}$ is constructed from four sets of matrices representing the demographic and genetic processes: The matrices $\mbf{U}^{S}_{i}$ and $\mbf{U}^{X}_{i}$ contain transition and survival probabilities for each genotype, produced by selfing and outcrossing respectively.
-The matrices $\mbf{F}_{i}$ and $\mbf{F}^{\prime}_{i}$ contain the genotype $\times$ stage specific contribution of genotype $i$ to the female and male gamete pools, respectively, and therefore to zygotes in the next generation.
+The matrices $\mbf{F}_{i}$ and $\mbf{F}^{\prime}_{i}$ contain the genotype $\times$ stage specific contributions of genotype $i$ to the female and male gamete pools, respectively, and therefore to zygotes in the next generation.
 We assume that whether individuals were produced through selfing or outcrossing does not affect their fecundity or mating success, that is, we assume  $\mbf{F}^X_{i}=\mbf{F}^S_{i}=\mbf{F}_{i}$ and $\mbf{F}^{\prime X}_{i}=\mbf{F}^{\prime S}_{i}=\mbf{F}^{\prime}_{i}$, respectively.
@@
 The population projection matrix is then only a function of the proportional population vector, $\tilde{\mbf{A}}[\tilde{\mbf{p}}(t)]$, and the model is referred to as a frequency-dependent model.
-In \ref{App:DensDep} we construct and analyze a density-dependent version of the model and show that with stage-independent density-dependent growth, our main conclusions remain intact.}
+In \ref{App:DensDep} we construct and analyze a density-dependent version of the model and show that with density-dependent survival affecting all stages equally, our main conclusions remain intact.}
 
@@
 
-\textcolor{red}{Lotte, I've reworded this to avoid dioecious language (males/females) and to emphasize that what really matters is that outcross matings are random, and selfing rates are genotype-independent...
-please check to make sure you agree with my edits} \textcolor{blue}{ For the purpose of this article, we assume that all outcross matings are random.
-The model can be extended to consider various forms of assortative mating beyond genotype-dependent selfing (see Online Supplement).
-In addition, we assume that individuals produced by selfing or by outcrossing do not differ in their mating success, though they can differ in their surival (see below).
+ \textcolor{blue}{ For the purpose of this article, we assume that all outcross matings are random.
+In addition, we assume that individuals produced by selfing or by outcrossing do not differ in their mating success, though they can differ in their surival (see above).
 Given these assumptions, the allele frequencies in the male mating population can obtained by simply summing the vectors of individuals produced through selfing and outcrossing:
@@
 where equality is obtained by normalizing the frequencies such that they add to one.
-The first entry sums the the stage-specific contributions to the gamete pool of heterozygotes and $AA$ homozygotes; the second entry sums the stage-specific contributions to the gamete pool of heterozygotes and $aa$ homozygotes.
+The first entry sums the  stage-specific contributions to the gamete pool of heterozygotes and $AA$ homozygotes; the second entry sums the stage-specific contributions to the gamete pool of heterozygotes and $aa$ homozygotes.
 }
@@
 \textcolor{blue}{ We now have all the ingredients to calculate the allele frequencies in the male gamete pool for the two-stage model in Eq(\ref{eq:F_male}), Eq(\ref{eq:maleGametePool}), and Table \ref{tab:Fitness}.
-Denoting the proportion of adults of genotype $i$ with $p^A_i$, the allele frequencies in the male gamete pool are given by
+Denoting the proportion of reproducing adults of genotype $i$ with $p^R_i$, the allele frequencies in the male gamete pool are given by
 \begin{linenomath*}
@@
 		\end{array} \right) = 
-				\frac{1}{(1-s_m)p^A_{aa}+ (1-hs_m)p^A_{Aa}+p_{aa}}\left(
+				\frac{1}{(1-s_m)p^R_{aa}+ (1-hs_m)p^R_{Aa}+p_{aa}}\left(
 		\begin{array}{c}
-(1-s_m)p^A_{AA}	+\frac{1}{2}(1-hs_m)p^A_{Aa}\\
-p^A_{aa}+\frac{1}{2}(1-hs_m)p^A_{Aa}
+(1-s_m)p^R_{AA}	+\frac{1}{2}(1-hs_m)p^R_{Aa}\\
+p^R_{aa}+\frac{1}{2}(1-hs_m)p^R_{Aa}
 		\end{array} \right).
@@
 The intrinsic population growth rate after convergence, $\lambda$, can be calculated as $\tilde{\mbf{n}}(t)/\tilde{\mbf{n}}(t-1)$.
-We note, however, that if the ecological component of the model is non-linear, more exotic dynamics are possible \citep{de2020matrix}.
 
@@
 "Demographically viable polymorphic parameter space" corresponds to the area inside the funnel that is also to the left of the extinction threshold for a given fertility value.
-Results are shown for three different population selfing rates ($C = \{0,\,1/4,\,1/2\}$), and two dominance scenarios (additivity, where $h = 1/2$, and dominance reversal, where $h = 1/4$); extinction thresholds are illustrated for three different values of female fecundity ($f$ values annotated on each panel).
-\textcolor{blue}{Predicted population intrinsic growth rates ($\lambda$) for the same parameter conditions are presented in fig.~S1, in the Online Supplementary Material.}}}
+Results are shown for three different population selfing rates ($C = \{0,\,1/4,\,1/2\}$), and two dominance scenarios (additivity, where $h = 1/2$, and dominance reversal, where $h = 1/4$).
+Extinction thresholds are illustrated for three different values of female fecundity ($f$ values annotated on each panel); \textcolor{blue}{ the coexistence boundaries are unaffected by the change in $f$.} \textcolor{blue}{Surface plots of predicted population intrinsic growth rates ($\lambda$) for the same parameter conditions are presented in fig.~S1, in the Online Supplementary Material.}}}
  \label{fig:extThresholds}
@@
 However, a key prediction from the evolutionary demographic model is that large fractions of SA polymorphic parameter space can be demographically inviable (fig.~\ref{fig:extThresholds}).
-The location of the extinction threshold, where the population intrinsic growth rate $\lambda = 1$, is primarily determined by the fertility parameter ($f$) but is also influenced by the population selfing rate ($C$) and dominance of the SA alleles ($h$).
+The location of the extinction threshold, where the population intrinsic growth rate $\lambda = 1$, is primarily determined by the fertility parameter ($f$) but is also influenced by the population selfing rate ($C$), and dominance of the SA alleles ($h$).
 
@@
 inv6 segregates at moderate frequency (about $8\%$), and carriers suffer an approximately $30 \%$ loss in pollen viability, but also increased flower (and therefore ovule and pollen) production.
-The genetic basis of these effects remain unclear -- though they likely polygenic -- but the net result is a "supergene" with remarkably strong effects on both female and male fertility that segregates as a single diallelic locus \textcolor{blue}{(i.e., with two alleles, wild-type and inverted)}.
+The genetic basis of these effects remain unclear -- though they are likely polygenic -- but the net result is a "supergene" with remarkably strong effects on both female and male fertility that segregates as a single diallelic locus \textcolor{blue}{(i.e., with two alleles, wild-type and inverted)}.
 We estimated selection coefficients for the effect of inv6 on pollen production (i.e., taking into account the simultaneous effect on flower number) and ovule production from the data reported in \citet{LeeKelly2015} under the relatively conservative assumptions that pollen/ovule production is proportional to flower number, and  additive SA fitness effects (dominance coefficients could not be estimated from field data; see \ref{App:Mimulus} for details of estimating selection coefficients).
@@
 
+\textcolor{blue}{In Appendix B we recreate Figure \ref{fig:mimulusFig} for a density-dependent model, replacing the heatmap of population growth rate with a heatmap of equilibrium population densities.
+We find that for a density-dependent term that affects survival of all (st)age classes in the same way, the coexistence boundaries are unchanged and the heatmap has the same shape.
+}
+
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 This demographic consequence of masculinization is rarely considered \citep{hitchcock2020gene}.
-Moreover, these findings complement recent theoretical and empirical studies indicating that SA selection is likely to be both condition dependent, and stronger in locally-adapted populations near the center of a species' range, where population growth rates are expected to be high \citep{BergerEtAl2014, Connallon2015}.
+Moreover, these findings complement recent theoretical and empirical studies indicating that SA selection is likely to be both condition dependent, and stronger in locally-adapted populations near the center of a species' range, where population growth rates are expected to be higher \citep{BergerEtAl2014, Connallon2015}.
 
@@
 Despite the demographic pitfalls associated with SA alleles, our case study using {\itshape M.~guttatus} appears to show that demographic rates observed in some real populations are capable of sustaining large regions of viable SA polymorphic space.
-The example also appears to provide some empirical support for the conjecure that locally-adapted populations are more likely to harbor SA polymorphisms than marginal or non-locally adapted ones; inviable polymorphic parameter space only occured when using the demographic data for non-local low-elevation perennial populations.
+The example also appears to provide some empirical support for the conjecture that locally-adapted populations are more likely to harbor SA polymorphisms than marginal or non-locally adapted ones; inviable polymorphic parameter space only occured when using the demographic data for non-local low-elevation perennial populations.
 Although we cannot make concrete predictions for inv6 in the Iron Mountain population in which it was observed, it is interesting that the estimated SA fitness effects place this polymophic inversion squarely in demographically viable polymorphic parameter space predicted by our model.
@@
 The density-dependent version of the model predicts equilibrium population densities rather than population intrinsic growth rates (see fig.~ \ref{fig:mimulusFig_dd}).
-This brief example shows that the shape of the polymorphic region is unchanged by density-dependence, and that the shape of the population density surface in fig.~\ref{fig:mimulusFig_dd} is very similar that for population growth rate in fig.~\ref{fig:mimulusFig}.
+This brief example shows that the shape of the polymorphic region is unchanged by this kind of density-dependence, and that the shape of the population density surface in fig.~\ref{fig:mimulusFig_dd} is very similar to that for population growth rate in fig.~\ref{fig:mimulusFig}.
 The similarity between the model predictions underscores an important point: density-dependence that acts age-independently does not alter the selection gradients (see also \citealt{mylius1995evolutionarily, caswell2017}).}
@@
 Likewise, in plants density-dependent competition for soil nutrients will likely influence seed germination rates differently than seedling growth rates or flower/fruit production in adults (CITATION).
-These types of age or stage specific density-dependence will change the selection gradients and therefore the shape of polymorphic parameter space as well as the equilibrium population densities predicted by the model.
-A thorough exploration of these more complicated forms of density dependence remains an important avenue for future work.}
+These types of age or stage specific density-dependence will change the selection gradients, and therefore the shape of polymorphic parameter space as well as the equilibrium population densities predicted by the model.
+\citep{de2020matrix} discuss some of the possible dynamical outcomes of a density-dependent genetic model for \textit{Tribolium} but a thorough exploration of the consequences of these more complicated forms of density-dependence for the maintenance of sexually antagonistic polymorphisms remains an important avenue for future work.}
 
@@
 We found that including basic demography can have a significant impact on traditional population genetic results, as has been suggested previously both by theoretical and empirical work \citep{kokko2003sexy, berger2016intralocus, grieshop2017male}.
-Although the potential negative consequences of sexual conflict for population viability have been known for some time, this aspect of sexual antagonism has rarely been considered in models investigating the scope for intralocus sexual antagonism to maintain genetic variation.
 
@@
  \centering
- \includegraphics[scale=0.8]{../output/figs/MimulusInv6NstarFig2.pdf}
+ \includegraphics[scale=0.8]{../output/figs/MimulusInv6NstarFig.pdf}
  \caption{\footnotesize{Illustration of density-dependent model predictions using empirically estimated demographic rates for {\itshape M.~guttatus}.

### 2021-12-29 doc/ResponseLetter-R1.tex
minor edits
2nd pass edits
2nd complete pass at edits in response to reviewer comments.

@@
 guttatus} example; and ({\itshape ii}) expanding our discussion of density-dependence to address these additions.
-Overall, these various additions have lengthened the manuscript \textcolor{blue}{by about a half-page}, but we feel that they have greatly improved the paper.
+Overall, we feel that these various additions have greatly improved the paper.
 \bigskip
@@
 guttatus} example with a simple DD version of the model.
-Second, we have expanded our results and discussion to more fully address the implications of DD, with reference to our modified figures and the new Appendix ({\bf L.XXX--XXX}).}
+Second, we have expanded our methods, results, and discussion to more fully address the implications of DD, with reference to our modified figures and the new Appendix ({\bf L.150--156; 478--481; 564--585}).}
 \end{quote}
@@
 For simplicity, we have decided to retain our original version of fig.1 (now {\bf fig.
-2}) in the main text and include the modified version showing $\lambda$ as a supplementary figure (now {\bf fig.~XXX}); but will swap this figure if the editor/reviewers feel differently.
+2}) in the main text and include the modified version showing $\lambda$ as a supplementary figure (now {\bf fig.~S1}); but will swap this figure if the editor/reviewers feel differently.
 However, we agree that the {\itshape M.
@@
 	{\itshape We agree that in our efforts to appeal to both demographers and evolutionary biologists studying sexual conflict, we ended up insufficiently discussing the relevant literature on both counts (but especially regarding evolutionarily explicit structured models).
-We have made a variety of changes to our introduction to improve this ({\bf L.XX--XX, L.XX-XX}).
+We have made a variety of changes to our introduction to improve this ({\bf L.45--51, L.69--71, L.76--85}).
 See also our specific responses to Reviewer 1.
@@
 \begin{quote}
-	{\itshape ...}
+	{\itshape We have added a brief general explanation of sexually antagonistic selection, which we felt was lacking in the first draft, and also allows us to more clearly link SA polymorphism with its' demographic consequences ({\bf L.2--3, L.6--7}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape We now briefly discuss this possibility in the Discussion ({\bf L.586--593}).}
+	{\itshape We now briefly discuss this possibility in the Discussion ({\bf L.605--611}).}
 \end{quote}
@@
 	{\itshape This is a really important comment and we thank you for pointing out this major shortcoming.
-One of our main goals while crafting the first draft of the manuscript was to find a middle ground that would appeal to both demographers and people studying sexual conflict.
+One of our main goals when crafting the first draft of the manuscript was to find a middle ground that would appeal to both demographers and people studying sexual conflict.
 It is clear that we didn't quite achieve that goal.
 Rather, by making compromises in the introduction we ended up lacking scholarship on both sides.
-To address your main concern, we have extended the Introduction to more directly connect with the broader literature on evolutionarily explicit structured models (see {\bf L.66-83}) and now also provide a more thorough link with the empirical sexually antagonistic selection literature (see {\bf L.42--49}).
-We hope these changes succeed in broadening the appeal of our manuscript to the wider readership of AmNat.}
+To address your main concern, we have extended the Introduction to more directly connect with the broader literature on evolutionarily explicit structured models (see {\bf L.76--85}) and now also provide a more thorough link with the empirical sexually antagonistic selection literature (see {\bf L.45--52}).
+We hope these changes succeed in broadening the appeal of our manuscript to the wider AmNat readership.}
 \end{quote}
@@
 
-I have one main comment:
--- Discussion: Conclusions are drawn before limitations of the model are mentioned.
+\noindent I have one main comment:
+
+\noindent-- Discussion: Conclusions are drawn before limitations of the model are mentioned.
 However, it seems to me that it would be very important to discuss the effect of those limitations on the results themselves and therefore discuss the generality of the results.
@@
 	{\itshape While we feel that there is strong precedence for discussing the key results of a study before diving into the various caveats that come with the analysis, the point is well taken that we need to clearly direct readers to the places in our paper where these limitations are discussed.
-In our Methods section, we are now careful to point readers to the various Appendixes, Supplements, and places in the Discussion where these limitations are dealt with (e.g., see {\bf 100-105; 112--117; 149--155; 210; fig.~2 caption}), and we have added more explicit discussion of several of our key assumptions in the discussion ({\bf e.g., L.556--574; 594--601})}
+In our Methods section, we are now careful to point readers to the various Appendixes, Supplements, and places in the Discussion where these limitations are dealt with (e.g., see {\bf L.101--103, 114--119, 151--157, 215--216, fig.~2 caption}), and we have added more explicit discussion of several of our key assumptions in the discussion ({\bf e.g., L.565--586; 606--612}).
+We discuss both the causes and potential effects of incorporating male-limitation of reproductive success on {\bf L.587--599}}.
 \end{quote}
@@
 \begin{quote}
-	{\itshape Response to the above 5 points: As per the reviewers (and the editors) suggestion, we have moved a lot of the derivations into the supplementary materials, and now no longer use H or W in the main text.
+	{\itshape \ul{Response to the above 5 points}: As per the reviewers (and the editors) suggestion, we have moved a lot of the derivations into the supplementary materials, and now no longer use H or W in the main text.
 In doing so, we hope to shift the focus from mathematical derivations, which are probably unhelpful to many readers, towards explaining the end result instead.
+However, we have made the requested changes to clarify the derivation in the Supplement.
 }
@@
 	\begin{quote}
-	{\itshape That's a mistake on our part, apologies. We have corrected this in {\bf Appendix A}, where it now appears.}
+	{\itshape That's a mistake on our part, apologies. We have corrected this in {\bf Online Supplement 1}, where it now appears.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape Very good point, that was awkward wording on our part.
-We have rephrased to more clearly explain that it is primarily the speed/magnitude of environmental change that we are referring to ({\bf L.20}).
-We have also included a new citation of a recent AmNat paper that explores some of these issues explicitly.}
+	{\itshape Very good points, that was awkward wording on our part.
+We have rephrased to more clearly explain that it is primarily the speed/magnitude of environmental change that we are referring to ({\bf L.23}).
+We have also included a new citation of a timely AmNat paper that explores some of these issues explicitly.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape Agreed. We have rephrased to avoid the term "trade-off" which tends to over-simplify the concept in or opinion ({\bf L.25}).}
+	{\itshape Agreed. We have rephrased to avoid the term "trade-off" which tends to over-simplify the concept in our opinion ({\bf L.28}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape We have rephrased the first sentence of this paragraph to be more specific ({\bf L.27}), which we hope also addresses this comment.}
+	{\itshape We have rephrased the first sentence of this paragraph to be more specific ({\bf L.30}), which we hope also addresses this comment.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape \textcolor{red}{Lotte, double check here?} We have added a few lines to discuss frequency vs density dependence when we first introduce the population projection matrix in equation 3 (CHECK LINE NUMBERS AFTER WE"RE DONE EDITING ).
+	{\itshape \textcolor{red}{Lotte, double check here?} We have added a few lines to discuss frequency vs density dependence when we first introduce the population projection matrix in equation 3 ({\bf L.151--157}).
 In addition, we have tried to clarify to improve this section by discussing that the survival and fertility rates ("demographic rates") are density-indepedent, and that the allele frequency dynamics make the model frequency dependent.}
@@
 \begin{quote}
-	{\itshape We have rephrased this sentence to clarify and be more consistent with our language earlier in the preceding paragraphs ({\bf L.37--39}).}
+	{\itshape We have rephrased this sentence to clarify and be more consistent with our language earlier in the preceding paragraphs ({\bf L.40--42}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape We have added a sentence earlier in this paragraph where we remind readers what dioecy is ({\bf L.54}).
+	{\itshape We have added a sentence earlier in this paragraph where we remind readers what dioecy is ({\bf L.57}).
 Hermaphroditism is defined earlier in the paragraph, so an additional reference seemed redundant.}
@@
 \begin{quote}
-	{\itshape Fixed ({\bf L.103-104}).}
+	{\itshape Fixed ({\bf L.105-106}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape Good point, fixed ({\bf L.161, 185}).}
+	{\itshape Good point, fixed ({\bf L.193}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape We want to thank the reviewer for this suggestion.
-We now start with a figure that only has one f, and that separately explains what happens to the allele frequencies (Population genetic outcome) and to the population growth rate (Demographic Outcome), see new {\bf fig.
+	{\itshape Thank you for this suggestion.
+To address this issue, as well as the Editor's request that we present population growth rates, we now start with an illustrative figure that only has one $f$, and that shows separately what happens to the allele frequencies (Population Genetic Outcome) and to the population growth rate (Demographic Outcome), see new {\bf fig.
 1}.
@@
 We have also included a new supplementary figure, {\bf fig.
-S1}, with detailed heatmaps for each of the parameter combinations presented in the this figure (i.e., one $f$ per panel).
+S1}, with detailed heatmaps for each of the parameter combinations presented in this figure (i.e., one $f$ per panel).
 However, we felt that this revised figure was to large and complicated to present in the main text (18 vs.
@@
 \begin{quote}
-	{\itshape We have added a brief parenthesis to clarify ({\bf L.434--435}).}
+	{\itshape We have added a brief parenthesis to clarify ({\bf L.447--448}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape This figure now appears as four separate heatmaps, each with one set of invasion conditions ({\bf fig.~5}). We hope this eases interpretation of the figure.}
+	{\itshape This figure now presents four separate panels with heatmaps of $\lambda$, each with one set of invasion conditions ({\bf fig.~5}). We hope this eases interpretation.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape We have re-written this sentence to avoid this awkward phrasing ({\bf L.579--580}).}
+	{\itshape We have re-written this sentence to avoid this awkward phrasing ({\bf L.600--601}).}
 \end{quote}

### 2021-12-29 doc/SA-Hermaphrodites-wDemography-R1.tex
2nd pass edits
2nd complete pass at edits in response to reviewer comments.

@@
 % Alternative opening sentence to the abstract:sentence
-Classic population genetics theory has been fundamental to understanding the evolution of sex-differences and the maintenance of sexually antagonistic (SA) genetic variation, but these models have rarely considered the demographic consequences of intralocus sexual antagonism.
-In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (i.e., dioecious/gonochorous) and partially selfing hermaphrodite populations.
-We focus on identifying parameter conditions under which SA polymorphism is maintained {\itshape and} the population growth rate remains positive.
-\textcolor{blue}{A sexually antagonistic polymorphism is a polymorphism of two alleles with antagonistic effects on the two sexes, that is, one of the two coexisting allele confers a fitness benefit to females and reduces male fitness, and the other allele confers a fitness benefit to males and reduces female fitness.
-} Additionally, we analyze the effects of inbreeding depression manifesting at different life-history stages and give an illustrative example of the potential for SA polymorphism in real populations using empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus}.
+
+\textcolor{blue}{When selection differs between the sexes, genes expressed by both males and females can experience sexually antagonistic (SA) selection, where beneficial alleles for one sex are deleterious for the other.} Classic population genetics theory has been fundamental to understanding the evolution of sex-differences and the maintenance of such SA genetic polymorphisms, but these models have rarely considered the demographic consequences \textcolor{blue}{ of the maladaptation imposed by coexisting alleles with deleterious fitness effects in each sex.} In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (i.e., dioecious/gonochorous) and partially selfing hermaphrodite populations.
+We focus on identifying when SA polymorphisms are maintained by balancing selection {\itshape and} the population growth rate remains positive.
+Additionally, we analyze the effects of inbreeding depression manifesting at different life-history stages and give an illustrative example of the potential for SA polymorphism in real populations using empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus}.
 Our results show that when population intrinsic growth rates approach one, extinction occurs across large swathes of parameter space favoring SA polymorphism or the fixation of male-beneficial alleles, and that inbreeding depression is a significant problem for maintaining SA polymorphism in partially selfing populations.
-Despite these demographic challenges, our example with {\itshape M.~guttatus} appears to show that demographic rates observed in some real populations are capable of sustaining large regions of viable SA polymorphic space.
+Despite these demographic challenges, our example with {\itshape M.~guttatus} appears to show that demographic rates observed in some real populations can sustain large regions of viable SA polymorphic space.
+
+%\textcolor{blue}{A sexually antagonistic polymorphism is a polymorphism of two alleles with antagonistic effects on the two sexes, that is, one of the two coexisting allele confers a fitness benefit to females and reduces male fitness, and the other allele confers a fitness benefit to males and reduces female fitness.
+}
 
@@
 
-% GOOD STUFF HERE...
-KEEPING THIS IN CASE WE WANT TO WORK MORE OF THIS INTO THE ABSTRACT...In this paper we developed a stage-structured mendelian matrix model that accommodates selection through male and female sex-functions in populations of simultaneous hermaphrodites with mixed mating systems.
-Using the model, we jointly analyzed the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations.
-We focused our analyses on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population is also demographically viable (i.e., has a positive intrinsic growth rate).
-We also looked at the effects of inbreeding depression manifesting in different life-history stages, exploring the fitness consequences of early- vs.~late-acting inbreeding depression on the model predictions.
-Finally, we gave an illustrative example of the potential for SA polymorphism in real populations by parameterizing our model with empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus} (now {\itshape Erythranthe guttata}).
-
 \newpage{}
@@
 Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, such as deleterious mutations \citep{Haldane1937}, \textcolor{blue}{rapid, unpredictable, or extreme changes} in environmental conditions \citep{Maynard-Smith1976, LandeShannon1996, OrrUnckless2008,Lyberger-etal-2021}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, and genetic constraints \citep{ConnallonHall2018, MatthewsConnallon2019}.
-Genetic constraints on adaptation can arise as a consequence of conflicting selection and gene-flow between different classes of individuals, or more generally, when \textcolor{blue}{there is a negative correlation between different fitness components} \citep{CharlesworthHughes2000, ConnallonHall2018}.
-
-\textcolor{blue}{Genetic constraints on adaptation arising from conflicting selection and gene flow between different classes of individuals} have particularly interesting evolutionary consequences \citep{ConnallonDebarreLi2018}.
-On the one hand, they provide an effective mechanism for the maintenance of genetic variation, which can increase a population's capacity for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
+Genetic constraints on adaptation can arise as a consequence of conflicting selection and gene-flow between different classes of individuals, or more generally, when \textcolor{blue}{different fitness components are negatively correlated} \citep{CharlesworthHughes2000, ConnallonHall2018}.
+
+\textcolor{blue}{Genetic constraints on adaptation arising from conflicting selection and gene flow} have particularly interesting evolutionary consequences \citep{ConnallonDebarreLi2018}.
+On one hand, they provide an effective mechanism for the maintenance of genetic variation, which can increase a population's capacity for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
 On the other hand, they prevent individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase a population's overall extinction risk \citep{kokko2003sexy,harts2014demography}.
@@
 
-Sexually antagonistic selection (abbreviated SA hereafter) \textcolor{blue}{is an important form of conflicting selection that arises when beneficial alleles for one sex are deleterious when expressed in the other.
+Sexually antagonistic selection (abbreviated SA hereafter) \textcolor{blue}{is an important form of conflicting selection that occurs when beneficial alleles for one sex are deleterious when expressed in the other.
 SA selection appears to be} a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and the evolution of sexual dimorphism and sex-related traits \citep{Kidwell1977, Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, ConnallonClark2012, BondurianskyChenoweth2009, Olito2019}.
 \textcolor{blue}{For instance, phenotypic selection studies commonly find sex-differences in directional selection \citep[e.g.,][]{DeLisle-etal-2018,Lewis-etal-2011,CoxCalsbeek2009,SinghPunzalan2018}, implying that many genetic variants underlying quantitative traits have SA fitness effects.
-Likewise, quantitative genetic studies frequently find negative cross-sex genetic correlations for fitness, indicating SA fitness effects for multilocus genotypes (\citealt{Chippendale-etal-2001,Delph-etal-2011}; reviewed in \citealt{ConnallonMatthews2019}).
-Moreover, polygenic signals of SA selection are evident in recent population genetic analyses of large genomic datasets \citep{RuzickaESEB2020, RuzickaConnallon2022}.}
+Likewise, quantitative genetic studies frequently find negative cross-sex genetic correlations for fitness (\citealt{Chippendale-etal-2001,Delph-etal-2011}; reviewed in \citealt{ConnallonMatthews2019}), and negative correlationpolygenic signals of SA selection are evident in recent population genetic analyses of large genomic datasets \citep{RuzickaESEB2020, RuzickaConnallon2022}.}
 
@@
 Most population genetic models, including models of SA selection, model the relative fitness of genotypes in populations of constant (and often infinite) size, and do not consider the population dynamical consequences of evolutionary change.
-In reality, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals \textcolor{blue}{that link genotype distributions to population dynamics (e.g., \citealt {johnston2013life, merot2020balancing}).
+In reality, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals \textcolor{blue}{which link genotype distributions to population dynamics (e.g., \citealt {johnston2013life, merot2020balancing}).
 Over the past 50 years, the field of eco-evolutionary dynamics has sought to jointly model the genetic and demographic processes involved in evolutionary change.
 In particular, theoretical models combining population genetics with density-dependent population growth \citep{roughgarden1971density} and stage-structured populations (consolidated in \citealt{Charlesworth_1994}) paved the way toward understanding the demographic changes accompanying allele frequency changes due to selection.
-Extensions of these theories to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and quantitative traits \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects} all highlight the sometimes counterintuitive demographic consequences of adaptive evolution that are missed by classical population genetics models.
-}
-
-%\textcolor{blue}{One of the earliest such models \citep{roughgarden1971density} combined a population genetic model with a density-dependent population model with no intraspecific demographic structure.
-Charlesworth then combined an age-structured population model with population genetics in a series of papers, consolidated in \cite{Charlesworth_1994}.
-\cite{orive1995senescence} extended this work to study life-history evolution in organisms with both clonal and sexual reproduction (see also \cite{orive2001somatic}).
-\cite{ronce2000kin} added spatial structure and stochasticity to age-structured genetic models to model kin selection, a body of work consolidated in \cite{rousset2004genetic}.
-Of course there are many other approaches to studying eco-evolutionary dynamics, e.g.
-combining quantitative genetics models with demographic structure (e.g.
-\cite{coulson2006putting,barfield2011evolution,childs2016evolution,orive2017effects}) or adaptive dynamics approaches (e.g.
-\cite{dieckmann2002adaptive,durinx2008adaptive,ten2020habitat}).
-For a comprehensive review of the eco-evolutionary literature, see \cite{fussmann2007eco}, or more recently, \cite{govaert2019eco}.}
+Extensions of these theories to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and quantitative traits \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects} all highlight that a fuller understanding of the genetic basis of fitness variation, and the processes that maintain it, requires an explicit treatment of the demographic consequences of evolutionary change.}
 
@@
 Simultaneous hermaphrodites can transmit genes to the next generation via both sperm/pollen and eggs/ovules, and have the potential to reproduce by a combination of self- and outcross-fertilizations.
-Maternal outcrossing involves receiving male gametes from another individual in the population, while paternal outcrossing involves exporting male gametes to another individual and fertilizing their ovules.
-Self-fertilization is achieved when an individual's male gametes fertilize their own ovules.
-To distinguish between parameters relating to male and female function, we denote matrices or vectors relating to the male sex function with a prime.
+Maternal outcrossing involves receiving male gametes from other individuals in the population, while paternal outcrossing involves exporting male gametes to other individuals and fertilizing their ovules.
+Self-fertilization occurs when an individual's male gametes fertilize their own ovules.
+To distinguish between parameters relating to male and female reproductive functions, we denote matrices or vectors relating to the male sex function with a prime.
 \textcolor{blue}{Following previous theory for partially selfing populations \cite[e.g.,][]{Charlesworth2010,JordanConnallon2014,Glemin2021}, we use a 'fixed prior selfing' model where all individuals are assumed to self-fertilize at a constant rate, $C$, prior to receiving outcross pollen, for reasons of analytic tractability.
-We present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}) in the Online Supplement, and briefly discuss implications of this approach in the Discussion.}
+We present a more general model of genotype-specific self-fertilization rates in the Online Supplement, and briefly discuss implications of this approach in the Discussion.}
 
@@
 
-\textcolor{blue}{To demonstrate how to construct and analyze a genotype $\times$ stage-classified model, we will now construct an example model for a hypothetical species with intralocus sexual conflict via the two sex functions.
+\textcolor{blue}{To demonstrate how to construct and analyze a genotype $\times$ stage-classified model, we will now construct an example model for a hypothetical hermaphrodite species with intralocus sexual conflict via the two sex functions.
 }For the sake of simplicity, we assume our hypothetical species has a life cycle with only two stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active.
@@
  \textcolor{blue}{ For the purpose of this article, we assume that all outcross matings are random.
-In addition, we assume that individuals produced by selfing or by outcrossing do not differ in their mating success, though they can differ in their surival (see above).
-Given these assumptions, the allele frequencies in the male mating population can obtained by simply summing the vectors of individuals produced through selfing and outcrossing:
+In addition, we assume that individuals produced by selfing or by outcrossing do not differ in their mating success, though they can differ in their survival (see above).
+Given these assumptions, the allele frequencies in the mating population can obtained by summing the vectors of individuals produced through selfing and outcrossing:
 \begin{linenomath*}
@@
 \textcolor{blue}{Individuals contribute  gametes to the population outcrossing gamete pool according to their fertility, which depends on their stage and genotype at the SA locus.
-The allele frequencies in the male gamete pool are then calculated as follows (see Online Supplementary Materials for a derivation of this equation),
+The allele frequencies in the male gamete pool are then calculated as
 \begin{linenomath*}
@@
 \end{linenomath*}
-where equality is obtained by normalizing the frequencies such that they add to one.
+where equality is obtained by normalizing the frequencies such that they add to one (see Online Supplementary Materials for a derivation of Eq(\ref{eq:maleGametePool})).
 The first entry sums the  stage-specific contributions to the gamete pool of heterozygotes and $AA$ homozygotes; the second entry sums the stage-specific contributions to the gamete pool of heterozygotes and $aa$ homozygotes.
@@
 \end{linenomath*}
-Note that this equation is identical to the classical population genetic results for frequencies of male gametes in a sexually antagonistic model, with the caveat that only reproductively active adults contribute to the gamete pool.
+Note that this equation is identical to the classical population genetic result for allele frequencies after selection among males in a two-sex SA model, with the caveat that only reproductively active adults contribute to the gamete pool \citep[e.g., see Eq(1a) of][]{Kidwell1977}.
 Eq(\ref{eq:maleGametePool_example}) increasingly deviates from the classical population genetics result if additional reproductive adult age or stage classes are included (see Eq(\ref{eq:maleGametePool})).
@@
 \subsubsection*{Population projection}
-Using the component matrices described above (the survival matrices, $\mathbf{U}^X_i$, $\mathbf{U}^S_i$, the fertility and mating success matrices, $\mathbf{F}_i$, $\mathbf{F}^\prime_i$) and the allele frequencies in the gamete pool, we can construct the population projection matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$.
+Using the component matrices described above (the survival matrices, $\mathbf{U}^X_i$, $\mathbf{U}^S_i$, the fertility and mating success matrices, $\mathbf{F}_i$, $\mathbf{F}^\prime_i$) and the allele frequencies in the male gamete pool, we can construct the population projection matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$.
 The resulting matrix that projects the eco-evolutionary dynamics is:
@@
 
-Iterating the projection matrix, Eq(\ref{eq:Atilde}) with the above demographic matrices, given an initial population state vector, allows numerical simulation of the eco-evolutionary dynamics for selection operating on any of the sex-function, or stage-specific demographic parameters.
+Iterating the projection matrix, Eq(\ref{eq:Atilde}), with the above demographic matrices, given an initial population state vector, allows numerical simulation of the eco-evolutionary dynamics for selection operating on any of the stage- or sex-function-specific demographic parameters.
 As we outline below, we use numerical techniques together with mathematical analyses to study the conditions for the maintenance of SA polymorphisms, and the demographic fate of the populations (i.e., positive growth, or extinction).
@@
 These chosen values are similar to those used in \citet{deVriesCaswell2019b}, facilitating comparison between models, and correspond to a life-history in which individuals spend multiple timesteps in the juvenile phase prior to maturing into reproductively active adults, but are otherwise arbitrary.
-Our parameters of interest include fertility, $f$, the inbreeding depression parameters, $\delta$, $\delta_j$, $\delta_a$, $\delta_{\gamma}$, and the selection parameters $h_f$, $s_f$, $h_m$, and $s_m$, which are given different values for each analyses as described in the figure captions.
+Our parameters of interest include \textcolor{blue}{maximum female} fertility, $f$, the inbreeding depression parameters, $\delta$, $\delta_j$, $\delta_a$, $\delta_{\gamma}$, and the selection parameters $h_f$, $s_f$, $h_m$, and $s_m$, which are given different values for each analyses as described in the figure captions.
 
@@
 Diverse eco-evolutionary outcomes are possible in the model, including fixation of either allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and suicide (e.g., see Figure 1 of \citealt{deVriesCaswell2019a, deVriesCaswell2019b} for examples of the possible outcomes).
-We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection \textcolor{blue}{(Figure \ref{fig:didactic}, left-hand panel)} and ({\itshape ii}) the intrinsic population growth rate at equilibrium is positive \textcolor{blue}{(Figure \ref{fig:didactic}, right-hand panel)}; a situation that we refer to as a 'demographically viable SA polymorphism'.
+We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection \textcolor{blue}{(Figure \ref{fig:didactic}, left-hand panel)} and ({\itshape ii}) the intrinsic population growth rate at equilibrium is greater than $1$ \textcolor{blue}{(Figure \ref{fig:didactic}, right-hand panel)}; a situation that we refer to as a 'demographically viable SA polymorphism'.
 
@@
 Specifically, for each boundary equilibrium we introduced the rare allele at low initial frequency and iterated Eq(\ref{eq:Atilde}) until the population had reached demographic and genotypic equilibrium.
-\textcolor{blue}{Although we use density-independent survival and fertility rates, the allele frequency dynamics introduce frequency-dependence into the model.
-As a consequence, }the population state vector will grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see chapter 17 in \citealt{Caswell2001}).
-The intrinsic population growth rate after convergence, $\lambda$, can be calculated as $\tilde{\mbf{n}}(t)/\tilde{\mbf{n}}(t-1)$.
+\textcolor{blue}{Although we use density-independent survival and fertility rates, the allele frequency dynamics introduce frequency-dependence into the model.} The population state vector will therefore grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see chapter 17 in \citealt{Caswell2001}).
+The intrinsic population growth rate after convergence, $\lambda$, can be calculated as \textcolor{blue}{ $n(t)/n(t+1)$, where $n = \sum_i n_i $ sums over all entries of the population state vector $\tilde{\mbf{n}}$.}
 
@@
 We present scenarios of equal dominance in the main text (i.e., $h_f = h_m = h$).
-Specifically, we examine scenarios of ({\itshape i}) additive SA fitness effects ($h = 1/2$), which are commonly observed for small-effect alleles or quantitative traits \citep{AgrawalWhitlock2011}; and ({\itshape ii}) dominance reversals, where the deleterious fitness effect of each SA allele is partially recessive in each sex ($h = 1/4$), which are predicted under fitness landscape models provided the population is not too far from the phenotypic optimum \citep{Manna2011, ConnallonClark2014}.
+Specifically, we examine scenarios of ({\itshape i}) additive SA fitness effects ($h = 1/2$), which are commonly observed for small-effect alleles or quantitative traits \citep{AgrawalWhitlock2011}; and ({\itshape ii}) dominance reversals, where the deleterious fitness effect of each SA allele is partially recessive through each sex function ($h = 1/4$), which are predicted under fitness landscape models provided the population is not too far from the phenotypic optimum \citep{Manna2011, ConnallonClark2014}.
 We briefly explore the consequences of sex-specific dominance in the Online Supplementary Material.
@@
 Following \citet{Olito2019}, we incorporate such negative covariance by constraining the inbreeding depression terms in the model ($\delta$ and $\delta_i$, where $i \in \{j,a,\gamma\}$) to follow a simple declining function of the selfing rate: \textcolor{blue}{$\delta = \delta^{\ast} (1 - b (1 - a(1 - C)/(C + a(1 - C))))$}, where $\delta^{\ast}$ is the hypothetical severity of inbreeding depression for a completely outcrossing population, $b$ is a shape parameter determining how far $\delta$ will decline under complete selfing (when $C = 1$), \textcolor{blue}{and $a$ is a shape parameter} controlling the curvature of the overall function for $\delta$ (see Appendix E in \citealt{Olito2019} for additional details).
-We set $\delta^{\ast} = 0.8$, $b = 0.5$, and $a = 0.2$ for all analyses, values chosen to be consistent with empirical estimates of inbreeding depression (e.g., fig.~2 in \citealt{HusbandSchemske1996}).
+We set $\delta^{\ast} = 0.8$, $b = 0.5$, and $a = 0.2$ for all analyses involving inbreeding depression, values chosen to be consistent with empirical estimates of inbreeding depression (e.g., fig.~2 in \citealt{HusbandSchemske1996}).
 
@@
 Results are shown for three different population selfing rates ($C = \{0,\,1/4,\,1/2\}$), and two dominance scenarios (additivity, where $h = 1/2$, and dominance reversal, where $h = 1/4$).
-Extinction thresholds are illustrated for three different values of female fecundity ($f$ values annotated on each panel); \textcolor{blue}{ the coexistence boundaries are unaffected by the change in $f$.} \textcolor{blue}{Surface plots of predicted population intrinsic growth rates ($\lambda$) for the same parameter conditions are presented in fig.~S1, in the Online Supplementary Material.}}}
+Extinction thresholds are illustrated for three different values of female fecundity ($f$ values annotated on each panel); \textcolor{blue}{ the coexistence boundaries are unaffected by the change in $f$.} \textcolor{blue}{Surface plots of predicted population intrinsic growth rates ($\lambda$) for the same parameter conditions are presented in fig.~S1 of the Online Supplementary Material.}}}
  \label{fig:extThresholds}
@@
 This causes a greater proportion of SA polymorphic parameter space to be demographically viable, with the demographic model predictions converging on those from population genetic models in high-fertility populations (fig.~\ref{fig:deltaPolySpace}).
-For populations with lower fertility rates, the proportion of demographically viable polymorphic parameter space matches the population genetic model predictions under complete outcrossing; however, as the selfing rate increases, demographic viability eventually crashes when the population can no longer sustain the concomitant increase in mortality due to inbreeding depression (fig.~\ref{fig:deltaPolySpace}).
-
-Mortality from inbreeding depression quickly outweighs the beneficial effect of selfing on the equilbirium female load that was apparent in the previous section.
-In contrast to our earlier results, predominantly outcrossing populations are predicted to have the highest proportion of demographically viable polymorphic space when inbreeding depression is taken into acount (compare fig.~\ref{fig:polySpace} with fig.~\ref{fig:deltaPolySpace}).
-Regardless of the life-history stage at which inbreeding depression affects survival, populations with intermediate to high selfing rates are unlikely to harbour SA polymorphism unless they can afford the resulting loss of self-fertilized ovules/offspring.
-An interesting alternative interpetation of these results is that populations with intermediate to high selfing rates and population intrinsic growth rates near one are vulnerable to extinction if a sexually antagonistic allele invades the population.
+For populations with lower fertility rates, the proportion of demographically viable polymorphic parameter space \textcolor{blue}{is very similar to} the population genetic model predictions \textcolor{blue}{for low selfing rates}.
+\textcolor{blue}{Note, however, that in our model, mortality due to late-acting inbreeding depression also reduces the population effective selfing rate, and therefore shifts the invasion boundaries for the SA alleles to more closely resemble an outcrossing population.
+In some cases, this can lead to larger regions of viable polymorphic parameter space than predicted by the population genetic models, which can only include early-acting inbreeding depression (e.g., see cross symbols in fig.~\ref{fig:deltaPolySpace})}.
+However, as the selfing rate increases, demographic viability eventually crashes when the population can no longer sustain the concomitant increase in mortality due to inbreeding depression (fig.~\ref{fig:deltaPolySpace}).
 
@@
 
+Mortality from inbreeding depression quickly outweighs the beneficial effect of selfing on the equilbirium female load that was apparent in the previous section.
+In contrast to our earlier results, predominantly outcrossing populations are predicted to have the highest proportion of demographically viable polymorphic space when inbreeding depression is taken into acount (compare fig.~\ref{fig:polySpace} with fig.~\ref{fig:deltaPolySpace}).
+Regardless of the life-history stage at which inbreeding depression affects survival, populations with intermediate to high selfing rates are unlikely to harbour SA polymorphism unless they can afford the resulting loss of self-fertilized ovules/offspring.
+An interesting alternative interpetation of these results is that populations with intermediate to high selfing rates and population intrinsic growth rates near one are vulnerable to extinction if a sexually antagonistic allele invades the population.
+
 The point in the life-cycle where inbreeding depression manifests can influence the threshold selfing rate at which demographically viable polymorphic parameter space crashes.
@@
 
-As an illustrative example of how our model can be used to explore whether demographic rates observed in natural populations appear likely to support balanced SA polymorphisms, we parameterized the model using empirically estimated demographic rates and fitness data for natural populations of the hermaphroditic flowering plant {\itshape Mimulus guttatus} (Scrophulariaceae; now known as {\itshape Erythranthe guttata}).
+To illustrate how our model can be used to explore whether demographic rates observed in natural populations appear likely to support balanced SA polymorphisms, we parameterized the model using empirically estimated demographic rates and fitness data for natural populations of the hermaphroditic flowering plant {\itshape Mimulus guttatus} (Scrophulariaceae; now known as {\itshape Erythranthe guttata}).
 {\itshape M.~guttatus} is an herbaceous, self-compatible wildflower native to western North America that exhibits remarkable among-population variation in numerous life-history and reproductive traits including selfing rates, inbreeding depression, floral morphology, and annual-to-perennial life-history \citep[e.g.,][]{RitlandGanders1987, Ritland1990, Willis1993, Willis1999a, Willis1999b, WuWillis2008}.
@@
 In contrast, when the selfing rate is at the higher end of empirical estimates for the Iron Mountain population in which inv6 has been documented ($C = 0.29$), inv6 falls nearer the upper boundary but still within SA polymorhic space when using the Low-Elevation Perennial demographic rates (fig.~\ref{fig:mimulusFig}B,D).
-This happens because under the empirical estimates of selfing and inbreeding depression, the polymorphic parameter space shifts downwards \textcolor{blue}{towards the x-axis}.
+This happens because, under the empirical estimates of selfing and inbreeding depression, the polymorphic parameter space shifts downwards \textcolor{blue}{towards the x-axis}.
 Additionally, inv6 falls much closer to the extinction threshold under partial selfing, suggesting that even relatively small perturbations to demographic rates or selection coefficients could result in non-locally adapted populations being unable to support the demographic costs associated with segregating SA alleles with selection coefficients of similar magnitude to inv6.
 
-\textcolor{blue}{In Appendix B we recreate Figure \ref{fig:mimulusFig} for a density-dependent model, replacing the heatmap of population growth rate with a heatmap of equilibrium population densities.
-We find that for a density-dependent term that affects survival of all (st)age classes in the same way, the coexistence boundaries are unchanged and the heatmap has the same shape.
+\textcolor{blue}{In Appendix B we reanalyze the {\itshape M.~guttatus} example using a simple density-dependent version of our model.
+We find that for (st)age-independent density-dependent growth, the coexistence boundaries are unchanged, and the equilibrium population densities closely resemble the shape of the density-independent population growth rates presented above (see fig.~\ref{fig:mimulusFig_dd}).
 }
@@
 The first and central finding of our study is that when population intrinsic growth rates approach one, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space where SA polymorphism is maintained.
-Since intrinsic growth rates far exceeding one suggest rapid exponential growth, they are generally rare (with the notable exception of recently introduced, invasive populations), suggesting that our model predictions may be highly relevant for SA polymorphism in many real-world populations.
+Since intrinsic growth rates far exceeding one suggest rapid exponential growth, they are generally rare (with the notable exceptions of recently introduced invasive populations, \textcolor{blue}{and locally adapted ones}), suggesting that our model predictions may be highly relevant for SA polymorphism in many real-world populations.
 In addition, we find that much of the parameter space where a male-beneficial allele \textcolor{blue}{fixes} is demographically inviable for populations whose intrinsic growth rates are close to one prior to invasion.
-This demographic consequence of masculinization is rarely considered \citep{hitchcock2020gene}.
+This demographic consequence of masculinization has rarely been considered \citep{hitchcock2020gene}.
 Moreover, these findings complement recent theoretical and empirical studies indicating that SA selection is likely to be both condition dependent, and stronger in locally-adapted populations near the center of a species' range, where population growth rates are expected to be higher \citep{BergerEtAl2014, Connallon2015}.
@@
 This key prediction is supported by a series of experimental results in seed beetles ({\itshape Callosobruchus maculatus}) which show that male-beneficial SA genotypes are less likely to contribute positively to population growth rates, and are more susceptible to extinction under environmental stress or inbreeding \citep{BergerEtAl2014, berger2016intralocus, grieshop2017male}.
-An interesting corrollary of our findings is that, since strong SA fitness effects often lead to extinction, the observable SA genetic variation in natural populations may often be under weak selection, and therefore strongly susceptible to genetic drift whether or not selection favours the fixation of one allele or balancing selection \citep{ConnallonClark2012}.
+An interesting corrollary of our findings is that, since strong SA fitness effects can often lead to extinction, the observable SA genetic variation in natural populations may often be under weak selection, and therefore strongly susceptible to genetic drift whether or not selection favours the fixation of one allele or balancing selection \citep{ConnallonClark2012}.
 
@@
 By combining the tools of demography and population genetics, the framework we present here enables the exploration of interactions between life cycle complexity, mating system, and sexual antagonism.
-For the sake of simplicity, we made a variety of simplifying assumptions in our analyses, and \textcolor{blue}{many extensions of the model are} possible.
+For the sake of simplicity, we made a variety of assumptions in our analyses, and \textcolor{blue}{many extensions of the model are} possible.
 For example, we used a simple life cycle with just two stages for most of this paper, adults and juveniles (the \textit{Mimulus} example has $3$ stages).
@@
 
-\textcolor{blue}{One particularly important limitation of our results is that we modeled density-independent growth.
+\textcolor{blue}{One particularly important limitation of our analysis is that we modeled density-independent growth.
 Of course, in natural populations density-dependent processes like resource competition or competition for breeding sites will often prevent populations from growing (or shrinking) exponentially.
@@
 This brief example shows that the shape of the polymorphic region is unchanged by this kind of density-dependence, and that the shape of the population density surface in fig.~\ref{fig:mimulusFig_dd} is very similar to that for population growth rate in fig.~\ref{fig:mimulusFig}.
-The similarity between the model predictions underscores an important point: density-dependence that acts age-independently does not alter the selection gradients (see also \citealt{mylius1995evolutionarily, caswell2017}).}
-
-\textcolor{blue}{In many natural populations, however, density-dependence does not have the same effect on the survival of individuals of all ages or developmental stages.
+The similarity between the density-independent and -dependent model predictions underscores an important point: density-dependence that acts age-independently does not alter selection gradients \citep{mylius1995evolutionarily, caswell2017}.}
+
+\textcolor{blue}{In many natural populations, however, density-dependence will not have the same effect on the survival of individuals of all ages or developmental stages.
 For example, competition for breeding sites will cause a reduction in fecundity rather than survival.
-Likewise, in plants density-dependent competition for soil nutrients will likely influence seed germination rates differently than seedling growth rates or flower/fruit production in adults (CITATION).
-These types of age or stage specific density-dependence will change the selection gradients, and therefore the shape of polymorphic parameter space as well as the equilibrium population densities predicted by the model.
-\citep{de2020matrix} discuss some of the possible dynamical outcomes of a density-dependent genetic model for \textit{Tribolium} but a thorough exploration of the consequences of these more complicated forms of density-dependence for the maintenance of sexually antagonistic polymorphisms remains an important avenue for future work.}
+Likewise, in plants density-dependent competition for soil nutrients will likely influence seed germination rates differently than seedling growth rates or flower/fruit production in adults \citep{AntonovicsLevin1980}.
+These types of age- or stage-specific density-dependence will change the selection gradients, and therefore the shape of polymorphic parameter space as well as the equilibrium population densities predicted by the model.
+\citet{de2020matrix} discuss some of the possible dynamical outcomes of a density-dependent genetic model for \textit{Tribolium} seed beetles but a thorough exploration of the consequences of more complicated forms of density-dependence for the maintenance of sexually antagonistic polymorphisms remains an important avenue for future work.}
 
@@
 Yet, many hermaphroditic populations experience limitation of reproductive success due to both quantity and quality of male gametes \citep[e.g.,][]{Yund2000,AizenHarder2007, Harder2016}.
-Explicitly modeling male gamete production is a natural extension to our modeling framework, and would enable us to analyze how tension between the demographic consequences of SA fitness variation through both female and male function alters or abolishes asymmetries in the extinction thresholds and polymorphic parameter space \citep[e.g.,][]{Tazzyman2015}.
+Explicitly modeling male gamete production is a natural extension to our modeling framework, and would enable us to analyze how tension between the demographic consequences of SA fitness variation through both female and male function alters or abolishes asymmetries in the extinction thresholds and polymorphic parameter space \textcolor{blue}{caused by our assumption of female demographic dominance} \citep[e.g.,][]{Tazzyman2015}.
 Interestingly, different forms of self-fertilization can aid in reproductive assurance, and should therefore have demographic consequences -- for example, under pollen limitation and delayed selfing (where only ovules that fail to receive outcross pollen are selfed), the selfing rate will be a function of genotype frequencies because these will directly influence pollen production \citep{HarderBarrett2006}.
@@
 It would be an interesting and feasible extension of our modeling framework to expand the individual state space to include selfing cohorts (i.e., first generation selfing, second generation selfing, etc.), as in the models of \citet{kelly1999response,kelly2007mutation}, thereby enabling a more biologically plausible approach to modelling self-fertilization.
-\textcolor{blue}{Likewise, we implemented a simple "fixed selfing" model, where all individuals in the population self-fertilize at the same rate.
-In reality, individuals will vary in rate of self-fertilization, which will influence both the expression of inbreeding depression and the scope for SA polymorphism \citep{JordanConnallon2014}.
-A naive expectation might be that among-individual variation in selfing might reduce the effective selfing rate in a population, expanding the scope for balanced SA polymorphisms, as it does in genotype-dependent mass-action selfing models \citep{JordanConnallon2014}.
+\textcolor{blue}{Likewise, we implemented a simple "fixed prior selfing" model, where all individuals in the population self-fertilize at the same rate, which is constant over time.
+In reality, environmental stochasticity will likely influence the population selfing rate and therefore both the expression of inbreeding depression and the scope for SA polymorphism \citep{JordanConnallon2014,Olito-etal-2018}.
+In this scenario, a naive expectation might be that the geometric mean selfing rate over time would represent a more biologically meaningful parameter influencing the genetic and demographic outcomes of SA selection.}
+
+% that among-individual variation in selfing might reduce the effective selfing rate in a population, expanding the scope for balanced SA polymorphisms, as it does in genotype-dependent mass-action selfing models \citep{JordanConnallon2014}.
 However, it is difficult to speculate how the associated demographic effects of such variation via reproductive assurance and inbreeding depression  would play out.}
@@
 In contrast, models of the population dynamical consequences of interlocus sexual conflict are more common \citep[e.g.,][]{tanaka1996sexual, martinez2017sexual}.
-We found that including basic demography can have a significant impact on traditional population genetic results, as has been suggested previously both by theoretical and empirical work \citep{kokko2003sexy, berger2016intralocus, grieshop2017male}.
+We found that including basic demography can have a significant impact on traditional population genetic results, as has been suggested previously by both theoretical and empirical studies \citep{kokko2003sexy, berger2016intralocus, grieshop2017male}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section{Example with density-dependence}\label{App:DensDep}
+\section{Stage-indepedent density-dependence}\label{App:DensDep}
 \renewcommand{\theequation}{B\arabic{equation}}
@@
 Results are shown for two hypothetical populations using demographic rates for locally adapted (Eagle Meadows; panels A,C), and non-local (Low-Elevation Perennial; panels B,D) populations reported by \citet{PetersonEtAl2016}.
-Invasion conditions and equilibrium densities ($N^{\ast}$) were calculated using the conservative assumption of additive fitness effects in both sexes ($h_f = h_m = 0.5$) for two parameter conditions: obligate outcrossing (panels A,B) and partial selfing with inbreeding depression using the highest field-estimate of selfing and inbreeding depression parameters ($\delta_i$) calculated for the Iron Mountain population of \citet{Willis1993} (panels C,D).
+Invasion conditions and equilibrium densities ($N^{\ast}$, for $\alpha = 10^{-4}$) were calculated using the conservative assumption of additive fitness effects in both sexes ($h_f = h_m = 0.5$) for two parameter conditions: obligate outcrossing (panels A,B) and partial selfing with inbreeding depression using the highest field-estimate of selfing and inbreeding depression parameters ($\delta_i$) calculated for the Iron Mountain population of \citet{Willis1993} (panels C,D).
 The location of inv6 is also shown on both plots, using selection coefficients calculated from field estimates of male and female fitness components from \citet{LeeKelly2015} under the relatively conservative assumption of additive fitness effects in both sexes, where $(s_f,\,s_m) = (0.30,\, 0.31)$.}}
@@
 
-\subsection{Demographic data}
+\subsection*{Demographic data}
 
@@
 
-\subsection{Selection on inv6}
+\subsection*{Selection on inv6}
 

### 2021-12-29 R/functions-Figs.R

@@ -985,8 +985,8 @@ resultsIllustrationFig  <-  function() {
 #        proportionalLabel(0.6, 0.9, expression(paste("Extinction")), cex=1.5, adj=c(0.5, 0.5), xpd=NA, srt=0)
 #        proportionalArrows(px1=0.6, py1=0.85, px2=0.9, py2=0.6, length=0.1)
 #        proportionalLabel(0.4, 0.35, expression(paste("Exp. Growth (", lambda," > 1)")), cex=1.5, adj=c(0.5, 0.5), xpd=NA, srt=0)
-        proportionalLabel(0.45, 0.5, expression(paste("Exp. Growth")), cex=1.5, adj=c(0.5, 0.5), xpd=NA, srt=0)
-        proportionalLabel(0.45, 0.425, expression(paste("(", lambda," > 1)")), cex=1.5, adj=c(0.5, 0.5), xpd=NA, srt=0)
+        proportionalLabel(0.55, 0.425, expression(paste("Exp. Growth")), cex=1.5, adj=c(0.5, 0.5), xpd=NA, srt=0)
+        proportionalLabel(0.55, 0.35, expression(paste("(", lambda," > 1)")), cex=1.5, adj=c(0.5, 0.5), xpd=NA, srt=0)
 
     # Color ScaleBar
     image.scale(A_f6.2, col=HEAT, breaks=breaks, horiz=FALSE, yaxt="n")

### 2021-12-29 run-Simulations.R

@@ -504,7 +504,33 @@ makeDataDeltaPolyParamSpace(sMax=0.15, res=0.003, precision = 1e-4,
 							om = 2, g = 3, theta = c(0.6,0.6,0.05,8.5), theta_prime = 8.5, 
 							hVals= c(1/2, 1/4), dStar = 0.8, 
 							tlimit = 10^5, eqThreshold = 1e-8)
-
+ 
+# Cs[11]
+# deltaSeq  <-  predDelta(dStar=dStar, b=1/2, a=0.2, C=Cs)
+ #
+# j=11
+# invData_f8.5  <-  titrateInvBoundaries(sMax=0.15, res=0.003, precision = 1e-4, alpha = 0,
+# 					 om = 2, g = 3, theta = c(0.6,0.6,0.05,8.5), theta_prime = 8.5, 
+# 					 hf = 1/2, hm = 1/2, C = Cs[j], 
+# 					 delta = 0, delta_j = deltaSeq[j], delta_a = 0, delta_gamma = 0,
+# 					 tlimit = 10^5, eqThreshold = 1e-8, verbose=TRUE, writeFile=FALSE)
+# invData_f6.5  <-  titrateInvBoundaries(sMax=0.15, res=0.003, precision = 1e-4, alpha = 0,
+# 					 om = 2, g = 3, theta = c(0.6,0.6,0.05,8.5), theta_prime = 6.5, 
+# 					 hf = 1/2, hm = 1/2, C = Cs[j], 
+# 					 delta = 0, delta_j = deltaSeq[j], delta_a = 0, delta_gamma = 0,
+# 					 tlimit = 10^5, eqThreshold = 1e-8, verbose=TRUE, writeFile=FALSE)
+ #
+# pgAinv  <-  popGen_A_invade_Delta_Add(C=Cs[j], delta=deltaSeq[j], sm=invData$sms)
+# pgainv  <-  popGen_a_invade_Delta_Add(C=Cs[j], delta=deltaSeq[j], sm=invData$sms)
+ #
+# plot(pgAinv ~ invData_f8.5$sms, type='l', lwd=2, col=1, ylim=c(0,0.15))
+# lines(pgainv ~ invData_f8.5$sms, lwd=2, col=1)
+ #
+# lines(invData_f8.5$AInvBound ~ invData_f8.5$sms, lwd=2, col=2)
+# lines(invData_f8.5$aInvBound ~ invData_f8.5$sms, lwd=2, col=2)
+ #
[…]

### 2021-12-30 doc/SA-Hermaphrodites-wDemography-R1.tex
Updated table 1 and added one sentence about maturation rate

@@
 \centering
-\caption{\bf Definition of terms.}
+\caption{\bf Mathematical notation and parameter values used in this article.}
 \label{tab:Terms}
@@
  \toprule
+\multicolumn{3}{l}{Definition of symbols used in the general model description} \\
+
 Symbol & Definition & Dimension \\
@@
 $\tilde{\mbf{A}}(\tilde{\mbf{p}})$ & Population projection matrix & $2 \omega g \times 2\omega g$ \\
-
-$s_f,\,s_m$ & Selection coefficients through female and male reproductive function & \\
-$h_f,\,h_m$ & Dominance coefficients through female and male reproductive function & \\
-$\sigma_j,\, \sigma_a$ & Survival rates for juvenile and adult stages & \\
-$\gamma$ & Transition rate from juvenile to adult stages \\
-$C$ & The population rate of self-fertilization & \\
-$\delta$ &\textcolor{blue}{ Inbreeding depression effect on ovule viability}&\\
-$ \delta_j$ & \textcolor{blue}{ Inbreeding depression effect on juvenile survival}& \\
-$\delta_a$ & \textcolor{blue}{ Inbreeding depression effect on adult survival}& \\
-$\delta_{\gamma}$ &\textcolor{blue}{Inbreeding depression effect on maturation rate} & \\
-$f_i,\,f^{\prime}_i$ & Adult fertilities through female and male sex functions respectively & \\
+\hline
+\multicolumn{3}{l}{Parameters used in the two stage example} \\
+Symbol & Definition & Value/Range\\
+\hline
+$s_f,\,s_m$ & Selection coefficients through female and male reproductive function & 0-0.15\\
+$h_f,\,h_m$ & Dominance coefficients through female and male reproductive function & $\frac{1}{2}$ or $\frac{1}{4}$\\
+$\sigma_j,\, \sigma_a$ & Survival rates for juvenile and adult stages & 0.6\\
+$\gamma$ & Transition rate from juvenile to adult stages & 0.05\\
+$C$ & The population rate of self-fertilization &0-1 \\
+$\delta$ &\textcolor{blue}{ Inbreeding depression effect on ovule viability}&0-0.8\\
+$ \delta_j$ & \textcolor{blue}{ Inbreeding depression effect on juvenile survival}& 0-0.8\\
+$\delta_a$ & \textcolor{blue}{ Inbreeding depression effect on adult survival}& 0-0.8\\
+$\delta_{\gamma}$ &\textcolor{blue}{Inbreeding depression effect on maturation rate} & 0-0.8\\
+$f_i,\,f^{\prime}_i$ & Adult fertilities through female and male sex functions respectively & 5.8-8.5 \\
 
@@
 \noindent where $f$ and $f^{\prime}$ represent \textcolor{blue}{maximum adult fertilities}, and $w_{i}$ and $w^{\prime}_{i}$ the genotypic relative scaling factors for female and male sex-functions (see Table \ref{tab:Fitness}).
+
+Unless stated otherwise, we use the following parameter values for the demographic rates in the model: $\sigma_j = \sigma_a = 0.6$ and $\gamma = 0.05$.
+These chosen values are similar to those used in \citet{deVriesCaswell2019b}, facilitating comparison between models, and correspond to a life-history in which individuals spend multiple timesteps in the juvenile phase prior to maturing into reproductively active adults, but are otherwise arbitrary.
+Our parameters of interest include \textcolor{blue}{maximum female} fertility, $f$, the inbreeding depression parameters, $\delta$, $\delta_j$, $\delta_a$, $\delta_{\gamma}$, and the selection parameters $h_f$, $s_f$, $h_m$, and $s_m$, which are given different values for each analyses as described in the figure captions.
 
@@
 
-Unless stated otherwise, we use the following parameter values for the demographic rates in the model: $\sigma_j = \sigma_a = 0.6$ and $\gamma = 0.05$.
-These chosen values are similar to those used in \citet{deVriesCaswell2019b}, facilitating comparison between models, and correspond to a life-history in which individuals spend multiple timesteps in the juvenile phase prior to maturing into reproductively active adults, but are otherwise arbitrary.
-Our parameters of interest include \textcolor{blue}{maximum female} fertility, $f$, the inbreeding depression parameters, $\delta$, $\delta_j$, $\delta_a$, $\delta_{\gamma}$, and the selection parameters $h_f$, $s_f$, $h_m$, and $s_m$, which are given different values for each analyses as described in the figure captions.
-
 % Notational note: The superscript on a matrix in age$\times$stage models conventionally indicates the stage or age of the subsection of the population vector that the matrix is acting on (see for example \citep{CaswellEtAl2018}).
@@
 Second, we assume that if inbreeding depression is caused primarily by recessive deleterious mutations (as suggested by empirical data), it should covary negatively with the population selfing rate due to purging, provided the population selfing rate has been relatively constant in recent evolutionary time (\citealt{Charlesworth2009}; though we note that other processes could give rise to this pattern, e.g., \citealt{CrnokrakBarrett2002, Charlesworth2009,HedrickGarcia-Dorado2016}).
-Following \citet{Olito2019}, we incorporate such negative covariance by constraining the inbreeding depression terms in the model ($\delta$ and $\delta_i$, where $i \in \{j,a,\gamma\}$) to follow a simple declining function of the selfing rate: \textcolor{blue}{$\delta = \delta^{\ast} (1 - b (1 - a(1 - C)/(C + a(1 - C))))$}, where $\delta^{\ast}$ is the hypothetical severity of inbreeding depression for a completely outcrossing population, $b$ is a shape parameter determining how far $\delta$ will decline under complete selfing (when $C = 1$), \textcolor{blue}{and $a$ is a shape parameter} controlling the curvature of the overall function for $\delta$ (see Appendix E in \citealt{Olito2019} for additional details).
+Following \citet{Olito2019}, we incorporate such negative covariance by constraining the inbreeding depression terms in the model ($\delta$ and $\delta_i$, where $i \in \{j,a,\gamma\}$) to follow a simple declining function of the selfing rate: \textcolor{blue}{
+\begin{equation}
+  \delta = \delta^{\ast} \left(1 - b \left(1 - \frac{a(1 - C)}{(C + a(1 - C))}\right)\right),
+  \label{eq:Canddelta}
+\end{equation}}
+where $\delta^{\ast}$ is the hypothetical severity of inbreeding depression for a completely outcrossing population, $b$ is a shape parameter determining how far $\delta$ will decline under complete selfing (when $C = 1$), \textcolor{blue}{and $a$ is a shape parameter} controlling the curvature of the overall function for $\delta$ (see Appendix E in \citealt{Olito2019} for additional details).
 We set $\delta^{\ast} = 0.8$, $b = 0.5$, and $a = 0.2$ for all analyses involving inbreeding depression, values chosen to be consistent with empirical estimates of inbreeding depression (e.g., fig.~2 in \citealt{HusbandSchemske1996}).
@@
 This causes a greater proportion of SA polymorphic parameter space to be demographically viable, with the demographic model predictions converging on those from population genetic models in high-fertility populations (fig.~\ref{fig:deltaPolySpace}).
-For populations with lower fertility rates, the proportion of demographically viable polymorphic parameter space \textcolor{blue}{is very similar to} the population genetic model predictions \textcolor{blue}{for low selfing rates}.
 \textcolor{blue}{Note, however, that in our model, mortality due to late-acting inbreeding depression also reduces the population effective selfing rate, and therefore shifts the invasion boundaries for the SA alleles to more closely resemble an outcrossing population.
@@
 However, as the selfing rate increases, demographic viability eventually crashes when the population can no longer sustain the concomitant increase in mortality due to inbreeding depression (fig.~\ref{fig:deltaPolySpace}).
+For populations with lower fertility rates, the proportion of demographically viable polymorphic parameter space \textcolor{blue}{is very similar to} the population genetic model predictions \textcolor{blue}{for low selfing rates}.
 
@@
 Inbreeding effects on juvenile survival had the strongest effect on population viability because on average individuals will spend multiple time steps in the juvenile stage before they mature.
-At each time, juveniles have a probability $\gamma$ to mature and a probability $\sigma_j(1-\delta_j)$ to survive.
+At each time step, juveniles have a probability $\gamma$ to mature and a probability $\sigma_j(1-\delta_j)$ to survive.
 Inbreeding depression at the juvenile stage therefore makes it harder to survive long enough to mature.
-Although early-acting inbreeding depression ($\delta$) actually manifests earlier in the life-cycle than juvenile survival, it acts only once by influencing the total number of self-fertilized zygotes that become juveniles.
+\textcolor{blue}{Of course the strength of this effect is influenced by the maturation rate, $\gamma$, which therefore also affects the shape of the curves in Figure \ref{fig:deltaPolySpace}.} Although early-acting inbreeding depression ($\delta$) actually manifests earlier in the life-cycle than juvenile survival, it acts only once by influencing the total number of self-fertilized zygotes that become juveniles.
 Note, however, that the relative strength of inbreeding depression at different stages of the life history will depend on the interaction between the various inbreeding depression parameters, which we have precluded from our analyses.

### 2021-12-31 doc/SA-Hermaphrodites-wDemography-R1.tex
Shortened introduction
Intro magic

@@
 
-\textit{Manuscript elements}: figure~1, figure~2, figure~3, figure~4, table~1, table~2, appendices~A, B, and C, Online Supplementary Material. All figures are to print in color.
+\textit{Manuscript elements}: figure~1, figure~2, figure~3, figure~4, figure~5, table~1, table~2, Appendices~A, B, C, and D, Online Supplementary Material. All figures are to print in color.
 
@@
 
-\textcolor{blue}{When selection differs between the sexes, genes expressed by both males and females can experience sexually antagonistic (SA) selection, where beneficial alleles for one sex are deleterious for the other.} Classic population genetics theory has been fundamental to understanding the evolution of sex-differences and the maintenance of such SA genetic polymorphisms, but these models have rarely considered the demographic consequences \textcolor{blue}{ of the maladaptation imposed by coexisting alleles with deleterious fitness effects in each sex.} In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (i.e., dioecious/gonochorous) and partially selfing hermaphrodite populations.
+\textcolor{blue}{When selection differs between the sexes, genes expressed by both males and females can experience sexually antagonistic (SA) selection, where beneficial alleles for one sex are deleterious for the other.} Classic population genetics theory has been fundamental to understanding the evolution of sex-differences and the maintenance of such SA genetic polymorphisms, but these models have rarely considered the demographic consequences \textcolor{blue}{ of coexisting alleles with deleterious fitness effects in each sex.} In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (i.e., dioecious/gonochorous) and partially selfing hermaphrodite populations.
 We focus on identifying when SA polymorphisms are maintained by balancing selection {\itshape and} the population growth rate remains positive.
@@
 
-To persist in the long term, a population must be able to adapt to its environment.
+%To persist in the long term, a population must be able to adapt to its environment.
 Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, such as deleterious mutations \citep{Haldane1937}, \textcolor{blue}{rapid, unpredictable, or extreme changes} in environmental conditions \citep{Maynard-Smith1976, LandeShannon1996, OrrUnckless2008,Lyberger-etal-2021}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, and genetic constraints \citep{ConnallonHall2018, MatthewsConnallon2019}.
 Genetic constraints on adaptation can arise as a consequence of conflicting selection and gene-flow between different classes of individuals, or more generally, when \textcolor{blue}{different fitness components are negatively correlated} \citep{CharlesworthHughes2000, ConnallonHall2018}.
-
-\textcolor{blue}{Genetic constraints on adaptation arising from conflicting selection and gene flow} have particularly interesting evolutionary consequences \citep{ConnallonDebarreLi2018}.
+\textcolor{blue}{A population's ability to adapt to its environment and persist in the long term depends upon the nature of the genetic variation that it harbors \citep{Fisher1930, Lewontin1975}.
+Yet, species exist in a complex world where the process of adaptation can be impeded by conflicting selection and gene-flow between different classes of individuals, or more generally, when different fitness components are negatively correlated \citep{CharlesworthHughes2000, ConnallonHall2018}.}
+
+Genetic constraints on adaptation due to conflicting selection have particularly interesting evolutionary consequences \citep{ConnallonDebarreLi2018}.
+On one hand, they provide an effective mechanism for the maintenance of genetic variation, which shapes a population's capacity for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
+On the other hand, they prevent individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase a population's overall extinction risk \citep{kokko2003sexy, harts2014demography}.
+Hence, for traits under conflicting selection, the nature and extent of genetic variation observed in natural populations should reflect a balance between the maintainance of genetic polymorphisms, and the population dynamical consequences of the resulting maladaptation.
+
+%\textcolor{blue}{Genetic constraints on adaptation arising from conflicting selection and gene flow} have particularly interesting evolutionary consequences \citep{ConnallonDebarreLi2018}.
 On one hand, they provide an effective mechanism for the maintenance of genetic variation, which can increase a population's capacity for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
@@
 Sexually antagonistic selection (abbreviated SA hereafter) \textcolor{blue}{is an important form of conflicting selection that occurs when beneficial alleles for one sex are deleterious when expressed in the other.
+Likewise, alleles with opposing fitness effects through male and female sex functions can cause SA selection in hermaphrodite populations, where individuals' fitness is determined jointly by their maternal and paternal reproductive success \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
 SA selection appears to be} a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and the evolution of sexual dimorphism and sex-related traits \citep{Kidwell1977, Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, ConnallonClark2012, BondurianskyChenoweth2009, Olito2019}.
 \textcolor{blue}{For instance, phenotypic selection studies commonly find sex-differences in directional selection \citep[e.g.,][]{DeLisle-etal-2018,Lewis-etal-2011,CoxCalsbeek2009,SinghPunzalan2018}, implying that many genetic variants underlying quantitative traits have SA fitness effects.
-Likewise, quantitative genetic studies frequently find negative cross-sex genetic correlations for fitness (\citealt{Chippendale-etal-2001,Delph-etal-2011}; reviewed in \citealt{ConnallonMatthews2019}), and negative correlationpolygenic signals of SA selection are evident in recent population genetic analyses of large genomic datasets \citep{RuzickaESEB2020, RuzickaConnallon2022}.}
-
-Alleles with opposing fitness effects through male and female sex functions can also cause SA selection in hermaphrodite populations, where individuals' overall fitness is determined jointly by their maternal and paternal reproductive success \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
+Likewise, quantitative genetic studies frequently find negative cross-sex genetic correlations for fitness (\citealt{Chippendale-etal-2001,Delph-etal-2011}; reviewed in \citealt{ConnallonMatthews2019}), and polygenic signals of SA selection are evident in recent population genetic analyses of large genomic datasets \citep{RuzickaESEB2020, RuzickaConnallon2022}.
+Overall, both theory and current empirical data suggest that there is ample scope for the maintenance of SA \textcolor{blue}{genetic variation} in both dioecious \textcolor{blue}{(separate-sexed)} and hermaphrodite populations \citep{BondurianskyChenoweth2009, Abbott2011, RuzickaESEB2020}.}
+
+% Sexually antagonistic selection (abbreviated SA hereafter) \textcolor{blue}{is an important form of conflicting selection that occurs when beneficial alleles for one sex are deleterious when expressed in the other.
+SA selection appears to be} a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and the evolution of sexual dimorphism and sex-related traits \citep{Kidwell1977, Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, ConnallonClark2012, BondurianskyChenoweth2009, Olito2019}.
+\textcolor{blue}{For instance, phenotypic selection studies commonly find sex-differences in directional selection \citep[e.g.,][]{DeLisle-etal-2018,Lewis-etal-2011,CoxCalsbeek2009,SinghPunzalan2018}, implying that many genetic variants underlying quantitative traits have SA fitness effects.
+Likewise, quantitative genetic studies frequently find negative cross-sex genetic correlations for fitness (\citealt{Chippendale-etal-2001,Delph-etal-2011}; reviewed in \citealt{ConnallonMatthews2019}), and polygenic signals of SA selection are evident in recent population genetic analyses of large genomic datasets \citep{RuzickaESEB2020, RuzickaConnallon2022}.}
+
+% Alleles with opposing fitness effects through male and female sex functions can also cause SA selection in hermaphrodite populations, where individuals' overall fitness is determined jointly by their maternal and paternal reproductive success \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
 \textcolor{blue}{In obligately outcrossing hermaphrodite populations, SA selection operates similarly as in seperate-sexed ones (dioecious species).
@@
 
-\textcolor{blue}{While significant efforts have been made to understand when balanced SA polymophisms can be maintained and to detect SA genetic variation from empirical data, much less consideration has been given to the demographic consequences of the maladaptation caused by SA selection}.
-Most population genetic models, including models of SA selection, model the relative fitness of genotypes in populations of constant (and often infinite) size, and do not consider the population dynamical consequences of evolutionary change.
+\textcolor{blue}{While significant efforts have been made to predict when SA polymorphisms should be maintained and to detect SA genetic variation from empirical data, less attention has been paid to the demographic consequences of SA selection \citep{MatthewsConnallon2019}}.
+Most population genetic theory, including SA selection theory, models the relative fitness of genotypes in populations of constant (and often infinite) size.
 In reality, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals \textcolor{blue}{which link genotype distributions to population dynamics (e.g., \citealt {johnston2013life, merot2020balancing}).
-Over the past 50 years, the field of eco-evolutionary dynamics has sought to jointly model the genetic and demographic processes involved in evolutionary change.
-In particular, theoretical models combining population genetics with density-dependent population growth \citep{roughgarden1971density} and stage-structured populations (consolidated in \citealt{Charlesworth_1994}) paved the way toward understanding the demographic changes accompanying allele frequency changes due to selection.
-Extensions of these theories to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and quantitative traits \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects} all highlight that a fuller understanding of the genetic basis of fitness variation, and the processes that maintain it, requires an explicit treatment of the demographic consequences of evolutionary change.}
+For over 50 years, the field of eco-evolutionary dynamics has sought to jointly model the genetic and demographic processes involved in evolutionary change.
+Theoretical models combining population genetics with density-dependent population growth \citep{roughgarden1971density} and stage-structured populations (consolidated in \citealt{Charlesworth_1994}) paved the way toward understanding the demographic consequences of allele frequency changes due to selection.
+Various extensions of these theories have been made to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and the evolution of quantitative traits \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects}.
+Together, these eco-evolutionary models highlight that a fuller understanding of the genetic basis of fitness variation, and the processes maintaining it, requries an integration of genetic and demographic evolutionary processes.}
+
+%a fuller understanding of the genetic basis of fitness variation, and the processes maintaining it, requires an explicit treatment of the demographic processes involved in evolutionary change.
 
 The potential demographic costs of sexual antagonism were pointed out by \cite{kokko2003sexy}, but few studies since have explicitly incorporated demography into models of sexual antagonism (with the notable exception of \citealt{harts2014demography}).
-\citet{deVriesCaswell2019a} introduced a mendelian matrix model with intralocus sexual antagonism and population dynamics, but did not analyze the consequences of demographic viability for the scope of sexual antagonism to maintain polymorphism.
-In this paper, we extend the model of \citet{deVriesCaswell2019a,deVriesCaswell2019b} to include both dioecious and hermaphroditic \textcolor{blue}{populations} in order to study when different genetic outcomes of SA selection, including fixation of male- or female-beneficial alleles as well as balanced polymorphisms, are also demographically viable.
-A major strength of our approach is that the model can be parameterized using empirically estimated demographic rates, enabling us to make predictions about the scope for demographically viable SA polymorphisms that are grounded in the biology of real populations.
+\citet{deVriesCaswell2019a} introduced a mendelian matrix model with SA selection and population dynamics, but did not analyze how demography influenced the maintenance of SA polymorphism.
+In this paper, we extend the model of \citet{deVriesCaswell2019a,deVriesCaswell2019b} to \textcolor{blue}{include both dioecious and hermaphroditic populations and jointly analyze the genetic (fixation of male- or female-beneficial alleles, or balanced polymorphism) and demographic (population persistence or extinction) outcomes of SA selection}.
+A major strength of our approach is that the model can be parameterized using empirically estimated demographic rates, enabling us to make predictions about the scope for SA polymorphisms that are grounded in the biology of real populations.
 We demonstrate this with a case study of {\itshape Mimulus guttatus} (now {\itshape Erythranthe guttata}).

### 2021-12-31 doc/SA-Hermaphrodites-wDemography-R1.tex
Minor tweaks to intro
Realized I didn't like that last sentence about eco-evo, so edited.

@@
 \textcolor{blue}{A population's ability to adapt to its environment and persist in the long term depends upon the nature of the genetic variation that it harbors \citep{Fisher1930, Lewontin1975}.
-Yet, species exist in a complex world where the process of adaptation can be impeded by conflicting selection and gene-flow between different classes of individuals, or more generally, when different fitness components are negatively correlated \citep{CharlesworthHughes2000, ConnallonHall2018}.}
+Yet, species exist in a complex world where the process of adaptation can be impeded by conflicting selection and gene-flow between different classes of individuals, or more generally, when different fitness components are negatively correlated \citep{Prout2000, CharlesworthHughes2000, ConnallonHall2018}.}
 
@@
 Likewise, quantitative genetic studies frequently find negative cross-sex genetic correlations for fitness (\citealt{Chippendale-etal-2001,Delph-etal-2011}; reviewed in \citealt{ConnallonMatthews2019}), and polygenic signals of SA selection are evident in recent population genetic analyses of large genomic datasets \citep{RuzickaESEB2020, RuzickaConnallon2022}.
-Overall, both theory and current empirical data suggest that there is ample scope for the maintenance of SA \textcolor{blue}{genetic variation} in both dioecious \textcolor{blue}{(separate-sexed)} and hermaphrodite populations \citep{BondurianskyChenoweth2009, Abbott2011, RuzickaESEB2020}.}
+Overall, both theory and current empirical data suggest that there is ample scope for the maintenance of SA \textcolor{blue}{genetic variation} in both dioecious \textcolor{blue}{(separate-sexed)} and hermaphrodite populations \citep{BondurianskyChenoweth2009, Abbott2011, RuzickaESEB2020, WangBarrett2020}.}
 
@@
 Various extensions of these theories have been made to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and the evolution of quantitative traits \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects}.
-Together, these eco-evolutionary models highlight that a fuller understanding of the genetic basis of fitness variation, and the processes maintaining it, requries an integration of genetic and demographic evolutionary processes.}
+This growing body of eco-evolutionary theory highlights that demographic processes must be taken into account explicitly in order to fully understand how genetic variation for fitness is shaped and maintained in natural populations.}
 

### 2022-01-03 doc/SA-Hermaphrodites-wDemography-R1.tex
more minor tweaks

@@
 Genetic constraints on adaptation can arise as a consequence of conflicting selection and gene-flow between different classes of individuals, or more generally, when \textcolor{blue}{different fitness components are negatively correlated} \citep{CharlesworthHughes2000, ConnallonHall2018}.
-\textcolor{blue}{A population's ability to adapt to its environment and persist in the long term depends upon the nature of the genetic variation that it harbors \citep{Fisher1930, Lewontin1975}.
-Yet, species exist in a complex world where the process of adaptation can be impeded by conflicting selection and gene-flow between different classes of individuals, or more generally, when different fitness components are negatively correlated \citep{Prout2000, CharlesworthHughes2000, ConnallonHall2018}.}
+\textcolor{blue}{A population's ability to adapt to its environment and persist in the long term depends upon the genetic variation that it carries \citep{Fisher1930, Lewontin1975}.
+Yet, populations are heterogenous collections of individuals existing in a complex world where adaptation can be impeded by conflicting selection and gene-flow between different classes of individuals, or more generally, when different fitness components are negatively correlated \citep{Prout2000, CharlesworthHughes2000, ConnallonHall2018}.}
 
@@
 \textcolor{blue}{For instance, phenotypic selection studies commonly find sex-differences in directional selection \citep[e.g.,][]{DeLisle-etal-2018,Lewis-etal-2011,CoxCalsbeek2009,SinghPunzalan2018}, implying that many genetic variants underlying quantitative traits have SA fitness effects.
-Likewise, quantitative genetic studies frequently find negative cross-sex genetic correlations for fitness (\citealt{Chippendale-etal-2001,Delph-etal-2011}; reviewed in \citealt{ConnallonMatthews2019}), and polygenic signals of SA selection are evident in recent population genetic analyses of large genomic datasets \citep{RuzickaESEB2020, RuzickaConnallon2022}.
+Moreover, quantitative genetic studies frequently find negative cross-sex genetic correlations for fitness (\citealt{Chippendale-etal-2001,Delph-etal-2011}; reviewed in \citealt{ConnallonMatthews2019}), and polygenic signals of SA selection are evident in recent population genetic analyses of large genomic datasets \citep{RuzickaESEB2020, RuzickaConnallon2022}.
 Overall, both theory and current empirical data suggest that there is ample scope for the maintenance of SA \textcolor{blue}{genetic variation} in both dioecious \textcolor{blue}{(separate-sexed)} and hermaphrodite populations \citep{BondurianskyChenoweth2009, Abbott2011, RuzickaESEB2020, WangBarrett2020}.}
@@
 Various extensions of these theories have been made to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and the evolution of quantitative traits \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects}.
-This growing body of eco-evolutionary theory highlights that demographic processes must be taken into account explicitly in order to fully understand how genetic variation for fitness is shaped and maintained in natural populations.}
+This growing body of eco-evolutionary theory makes plain that demographic processes must be taken into account explicitly to fully understand how genetic variation for fitness is shaped and maintained in natural populations.}
 
@@
 
-\textcolor{blue}{One particularly important limitation of our analysis is that we modeled density-independent growth.
+\textcolor{blue}{One particularly important limitation of our main analysis is that we modeled density-independent growth.
 Of course, in natural populations density-dependent processes like resource competition or competition for breeding sites will often prevent populations from growing (or shrinking) exponentially.
-In \ref{App:DensDep} we extend our model to include a simple form of density-dependence that acts on the survival of all life history stages equally, and reanalyse our case study of \textit{M.
+In the simple density-dependent version of our model presented in \ref{App:DensDep}, we included density-dependence acting on the survival of all life history stages equally, and reanalysed our case study of \textit{M.
 guttatus}.
 The density-dependent version of the model predicts equilibrium population densities rather than population intrinsic growth rates (see fig.~ \ref{fig:mimulusFig_dd}).
-This brief example shows that the shape of the polymorphic region is unchanged by this kind of density-dependence, and that the shape of the population density surface in fig.~\ref{fig:mimulusFig_dd} is very similar to that for population growth rate in fig.~\ref{fig:mimulusFig}.
-The similarity between the density-independent and -dependent model predictions underscores an important point: density-dependence that acts age-independently does not alter selection gradients \citep{mylius1995evolutionarily, caswell2017}.}
+This brief example shows that the shape of the polymorphic region is unchanged by this form of density-dependence, and that the shape of the population density surface in fig.~\ref{fig:mimulusFig_dd} is very similar to that for population growth rate in fig.~\ref{fig:mimulusFig}.
+The similarity between the density-independent and -dependent model predictions underscores an important point: density-dependence that acts (st)age-independently does not alter selection gradients \citep{mylius1995evolutionarily, caswell2017}.}
 
@@
 For example, competition for breeding sites will cause a reduction in fecundity rather than survival.
-Likewise, in plants density-dependent competition for soil nutrients will likely influence seed germination rates differently than seedling growth rates or flower/fruit production in adults \citep{AntonovicsLevin1980}.
+In plants, density-dependent competition for soil nutrients will likely influence seed germination rates differently than seedling growth rates or flower/fruit production in adults \citep{AntonovicsLevin1980}.
 These types of age- or stage-specific density-dependence will change the selection gradients, and therefore the shape of polymorphic parameter space as well as the equilibrium population densities predicted by the model.
-\citet{de2020matrix} discuss some of the possible dynamical outcomes of a density-dependent genetic model for \textit{Tribolium} seed beetles but a thorough exploration of the consequences of more complicated forms of density-dependence for the maintenance of sexually antagonistic polymorphisms remains an important avenue for future work.}
+\citet{de2020matrix} discuss some of the possible dynamical outcomes of a density-dependent genetic model for \textit{Tribolium} seed beetles but a thorough analysis of more complicated forms of density-dependence for the maintenance of SA polymorphisms remains an important avenue for future work.}
 
@@
 It would be an interesting and feasible extension of our modeling framework to expand the individual state space to include selfing cohorts (i.e., first generation selfing, second generation selfing, etc.), as in the models of \citet{kelly1999response,kelly2007mutation}, thereby enabling a more biologically plausible approach to modelling self-fertilization.
-\textcolor{blue}{Likewise, we implemented a simple "fixed prior selfing" model, where all individuals in the population self-fertilize at the same rate, which is constant over time.
-In reality, environmental stochasticity will likely influence the population selfing rate and therefore both the expression of inbreeding depression and the scope for SA polymorphism \citep{JordanConnallon2014,Olito-etal-2018}.
-In this scenario, a naive expectation might be that the geometric mean selfing rate over time would represent a more biologically meaningful parameter influencing the genetic and demographic outcomes of SA selection.}
+\textcolor{blue}{Likewise, the simple 'fixed prior selfing' model we implemented assumes constant selfing rates among individuals over time.
+In reality, environmental stochasticity will influence the population selfing rate and therefore both the expression of inbreeding depression and the scope for SA polymorphism \citep{JordanConnallon2014,Olito-etal-2018}.
+A naive expectation might be that some form of weighted geometric mean selfing rate would represent a more biologically meaningful parameter influencing the genetic and demographic outcomes of SA selection in more complicated 'selfing environments' \citep[e.g., ][]{ConnallonSharmaOlito2019}.}
 
@@
 
-\cite{grieshop2017male} found that genotypes with sexually antagonistic alleles that are male-beneficial experience higher levels of inbreeding depression than genotypes with female beneficial SA alleles.
+Finally, \cite{grieshop2017male} found that genotypes with SA alleles that are male-beneficial experience higher levels of inbreeding depression than genotypes with female beneficial SA alleles.
 This effect can easily be included in our model by making the inbreeding depression parameters, $\delta$ and $\delta_i$, a function of the genotype of the individual.

### 2022-01-03 doc/SA-Hermaphrodites-wDemography-R1.tex
Minor edits
Minor tweaks

@@
 Genetic constraints on adaptation can arise as a consequence of conflicting selection and gene-flow between different classes of individuals, or more generally, when \textcolor{blue}{different fitness components are negatively correlated} \citep{CharlesworthHughes2000, ConnallonHall2018}.
-\textcolor{blue}{A population's ability to adapt to its environment and persist in the long term depends upon the genetic variation that it carries \citep{Fisher1930, Lewontin1975}.
+\textcolor{blue}{A population's ability to adapt to its environment and persist in the long term depends on the genetic variation that it carries \citep{Fisher1930, Lewontin1975}.
 Yet, populations are heterogenous collections of individuals existing in a complex world where adaptation can be impeded by conflicting selection and gene-flow between different classes of individuals, or more generally, when different fitness components are negatively correlated \citep{Prout2000, CharlesworthHughes2000, ConnallonHall2018}.}
@@
 Various extensions of these theories have been made to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and the evolution of quantitative traits \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects}.
-This growing body of eco-evolutionary theory makes plain that demographic processes must be taken into account explicitly to fully understand how genetic variation for fitness is shaped and maintained in natural populations.}
+This growing body of eco-evolutionary theory makes plain that demographic processes must be taken into account explicitly to fully understand how genetic variation for fitness is shaped, and maintained in natural populations \citep{metcalf2007evolutionary,coulson2006putting,rueffler2006evolutionary}.}
 
@@
 where equality is obtained by normalizing the frequencies such that they add to one (see Online Supplementary Materials for a derivation of Eq(\ref{eq:maleGametePool})).
-The first entry sums the  stage-specific contributions to the gamete pool of heterozygotes and $AA$ homozygotes; the second entry sums the stage-specific contributions to the gamete pool of heterozygotes and $aa$ homozygotes.
+The first entry sums the  stage-specific contributions of allele A to the gamete pool by heterozygotes and $AA$ homozygotes; the second entry sums the stage-specific contributions of allele a to the gamete pool by heterozygotes and $aa$ homozygotes.
 }
@@
 		\end{array} \right) = 
-				\frac{1}{(1-s_m)p^R_{aa}+ (1-hs_m)p^R_{Aa}+p_{aa}}\left(
+				\frac{1}{(1-s_m)p^R_{aa}+ (1-hs_m)p^R_{Aa}+p^R_{aa}}\left(
 		\begin{array}{c}
@@
 
-We identify conditions where SA polymorphism is 'protected' by evaluating the stability of populations initially fixed for either SA allele to invasion by the other (i.e., we assessed stability at the boundary equilibrium genotype frequencies of $\hat{\mbf{p}}_{AA} = 1$ and $\hat{\mbf{p}}_{aa} = 1$; \citealt{Levene1953, Prout1968, deVriesCaswell2019b}).
+We identify conditions where SA polymorphism is 'protected' by evaluating the stability of populations initially fixed for either SA allele to invasion by the other allele (i.e., we assessed stability at the boundary equilibrium genotype frequencies of $\hat{\mbf{p}}_{AA} = 1$ and $\hat{\mbf{p}}_{aa} = 1$; \citealt{Levene1953, Prout1968, deVriesCaswell2019b}).
 The formal conditions for a protected polymorphism are determined by linearizing the model in the vicinity of the boundary equilibria ($\hat{\mbf{p}}$), and evaluating the magnitude of the largest eigenvalue of the Jacobian matrix of the linearization.
@@
 
-Mortality from inbreeding depression quickly outweighs the beneficial effect of selfing on the equilbirium female load that was apparent in the previous section.
+Mortality from inbreeding depression quickly outweighs the beneficial effect of selfing on the equilibrium female load that was apparent in the previous section.
 In contrast to our earlier results, predominantly outcrossing populations are predicted to have the highest proportion of demographically viable polymorphic space when inbreeding depression is taken into acount (compare fig.~\ref{fig:polySpace} with fig.~\ref{fig:deltaPolySpace}).
@@
 The first and central finding of our study is that when population intrinsic growth rates approach one, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space where SA polymorphism is maintained.
-Since intrinsic growth rates far exceeding one suggest rapid exponential growth, they are generally rare (with the notable exceptions of recently introduced invasive populations, \textcolor{blue}{and locally adapted ones}), suggesting that our model predictions may be highly relevant for SA polymorphism in many real-world populations.
-In addition, we find that much of the parameter space where a male-beneficial allele \textcolor{blue}{fixes} is demographically inviable for populations whose intrinsic growth rates are close to one prior to invasion.
+We find that much of the parameter space where a male-beneficial allele \textcolor{blue}{fixes} is demographically inviable for populations whose intrinsic growth rates are close to one prior to invasion.
+Since intrinsic growth rates far exceeding one suggest rapid exponential growth, they are generally rare (with the notable exceptions of recently introduced invasive populations), suggesting that our model predictions may be highly relevant for SA polymorphism in many real-world populations.
 This demographic consequence of masculinization has rarely been considered \citep{hitchcock2020gene}.
-Moreover, these findings complement recent theoretical and empirical studies indicating that SA selection is likely to be both condition dependent, and stronger in locally-adapted populations near the center of a species' range, where population growth rates are expected to be higher \citep{BergerEtAl2014, Connallon2015}.
+Moreover, these findings complement recent theoretical and empirical studies indicating that SA selection is likely to be both condition dependent, and stronger in locally-adapted populations near the center of a species' range, where population growth rates are expected to be higher and populations might be more likely to be able to suffer the demographic cost of a SA polymorphism while remaining viable \citep{BergerEtAl2014, Connallon2015}.
 
@@
 guttatus}.
-The density-dependent version of the model predicts equilibrium population densities rather than population intrinsic growth rates (see fig.~ \ref{fig:mimulusFig_dd}).
+The density-dependent version of the model predicts equilibrium population densities rather than population intrinsic growth rates (see fig.~\ref{fig:mimulusFig_dd}).
 This brief example shows that the shape of the polymorphic region is unchanged by this form of density-dependence, and that the shape of the population density surface in fig.~\ref{fig:mimulusFig_dd} is very similar to that for population growth rate in fig.~\ref{fig:mimulusFig}.
-The similarity between the density-independent and -dependent model predictions underscores an important point: density-dependence that acts (st)age-independently does not alter selection gradients \citep{mylius1995evolutionarily, caswell2017}.}
+The similarity between the density-independent and -dependent model predictions underscores an important point: density-dependence that acts (st)age-independently does not alter selection gradients \citep{mylius1995evolutionarily, caswell2017senescence}.}
 
@@
 \begin{equation}
-	\tilde{\mbf{n}}(t + 1) = \exp\left(-\alpha \sum_i n_i \right) \tilde{\mbf{A}}[\tilde{\mbf{n}}(t)] \, \tilde{\mbf{n}}(t),
+	\tilde{\mbf{n}}(t + 1) = \exp\left(-\alpha \sum_i n_i \right) \tilde{\mbf{A}}[\tilde{\mbf{p}}(t)] \, \tilde{\mbf{n}}(t),
 \end{equation}
@@
 \begin{equation}
-	\tilde{\mbf{n}}(t + 1) = \exp\left(-\alpha \| \tilde{\mbf{n}}(t) \|\right) \tilde{\mbf{A}}[\tilde{\mbf{n}}(t)] \, \tilde{\mbf{n}}(t),
+	\tilde{\mbf{n}}(t + 1) = \exp\left(-\alpha \| \tilde{\mbf{n}}(t) \|\right) \tilde{\mbf{A}}[\tilde{\mbf{p}}(t)] \, \tilde{\mbf{n}}(t),
 \end{equation}

### 2022-01-03 R/functions-Figs.R
Minor fig edits, added couple bibs

@@ -647,7 +647,7 @@ lambdaPolySpaceFig  <-  function() {
     lines(AInvBound ~ sms, lty=1, lwd=1.5, col=COLS$line, data=invA)
     # Labels/Annotations
     proportionalLabel(0.04, 1.075, 'D', cex=1.2, adj=c(0.5, 0.5), xpd=NA)
-    proportionalLabel(-0.3, 0.5, expression(paste("Selection through female reproductive function (", italic(s[f]), ")")), cex=1.5, adj=c(0.5, 0.5), xpd=NA, srt=90)
+    proportionalLabel(-0.3, 0.5, expression(paste("Selection on female reproductive function (", italic(s[f]), ")")), cex=1.5, adj=c(0.5, 0.5), xpd=NA, srt=90)
     proportionalLabel(-0.75, 0.5, expression(paste(italic(h), " = 1/2")), cex=2.0, adj=c(0.5, 0.5), xpd=NA, srt=90)
     proportionalLabel(-0.5, 0.5, expression(paste(italic(f), " = 6.0")), cex=1.5, adj=c(0.5, 0.5), xpd=NA, srt=90)
 
@@ -711,7 +711,7 @@ lambdaPolySpaceFig  <-  function() {
     lines(AInvBound ~ sms, lty=1, lwd=1.5, col=COLS$line, data=invB)
     # Labels/Annotations
     proportionalLabel(0.04, 1.075, 'H', cex=1.2, adj=c(0.5, 0.5), xpd=NA)
-    proportionalLabel(0.5, -0.3, expression(paste("Selection through male reproductive function (", italic(s[m]), ")")), cex=1.5, adj=c(0.5, 0.5), xpd=NA, srt=0)
+    proportionalLabel(0.5, -0.3, expression(paste("Selection on male reproductive function (", italic(s[m]), ")")), cex=1.5, adj=c(0.5, 0.5), xpd=NA, srt=0)
 
     # Panel 9: h = 1/2,  C = 1/2, f = 5.8
     C_f5.8 <- interp(x=heatC_f5.8$sm, y=heatC_f5.8$sf, z=heatC_f5.8$lambda_sim, 
@@ -802,7 +802,7 @@ plot(0, xaxt = 'n', yaxt = 'n', bty = 'n', pch = '', ylab = '', xlab = '')
     # Labels/Annotations
     proportionalLabel(0.04, 1.075, 'M', cex=1.2, adj=c(0.5, 0.5), xpd=NA)
[…]

### 2022-01-03 makeFigs.R

@@ -71,8 +71,8 @@ embed_fonts(figPath(name='MimulusInv6LambdaFig.pdf'))
 # DENSITY DEPENDENT DYNAMICS
 # Mimulus Case study: Nstar heatmap w/ Inv6
 toPdf(MimulusInv6NstarFig(), 
-			figPath(name='MimulusInv6NstarFig2.pdf'), width=8, height=8)
-embed_fonts(figPath(name='MimulusInv6NstarFig2.pdf'))
+			figPath(name='MimulusInv6NstarFig.pdf'), width=8, height=8)
+embed_fonts(figPath(name='MimulusInv6NstarFig.pdf'))
 
 
 

### 2022-01-05 doc/ResponseLetter-R1.tex
Finalized revisions, response letter, etc.

@@
 
-\noindent First of all, we want to thank the Ed., AE, and EiC for their thoughtful and forthright responses (and apologies) regarding the editorial delay on our manuscript.
-We understand completely that pandemic-related personal issues have been and continue to be extremely disruptive in these difficult times, and that editorial duties for our journals is a volunteer service.
-To our AE, Dr.~Orive, we sincerely hope that you and your family are well, and we truly appreciate the time and effort you took to evaluate our paper when you clearly had more imporant issues to deal with.
+\noindent First of all, we want to thank the Ed., AE, and EiC for their thoughtful and forthright responses regarding the editorial delay on our manuscript.
+We completely understand that pandemic-related personal issues have been and continue to be extremely disruptive in these difficult times, and that editorial duties for our journals is a volunteer service.
+To our AE, Dr.~Orive, we sincerely hope that you and your family are well, and we truly appreciate the time and effort you took to evaluate our paper when you clearly had more imporant things to deal with.
 \bigskip
@@
 \begin{quote}
-	{\itshape Thank you, again, for your candid and forthright apology and explanation.
-We truly appreciate it, and sympathize with the growing challenges that seem to creep up on us all during these difficult times.}
+	{\itshape Thank you, again, for being so candid and forthright. We truly appreciate it, and sympathize with the growing challenges that have crept up on us all during the pandemic.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape We agree entirely. Please see our detailed responses to Dr.~Orive and R1 on this count.}
+	{\itshape We agree entirely and have restructured our introduction accordingly. Please see our detailed responses to Dr.~Orive and R1 on this count.}
 \end{quote}
@@
 	{\itshape This is a very good and totally valid point.
-To be honest, we believe that a separate paper is necessary to deal adequately with the many consequences of density dependence for our model predictions.
+Honestly, we believe a separate paper is necessary to deal with the many consequences of density dependence for our model predictions.
 However, we also agree that sidestepping this issue completely isn't good enough.
-In the end, we have made two main changes to address density dependence.
-First, have added a new {\bf Appendix B} where we briefly explain how density dependence can be incorporated into the model, and reanalyze our {\itshape M.
-guttatus} example with a simple DD version of the model.
-Second, we have expanded our methods, results, and discussion to more fully address the implications of DD, with reference to our modified figures and the new Appendix ({\bf L.150--156; 478--481; 564--585}).}
+We have now made two main changes to address density dependence.
+First, have added a new {\bf Appendix B} where we briefly show how density dependence can be incorporated into the model, and reanalyze our {\itshape M.
+guttatus} example with a simple density-dependent version of the model.
+Second, we have expanded our methods, results, and especially the discussion to more fully address the implications of density-dependence, with reference to our modified figures and the new {\bf Appendix B} (see {\bf L.128--134; 444--447; 529--549}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape This is great suggestion.
-We have now included a new illustrative figure ({\bf fig.~1}), and revised versions of figs.
-1 \& 4 (now {\bf figs.~S1 \& 5}) to show heatmaps of absolute values for predicted $\lambda$ instead of the extinction boundaries.
-However, one unfortunate consequence of presenting $\lambda$ instead of the extinction boundaries is that it is no longer possible to illustrate multiple values of maximum fertility ($f$) on each figure panel.
+	{\itshape This is great suggestion which inspired us to make several major changes.
+We now include a new illustrative figure ({\bf fig.~1}), and revised versions of figs.
+1 \& 4 (now {\bf figs.~S1 \& 5}) which show heatmaps of absolute values for predicted $\lambda$ instead of just extinction boundaries.
+Unfortunately, when presenting $\lambda$ in this way, it is no longer possible to show multiple values of maximum fertility ($f$) on a single panel.
 In particular, our old fig.
 1 now requires 18 panels to illustrate $\lambda$ for the same set of parameter combinations.
-We felt that this resulted in an unnecessarily complicated figure, especially since fig.
-1's main purpose was to illustrate the general behaviour of the model for different fertility values.
+We felt that this resulted in an unnecessarily large and complicated figure.
 For simplicity, we have decided to retain our original version of fig.1 (now {\bf fig.
-2}) in the main text and include the modified version showing $\lambda$ as a supplementary figure (now {\bf fig.~S1}); but will swap this figure if the editor/reviewers feel differently.
-However, we agree that the {\itshape M.
-guttatus} example was greatly improved by showing $\lambda$, and we now present the modified version of fig.
+2}) in the main text and include the modified version showing 18 heatmaps for $\lambda$ as a supplementary figure (now {\bf fig.~S1}).
+We hope that, with the new illustrative {\bf fig.~1}, the extinction boundaries presented in {\bf fig.~2} will be easier for readers to understand.
+However, if the editors and reviewers prefer the full heatmaps, we will happily swap {\bf fig.~2} with {\bf fig.~S1}.
+Lastly, we agree that the {\itshape M.
+guttatus} example is greatly improved by showing $\lambda$, and we now present the modified version of fig.
 4 (now {\bf fig.
@@
 \begin{quote}
-	{\itshape We agree that in our efforts to appeal to both demographers and evolutionary biologists studying sexual conflict, we ended up insufficiently discussing the relevant literature on both counts (but especially regarding evolutionarily explicit structured models).
-We have made a variety of changes to our introduction to improve this ({\bf L.45--51, L.69--71, L.76--85}).
+	{\itshape We can see now that in our efforts to appeal to both demographers and evolutionary biologists studying sexual conflict, we insufficiently discussed the relevant literature on both counts.
+We have restructured our introduction to better engage with these fields without lengthening it (see, in particular, {\bf L.43--52, L.58--67, L.76--85}).
 See also our specific responses to Reviewer 1.
@@
 \begin{quote}
-	{\itshape We have added a brief general explanation of sexually antagonistic selection, which we felt was lacking in the first draft, and also allows us to more clearly link SA polymorphism with its' demographic consequences ({\bf L.2--3, L.6--7}).}
+	{\itshape We have added a brief general explanation of sexually antagonistic selection, which we felt was lacking in the first draft, and also allows us to more clearly link SA polymorphism with its' demographic consequences ({\bf L.2--6}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape We have changed our axes labels to be more descriptive }
+	{\itshape Good point. We have changed our axes labels to be more descriptive }
 \end{quote}
@@
 \begin{quote}
-	{\itshape We now briefly discuss this possibility in the Discussion ({\bf L.605--611}).}
+	{\itshape We now briefly discuss this possibility in the Discussion ({\bf L.564--570}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape This is a really important comment and we thank you for pointing out this major shortcoming.
-One of our main goals when crafting the first draft of the manuscript was to find a middle ground that would appeal to both demographers and people studying sexual conflict.
+	{\itshape This is a very important comment and we thank you for pointing out this major shortcoming.
+Our goal when crafting the first draft was to find a middle ground that would appeal to both demographers and people studying sexual conflict.
 It is clear that we didn't quite achieve that goal.
 Rather, by making compromises in the introduction we ended up lacking scholarship on both sides.
-To address your main concern, we have extended the Introduction to more directly connect with the broader literature on evolutionarily explicit structured models (see {\bf L.76--85}) and now also provide a more thorough link with the empirical sexually antagonistic selection literature (see {\bf L.45--52}).
-We hope these changes succeed in broadening the appeal of our manuscript to the wider AmNat readership.}
+To address your main concern, we have restructured the Introduction to more directly connect with the broader literature on evolutionarily explicit structured models (see {\bf L.58--67}) and now also provide a stronger link with the empirical sexually antagonistic selection literature (see {\bf L.43--51}).
+We hope these changes, along with a more streamlined presentation of SA selection (see {\bf L.34--39; 52--54}) succeed in broadening the appeal of our manuscript to the wider AmNat readership.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape Thank you for the encouraging assessment.}
+	{\itshape Thank you for the very encouraging assessment.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape While we feel that there is strong precedence for discussing the key results of a study before diving into the various caveats that come with the analysis, the point is well taken that we need to clearly direct readers to the places in our paper where these limitations are discussed.
-In our Methods section, we are now careful to point readers to the various Appendixes, Supplements, and places in the Discussion where these limitations are dealt with (e.g., see {\bf L.101--103, 114--119, 151--157, 215--216, fig.~2 caption}), and we have added more explicit discussion of several of our key assumptions in the discussion ({\bf e.g., L.565--586; 606--612}).
-We discuss both the causes and potential effects of incorporating male-limitation of reproductive success on {\bf L.587--599}}.
+	{\itshape While we feel that there is strong precedence for discussing the key results of a study before diving into the various caveats for with the analysis, yoru point is well taken that we need to more clearly acknowledge these limitations in the methods and direct readers to where these limitations are discussed.
+In our Methods section, we are now careful to point readers to the various Appendixes, Supplements, and places in the Discussion where these limitations are dealt with (e.g., see {\bf L.94--98, 103--109, 128--134, 183--186, 187--189, 196--197, 207--212, 274--275}), and we have added more explicit discussion of several of our key assumptions in the discussion ({\bf e.g., L.519--581}).
+We discuss both the causes and potential effects of incorporating male-limitation of reproductive success on {\bf L.550--561}}.
 \end{quote}
@@
 \begin{quote}
-	{\itshape \ul{Response to the above 5 points}: As per the reviewers (and the editors) suggestion, we have moved a lot of the derivations into the supplementary materials, and now no longer use H or W in the main text.
-In doing so, we hope to shift the focus from mathematical derivations, which are probably unhelpful to many readers, towards explaining the end result instead.
-However, we have made the requested changes to clarify the derivation in the Supplement.
+	{\itshape {\bf \ul{Response to the above 5 points}}: As per the reviewers (and the editors) suggestion, we have moved most of the derivations of the general model into the supplementary materials, and now no longer use H or W in the main text.
+In doing so, we hope to shift the focus from mathematical derivations towards explaining the end result instead.
+However, we have made the requested changes to clarify the derivation in {\bf Online Supplement 1}.
 }
@@
 \begin{quote}
-	{\itshape Very good points, that was awkward wording on our part.
-We have rephrased to more clearly explain that it is primarily the speed/magnitude of environmental change that we are referring to ({\bf L.23}).
-We have also included a new citation of a timely AmNat paper that explores some of these issues explicitly.}
+	{\itshape These are good points, that was awkward wording on our part.
+While revising the introduction, we felt that much of this first paragraph was tangential to the key concepts we wanted to introduce.
+We have now re-written and shortened it to more quickly and concisely introduce several concepts which we feel are more relevant to the rest of the introduction ({\bf L.20--24}).
+The offending sentence has been removed.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape Agreed. We have rephrased to avoid the term "trade-off" which tends to over-simplify the concept in our opinion ({\bf L.28}).}
+	{\itshape Agreed. We have rephrased to avoid the term "trade-off" which tends to over-simplify the concept in our opinion ({\bf L.23--24}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape We have rephrased the first sentence of this paragraph to be more specific ({\bf L.30}), which we hope also addresses this comment.}
+	{\itshape We have rephrased the first sentence of this paragraph to be more specific ({\bf L.25--26}), which we hope also addresses this comment.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape \textcolor{red}{Lotte, double check here?} We have added a few lines to discuss frequency vs density dependence when we first introduce the population projection matrix in equation 3 ({\bf L.151--157}).
-In addition, we have tried to clarify to improve this section by discussing that the survival and fertility rates ("demographic rates") are density-indepedent, and that the allele frequency dynamics make the model frequency dependent.}
+	{\itshape We have added a few lines to discuss frequency vs density dependence when we first introduce the population projection matrix in equation 3 ({\bf L.131--132}).
+In addition, we have tried to clarify to improve this section by discussing that the survival and fertility rates ("demographic rates") are density-indepedent, and that the allele frequency dynamics make the model frequency dependent ({\bf L.128--134}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape We have rephrased this sentence to clarify and be more consistent with our language earlier in the preceding paragraphs ({\bf L.40--42}).}
+	{\itshape We have rephrased this sentence to clarify and be more consistent with our language earlier in the preceding paragraphs ({\bf L.39--42}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape We have removed $L$ and written the full expression involving $C$.}
+	{\itshape We now present the full expression involving $C$ as Eq(16).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape We have added a sentence earlier in this paragraph where we remind readers what dioecy is ({\bf L.57}).
-Hermaphroditism is defined earlier in the paragraph, so an additional reference seemed redundant.}
+	{\itshape We have added a brief parenthesis to remind readers what dioecy is ({\bf L.50}). Hermaphroditism is defined earlier in the paragraph, so an additional reference seemed redundant.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape Fixed ({\bf L.105-106}).}
+	{\itshape Fixed ({\bf L.85-86}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape Good point, fixed ({\bf L.193}).}
+	{\itshape Good point, fixed ({\bf L.169, 175, Table 2}).}
 \end{quote}
@@
 	{\itshape Thank you for this suggestion.
-To address this issue, as well as the Editor's request that we present population growth rates, we now start with an illustrative figure that only has one $f$, and that shows separately what happens to the allele frequencies (Population Genetic Outcome) and to the population growth rate (Demographic Outcome), see new {\bf fig.
+To address this issue, as well as the Editor's request that we present population growth rates, we now start with an illustrative figure that has only one $f$, and that shows separately what happens to the allele frequencies (Population Genetic Outcome) and to the population growth rate (Demographic Outcome), see new {\bf fig.
 1}.
@@
 \begin{quote}
-	{\itshape We have added a brief parenthesis to clarify ({\bf L.447--448}).}
+	{\itshape We have added a brief parenthesis to clarify ({\bf L.424--425}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape This figure now presents four separate panels with heatmaps of $\lambda$, each with one set of invasion conditions ({\bf fig.~5}). We hope this eases interpretation.}
+	{\itshape This figure now has four separate panels with heatmaps of $\lambda$, each with one set of invasion conditions ({\bf fig.~5}). We hope this eases interpretation.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape We have re-written this sentence to avoid this awkward phrasing ({\bf L.600--601}).}
+	{\itshape We have re-written this sentence to avoid this awkward phrasing ({\bf L.574--575}).}
 \end{quote}

### 2022-01-05 doc/SA-Hermaphrodites-wDemography-R1.tex
Finalized revisions, response letter, etc.

@@
 
-\textcolor{blue}{When selection differs between the sexes, genes expressed by both males and females can experience sexually antagonistic (SA) selection, where beneficial alleles for one sex are deleterious for the other.} Classic population genetics theory has been fundamental to understanding the evolution of sex-differences and the maintenance of such SA genetic polymorphisms, but these models have rarely considered the demographic consequences \textcolor{blue}{ of coexisting alleles with deleterious fitness effects in each sex.} In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (i.e., dioecious/gonochorous) and partially selfing hermaphrodite populations.
+\textcolor{blue}{When selection differs between the sexes, genes expressed by both males and females can experience sexually antagonistic (SA) selection, where beneficial alleles for one sex are deleterious for the other.} Classic population genetics theory has been fundamental to understanding \textcolor{blue}{how and when SA genetic variation can be maintained by balancing selection}, but these models have rarely considered the demographic consequences \textcolor{blue}{of coexisting alleles with deleterious fitness effects in each sex.} In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (i.e., dioecious/gonochorous) and partially selfing hermaphrodite populations.
 We focus on identifying when SA polymorphisms are maintained by balancing selection {\itshape and} the population growth rate remains positive.
@@
 Genetic constraints on adaptation can arise as a consequence of conflicting selection and gene-flow between different classes of individuals, or more generally, when \textcolor{blue}{different fitness components are negatively correlated} \citep{CharlesworthHughes2000, ConnallonHall2018}.
-\textcolor{blue}{A population's ability to adapt to its environment and persist in the long term depends on the genetic variation that it carries \citep{Fisher1930, Lewontin1975}.
+\textcolor{blue}{A population's ability to adapt to its environment and persist in the long term depends upon the genetic variation that it carries \citep{Fisher1930, Lewontin1975}.
 Yet, populations are heterogenous collections of individuals existing in a complex world where adaptation can be impeded by conflicting selection and gene-flow between different classes of individuals, or more generally, when different fitness components are negatively correlated \citep{Prout2000, CharlesworthHughes2000, ConnallonHall2018}.}
 
-Genetic constraints on adaptation due to conflicting selection have particularly interesting evolutionary consequences \citep{ConnallonDebarreLi2018}.
-On one hand, they provide an effective mechanism for the maintenance of genetic variation, which shapes a population's capacity for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
+Genetic constraints on adaptation \textcolor{blue}{due to conflicting selection }have particularly interesting evolutionary consequences.
+On one hand, they provide an effective mechanism for the maintenance of genetic variation, which shapes a population's capacity for future adaptation \citep{Fisher1930, Prout2000, CharlesworthHughes2000, ConnallonHall2018}.
 On the other hand, they prevent individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase a population's overall extinction risk \citep{kokko2003sexy, harts2014demography}.
@@
 Sexually antagonistic selection (abbreviated SA hereafter) \textcolor{blue}{is an important form of conflicting selection that occurs when beneficial alleles for one sex are deleterious when expressed in the other.
-Likewise, alleles with opposing fitness effects through male and female sex functions can cause SA selection in hermaphrodite populations, where individuals' fitness is determined jointly by their maternal and paternal reproductive success \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
+Likewise, alleles with opposing fitness effects on male and female sex functions can experience SA selection in hermaphrodite populations, where individual fitness is determined jointly by maternal and paternal reproductive success \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
 SA selection appears to be} a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and the evolution of sexual dimorphism and sex-related traits \citep{Kidwell1977, Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, ConnallonClark2012, BondurianskyChenoweth2009, Olito2019}.
-\textcolor{blue}{For instance, phenotypic selection studies commonly find sex-differences in directional selection \citep[e.g.,][]{DeLisle-etal-2018,Lewis-etal-2011,CoxCalsbeek2009,SinghPunzalan2018}, implying that many genetic variants underlying quantitative traits have SA fitness effects.
-Moreover, quantitative genetic studies frequently find negative cross-sex genetic correlations for fitness (\citealt{Chippendale-etal-2001,Delph-etal-2011}; reviewed in \citealt{ConnallonMatthews2019}), and polygenic signals of SA selection are evident in recent population genetic analyses of large genomic datasets \citep{RuzickaESEB2020, RuzickaConnallon2022}.
+\textcolor{blue}{For instance, sex-differences in directional selection are commonly detected in phenotypic selection studies \citep[e.g.,][]{DeLisle-etal-2018,Lewis-etal-2011,CoxCalsbeek2009,SinghPunzalan2018}, suggesting that many genetic variants contributing to quantitative traits have SA effects.
+Moreover, quantitative genetic studies frequently find negative cross-sex genetic correlations for fitness (\citealt{Chippendale-etal-2001,Delph-etal-2011}; reviewed in \citealt{ConnallonMatthews2019}), and polygenic signals of SA selection are evident in several recent population genetic analyses of large genomic datasets \citep{RuzickaESEB2020, RuzickaConnallon2022}.
 Overall, both theory and current empirical data suggest that there is ample scope for the maintenance of SA \textcolor{blue}{genetic variation} in both dioecious \textcolor{blue}{(separate-sexed)} and hermaphrodite populations \citep{BondurianskyChenoweth2009, Abbott2011, RuzickaESEB2020, WangBarrett2020}.}
@@
 
-\textcolor{blue}{While significant efforts have been made to predict when SA polymorphisms should be maintained and to detect SA genetic variation from empirical data, less attention has been paid to the demographic consequences of SA selection \citep{MatthewsConnallon2019}}.
+\textcolor{blue}{While significant efforts have been made to predict when SA polymorphisms can be maintained and to detect SA genetic variation from empirical data, the demographic consequences of SA selection have often been overlooked \citep{MatthewsConnallon2019}}.
 Most population genetic theory, including SA selection theory, models the relative fitness of genotypes in populations of constant (and often infinite) size.
-In reality, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals \textcolor{blue}{which link genotype distributions to population dynamics (e.g., \citealt {johnston2013life, merot2020balancing}).
+In reality, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals \textcolor{blue}{that link genotype distributions to population dynamics (e.g., \citealt {johnston2013life, merot2020balancing}).
 For over 50 years, the field of eco-evolutionary dynamics has sought to jointly model the genetic and demographic processes involved in evolutionary change.
 Theoretical models combining population genetics with density-dependent population growth \citep{roughgarden1971density} and stage-structured populations (consolidated in \citealt{Charlesworth_1994}) paved the way toward understanding the demographic consequences of allele frequency changes due to selection.
-Various extensions of these theories have been made to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and the evolution of quantitative traits \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects}.
-This growing body of eco-evolutionary theory makes plain that demographic processes must be taken into account explicitly to fully understand how genetic variation for fitness is shaped, and maintained in natural populations \citep{metcalf2007evolutionary,coulson2006putting,rueffler2006evolutionary}.}
+Various extensions of these theories have been made to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and quantitative trait evolution \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects}.
+This growing body of eco-evolutionary theory makes plain that demographic processes must be taken into account to fully understand how genetic variation for fitness is shaped in natural populations \citep{metcalf2007evolutionary,coulson2006putting,rueffler2006evolutionary}.}
 
@@
 The potential demographic costs of sexual antagonism were pointed out by \cite{kokko2003sexy}, but few studies since have explicitly incorporated demography into models of sexual antagonism (with the notable exception of \citealt{harts2014demography}).
-\citet{deVriesCaswell2019a} introduced a mendelian matrix model with SA selection and population dynamics, but did not analyze how demography influenced the maintenance of SA polymorphism.
-In this paper, we extend the model of \citet{deVriesCaswell2019a,deVriesCaswell2019b} to \textcolor{blue}{include both dioecious and hermaphroditic populations and jointly analyze the genetic (fixation of male- or female-beneficial alleles, or balanced polymorphism) and demographic (population persistence or extinction) outcomes of SA selection}.
+\citet{deVriesCaswell2019a} introduced a mendelian matrix model with SA selection and population dynamics, but did not analyze how demography influenced SA polymorphism.
+In this paper, we extend the model of \citet{deVriesCaswell2019a,deVriesCaswell2019b} to \textcolor{blue}{include both dioecious and hermaphroditic populations and jointly analyze the genetic (fixation/polymorphism) and demographic (persistence/extinction) outcomes of SA selection}.
 A major strength of our approach is that the model can be parameterized using empirically estimated demographic rates, enabling us to make predictions about the scope for SA polymorphisms that are grounded in the biology of real populations.
@@
 
-Simultaneous hermaphrodites can transmit genes to the next generation via both sperm/pollen and eggs/ovules, and have the potential to reproduce by a combination of self- and outcross-fertilizations.
-Maternal outcrossing involves receiving male gametes from other individuals in the population, while paternal outcrossing involves exporting male gametes to other individuals and fertilizing their ovules.
+Simultaneous hermaphrodites can transmit genes to the next generation via both sperm/pollen and eggs/ovules, and can potentially reproduce by a combination of self- and outcross-fertilization.
+Maternal outcrossing involves receiving male gametes from other individuals in the population, while paternal outcrossing involves exporting male gametes to fertilize other individuals' ovules.
 Self-fertilization occurs when an individual's male gametes fertilize their own ovules.
 To distinguish between parameters relating to male and female reproductive functions, we denote matrices or vectors relating to the male sex function with a prime.
-\textcolor{blue}{Following previous theory for partially selfing populations \cite[e.g.,][]{Charlesworth2010,JordanConnallon2014,Glemin2021}, we use a 'fixed prior selfing' model where all individuals are assumed to self-fertilize at a constant rate, $C$, prior to receiving outcross pollen, for reasons of analytic tractability.
+\textcolor{blue}{Following previous theory for partially selfing populations \cite[e.g.,][]{Charlesworth2010,JordanConnallon2014,Glemin2021}, we use a 'fixed prior selfing' model where all individuals are assumed to self-fertilize at a constant rate, $C$, prior to receiving outcross pollen.
 We present a more general model of genotype-specific self-fertilization rates in the Online Supplement, and briefly discuss implications of this approach in the Discussion.}
 
-Individuals in the model are jointly classified by life-cycle stage ($1, \ldots, \omega$), genotype ($1, \ldots, g$), and whether they were produced by self- or outcross fertilization (denoted by $S$ and $X$ superscripts, respectively; see Table \ref{tab:Terms} for a full description of terms included in the model).
-Whether individuals were produced through selfing or outcrossing is included in the individual state description because individuals produced through selfing might experience reduced survival, growth, or maturation rates as a consequence of inbreeding depression.
-In reality, the severity of inbreeding depression will not only be a function of whether an individual is produced by selfing but also of how many consecutive generations of inbreeding have occurred in their lineage (e.g., \citealt{kelly1999response, kelly2007mutation}).
-However, tracking lineage-specific inbreeding histories is beyond the scope of this paper, and we therefore make the common assumption that the severity of inbreeding depression is the same for all individuals (e.g., \citealt{Charlesworth1987, Charlesworth2009, Charlesworth2010,JordanConnallon2014}).
+Individuals are jointly classified by life-cycle stage ($1, \ldots, \omega$), genotype ($1, \ldots, g$), and whether they were produced by self- or outcross fertilization (denoted by $S$ and $X$ superscripts, respectively; see Table \ref{tab:Terms} for a full description of terms included in the model).
+We include how individuals were produced in the state description because individuals produced by selfing might experience reduced survival, growth, or maturation rates due to inbreeding depression.
+\textcolor{blue}{This approach makes the common simplifying assumption that all individuals in the population experience the same severity of inbreeding depression (e.g., \citealt{Charlesworth1987, Charlesworth2009, Charlesworth2010, JordanConnallon2014}).
+We address potential consequences of relaxing this assumption, and how our model can be extended to do so, in the discussion.
+}
+
+%In reality, the severity of inbreeding depression will not only be a function of whether an individual is produced by selfing but also of how many consecutive generations of inbreeding have occurred in their lineage (e.g., \citealt{kelly1999response, kelly2007mutation}).
+However, tracking lineage-specific inbreeding histories is beyond the scope of this paper, and we therefore make the common assumption
 
@@
 Deviations from this assumption are straightforward to incorporate but beyond the scope of this paper.
-
-\textcolor{blue}{If the survival or fertility rates are a function of the population density, $\tilde{\mbf{n}}(t)$, then the model will be a density dependent model and the population projection matrix becomes a function of the full population vector, $\tilde{\mbf{A}}[\tilde{\mbf{n}}(t)]$.
-Below, we limit our analyses to consider density-independent survival and fertility rates.
-The population projection matrix is then only a function of the proportional population vector, $\tilde{\mbf{A}}[\tilde{\mbf{p}}(t)]$, and the model is referred to as a frequency-dependent model.
-In \ref{App:DensDep} we construct and analyze a density-dependent version of the model and show that with density-dependent survival affecting all stages equally, our main conclusions remain intact.}
 
@@
 $C$ & The population rate of self-fertilization &0-1 \\
-$\delta$ &\textcolor{blue}{ Inbreeding depression effect on ovule viability}&0-0.8\\
+$\delta$ &\textcolor{blue}{ Inbreeding depression effect on zygote viability}&0-0.8\\
 $ \delta_j$ & \textcolor{blue}{ Inbreeding depression effect on juvenile survival}& 0-0.8\\
@@
 $\delta_{\gamma}$ &\textcolor{blue}{Inbreeding depression effect on maturation rate} & 0-0.8\\
-$f_i,\,f^{\prime}_i$ & Adult fertilities through female and male sex functions respectively & 5.8-8.5 \\
+$f_i,\,f^{\prime}_i$ & Adult \textcolor{blue}{maximum} fertilities through female and male sex functionsxi & 5.8-8.5 \\
 
@@
 
+\textcolor{blue}{If the survival or fertility rates are a function of the population density, $\tilde{\mbf{n}}(t)$, then the model will be density-dependent and the population projection matrix becomes a function of the full population vector, $\tilde{\mbf{A}}[\tilde{\mbf{n}}(t)]$.
+Below, we limit our analyses to consider density-independent survival and fertility rates.
+The population projection matrix is then only a function of the proportional population vector, $\tilde{\mbf{A}}[\tilde{\mbf{p}}(t)]$, and the model is referred to as frequency-dependent.
+In \ref{App:DensDep} we construct and analyze a simple density-dependent version of the model and show that with density-dependent survival affecting all stages equally, our main conclusions remain intact.}
+
 \subsubsection*{Sexually antagonistic selection \& inbreeding depression} \label{sec:SAsel}
 
-\textcolor{blue}{To demonstrate how to construct and analyze a genotype $\times$ stage-classified model, we will now construct an example model for a hypothetical hermaphrodite species with intralocus sexual conflict via the two sex functions.
-}For the sake of simplicity, we assume our hypothetical species has a life cycle with only two stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active.
+\textcolor{blue}{We now construct and analyze an example genotype $\times$ stage-classified model for a hypothetical hermaphrodite species with intralocus sexual conflict via the two sex functions.} For the sake of simplicity, we assume our hypothetical species has a life cycle with only two stages: juveniles and adults (i.e., $\omega = 2$), and that only adults can reproduce.
 Suppose that there is a genetic trade-off between the sex-functions at a single diallelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect.
@@
 \subsubsection*{Mating and offspring production under partial selfing}
+
  Outcrossing in our model proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}.
@@
 
- \textcolor{blue}{ For the purpose of this article, we assume that all outcross matings are random.
-In addition, we assume that individuals produced by selfing or by outcrossing do not differ in their mating success, though they can differ in their survival (see above).
+ \textcolor{blue}{ For the purposes of this article, we assume that all outcross matings are random, and that individuals produced by selfing or outcrossing have the same mating success, though they can differ in their survival (see above).
 Given these assumptions, the allele frequencies in the mating population can obtained by summing the vectors of individuals produced through selfing and outcrossing:
@@
 \end{linenomath*}}
-
 \textcolor{blue}{Individuals contribute gametes to the population outcrossing gamete pool according to their fertility, which depends on their stage and genotype at the SA locus.
@@
 where equality is obtained by normalizing the frequencies such that they add to one (see Online Supplementary Materials for a derivation of Eq(\ref{eq:maleGametePool})).
-The first entry sums the  stage-specific contributions of allele A to the gamete pool by heterozygotes and $AA$ homozygotes; the second entry sums the stage-specific contributions of allele a to the gamete pool by heterozygotes and $aa$ homozygotes.
+The first entry sums the  stage-specific contributions of allele $A$ to the gamete pool by $AA$ homozygotes and heterozygotes; the second entry does the same for the stage-specific contributions of allele $a$ to the male gamete pool.
 }
@@
 Note that this equation is identical to the classical population genetic result for allele frequencies after selection among males in a two-sex SA model, with the caveat that only reproductively active adults contribute to the gamete pool \citep[e.g., see Eq(1a) of][]{Kidwell1977}.
-Eq(\ref{eq:maleGametePool_example}) increasingly deviates from the classical population genetics result if additional reproductive adult age or stage classes are included (see Eq(\ref{eq:maleGametePool})).
-}
+However, Eq(\ref{eq:maleGametePool}) increasingly deviates from the classical population genetics result if additional reproductive stages are included.
+Note also that by modeling allele frequencies in the male gamete pool rather than sperm/pollen production, we are making the common assumption of female demographic dominance (i.e., population growth rates are determined by female fecundity, not production of male gametes; \citealt{Caswell2001,iannelli2005gender,pollard1975mathematical}).
+We elaborate on several implications of this assumption for hermaphroditic populations in the discussion.}
 
 \subsubsection*{Population projection}
+
 Using the component matrices described above (the survival matrices, $\mathbf{U}^X_i$, $\mathbf{U}^S_i$, the fertility and mating success matrices, $\mathbf{F}_i$, $\mathbf{F}^\prime_i$) and the allele frequencies in the male gamete pool, we can construct the population projection matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$.
@@
 
-Diverse eco-evolutionary outcomes are possible in the model, including fixation of either allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and suicide (e.g., see Figure 1 of \citealt{deVriesCaswell2019a, deVriesCaswell2019b} for examples of the possible outcomes).
-We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection \textcolor{blue}{(Figure \ref{fig:didactic}, left-hand panel)} and ({\itshape ii}) the intrinsic population growth rate at equilibrium is greater than $1$ \textcolor{blue}{(Figure \ref{fig:didactic}, right-hand panel)}; a situation that we refer to as a 'demographically viable SA polymorphism'.
-
-We identify conditions where SA polymorphism is 'protected' by evaluating the stability of populations initially fixed for either SA allele to invasion by the other allele (i.e., we assessed stability at the boundary equilibrium genotype frequencies of $\hat{\mbf{p}}_{AA} = 1$ and $\hat{\mbf{p}}_{aa} = 1$; \citealt{Levene1953, Prout1968, deVriesCaswell2019b}).
+Diverse eco-evolutionary outcomes are possible in the model, including fixation of either allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and suicide \citep[see][for examples of the possible outcomes]{deVriesCaswell2019a, deVriesCaswell2019b}.
+We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection \textcolor{blue}{(Figure \ref{fig:didactic}, left panel)} and ({\itshape ii}) the intrinsic population growth rate at equilibrium, $\lambda$, is greater than $1$ \textcolor{blue}{(Figure \ref{fig:didactic}, right panel)}; a situation that we refer to as a 'demographically viable SA polymorphism'.
+
+We identify conditions where SA polymorphism is 'protected' by evaluating the stability of populations initially fixed for either allele to invasion by the other (i.e., we assessed stability at the boundary equilibrium genotype frequencies of $\hat{\mbf{p}}_{AA} = 1$ and $\hat{\mbf{p}}_{aa} = 1$; \citealt{Levene1953, Prout1968, deVriesCaswell2019b}).
 The formal conditions for a protected polymorphism are determined by linearizing the model in the vicinity of the boundary equilibria ($\hat{\mbf{p}}$), and evaluating the magnitude of the largest eigenvalue of the Jacobian matrix of the linearization.
@@
 Second, we assume that if inbreeding depression is caused primarily by recessive deleterious mutations (as suggested by empirical data), it should covary negatively with the population selfing rate due to purging, provided the population selfing rate has been relatively constant in recent evolutionary time (\citealt{Charlesworth2009}; though we note that other processes could give rise to this pattern, e.g., \citealt{CrnokrakBarrett2002, Charlesworth2009,HedrickGarcia-Dorado2016}).
-Following \citet{Olito2019}, we incorporate such negative covariance by constraining the inbreeding depression terms in the model ($\delta$ and $\delta_i$, where $i \in \{j,a,\gamma\}$) to follow a simple declining function of the selfing rate: \textcolor{blue}{
+Following \citet{Olito2019}, we incorporate such negative covariance by constraining the inbreeding depression terms in the model ($\delta$ and $\delta_i$, where $i \in \{j,a,\gamma\}$) to follow a simple declining function of the selfing rate:
+\begin{linenomath*}
+\textcolor{blue}{
 \begin{equation}
@@
 \end{equation}}
+\end{linenomath*}
 where $\delta^{\ast}$ is the hypothetical severity of inbreeding depression for a completely outcrossing population, $b$ is a shape parameter determining how far $\delta$ will decline under complete selfing (when $C = 1$), \textcolor{blue}{and $a$ is a shape parameter} controlling the curvature of the overall function for $\delta$ (see Appendix E in \citealt{Olito2019} for additional details).
@@
 We begin with an illustration of demographically viable polymorphic parameter space in the absence of inbreeding depression in Figure \ref{fig:extThresholds} (i.e., $\delta = \delta_i = 0$).
-The red regions in Figure \ref{fig:extThresholds} indicate areas where the population goes extinct (\textcolor{blue}{i.e., where $\lambda < 1$}) for different values of the fertility parameter.
-Lower female fertility values correspond to a larger demographically inviable area.
+The red regions in Figure \ref{fig:extThresholds} indicate areas where the population goes extinct (\textcolor{blue}{i.e., where $\lambda < 1$}) for different \textcolor{blue}{maximum female fertility values}.
+Lower female fertility values correspond to larger areas where extinction occurs.
 The funnel between the thick black lines indicates the area where the sexually antagonistic alleles coexist.
@@
 The combination of protection from reduced female fertility and reduced total polymorphic parameter space caused by selfing results in populations with intermediate selfing rates having the greatest proportion of demographically viable parameter space at medium and low fertilities (fig.~\ref{fig:polySpace}, med.~and low.~fertility values).
-
-%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Demographic effects of inbreeding depression} \label{subsec:InbreedEffects}
-
-Unlike previous population genetic models, which assume constant population sizes \citep{JordanConnallon2014,Olito2017}, mortality caused by inbreeding depression can strongly influence population persistence in our evolutionary demographic model.
-Populations with high fertility rates can maintain positive population growth rates despite this additional mortality.
-This causes a greater proportion of SA polymorphic parameter space to be demographically viable, with the demographic model predictions converging on those from population genetic models in high-fertility populations (fig.~\ref{fig:deltaPolySpace}).
-\textcolor{blue}{Note, however, that in our model, mortality due to late-acting inbreeding depression also reduces the population effective selfing rate, and therefore shifts the invasion boundaries for the SA alleles to more closely resemble an outcrossing population.
-In some cases, this can lead to larger regions of viable polymorphic parameter space than predicted by the population genetic models, which can only include early-acting inbreeding depression (e.g., see cross symbols in fig.~\ref{fig:deltaPolySpace})}.
-However, as the selfing rate increases, demographic viability eventually crashes when the population can no longer sustain the concomitant increase in mortality due to inbreeding depression (fig.~\ref{fig:deltaPolySpace}).
-For populations with lower fertility rates, the proportion of demographically viable polymorphic parameter space \textcolor{blue}{is very similar to} the population genetic model predictions \textcolor{blue}{for low selfing rates}.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Demographic effects of inbreeding depression} \label{subsec:InbreedEffects}
+
+Unlike previous population genetic models \textcolor{blue}{of SA selection in hermaphrodites}, which assume constant population sizes \citep{JordanConnallon2014,Olito2017}, mortality caused by inbreeding depression can strongly influence population persistence in our evolutionary demographic model.
+Populations with high fertility rates can maintain positive population growth rates despite this additional mortality.
+This causes a greater proportion of SA polymorphic parameter space to be demographically viable, with the demographic model predictions converging on those from population genetic models in high-fertility populations (fig.~\ref{fig:deltaPolySpace}).
+\textcolor{blue}{Note, however, that mortality due to late-acting inbreeding depression also reduces the population effective selfing rate in our model, and therefore shifts the invasion boundaries for the SA alleles to more closely resemble an outcrossing population.
+In some cases, this leads to larger regions of viable polymorphic space than predicted by the population genetic models, which only include early-acting inbreeding depression (e.g., see cross symbols in fig.~\ref{fig:deltaPolySpace})}.
+However, as the selfing rate increases, demographic viability eventually crashes when the population can no longer sustain the concomitant increase in mortality due to inbreeding depression (fig.~\ref{fig:deltaPolySpace}).
+
 Mortality from inbreeding depression quickly outweighs the beneficial effect of selfing on the equilibrium female load that was apparent in the previous section.
@@
 Our results indicate that population viability was most sensitive to inbreeding depression affecting juvenile survival rates ($\delta_j$; fig.~\ref{fig:deltaPolySpace}, dark blue points), while early-acting inbreeding depression ($\delta$, ovule abortion shortly after fertilization) had a similar effect on population viability as late-acting inbreeding depression affecting adult survial ($\delta_a$) and juvenile-to-adult transition rates ($\delta_{\gamma}$).
-Inbreeding effects on juvenile survival had the strongest effect on population viability because on average individuals will spend multiple time steps in the juvenile stage before they mature.
-At each time step, juveniles have a probability $\gamma$ to mature and a probability $\sigma_j(1-\delta_j)$ to survive.
+Inbreeding effects on juvenile survival had the strongest effect on population viability because, on average, individuals spend multiple time steps as juveniles before they mature.
+At each time step, juveniles mature with probability $\gamma$ and survive with probability $\sigma_j(1-\delta_j)$.
 Inbreeding depression at the juvenile stage therefore makes it harder to survive long enough to mature.
-\textcolor{blue}{Of course the strength of this effect is influenced by the maturation rate, $\gamma$, which therefore also affects the shape of the curves in Figure \ref{fig:deltaPolySpace}.} Although early-acting inbreeding depression ($\delta$) actually manifests earlier in the life-cycle than juvenile survival, it acts only once by influencing the total number of self-fertilized zygotes that become juveniles.
+\textcolor{blue}{Of course the strength of this effect is influenced by the maturation rate, $\gamma$, which therefore also affects the shape of the curves in Figure \ref{fig:deltaPolySpace}.} Although early-acting inbreeding depression ($\delta$) manifests earlier in the life-cycle than juvenile survival, it acts only once by influencing the total number of self-fertilized zygotes that become juveniles.
 Note, however, that the relative strength of inbreeding depression at different stages of the life history will depend on the interaction between the various inbreeding depression parameters, which we have precluded from our analyses.
@@
 
-To illustrate how our model can be used to explore whether demographic rates observed in natural populations appear likely to support balanced SA polymorphisms, we parameterized the model using empirically estimated demographic rates and fitness data for natural populations of the hermaphroditic flowering plant {\itshape Mimulus guttatus} (Scrophulariaceae; now known as {\itshape Erythranthe guttata}).
+To illustrate how our model can be used to explore whether natural populations appear likely to harbor balanced SA polymorphisms, we parameterized our model using empirically estimated demographic rates and fitness data for natural populations of the hermaphroditic flowering plant {\itshape Mimulus guttatus} (Scrophulariaceae; now known as {\itshape Erythranthe guttata}).
 {\itshape M.~guttatus} is an herbaceous, self-compatible wildflower native to western North America that exhibits remarkable among-population variation in numerous life-history and reproductive traits including selfing rates, inbreeding depression, floral morphology, and annual-to-perennial life-history \citep[e.g.,][]{RitlandGanders1987, Ritland1990, Willis1993, Willis1999a, Willis1999b, WuWillis2008}.
@@
 inv6 segregates at moderate frequency (about $8\%$), and carriers suffer an approximately $30 \%$ loss in pollen viability, but also increased flower (and therefore ovule and pollen) production.
-The genetic basis of these effects remain unclear -- though they are likely polygenic -- but the net result is a "supergene" with remarkably strong effects on both female and male fertility that segregates as a single diallelic locus \textcolor{blue}{(i.e., with two alleles, wild-type and inverted)}.
-We estimated selection coefficients for the effect of inv6 on pollen production (i.e., taking into account the simultaneous effect on flower number) and ovule production from the data reported in \citet{LeeKelly2015} under the relatively conservative assumptions that pollen/ovule production is proportional to flower number, and  additive SA fitness effects (dominance coefficients could not be estimated from field data; see \ref{App:Mimulus} for details of estimating selection coefficients).
+The genetic basis of these effects remain unclear -- though they are likely polygenic -- but the net result is a "supergene" with remarkably strong effects on both female and male fertility that segregates as a single diallelic locus \textcolor{blue}{(i.e., with two alleles: wild-type and inverted)}.
+We estimated selection coefficients for the effect of inv6 on pollen production (i.e., taking into account the simultaneous effect on flower number) and ovule production from the data reported in \citet{LeeKelly2015} under the relatively conservative assumptions that pollen/ovule production is proportional to flower number, and  additive SA fitness effects (dominance coefficients could not be estimated from field data; see \ref{App:Mimulus} for details).
 Under these assumptions, the average selection coefficient for inv6 on flower production across two years in the field was $(s_f,s_m) = (0.3,0.31)$.
@@
 Invasion conditions and population intrinsic growth rates ($\lambda$) were calculated using the conservative assumption of additive fitness effects in both sexes ($h_f = h_m = 0.5$) for two parameter conditions: obligate outcrossing (panels A,B) and partial selfing with inbreeding depression using the highest field-estimate of selfing and inbreeding depression parameters ($\delta_i$) calculated for the Iron Mountain population of \citet{Willis1993} (panels C,D).
-The location of inv6 is also shown on both plots, using selection coefficients calculated from field estimates of male and female fitness components from \citet{LeeKelly2015} under the relatively conservative assumption of additive fitness effects in both sexes, where $(s_f,\,s_m) = (0.30,\, 0.31)$.}}
+The location of inv6 is also shown on both plots, using selection coefficients calculated from field estimates of male and female fitness components from \citet{LeeKelly2015}: $(s_f,\,s_m) = (0.30,\, 0.31)$.}}
  \label{fig:mimulusFig}
@@
 The Eagle Meadows and Low-Elevation Perennial populations of \citet{PetersonEtAl2016} had contrasting demographic rates that strongly affected the scope for demographically viable SA polymorphism.
-The Eagle Meadows population was locally adapted with a very high intrinsic growth rate ($\lambda \approx 1.7$).
+The locally adapted Eagle Meadows population had a very high intrinsic growth rate ($\lambda \approx 1.7$).
 This growth rate was sufficiently high that all $s_f \times s_m$ selection parameter space (where $s_f, s_m \in (0,1]$) remained demographically viable, regardless of the selfing rate and effects of inbreeding depression ($C = 0$ or $0.29$; fig.~\ref{fig:mimulusFig}A,C).
@@
 
-\textcolor{blue}{In Appendix B we reanalyze the {\itshape M.~guttatus} example using a simple density-dependent version of our model.
+\textcolor{blue}{In Appendix B we reanalyze this {\itshape M.~guttatus} example using a simple density-dependent version of our model.
 We find that for (st)age-independent density-dependent growth, the coexistence boundaries are unchanged, and the equilibrium population densities closely resemble the shape of the density-independent population growth rates presented above (see fig.~\ref{fig:mimulusFig_dd}).
@@
 
-Classic population genetics theory predicts that sexually antagonistic selection is unlikely to maintain genetic variation except under narrow conditions, with polymorphism requiring either finely balanced or unusually strong selection, or partially recessive fitness effects through each sex \citep{Kidwell1977,Pamilo1979,Prout2000, ConnallonClark2014}.
+Classic population genetics theory predicts that sexually antagonistic selection can only maintain genetic variation under narrow conditions, with polymorphism requiring either finely balanced or unusually strong selection, or partially recessive fitness effects through each sex \citep{Kidwell1977,Pamilo1979,Prout2000, ConnallonClark2014}.
 Extensions of the theory have identified numerous ways in which the conditions for polymorphism become more permissive in both dioecious and hermaphroditic organisms, including genetic linkage of SA loci, the evolution of sex-specific dominance, population subdivision, and life-cycle complexity \citep[e.g.,][]{Patten2010,JordanConnallon2014,JordanCharlesworth2012,SpencerPriest2016, ConnallonSharmaOlito2019, Olito-etal-2018}.
@@
 The first and central finding of our study is that when population intrinsic growth rates approach one, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space where SA polymorphism is maintained.
-We find that much of the parameter space where a male-beneficial allele \textcolor{blue}{fixes} is demographically inviable for populations whose intrinsic growth rates are close to one prior to invasion.
+Additionally, large regions where male-beneficial alleles are predicted to \textcolor{blue}{fix} become demographically inviable for populations whose intrinsic growth rates are close to one prior to invasion.
+This demographic consequence of masculinization has rarely been considered \citep{hitchcock2020gene}.
 Since intrinsic growth rates far exceeding one suggest rapid exponential growth, they are generally rare (with the notable exceptions of recently introduced invasive populations), suggesting that our model predictions may be highly relevant for SA polymorphism in many real-world populations.
-This demographic consequence of masculinization has rarely been considered \citep{hitchcock2020gene}.
-Moreover, these findings complement recent theoretical and empirical studies indicating that SA selection is likely to be both condition dependent, and stronger in locally-adapted populations near the center of a species' range, where population growth rates are expected to be higher and populations might be more likely to be able to suffer the demographic cost of a SA polymorphism while remaining viable \citep{BergerEtAl2014, Connallon2015}.
+Moreover, these findings complement recent theoretical and empirical studies indicating that SA selection is likely to be both condition dependent, and stronger in locally-adapted populations near the center of a species' range, where population fertility rates are expected to be higher and populations might be more likely to be able to suffer the demographic cost of a SA polymorphism while remaining viable \citep{BergerEtAl2014, Connallon2015}.
 
@@
 This key prediction is supported by a series of experimental results in seed beetles ({\itshape Callosobruchus maculatus}) which show that male-beneficial SA genotypes are less likely to contribute positively to population growth rates, and are more susceptible to extinction under environmental stress or inbreeding \citep{BergerEtAl2014, berger2016intralocus, grieshop2017male}.
-An interesting corrollary of our findings is that, since strong SA fitness effects can often lead to extinction, the observable SA genetic variation in natural populations may often be under weak selection, and therefore strongly susceptible to genetic drift whether or not selection favours the fixation of one allele or balancing selection \citep{ConnallonClark2012}.
+An interesting corrollary of our findings is that, since strong SA fitness effects can often lead to extinction, the observable SA genetic variation in natural populations may often be under weak selection, and therefore strongly susceptible to genetic drift whether selection favours the fixation of one allele or polymorphism \citep{ConnallonClark2012}.
 
@@
 
-In hermaphroditic populations, self-fertilization can alleviate the demographic costs of balanced SA polymorphisms under some conditions, however, the concommitant effects of inbreeding depression generally exacerbate them in populations with mixed-mating systems.
+In hermaphroditic populations, self-fertilization can alleviate the demographic costs of balanced SA polymorphisms under some conditions; however, our model indicates that the concommitant effects of inbreeding depression will generally exacerbate them in populations with mixed-mating systems.
 This prediction is in stark contrast to previous population genetics models of SA selection in hermaphrodites, where the sole effect of inbreeding depression is to reduce the population effective selfing rate through the loss of selfed-zygotes, thereby expanding polymorphic parameter space \citep{JordanConnallon2014, Olito2017}.
 In our model, this reduction in the effective selfing rate is accompanied by significant mortality due to inbreeding depression (whether early- or late-acting), which can quickly tip partially-selfing populations over the brink to extinction.
-Beyond the maintenance of SA polymorphisms, this finding underscores a simple but important point that is often overlooked in studies of the evolution of self-fertilization and selfing-syndromes, which tend to emphasize the coevolution of the deleterious mutation load and mating system \citep[e.g.,][]{Charlesworth1987, LandeSchemske1985, Goodwillie2005}: highly fertile populations can better afford the severe demographic costs of inbreeding.
-This suggests that traits related to female fecundity, such as ovule and flower production, may strongly influence the distribution of successful transitions to self-fertilization among hermaphroditic taxa, as well as a variety of ecological correlates of selfing and mixed-mating \citep{Goodwillie2005, Igic2006, Grossenbacher2015}.
+Beyond the maintenance of SA polymorphisms, this finding underscores a simple but important point that is often overlooked in studies of the evolution of self-fertilization, which tend to emphasize the coevolution of the deleterious mutation load and mating system \citep[e.g.,][]{Charlesworth1987, LandeSchemske1985, Goodwillie2005}: highly fertile populations can better afford the severe demographic costs of inbreeding.
+Thus, traits related to female fecundity, such as ovule and flower production, may strongly influence the distribution of successful transitions to self-fertilization among hermaphroditic taxa, along with other ecological correlates of selfing \citep{Goodwillie2005, Igic2006, Grossenbacher2015}.
 
@@
 
-\subsubsection*{Extensions and future directions}\label{subsubsec:extensions}
+\subsubsection*{Assumptions, extensions, and future directions}\label{subsubsec:extensions}
 
@@
 For the sake of simplicity, we made a variety of assumptions in our analyses, and \textcolor{blue}{many extensions of the model are} possible.
-For example, we used a simple life cycle with just two stages for most of this paper, adults and juveniles (the \textit{Mimulus} example has $3$ stages).
+For example, we used a simple life cycle with just two stages for most of this paper (the \textit{Mimulus} example has $3$ stages).
 However, it is possible to include additional age classes, allowing us to explore whether the scope for SA selection to maintain polymorphisms is affected by whether a species exhibits positive or negative senescence \citep{jones2014diversity}.
-Put another way, does the shape of the survival curve affect the demographic and population genetic consequences of SA selection, and how does this interact with the age at which a sexually antagonistic allele is expressed?
-What if the allele not only affects male and female fertility but also affects survival of individuals?
+Put another way, our model can be used ask how the shape of the survival curve affects the demographic and population genetic consequences of SA selection, and how this interacts with the age at which SA alleles are expressed?
+or what happens if SA alleles affect both male and female fertility and survival?
 
@@
 Of course, in natural populations density-dependent processes like resource competition or competition for breeding sites will often prevent populations from growing (or shrinking) exponentially.
-In the simple density-dependent version of our model presented in \ref{App:DensDep}, we included density-dependence acting on the survival of all life history stages equally, and reanalysed our case study of \textit{M.
+In \ref{App:DensDep}, we incorporated a simple form of density-dependence acting on the survival of all life history stages equally, and reanalysed our case study of \textit{M.
 guttatus}.
 The density-dependent version of the model predicts equilibrium population densities rather than population intrinsic growth rates (see fig.~\ref{fig:mimulusFig_dd}).
-This brief example shows that the shape of the polymorphic region is unchanged by this form of density-dependence, and that the shape of the population density surface in fig.~\ref{fig:mimulusFig_dd} is very similar to that for population growth rate in fig.~\ref{fig:mimulusFig}.
-The similarity between the density-independent and -dependent model predictions underscores an important point: density-dependence that acts (st)age-independently does not alter selection gradients \citep{mylius1995evolutionarily, caswell2017senescence}.}
-
-\textcolor{blue}{In many natural populations, however, density-dependence will not have the same effect on the survival of individuals of all ages or developmental stages.
+Our reanalysis shows that the shape of the polymorphic region is unchanged by this form of density-dependence, and that the population density surface in fig.~\ref{fig:mimulusFig_dd} is very similar to that for population growth rate in fig.~\ref{fig:mimulusFig}.
+The similarity between the density-independent and -dependent model predictions underscores a key point made by earlier studies: density-dependence acting (st)age-independently does not alter selection gradients \citep{mylius1995evolutionarily, caswell2017senescence}.}
+
+\textcolor{blue}{Yet, in many natural populations density-dependence will not have the same effect on individual survival at all ages or developmental stages.
 For example, competition for breeding sites will cause a reduction in fecundity rather than survival.
-In plants, density-dependent competition for soil nutrients will likely influence seed germination rates differently than seedling growth rates or flower/fruit production in adults \citep{AntonovicsLevin1980}.
-These types of age- or stage-specific density-dependence will change the selection gradients, and therefore the shape of polymorphic parameter space as well as the equilibrium population densities predicted by the model.
+In plants, density-dependent competition for soil nutrients will influence seed germination rates differently than seedling growth rates or adult flower/fruit production \citep{AntonovicsLevin1980}.
+These types of age- or stage-specific density-dependence will change selection gradients, and therefore the shape of polymorphic parameter space as well as the equilibrium population densities predicted by the model.
 \citet{de2020matrix} discuss some of the possible dynamical outcomes of a density-dependent genetic model for \textit{Tribolium} seed beetles but a thorough analysis of more complicated forms of density-dependence for the maintenance of SA polymorphisms remains an important avenue for future work.}
 
-Like most demographic matrix models, ours also assumes female demographic dominance, where population growth rates are determined entirely by female fecundity \citep{pollard1975mathematical,Caswell2001,iannelli2005gender}.
-Yet, many hermaphroditic populations experience limitation of reproductive success due to both quantity and quality of male gametes \citep[e.g.,][]{Yund2000,AizenHarder2007, Harder2016}.
-Explicitly modeling male gamete production is a natural extension to our modeling framework, and would enable us to analyze how tension between the demographic consequences of SA fitness variation through both female and male function alters or abolishes asymmetries in the extinction thresholds and polymorphic parameter space \textcolor{blue}{caused by our assumption of female demographic dominance} \citep[e.g.,][]{Tazzyman2015}.
+Like most demographic matrix models, ours assumed female demographic dominance, where population growth rates are determined entirely by female fecundity \citep{pollard1975mathematical,Caswell2001,iannelli2005gender}.
+Yet, reproductive success in many hermaphroditic populations is limited by both the quantity and quality of male gametes \citep[e.g.,][]{Yund2000, AizenHarder2007, Harder2016}.
+Modeling male gamete production is a natural extension to our framework, and would enable us to analyze how tension between the demographic consequences of SA fitness variation through both female and male function alters or abolishes asymmetries in the extinction thresholds and polymorphic space \textcolor{blue}{caused from female demographic dominance} \citep[e.g.,][]{Tazzyman2015}.
 Interestingly, different forms of self-fertilization can aid in reproductive assurance, and should therefore have demographic consequences -- for example, under pollen limitation and delayed selfing (where only ovules that fail to receive outcross pollen are selfed), the selfing rate will be a function of genotype frequencies because these will directly influence pollen production \citep{HarderBarrett2006}.
 
-\textcolor{blue}{Our assumption that how individuals themselves were produced (i.e., by selfing or outcrossing) affects the level of inbreeding depression they suffer also represents a major simplification}.
+\textcolor{blue}{Our assumption that only how individuals themselves were produced (by selfing or outcrossing) affects the level of inbreeding depression they suffer also represents a major simplification}.
 In reality, the history of consecutive generations of inbreeding in each individual's lineage will influence the severity of inbreeding depression they experience, particularly when inbreeding depression is caused primarily by recessive deleterious mutations.
-It would be an interesting and feasible extension of our modeling framework to expand the individual state space to include selfing cohorts (i.e., first generation selfing, second generation selfing, etc.), as in the models of \citet{kelly1999response,kelly2007mutation}, thereby enabling a more biologically plausible approach to modelling self-fertilization.
+It would be an interesting and feasible extension of our modeling framework to expand the individual state space to include selfing cohorts (i.e., first generation selfing, second generation selfing, etc.), as in \citet{kelly1999response,kelly2007mutation}, and thereby provide a more biologically meaningful approach to modelling self-fertilization and inbreeding depression.
 \textcolor{blue}{Likewise, the simple 'fixed prior selfing' model we implemented assumes constant selfing rates among individuals over time.
 In reality, environmental stochasticity will influence the population selfing rate and therefore both the expression of inbreeding depression and the scope for SA polymorphism \citep{JordanConnallon2014,Olito-etal-2018}.
-A naive expectation might be that some form of weighted geometric mean selfing rate would represent a more biologically meaningful parameter influencing the genetic and demographic outcomes of SA selection in more complicated 'selfing environments' \citep[e.g., ][]{ConnallonSharmaOlito2019}.}
+A naive expectation might be that a weighted geometric mean selfing rate would represent a more biologically meaningful parameter influencing the genetic and demographic outcomes of SA selection in more complicated 'selfing environments' \citep[e.g., ][]{ConnallonSharmaOlito2019}.}
 
@@
 
-Finally, \cite{grieshop2017male} found that genotypes with SA alleles that are male-beneficial experience higher levels of inbreeding depression than genotypes with female beneficial SA alleles.
-This effect can easily be included in our model by making the inbreeding depression parameters, $\delta$ and $\delta_i$, a function of the genotype of the individual.
-Such an effect would further reduce the demographically viable polymorphic parameter space, and increase the bias in viable SA polymorphisms towards alleles with weaker selection in females.
+Finally, \cite{grieshop2017male} found that genotypes carrying male-beneficial SA alleles experience more severe inbreeding depression than those with female-beneficial SA alleles.
+This effect can easily be included in our model by making the inbreeding depression parameters, $\delta$ and $\delta_i$, a function of the an individual's genotype.
+Such SA genotype-dependent inbreeding depression would further reduce the demographically viable polymorphic parameter space and increase the bias in viable SA polymorphisms towards alleles with weaker selection in females.
 
@@
 \textcolor{blue}{Here we extend the model presented in the main text by incorporating density-dependent population regulation.
-We chose the simplest population regulation possible in a structured population model: population regulation affecting all life history stages in the same way.
+For brevity, we chose the simplest population regulation possible in a structured population model: population regulation affecting all life history stages in the same way.
 Note that this is the only type of population regulation that is possible in unstructured population models, like the classical Lotka-Volterra or Rosenzweig-MacArthur models.
@@
 \end{linenomath*}
-where $\sum_i n_i$ sums over all entries of the population state vector, and hence over all genotypes in all life-history stages. Equivalently, in matrix algebra notation,
+where $\sum_i n_i$ sums over all entries of the population state vector, $\tilde{\mbf{n}}$, and hence over all genotypes in all life-history stages. Equivalently, in matrix algebra notation,
 \begin{linenomath*}
@@
 \end{linenomath*}
-where $\| \tilde{\mbf{n}}(t) \|$ is the one-norm which is identical to the sum of the entries of $\tilde{\mbf{n}}(t)$ because the population vector only has nonnegative entries. }
+where $\| \tilde{\mbf{n}}(t) \|$ is the one-norm which is identical to the sum of the entries of $\tilde{\mbf{n}}(t)$ because the population vector only has non-negative entries. }
 
@@
 
+\renewcommand\thesection{}
+\renewcommand\thesubsection{}
+
 %%%%%%%%%%%%%%%%%%%%%

### 2022-01-05 doc/ResponseLetter-R1.tex
Minor typos response letter, second sentence intro
Yay :)

@@
 \begin{quote}
-	{\itshape Agreed. Please see our sepecific responses to Reviewer 2 below.}
+	{\itshape Agreed. Please see our specific responses to Reviewer 2 below.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape While we feel that there is strong precedence for discussing the key results of a study before diving into the various caveats for with the analysis, yoru point is well taken that we need to more clearly acknowledge these limitations in the methods and direct readers to where these limitations are discussed.
+	{\itshape While we feel that there is strong precedence for discussing the key results of a study before diving into the various caveats for with the analysis, your point is well taken that we need to more clearly acknowledge these limitations in the methods and direct readers to where these limitations are discussed.
 In our Methods section, we are now careful to point readers to the various Appendixes, Supplements, and places in the Discussion where these limitations are dealt with (e.g., see {\bf L.94--98, 103--109, 128--134, 183--186, 187--189, 196--197, 207--212, 274--275}), and we have added more explicit discussion of several of our key assumptions in the discussion ({\bf e.g., L.519--581}).

### 2022-01-05 doc/SA-Hermaphrodites-wDemography-R1.tex
Minor typos response letter, second sentence intro
Yay :)

@@
 \textcolor{blue}{A population's ability to adapt to its environment and persist in the long term depends upon the genetic variation that it carries \citep{Fisher1930, Lewontin1975}.
-Yet, populations are heterogenous collections of individuals existing in a complex world where adaptation can be impeded by conflicting selection and gene-flow between different classes of individuals, or more generally, when different fitness components are negatively correlated \citep{Prout2000, CharlesworthHughes2000, ConnallonHall2018}.}
+Yet, populations are heterogenous collections of individuals existing in a complex world where adaptation can be impeded by conflicting selection and gene-flow between different classes of individuals \citep{Prout2000, CharlesworthHughes2000, ConnallonHall2018}.}
 

### 2022-01-05 doc/ResponseLetter-R1.tex
Final check before resubmitting

@@
 guttatus} example; and ({\itshape ii}) expanding our discussion of density-dependence to address these additions.
-Overall, we feel that these various additions have greatly improved the paper.
+Overall, we feel that these various revisions have greatly improved the paper.
 \bigskip
@@
 %%%%%%%%%%%%%%%%%%%%
-\section*{Decision letter with our responses}
+\section*{Decision letter with our responses (in italics)}
 
@@
 guttatus} example with a simple density-dependent version of the model.
-Second, we have expanded our methods, results, and especially the discussion to more fully address the implications of density-dependence, with reference to our modified figures and the new {\bf Appendix B} (see {\bf L.128--134; 444--447; 529--549}).}
+Second, we have expanded our methods, results, and especially the discussion to more fully address the implications of density-dependence, with reference to our modified figures and the new {\bf Appendix B} (see {\bf L.128--134; 456--459; 529--561}).}
 \end{quote}
@@
 	{\itshape This is great suggestion which inspired us to make several major changes.
-We now include a new illustrative figure ({\bf fig.~1}), and revised versions of figs.
-1 \& 4 (now {\bf figs.~S1 \& 5}) which show heatmaps of absolute values for predicted $\lambda$ instead of just extinction boundaries.
+We now include a new illustrative figure ({\bf fig.~1}), and revised versions of figs.~1 \&~4 (now {\bf figs.~S1 \& 5}) which show heatmaps of absolute values for predicted $\lambda$ instead of just extinction boundaries.
 Unfortunately, when presenting $\lambda$ in this way, it is no longer possible to show multiple values of maximum fertility ($f$) on a single panel.
-In particular, our old fig.
-1 now requires 18 panels to illustrate $\lambda$ for the same set of parameter combinations.
+In particular, our old fig.~1 now requires 18 panels to illustrate $\lambda$ for the same set of parameter combinations.
 We felt that this resulted in an unnecessarily large and complicated figure.
@@
 	{\itshape We can see now that in our efforts to appeal to both demographers and evolutionary biologists studying sexual conflict, we insufficiently discussed the relevant literature on both counts.
-We have restructured our introduction to better engage with these fields without lengthening it (see, in particular, {\bf L.43--52, L.58--67, L.76--85}).
+We have restructured our introduction to better engage with these fields without lengthening it (see, in particular, {\bf L.43--51, L.58--67}).
 See also our specific responses to Reviewer 1.
@@
 \begin{quote}
-	{\itshape We have added a brief general explanation of sexually antagonistic selection, which we felt was lacking in the first draft, and also allows us to more clearly link SA polymorphism with its' demographic consequences ({\bf L.2--6}).}
+	{\itshape We have added a brief explanation of sexually antagonistic selection to the beginning of the Abstract, which we felt was lacking in the first draft, and also allows us to more clearly link SA polymorphism with its' demographic consequences ({\bf L.2--6}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape We now briefly discuss this possibility in the Discussion ({\bf L.564--570}).}
+	{\itshape We now briefly discuss this possibility in the Discussion ({\bf L.581--587}).}
 \end{quote}
@@
 To address your main concern, we have restructured the Introduction to more directly connect with the broader literature on evolutionarily explicit structured models (see {\bf L.58--67}) and now also provide a stronger link with the empirical sexually antagonistic selection literature (see {\bf L.43--51}).
-We hope these changes, along with a more streamlined presentation of SA selection (see {\bf L.34--39; 52--54}) succeed in broadening the appeal of our manuscript to the wider AmNat readership.}
+We hope these changes, along with a more streamlined opening ({\bf L.20--23}) and presentation of SA selection ({\bf L.34--54}) succeed in engaging with the broader AmNat readership.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape While we feel that there is strong precedence for discussing the key results of a study before diving into the various caveats for with the analysis, your point is well taken that we need to more clearly acknowledge these limitations in the methods and direct readers to where these limitations are discussed.
+	{\itshape While we feel that there is strong precedence for discussing the key results of a study before diving into the various caveats for with the analysis, your point is well taken that we need to more clearly acknowledge these limitations in the Methods section and direct readers to where they are discussed.
 In our Methods section, we are now careful to point readers to the various Appendixes, Supplements, and places in the Discussion where these limitations are dealt with (e.g., see {\bf L.94--98, 103--109, 128--134, 183--186, 187--189, 196--197, 207--212, 274--275}), and we have added more explicit discussion of several of our key assumptions in the discussion ({\bf e.g., L.519--581}).
-We discuss both the causes and potential effects of incorporating male-limitation of reproductive success on {\bf L.550--561}}.
+We discuss the causes and potential effects of incorporating male-limitation of reproductive success on {\bf L.550--561}.
+We are reluctant to speculate much further on the potential effects of incorporating male gamete production because this will simultaneously influence both the invasion conditions and population growth rates in different regions of parameter space}.
 \end{quote}
@@
 \begin{quote}
-	{\itshape These are good points, that was awkward wording on our part.
+	{\itshape These are good points, and that was awkward wording on our part.
 While revising the introduction, we felt that much of this first paragraph was tangential to the key concepts we wanted to introduce.
@@
 \begin{quote}
-	{\itshape We have rephrased the first sentence of this paragraph to be more specific ({\bf L.25--26}), which we hope also addresses this comment.}
+	{\itshape We have combined this paragraph with the previous one in this revision.
+We have edited the opening of this combined paragraph to be more specific ({\bf L.34--39}), which we hope also addresses this comment.}
 \end{quote}
@@
 \begin{quote}
-	{\itshape We have added a few lines to discuss frequency vs density dependence when we first introduce the population projection matrix in equation 3 ({\bf L.131--132}).
-In addition, we have tried to clarify to improve this section by discussing that the survival and fertility rates ("demographic rates") are density-indepedent, and that the allele frequency dynamics make the model frequency dependent ({\bf L.128--134}).}
+	{\itshape We have added a few lines to discuss frequency vs density dependence shortly after introducing the population projection matrix in Eq(3).
+Additionally, we have tried to improve this section by discussing that the survival and fertility rates ("demographic rates") are density-indepedent, and that the allele frequency dynamics make the model frequency dependent ({\bf L.128--134}).}
 \end{quote}
@@
 \begin{quote}
-	{\itshape We have rephrased this sentence to clarify and be more consistent with our language earlier in the preceding paragraphs ({\bf L.39--42}).}
+	{\itshape See comment above regarding {\bf L.34--39}.}
 \end{quote}

### 2022-01-05 doc/SA-Hermaphrodites-wDemography-R1.tex
Final check before resubmitting
update expired anonymous GitHub link

@@
 The derivation of the general model for any number of age or stage classes is provided in the Online Supplementary Material.} The derivation follows closely the logic presented in \citet{deVriesCaswell2019a, deVriesCaswell2019b}, and in fact, it reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019a} under obligate outcrossing \textcolor{blue}{(when selfing is not allowed)}.
-All computer code necessary to reproduce the results are available at \url{https://anonymous.4open.science/r/SA-Hermaphrodites-wDemography-BF66}.%\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
+All computer code necessary to reproduce the results are available at \url{https://anonymous.4open.science/r/SA-Hermaphrodites-wDemography-4FFF}.%\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
 

### 2022-01-22 doc/SA-Hermaphrodites-wDemography-R1.tex
Update ms with Dryad info

@@
 
-% ALTERNATIVE TITLES:
-%Evolutionary demography and the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites
-
-% Population viability affects/impacts the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites
-
- % Demographic viability of antagonistic alleles impacts conditions for the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites
-
- %Conditions for the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites impacted by considering population viability
-
-%\author{Colin Olito$^{1,\ast}$ \\ 
-%Charlotte de Vries$^{2}$}
+\author{Colin Olito$^{1,\ast}$ \\ 
+Charlotte de Vries$^{2}$}
 \date{\today}
@@
 
-%\noindent{} 1. Department of Biology, Lund University, Lund 223 62, Sweden;
-
-%\noindent{} 2.  Department of Evolutionary Biology and Environmental Studies, University of Zurich, Zurich CH-8057, Switzerland;
-
-%\noindent{} $\ast$ Corresponding author; e-mail: colin.olito@gmail.com
+\noindent{} 1. Department of Biology, Lund University, Lund 223 62, Sweden;
+
+\noindent{} 2.  Department of Evolutionary Biology and Environmental Studies, University of Zurich, Zurich CH-8057, Switzerland;
+
+\noindent{} $\ast$ Corresponding author; e-mail: \url{colin.olito@gmail.com}
 
@@
 \textit{Manuscript type}: Article. %Or e-article, note, e-note, natural history miscellany, e-natural history miscellany, comment, reply, invited symposium, or historical perspective.
+
+\bigskip
+
+\textit{Online enhancements}: Dryad digital repository ({\itshape https://doi.org/10.5061/dryad.c2fqz619t}).
 
@@
 
-\textcolor{blue}{When selection differs between the sexes, genes expressed by both males and females can experience sexually antagonistic (SA) selection, where beneficial alleles for one sex are deleterious for the other.} Classic population genetics theory has been fundamental to understanding \textcolor{blue}{how and when SA genetic variation can be maintained by balancing selection}, but these models have rarely considered the demographic consequences \textcolor{blue}{of coexisting alleles with deleterious fitness effects in each sex.} In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (i.e., dioecious/gonochorous) and partially selfing hermaphrodite populations.
+When selection differs between the sexes, genes expressed by both males and females can experience sexually antagonistic (SA) selection, where beneficial alleles for one sex are deleterious for the other.
+Classic population genetics theory has been fundamental to understanding how and when SA genetic variation can be maintained by balancing selection, but these models have rarely considered the demographic consequences of coexisting alleles with deleterious fitness effects in each sex.
+In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (i.e., dioecious/gonochorous) and partially selfing hermaphrodite populations.
 We focus on identifying when SA polymorphisms are maintained by balancing selection {\itshape and} the population growth rate remains positive.
@@
 
-%\textcolor{blue}{A sexually antagonistic polymorphism is a polymorphism of two alleles with antagonistic effects on the two sexes, that is, one of the two coexisting allele confers a fitness benefit to females and reduces male fitness, and the other allele confers a fitness benefit to males and reduces female fitness.
-}
-
-%\section*{Alternative Abstract -- currently 215/200 word max.}
-%%%%%%%%%%%%%%%%%%%%
-%Genetic trade-offs between male and female fitness impose fascinating genetic constraints on adaptation.
-On one hand they can maintain genetic variation for a variety of life-history and gender-related traits, while on the other hand they cause unavoidable maladaptation as individuals are prevented from reaching their sex-specific fitness optima.
-In this paper we develop a stage-structured mendelian matrix model and jointly analyze the evolutionary and demographic consequences of sexually antagonistic (SA) selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations.
-We focus on identifying the parameter conditions under which SA polymorphisms are maintained {\itshape and} the population intrinsic growth rate remains positive.
-We also analyze the effects of inbreeding depression manifesting in different life-history stages.
-Finally, we show an illustrative example of the potential for SA polymorphism in real populations using empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus}.
-Our results indicate that when population growth rates are near one, large swathes of polymorphic parameter space become demographically inviable, and that inbreeding depression poses a significant challenge to the maintenance of SA polymorphism in partially selfing populations.
-Despite these demographic challenges, our example with {\itshape M.~guttatus} appears to show that demographic rates observed in some real populations are capable of sustaining large regions of viable SA polymorphic space.
-
 \newpage{}
@@
 
-%To persist in the long term, a population must be able to adapt to its environment.
-Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, such as deleterious mutations \citep{Haldane1937}, \textcolor{blue}{rapid, unpredictable, or extreme changes} in environmental conditions \citep{Maynard-Smith1976, LandeShannon1996, OrrUnckless2008,Lyberger-etal-2021}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, and genetic constraints \citep{ConnallonHall2018, MatthewsConnallon2019}.
-Genetic constraints on adaptation can arise as a consequence of conflicting selection and gene-flow between different classes of individuals, or more generally, when \textcolor{blue}{different fitness components are negatively correlated} \citep{CharlesworthHughes2000, ConnallonHall2018}.
-\textcolor{blue}{A population's ability to adapt to its environment and persist in the long term depends upon the genetic variation that it carries \citep{Fisher1930, Lewontin1975}.
-Yet, populations are heterogenous collections of individuals existing in a complex world where adaptation can be impeded by conflicting selection and gene-flow between different classes of individuals \citep{Prout2000, CharlesworthHughes2000, ConnallonHall2018}.}
-
-Genetic constraints on adaptation \textcolor{blue}{due to conflicting selection }have particularly interesting evolutionary consequences.
+A population's ability to adapt to its environment and persist in the long term depends upon the genetic variation that it carries \citep{Fisher1930, Lewontin1975}.
+Yet, populations are heterogenous collections of individuals existing in a complex world where adaptation can be impeded by conflicting selection and gene-flow between different classes of individuals \citep{Prout2000, CharlesworthHughes2000, ConnallonHall2018}.
+
+Genetic constraints on adaptation due to conflicting selection have particularly interesting evolutionary consequences.
 On one hand, they provide an effective mechanism for the maintenance of genetic variation, which shapes a population's capacity for future adaptation \citep{Fisher1930, Prout2000, CharlesworthHughes2000, ConnallonHall2018}.
@@
 
-%\textcolor{blue}{Genetic constraints on adaptation arising from conflicting selection and gene flow} have particularly interesting evolutionary consequences \citep{ConnallonDebarreLi2018}.
-On one hand, they provide an effective mechanism for the maintenance of genetic variation, which can increase a population's capacity for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}.
-On the other hand, they prevent individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase a population's overall extinction risk \citep{kokko2003sexy,harts2014demography}.
-Hence, for traits under conflicting selection, the nature and extent of genetic variation observed in natural populations should reflect a balance between the maintainance of genetic polymorphisms, and the population dynamical consequences of the resulting maladaptation.
-
-Sexually antagonistic selection (abbreviated SA hereafter) \textcolor{blue}{is an important form of conflicting selection that occurs when beneficial alleles for one sex are deleterious when expressed in the other.
+Sexually antagonistic selection (abbreviated SA hereafter) is an important form of conflicting selection that occurs when beneficial alleles for one sex are deleterious when expressed in the other.
 Likewise, alleles with opposing fitness effects on male and female sex functions can experience SA selection in hermaphrodite populations, where individual fitness is determined jointly by maternal and paternal reproductive success \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
-SA selection appears to be} a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and the evolution of sexual dimorphism and sex-related traits \citep{Kidwell1977, Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, ConnallonClark2012, BondurianskyChenoweth2009, Olito2019}.
-\textcolor{blue}{For instance, sex-differences in directional selection are commonly detected in phenotypic selection studies \citep[e.g.,][]{DeLisle-etal-2018,Lewis-etal-2011,CoxCalsbeek2009,SinghPunzalan2018}, suggesting that many genetic variants contributing to quantitative traits have SA effects.
+SA selection appears to be a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and the evolution of sexual dimorphism and sex-related traits \citep{Kidwell1977, Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, ConnallonClark2012, BondurianskyChenoweth2009, Olito2019}.
+For instance, sex-differences in directional selection are commonly detected in phenotypic selection studies \citep[e.g.,][]{DeLisle-etal-2018,Lewis-etal-2011,CoxCalsbeek2009,SinghPunzalan2018}, suggesting that many genetic variants contributing to quantitative traits have SA effects.
 Moreover, quantitative genetic studies frequently find negative cross-sex genetic correlations for fitness (\citealt{Chippendale-etal-2001,Delph-etal-2011}; reviewed in \citealt{ConnallonMatthews2019}), and polygenic signals of SA selection are evident in several recent population genetic analyses of large genomic datasets \citep{RuzickaESEB2020, RuzickaConnallon2022}.
-Overall, both theory and current empirical data suggest that there is ample scope for the maintenance of SA \textcolor{blue}{genetic variation} in both dioecious \textcolor{blue}{(separate-sexed)} and hermaphrodite populations \citep{BondurianskyChenoweth2009, Abbott2011, RuzickaESEB2020, WangBarrett2020}.}
-
-% Sexually antagonistic selection (abbreviated SA hereafter) \textcolor{blue}{is an important form of conflicting selection that occurs when beneficial alleles for one sex are deleterious when expressed in the other.
-SA selection appears to be} a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and the evolution of sexual dimorphism and sex-related traits \citep{Kidwell1977, Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, ConnallonClark2012, BondurianskyChenoweth2009, Olito2019}.
-\textcolor{blue}{For instance, phenotypic selection studies commonly find sex-differences in directional selection \citep[e.g.,][]{DeLisle-etal-2018,Lewis-etal-2011,CoxCalsbeek2009,SinghPunzalan2018}, implying that many genetic variants underlying quantitative traits have SA fitness effects.
-Likewise, quantitative genetic studies frequently find negative cross-sex genetic correlations for fitness (\citealt{Chippendale-etal-2001,Delph-etal-2011}; reviewed in \citealt{ConnallonMatthews2019}), and polygenic signals of SA selection are evident in recent population genetic analyses of large genomic datasets \citep{RuzickaESEB2020, RuzickaConnallon2022}.}
-
-% Alleles with opposing fitness effects through male and female sex functions can also cause SA selection in hermaphrodite populations, where individuals' overall fitness is determined jointly by their maternal and paternal reproductive success \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.
-\textcolor{blue}{In obligately outcrossing hermaphrodite populations, SA selection operates similarly as in seperate-sexed ones (dioecious species).
-However, the possibility of self-fertilization has important consequences for the maintenance of SA genetic variation in hermaphrodites}.
-Self-fertilization reduces \textcolor{blue}{opportunities for} balanced SA polymorphisms, while simultaneously creating a bias in selection through the female sex function (\citealt{JordanConnallon2014,Glemin2021}; but see \citealt{Tazzyman2015}).
-Yet, other factors, such as genetic linkage to other SA loci or a sex-determining region \citep{Otto2011, JordanCharlesworth2012, Olito2017, Olito2019}, antagonism between viability and fecundity selection \citep{Glemin2021}, or spatial heterogeneity and complexity of the life-cycle \textcolor{blue}{can all facilitate the maintenance of} SA polymorphisms \citep{Olito-etal-2018,ConnallonSharmaOlito2019, Glemin2021}.
-Overall, both theoretical predictions and current empirical data suggest that there is ample scope for SA \textcolor{blue}{fitness variation} and the maintenance of SA polymorphisms in both dioecious and hermaphrodite populations \citep{Abbott2011, DelphHerlihy2012, WangBarrett2020, RuzickaESEB2020, RuzickaConnallon2022}.
-
-\textcolor{blue}{While significant efforts have been made to predict when SA polymorphisms can be maintained and to detect SA genetic variation from empirical data, the demographic consequences of SA selection have often been overlooked \citep{MatthewsConnallon2019}}.
+Overall, both theory and current empirical data suggest that there is ample scope for the maintenance of SA genetic variation in both dioecious (separate-sexed) and hermaphrodite populations \citep{BondurianskyChenoweth2009, Abbott2011, RuzickaESEB2020, WangBarrett2020}.
+
+While significant efforts have been made to predict when SA polymorphisms can be maintained and to detect SA genetic variation from empirical data, the demographic consequences of SA selection have often been overlooked \citep{MatthewsConnallon2019}.
 Most population genetic theory, including SA selection theory, models the relative fitness of genotypes in populations of constant (and often infinite) size.
-In reality, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals \textcolor{blue}{that link genotype distributions to population dynamics (e.g., \citealt {johnston2013life, merot2020balancing}).
+In reality, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals that link genotype distributions to population dynamics (e.g., \citealt {johnston2013life, merot2020balancing}).
 For over 50 years, the field of eco-evolutionary dynamics has sought to jointly model the genetic and demographic processes involved in evolutionary change.
@@
 Various extensions of these theories have been made to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and quantitative trait evolution \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects}.
-This growing body of eco-evolutionary theory makes plain that demographic processes must be taken into account to fully understand how genetic variation for fitness is shaped in natural populations \citep{metcalf2007evolutionary,coulson2006putting,rueffler2006evolutionary}.}
-
-%a fuller understanding of the genetic basis of fitness variation, and the processes maintaining it, requires an explicit treatment of the demographic processes involved in evolutionary change.
+This growing body of eco-evolutionary theory makes plain that demographic processes must be taken into account to fully understand how genetic variation for fitness is shaped in natural populations \citep{metcalf2007evolutionary,coulson2006putting,rueffler2006evolutionary}.
 
@@
 \citet{deVriesCaswell2019a} introduced a mendelian matrix model with SA selection and population dynamics, but did not analyze how demography influenced SA polymorphism.
-In this paper, we extend the model of \citet{deVriesCaswell2019a,deVriesCaswell2019b} to \textcolor{blue}{include both dioecious and hermaphroditic populations and jointly analyze the genetic (fixation/polymorphism) and demographic (persistence/extinction) outcomes of SA selection}.
+In this paper, we extend the model of \citet{deVriesCaswell2019a,deVriesCaswell2019b} to include both dioecious and hermaphroditic populations and jointly analyze the genetic (fixation/polymorphism) and demographic (persistence/extinction) outcomes of SA selection.
 A major strength of our approach is that the model can be parameterized using empirically estimated demographic rates, enabling us to make predictions about the scope for SA polymorphisms that are grounded in the biology of real populations.
@@
 
-% \bigskip
-
-% \begin{enumerate}
-% 	\item Introduce Evolutionary Demography, pivot to focus on the consequences of taking demography into account explicitly for the maintenance of SA polymorphism. [Lotte]
-% 	\item Introduce key difference between existing population genetic models of SA polymorphism and our models: possibility of diverse population and evolutionary dynamics. [Lotte then Colin]
-% 	\item Explain and highlight "demographically viable" parameter space, indicate that we are interested in whether parameter space predicted to be polymorphic by Pop.~gen.~models is, indeed, "demographically viable" [Lotte then Colin]
-% 		\begin{enumerate}
-% 			\item Pop.
-Gen.
-models tacitly assume large and/or stable population size, parameterize genotypic fitnesses relative to the 'best performing' genotype.
-Non-overlapping generations.
-Casually define fitness as lifetime reproductive fitness of a given genotype.
-% 			\item "In practice, these genotypic fitnesses emerge from myriad process acting throughout the life-history of individuals"
-% 			\item these processes contribute to key demographic rates such as the intrinsic population growth rate, 
-% 		\end{enumerate}
-% 		\item A major strength of the model is that we can use empirically estimated demographic rates to parameterize the model, enabling us to make predictions about the scope for demographically viable SA polymorphisms that are grounded in the biology of real populations.
-% \end{enumerate}
-% \bigskip
-
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 Here, we briefly describe a matrix model incorporating multiple life-cycle stages and a single diallelic locus under SA selection on male and female fertility for populations of partially-selfing simultaneous hermaphrodites.
-\textcolor{blue}{We focus our presentation on a two-stage example of the model and discuss its' underlying biological assumptions.
-The derivation of the general model for any number of age or stage classes is provided in the Online Supplementary Material.} The derivation follows closely the logic presented in \citet{deVriesCaswell2019a, deVriesCaswell2019b}, and in fact, it reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019a} under obligate outcrossing \textcolor{blue}{(when selfing is not allowed)}.
-All computer code necessary to reproduce the results are available at \url{https://anonymous.4open.science/r/SA-Hermaphrodites-wDemography-4FFF}.%\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.
+We focus our presentation on a two-stage example of the model and discuss its' underlying biological assumptions.
+The derivation of the general model for any number of age or stage classes is provided in the Online Supplementary Material.
+The derivation follows closely the logic presented in \citet{deVriesCaswell2019a, deVriesCaswell2019b}, and in fact, it reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019a} under obligate outcrossing (when selfing is not allowed).
+All computer code and data necessary to reproduce the results are available on Dryad (\url{https://doi.org/10.5061/dryad.c2fqz619t}) and on GitHub (\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}).
 
@@
 To distinguish between parameters relating to male and female reproductive functions, we denote matrices or vectors relating to the male sex function with a prime.
-\textcolor{blue}{Following previous theory for partially selfing populations \cite[e.g.,][]{Charlesworth2010,JordanConnallon2014,Glemin2021}, we use a 'fixed prior selfing' model where all individuals are assumed to self-fertilize at a constant rate, $C$, prior to receiving outcross pollen.
-We present a more general model of genotype-specific self-fertilization rates in the Online Supplement, and briefly discuss implications of this approach in the Discussion.}
+Following previous theory for partially selfing populations \cite[e.g.,][]{Charlesworth2010,JordanConnallon2014,Glemin2021}, we use a 'fixed prior selfing' model where all individuals are assumed to self-fertilize at a constant rate, $C$, prior to receiving outcross pollen.
+We present a more general model of genotype-specific self-fertilization rates in the Online Supplement, and briefly discuss implications of this approach in the Discussion.
 
@@
 We include how individuals were produced in the state description because individuals produced by selfing might experience reduced survival, growth, or maturation rates due to inbreeding depression.
-\textcolor{blue}{This approach makes the common simplifying assumption that all individuals in the population experience the same severity of inbreeding depression (e.g., \citealt{Charlesworth1987, Charlesworth2009, Charlesworth2010, JordanConnallon2014}).
+This approach makes the common simplifying assumption that all individuals in the population experience the same severity of inbreeding depression (e.g., \citealt{Charlesworth1987, Charlesworth2009, Charlesworth2010, JordanConnallon2014}).
 We address potential consequences of relaxing this assumption, and how our model can be extended to do so, in the discussion.
-}
 
@@
 $C$ & The population rate of self-fertilization &0-1 \\
-$\delta$ &\textcolor{blue}{ Inbreeding depression effect on zygote viability}&0-0.8\\
-$ \delta_j$ & \textcolor{blue}{ Inbreeding depression effect on juvenile survival}& 0-0.8\\
-$\delta_a$ & \textcolor{blue}{ Inbreeding depression effect on adult survival}& 0-0.8\\
-$\delta_{\gamma}$ &\textcolor{blue}{Inbreeding depression effect on maturation rate} & 0-0.8\\
-$f_i,\,f^{\prime}_i$ & Adult \textcolor{blue}{maximum} fertilities through female and male sex functionsxi & 5.8-8.5 \\
+$\delta$ & Inbreeding depression effect on zygote viability&0-0.8\\
+$ \delta_j$ &  Inbreeding depression effect on juvenile survival & 0-0.8\\
+$\delta_a$ &  Inbreeding depression effect on adult survival & 0-0.8\\
+$\delta_{\gamma}$ &Inbreeding depression effect on maturation rate & 0-0.8\\
+$f_i,\,f^{\prime}_i$ & Adult maximum fertilities through female and male sex functions & 5.8-8.5 \\
 
@@
 
-\textcolor{blue}{If the survival or fertility rates are a function of the population density, $\tilde{\mbf{n}}(t)$, then the model will be density-dependent and the population projection matrix becomes a function of the full population vector, $\tilde{\mbf{A}}[\tilde{\mbf{n}}(t)]$.
+If the survival or fertility rates are a function of the population density, $\tilde{\mbf{n}}(t)$, then the model will be density-dependent and the population projection matrix becomes a function of the full population vector, $\tilde{\mbf{A}}[\tilde{\mbf{n}}(t)]$.
 Below, we limit our analyses to consider density-independent survival and fertility rates.
 The population projection matrix is then only a function of the proportional population vector, $\tilde{\mbf{A}}[\tilde{\mbf{p}}(t)]$, and the model is referred to as frequency-dependent.
-In \ref{App:DensDep} we construct and analyze a simple density-dependent version of the model and show that with density-dependent survival affecting all stages equally, our main conclusions remain intact.}
+In \ref{App:DensDep} we construct and analyze a simple density-dependent version of the model and show that with density-dependent survival affecting all stages equally, our main conclusions remain intact.
 
@@
 
-\textcolor{blue}{We now construct and analyze an example genotype $\times$ stage-classified model for a hypothetical hermaphrodite species with intralocus sexual conflict via the two sex functions.} For the sake of simplicity, we assume our hypothetical species has a life cycle with only two stages: juveniles and adults (i.e., $\omega = 2$), and that only adults can reproduce.
+We now construct and analyze an example genotype $\times$ stage-classified model for a hypothetical hermaphrodite species with intralocus sexual conflict via the two sex functions.
+For the sake of simplicity, we assume our hypothetical species has a life cycle with only two stages: juveniles and adults (i.e., $\omega = 2$), and that only adults can reproduce.
 Suppose that there is a genetic trade-off between the sex-functions at a single diallelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect.
@@
 
-\noindent where $f$ and $f^{\prime}$ represent \textcolor{blue}{maximum adult fertilities}, and $w_{i}$ and $w^{\prime}_{i}$ the genotypic relative scaling factors for female and male sex-functions (see Table \ref{tab:Fitness}).
+\noindent where $f$ and $f^{\prime}$ represent maximum adult fertilities, and $w_{i}$ and $w^{\prime}_{i}$ the genotypic relative scaling factors for female and male sex-functions (see Table \ref{tab:Fitness}).
 
@@
 These chosen values are similar to those used in \citet{deVriesCaswell2019b}, facilitating comparison between models, and correspond to a life-history in which individuals spend multiple timesteps in the juvenile phase prior to maturing into reproductively active adults, but are otherwise arbitrary.
-Our parameters of interest include \textcolor{blue}{maximum female} fertility, $f$, the inbreeding depression parameters, $\delta$, $\delta_j$, $\delta_a$, $\delta_{\gamma}$, and the selection parameters $h_f$, $s_f$, $h_m$, and $s_m$, which are given different values for each analyses as described in the figure captions.
+Our parameters of interest include maximum female fertility, $f$, the inbreeding depression parameters, $\delta$, $\delta_j$, $\delta_a$, $\delta_{\gamma}$, and the selection parameters $h_f$, $s_f$, $h_m$, and $s_m$, which are given different values for each analyses as described in the figure captions.
 
@@
 
- \textcolor{blue}{ For the purposes of this article, we assume that all outcross matings are random, and that individuals produced by selfing or outcrossing have the same mating success, though they can differ in their survival (see above).
+ For the purposes of this article, we assume that all outcross matings are random, and that individuals produced by selfing or outcrossing have the same mating success, though they can differ in their survival (see above).
 Given these assumptions, the allele frequencies in the mating population can obtained by summing the vectors of individuals produced through selfing and outcrossing:
@@
  \end{equation}
-\end{linenomath*}}
-\textcolor{blue}{Individuals contribute gametes to the population outcrossing gamete pool according to their fertility, which depends on their stage and genotype at the SA locus.
+\end{linenomath*}
+Individuals contribute gametes to the population outcrossing gamete pool according to their fertility, which depends on their stage and genotype at the SA locus.
 The allele frequencies in the male gamete pool are then calculated as
@@
 The first entry sums the  stage-specific contributions of allele $A$ to the gamete pool by $AA$ homozygotes and heterozygotes; the second entry does the same for the stage-specific contributions of allele $a$ to the male gamete pool.
-}
-
-\textcolor{blue}{ We now have all the ingredients to calculate the allele frequencies in the male gamete pool for the two-stage model in Eq(\ref{eq:F_male}), Eq(\ref{eq:maleGametePool}), and Table \ref{tab:Fitness}.
+
+We now have all the ingredients to calculate the allele frequencies in the male gamete pool for the two-stage model in Eq(\ref{eq:F_male}), Eq(\ref{eq:maleGametePool}), and Table \ref{tab:Fitness}.
 Denoting the proportion of reproducing adults of genotype $i$ with $p^R_i$, the allele frequencies in the male gamete pool are given by
@@
 Note also that by modeling allele frequencies in the male gamete pool rather than sperm/pollen production, we are making the common assumption of female demographic dominance (i.e., population growth rates are determined by female fecundity, not production of male gametes; \citealt{Caswell2001,iannelli2005gender,pollard1975mathematical}).
-We elaborate on several implications of this assumption for hermaphroditic populations in the discussion.}
+We elaborate on several implications of this assumption for hermaphroditic populations in the discussion.
 
@@
 Diverse eco-evolutionary outcomes are possible in the model, including fixation of either allele, balanced polymorphism, population growth or extinction, and even evolutionary rescue and suicide \citep[see][for examples of the possible outcomes]{deVriesCaswell2019a, deVriesCaswell2019b}.
-We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection \textcolor{blue}{(Figure \ref{fig:didactic}, left panel)} and ({\itshape ii}) the intrinsic population growth rate at equilibrium, $\lambda$, is greater than $1$ \textcolor{blue}{(Figure \ref{fig:didactic}, right panel)}; a situation that we refer to as a 'demographically viable SA polymorphism'.
+We focus on identifying parameter conditions where two criteria are satisfied: ({\itshape i}) SA polymorphism is maintained under balancing selection (Figure \ref{fig:didactic}, left panel) and ({\itshape ii}) the intrinsic population growth rate at equilibrium, $\lambda$, is greater than $1$ (Figure \ref{fig:didactic}, right panel); a situation that we refer to as a 'demographically viable SA polymorphism'.
 
@@
 Specifically, for each boundary equilibrium we introduced the rare allele at low initial frequency and iterated Eq(\ref{eq:Atilde}) until the population had reached demographic and genotypic equilibrium.
-\textcolor{blue}{Although we use density-independent survival and fertility rates, the allele frequency dynamics introduce frequency-dependence into the model.} The population state vector will therefore grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see chapter 17 in \citealt{Caswell2001}).
-The intrinsic population growth rate after convergence, $\lambda$, can be calculated as \textcolor{blue}{ $n(t)/n(t+1)$, where $n = \sum_i n_i $ sums over all entries of the population state vector $\tilde{\mbf{n}}$.}
-
-% Unfortunately, closed-form solutions for the invasion conditions for each allele and the equilibrium frequencies under different selection scenarios are analytically intractable except for restricted conditions (see \hl{Online Appendix X}).
-We therefore used numerical techniques to identify invasion and extinction thresholds, which could then be used to define regions of demographically viable polymorphism across $s_f \times s_m$ parameter space.
+Although we use density-independent survival and fertility rates, the allele frequency dynamics introduce frequency-dependence into the model.
+The population state vector will therefore grow or shrink exponentially after converging to stable population structure and genotypic frequencies (see chapter 17 in \citealt{Caswell2001}).
+The intrinsic population growth rate after convergence, $\lambda$, can be calculated as $n(t)/n(t+1)$, where $n = \sum_i n_i $ sums over all entries of the population state vector $\tilde{\mbf{n}}$.
+
 Because single-locus selection coefficients are generally weak (e.g., \citealt{Eyre-WalkerKeightly2007}) and strongly skewed, we limit our analyses to coefficients within $0 < s_f,s_m \leq 0.15$, unless stated otherwise.
@@
 \begin{linenomath*}
-\textcolor{blue}{
 \begin{equation}
@@
   \label{eq:Canddelta}
-\end{equation}}
-\end{linenomath*}
-where $\delta^{\ast}$ is the hypothetical severity of inbreeding depression for a completely outcrossing population, $b$ is a shape parameter determining how far $\delta$ will decline under complete selfing (when $C = 1$), \textcolor{blue}{and $a$ is a shape parameter} controlling the curvature of the overall function for $\delta$ (see Appendix E in \citealt{Olito2019} for additional details).
+\end{equation}
+\end{linenomath*}
+where $\delta^{\ast}$ is the hypothetical severity of inbreeding depression for a completely outcrossing population, $b$ is a shape parameter determining how far $\delta$ will decline under complete selfing (when $C = 1$), and $a$ is a shape parameter controlling the curvature of the overall function for $\delta$ (see Appendix E in \citealt{Olito2019} for additional details).
 We set $\delta^{\ast} = 0.8$, $b = 0.5$, and $a = 0.2$ for all analyses involving inbreeding depression, values chosen to be consistent with empirical estimates of inbreeding depression (e.g., fig.~2 in \citealt{HusbandSchemske1996}).
@@
 Results are shown for three different population selfing rates ($C = \{0,\,1/4,\,1/2\}$), and two dominance scenarios (additivity, where $h = 1/2$, and dominance reversal, where $h = 1/4$).
-Extinction thresholds are illustrated for three different values of female fecundity ($f$ values annotated on each panel); \textcolor{blue}{ the coexistence boundaries are unaffected by the change in $f$.} \textcolor{blue}{Surface plots of predicted population intrinsic growth rates ($\lambda$) for the same parameter conditions are presented in fig.~S1 of the Online Supplementary Material.}}}
+Extinction thresholds are illustrated for three different values of female fecundity ($f$ values annotated on each panel); the coexistence boundaries are unaffected by the change in $f$.} Surface plots of predicted population intrinsic growth rates ($\lambda$) for the same parameter conditions are presented in fig.~S1 of the Online Supplementary Material.}
  \label{fig:extThresholds}
@@
 We begin with an illustration of demographically viable polymorphic parameter space in the absence of inbreeding depression in Figure \ref{fig:extThresholds} (i.e., $\delta = \delta_i = 0$).
-The red regions in Figure \ref{fig:extThresholds} indicate areas where the population goes extinct (\textcolor{blue}{i.e., where $\lambda < 1$}) for different \textcolor{blue}{maximum female fertility values}.
+The red regions in Figure \ref{fig:extThresholds} indicate areas where the population goes extinct (i.e., where $\lambda < 1$) for different maximum female fertility values.
 Lower female fertility values correspond to larger areas where extinction occurs.
 The funnel between the thick black lines indicates the area where the sexually antagonistic alleles coexist.
-Above the funnel, the female-benefit/male-detriment allele goes to fixation; below the funnel, the male-benefit/female-detriment allele \textcolor{blue}{goes to fixation}.
-Much of the region where the male-benefit/female-detriment allele \textcolor{blue}{fixes} is demographically inviable due to the demographic consequences of reduced female fertility in this region.
+Above the funnel, the female-benefit/male-detriment allele goes to fixation; below the funnel, the male-benefit/female-detriment allele goes to fixation.
+Much of the region where the male-benefit/female-detriment allele fixes is demographically inviable due to the demographic consequences of reduced female fertility in this region.
 The SA polymorphisms that remain viable at lower fertility values correspond to regions where the female-deleterious allele is predicted to segregate at low frequencies, or in regions with weak selection through both sex functions, an asymmetry that reflects the assumption of female demographic dominance.
@@
 
-Unlike previous population genetic models \textcolor{blue}{of SA selection in hermaphrodites}, which assume constant population sizes \citep{JordanConnallon2014,Olito2017}, mortality caused by inbreeding depression can strongly influence population persistence in our evolutionary demographic model.
+Unlike previous population genetic models of SA selection in hermaphrodites, which assume constant population sizes \citep{JordanConnallon2014,Olito2017}, mortality caused by inbreeding depression can strongly influence population persistence in our evolutionary demographic model.
 Populations with high fertility rates can maintain positive population growth rates despite this additional mortality.
 This causes a greater proportion of SA polymorphic parameter space to be demographically viable, with the demographic model predictions converging on those from population genetic models in high-fertility populations (fig.~\ref{fig:deltaPolySpace}).
-\textcolor{blue}{Note, however, that mortality due to late-acting inbreeding depression also reduces the population effective selfing rate in our model, and therefore shifts the invasion boundaries for the SA alleles to more closely resemble an outcrossing population.
-In some cases, this leads to larger regions of viable polymorphic space than predicted by the population genetic models, which only include early-acting inbreeding depression (e.g., see cross symbols in fig.~\ref{fig:deltaPolySpace})}.
+Note, however, that mortality due to late-acting inbreeding depression also reduces the population effective selfing rate in our model, and therefore shifts the invasion boundaries for the SA alleles to more closely resemble an outcrossing population.
+In some cases, this leads to larger regions of viable polymorphic space than predicted by the population genetic models, which only include early-acting inbreeding depression (e.g., see cross symbols in fig.~\ref{fig:deltaPolySpace}).
 However, as the selfing rate increases, demographic viability eventually crashes when the population can no longer sustain the concomitant increase in mortality due to inbreeding depression (fig.~\ref{fig:deltaPolySpace}).
@@
 Inbreeding depression at the juvenile stage therefore makes it harder to survive long enough to mature.
-\textcolor{blue}{Of course the strength of this effect is influenced by the maturation rate, $\gamma$, which therefore also affects the shape of the curves in Figure \ref{fig:deltaPolySpace}.} Although early-acting inbreeding depression ($\delta$) manifests earlier in the life-cycle than juvenile survival, it acts only once by influencing the total number of self-fertilized zygotes that become juveniles.
+Of course the strength of this effect is influenced by the maturation rate, $\gamma$, which therefore also affects the shape of the curves in Figure \ref{fig:deltaPolySpace}.
+Although early-acting inbreeding depression ($\delta$) manifests earlier in the life-cycle than juvenile survival, it acts only once by influencing the total number of self-fertilized zygotes that become juveniles.
 Note, however, that the relative strength of inbreeding depression at different stages of the life history will depend on the interaction between the various inbreeding depression parameters, which we have precluded from our analyses.
@@
 inv6 segregates at moderate frequency (about $8\%$), and carriers suffer an approximately $30 \%$ loss in pollen viability, but also increased flower (and therefore ovule and pollen) production.
-The genetic basis of these effects remain unclear -- though they are likely polygenic -- but the net result is a "supergene" with remarkably strong effects on both female and male fertility that segregates as a single diallelic locus \textcolor{blue}{(i.e., with two alleles: wild-type and inverted)}.
+The genetic basis of these effects remain unclear -- though they are likely polygenic -- but the net result is a "supergene" with remarkably strong effects on both female and male fertility that segregates as a single diallelic locus (i.e., with two alleles: wild-type and inverted).
 We estimated selection coefficients for the effect of inv6 on pollen production (i.e., taking into account the simultaneous effect on flower number) and ovule production from the data reported in \citet{LeeKelly2015} under the relatively conservative assumptions that pollen/ovule production is proportional to flower number, and  additive SA fitness effects (dominance coefficients could not be estimated from field data; see \ref{App:Mimulus} for details).
@@
 In contrast, when the selfing rate is at the higher end of empirical estimates for the Iron Mountain population in which inv6 has been documented ($C = 0.29$), inv6 falls nearer the upper boundary but still within SA polymorhic space when using the Low-Elevation Perennial demographic rates (fig.~\ref{fig:mimulusFig}B,D).
-This happens because, under the empirical estimates of selfing and inbreeding depression, the polymorphic parameter space shifts downwards \textcolor{blue}{towards the x-axis}.
+This happens because, under the empirical estimates of selfing and inbreeding depression, the polymorphic parameter space shifts downwards towards the x-axis.
 Additionally, inv6 falls much closer to the extinction threshold under partial selfing, suggesting that even relatively small perturbations to demographic rates or selection coefficients could result in non-locally adapted populations being unable to support the demographic costs associated with segregating SA alleles with selection coefficients of similar magnitude to inv6.
 
-\textcolor{blue}{In Appendix B we reanalyze this {\itshape M.~guttatus} example using a simple density-dependent version of our model.
+In Appendix B we reanalyze this {\itshape M.~guttatus} example using a simple density-dependent version of our model.
 We find that for (st)age-independent density-dependent growth, the coexistence boundaries are unchanged, and the equilibrium population densities closely resemble the shape of the density-independent population growth rates presented above (see fig.~\ref{fig:mimulusFig_dd}).
-}
 
@@
 The first and central finding of our study is that when population intrinsic growth rates approach one, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space where SA polymorphism is maintained.
-Additionally, large regions where male-beneficial alleles are predicted to \textcolor{blue}{fix} become demographically inviable for populations whose intrinsic growth rates are close to one prior to invasion.
+Additionally, large regions where male-beneficial alleles are predicted to fix become demographically inviable for populations whose intrinsic growth rates are close to one prior to invasion.
 This demographic consequence of masculinization has rarely been considered \citep{hitchcock2020gene}.
@@
 By combining the tools of demography and population genetics, the framework we present here enables the exploration of interactions between life cycle complexity, mating system, and sexual antagonism.
-For the sake of simplicity, we made a variety of assumptions in our analyses, and \textcolor{blue}{many extensions of the model are} possible.
+For the sake of simplicity, we made a variety of assumptions in our analyses, and many extensions of the model are possible.
 For example, we used a simple life cycle with just two stages for most of this paper (the \textit{Mimulus} example has $3$ stages).
@@
 
-\textcolor{blue}{One particularly important limitation of our main analysis is that we modeled density-independent growth.
+One particularly important limitation of our main analysis is that we modeled density-independent growth.
 Of course, in natural populations density-dependent processes like resource competition or competition for breeding sites will often prevent populations from growing (or shrinking) exponentially.
@@
 Our reanalysis shows that the shape of the polymorphic region is unchanged by this form of density-dependence, and that the population density surface in fig.~\ref{fig:mimulusFig_dd} is very similar to that for population growth rate in fig.~\ref{fig:mimulusFig}.
-The similarity between the density-independent and -dependent model predictions underscores a key point made by earlier studies: density-dependence acting (st)age-independently does not alter selection gradients \citep{mylius1995evolutionarily, caswell2017senescence}.}
-
-\textcolor{blue}{Yet, in many natural populations density-dependence will not have the same effect on individual survival at all ages or developmental stages.
+The similarity between the density-independent and -dependent model predictions underscores a key point made by earlier studies: density-dependence acting (st)age-independently does not alter selection gradients \citep{mylius1995evolutionarily, caswell2017senescence}.
+
+Yet, in many natural populations density-dependence will not have the same effect on individual survival at all ages or developmental stages.
 For example, competition for breeding sites will cause a reduction in fecundity rather than survival.
@@
 These types of age- or stage-specific density-dependence will change selection gradients, and therefore the shape of polymorphic parameter space as well as the equilibrium population densities predicted by the model.
-\citet{de2020matrix} discuss some of the possible dynamical outcomes of a density-dependent genetic model for \textit{Tribolium} seed beetles but a thorough analysis of more complicated forms of density-dependence for the maintenance of SA polymorphisms remains an important avenue for future work.}
+\citet{de2020matrix} discuss some of the possible dynamical outcomes of a density-dependent genetic model for \textit{Tribolium} seed beetles but a thorough analysis of more complicated forms of density-dependence for the maintenance of SA polymorphisms remains an important avenue for future work.
 
@@
 Yet, reproductive success in many hermaphroditic populations is limited by both the quantity and quality of male gametes \citep[e.g.,][]{Yund2000, AizenHarder2007, Harder2016}.
-Modeling male gamete production is a natural extension to our framework, and would enable us to analyze how tension between the demographic consequences of SA fitness variation through both female and male function alters or abolishes asymmetries in the extinction thresholds and polymorphic space \textcolor{blue}{caused from female demographic dominance} \citep[e.g.,][]{Tazzyman2015}.
+Modeling male gamete production is a natural extension to our framework, and would enable us to analyze how tension between the demographic consequences of SA fitness variation through both female and male function alters or abolishes asymmetries in the extinction thresholds and polymorphic space caused from female demographic dominance \citep[e.g.,][]{Tazzyman2015}.
 Interestingly, different forms of self-fertilization can aid in reproductive assurance, and should therefore have demographic consequences -- for example, under pollen limitation and delayed selfing (where only ovules that fail to receive outcross pollen are selfed), the selfing rate will be a function of genotype frequencies because these will directly influence pollen production \citep{HarderBarrett2006}.
 
-\textcolor{blue}{Our assumption that only how individuals themselves were produced (by selfing or outcrossing) affects the level of inbreeding depression they suffer also represents a major simplification}.
+Our assumption that only how individuals themselves were produced (by selfing or outcrossing) affects the level of inbreeding depression they suffer also represents a major simplification.
 In reality, the history of consecutive generations of inbreeding in each individual's lineage will influence the severity of inbreeding depression they experience, particularly when inbreeding depression is caused primarily by recessive deleterious mutations.
 It would be an interesting and feasible extension of our modeling framework to expand the individual state space to include selfing cohorts (i.e., first generation selfing, second generation selfing, etc.), as in \citet{kelly1999response,kelly2007mutation}, and thereby provide a more biologically meaningful approach to modelling self-fertilization and inbreeding depression.
-\textcolor{blue}{Likewise, the simple 'fixed prior selfing' model we implemented assumes constant selfing rates among individuals over time.
+Likewise, the simple 'fixed prior selfing' model we implemented assumes constant selfing rates among individuals over time.
 In reality, environmental stochasticity will influence the population selfing rate and therefore both the expression of inbreeding depression and the scope for SA polymorphism \citep{JordanConnallon2014,Olito-etal-2018}.
-A naive expectation might be that a weighted geometric mean selfing rate would represent a more biologically meaningful parameter influencing the genetic and demographic outcomes of SA selection in more complicated 'selfing environments' \citep[e.g., ][]{ConnallonSharmaOlito2019}.}
-
-% that among-individual variation in selfing might reduce the effective selfing rate in a population, expanding the scope for balanced SA polymorphisms, as it does in genotype-dependent mass-action selfing models \citep{JordanConnallon2014}.
-However, it is difficult to speculate how the associated demographic effects of such variation via reproductive assurance and inbreeding depression  would play out.}
+A naive expectation might be that a weighted geometric mean selfing rate would represent a more biologically meaningful parameter influencing the genetic and demographic outcomes of SA selection in more complicated 'selfing environments' \citep[e.g., ][]{ConnallonSharmaOlito2019}.
 
@@
 
-\textcolor{blue}{Here we extend the model presented in the main text by incorporating density-dependent population regulation.
+Here we extend the model presented in the main text by incorporating density-dependent population regulation.
 For brevity, we chose the simplest population regulation possible in a structured population model: population regulation affecting all life history stages in the same way.
@@
 \end{linenomath*}
-where $\| \tilde{\mbf{n}}(t) \|$ is the one-norm which is identical to the sum of the entries of $\tilde{\mbf{n}}(t)$ because the population vector only has non-negative entries. }
-
-\textcolor{blue}{Using this density-dependent version of the model, we reanalyzed the {\itshape M.
+where $\| \tilde{\mbf{n}}(t) \|$ is the one-norm which is identical to the sum of the entries of $\tilde{\mbf{n}}(t)$ because the population vector only has non-negative entries.
+
+Using this density-dependent version of the model, we reanalyzed the {\itshape M.
 guttatus} example from the main text.
 Fig.~\ref{fig:mimulusFig_dd} has the same structure as fig.~\ref{fig:mimulusFig}, but now shows the equilibrium population size, $N^{\ast} = \sum_i n_i $, rather than the population growth rate.
-}
-
-\textcolor{blue}{With this simplest form of density-dependence, our our main conclusions remain intact.
+
+With this simplest form of density-dependence, our our main conclusions remain intact.
 By introducing a density-dependent factor acting on survival of all life-history stages in the same way, we did not change any of the selection gradients.
@@
 Adding this simple form of density-dependence does not change the relative importance of any of the life history stages, nor any of the demographic rates in the population projection matrix.
-}
-
-\textcolor{blue}{Of course, many other more realistic (and more interesting) density-dependent scenarios are possible that will change the results presented in fig.~\ref{fig:mimulusFig_dd} (see Discussion).
+
+Of course, many other more realistic (and more interesting) density-dependent scenarios are possible that will change the results presented in fig.~\ref{fig:mimulusFig_dd} (see Discussion).
 In these more complicated scenarios, we would expect significant qualitative as well as quantitative changes to our model predictions.
-}
 
@@
 Specifically, we used data from two experimental populations with contrasting demographic rates: the locally adapted 'Eagle Meadows' population (data from 2012), and a non-local 'low-elevation perennials' population (data from 2013).
-The vital rates used in theirs, and our, calculations are summarized below in Table \ref{tab:MimDemData} (see corrected Tables.~1 and S2 in \citet{PetersonEtAl2016, PetersonEtAl2017}).
+The vital rates used in their, and our, calculations are summarized below in Table \ref{tab:MimDemData} (see corrected Tables.~1 and S2 in \citet{PetersonEtAl2016, PetersonEtAl2017}).
 Note that the same estimates for seed bank survival ($D$) and Seedling recruitment ($A$) were used for all populations in the study.
@@
 Given the obvious pitfalls of estimating selection on inv6 from the available data, it goes without saying that the location of inv6 in demographically viable polymorphic parameter space is a highly speculative, but nevertheless interesting, proof of concept for our model predictions.
+Supplementary data files showing how selection coefficients were calculated are publicly available on Dryad (\citealt{Olito-DeVries-2022-Dryad}; \url{https://doi.org/10.5061/dryad.c2fqz619t}).
 

### 2022-01-22 Scratch.R

@@ -1,1 +1,0 @@
-## Scratch

### 2022-01-22 fix-Zeta-MWE.R
untrack ancillary files
This commit untracks two .R scripts that were used to troubleshoot
during the development of the models, but are not essential for
reproducing the main results.

@@ -1,338 +1,0 @@
-##  Minimum working example of disagreement b/w
-##   Simulations and Zeta_i
-
-
-# run simulations necessary to show disagreement
-# (can take 10 min  or so)
-rm(list=ls())
-source('R/functions-Simulations.R')
-
-	# Obligate Outcrossing
-	# C = 0
-	pars  <-  list(
-					"hf"     =  1/2,
-					"hm"     =  1/2,
-					"C"      =  0,
-					"delta"  =  0
-					)
-
-	selLoop(sMax = 0.15, nSamples=1000,
-			om = 2, g = 3, theta = c(0.6,0.6,0.6,0.6,0.05,0.05,6), theta_prime = c(0.6,0.6,0.6,0.6,0.05,0.05,6), 
-			hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, tlimit = 10^5, returnRes = FALSE)
-
-	# Intermediate Selfing
-	# C = 1/2
-	pars  <-  list(
-					"hf"     =  1/2,
-					"hm"     =  1/2,
-					"C"      =  1/2,
-					"delta"  =  0
-					)
-
-	selLoop(sMax = 0.15, nSamples=1000,
-			om = 2, g = 3, theta = c(0.6,0.6,0.6,0.6,0.05,0.05,6), theta_prime = c(0.6,0.6,0.6,0.6,0.05,0.05,6), 
-			hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, tlimit = 10^5, returnRes = FALSE)
-
-
-	# Obligate Outcrossing
-	# C = 0
-	pars  <-  list(
-					"hf"     =  1/4,
-					"hm"     =  1/4,
-					"C"      =  0,
-					"delta"  =  0
-					)
-
-	selLoop(sMax = 0.15, nSamples=1000,
-			om = 2, g = 3, theta = c(0.6,0.6,0.6,0.6,0.05,0.05,6), theta_prime = c(0.6,0.6,0.6,0.6,0.05,0.05,6), 
-			hf = pars$hf, hm = pars$hm, C = pars$C, delta = pars$delta, tlimit = 10^5, returnRes = FALSE)
-
-	# Intermediate Selfing
-	# C = 1/2
-	pars  <-  list(
-					"hf"     =  1/4,
-					"hm"     =  1/4,
-					"C"      =  1/2,
-					"delta"  =  0
-					)
-
-	selLoop(sMax = 0.15, nSamples=1000,
[…]

### 2022-01-22 test_SA_Hermaphrodite_Model.R

@@ -1,272 +1,0 @@
-# First attempt to implement a Mendelian Matrix Population Model
-# for a partially selfing population of hermaphrodites with selection
-# through both sex functions.
-#
-#	For simplicity I'm just hacking Lotte's code here to see if
-#	all the component matrices fit, and if the model does what 
-#   I expect it to do.
-#
-# The model has:
-# 	Two stages: juvenile and adult
-# 	Three genotypes: AA, Aa, aa
-#   A constant selfing rate: C_AA = C_Aa = C_aa = C, where 0 <= C <= 1
-#   Constant inbreeding depression: 0 < delta < 1
-
-
-rm(list=ls())
-#####################
-##  Dependencies
-source('R/functions-MatModels.R')
-
-Ainvade  <-  TRUE
-intInit  <-  FALSE
-
-# Simulation time"
-tlimit <- 10^4
-
-# dimensions
-om  <-  2
-g   <-  3
-
-# Create identity and ones matrices
-Ig   <-  diag(g)
-Iom  <-  diag(om)
-eg   <-  ones(c(g,1))
-eom  <-  ones(c(om,1))
-K    <-  vecperm(om,g)
-
-# Population Selfing Rate and inbreeding depression
-C      <-  1/4
-delta  <-  0
-
-#####################################################"
-
-# PARAMETERS AND INITIAL CONDITIONS 
-
-####################################################"
-
-# BASELINE PARAMETERS for survival and 
-# fertility through Female and Male function
-# theta=[sigma_J, sigma_A, gamma, f_ii]
-theta_temp  <-  c(0.6,0.6,0.05,6.8)
-theta       <-  rep.col(theta_temp,3)
-
-theta_primetemp <- c(0.6,0.6,0.05,6.8)
-theta_prime <- rep.col(theta_primetemp,3)
-
-## SELECTION PARAMETERS
-hf  <-  1/2
-hm  <-  1/2
-sf  <-  0.1
-sm  <-  0.12
-fii        <-  c(1, 1 - hf*sf, 1 - sf)
-fii_prime  <-  c(1 - sm, 1 - hm*sm, 1)
-
-##SELECTION DIFFERENTIAL FEMALE-FUNCTION
-##female function
-theta[4,]  <-  theta[4,]*fii
[…]

### 2022-01-29 R/inv6-selection-coefficients.R
Add inv6 sel coeff calculations script
This commit adds the short R script to calculate the selection
coefficients on inv6 from the Lee et al. (2016) study.

@@ -1,0 +1,93 @@
+################################################################
+#' Load measurements for estimating selection coefficients for
+#' inv6 from figures in Lee et al. (2016), back-calculate
+#' selection coefficients, and save as objects to be called 
+#' by ./makeFigs.R
+#'	
+#'  Author: Colin Olito
+#' 
+#' 	Notes:	Estimates for male and female fitness components
+#' were back-calculated from direct measures of figures 2, 5,
+#' & 6 in:
+#' 
+#' Lee, Y. W., L. Fishman, J. K. Kelly, and J. H. Willis. 2016.
+#' A segregating inversion generates fitness variation in 
+#' yellow monkeyflower (Mimulus guttatus). Genetics 202:1473–1484.
+#' 
+#' Raw measurements of the figures are provided as supplementary
+#' files  inv6-measurements.pptx and inv6-measurements.odp, 
+#' which are available on DRYAD at:
+#' 
+#' DRYAD Reviewer link (stable link to be provided when assigned)
+#' https://datadryad.org/stash/share/81sAuXGEg8cSh-S9VVL0PfBCsl6YLkG1OIFBCvOefac 
+
+
+
+
+
+###############
+# Import data
+###############
+
+inv6Data  <-  read.csv('./data/inv6_Figure_Measurements.csv', header=TRUE)
+
+#############################################
+# Rescale back to original measurement units
+#############################################
+
+# Female fitness components
+no_inv6_flowers   <-  (inv6Data$no_inv6_flowers_inches / inv6Data$pptInches[inv6Data$scaleUnits == 'flowers']) * inv6Data$scaleAxis[inv6Data$scaleUnits == 'flowers'] 
+inv6_het_flowers  <-  (inv6Data$inv6_het_flowers_inches / inv6Data$pptInches[inv6Data$scaleUnits == 'flowers']) * inv6Data$scaleAxis[inv6Data$scaleUnits == 'flowers'] 
+
+no_inv6_fruits   <-  (inv6Data$no_inv6_fruits_inches / inv6Data$pptInches[inv6Data$scaleUnits == 'fruits']) * inv6Data$scaleAxis[inv6Data$scaleUnits == 'fruits'] 
[…]

### 2022-01-29 R/functions-Figs.R
Add reference to inv6 coeff calcs
This commit adds minor changes to ./R/functions-Figs.R file so that
it automatically calculates the selection coefficients on inv6 from
the data file ./data/inv6_Figure_Measurements.csv, using
./R/inv6-selection-coefficients.R, and then calls these values in
all the relevant plotting functions.

@@ -13,6 +13,7 @@ library(raster)
 library(akima)
 
 source('./R/functions-Simulations.R')
+source('./data/inv6-selection-coefficients.R')
 
 #######################
 # AUXILLIARY FUNCTIONS
@@ -30,6 +31,7 @@ toDev <- function(expr, dev, filename, ..., verbose=TRUE) {
   if ( verbose )
     cat(sprintf('Creating %s\n', filename))
     dev(filename, family='CM Roman', ...)
+#    dev(filename, family='yourFONT', ...) # To embed figs with the font of your choice, uncomment this line (and comment out the previous one) and replace yourFONT with your chosen font family.
     on.exit(dev.off())
     eval.parent(substitute(expr))
 }
@@ -1394,9 +1396,6 @@ MimulusInv6Fig  <-  function() {
 
 #    ext4$sf[46]  <-  mean(ext4$sf[45], ext4$sf[47])
     ext4$smThreshold[46]  <-  mean(c(ext4$smThreshold[45], ext4$smThreshold[47]))
-    # Selection coefficients for Inv6 (including average flower # effect)
-    sfInv6  <-  0.31
-    smInv6  <-  0.30
 
 # Color scheme
     COLS  <-  list(
@@ -1426,11 +1425,11 @@ MimulusInv6Fig  <-  function() {
         lines(AInvBound ~ sms, lty=2, lwd=2, col=COLS$line, data=inv2)
         lines(c(inv2$aInvBound[inv2$aInvBound < max(inv2$sms)],max(inv2$sms)) ~ c(inv2$sms[inv2$aInvBound < max(inv2$sms)],0.50), lty=2, lwd=2, col=COLS$line, data=inv2)
         # no lines to plot for extinction threshold (all polymorphic space is viable)
-        points(sfInv6 ~ smInv6, pch=21, col=COLS$line, bg=COLS$fill, cex=1.25)
+        points(inv6_sf ~ inv6_sm, pch=21, col=COLS$line, bg=COLS$fill, cex=1.25)
         # axes        
         axis(1, las=1)
         axis(2, las=1)
[…]

### 2022-01-29 run-Simulations.R
Use data files in run-Simulations.R
This commit alters the run-Simulations.R script so that prior to
running any simulations, it imports the .csv data files for the
Peterson et al. (2016) and Willis (1993) parameters reported in the
main text and Appendix D, and stores them as list objects. The lists
are then used for all simulations in the Mimulus case study.

@@ -8,7 +8,7 @@
 #'  Author: Colin Olito
 #'
 #'  NOTES:  To reproduce the analyses for the paper, run the 
-#' 			following R code up to L.725. This will generate
+#' 			following R code up to L.717. This will generate
 #' 			output data as .csv files saved to ./output/simData/
 #' 			Be sure to create this directory locally prior
 #' 			to running the simulations.
@@ -21,6 +21,36 @@ rm(list=ls())
 source('R/functions-Simulations.R')
 source('R/loadData-Compadre.R')
 
+###########################################
+##  Import parameter values from tables in
+##  Peterson et al. (2016) & Willis (1993)
+petersonData  <-  read.csv('./data/Peterson_2016_Data.csv', header=TRUE)
+WillisData    <-  read.csv('./data/Willis_1993_Data.csv', header=TRUE)
+Peterson2016.EM.theta.list  <-  list(D = petersonData$D[petersonData$population == "EM"], 
+						G = petersonData$G[petersonData$population == "EM"],
+						F = petersonData$F[petersonData$population == "EM"],
+						O = petersonData$O[petersonData$population == "EM"],
+						A = petersonData$A[petersonData$population == "EM"],
+						S = petersonData$S[petersonData$population == "EM"],
+						R = petersonData$R[petersonData$population == "EM"],
+						pop = petersonData$population[petersonData$population == "EM"]
+						)
+Peterson2016.LEP.theta.list  <-  list(D = petersonData$D[petersonData$population == "LEP"], 
+						 G = petersonData$G[petersonData$population == "LEP"],
+						 F = petersonData$F[petersonData$population == "LEP"],
+						 O = petersonData$O[petersonData$population == "LEP"],
+						 A = petersonData$A[petersonData$population == "LEP"],
+						 S = petersonData$S[petersonData$population == "LEP"],
+						 R = petersonData$R[petersonData$population == "LEP"],
[…]

### 2022-01-29 R/functions-Simulations.R
Minor declutter
This commit removed a deprecated version of the calc_zeta() function
which was commented out.

@@ -165,29 +165,6 @@ predDelta  <-  function(dStar, b=1/2, a=0.2, C) {
 ###########################################
 #' Eigenvalue Calculator for full demographic model
 #' 
-#calcZeta  <-  function(om, FSi, FXi, FXi_pr, USi, UXi, pHat_AA, pHat_aa, C, delta, lambda_i){
-#
-#	pn_AA   <-  ones(c(om)) %*% FXi_pr[,,1] %*% (pHat_AA[1:om] + pHat_AA[om+1:om])
-#	pn_aa   <-  ones(c(om)) %*% FXi_pr[,,3] %*% (pHat_aa[(4*om+1):(5*om)] + pHat_aa[(5*om+1):(6*om)])
-#	M22_AA  <-  rbind(
-#					cbind(USi[,,2] + C*(1 - delta)*FSi[,,2]/2, C*(1 - delta)*FSi[,,2]/2, zeros(c(om,om))),
-#					cbind(           ((1 - C)/2)*FXi[,,2] + kronecker(c(((1 - C)/(2*pn_AA)))*FXi[,,1]%*%(pHat_AA[1:om] + pHat_AA[om+1:om]), (ones(om)%*%FXi_pr[,,2])),
-#						  UXi[,,2] + ((1 - C)/2)*FXi[,,2] + kronecker(c(((1 - C)/(2*pn_AA)))*FXi[,,1]%*%(pHat_AA[1:om] + pHat_AA[om+1:om]), (ones(om)%*%FXi_pr[,,2])), 
-#						                 (1 - C)*FXi[,,3] + kronecker(c(((1 - C)/(  pn_AA)))*FXi[,,1]%*%(pHat_AA[1:om] + pHat_AA[om+1:om]), (ones(om)%*%FXi_pr[,,3]))),
-#					cbind(((C*(1 - delta))/4)*FSi[,,2], ((C*(1 - delta))/4)*FSi[,,2], USi[,,3] + C*(1 - delta)*FSi[,,3])
-#					)
-#	M22_aa  <-  rbind(
-#					cbind(USi[,,2] + C*(1 - delta)*FSi[,,2]/2, C*(1 - delta)*FSi[,,2]/2, zeros(c(om,om))),
[…]

### 2022-01-29 doc/SA-Hermaphrodites-wDemography-R1.tex
Add Dryad reviewer link to ms

@@
 
-\textit{Online enhancements}: Dryad digital repository ({\itshape https://doi.org/10.5061/dryad.c2fqz619t}).
+\textit{Online enhancements}: Dryad digital repository ({\itshape https://datadryad.org/stash/share/81sAuXGEg8cSh-S9VVL0PfBCsl6YLkG1OIFBCvOefac}).
 
@@
 The derivation follows closely the logic presented in \citet{deVriesCaswell2019a, deVriesCaswell2019b}, and in fact, it reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019a} under obligate outcrossing (when selfing is not allowed).
-All computer code and data necessary to reproduce the results are available on Dryad (\url{https://doi.org/10.5061/dryad.c2fqz619t}) and on GitHub (\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}).
+All computer code and data necessary to reproduce the results are available on Dryad (\url{https://datadryad.org/stash/share/81sAuXGEg8cSh-S9VVL0PfBCsl6YLkG1OIFBCvOefac}) and on GitHub (\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}).
 
@@
 Given the obvious pitfalls of estimating selection on inv6 from the available data, it goes without saying that the location of inv6 in demographically viable polymorphic parameter space is a highly speculative, but nevertheless interesting, proof of concept for our model predictions.
-Supplementary data files showing how selection coefficients were calculated are publicly available on Dryad (\citealt{Olito-DeVries-2022-Dryad}; \url{https://doi.org/10.5061/dryad.c2fqz619t}).
+Supplementary data files showing how selection coefficients were calculated are publicly available on Dryad (\citealt{Olito-DeVries-2022-Dryad}; \url{https://datadryad.org/stash/share/81sAuXGEg8cSh-S9VVL0PfBCsl6YLkG1OIFBCvOefac}).
 

### 2022-01-29 R/functions-Figs.R

@@ -13,7 +13,7 @@ library(raster)
 library(akima)
 
 source('./R/functions-Simulations.R')
-source('./data/inv6-selection-coefficients.R')
+source('./R/inv6-selection-coefficients.R')
 
 #######################
 # AUXILLIARY FUNCTIONS

### 2022-02-08 doc/SA-Hermaphrodites-wDemography-R1.tex
Add Dryad doi, update .gitignore, etc
This commit added the official dryad doi to the references,
added the ./notes/ folder to the list of untracked folders for the
public repo, and made minor edits to the README.md

@@
 
-\textit{Online enhancements}: Dryad digital repository ({\itshape https://datadryad.org/stash/share/81sAuXGEg8cSh-S9VVL0PfBCsl6YLkG1OIFBCvOefac}).
+\textit{Online enhancements}: Dryad digital repository (\citealt{Olito-DeVries-2022-Dryad}; reviewer link: \url{https://datadryad.org/stash/share/81sAuXGEg8cSh-S9VVL0PfBCsl6YLkG1OIFBCvOefac}).
 
@@
 The derivation follows closely the logic presented in \citet{deVriesCaswell2019a, deVriesCaswell2019b}, and in fact, it reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019a} under obligate outcrossing (when selfing is not allowed).
-All computer code and data necessary to reproduce the results are available on Dryad (\url{https://datadryad.org/stash/share/81sAuXGEg8cSh-S9VVL0PfBCsl6YLkG1OIFBCvOefac}) and on GitHub (\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}).
+All computer code and data necessary to reproduce the results are available on Dryad (\citealt{Olito-DeVries-2022-Dryad}, reviewer link: \url{https://datadryad.org/stash/share/81sAuXGEg8cSh-S9VVL0PfBCsl6YLkG1OIFBCvOefac}) and on GitHub (\url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}).
 

### 2022-02-08 matlab/deVries_Olito_selfing.m
removed notes and matlab folders from github

@@ -1,276 +1,0 @@
-%Program to implement Mendelian Matrix Population Models
-%for two-stage model with selfing
-%Version where how-you-were-produced does not affect your demographic rates
-%Authors: C. de Vries and C. Olito
-% July 2020
-clearvars; 
-%Dimensions, 2 stages, 3 genotypes
-om=2;
-g=3;
-
-%Identity and ones matrices
-Ig=eye(g);
-Iom=eye(om);
-eg=ones(g,1);
-eom=ones(om,1);
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Parameters for EVOLUTIONARY SUICIDE (Fig 1A and 1C). 
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%Params from Colins R code: fwdDemModelSim  <-  function(	om = 2, g = 3, theta = c(0.6,0.6,0.05,5.9), theta_prime = c(0.6,0.6,0.05,5.9), 
-%								hf = 1/2, hm = 1/2, sf = 0.1, sm = 0.105, C = 0, delta = 0, tlimit = 10^4, Ainvade = FALSE, intInit = FALSE, ...) {
-
-Nzero=[99.9000 0 0.1000
-    99.9000 0 0.1000];
-%Baseline parameters
-%theta=[s1, s2, g, f]
-
-theta=[0.6 0.6 0.05 6.1]'*ones(1,3);
-%Selection differentials
-
-% %s1, juvenile survival
-% ds1=0.032;
-%  seldiff=[-ds1+0.01 ds1+0.008 -ds1+0.005];
-%  theta(1,:)=theta(1,:)+seldiff;
-
-%s2 differential, adult survival
-% s2
-%     ds2=0.0001;
-% 
-% seldiff=[0 0 ds2];
-% theta(2,:)=theta(2,:)+seldiff;
-
-%this is the f differential for the pleiotropy example
-% f
-    hf=.5;
-    hm=.5;
-    sf=0.1;
-    sm=0.105;
-    seldiff=[1 (1 - hf*sf) (1 - sf)];
-theta(4,:)=theta(4,:).*seldiff;
-
-seldiff_prime=[(1 - sm) (1 - hm*sm) 1];
-f_prime=theta(4,:).*seldiff_prime;
[…]

### 2022-02-08 matlab/deVries_Olito_selfing_titrate.m

@@ -1,315 +1,0 @@
-%Program to implement Mendelian Matrix Population Models
-%for two-stage model with selfing
-%Version where how-you-were-produced does not affect your demographic rates
-%Authors: C. de Vries and C. Olito
-% July 2020
-clearvars -except lambda_eq; 
-%Dimensions, 2 stages, 3 genotypes
-om=2;
-g=3;
-
-%Identity and ones matrices
-Ig=eye(g);
-Iom=eye(om);
-eg=ones(g,1);
-eom=ones(om,1);
-
-%Params from Colins R code: fwdDemModelSim  <-  function(	om = 2, g = 3, theta = c(0.6,0.6,0.05,5.9), theta_prime = c(0.6,0.6,0.05,5.9), 
-%								hf = 1/2, hm = 1/2, sf = 0.1, sm = 0.105, C = 0, delta = 0, tlimit = 10^4, Ainvade = FALSE, intInit = FALSE, ...) {
-
-Nzero=[99.9000 0 0.1000
-    99.9000 0 0.1000];
-%Baseline parameters
-%theta=[s1, s2, g, f]
-
-theta=[0.6 0.6 0.05 5.8]'*ones(1,3);
-sm=0:0.001:0.15;
-lambda_save_bigA=nan*ones(length(sm),3);
-lambda_save_a=nan*ones(length(sm),3);
-
-%The precision below which the algorithm stops searching for the boundary sm
-%value
-deltathres=0.0001;
-
-%Selection differentials
-%this is the f differential for the pleiotropy example
-% f
-    hf=.25;
-    hm=.25;
-    
-Thres_bigA=nan*ones(length(sm),1);
-Thres_a=nan*ones(length(sm),1);
-
-delta=0;
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Find threshold zeta_AA=1 for each sm
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-for j=1:length(sm)
-
-    if isnan(Thres_bigA(j)) 
-        sf_temp=[0 0.15]; delta_titrate=1;
-                    % first just a  check
-                    [x_bigA0,x_a0,lambda_bigA,lambda_heterozyg,lambda_a]=eigenvalues_Jacobian(theta,hf,hm,0,sm(j),delta);
[…]

### 2022-02-08 matlab/deVries_Olito_selfing_titrate_flip.m

@@ -1,319 +1,0 @@
-%Program to implement Mendelian Matrix Population Models
-%for two-stage model with selfing
-%Version where how-you-were-produced does not affect your demographic rates
-%Authors: C. de Vries and C. Olito
-% July 2020
-clearvars ;
-%Dimensions, 2 stages, 3 genotypes
-om=2;
-g=3;
-
-%Identity and ones matrices
-Ig=eye(g);
-Iom=eye(om);
-eg=ones(g,1);
-eom=ones(om,1);
-
-%Params from Colins R code: fwdDemModelSim  <-  function(	om = 2, g = 3, theta = c(0.6,0.6,0.05,5.9), theta_prime = c(0.6,0.6,0.05,5.9), 
-%								hf = 1/2, hm = 1/2, sf = 0.1, sm = 0.105, C = 0, delta = 0, tlimit = 10^4, Ainvade = FALSE, intInit = FALSE, ...) {
-
-Nzero=[99.9000 0 0.1000
-    99.9000 0 0.1000];
-%Baseline parameters
-%theta=[s1, s2, g, f]
-
-theta=[0.6 0.6 0.05 5.8]'*ones(1,3);
-sf=0:0.001:0.15;
-lambda_save_bigA=nan*ones(length(sf),3);
-lambda_save_a=nan*ones(length(sf),3);
-
-%The precision below which the algorithm stops searching for the boundary sm
-%value
-deltathres=0.0001;
-
-%Selection differentials
-%this is the f differential for the pleiotropy example
-% f
-    hf=.25;
-    hm=.25;
-    
-Thres_bigA=nan*ones(length(sf),1);
-Thres_a=nan*ones(length(sf),1);
-
-delta=0;
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Find threshold zeta_AA=1 for each sm
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-for j=1:length(sf)
-
-    if isnan(Thres_bigA(j)) 
-        sm_temp=[0 0.15]; delta_titrate=1;
-                    % first just a  check
-                    [x_bigA0,x_a0,lambda_bigA,lambda_heterozyg,lambda_a]=eigenvalues_Jacobian(theta,hf,hm,sf(j),0,delta);
[…]

### 2022-02-08 matlab/eigenvalues_Jacobian.m

@@ -1,41 +1,0 @@
-function [x_bigA,x_a,lambda_bigA,lambda_heterozyg,lambda_a]=eigenvalues_Jacobian(theta,hf,hm,sf,sm,delta)
-                              
-s1=theta(1,:);
-s2=theta(2,:);
-gam=theta(3,:);
-lambda=zeros(3,1);
-%selfing rate
-C=0*ones(1,3);
-
-    seldiff=[1 (1 - hf*sf) (1 - sf)];
-    f=theta(4,:).*seldiff;
-    seldiff_prime=[(1 - sm) (1 - hm*sm) 1];
-    f_prime=theta(4,:).*seldiff_prime;
-
-    for i=1:3
-        U{i}=[s1(i)*(1-gam(i)) 0 ;
-        s1(i)*gam(i) s2(i)];
-        F{i}=(1-C(i))*[0 f(i);
-        0 0];
-        F_S{i}=C(i)*(1-delta)*[0 f(i);
-        0 0];
-        F_prime{i}=[0  f_prime(i)
-        0 0];
-    %Genotype specific population growth rate
-        lambda(i)=max(eig(F{i}+U{i}+F_S{i}));
-    end % for i
-%Calculating the largest eigenvalue of the Jacobian for both boundaryies
-Jac_bigA=[U{2}+.5.*F{2}+.5.*F_S{2}+.5.*(f_prime(2)/f_prime(1)).*F{1} F{3}+(f_prime(3)/f_prime(1)).*F{1}
-            .25.*F_S{2} U{3}+F_S{3}];
-Jac_a=[U{2}+.5.*F{2}+.5.*F_S{2}+.5*(f_prime(2)/f_prime(3)).*F{3} F{1}+(f_prime(1)/f_prime(3)).*F{3}
-            .25.*F_S{2} U{1}+F_S{1}];    
-%Dominant eigenvalue of the Jacobian at the AA boundary
-x_bigA=max(eig(Jac_bigA))/lambda(1);
-%Dominant eigenvalue of the Jacobian at the aa boundary
-x_a=max(eig(Jac_a))/lambda(3);
-
-lambda_bigA=lambda(1);
-lambda_heterozyg=lambda(2);
-lambda_a=lambda(3);
-
-end

### 2022-02-08 matlab/emat.m

@@ -1,10 +1,0 @@
-function [E] = emat(m,n,i,j)
-%emat(m,n,i,j)
-%matrix mXn, with 1 in i,j entry and zeros elsewhere
-
-E=zeros(m,n);
-E(i,j)=1;
-
-
-end
-

### 2022-02-08 matlab/simulate_dyn.m

@@ -1,130 +1,0 @@
-function [lambda_sim]=simulate_dyn(theta, hf,hm,sf,sm,delta,nzero)
- %[lambda_sim,pout,nout]=simulate_dyn(theta,hf,hm,sf,sm,delta,nzero)
-                      
-s1=theta(1,:);
-s2=theta(2,:);
-gam=theta(3,:);
-lambda=zeros(3,1);
-%selfing rate
-C=0*ones(1,3);
-g=3;
-om=2;
-Ig=eye(g);
-Iom=eye(om);
-eg=ones(g,1);
-eom=ones(om,1);
-
-    seldiff=[1 (1 - hf*sf) (1 - sf)];
-    f=theta(4,:).*seldiff;
-    seldiff_prime=[(1 - sm) (1 - hm*sm) 1];
-    f_prime=theta(4,:).*seldiff_prime;
-
-    for i=1:3
-        U{i}=[s1(i)*(1-gam(i)) 0 ;
-        s1(i)*gam(i) s2(i)];
-        F{i}=(1-C(i))*[0 f(i);
-        0 0];
-        F_S{i}=C(i)*(1-delta)*[0 f(i);
-        0 0];
-        F_prime{i}=[0  f_prime(i)
-        0 0];
-    %Genotype specific population growth rate
-        lambda(i)=max(eig(F{i}+U{i}+F_S{i}));
-    end % for i
-
-
-d=eye(g);
-c=[0; 1]; %Indicator vector, with a 1 in positions corresponding to breeding stages
-
-bbD=kron(Iom,d); %Equation 27 main text
-bbU=zeros(g*om); 
-bbF=zeros(g*om);
-bbF_S=zeros(g*om);
-bbF_prime=zeros(g*om);
-for i=1:g
-    bbU=bbU+kron(emat(g,g,i,i),U{i}); %Equation 25 main text
-    bbF=bbF+kron(emat(g,g,i,i),F{i}); %Equation 26 main text
-    bbF_prime=bbF_prime+kron(emat(g,g,i,i),F_prime{i});
-    bbF_S=bbF_S+kron(emat(g,g,i,i),F_S{i});
-end % for i
-
-%Matrices
-Ig=eye(g);
-K=vecperm(om,g);
-
-X=kron(Ig,c'); %Equation 18
-
-W=[1 .5 0
-    0 .5 1];
-
[…]

### 2022-02-08 matlab/vec.m

@@ -1,3 +1,0 @@
-function y=vec(x);
-
-y=x(:);

### 2022-02-08 matlab/vecperm.m

@@ -1,19 +1,0 @@
-
-function p = vecperm(m,n)
-% vecperm
-% function to calculate the vec permutation matrix of size m,n
-% 4/9/03
-%function p = vecperm(m,n)
-
-p = zeros(m*n);
-a = zeros(m,n);
-for i = 1:m
-    for j = 1:n
-        e = a;
-        e(i,j) = 1;
-        p = p + kron(e,e');
-    end
-end
-
-
-        
